### 2024-07-08 scripts/chicken_scratch/cb_geomarket_ch.Rmd
initial commit, copying files from messed up repo

---
title: 'CB Geomarket'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include = F}
knitr::opts_chunk$set(message = F, warning = F)

options(scipen = 10000)
```

```{r}
library(tidyverse)
library(readxl)
library(lubridate)
library(haven)
library(labelled)
library(tidycensus)
library(sf)
library(tigris)
library(maptools)
```

# Load data

```{r}
data_dir <- file.path('..', '..', 'data')
eps_data_dir <- file.path(data_dir, 'eps_market')
```

```{r}
# Geomarket to zip conversion
eps_zip <- read_excel(path = file.path(eps_data_dir, 'eps.xls'), sheet = 'Sheet 1') %>% 
  mutate(eps = if_else(str_sub(eps, 3, 3) == '0', str_c(str_sub(eps, 1, 2), ' ', str_sub(eps, 4, 4)), eps))

eps_zip
```

```{r}
# Zip code shapes (2010)
zcta_2010 <- zctas(
  cb = T, 
  year = 2010
)

zcta_2010
```

```{r}
# Zip code income data (2015)
income_2015_zcta <- get_acs(
  geography = 'zcta',
  variables = 'B19013_001',
  year = 2015,
  geometry = F,
  survey = 'acs5'
)

income_2015_zcta
```

```{r}
# Aggregate income data to geomarket level
income_2015_eps <- eps_zip %>% 
  inner_join(income_2015_zcta, by = c('zip' = 'GEOID')) %>%
  group_by(eps) %>% 
  summarize(
    mean_inc = mean(estimate, na.rm = T)
  ) %>% 
  ungroup()

income_2015_eps
income_2015_eps %>% arrange(desc(mean_inc)) %>% print(n=100)
```

```{r}
# CA data
eps_zip_CA <- zcta_2010 %>%  # has to be left join (right/inner loses attributes on geometry)
  left_join(eps_zip, by = c('ZCTA5' = 'zip')) %>% 
  filter(str_detect(eps, '^CA'))

eps_zip_CA
```

```{r}
myzip <- readShapePoly(file.path(eps_data_dir, 'US_Zip_-_Lower_48.shp'))

myzip@data <- data.frame(myzip@data, eps_zip[match(as.character(myzip@data$ZIP), eps_zip$zip),])
myzip@data
myzip_CA <- myzip[!is.na(myzip@data$eps) & str_detect(myzip@data$eps, '^CA'), ]
```

# Plot CA (zip-code level)

## Data from `US_Zip_-_Lower_48.shp`

Matches zip codes from [CA state geoportal](https://gis.data.ca.gov/datasets/CDEGIS::california-zip-codes/explore?location=38.510705%2C-120.973098%2C6.71):

```{r}
plot(myzip_CA, col = 'pink')
```

## Data from Census Bureau

Matches zip codes from [Census Bureau](https://hub.arcgis.com/datasets/7f3aa1bd8d1f4915b928ff4da5f3f3bb/explore?location=38.818303%2C-119.108380%2C6.90):

```{r}
eps_zip_CA %>% 
  ggplot(aes(fill = 'pink')) +
  geom_sf()
```

Color by zip code-level income:

```{r}
eps_zip_CA %>% 
  left_join(income_2015_zcta, by = c('NAME' = 'GEOID')) %>% 
  ggplot(aes(fill = estimate)) +
  geom_sf(color = NA) +
  scale_fill_gradient(low = 'lightyellow', high = 'red')
```

# Plot CA (geomarket level)

No difference between aggregate functions:

```{r}
# Aggregate by identity function
eps_shapes <- eps_zip_CA %>%
  aggregate(by = list(eps_zip_CA$eps), FUN = identity) %>% 
  select(-eps) %>% rename(eps = Group.1)

eps_shapes %>% 
  ggplot(aes(fill = 'pink')) +
  geom_sf()
```

```{r}
# Aggregate by mean function
eps_shapes <- eps_zip_CA %>%
  aggregate(by = list(eps_zip_CA$eps), FUN = mean) %>% 
  select(-eps) %>% rename(eps = Group.1)

eps_shapes %>% 
  ggplot(aes(fill = 'pink')) +
  geom_sf()
```

Color by geomarket-level income:

```{r}
eps_shapes %>% 
  left_join(income_2015_eps, by = 'eps') %>% 
  ggplot(aes(fill = mean_inc)) +
  geom_sf(color = NA) +
  geom_sf_label(aes(label = eps)) +
  scale_fill_gradient(low = 'lightyellow', high = 'red')
```

Color by zip code-level income w/ geomarket outline (i.e., within geomarket):

```{r}
eps_zip_CA %>% 
    left_join(income_2015_zcta, by = c('NAME' = 'GEOID')) %>% 
    ggplot() +
    geom_sf(aes(fill = estimate), color = NA) +
    geom_sf(data = eps_shapes, fill = NA, color = 'black') +
    scale_fill_gradient(low = 'lightyellow', high = 'red')
```

### 2024-07-08 text/front_end.qmd
initial commit, copying files from messed up repo

---
title: "Structuring College Access: The Market Segment Model and College Board Geo-markets"
author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
bibliography: ./../assets/bib/cb_geomarket_bib.bib
csl: ./../assets/bib/apa.csl
format:
  pdf:
    documentclass: article
    #papersize: letter
    number-sections: true
    number-depth: 2
    #toc: FALSE
    #pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #keep_tex: true
    extensions: [andrewheiss/wordcount]
editor: visual
geometry:
  margin=1in
fontsize: 11pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
library(knitr)
#library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)

#hey ho This many words: {{< words-total >}}
```

```{=tex}
\pagenumbering{roman}
\newpage
\pagenumbering{gobble}
\pagenumbering{arabic}
```

# Introduction

In 1983, Zemsky and Oedel authored *The Structure of College Choice*, published by *The College Board.* This book develops the Market Segment Model and introduces local "community-based enrollment markets" (p.14) -- later called "geomarkets" -- which are the basis for the College Board market Enrollment Planning Service (EPS), a market research product. @RN4982[pp., 9-10] write that the Market Segment Model sought to quantify the knowledge that college admissions officers have about student demand:

> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting\ldots We believe that the institutions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research, thus, is based on listening carefully to what admissions officers have to say."

@RN4982 [p. 10] stress the importance of "knowing the territory":

> A good recruiter knows where to look for prospective applicants, as seen in the students' willingness or eluctance to travel. It is necessary to identify not only the most promising communities or poools, but also the specific neighborhoods within those communities -- hence the recruiter's emphasis on feeder high schools. From the beginning, we have sought to mold our research to these concepts, to capture and quantify the phenomenon behind the folklore. Thus, it is the admissions officers' notion of admissions pools that dictates the geographic units of our analysis." \[p. 11\]

The Market Segment Model was created by analyzing College Board data about which colleges SAT test-takers from 1980 sent scores to. High school students who sent scores mostly to colleges in their local geomarket were defined as "local" students. "In-state" students were those who sent most scores to colleges outside their local market but in their home state. "Regional" students were those who sent most scores to colleges outside their home state but in their region (e.g., New England, Midwest). "National" students were those who sent most scores to colleges outside their region. Having categorized students, each local geomarket could be evaluated by the number of students from each of the four market segments, yielding the Market Segment Profile (see @tbl-market-segment-characteristics). Next, for each local geomarket, the Institutional Profile (see @tbl-sample-institutional-profile) the number of students from each market segment that sent scores to a particular college, which could be your college or a competitor.

Collectively, the quantitative outputs of the Market Segment Model provided intelligence about where are the desirable students and which students might consider your college or a competitor. The Market Segment model and geomarkets became the basis or the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today. @RN5019_2010 says "Enrollment Planning Service (EPS\textsuperscript{\textregistered}) the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found." @RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.

In 19XX, geomarkets were incorporated into the College Board Student Search Service student list product, which College Board created in 1972 [@belkin2019-studata]. Student lists contain the contact information of prospective students, the basic input for recruiting campaigns that target individual prospects using mail, email, and targeted social media. RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists. The Student Search Service database consists of College Board test-takers. Colleges pay a fee for each prospect (e.g., \$0.50 in 20XX) and control which prospect profiles through purchase by selecting search filters, such as high school graduation year, SAT score, AP test score, state, etc. Geomarket filters enable colleges to include/exclude prospects from particular geomarkets.

Scholarship in the sociology of education describes schools and colleges as "sorting machines" that reproduce inequality by incorporating, creating, and allocating students to different categories [@RN4873]. Given the consequences subsequent outcomes, the sociology of education is particularly concerned with the transition from high school to college. Over the last decade or so, scholarship on college access has begun to take more seriously the idea that colleges are not passive recipients of applications; rather they expend substantial resources soliciting demand from desirable applicants [@RN3519; @RN4774; @RN4324]. Most scholarship on recruiting analyzes the behavior of individual colleges, thereby assuming that recruiting behavior is a function of individual colleges [e.g., @RN4758; @RN4759]. @RN5012 analyze the College Board *Student Search Service* product, highlighting the role that third-party products play in structuring the recruiting behavior of colleges. The authors show that the utilization of particular search filters -- particularly geographic filters, such as zip code -- is associated with systematic racial exclusion that does not depend on which college purchased the list.

However, prior research has not examined the role of third-party market research in reproducing socioeconomic and racial inequality. This manuscript analyzes the Market Segment Model and College Board Geomarkets as a case study of quantification. We develop arguments from the broader literature on quantification [e.g.,] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975. Predictive analytics are based on correlation. When we use data on past correlations to make recommendations about the future, we amplify the effects of historic structural inequality. The Market Segment Model analyzes SAT score-sending data from 1980 and concludes that student demand for higher education is primarily a function of social class= [@RN4982]. Homophily is the idea that actors that share common characteristics are likely to form connections with one another. The Market Segment Model argues that homophily is the organizing principle of student demand and competition between colleges, that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students" [@RN4982, p. 72]. Subsequently, the assumption of homophily is programmed into the EPS and Student Service Service products that colleges utilize to decide where to focus recruiting energy.

Empirically, our analyses address the following research questions, which speak to how geomarkets are utilized within the Enrollment Planning Service and within the Student Search Service student list product. First, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time? We address this question by spatially joining the geomarket shapefile to Census data about socioeconomic and racial characteristics. Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets, in concert with other commonly used search filters? We address this question by analyzing student list purchases that do not filter geomarkets and simulating which prospects would have been included had they filtered on particular geomarkets. We obtained these data by issuing public records requests to public universities in four states. Third, in student list purchases that filtered on geomarkets, how does the socioeconomic and racial composition of included students compare the that of the geomarket and how does it compare to that of the broader metropolitan area?

Scholarship on quantification in education has focused on consumer-facing measures such as U.S. News Rankings [@RN5003] and UK "league tables" about school quality [@RN5005]. To our knowledge, prior knowledge has not examined business-facing measures produced by third-party market research. Here, the desirable attributes of students and places they live are quantified with the purpose of informing colleges about where they should allocate recruiting resources to efficiently enroll the most desirable students. We reason that market research reproduces inequality by encouraging colleges to plan their recruiting efforts based on a snapshot of social stratification in student demand that is substantially a function of historic, structural inequality in educational opportunity. In contrast to @RN5005, we cannot demonstrate the effect of geomarkets on inequality. However, we show the mechanisms of how geomarkets reproduce inequality by simulating who is included/excluded when geomarkets are utilized within student list purchases as recommended by @RN4982.

In the following section we provide background information about enrollment management and scholarship on recruiting. Second, we introduce core ideas and concepts from scholarship on quantification. Third, we apply these ideas to a close read of @RN4982 and motivate hypotheses. Four, we describe data and methods. Fifth, we present results. Finally, we discuss implications for scholarship, policy, and practice.

# Background: Enrollment Management

The term "enrollment management" (EM) refers to a profession, an administrative structure, and an industry. As a profession, EM integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]. As an administrative structure, the EM office typically controls the activities of admissions, financial aid, and recruiting [@RN2406]. The EM industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers that produce market research, software, and and consulting services (e.g., College Board, ACT, EAB, Huron, Slate, PowerSchool).

Figure @fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004]. The funnel begins with a large pool of "prospects" (i.e., prospective students) that the university would like to enroll. "Leads" are prospects whose contact information has been obtained. "Inquiries" are prospects that contact the institution and consist of two types: first, inquiries who respond to an initial solicitation (e.g., email) from the university; and second, "student-as-first-contact" inquiries who reach out to the university on their own (e.g., sending ACT scores). <!--  Applicants consist of inquiries who apply plus "stealth applicants" who do not contact the university before applying.  -->The funnel narrows at each successive stage in order to convey the assumption of "melt" (e.g., a subset of "inquiries" will apply). Practically, the purpose of the enrollment funnel is to inform recruiting interventions that target one or more stages. These interventions seek to increase the probability of "conversion" across stages [@RN4322]. At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz]. Purchased leads are served emails, brochures, and targeted social media advertisements designed to solicit inquiries and applications [@RN4895].

Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel they focus on. The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking]. A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications. Scholarship on recruiting has conceptualized recruiting behavior as an indicator of college enrollment priorities [@RN4758] and identifies recruiting practices as a mechanism for social reproduction in college access [@RN4324; @RN3519].

Ethnographies by @RN3519 and by @RN4407 analyze the connections between privileged colleges and privileged high schools. @RN4407 shows how guidance counselors at elite private school guidance counselors lobby admissions counselors on behalf of marginal students. @RN3519 finds that college admissions officers visit high schools as a means of maintaining positive relationships with guidance counselors at affluent feeder schools. Consistent with this finding, quantitative case studies of off-campus recruiting visits show that selective private colleges visit affluent, predominantly white schools and communities, with a disproportionate number of visits to private high schools [@RN5014; @RN4330]. Additionally, public research universities often make make more out-of-state recruiting visits than in-state visits, and these out-of-state visits focus on the same set sorts of schools targeted by selective private universities [@RN4759; @RN4758].

Other studies investigate recruiting behavior by non-selective institutions [@RN4520; @cottom2017lower; @dache2018dangling]. Analyzing recruiting from the perspective of high school students, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options. @cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities. <!--  WHAT YOU SAID ABOUT POSECZNICK Posecznick (2017) studied the “Ravenwood College,” a private nonprofit college founded to serve working Black women. Consistent with analyses of tuition discounting by Lapovsky (2001), the survival of Ravenwood depends on the success of admissions officers in identifying prospects (e.g., through on- and off-campus recruiting events) and “converting” these prospects to enrolled students.  -->

We make two observations about the recruiting literature. First, consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015] and industrial organization [@RN1549; @RN2133], the industry is characterized by a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches, a process that exemplifies the idea that higher education is mechanism of social reproduction [@RN2377; @RN4873].<!--  For example, for-profit institutions target adults in communities not being targeted by traditional colleges [@cottom2017lower]. Non-selective private non-profits target prospects who might be considering community college or a local state colleges [@RN4324]. Selective private colleges target private high schools, with a particular link between sectarian high schools and colleges of the same faith [@RN5014], but these schools are increasingly targeted by out-of-state public research universities, suggesting growing competition for a coveted market niche. --> Second, by focusing on the behaviors of individual colleges, the literature implicitly states that recruiting is something done by individual colleges. 
<!--  
Why then, does recruiting behavior look so similar across colleges of a particular "type"?[^1]
[^1]: New institutional theory offers one explanation [@RN527]. In highly developed organizational fields like higher education, forces of competitive and institutional isomorphism -- including institutional locics -- compel organizations adopt similar practices as their peers. Practices deemed advantageous spread quickly, for example, public research universities targeting nonresident prospects to compensate for declines in state appropriations. These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.
 -->

<!-- make or buy  
The process of recruiting students involves many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor. EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success. As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns. The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm]. In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
-->

However, the recruiting behavior of colleges is structured by third-party actors. Drawing from scholarship on organizational theory [@RN4738], enrollment management processes involve many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz]. EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success. <!-- As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns. --> The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].

Moreover, process of college access is structured by third-party recruiting products utilized by colleges. @RN5012 deconstruct the College Board *Student Search Service* product. Selection devices are procedures or routines for making categorical decisions based on input values [@RN4778]. Student list products are selection devices that enable colleges to choose which prospective students they target or ignore based on search filters that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students. "Racialized inputs" are ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786]. @RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) are racialized inputs. Using a national sample of high school students, and also using data from actual student lists purchased by public universities, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects. 
 <!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted. -->

Student list products offer many geographic filters. Geographic borders are perhaps the most commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887], owing to extreme racial segregation in the U.S. In addition to offering geographic filters based on known geographic borders (e.g., zip code, county, CBSA, state), the College Board created new geographic borders that subsequently became filters in the *Student Search Service*. The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565]. Geodemographic search filters can be read as the mimicry of corporate market logic [@RN4004] in that they apply the logic and methods of commercial geodemographic market segmentation products, such as PRIZM by the Claritas Corporation [@RN4957], to the case of college recruiting.

Geomarket search filters, the focus of this manuscript, slice states and metropolitan areas into smaller, local recruiting markets. For example, geomarket "CA08" refers to "Alameda Country excluding Oakland." A university might purchase prospect contact information by filtering on this geographic region in combination with additional filters, such as year of high school graduation and SAT test score. When we initiated this manuscript, we asked ourselves, "where did geomarket filters come from?" The answer surprised us. We had imagined ourselves studying the application of the "data science revolution" to college recruiting, but it turned out we were studying the 1970s and 1980s! The Market Segment Model and geomarkets can be read as the application of social science research methods to "capture and quantify the phenomenon behind the folklore" [@RN4982, p. 11] of college admissions officers. This work, comissioned by College Board, was subsequently turned into a market research tool -- The College Board Enrollment Planning Service -- that helps colleges decide where to look for prospective students.

```{=html}
<!--The enrollment management industry emphasizes market research.
Market research is the process of producing information that informs decisions about product design or how to find customers.   -->
```

The enrollment management industry puts great stock in market research. @RN4986[p. 17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].^[@RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college. In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited. In the chapter entitled "Travel," @RN3519 [p. 69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."] However, most market research in the contemporary enrollment management industry is produced by third-party organizations. For example, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development [e.g., @RN4741; @RN4728, but most only available to clients. EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.

Despite being central to enrollment management for at least 40 years [@RN4986], the role of third-party market research has been ignored by scholarship on college access. Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935;@RN1814]. However, market research informs where colleges look for students and which prospective students to target. These recruiting interventions intervene on student demand [@RN4324; @cottom2017lower], thereby structuring college access pathways in ways that could amplify or diminish historic inequality in college access. The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals [@RN4986]. This article offers one story about the quiet revolution that transformed college access into a sophisticated, supply-led enterprise. We argue that market research structures who colleges recruit, in ways that reinforce existing social stratification. The case of Market Segment Model and College Board geomarkets shows how market research -- and its commodification into recruiting products -- is a mechanism for social reproduction. 
[LAST TWO SENTENCES ARE REDUNDANT BUT BUT ARE TRYING TO DO SLIGHTLY DIFFERENT THINGS AND I LIKE THEM BOTH]

# Quantification

```{=html}
<!--
CUT GENERAL PARAGRAPH ON REPRODUCTION

Within the sociology of education, @RN2377 describes education as a sieve, allocating people to opportunities in socially stratified, capitalist societies. This role highlights "Max Weber's core insight that education has a dual character" [@RN2377, p. 129], on one hand, serving social mobility by redistributing access to privileged positions and, on the other, serving social reproduction by granting privileged families disproportionate access to the educational institutions that confer pathways to privileged positions. Similarly, @RN4873 describe schools and colleges as "sorting machines" that incorporate external categories (e.g., race, class), create categories (e.g., instructional tracks), and allocate students to categories. Initial category allocation affects subsequent allocation, such that school assignments and instructional track assignments affect college access, which affects employment assignments. "In evaluating schools as inequality regimes" @RN4873 \[p. 322\] argue, "the relevant question is not whether schools produce inequality; rather, the relevant questions are, when and where do schools produce inequality, \[and\] how do they do so." As @RN2377 show, whereas early research on social stratification sought to identify patterns of inequality [e.g., @RN4873; @RN290; @RN1268], the sociology of education is now more concerned with identifying the mechanisms underlying these patterns. 
-->
```
We situate our study amidst scholarship on within scholarship on quantification as a mechanism that reproduces inequality. @RN5007 [p. 402] define quantification as "the production and communication of numbers" (p. 402) and introduce reacivity, discipline, and authority as three interrelated themes that describe the effects of quantification. Reactivity is the idea that salient quantitative measures cause people and organizations to change their behavior. Discipline is the idea that quantification causes actors to react in a particular way. Authority is the idea that quantification changes decision-making power, often weakening the discretion of local decision-makers.

Scholarship about U.S. News & World Report (USNWR) Law School rankings demonstrate the reactivity, discipline, and authority effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375]. Developed as a means of informing prospective students about quality, rankings had the intended effect of affecting. Rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers. Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p. 415]. Rankings *disciplined* the behavior of law schools. For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003]. Rankings also weakened the *authority* of admissions directors and admissions officers to make independent decisions about which applicants to admit, ideal class size, and curricular offerings [@RN5003]. <!--
In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.    -->

The analysis of UK League Tables by @RN5005 show how quantification can be a mechanism of social reproduction. As part of a broader *New Public Management* effort to infuse market competition within public providers [@RN161], the UK government began publishing school "league tables" in 1992. League tables ranked schools based on student performance on high-stakes national exams taken at age 16. Using Census data on occupational class measured at the local authority level (similar to a U.S. county), @RN5005 find that localities with higher performing schools experienced growth in the share of managerial/professional households following the adoption of league tables and a decline in the share of working-class households. Additionally, using longitudinal household survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables. With respect to mechanisms, league tables facilitate making evaluative comparisons between schools regardless of geographic proximity. They also discipline households and schools to conceive of school quality in terms of test scores, which are substantially a function of the class composition of schools. Compared to working-class parents, parents with professional/managerial occupations are more aware of school league tables and have financial resources necessary to move to neighborhoods near a "good" school. Therefore, league tables contributed to social reproduction because managerial/professional households respond to consumer-facing metrics by moving to more expensive neighborhoods, near higher performing schools. <!-- THIS IS RATCHET EFFECT STUFF. I CUT IT BECAUSE YOU HAVEN'T INTRODUCED THE IDEA OF RATCHET EFFECT
Schools that performed well on initial league tables tended to be schools with a high share of professional/managerial families. Therefore, quantification takes a snapshot of existing social stratification and that snapshot becomes a power unto itself, amplifying the effect of previous strctural inequality on subsequent structural inequality.
-->

```{=html}
<!-- THIS IS THE CUT TRANSITION PARAGRAPH
In the broader quantification literature, critical data studies [e.g., @RN4975; @RN4849; @RN4772] and sociological analyses of technology and inequality [e.g., @RN4778; @RN4786; @RN4877; @RN4810] introduce concepts salient to the analysis of market research. Our focus is on "selection devices," which utilize some combination of inputs and discretion to categorize places, people, and organizations for the purpose of allocating opportunities or interventions [@RN4778; @RN5012].
-->
```
In the interdisciplinary literature on quantification [@RN5008], the discussions of correlation and homophily by @RN4975 introduce concepts that are salient to the analysis of market research. Chapter 1 of @RN4975 is *Correlating Eugenics*. Correlation measures the extent to which two or more variables move together. The Pearson Correlation Coefficient was developed by the 19th Century statisticians and "biometric eugenicists" Francis Galton and his student Karl Pearson [@RN4975]. @RN4975 [p. 50] writes that "correlation grounds big data's so-called revolutionary potential. As Wired editor Chris Anderson infamously declared in his 2008 editorial 'The End of Theory,' big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'" Predictive analytics are developed in two steps. First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest. Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and to assign levels of risk to each case. These models predict outcomes based on correlations without requiring knowledge about underlying causal relationships. @RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.

Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated to a variable of interest. For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of these others. @RN4849 [p. 146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'

Many studies show that using correlational models make predictions reproduces structural inequality. The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time. These correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models. @RN4794 [p. 224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Therefore, disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies disproportionate targeting/exclusion. This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849]. An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849]. Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests. Note that predictive models such as the LSI-R model are not merely used for social science purposes. Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.

Chapter 2 of @RN4975, *Homophily, or the Swarming of the Segregated Neighborhood*, argues that homophily often underlies the use of correlation to predict future behavior. Homophily means that actors who share common characteristics are likely to form connections with one another. The ubiquitous colloquialism is "birds of a feather flock together." Homophily is a core assumption for network science [@mcpherson2001birds], in which actors (e.g., high school students) are linked to one another through direct and indirect network ties (e.g., submit SAT scores to the same college). Because network science often draws from rational choice theory, it assumes that homophily is the result of voluntary action by individuals, thereby "eras\[ing\] historical contingencies, institutional discrimination, and economic realities" [@RN4975, p. 95] that underlie behavior consistent with homophily. @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon. In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.[^4]

[^4]: Thus, @RN4975 [p. 82] writes, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together."

Within market research, homophily is central to geodemography and the creation of market segments. Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878]. Market segments are subgroups within a larger market that have similar consumer demand. Early geodemographic classifications of consumers (e.g., PRIZM by Claritas Corporation) were derived from publicly available Census data, which disaggregated data to the zip code level [@RN4957]. The Claritas Corporation argued -- and had a financial incentive to do so -- that people living near one another share similar consumer preferences. Therefore, geographic localities could be categorized into market segments that would be useful for direct mail marketing campaigns. Later, the development of individual credit scores (e.g., FICO score) enabled merchants to classify consumers into many, fine-grained groups [@RN4877]. @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups. Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015]. For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether. Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers. <!--
ORIGINAL TEXT ABOUT GEODEMOGRAPHIC CLASSIFICATION SYSTEMS

Clusters were given evocative names like "Blue Blood Estates," "\[Archie\] Bunker's Neighbors", and "Share Croppers." Because U.S. publicly available U.S. Census data disaggregated data only to the zip code-level, the Claritas Corporation had a strong incentive to argue that consumer behavior within zip codes was very similar [@RN4957]. @RN4565 emulated the behavior of Claritas by developing the *Segment Analysis Service: An Educationally Relevant Geodemographic Tagging Service*, which was an "add-on" to the *Student Search Service* student list product. Describing its conceptual underpinnings, @RN4565 [p. 1] states, "the basic tenet of geodemography is that people with similar cultural backgrounds, means, and perspectives naturally gravitate toward one another or form relatively homogeneous communities; in other words, birds of a feather flock together." The Segment product allocated each high school and each census tract into a distinct "educational neighborhoods" thought to reflect useful market segments for colleges looking for students.

PREVIOUS TEXT ABOUT CLASSIFICATION SITUATIONS
Related to the creation of market segments, @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups. Historically, classifications were binary. Consumers with "good" credit were offered loans and those with bad credit were not. The innovation of individual FICO scores -- and, later, other credit scores -- enabled merchants to classify consumers into many, fine-grained groups or along a continuum. Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015]. People previously denied credit were now granted credit on different terms. For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether. Thus, at the bottom of the continuum, contemporary classification situations produce "predatory inclusion" -- exemplified by payday loans and for-profit colleges -- which is the logic of "including marginalized consumer-citizens into ostensibly democratizing mobility schemes on extractive terms" [@RN4774, p. 443].

Negative consequences of homophily
Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
-->

Reviewing recent scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers. Law school rankings [e.g., @RN5003] and school league tables [e.g., @RN5005] seem to exemplify this sort of salience. Widely known studies of quantification in education tend to study consumer-facing quantification. To our knowledge, prior research on quantification in education has not examined how third-party market research classifies consumers. This manuscript analyzes "market segment model" [@RN4982], which categorized high school students into vertical market segments and simultaneously created local "geo-markets" that were evaluated based on their composition of student market segments. The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target. @RN5021 reports that in 1995, 37% of 4-year public universities and 49% of 4-year private universities used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS. Geo-markets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities. <!--
CUT COMMENTS ON SALIENCE
By contrast, many quantification projects -- for example, human rights indicators -- have "little oomph" [@RN5006, p. 265]. FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created." 
--> Scholarship on predictive analytics [@RN4975; @RN4849] suggests the mechanisms by which market research reproduces structural inequality. Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification. Unlike studies like @RN5005 we cannot show the effect of quantification. However, this study provides inisight about the mechanism underlying the effect. By analyzing and simulating *Student Search Service* purchases that filter on geo-markets we can show *how* the geo-markets reproduce historical class-based and race-based inequality in college access.

# The Market Segment Model and College Board Geomarkets

 <!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.-->
As a member organization, several College Board products developed from joint collaborations with administrators or professors from member institutions. During the 1970s, under the leadership of Larry Litten, then director of institutional research, Carleton College pionered applying market research to student recruiting. In the late 1970s, College Board provided Carleton College with funding and data for a six-year project that analyzed college choice decisions by households in six metropolitan areas. This project was the basis for the landmark book *Applying Market Research in College Admissions* [@RN4986], published by College Board, and was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll. Litten would later become Associate Director of the Consortium of Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.

In 1978, Robert Zemsky, a University of Pennsylvania professor, was asked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p. x]. To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE). As the project project became more ambitious, the researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in tehse analyses was the state' [@RN4982, p. x]. In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.^[say that Pennsylvania Department of Education was also a partner; see page x of Zemsky] @RN4982[p. 4] write that,

> For our analysis, we sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice. The Market Segment model that we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups. The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways.

**Creating the Market Segment Model.** Considering the @RN5007 \[p. 402\] definition of quantification as "the production and communication of numbers," the Market Segment Model is an effort to quantify and communicate the knowledge that admissions officers have about student demand. @RN4982[p. 4] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." They created  "three types of boundaries" [@RN4982, p. 11], region, state, and community. The three regions were New England, Middle States, and the South. Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" \[in New England, Middle States, and the South\].

These enrollment markets, later called "geomarkets" by the College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors. We know little about how geomarkets were created. @RN4982 write that,

> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication. Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs. In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."

Having defined these geographic boundaries, high school students were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior. Each institution a student sends SAT scores to can be defined as "local" (institution is in the same local market that the student lives in), "in-state" (same state but different local market), "regional" (same region but different state), or "national." In turn, a "local" student submits more SAT scores to local institutions than they do to in-state, regional, or national institutions. An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc. <!--
A "regional" student submits more SAT scores to regional institutions than they do to local, in-state, or national institutions. A "national" student submits more SAT scores to national institutions than they do to local, in-state, or regional institutions.
-->

The two primary outputs of the market segment model are the (1) Market Segment Profile and (2) the Institutional Profile. The Market Segment Profile, which was produced separately for each of the 143 geo-markets, provides information about the number and characteristics of students in each market segment. @tbl-market-segment-characteristics reproduces a partial, simplified version of the bottom panel of @RN4982 [Figure 2.1] which represents the Market Segment Profile for "Connecticut Market 3: Fairfield County." @tbl-market-segment-characteristics has separate columns for each market segment (local, in-state, regional, national) and rows show the number of test-takers and characteristics of test-takers. For example, there were 550 "local" students in Fairfield county and these students submitted SAT scores to 2.8 postsecondary institutions on average. By contrast, there were 1,664 "regional" students who submitted test scores to 4.8 institutions on average. For local students, 11.2% had family income greater than \$35,000 and 9% had both parents with a BA. For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA. Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.[^5]

[^5]: @tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market. For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution. Of the 1,199 in-state students, 757 (63.1) sent scores to institution #1, a public doctoral granting institution, and 515 (43.0%) sent scores to institution #2, a public master's granting institution.

INSERT @tbl-market-segment-characteristics ABOUT HERE

The *Institutional Profile* describes students who send scores to a particular institution and which majors these students are interested in. For each institution, institutional profiles are created separately for students from a particular local market, for all students in a state, or all students in a region. @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the institutional profile of an anonymous institution for students from Fairfield County, CT. @tbl-sample-institutional-profile shows that 58 in-state students submitted SAT scores to the institution. These 58 students represented 4.8% of the total 1,199 in-state students from Fairfield County. 69 regional students sent scores to the institution, representing 4.1% of all 1,664 regional test-takers. Of these 69 regional students, 35 expressed interest in majoring in the liberal arts. These 35 students represent 7.0% of all regional students from Fairfield County who expressed interest in the liberal arts.

INSERT @tbl-sample-institutional-profile ABOUT HERE

@RN4982 argue that the Institutional Profile and the Market Segment Profile enable admissions officers to know where to look for students and which institutions are competing for those students (p. 25):

> The Institutional Profile and the Market Segment Profile quantify the admission officers' intuitive grasp of market structure. Structure here carries a dual meaning, connoting both the structure of student choice and the structure of institutional competition\ldots This two-sided interpretation furnishes the essential framework for planning by individual colleges and universities\ldots To draw effectively on its own natural constituency, a college not only must contact the "right" kind of students --- that is, students who are predisposed toward that type of institution -- but also must persuade them of its special character. This means knowing the competition as well as the clientele.

**Correlations**. Chapter 3 of @RN4982 -- "A Sense of Place: Students, Families, and Communities" -- identifies the student characteristics associated with being in the local, in-state, regional, or national market segment and examines the extent to which these relationships hold across states and local markets. The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination. [@RN4982, p. 33] state that these four variables "reflect the basic social patterns of the nation. It would have been surprising if these were not the four social variables that best explained the patterns of college choice."

@RN4982 summarize their correlational analyses using authoritative language. @RN4982[pp. 34-35] asserts, "it is hard to overemphasize the importance of this statistical pattern. Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice." With respect to family income, @RN4982[p. 33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982[p. 33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time." Further analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations. Considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model implicitly and explicitly states that student demand for higher education is mostly a function of social origin. @RN4982 also find that the SAT score, parental education, and family income predict student score-sending behavior at the local geo-market level.

However, geomarkets differ in the relative abundance of students with particular socioeconomic characteristics and this has practical implications for recruiting. @RN4982 recommend that colleges target geo-markets with desirable compositions of socioeconomic characteristics and student market segment. @RN4982[p. 44] describes a hypothetical college that wants to target regional and national students in New England and is considering whether to expend limited recruiting efforts in the Boston geo-market -- which was relatively low-income in 1980 -- or in more affluent geo-markets nearby:

> Where would you concentrate your energies? Ideally, you would seek communities with a high proportion of students already predisposed toward institutions such as your own. The Market Segment Model would provide this information through segment percentages for the community in question. Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots. Your efforts would surely be better directed toward three of the four communities in the bottom band, Manchester, Hartford, and Fairfield County. Simply by knowing a little bit about the students' backgrounds and academic records you could quickly focus your attention on those most likely to consider your kind of institution. Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own."

**EPS and College Board Geomarkets**. In 1984, College Board created the Enrollment Planning Service (EPS) based on the Market Segment Model [@RN4987; @RN4980]. Each institution purchasing EPS information services would obtain access to the Market Segment Report for each local market. They also received the Institutional Profile -- their own and that of competitors -- for each local market. Promotional material highlights the themes of "locate the students you want" [@RN5019_2003] and "focus your valuable time and resources on the right prospects" [@RN5019]. @RN5019_2010 promotional material describes EPS as:

> the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found. You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, Zip codes, high schools and international regions. EPS provides you with comprehensive reports on your markets, your position in those markets, and your copetition."

```{=html}
<!--
@RN5019 [2005]
> The marketing software that pinpoints the schools and geomarkets where your best prospects are most likely to be found. With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition. Focus your valuable time and resources on the right prospects

Software documentation by @RN5017 states that "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States. Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
-->
```
Whereas @RN4982[Appendices B1-B3] identified 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S. states, with 304 geo-markets in total. @RN4988 shows the College Board geo-markets circa 2023. Interestingly, the geo-markets for New England, Middle States, and the South remain nearly identical to those developed by @RN4982. Concrete information about how EPS created the remaining geomarkets is elusive. Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^7]

[^7]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population. Students from California don't all share the same college-going behaviors. We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."

Although how we lack clarity about how geomarkets were created, we expect substantial socioeconomic inequality between geomarkets in large metropolitan areas. Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality. Additionally, @RN4982 created geomarkets in service of the Market Segment Model. The Market Segment Model argues that socioeconomic status drives whether students belong to the local, in-state, regional, or national market segment. In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status. The language of @RN4982[pp. 11-12, quoted above] suggests that geomarket borders within large metropolitan areas may have been drawn in ways that separate affluent from less affluent communities. We present the following hypothesis:

**H1**. In large metropolitan areas that contain multiple geomarkets, we expect significant socioeconomic inequality (e.g., income, parental education) between geomarkets.

We expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons. Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race once. However, U.S. cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779; @rothstein2017color]. Structures built upon racialized structures are racialized structures. @RN4975 argues that unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity. Second, geomarket borders may have been drawn along class divides. There is a strong correlation between race and wealth [@RN4884]. Third, geo-market borders may have been drawn in a way that follows the countours of racial segregation in residential housing. Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets. Although these arguments are speculative, the extent of racial inequality between geomarkets can be assessed empirically. We present the following hypothesis:

**H2**. In large metropolitan areas that contain multiple geomarkets, we expect significant racial inequality (e.g., income, parental education) between geomarkets.

We argue that the Market Segment Model, geomarkets, and their commodification within EPS exemplify concerns about the application of correlational analysis. The data on 1980 SAT score-sending behavior analyzed by @RN4982 can be conceived as training data. These training data were used to define geomarkets (e.g., how many geomarkets in a metropolitan area is partially a function of the number of test-takers) and to identify student characteristics associated with the local, in-state, regional, and national market segments. These analyses found that student demand for different kinds of colleges was largely a function of social origin, and that student demand within a geo-market was largely a function of the class composition of the geo-market. However, @RN4982 did not interrogate the historical and contemporary structural inequalities that produced observed patterns of social stratification. Instead, the Market Segment Model encouraged colleges to use this snapshot of student demand as the basis for future decisions about which communities and prospective students should be targeted by college recruiting campaigns. This process exeemplifies the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849] whereby predicting the behavior of future individuals based on the behavior of past individuals "threatens to reify and reporduce existing inequalities" [@RN4794, p. 224]. This process also exemplifies concerns about proxy variables whereby selective colleges view privileged social class as a proxy for desirable students and this creates an incentive to focus recruiting efforts on communities that have these characteristics. @RN4982 [p. 44] corroborates these concerns:

> On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged. Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good. The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.

**Homophily and Utilization**. *The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] shows that homophily is a core tenet of the Market Segment Model. The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^8] @RN4982[p. 42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^9] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another. Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status. @RN4982[p. 72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:

[^8]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2). Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another. These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.

[^9]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p. 54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector. For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."

<!--[^10]: These patterns are similar to different offers for credit products in The Ordinal Society [@RN5015], except @RN5015 show that the matching of consumers to products is engineered by firms rather than naturally occurring.-->

> We now know why students so seldom speak of their own social or family backgrounds in explaining how they go about choosing a college. They have no need to. Students describe themselves socially simply by telling us the colleges and universities in which they are interested. The layering of collegiate competition is primarily a socioeconomic layering. The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students. Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.

Although this perspective is broadly consistent with mid-20th Century scholarship on social mobility [@RN915; @RN1273], the Market Segment Model was the basis for the EPS market research product. Later, geomarkets were incorporated into the *Student Search Service* student list product. Commodification of the Market Segment Model "engineers homophily" [@RN4975] by taking a snapshot of existing stratification and then recommending that colleges utilize this snapshot as the basis for decisions about where to concentrate recruiting efforts. @RN4982 [p. 58] states that "correlations, do not simply predict actions; they also form them." In itself, the Market Segment Model is simply a stylized depiction [HIRSCHMAN 2016 STYLIZED FACTS] of the structure of college choice, one that does not consider class-based and race-based structures that produced these patterns. Once the College Board inscribed the Market Segment Model into the EPS and Student Search Service Products, these products are likely to amplify the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.

Utilization of the Market Segment Model and geomarkets within the Enrollment Planning Service (EPS) raises concerns related to the quantification themes of reactivity, discipline, and authority introduced [@RN5007]. @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers. Although EPS is a business-facing rather than a consumer-facing product, it has been a highly salient one. @RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product. We suggest that EPS disciplines colleges to approach recruiting in a manner consistent with the Market Segment Model. Colleges are encouraged to, first, choose geomarkets with a high number of desirable prospects and, second, plan recruiting efforts within selected geomarkets. Under the principle of homophily, selective colleges should target geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges. Consistent with this perspective, EPS software documentation from @RN5017 describes how geomarkets -- here, called "EPS market codes" -- can be utlized:

> EPS\textsuperscript{\texttrademark} (Enrollment Planning Service) is a geographic and demographic data service offered annually by the College Board to Colleges and Universities. EPS provides information to subscribing institutions about competitors, feeder schools, and demographic strengths and weaknesses. EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States. Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful.

EPS may also weaken the authority of local decision-makers. The Market Segment model sought to develop a concrete, data-driven framework that replicated the aggregate knowledge of local admissions officers [@RN4982]. Once this knowledge is quantified and commodified into a product stored on a CD-ROM, the local expertise of admissions officers becomes less valuable. The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally. In turn, admissions officers are relegated to a foot-soldier implementation role rather than being sought out for their local expertise about student demand.

Finally, we suggest that utilization of geomarket filters within the Student Search Service product is likely to amplify class-based and race-based disparities in college opportunity. In contrast to standardized selection devices (e.g., the FICO score algorithm), student list products are discretionary selection devices in that colleges have discretion of which filters they elect to select prospects. Compared to standardized selection devices, scholars tend to find that administrative discretion over selection devices causes structural inequality to increase [@RN4883; @RN4881; @RN4795]. Even in the absence of explicit or implicit bias, discretionary selection devices can increase racial inequality when decision-makers have incomplete knowledge about how the inputs they choose are correlated with race. For example, research shows that Americans dramatically underestimate the magnitude of racial income inequality [@RN4884]. As such, discretionary selection devices are particularly sensitive to inputs that are correlated with race but are perceived to be race-neutral.

Geomarkets are opaque in that geomarket borders may be unclear to someone purchasing student lists and very few people know how these borders were created. @RN4849 argues that opacity in algorithmic products is intentional and is a core criterion of "weapons of math destruction." Enrollment management professionals utilizing EPS and *Student Search Service* may interpret geomarkets as a race-neutral recruiting tool. When purchasing student lists, @RN4982 [p. X] suggest that affluent colleges should consider targeting affluent geomarkets. Student list purchases do not show how the characteristics of targeted prospects compare to the characteristics of the surrounding community. From this perspective, racial inequality from the use of geomarkets in student list purchases may be the unintentional result of using a seemingly race-neutral product. Furthermore, purchases that specify multiple filters (e.g., test scores, grades, intended major) including geomarkets can yield unintended racial inequality because administrators have incomplete knowledge about how the intersection of these filters interact with local patterns of segregation.

<!-- CUT TEXT DEFINING COLORBLIND RACISM AND RACE-NEUTRAL RACISM
Colorblind racism describes ideology and associated behaviors that ignore the significance of race and racial discrimination [@RN5020]. The Market Segment Model can be seen as an example of colorblind racism. @RN4982 develop an analysis of the structure of college choice that ignores race but we contend that geo-markets are correlated with race. Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975]. 
-->

 <!-- CUT PARAGRAPH: WHY GEOMARKETS ARE WEAPONS OF MATH DESTRUCTION
The incorporation of geo-markets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction. First, geo-markets are opaque. The borders of a geomarket may be unclear to someone purchasing student lists and very few people know how these borders were created \[COULD TALK ABOUT PRODUCT UTILIZATION HERE; NOT CLEAR HOW GEOMARKETS INTERACT WITH OTHER STUDENT LIST FILTERS\]. Second, geo-markets scale. Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on geo-markets in concert with other filters (e.g., SAT score, GPA). Third, geo-markets can cause damage. When a college decides to purchase names from one geo-market but exclude prospects from an adjacent geo-market, this affects which college students have an opportunity to attend. @RN4982 suggests that selective colleges focus on geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-->

# Methods

Below, we use simulations and actual student list purchases that filtered on geomarkets. In purchases that filtered on test-score and/or GPA thresholds, we can simluate who would be included and excluded had certain geo-markets were been selected. In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.

# Discussion

In their analysis of quantifying school quality in England, @RN5005 [p. 517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model argues that student demand for higher education is mostly a function of social origin.

# References

::: {#refs}
:::

\clearpage

<!-- TABLES -->

```{r fig-em-funnel, echo = FALSE, fig.align = 'center', fig.cap = "The Enrollment Funnel", out.width = "45%"}
knitr::include_graphics('./../assets/images/enroll_funnelv2.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

\clearpage

<!-- TABLES -->

```{r market-segment-characteristics, echo = F}
#| label: tbl-market-segment-characteristics
#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'

market_segment <- data.frame(
  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
)

kable(
  x = market_segment,
  col.names = c('Characteristic','Local','In-state','Regional','National'),
  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
)
```

\clearpage

```{r sample-institutional-profile, echo = F}
#| label: tbl-sample-institutional-profile
#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'

sample_inst <- data.frame(
  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
  local = c(1,0.2,0,0.0,0,0.0),
  in_state = c(58,4.8,25,8.4,2,3.0),
  regional= c(69,4.1,35,7.0,5,5.8),
  national = c(109,2.9,61,5.3,3,0.8),
  total = c(237,3.3,121,5.9,10,1.8)
)

kable(
  x = sample_inst,
  col.names = c('','Local','In-state','Regional','National','Total'),
  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
)
```

\clearpage

```{r market-segment-institutions, echo = F}
#| label: tbl-market-segment-institutions
#| tbl-cap: 'Top 5 institutions in terms of number of scores sent by segment, Connecticut Market 3: Fairfield County'

market_segment_inst <- data.frame(
  rank = c(1,2,3,4,5),
  local_num_scores_sent = c(277,261,183,103,100),
  local_pct_scores_sent = c(50.4,47.5,33.3,18.7,18.2),
  local_inst_type = c('priv ma','priv ma','priv ma','pub doct','pub ma'),
  state_num_scores_sent = c(757,515,438,183,177),
  state_pct_scores_sent = c(63.1,43.0,36.5,15.3,14.8),
  state_inst_type = c('pub doct','pub ma','pub ma','pub ma','pub ma'),
  regional_num_scores_sent = c(610,348,272,248,197),
  regional_pct_scores_sent = c(36.7,20.9,16.3,14.9,11.8),
  regional_inst_type = c('pub doct','priv doct','priv doct','pub doct','pub doct'),
  national_num_scores_sent = c(1226,371,327,312,308),
  national_pct_scores_sent = c(32.6,9.9,8.7,8.3,8.2),
  national_inst_type = c('pub doc','priv doct','priv res','priv doct','priv doct')
)

kable(
  x = market_segment_inst,
  col.names = c('','Num','Pct', 'Type','Num','Pct','Type','Num','Pct','Type','Num','Pct','Type'),
  #caption = 'Top 5 institutions in terms of number of scores sent, by segment \\\\Connecticut Market 3: Fairfield County'
) %>% 
  #kable_styling(full_width = F) %>% 
  add_header_above(c(' '=1,'Local (N=550)' = 3, 'In-state (N=1,199)' = 3,'Regional (N=1,664)' = 3, 'National (N=3,766)' = 3))
```

### 2024-07-08 text/old_spring2024/background_em_long_theory_5_30_2024.qmd
initial commit, copying files from messed up repo

--- 2024-07-08 text/front_end.qmd
@@
 ---
-title: "Structuring College Access: The Market Segment Model and College Board Geo-markets"
+title: "Structuring College Access: College Board Geo-markets and the Market Segment Model"
 author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
@@
     #keep_tex: true
-    extensions: [andrewheiss/wordcount]
 editor: visual
@@
 library(kableExtra)
+```
 
-#hey ho This many words: {{< words-total >}}
-```
+# Introduction
+
+hey ho This many words: {{< words-total >}}
+
+COULD START WITH HISTORICAL VIGNETTE OF ZEMSKY PARTNERING WITH COLLEGE BOARD TO CREATE THE MARKET SEGMENT MODEL.
+
+-   quotes on "admissions officers are invariably tellers of stories" and the goal of the market segment model is to quantify these stories.
+
+TEXT TEXT
+
+Thinking about a revised outline, 4/19/2024
+
+-   Intro/lit review
+-   theory \[e.g., quantification, whatever\]
+-   then do a close read of the zemsky book, which ends in creation of EPS, incorporating geomarkets in student list products
+    -   hypotheses would be motivated here
+-   methods/data
+
+the alternative (typical education journal) structure would be this \[which has one more level 1 heading and feels to inefficient\]:
+
+-   introduction \[topic importance\]
+-   literature review
+-   background
+-   conceptual framework
+-   methods data
+
+IMPORTANT (defense/offense): One important limitation of this manuscript is that we do not study the effect of geomarkets on college access.
+However, we do show empirically how geomarkets work as a mechanism of exclusion and social reproduction.
+
+Prior research on quantification in education focuses on consumer-facing products.
+For example, Espeland and Sauder study US News Rankings.
+McArthur, D., & Reeves, A.
+(2022) study School League Tables.
+Prior research has not studied quantification within the context of market research.
+Here, the desirable attributes of students and places they live are quantified with the purpose of informing colleges about where they should allocate recruiting resources to efficiently enroll the most desirable students.
 
@@
 ```
-
-# Introduction
-
-In 1983, Zemsky and Oedel authored *The Structure of College Choice*, published by *The College Board.* This book develops the Market Segment Model and introduces local "community-based enrollment markets" (p.14) -- later called "geomarkets" -- which are the basis for the College Board market Enrollment Planning Service (EPS), a market research product.
-@RN4982[pp., 9-10] write that the Market Segment Model sought to quantify the knowledge that college admissions officers have about student demand:
-
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting\ldots We believe that the institutions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research, thus, is based on listening carefully to what admissions officers have to say."
-
-@RN4982 [p. 10] stress the importance of "knowing the territory":
-
-> A good recruiter knows where to look for prospective applicants, as seen in the students' willingness or eluctance to travel.
-It is necessary to identify not only the most promising communities or poools, but also the specific neighborhoods within those communities -- hence the recruiter's emphasis on feeder high schools.
-From the beginning, we have sought to mold our research to these concepts, to capture and quantify the phenomenon behind the folklore.
-Thus, it is the admissions officers' notion of admissions pools that dictates the geographic units of our analysis." \[p.
-11\]
-
-The Market Segment Model was created by analyzing College Board data about which colleges SAT test-takers from 1980 sent scores to.
-High school students who sent scores mostly to colleges in their local geomarket were defined as "local" students.
-"In-state" students were those who sent most scores to colleges outside their local market but in their home state.
-"Regional" students were those who sent most scores to colleges outside their home state but in their region (e.g., New England, Midwest).
-"National" students were those who sent most scores to colleges outside their region.
-Having categorized students, each local geomarket could be evaluated by the number of students from each of the four market segments, yielding the Market Segment Profile (see @tbl-market-segment-characteristics).
-Next, for each local geomarket, the Institutional Profile (see @tbl-sample-institutional-profile) the number of students from each market segment that sent scores to a particular college, which could be your college or a competitor.
-
-Collectively, the quantitative outputs of the Market Segment Model provided intelligence about where are the desirable students and which students might consider your college or a competitor.
-The Market Segment model and geomarkets became the basis or the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
-@RN5019_2010 says "Enrollment Planning Service (EPS\textsuperscript{\textregistered}) the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found." @RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
-
-In 19XX, geomarkets were incorporated into the College Board Student Search Service student list product, which College Board created in 1972 [@belkin2019-studata].
-Student lists contain the contact information of prospective students, the basic input for recruiting campaigns that target individual prospects using mail, email, and targeted social media.
-RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
-The Student Search Service database consists of College Board test-takers.
-Colleges pay a fee for each prospect (e.g., \$0.50 in 20XX) and control which prospect profiles through purchase by selecting search filters, such as high school graduation year, SAT score, AP test score, state, etc.
-Geomarket filters enable colleges to include/exclude prospects from particular geomarkets.
-
-Scholarship in the sociology of education describes schools and colleges as "sorting machines" that reproduce inequality by incorporating, creating, and allocating students to different categories [@RN4873].
-Given the consequences subsequent outcomes, the sociology of education is particularly concerned with the transition from high school to college.
-Over the last decade or so, scholarship on college access has begun to take more seriously the idea that colleges are not passive recipients of applications; rather they expend substantial resources soliciting demand from desirable applicants [@RN3519; @RN4774; @RN4324].
-Most scholarship on recruiting analyzes the behavior of individual colleges, thereby assuming that recruiting behavior is a function of individual colleges [e.g., @RN4758; @RN4759].
-@RN5012 analyze the College Board *Student Search Service* product, highlighting the role that third-party products play in structuring the recruiting behavior of colleges.
-The authors show that the utilization of particular search filters -- particularly geographic filters, such as zip code -- is associated with systematic racial exclusion that does not depend on which college purchased the list.
-
-However, prior research has not examined the role of third-party market research in reproducing socioeconomic and racial inequality.
-This manuscript analyzes the Market Segment Model and College Board Geomarkets as a case study of quantification.
-We develop arguments from the broader literature on quantification [e.g.,] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
-Predictive analytics are based on correlation.
-When we use data on past correlations to make recommendations about the future, we amplify the effects of historic structural inequality.
-The Market Segment Model analyzes SAT score-sending data from 1980 and concludes that student demand for higher education is primarily a function of social class= [@RN4982].
-Homophily is the idea that actors that share common characteristics are likely to form connections with one another.
-The Market Segment Model argues that homophily is the organizing principle of student demand and competition between colleges, that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students" [@RN4982, p.
-72].
-Subsequently, the assumption of homophily is programmed into the EPS and Student Service Service products that colleges utilize to decide where to focus recruiting energy.
-
-Empirically, our analyses address the following research questions, which speak to how geomarkets are utilized within the Enrollment Planning Service and within the Student Search Service student list product.
-First, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
-We address this question by spatially joining the geomarket shapefile to Census data about socioeconomic and racial characteristics.
-Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets, in concert with other commonly used search filters?
-We address this question by analyzing student list purchases that do not filter geomarkets and simulating which prospects would have been included had they filtered on particular geomarkets.
-We obtained these data by issuing public records requests to public universities in four states.
-Third, in student list purchases that filtered on geomarkets, how does the socioeconomic and racial composition of included students compare the that of the geomarket and how does it compare to that of the broader metropolitan area?
-
-Scholarship on quantification in education has focused on consumer-facing measures such as U.S.
-News Rankings [@RN5003] and UK "league tables" about school quality [@RN5005].
-To our knowledge, prior knowledge has not examined business-facing measures produced by third-party market research.
-Here, the desirable attributes of students and places they live are quantified with the purpose of informing colleges about where they should allocate recruiting resources to efficiently enroll the most desirable students.
-We reason that market research reproduces inequality by encouraging colleges to plan their recruiting efforts based on a snapshot of social stratification in student demand that is substantially a function of historic, structural inequality in educational opportunity.
-In contrast to @RN5005, we cannot demonstrate the effect of geomarkets on inequality.
-However, we show the mechanisms of how geomarkets reproduce inequality by simulating who is included/excluded when geomarkets are utilized within student list purchases as recommended by @RN4982.
-
-In the following section we provide background information about enrollment management and scholarship on recruiting.
-Second, we introduce core ideas and concepts from scholarship on quantification.
-Third, we apply these ideas to a close read of @RN4982 and motivate hypotheses.
-Four, we describe data and methods.
-Fifth, we present results.
-Finally, we discuss implications for scholarship, policy, and practice.
-
 # Background: Enrollment Management
@@
 The EM industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers that produce market research, software, and and consulting services (e.g., College Board, ACT, EAB, Huron, Slate, PowerSchool).
+
+The process of recruiting students involves many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor.
+EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
+As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
+The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
+In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
 
@@
 Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel they focus on.
-The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
+The majority of scholarship focuses on the admissions stage, for example which admissions criteria are used [@RN4874; @RN4993] and which applicants are admitted [@RN4131; @RN3522] \[CITE SOMETHING BY BASTEDO\].
+
 A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
@@
 
-Other studies investigate recruiting behavior by non-selective institutions [@RN4520; @cottom2017lower; @dache2018dangling].
+Other studies have investigated recruiting behavior by non-selective institutions [@RN4520; @cottom2017lower].
 Analyzing recruiting from the perspective of high school students, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
-@cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.
+@cottom2017lower and @dache2018dangling show that for-profit colleges found a niche in Black and Hispanic communities because traditional colleges ignored these communities.
 <!--  WHAT YOU SAID ABOUT POSECZNICK Posecznick (2017) studied the “Ravenwood College,” a private nonprofit college founded to serve working Black women.
@@
 We make two observations about the recruiting literature.
-First, consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015] and industrial organization [@RN1549; @RN2133], the industry is characterized by a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches, a process that exemplifies the idea that higher education is mechanism of social reproduction [@RN2377; @RN4873].<!--  For example, for-profit institutions target adults in communities not being targeted by traditional colleges [@cottom2017lower].
+First, consistent with scholarship on political economy [@RN4517; @cottom2017lower] \[cite Stevens; cottom\] and industrial organization \[cite Winston\], the recruiting literature suggests a hierarchical competition in which different "types" of postsecondary institutions target hierarchically arranged customer niches.
+For example, for-profit institutions target adults in communities not being targeted by traditional colleges [@cottom2017lower].
 Non-selective private non-profits target prospects who might be considering community college or a local state colleges [@RN4324].
 Selective private colleges target private high schools, with a particular link between sectarian high schools and colleges of the same faith [@RN5014], but these schools are increasingly targeted by out-of-state public research universities, suggesting growing competition for a coveted market niche.
---> Second, by focusing on the behaviors of individual colleges, the literature implicitly states that recruiting is something done by individual colleges.
-<!--  
+This process of hierarchically organized colleges matching to hierarchical student segments exemplifies the idea that higher education is mechanism of social reproduction [@RN2377; @RN4873]
+
+Second, by focusing on the behaviors of individual colleges, the literature makes the implicit assumption that recruiting is something done by individual colleges.
 Why then, does recruiting behavior look so similar across colleges of a particular "type"?[^1]
+
 [^1]: New institutional theory offers one explanation [@RN527].
@@
 Practices deemed advantageous spread quickly, for example, public research universities targeting nonresident prospects to compensate for declines in state appropriations.
-These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.
- -->
+These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.<!--WHAT DOMINA (2017) SAID ABOUT LOGICS, P.
+322: Like other distinctively modern institutions, contemporary schools are governed by standardized and rationalized institutional logics (Baker 2014, Bowker & Star 2000, Timmermans & Epstein 2010, Meyer 1977).
+These logics have profound implications for educational category assignment processes, leading educators, policy makers, scholars, and others to debate the universality of selection-->
 
-<!-- make or buy  
-The process of recruiting students involves many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor.
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
--->
+@RN5012 argues that the process of college access is structured by third-party recruiting products utilized by colleges.
+@RN5012 deconstructs the College Board *Student Search Service* product.
+Student list products are "selection devices" [@RN4778] that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
+Drawing from @RN4786, @RN5012 argue that several search filters (e.g., AP test score, zip code, geodemographic segment) are "racialized inputs," defined as ostensibly race-neutral inputs that are correlated with race because disadvantaged racial groups have historically been excluded from the input.
+Using a national sample of high school students from the High School Longitudinal Study of 2009 (HSLS), they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
+Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
 
-However, the recruiting behavior of colleges is structured by third-party actors.
-Drawing from scholarship on organizational theory [@RN4738], enrollment management processes involve many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-<!-- As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
---> The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-
-Moreover, process of college access is structured by third-party recruiting products utilized by colleges.
-@RN5012 deconstruct the College Board *Student Search Service* product.
-Selection devices are procedures or routines for making categorical decisions based on input values [@RN4778].
-Student list products are selection devices that enable colleges to choose which prospective students they target or ignore based on search filters that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
-"Racialized inputs" are ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
-@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) are racialized inputs.
-Using a national sample of high school students, and also using data from actual student lists purchased by public universities, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
- <!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
--->
-
-Student list products offer many geographic filters.
-Geographic borders are perhaps the most commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887], owing to extreme racial segregation in the U.S.
-In addition to offering geographic filters based on known geographic borders (e.g., zip code, county, CBSA, state), the College Board created new geographic borders that subsequently became filters in the *Student Search Service*.
-The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
-Geodemographic search filters can be read as the mimicry of corporate market logic [@RN4004] in that they apply the logic and methods of commercial geodemographic market segmentation products, such as PRIZM by the Claritas Corporation [@RN4957], to the case of college recruiting.
-
-Geomarket search filters, the focus of this manuscript, slice states and metropolitan areas into smaller, local recruiting markets.
-For example, geomarket "CA08" refers to "Alameda Country excluding Oakland." A university might purchase prospect contact information by filtering on this geographic region in combination with additional filters, such as year of high school graduation and SAT test score.
-When we initiated this manuscript, we asked ourselves, "where did geomarket filters come from?" The answer surprised us.
-We had imagined ourselves studying the application of the "data science revolution" to college recruiting, but it turned out we were studying the 1970s and 1980s!
-The Market Segment Model and geomarkets can be read as the application of social science research methods to "capture and quantify the phenomenon behind the folklore" [@RN4982, p.
-11] of college admissions officers.
-This work, comissioned by College Board, was subsequently turned into a market research tool -- The College Board Enrollment Planning Service -- that helps colleges decide where to look for prospective students.
+[NEED CLEANER TRANSITION]Two geographic search filters -- geo-market and geodemographic segment -- that @RN5012 observe in lists purchased by public could not recreate using HSLS.
+Using geographic borders to determine selection for an intervention has been widely studied as a mechanism of structural inequality because of historic and contemporary racial segregation in American communities and schools [e.g., @RN4775; @RN4849; @RN4779].
+The geo-market and geodemographic segment search filters of the College Board *Student Search Service* are examples of creating new geographic borders for the purpose of targeting students for educational opportunities.
+The geo-market filter slices states and metropolitan areas into smaller, "local" markets.
+For example, CA08 refers to "Alameda Country excluding Oakdland" and a university might purchase prospect contact information by filtering on this geographic region, alongside additional filters.
+The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors.
+This manuscript focuses on geo-markets.
+The first question we asked ourselves was, "where did these geo-markets come from?" The answer is, market research.
 
@@
 ```
-
 The enrollment management industry puts great stock in market research.
-@RN4986[p.
-17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].^[@RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
+@RN4986 define market research as all "activities intended to increase understanding of a market and consumer behavior in that market" (p.
+17).
+At the most basic level, (in-house) market research is a function of intentional record-keeping.
+@RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
 In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited.
 In the chapter entitled "Travel," @RN3519 [p.
-69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."] However, most market research in the contemporary enrollment management industry is produced by third-party organizations.
-For example, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development [e.g., @RN4741; @RN4728, but most only available to clients.
+69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."
+
+The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals.
+In the late 1970s, College Board provided Carleton College with funding and data for a six-year project that analyzed college choice decisions by households in six metropolitan areas.[^2] This project was the basis for the landmark book *Applying Market Research in College Admissions* [@RN4986], published by College Board, and was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll.
+<!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.
+-->
+
+[^2]: In the 1970s, Carleton College became a leader in applying market research to student recruiting, under the leadership of Larry Litten, the then director of institutional research.
+Litten would later become Associate Director of the Consortium of Financing Higher Education (COFHE).
+
+Similarly, in 1979 The College Board provided Robert Zemsky, a professor of education at UPenn, with data and funding to develop the *Comprehensive Undergraduate Enrollment Planning Project* [@RN4982].
+In 1983, The College Board published *The Structure of College Choice* by Robert Zemsky and Penney Oedel, developed the "market segment model" and defined local "geo-markets" based on a comprehensive analysis of SAT score-sending by prospective students to coleges and universities.
+In 1984/5, the market segment model and geo-markets became the basis for the College Board Enrollment Planning Service (EPS), the market research division of College Board.
+
+This manuscript analyzes the market segment model and geo-markets as a case study in order to investigate market research -- and its commodification into recruiting products -- as a mechanism for social reproduction.
+At root, third-party market research helps universities identify who are the desirable students, where are they located, which institutions are they considering, and how best to reach them.
+Today, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development \[e.g., CITE RNL\], but most only available to clients.
 EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
-
-Despite being central to enrollment management for at least 40 years [@RN4986], the role of third-party market research has been ignored by scholarship on college access.
-Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935;@RN1814].
-However, market research informs where colleges look for students and which prospective students to target.
-These recruiting interventions intervene on student demand [@RN4324; @cottom2017lower], thereby structuring college access pathways in ways that could amplify or diminish historic inequality in college access.
-The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals [@RN4986].
-This article offers one story about the quiet revolution that transformed college access into a sophisticated, supply-led enterprise.
+Despite being central to enrollment management for at least 40 years [@RN4986], the role of third-party market research has been ignored by scholarship on college access, which views enrollment management behavior as the autonomous decisions of individual colleges, albeit structured by institutional logics [@RN4758; @RN4873; @RN3519].
 We argue that market research structures who colleges recruit, in ways that reinforce existing social stratification.
-The case of Market Segment Model and College Board geomarkets shows how market research -- and its commodification into recruiting products -- is a mechanism for social reproduction.
-[LAST TWO SENTENCES ARE REDUNDANT BUT BUT ARE TRYING TO DO SLIGHTLY DIFFERENT THINGS AND I LIKE THEM BOTH]
 
@@
 News & World Report (USNWR) Law School rankings demonstrate the reactivity, discipline, and authority effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
-Developed as a means of informing prospective students about quality, rankings had the intended effect of affecting.
-Rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
-Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p.
-415].
-Rankings *disciplined* the behavior of law schools.
-For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
-Rankings also weakened the *authority* of admissions directors and admissions officers to make independent decisions about which applicants to admit, ideal class size, and curricular offerings [@RN5003].
-<!--
-In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.    -->
+Developed as a means of informing prospective students about quality, rankings affected which law schools applicants applied to, the intended effect.
+However, rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
+(clients evaluated firms based on the prestige of the law schools that firm lawyers attended) and
+
+USNWR developed the rankings as a means to inform prospective students about law school quality. However, rankings affected
+
+US News developed law schools rankings as a means of informing prospective students about law school quality.
+In turn, rankings affected how prospective students conceive law school quality and which schools they applied to.
+Rankings also affected the cognition and behavior of other actors.
+Looking downstream, entities that received law school students as an input -- for example, law firms -- used rankings as a hiring criterion because their clients evaluated the organization based on the the prestige of the schools their firm's lawyers attended.
+
+Drawing from Foucault \[CITE DISCIPLINE AND PUNISH\], discipline is the idea that surveilance -- attributes of a law school being quantified and aggregated into a public-facing metric -- compels people and organizations to react in a particular way.
+For example, the reputational assessment of peer leaders as an important component of most US News rankings.
+Bastedo and Bowman showed that once (undergraduate) USNWR rankings became popular, the reputational assessment of peers primarily became a function of previous USNWR rankings.
+Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots schools felt pressured to take them seriously" \[QUOTE FROM ESPELAND AND STEVENS\].
+For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid \[CITE\].
+
+The theme of authority is that quantification destabilizes and weakens the authority of local decision-makers.
+For example, law school rankings weakened the authority of law schools to make independent decisions about which applicants to admit, ideal class size, and curricular offerings \[CITE THE BOOK\].
+In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.
 
 The analysis of UK League Tables by @RN5005 show how quantification can be a mechanism of social reproduction.
-As part of a broader *New Public Management* effort to infuse market competition within public providers [@RN161], the UK government began publishing school "league tables" in 1992.
-League tables ranked schools based on student performance on high-stakes national exams taken at age 16.
-Using Census data on occupational class measured at the local authority level (similar to a U.S.
-county), @RN5005 find that localities with higher performing schools experienced growth in the share of managerial/professional households following the adoption of league tables and a decline in the share of working-class households.
-Additionally, using longitudinal household survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables.
-With respect to mechanisms, league tables facilitate making evaluative comparisons between schools regardless of geographic proximity.
-They also discipline households and schools to conceive of school quality in terms of test scores, which are substantially a function of the class composition of schools.
-Compared to working-class parents, parents with professional/managerial occupations are more aware of school league tables and have financial resources necessary to move to neighborhoods near a "good" school.
-Therefore, league tables contributed to social reproduction because managerial/professional households respond to consumer-facing metrics by moving to more expensive neighborhoods, near higher performing schools.
-<!-- THIS IS RATCHET EFFECT STUFF.
-I CUT IT BECAUSE YOU HAVEN'T INTRODUCED THE IDEA OF RATCHET EFFECT
+In 1992, as part of a the broader *New Public Management* effort to instill market competition amongst public providers [@RN161], the UK government started publishing school "league tables" that ranked schools based on student performance on the high-stakes General Certificate of Secondary Education (GCSE) exams that students take at age 16.
+@RN5005 argue that quantifying the quality of schools entrenches class-based inequalities between schools particularly when the measure of quality -- standardized test scores -- is substantially a function of the class composition of schools.
+School league tables discipline households and schools to conceive of school quality in terms of test scores and they facilitate making evaluative comparisons between schools that are geographically close or distant.
+Compared to working-clas parents, parents with professional/managerial occupations are more aware of school league tables and have financial resources necessary to move to neighborhoods near a "good" school.
+
+Using school-level data and data on census data on occupational class measured at the local authority level (similar to a U.S.
+county), @RN5005 find that, following the adoption of league tables, local areas with higher performing schools experience growth in the share of managerial/professional occupational classes and a decline in the share of working class families.
+Additionally, using longitudinal survey data on 1% of the population of England and Wales and a diference-in-difference estimator, the authors find that managerial/professional households are more likely to move to local areas with higher ranked schools following the introduction of school league tables.
+Overall, @RN5005 show that quantifying which schools are high-performing contributes to social reproduction because managerial/professional households respond to consumer-facing metrics by moving to more expensive neighborhoods, near higher performing schools.
 Schools that performed well on initial league tables tended to be schools with a high share of professional/managerial families.
 Therefore, quantification takes a snapshot of existing social stratification and that snapshot becomes a power unto itself, amplifying the effect of previous strctural inequality on subsequent structural inequality.
--->
 
-```{=html}
-<!-- THIS IS THE CUT TRANSITION PARAGRAPH
 In the broader quantification literature, critical data studies [e.g., @RN4975; @RN4849; @RN4772] and sociological analyses of technology and inequality [e.g., @RN4778; @RN4786; @RN4877; @RN4810] introduce concepts salient to the analysis of market research.
 Our focus is on "selection devices," which utilize some combination of inputs and discretion to categorize places, people, and organizations for the purpose of allocating opportunities or interventions [@RN4778; @RN5012].
--->
-```
-In the interdisciplinary literature on quantification [@RN5008], the discussions of correlation and homophily by @RN4975 introduce concepts that are salient to the analysis of market research.
-Chapter 1 of @RN4975 is *Correlating Eugenics*.
-Correlation measures the extent to which two or more variables move together.
-The Pearson Correlation Coefficient was developed by the 19th Century statisticians and "biometric eugenicists" Francis Galton and his student Karl Pearson [@RN4975].
-@RN4975 [p.
-50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared in his 2008 editorial 'The End of Theory,' big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'" Predictive analytics are developed in two steps.
+
+A multi-disciplinary literature examines structural inequality produced by predictive analytics that capture correlational relationships [@RN4975].
+Predictive analytics utilize actuarial methods, which were pioneered by the insurance industry and proceed in two steps [@RN4835; @RN4778].
 First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
-Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and to assign levels of risk to each case.
-These models predict outcomes based on correlations without requiring knowledge about underlying causal relationships.
+Second, apply these results to future cases in order to make predictions and to assign levels of risk to each case.
+Importantly, these models predict outcomes based on correlations without requiring knowledge about causal relationships.
 @RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
 
-Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated to a variable of interest.
-For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of these others.
+The use of predictive analytics reproduces structural inequality.
+The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
+Although these correlations may b e a function of enduring structural inequality, the underlying causes are not considered by applications of predictive models.[^3].
+This algorithm recommends high purchase volume in zip codes that sent large number of students in previous years.
+@RN4794 \[p.
+224\] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities.
+@RN4851 refers to this phenomenon as the"ratchet effect," whereby disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies disproportionate targeting/exclusion.[HMM.
+PERNICIOUS FEEDBACK LOOPS WOULD SEEM TO FIT HERE]
+
+[^3]: For example, the College Board *Student Search Service* allows colleges to filter student list purchases based on zip code, amongst other filters.
+The consulting firm Ruffalo Noel Levitz offers an algorithm that recommends how many prospect profiles a college should purchase from each zip code, based on historic enrollment data from the college [@jmu_rnl]
+
+@RN4849 shows how predictive analytics create "weapons of math destruction" (WMD), which are defined as algorithms that have three characteristics: opacity, meaning that the algorithm is proprietary or unknown to the people it affects; scale, meaning that the algorithm can affect large numbers of people; and damage, meaning that the algorithm can reinforce past inequality.
+An exemplar WMD is the XXX model used to predict a prisoner's chances of re-arrest.
+How the algorithm uses inputs to predict the outcome is unclear (opacity).[^4] The algorithm is used by 24 states (scale).
+Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
+@RN5006 observe that such "pernicious feedback loops" [@RN4849] are similar to the "ratchet effect" [@RN4851] and are extreme versions of "reactivity" from @RN5003.
+
+[^4]: By contrast, U.S. law school rankings and UK leage tables are not WMDs because these algorithms are transparent [@RN5006].
+
+Due to data limitations, weapons of math destruction (WMD) -- and predictive analytics more broadly -- often utilize "proxy" variables, which are variables that are correlated to a variable of interest [@RN4849].
+FICO credit scores are based solely on individual repayment history and do not utilize proxies.
+By contrast, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of these others.
 @RN4849 [p.
@@
 
-Many studies show that using correlational models make predictions reproduces structural inequality.
-The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
-These correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
-@RN4794 [p.
-224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Therefore, disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies disproportionate targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].
-An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
-Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
-Note that predictive models such as the LSI-R model are not merely used for social science purposes.
-Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.
-
-Chapter 2 of @RN4975, *Homophily, or the Swarming of the Segregated Neighborhood*, argues that homophily often underlies the use of correlation to predict future behavior.
+@RN4975 argues that ideas about homophily often underlie the use of correlation to predict future behavior.
 Homophily means that actors who share common characteristics are likely to form connections with one another.
-The ubiquitous colloquialism is "birds of a feather flock together." Homophily is a core assumption for network science [@mcpherson2001birds], in which actors (e.g., high school students) are linked to one another through direct and indirect network ties (e.g., submit SAT scores to the same college).
-Because network science often draws from rational choice theory, it assumes that homophily is the result of voluntary action by individuals, thereby "eras\[ing\] historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
+The ubiquitous colloquialism is "birds of a feather flock together." Homophily has become a core assumption for network science [@mcpherson2001birds], in which actors (e.g., high school students) are linked to one another through direct and indirect network ties (e.g., submit SAT scores to the same college).
+Because network science often draws from rational choice theory, it assumes that homophily is the result of voluntary action by individuals, which "erases historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
 95] that underlie behavior consistent with homophily.
 @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.[^4]
+In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.
+Thus, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together" \[p.
+82\].
 
-[^4]: Thus, @RN4975 [p.
-82] writes, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together."
-
-Within market research, homophily is central to geodemography and the creation of market segments.
-Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
+Before the emergence of commercial social networks, homophily was the stated motivation behind geodemography, a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
+Geodemography emerged in the 1970s from parallel efforts in the U.S.
+and U.K to classify consumers based on publicly available census data.
 Market segments are subgroups within a larger market that have similar consumer demand.
-Early geodemographic classifications of consumers (e.g., PRIZM by Claritas Corporation) were derived from publicly available Census data, which disaggregated data to the zip code level [@RN4957].
-The Claritas Corporation argued -- and had a financial incentive to do so -- that people living near one another share similar consumer preferences.
-Therefore, geographic localities could be categorized into market segments that would be useful for direct mail marketing campaigns.
-Later, the development of individual credit scores (e.g., FICO score) enabled merchants to classify consumers into many, fine-grained groups [@RN4877].
-@RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].
-For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
-<!--
-ORIGINAL TEXT ABOUT GEODEMOGRAPHIC CLASSIFICATION SYSTEMS
-
+In the U.S., the Claritas Corporation classified zip codes into market segments that would be useful for direct mail marketing campaigns [@RN4957].
 Clusters were given evocative names like "Blue Blood Estates," "\[Archie\] Bunker's Neighbors", and "Share Croppers." Because U.S.
@@
 1] states, "the basic tenet of geodemography is that people with similar cultural backgrounds, means, and perspectives naturally gravitate toward one another or form relatively homogeneous communities; in other words, birds of a feather flock together." The Segment product allocated each high school and each census tract into a distinct "educational neighborhoods" thought to reflect useful market segments for colleges looking for students.
+<!--
+Negative consequences of homophily
+Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.-->
 
-PREVIOUS TEXT ABOUT CLASSIFICATION SITUATIONS
 Related to the creation of market segments, @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
@@
 443].
-
-Negative consequences of homophily
-Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
+Whereas FICO scores assign individuals a credit-worthiness value based solely on their individual record of financial transactions [@RN4877], other innovations classify consumers by integrating credit scores with demographic and geographic information [@RN4849].
+<!--
+The innovation of the FICO score assigned individuals a numeric credit-worthiness value, based solely on their individual record of financial transactions [@RN4877].
+In turn, the FICO score -- and, later, other credit scores -- enabled merchants to classify consumers into many, fine-grained groups or along a continuum.
 -->
 
+Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
+To our knowledge, prior scholarship on quantification within the sociology of education has not examined the role of third-party market research in classifying consumers.
+Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
+Furthermore, prior scholarship has not examined how market research becomes embedded in products that structure how consumers are matched to producers.
+This manuscript examines how the "market segment model" [@RN4982] categorized high school students into vertical market segments -- local, in-state, regional, and national -- and simultaneously created local "geo-markets" that were evaluated based on their composition of student market segments.
+These geo-markets are highly correlated with race and class.
+The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target, and geo-markets were later incorporated as a filter in the College Board *Student Search Service*.
+Prior scholarship in the sociology of education shows the effect of quantification on social reproduction [@RN5007; @RN5005].
+This study provides insight about the mechanism underlying this effect, by showing how geo-markets are highly correlated with race and class and by showing how geomarkets are utilized within student list purchases to reproduce historical class and racial inequality in college access.
+
+PARAGRAPH ABOUT CONTRIBUTION THAT IS MOVED HERE. WORK THESE IDEAS IN SOMEWHERE
+
+-   NOTE: THE TEXT FROM THE CASE STUDY WHERE YOU SAY HOW MANY COLLEGES WERE USING EPS CAN GO HERE; IT SPEAKS TO SALIENCE; WHEN TALKING ABOUT US NEWS AND UK LEAGUE TABLES ARE CONSUMER FACING AND HIGHLY SAIENT; CAN SAY THAT EPS IS BUSINESS FACING BUT HIGHLY SALIENT
+
 Reviewing recent scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Law school rankings [e.g., @RN5003] and school league tables [e.g., @RN5005] seem to exemplify this sort of salience.
-Widely known studies of quantification in education tend to study consumer-facing quantification.
-To our knowledge, prior research on quantification in education has not examined how third-party market research classifies consumers.
-This manuscript analyzes "market segment model" [@RN4982], which categorized high school students into vertical market segments and simultaneously created local "geo-markets" that were evaluated based on their composition of student market segments.
-The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target.
-@RN5021 reports that in 1995, 37% of 4-year public universities and 49% of 4-year private universities used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
-Geo-markets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities.
-<!--
-CUT COMMENTS ON SALIENCE
+Law school rankings and school league tables seem to exemplify this sort of salience.
 By contrast, many quantification projects -- for example, human rights indicators -- have "little oomph" [@RN5006, p.
 265].
-FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created."
---> Scholarship on predictive analytics [@RN4975; @RN4849] suggests the mechanisms by which market research reproduces structural inequality.
-Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
+FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created." The widely known studies of quantification in education \[e.g., \] [@RN5003; @RN5005] tend to study consumer-facing quantification.
+By contrast, market research targets businesses rather than consumers.
+To our knowledge, prior research on quantification in education does not examine market research.
 Unlike studies like @RN5005 we cannot show the effect of quantification.
-However, this study provides inisight about the mechanism underlying the effect.
-By analyzing and simulating *Student Search Service* purchases that filter on geo-markets we can show *how* the geo-markets reproduce historical class-based and race-based inequality in college access.
+However, by analyzing and simulating student list purchases we can show *how* the "geo-markets" reproduce inequality \[NOTE: THIS PARAGRAPH CONTAINS TWO IDEAS; MIGHT MOVE OR CUT ONE OF THEM LATER\].
 
-# The Market Segment Model and College Board Geomarkets
+# The Market Segment Model
 
- <!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.-->
-As a member organization, several College Board products developed from joint collaborations with administrators or professors from member institutions.
-During the 1970s, under the leadership of Larry Litten, then director of institutional research, Carleton College pionered applying market research to student recruiting.
-In the late 1970s, College Board provided Carleton College with funding and data for a six-year project that analyzed college choice decisions by households in six metropolitan areas.
-This project was the basis for the landmark book *Applying Market Research in College Admissions* [@RN4986], published by College Board, and was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll.
-Litten would later become Associate Director of the Consortium of Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
+The market segment model -- and its associated geo-markets -- is a case that incorporates several themes of the quantification literature and one that is simultaneously historical and contemporary.
+In 1978, Robert Zemsky, a University of Pennsylvania professor, began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE) -- a group of 30 selective private institutions -- to understand questions like, "'Who thinks about Penn?' \[and\] also, 'What other institutions do they think about when they think about us?'" [@RN4982, p.
+x].
+As the project project became more ambitious, the researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincientally the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in tehse analyses was the state' [@RN4982, p.
+x].
+In 1979, College Board began providing data and funding for what became the Comprehensive Undergraduate Enrollment Planning Project (CUEPP).
+In 1983, the College Board published *The Structure of College Choice* [@RN4982], which developed the market segment model, introduced geo-markets, and became the basis for the College Board Enrollment Planning Service (EPS).
 
-In 1978, Robert Zemsky, a University of Pennsylvania professor, was asked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
-x].
-To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE).
-As the project project became more ambitious, the researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in tehse analyses was the state' [@RN4982, p.
-x].
-In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.^[say that Pennsylvania Department of Education was also a partner; see page x of Zemsky] @RN4982[p.
-4] write that,
+Considering the @RN5007 \[p.
+402\] definition of quantification as "the production and communication of numbers," the market segment model is an effort to quantify and communicate the knowledge that admissions officers have about student demand.
+@RN4982 [pp.
+9-10] write that,
 
-> For our analysis, we sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
-The Market Segment model that we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
-The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways.
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting \ldots we believe that the institutions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process."
 
-**Creating the Market Segment Model.** Considering the @RN5007 \[p.
-402\] definition of quantification as "the production and communication of numbers," the Market Segment Model is an effort to quantify and communicate the knowledge that admissions officers have about student demand.
-@RN4982[p.
-4] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." They created  "three types of boundaries" [@RN4982, p.
+@RN4982 [p. 10] stress the importance of "knowing the territory":
+
+> A good recruiter knows where to look for prospective applicants, as seen in the students' willingness or eluctance to travel.
+It is necessary to identify not only the most promising communities or poools, but also the specific neighborhoods within those communities -- hence the recruiter's emphasis on feeder high schools.
+From the beginning, we have sought to mold our research to these concepts, to capture and quantify the phenomenon behind the folklore.
+Thus, it is the admissions officers' notion of admissions pools that dictates the geographic units of our analysis." \[p.
+11\]
+
+**Creating the market segment model.** The first step in developing the market segment model was creating "three types of boundaries" [@RN4982, p.
 11], region, state, and community.
@@
 Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" \[in New England, Middle States, and the South\].
-
-These enrollment markets, later called "geomarkets" by the College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
-We know little about how geomarkets were created.
+These enrollment markets, later called "geo-markets" by he College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
 @RN4982 write that,
@@
 
-**Correlations**.
+**Understanding and using the market segment model**.
 Chapter 3 of @RN4982 -- "A Sense of Place: Students, Families, and Communities" -- identifies the student characteristics associated with being in the local, in-state, regional, or national market segment and examines the extent to which these relationships hold across states and local markets.
-The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
-[@RN4982, p.
-33] state that these four variables "reflect the basic social patterns of the nation.
-It would have been surprising if these were not the four social variables that best explained the patterns of college choice."
+The data analyzed by @RN4982 can be seen as training data.
+In actual recruiting applications, colleges would have historic data on the number of students in each segment in a local market, but they would not know a current high school student is a local, in-state, regional, or national prospect.
+Therefore, knowing which observable student characteristics are associated with score sending outcomes can be useful for knowing which prospects an institution should target.
 
-@RN4982 summarize their correlational analyses using authoritative language.
-@RN4982[pp.
-34-35] asserts, "it is hard to overemphasize the importance of this statistical pattern.
-Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice." With respect to family income, @RN4982[p.
-33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982[p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time." Further analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
-Considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model implicitly and explicitly states that student demand for higher education is mostly a function of social origin.
-@RN4982 also find that the SAT score, parental education, and family income predict student score-sending behavior at the local geo-market level.
+Chapter 3 of @RN4982 is fundamentally about correlations.
+The authors identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that individually or in combination predict score-sending behavior.
+These four variables "reflect the basic social patterns of the nation"<!--.
+It would have been surprising if these were not the four social variables that best explained the patterns of college choice"--> [@RN4982, p.
+33].
+The text uses authoritative language to summarize the correlational analyses.[^6] @RN4982 state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" (p.
+33) and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations.
+Commenting on family income, @RN4982 write that"we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time" (p.
+33).
+Further analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
+In their analysis of quantifying school quality in England, @RN5005 [p.
+517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin \[CITE\], the Market Segment Model essentially argues that student demand for higher education is mostly a function of social origin.
 
-However, geomarkets differ in the relative abundance of students with particular socioeconomic characteristics and this has practical implications for recruiting.
+[^6]: For example, @RN4982, pp.
+34-35 state, "it is hard to overemphasize the importance of this statistical pattern.
+Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice."
+
+@RN4982 find that the SAT score, parental education, and family income also predict student score-sending behavior at the local geo-market level but geo-markets differ in the relative abundance of students with particular socioeconomic characteristics.
+These findings have practical implications for recruiting.
 @RN4982 recommend that colleges target geo-markets with desirable compositions of socioeconomic characteristics and student market segment.
-@RN4982[p.
+@RN4982 [p.
 44] describes a hypothetical college that wants to target regional and national students in New England and is considering whether to expend limited recruiting efforts in the Boston geo-market -- which was relatively low-income in 1980 -- or in more affluent geo-markets nearby:
@@
 
-**EPS and College Board Geomarkets**.
-In 1984, College Board created the Enrollment Planning Service (EPS) based on the Market Segment Model [@RN4987; @RN4980].
+In 1984, College Board created the Enrollment Planning Service (EPS) based on the @RN4982 market segment model [@RN4987; @RN4980].
 Each institution purchasing EPS information services would obtain access to the Market Segment Report for each local market.
@@
 ```
-Whereas @RN4982[Appendices B1-B3] identified 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S.
+Whereas @RN4982 [Appendices X-Y] created 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S.
 states, with 304 geo-markets in total.
@@
 Interestingly, the geo-markets for New England, Middle States, and the South remain nearly identical to those developed by @RN4982.
-Concrete information about how EPS created the remaining geomarkets is elusive.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^7]
+Concrete information about how EPS created the remaining geo-markets is elusive.
+Documentation and promotional material suggests that geo-market borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^7]
 
@@
 
-Although how we lack clarity about how geomarkets were created, we expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
+Although the methods utilized to create geomarket borders remain unclear, we expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
 Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
@@
 In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status.
-The language of @RN4982[pp.
-11-12, quoted above] suggests that geomarket borders within large metropolitan areas may have been drawn in ways that separate affluent from less affluent communities.
-We present the following hypothesis:
+The language of @RN4982 [pp.
+11-12] suggests that in large metropolitan areas, the borders of geomarkets may have been drawn in ways that separate affluent from less affluent communities: "each major metroplitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs." We present the following hypothesis:
 
@@
 
-We expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons.
-Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race once.
-However, U.S.
-cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779; @rothstein2017color].
-Structures built upon racialized structures are racialized structures.
-@RN4975 argues that unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity.
-Second, geomarket borders may have been drawn along class divides.
-There is a strong correlation between race and wealth [@RN4884].
-Third, geo-market borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
+We also expect substantial racial inequality between geomarkets in large metropolitan areas.
+Interestingly, whereas the Market Segment Model is explicitly based on socioeconomic stratification, @RN4982 do not mention race once.
+Structural racism is "a form of systematic racial bias embedded in the 'normal' functions of laws and social relations" [@RN4760, p.
+1143], whereby processes viewed as neutral or common sense systematically advantage dominant groups and disadvantage marginalized groups.
+Because of the extent of residential racial segregation in the U.S.
+[@RN4779; @rothstein2017color], algorithms that utilize geographic borders as an input have been shown to reproduce racial inequality [e.g., @RN4775; @RN4849; @RN4779].
+Unless they intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
+
+Geomarket borders may be correlated with race for two additional reasons.
+First, geomarket borders may have been drawn along class divides and there is a strong correlation between race and wealth [@RN4884].
+Second, geo-market borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
 Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
@@
 
-We argue that the Market Segment Model, geomarkets, and their commodification within EPS exemplify concerns about the application of correlational analysis.
+**Reproducing Inequality**.
+Finding that geomarkets are correlated with class and race is neither surprising nor noteworthy.
+However, we argue that the Market Segment Model and geomarkets are mechanisms for reproducing class-based and race-based inequality that exemplify many themes of the broader quantification literature.
+<!--
+In both their usage within the EPS market research product and the Student Search Service student list product, 
+-->
+
+the Market Segment Model and geo-markets exemplify concerns about the application of correlational analysis.
 The data on 1980 SAT score-sending behavior analyzed by @RN4982 can be conceived as training data.
@@
 These analyses found that student demand for different kinds of colleges was largely a function of social origin, and that student demand within a geo-market was largely a function of the class composition of the geo-market.
-However, @RN4982 did not interrogate the historical and contemporary structural inequalities that produced observed patterns of social stratification.
+but did not interrogate the historical and contemporary structural inequalities that produced observed patterns of social stratification.
 Instead, the Market Segment Model encouraged colleges to use this snapshot of student demand as the basis for future decisions about which communities and prospective students should be targeted by college recruiting campaigns.
@@
 
-**Homophily and Utilization**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] shows that homophily is a core tenet of the Market Segment Model.
-The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^8] @RN4982[p.
-42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^9] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
+The utilization of the Market Segment Model as the basis for Enrollment Planning Service (EPS) raises the themes of reactivity, discipline, and authority introduced by @RN5007.
+Reactivity means that quantitative measures cause people and organizations to change behavior.
+Both @RN5003 (law school rankings) and @RN5005 (UK leage tables) study highly salient consumer-facing products.
+Although EPS targeted colleges rather than students, we suggest that EPS -- and geomarkets -- were highly salient within the enrollment management community.
+@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
+Discipline is the idea that quantitative measures compel people and organizations to react in a particular way.
+To the best of our knowledge, EPS originally provided colleges permutations of the Market Segment Model -- how many students in each market segment -- and the Institutional Profile -- which students in a market segment were interested in my college or a competitor.
+Both of these outputs were provided at the geomarket level.
+Promotional material from College Board -[@RN5019; @RN5019_2010] suggests that EPS encouraged colleges to first choose geo-markets of interest and then to plan recruiting efforts within these selected geo-markets.
+
+Usage of EPS also raises the theme of authority, whereby quantification weakens the authority of local decision-makers [@RN5007].
+The Market Segment model sought to develop a concrete, data-driven framework that replicated the aggregate wisdom of admissions officers [@RN4982].
+This model makes the local expertise of admissions officers less valuable.
+EPS transformed the model into a product stored on a CD-ROM.
+In turn, the EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
+Individual admissions officers are relegated to implementation of the plan rather than being valued for their local expertise about student demand.
+
+The incorporation of geo-markets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction.
+First, geo-markets are opaque.
+The borders of a geomarket may be unclear to someone purchasing student lists and very few people know how these borders were created \[COULD TALK ABOUT PRODUCT UTILIZATION HERE; NOT CLEAR HOW GEOMARKETS INTERACT WITH OTHER STUDENT LIST FILTERS\].
+Second, geo-markets scale.
+Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on geo-markets in concert with other filters (e.g., SAT score, GPA).
+Third, geo-markets can cause damage.
+When a college decides to purchase names from one geo-market but exclude prospects from an adjacent geo-market, this affects which college students have an opportunity to attend.
+@RN4982 suggests that selective colleges focus on geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
+
+Chapter 4 of @RN4982 -- The Company We Keep: Colleges and Their Competition -- shows that the Market Segment Model is based on homophily.
+The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^8] @RN4982 [p.
+42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^9] Subsequent analyses investigate the tuition price and socioeconomic competition of institutions in competition with one another.
 Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-@RN4982[p.
-72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+Rather than considering the structural inequality that gives rise to these patterns, @RN4982 describe a naturally occuring process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities.[^10] @RN4982 [p.
+72] state that
 
@@
 
-<!--[^10]: These patterns are similar to different offers for credit products in The Ordinal Society [@RN5015], except @RN5015 show that the matching of consumers to products is engineered by firms rather than naturally occurring.-->
+[^10]: These patterns are similar to different offers for credit products in The Ordinal Society [@RN5015], except @RN5015 show that the matching of consumers to products is engineered by firms rather than naturally occurring.
 
@@
 
-Although this perspective is broadly consistent with mid-20th Century scholarship on social mobility [@RN915; @RN1273], the Market Segment Model was the basis for the EPS market research product.
-Later, geomarkets were incorporated into the *Student Search Service* student list product.
-Commodification of the Market Segment Model "engineers homophily" [@RN4975] by taking a snapshot of existing stratification and then recommending that colleges utilize this snapshot as the basis for decisions about where to concentrate recruiting efforts.
+Although this perspective is not entirely inconsistent with mid-20th Century scholarship on social mobility [@RN915; @RN1273], the Market Segment Model was the basis for the EPS market research product and, later, geomarkets were incorporated into the *Student Search Service* student list product.
+Drawing from @RN4975, commodification of the Market Segment Model "engineers homophily" [@RN4982] by taking a snapshot of existing stratification -- without considering the class-based or race-based structures that produced these patterns -- and then recommending that colleges utilize this snapshot as the basis for decisions about where to concentrate recruiting efforts.
 @RN4982 [p.
-58] states that "correlations, do not simply predict actions; they also form them." In itself, the Market Segment Model is simply a stylized depiction [HIRSCHMAN 2016 STYLIZED FACTS] of the structure of college choice, one that does not consider class-based and race-based structures that produced these patterns.
-Once the College Board inscribed the Market Segment Model into the EPS and Student Search Service Products, these products are likely to amplify the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.
+58] states that "correlations, do not simply predict actions; they also form them." The Market Segment Model not only describes the structure of college choice, it influences the future structure of college choice by becoming the basis for practices and products.
+Once the Market Segment Model is inscribed in EPS and Student Search Service, these products are likely to amplify the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.
 
-Utilization of the Market Segment Model and geomarkets within the Enrollment Planning Service (EPS) raises concerns related to the quantification themes of reactivity, discipline, and authority introduced [@RN5007].
-@RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Although EPS is a business-facing rather than a consumer-facing product, it has been a highly salient one.
-@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
-We suggest that EPS disciplines colleges to approach recruiting in a manner consistent with the Market Segment Model.
-Colleges are encouraged to, first, choose geomarkets with a high number of desirable prospects and, second, plan recruiting efforts within selected geomarkets.
-Under the principle of homophily, selective colleges should target geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-Consistent with this perspective, EPS software documentation from @RN5017 describes how geomarkets -- here, called "EPS market codes" -- can be utlized:
-
-> EPS\textsuperscript{\texttrademark} (Enrollment Planning Service) is a geographic and demographic data service offered annually by the College Board to Colleges and Universities.
-EPS provides information to subscribing institutions about competitors, feeder schools, and demographic strengths and weaknesses.
-EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful.
-
-EPS may also weaken the authority of local decision-makers.
-The Market Segment model sought to develop a concrete, data-driven framework that replicated the aggregate knowledge of local admissions officers [@RN4982].
-Once this knowledge is quantified and commodified into a product stored on a CD-ROM, the local expertise of admissions officers becomes less valuable.
-The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
-In turn, admissions officers are relegated to a foot-soldier implementation role rather than being sought out for their local expertise about student demand.
-
-Finally, we suggest that utilization of geomarket filters within the Student Search Service product is likely to amplify class-based and race-based disparities in college opportunity.
+UTILIZATION.
+Finally, within the Student Search Service product, we suggest that geomarket filters are likely to be used in ways that amplify class-based and race-based disparities in college opportunity.
 In contrast to standardized selection devices (e.g., the FICO score algorithm), student list products are discretionary selection devices in that colleges have discretion of which filters they elect to select prospects.
 Compared to standardized selection devices, scholars tend to find that administrative discretion over selection devices causes structural inequality to increase [@RN4883; @RN4881; @RN4795].
-Even in the absence of explicit or implicit bias, discretionary selection devices can increase racial inequality when decision-makers have incomplete knowledge about how the inputs they choose are correlated with race.
+Even in the absence of explicit or implicit bias, discretionary selection devices can increase structural inequality when decision-makers have incomplete knowledge about how the inputs they choose are correlated with race \[CAN MOVE O'NEIL TEXT ON OPACITY HERE\].
 For example, research shows that Americans dramatically underestimate the magnitude of racial income inequality [@RN4884].
@@
 
-Geomarkets are opaque in that geomarket borders may be unclear to someone purchasing student lists and very few people know how these borders were created.
-@RN4849 argues that opacity in algorithmic products is intentional and is a core criterion of "weapons of math destruction." Enrollment management professionals utilizing EPS and *Student Search Service* may interpret geomarkets as a race-neutral recruiting tool.
+Colorblind racism describes ideology and associated behaviors that ignore the significance of race and racial discrimination [@RN5020].
+The Market Segment Model can be seen as an example of colorblind racism.
+@RN4982 develop an analysis of the structure of college choice that ignores race but we contend that geo-markets are correlated with race.
+Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975].
+Enrollment management professionals utilizing EPS and *Student Search Service* may interpret geomarkets as a race-neutral recruiting tool.
 When purchasing student lists, @RN4982 [p.
@@
 Furthermore, purchases that specify multiple filters (e.g., test scores, grades, intended major) including geomarkets can yield unintended racial inequality because administrators have incomplete knowledge about how the intersection of these filters interact with local patterns of segregation.
-
-<!-- CUT TEXT DEFINING COLORBLIND RACISM AND RACE-NEUTRAL RACISM
-Colorblind racism describes ideology and associated behaviors that ignore the significance of race and racial discrimination [@RN5020].
-The Market Segment Model can be seen as an example of colorblind racism.
-@RN4982 develop an analysis of the structure of college choice that ignores race but we contend that geo-markets are correlated with race.
-Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975].
+<!-- We suggests that geomarkets are a colorblind racist market device.
+Futhermore, geomarkets may interact with other student list filters (e.g., test score, GPA) in ways that exacerbate exclusion of under-represented student populations.
 -->
-
- <!-- CUT PARAGRAPH: WHY GEOMARKETS ARE WEAPONS OF MATH DESTRUCTION
-The incorporation of geo-markets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction.
-First, geo-markets are opaque.
-The borders of a geomarket may be unclear to someone purchasing student lists and very few people know how these borders were created \[COULD TALK ABOUT PRODUCT UTILIZATION HERE; NOT CLEAR HOW GEOMARKETS INTERACT WITH OTHER STUDENT LIST FILTERS\].
-Second, geo-markets scale.
-Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on geo-markets in concert with other filters (e.g., SAT score, GPA).
-Third, geo-markets can cause damage.
-When a college decides to purchase names from one geo-market but exclude prospects from an adjacent geo-market, this affects which college students have an opportunity to attend.
-@RN4982 suggests that selective colleges focus on geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
--->
-
-# Methods
 
@@
 In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.
-
-# Discussion
-
-In their analysis of quantifying school quality in England, @RN5005 [p.
-517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model argues that student demand for higher education is mostly a function of social origin.
 

### 2024-07-08 text/old_spring2024/cb_geomarket_background.qmd
initial commit, copying files from messed up repo

---
title: "Structuring College Choice: The Curious, Mundane Case of College Board Geomarkets"
author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
bibliography: ./../../assets/bib/cb_geomarket_bib.bib
csl: ./../../assets/bib/apa.csl
format:
  pdf:
    documentclass: article
    #papersize: letter
    number-sections: true
    number-depth: 2
    #toc: FALSE
    #pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #keep_tex: true
editor: visual
geometry:
  margin=1in
fontsize: 12pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
library(knitr)
#library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)
```

# Introduction

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

the bal bla @RN2010

# Literature Review

Scholarship on college access can be categorized as focusing on demand-side versus supply-side factors. Broadly, economics defines the demand-side as factors that affect consumer demand for a product (consumer income, population trends, prices, consumer tastes) whereas the supply-side refers to factors that affect production of a product (e.g., cost of inputs and other production costs, technological advancement, government regulation). The majority of research on college access focuses on demand-side factors. In economics, demand-side studies include how students respond to tuition and financial aid, student academic preparation (e.g., racial and socioeconomic "test score gaps"), and information interventions. In sociology, demand-side topics include analyses of college-going culture \[cite Bourdieu, McDonough\]

# Background

According to its [bylaws](https://about.collegeboard.org/governance/bylaws),

> The College Board is a not-for-profit membership corporation operating for educational purposes \ldots Members of the College Board are secondary and postsecondary institutions, districts and systems, and nonprofit educational associations, organizations, and agencies serving secondary and/or postsecondary education

Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board. For example, in the 1970s, Carleton College became a leader in applying market research to student recruiting under the direction of Larry Litten, then director of the Office of Institutional Research. In the late 1970s Carleton College partnered with College Board to develop the "Carleton College-College Board Six-Market Study" about the college choice decisions of households, with College Board providing funding and data for the project @RN4986 \[and Litten became Associate Director of the Consortium of Financing Higher Education (COFHE). This research project was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll. Similarly, in 1979 The College Board partnered with Robert Zemsky, a professor at UPenn, to develop the *Comprehensive Undergraduate Enrollment Planning Project* [@RN4982]. This project was the basis for the College Board Enrollment Planning Service (EPS), the market research and enrollment consulting division \[diff word?\] of College Board.

## The Market Segment Model

In 1983, The College Board published *The Structure of College Choice* by Robert Zemsky and Penney Oedel. This book, based on the *Comprehensive Undergraduate Enrollment Planning Project* develops the Market Segment model, which introduced "geo-markets" and became the basis for EPS.

@RN4982 describes the Market Segment model as an effort to develop a systematic, quantitative description of which institutions are likely to be considered by which institutions, based on the shared stories and wisdom of college admissions officers. @RN4982 [p. 9] write that,

> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting \ldots. We have begun with this celebtration of storytelling for two reasons. First, we believe that the institutions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process. Second, much of that knowledge has been captured in the admissions officers' tales, which exhibit all the hallmarks of a classic folklore: language, symbol, and system. Our research thus is based on listening carefully to what admissions officers have to say."

> From the beginning, we have sought to mold our research to these concepts, to capture and quantify the phenomenon behind the folklore. Thus, it is the admissions officers' notion of admissions pools that dictates the geographic units of our analysis." \[p. 11\]

As such, the first step in developing the Market Segment model was to identify p. 11 "three types of boundaries -- region, state, and community" -- as follows: first, regions were defined as New Egland, Middle States, and the South; Second, states are states; and, third, each state was divided into "as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" \[in New England, Middle States, and the South\]. These community-based enrollment markets -- later called "geomarkets" by The College Board Enrollment Planning Service -- are the analytic focus of this manuscript. Enrollment markets are intended to be consistent with the conception of a catchment market from the perspective of admissions counselors. @RN4982 write that,

> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication. Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs. In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."

Having defined these geographic boundaries, each SAT score a student submits to an institution can be defined as "local" (institution is in the same local market that the student lives in), "in-state" (institution is located in different local market but the same state the that the student lives in), "regional" (institution is located in a different state but the same regional market that the student lives in), or "national" (institution is located in a different regional market than the student lives in).

Next, high school students are categorized into found different *segments* -- local, in-state, regional, and national -- based on their score-sending behavior. A "local" student submits more SAT scores to local institutions than they do to in-state, regional, or national institutions. An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc. <!--
A "regional" student submits more SAT scores to regional institutions than they do to local, in-state, or national institutions. A "national" student submits more SAT scores to national institutions than they do to local, in-state, or regional institutions.
-->

*The Market Segment Profile*. The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile. @RN4982 state that "the Market Segment Profile is each market's signature, a detailing of student choice and institutional standing within each segment of that market. In a single-page format, this information is presented for each of the 143 markets currently included in the Comprehensive Undergraduate Enrollment Planning Project." @tbl-market-segment-characteristics reproduces a partial, simplified version of the bottom panel of @RN4982 [Figure 2.1] which represents the Market Segment Profile for "Connecticut Market 3: Fairfield County."

@tbl-market-segment-characteristics has separate columns for each market segment (local, in-state, regional, national) and rows show the number of test-takers and characteristics of test-takers. For example, there were 550 "local" students in Fairfield county and these students submitted SAT scores to 2.8 postsecondary institutions on average. By contrast, there were 1,664 "regional" students who submitted test scores to 4.8 institutions on average. For local students, 11.2% had family income greater than \$35,000 and 9% had both parents with a BA. For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA. Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.[^1]

[^1]: @tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market. For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution. Of the 1,199 in-state students, 757 (63.1) sent scores to institution #1, a public doctoral granting institution, and 515 (43.0%) sent scores to institution #2, a public master's granting institution.

INSERT @tbl-market-segment-characteristics ABOUT HERE

*The Institutional Profile*. For each postsecondary institution, the Market Segment Model produces several Institutional Profiles, which describe the students who sent scores to the institution and which majors they are interested in. Institutional profiles are created separately for students from a particular local market, for all students in a state, or all students in a region. @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], which is the institutional profile of an anonymous institution for students from Fairfield County, CT. @tbl-sample-institutional-profile shows that 58 students submitted SAT scores to the institution. These 58 students represented 4.8% of the total 1,199 in-state students from Fairfield County. 69 regional students sent scores to the institution, representing 4.1% of all 1,664 regional test-takers. Of these 69 regional students, 35 expressed interest in majoring in the liberal arts. These 35 students represent 7.0% of all regional students from Fairfield County who expressed interest in the liberal arts.

INSERT @tbl-sample-institutional-profile ABOUT HERE

@RN4982 argue that the Institutional Profile and the Market Segment Profile enable admissions officers to know where to look for students and which institutions are competing for those students (p. 25):

> The Institutional Profile and the Market Segment Profile quantify the admission officers' intuitive grasp of market structure. Structure here carries a dual meaning, connoting both the structure of student choice and the structure of institutional competition\ldots This two-sided interpretation furnishes the essential framework for planning by individual colleges and universities\ldots To draw effectively on its own natural constituency, a college not only must contact the "right" kind of students --- that is, students who are predisposed toward that type of institution -- but also must persuade them of its special character. This means knowing the competition as well as the clientele.

**Enrollment Planning Service**. The College Board Enrollment Planning Service (EPS) was created in 1984 to provide institution-specific market research that would inform the recruiting strategy of member institutions [@RN4987].

Just as @RN4986 and the "six market area study" -- underwritten by College Board -- was basis for the College Board Admitted Student Questtionnaire product, @RN4982 and the the *Comprehensive Undergraduate Enrollment PlanningProject* -- underwritten by College Board -- was the basis for EPS [@RN4980]. Although information is spotty, the original EPS product seemed to be based fundamentally on the Market Segment Model developed by @RN4982. Each institution purchasing EPS information services would obtain access to the Market Segment Report for each local market. They also received the Institutional report for each state and for each local market. Collectively, these reports enabled institutions to determine where attractive students were located, which students were considering -- or might consider -- the institution, and which institutions were likely to be in competition for these students.

The appendix of @RN4982 maps the local markets for New England, Middle States, and the South, for a total of 143 local markets. At some point, College Board EPS created geo-markets for the remaining US states. @RN4988 shows the College Board geo-markets circa 2023. Incredibly, the geomarkets for New England, Middle States, and the South remain \[?ALMOST?\] exactly the same.

WHAT ELSE TO SAY ABOUT ENROLLMENT PLANNING SERVICE???

PX: think about how geomarkets are used More generally as part of recruitment planning by EPS NEED TO GET SOME CONCRETE INFORMATION ABOUT HOW GEOMARKETS ARE USED In student list purchases, geomarkets are used as a filter alongside other filters

/https://secure-media.collegeboard.org/homeOrg/content/pdf/welcome_to_the_college_board_fullversion.pdf This document is from 2012 "Enrollment Planning Service (EPS®) is an online database using a wide range of criteria, including geography, high schools, score-sending patterns and international regions. EPS allows staff to analyze data and find prospective students both nationally and internationally who are likely to succeed at their institution. This allows for better and more targeted recruiting and communications based on a wide range of criteria, including geography, demographics, academic preparation and collegiate aspirations. EPS also provides comprehensive reporting and fully integrates with Student Search Service.

This is interesting https://www.pdffiller.com/28049652-fillable-college-board-enrollment-planning-service-form?mode=view

EPS "is the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found. You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, Zip codes, high schools and international regions. EPS provides you with comprehensive reports on your markets, your position in those markets, and your copetition."

Market research by college board https://www.tandfonline.com/doi/pdf/10.1080/00091383.1984.10570121 https://eric.ed.gov/?id=ED238341 Druesne, B., Harvey, J. & Zavada, M. (Summer 1980). College mailings: What works. The College Board Review, 116,12-17.

ENROLLMENT PLANNING SERVICE From https://www.jstor.org/stable/pdf/40195699.pdf \[1986 RIHE\] Other Erollment Planning SerVIce EPS and Descriptor PLUS are data serVIces that make use of College Board infornMation to provide Institutions with intelligence on trends and conditions that directly affect their enrollment management strategy

**Student Search Service**.

SAY BASIC SHIT ABOUT STUDENT SEARCH SERVICE. COME BACK TO THIS LATER.

HOW GEOMARKETS ARE USED IN STUDENT LIST PRODUCTS

# Conceptual Framework

We develop a conceptual framework by drawing from *Discriminating Data: Correlation, Neighborhoods, and the New Politics of Recognition* by Wendy Hui Krong Chun (2021). *Discriminating Data* investigates how the correlation and assumptions about homophily have been utilized to create a future that amplifies the power of past discrimination. In particular, we draw from chapter one, "Correlating Eugenics," and chapter two "Homophily, or the Swarming of the Segregated Neighborhood."

## Correlations

Correlation measures the extent to which two or more variables move together. According to the Pearson correlation coefficient, if an observation $x_i$ greater than the mean $\bar{x}$ is associated with an observation $y_i$ greater than the mean $\bar{y}$ then variables $x$ and $y$ are positively correlated.

@RN4975 traces the origins of correlation to 19th century statisticions and "biometric eugenicists" Francis Galton and and his student Karl Pearson. These scholars developed the concepts of correlation and linear regression while conducting research on plants, animals, and peoples about the relationships between breeding and genetic inheritence. Biometric eugenicists viewed selective breeding as the most important tool for human progress. Eugenics, Pearson stated is "the science of improving stock, not only by judicious mating, but by all the influences which gave the more suitable strains a better chance" \[quoted in Chun page 59\]. Eugenicists believed that efforts to improve conditions -- for plants, animals, people -- would only improve performance for a single generation whereas selective breeding would improve performance for multiple generations. Therefore, Pearson stated, "Give educational facilities to all, limit the hours of labour to eight-a-day\ldots give a minimum wage with free medical advice, and yet you will find that the unemployables, the degenerates and the physical and mental weaklings increase rather than decrease" \[quoted from Chun page 65\]. The key to human progress, is to keep unfit bloodlines separate from pure bloodlines and to stop unpure from breeding.

The predictive analytics embedded in most algorithmic products are based on correlation. The creation of predictions proceeds in two steps: Using "training" data, apply statistical models to previous cases to determine the predictors of an outcome; second, apply the results of these analyses to predict the outcome for future cases \[give brief explanation of training data and then predict for subsequent datasets, can draw from your EEPA chapter\] Correlations can be used to predict outcomes without requiring knowledge about causal relationships. @RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes based on Facebook Likes. Their approach was subsequently utilized by political consulting firm Cambridge Analytica to identify and target subgroups that could be swayed to vote for their clients, such as Ted Cruz, who were seeking elected office.

In @RN4982, chapter 3 "A Sense of Place: Students, Families, and Communities" is fundamentally about correlations. Having categorized each high school student as local, in-state, regional, or national based on their score-sending behavior, chapter 3 analyzes which social and economic characteristics are associated with these behaviors and the extent to which these relationships hold across states and local markets. The data analyzed by @RN4982 can be seen as training data. In actual recruiting applications, colleges would have historic data on the number of students in each segment in a local market, but they would not know a current high school student is a local, in-state, regional, or national prospect. Therefore, knowing which observable student characteristics are associated with score sending outcomes can be useful for knowing which prospects an institution should target. \[NOTE: THEY WOULD ALSO KNOW THE NUMBER OF STUDENTS IN EACH SEGMENT FRONM PAST COHORTS; CUZ THIS IS IN THE MARKET SEGMENT REPORT\]

@RN4982 identify four variables that can be used individually or in combination to predict score sending behavior. "These four attributes -- educational aspirations, parental education, scholastic aptitude, and family income -- reflect the basic social patterns of the nation. It would have been surprising if these were not the four social variables that best explained the patterns of college choice" (p. 33). Commenting on parental education, @RN4982 [p. 33] state that, "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations." Commenting on family income, they write that we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time" (p. 33). Further analyses reveal that the SAT score is the most important predictor, followed by parental education, family income, and educational aspirations.

Many scholars of critical data studies have shown the negative consequences of using predictive analytics to make decisions [CITE SAFIYA, AMONGST OTHERS]. Proxy variables are variables that are highly correlated to a variable of interest (e.g., EXAMPLE). @RN4849 states that "weapons of math destruction" use proxy variables to make predictions in the absence of desired data. For example, the COMPAS algorithm for predicting re-arrest utilizes home zip-code as an input, which results in higher predicted probabilities of re-arrest for Black people because of racial segregation in housing. Weapons of math destruction often exclude race as an input and use inputs that are correlated with race, while simultaneously ignoring historical structural racism that produced the correlation in the first place. For example, Norris [CITE] shows that Moody's algorithm for assigning city credit scores uses median income as an input, but not racial composition. Given historic wage discrimination, the resulting system penalizes cities with large Black populations while appearing to be race-neutral \[or colorblind?\]. Another consequence of using historic correlations to make future decisions is that this amplifies the effects of past discrimination that produced these correlations \[some redundancy in this paragraph\]. Thus, "Correlations, do not simply predict actions; they also form them" [@RN4982, p. 58].

We argue that the designation and usage of local markets in the Market Segment Model exemplify the negative consequences of correlations. @RN4982 [p. 44] describes a hypothetical college that wants to target regional and national students in New England and is considering whether to expend limited recruiting efforts in the Boston geomarket -- which in 1980, was relatively low-income -- or in more affluent geomarkets nearby:

> Where would you concentrate your energies? Ideally, you would seek communities with a high proportion of students already predisposed toward institutions such as your own. The Market Segment Model would provide this information through segment percentages for the community in question. Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots if you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots. Your efforts would surely be better directed toward three of the four communities in the bottom band, Manchester, Hartford, and Fairfield County. Simply by knowing a little bit about the students' backgrounds and academic records you could quickly focus your attention on those most likely to consider your kind of institution. Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own."

In this example, historical data on the number of students by Segment is used to divert recruiting energies from lower-income to higher-income local markets. It also shows how geomarkets can be utilized in concert and within student list products. First, decide which geomarkets to concentrate on, then filter students in those geomarkets based on socioeconomic and academic characteristics.

SOMEWHERE: THIS QUOTE FROM CHUN PAGE 57 IS PRETTY SOLID; BUT MAYBE YOU ALREADY HAVE TOO MANY QUOTES?

> As this example makes clear, these models not only "discover" the effects of discrimination; they also automate and perpetuate them for they exploit, rather than remedy inequalities. These correlations are at the heart of what communications scholar Oscar Gandy, writing in 2009..... "identified as"technologies of rational discrimination": unless there is a clear determination not to discriminate, Gandy explained, these technologies perpetuate inequality by creating and comparing "analytically generated groups in terms of their expected value or risk." That homophily drives these groups and correlations "that work" is no accident.

## Homophily and Social Networks

In network science, actors (nodes) are connected to one another via network ties. The first stage in social network analyis is to decide who are the actors and what are the ties that link them. For example, in @RN4982 we can think of students sending SAT scores to colleges as a "two-mode" social network in which students (node 1) send SAT scores -- the network tie -- to colleges (node 2). The second, theorization, stage of social network analysis makes makes statements about which actors are likely to form direct or indirect ties with one another. @RN4975 [p. 90] observes that this stage "builds models that reproduce the abstractions produced in the first. Whatever repeats the initial mapping is true or causal: truth within these mathematical or logical systems as Arendt points out is consistency." For this reason, @RN4975 argues that network science is biased in the direction of its theoretical models and its assumptions -- which are previously proven theoretical models -- in that behavior consistent with the model is defined as "truth" and behavior inconsistent with the model is cast away as "noise."

Homophily has become a core assumption of network science [@mcpherson2001birds]. Homophily means that actors who share common characteristics are likely to form connections with one another, an idea captured by the phrases "birds of a feather flock together" and "similarity breeds connection" [@mcpherson2001birds, p. 415]. By contrast, heterophily is the idea that actors with different characteristics are likely to form connections with one another. According to the review by [@mcpherson2001birds, p. 417], "the classic citation [on homophily] in the sociological literature" was the @RN4990 study of friendship formation in Hilltown and Craftown, two mixed-race housing developments. As survey methods and social network methods advanced in the 1970s and 1980s, evidence in favor of homophily proliferated [@mcpherson2001birds], to the point where homophily became an assumption rather than a finding [@RN4975].

@RN4975 problematizes the idea that homophily is a naturally occurring phenomenon. In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users. Thus, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together" \[p. 82\].

@RN4975 also problematizes academic scholarship about homophily in two ways. First, network science defines behavior consistent with the theoretical model to be "truth" and behavior inconsistent with the theoretical model to be noise. @RN4975 re-analyzed data about friendship patterns in Hilltown and Craftown and found substantial evidence supporting heterophily in the Black residents, but @RN4990 focus on friendship formation for white residents. Second, drawing heavily from rational choice theory, network science treats homophily as the result of "voluntary" actions by individuals; "it erases historical contingencies, institutional discrimination, and economic realities" [@RN4975, p. 95] that underlie behavior consistent with homophily.

In @RN4982, Chapter 4 -- "The Company We Keep: Colleges and Their Competition" -- conducts a social network analysis to identify which institutions are in competition with one another. A network tie between two institutions occur if "15 percent of the students who sent SAT scores to the first institution also sent a score to the second that 15 percent of students who sent an SAT score to the second institution also sent one to the first" [@RN4982]. Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another. These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.

Homophily is the core finding of these network analyses. @RN4982 [p. 42] state that "On the basis of this analysis, we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^2] Subsequent analyses investigate competitive overlap across institutional type -- e.g., private flagship, private selective college, public flagship -- overlaying tuition price and student student socioeconomic status (SES). The analyses reveal substantial overlap between private selective colleges and private flagship universities, charge the highest prices and enroll students with much higher SES than other institutional types. In conclusion, @RN4982 [p. 72] states,

[^2]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p. 54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector. For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."

> We now know why students so seldom speak of their own social or family backgrounds in explaining how they go about choosing a college. They have no need to. Students describe themselves socially simply by telling us the colleges and universities in which they are interested. The layering of collegiate competition is primarily a socioeconomic layering. The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students. Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.

@RN4975 discusses the negative consequences arising from homophily. First, when we assume homophily is the naturally occurring behavior of individuals rather than a consequence of societal structures, then we ignore the structures that give rise to the behaviors we observe. Second, when we "engineer homophily" -- through algorithms or policies -- we make it less likely for heterophily to emerge, eliminating possibilities -- of friendships, of matches -- that may have occurred in the absence of such engineering.

The Market Segment Model manifests both negative consequences of homophily. First, @RN4982 observes hierarchical hierarchical matching between the socioeconomic status of students and the socioeconomic status of colleges, but interprets these findings as naturally occurring phenomena rather than the consequences of historic structural inequality in the opportunity structure of the U.S. Second, the Market Segment Model not only describes the structure of college choice, it influences the future structure of college choice by becoming the basis for products and practices. The Market Segment Model is incorporated into the College Board Enrollment Planning Service (EPS) and Student Search Service. This product amplifies the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.

## Geomarkets and Structural Inequality

**Geomarkets and socioeconomic inequality**. We argue that geomarkets -- and the Market Segment model more broadly -- are built on structural inequality, exhibit structural inequality, and are used in ways that amplify structural inequality.

The Market Segment Model is explicitly a function of socioeconomic status. @RN4982 state that socioeconomic status drives whether students belong to the local, in-state, regional, or national segment. In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status. Colleges prioritize geomarkets depending on the number of students from desired students. When purchasing student lists, @RN4982 suggest that affluent coleges should consider targeting affluent geomarkets. Within geomarkets targeted by a student list purchase, colleges may use additional filters to capture student market segment and these filters may systematically exclude low-income students. Finally, in large metropolitan areas, the borders of geomarkets may be drawn in ways that separate affluent from less affluent communities. @RN4982 [pp. 11-12] state that "each major metroplitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs." We present the following hypothesis.

**H1**. In large metropolitan areas that contain multiple geomarkets, we expect significant socioeconomic inequality (e.g., income, parental education) between geomarkets.

**Geomarkets and racial inequality**. The relationship between geomarkets and racial inequality is less obvious because race is not an explicit component of the Market Segment Model and @RN4982 make no mention of race. The question of whether geomarket borders are correlated with race is a question about structural racism. Structural racism is "a form of systematic racial bias embedded in the 'normal' functions of laws and social relations" [@RN4760, p. 1143], whereby processes viewed as neutral or common sense systematically advantage dominant groups and disadvantage marginalized groups.

Colorblind racism refers to behaviors and ideology that ignore the significance of race and racial discrimination [CITE LATER WORK BY BONILLA-SILVA]. The Market Segment Model can be seen as an example of colorblind racism because @RN4982 develop an analysis of the structure of college choice that completely ignores race. In turn, enrollment management professionals interpet the Market Segment Model and geomarkets as race-neutral recruiting tools. Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975]. DAN HIRSCHMAN CALLS IT A Dan calls geomarket colorblind racist market device.

We argue that geomarkets are seemingly race-neutral inputs that are correlated with race. Geographic borders are the most commonly studied racialized inputs [e.g., @RN4775; @RN4849; @RN4779]. These studies build on the fact that American communities and schools are racially segregated as a consequence of historic and contemporary laws, policies, and practices promoting racial segregation [@RN4551; @RN4801; @rothstein2017color; @RN4975]. Selection devices that categorize people based on geographic location without considering structures that produce segregation are likely to reproduce historical race-based inequality in opportunity. Although previous research has focused racial inequality stemming from the use of zip codes as an input [e.g., @RN4849], College Board geomarkets are much larger than individual zip codes.

Nevertheless, we argue that geomarket borders may be correlated with race for two reasons. First, borders may be drawn in a racicialized way that follows the countours of racial segregation in residential housing. Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets. Second, geomarket borders drawn along class divides may exhibit racial inequality because of race-based inequality in income and wealth \[CITE\]. We present the following hypothesis:

**H2**. In large metropolitan areas that contain multiple geomarkets, we expect significant racial inequality (e.g., income, parental education) between geomarkets.

MOTIVATE HYPOTHESIS ABOUT INCREASE/DECREASE OVER TIME IN BETWEEN-GEOMARKET RACIAL INEQUALITY. POTENTIAL FACTORS MIGHT BE GENTRIFICATION; RISING INCOME INEQUALITY

**Product utilization and racial inequality**. In addition to geomarket borders being correlated with class and race, we argue that geomarkets are likely to be used in ways that amplify socioeconomic and racial inequality in college access. Student lists are purchased by university enrollment personnel or by enrollment management consultants acting as agents on behalf of a university. Student list products are designed to filter on multiple search filters simultaneously, and they grant administrative discretion over which filters to select, which filter thresholds to select, and how many purchases to execute.

Compared to standardized selection devices, scholars tend to find that administrative discretion over selection devices causes structural inequality to increase [@RN4883; @RN4881; @RN4795]. Discretionary selection devices allow explicit or implicit individual bias to affect selection decisions [@RN4794; @RN4801]. Even without implicit bias, discretionary selection devices can increase structural inequality when decision-makers have incomplete knowledge about how the inputs they choose are correlated with race. As such, discretionary selection devices are particularly sensitive to racialized inputs that are perceived to be race-neutral. For example, research shows that Americans dramatically underestimate the magnitude of racial income inequality [@RN4884].

Student list purchases that filter on College Board geo-markets do not not inherently exclude communities of color. Indeed, a university could filter on geomarekts in order to explicitly target an area with a high number of low-SES students or students of color. As such, the incorporation of geomarkets into the student list product creates new ways to target underrepresented communities and also creates new ways to discriminate against these communities. For example, an enrollment professional can reason, there are a lot of 'good' students in the "Contra Cost County" (CA06) and the "Alameda County excluding Oakland" (CA08) geomarkets, but not so many in the "City of Oakland" (CA07) geomarket.

We conceive of geomarkets as a "colorbind racist market device" in that geomarkets are often correlated with race, but geomarkets are likely perceived as race-neutral by people making decisions about student list purchases.[^3] Student list purchases do not show how the characteristics of targeted prospects compare to the characteristics of the surrounding community. From this perspective, racial inequality from the use of geomarkets in student list purchases is likely to be an unintentional result of using a seemingly race-neutral product. Additionally, purchases that specify multiple filters (e.g., test scores, grades, intended major) including geomarkets can yield unintended racial inequality because administrators have incomplete knowledge about how the intersection of these filters interact with local patterns of segregation.'

[^3]: Colorblind racism is the result of ideology that attemps to minimize race; using geomarkets one doesn't have to minimize race, because race has already been eliminated from geomarkets and Market Segment Model.

We argue that unintentional racial inequality can arise because the Market Segment Model -- and the enrollment management profession more generally -- are oriented towards realizing enrollment goals rather than equality of opportunity for students. For example, @RN4982 recommends that instead of targeting the Boston geo-market, a (hypothetical) college targets Manchester-NH, Harford-CT, Burlington-VT, and Fairfield County-CT. The rationale is that students from these local markets are more likely to have socioeconomic characteristics associated with attending this college. By contrast, in 1980, the Boston geomarket was relatively low income and -- though ignored by @RN4982 -- was more racially diverse. @RN4982 [p. 44] comment on this recommendation:

> On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged. Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good. The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.

Consistent with @RN4982, prior research finds that selective private colleges and universities target affluent schools and communities [@RN4330; @RN4733]. The same is true for the out-of-state recruiting efforts of public research universities [@RN4759; @RN4758]. We argue that student list purchases that target affluent geomarkets of large metropolitan areas are likely to exclude more students of color compared to student list purchases that target the entire metropolitan area or target less-affluent geomarkets in the MSA. We test this argument by analyzing actual student list purchases of public reearch universities. We examine student list purchases that filter on geomarket. In student list purchases that did not filter on geomarket, we simulate what would happen if they had.

HX: STATE HYPOTHESIS???

# References

::: {#refs}
:::

\clearpage

<!-- TABLES -->

```{r market-segment-characteristics, echo = F}
#| label: tbl-market-segment-characteristics
#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'

market_segment <- data.frame(
  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
)

kable(
  x = market_segment,
  col.names = c('Characteristic','Local','In-state','Regional','National'),
  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
)
```

\clearpage

```{r sample-institutional-profile, echo = F}
#| label: tbl-sample-institutional-profile
#| tbl-cap: 'Simplified sample institutional profile for Connecticut Market 3: Fairfield County'

sample_inst <- data.frame(
  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
  local = c(1,0.2,0,0.0,0,0.0),
  in_state = c(58,4.8,25,8.4,2,3.0),
  regional= c(69,4.1,35,7.0,5,5.8),
  national = c(109,2.9,61,5.3,3,0.8),
  total = c(237,3.3,121,5.9,10,1.8)
)

kable(
  x = sample_inst,
  col.names = c('','Local','In-state','Regional','National','Total'),
  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
)
```

\clearpage

```{r market-segment-institutions, echo = F}
#| label: tbl-market-segment-institutions
#| tbl-cap: 'Top 5 institutions in terms of number of scores sent by segment, Connecticut Market 3: Fairfield County'

market_segment_inst <- data.frame(
  rank = c(1,2,3,4,5),
  local_num_scores_sent = c(277,261,183,103,100),
  local_pct_scores_sent = c(50.4,47.5,33.3,18.7,18.2),
  local_inst_type = c('priv ma','priv ma','priv ma','pub doct','pub ma'),
  state_num_scores_sent = c(757,515,438,183,177),
  state_pct_scores_sent = c(63.1,43.0,36.5,15.3,14.8),
  state_inst_type = c('pub doct','pub ma','pub ma','pub ma','pub ma'),
  regional_num_scores_sent = c(610,348,272,248,197),
  regional_pct_scores_sent = c(36.7,20.9,16.3,14.9,11.8),
  regional_inst_type = c('pub doct','priv doct','priv doct','pub doct','pub doct'),
  national_num_scores_sent = c(1226,371,327,312,308),
  national_pct_scores_sent = c(32.6,9.9,8.7,8.3,8.2),
  national_inst_type = c('pub doc','priv doct','priv res','priv doct','priv doct')
)

kable(
  x = market_segment_inst,
  col.names = c('','Num','Pct', 'Type','Num','Pct','Type','Num','Pct','Type','Num','Pct','Type'),
  #caption = 'Top 5 institutions in terms of number of scores sent, by segment \\\\Connecticut Market 3: Fairfield County'
) %>% 
  #kable_styling(full_width = F) %>% 
  add_header_above(c(' '=1,'Local (N=550)' = 3, 'In-state (N=1,199)' = 3,'Regional (N=1,664)' = 3, 'National (N=3,766)' = 3))
```

```{=html}
<!--
\begin{landscape}
\end{landscape}

\restoregeometry
-->
```

### 2024-07-08 text/old_spring2024/cb_geomarket_lit_review.qmd
initial commit, copying files from messed up repo

--- 2024-07-08 text/old_spring2024/background_em_long_theory_5_30_2024.qmd
@@
 ---
-title: "Structuring College Access: College Board Geo-markets and the Market Segment Model"
+title: "Structuring College Choice: The Curious, Mundane Case of College Board Geo-markets"
 author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
@@
   margin=1in
-fontsize: 11pt
+fontsize: 12pt
 header-includes:
@@
 
+```{=tex}
+\pagenumbering{roman}
+\newpage
+\pagenumbering{gobble}
+\pagenumbering{arabic}
+```
 # Introduction
-
-hey ho This many words: {{< words-total >}}
-
-COULD START WITH HISTORICAL VIGNETTE OF ZEMSKY PARTNERING WITH COLLEGE BOARD TO CREATE THE MARKET SEGMENT MODEL.
-
--   quotes on "admissions officers are invariably tellers of stories" and the goal of the market segment model is to quantify these stories.
 
@@
 
-```{=tex}
-\pagenumbering{roman}
-\newpage
-\pagenumbering{gobble}
-\pagenumbering{arabic}
-```
-# Background: Enrollment Management
+# Background: Enrollment Management and the Enrollment Funnel
 
@@
 The EM industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers that produce market research, software, and and consulting services (e.g., College Board, ACT, EAB, Huron, Slate, PowerSchool).
-
-The process of recruiting students involves many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor.
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
 
@@
 Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel they focus on.
-The majority of scholarship focuses on the admissions stage, for example which admissions criteria are used [@RN4874; @RN4993] and which applicants are admitted [@RN4131; @RN3522] \[CITE SOMETHING BY BASTEDO\].
+The majority of scholarship focuses on the admissions stage, for example which admissions criteria are used [@RN4874; @RN4993] and which applicants are admitted in practice [@RN4131; @RN3522] \[CITE SOMETHING BY BASTEDO\].
 
 A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
-Scholarship on recruiting has conceptualized recruiting behavior as an indicator of college enrollment priorities [@RN4758] and identifies recruiting practices as a mechanism for social reproduction in college access [@RN4324; @RN3519].
+Scholarship on recruiting has conceptualized recruiting behavior as an indicator of college enrollment priorities [@RN4758] and identifies recruiting practices as a mechanism for social reproduction in college access.
 
@@
 @RN3519 finds that college admissions officers visit high schools as a means of maintaining positive relationships with guidance counselors at affluent feeder schools.
-Consistent with this finding, quantitative case studies of off-campus recruiting visits show that selective private colleges visit affluent, predominantly white schools and communities, with a disproportionate number of visits to private high schools [@RN5014; @RN4330].
+Consistent with this finding, quantitative case studies of off-campus recruiting visits show that selective private colleges visit affluent, predominantly white schools and communities, with a disproportionate number of visits to private high schools [@RN4733; @RN4330].
 Additionally, public research universities often make make more out-of-state recruiting visits than in-state visits, and these out-of-state visits focus on the same set sorts of schools targeted by selective private universities [@RN4759; @RN4758].
@@
 We make two observations about the recruiting literature.
-First, consistent with scholarship on political economy [@RN4517; @cottom2017lower] \[cite Stevens; cottom\] and industrial organization \[cite Winston\], the recruiting literature suggests a hierarchical competition in which different "types" of postsecondary institutions target hierarchically arranged customer niches.
+First, looking across studies, the recruiting literature suggests a hierarchical competition in which different "types" of postsecondary institutions target different customer niches.
 For example, for-profit institutions target adults in communities not being targeted by traditional colleges [@cottom2017lower].
 Non-selective private non-profits target prospects who might be considering community college or a local state colleges [@RN4324].
-Selective private colleges target private high schools, with a particular link between sectarian high schools and colleges of the same faith [@RN5014], but these schools are increasingly targeted by out-of-state public research universities, suggesting growing competition for a coveted market niche.
-This process of hierarchically organized colleges matching to hierarchical student segments exemplifies the idea that higher education is mechanism of social reproduction [@RN2377; @RN4873]
+Selective private colleges target private high schools, with a particular link between sectarian high schools and colleges of the same faith [@RN4733], but these schools are increasingly targeted by out-of-state public research universities, suggesting growing competition for a coveted market niche.
 
-Second, by focusing on the behaviors of individual colleges, the literature makes the implicit assumption that recruiting is something done by individual colleges.
-Why then, does recruiting behavior look so similar across colleges of a particular "type"?[^1]
-
-[^1]: New institutional theory offers one explanation [@RN527].
+Second, the recruiting literature -- and scholarship on enrollment management more broadly -- assumes that recruiting is a function of individual organizational behavior \[alternatively, "assumes that recruiting is something done by individual colleges"\].
+Why then, does recruiting behavior look so similar across colleges of a particular "type"?
+Institutional theory offers a compelling explanation [@RN527].
 In highly developed organizational fields like higher education, forces of competitive and institutional isomorphism -- including institutional locics -- compel organizations adopt similar practices as their peers.
 Practices deemed advantageous spread quickly, for example, public research universities targeting nonresident prospects to compensate for declines in state appropriations.
-These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.<!--WHAT DOMINA (2017) SAID ABOUT LOGICS, P.
+These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.
+<!--WHAT DOMINA (2017) SAID ABOUT LOGICS, P.
 322: Like other distinctively modern institutions, contemporary schools are governed by standardized and rationalized institutional logics (Baker 2014, Bowker & Star 2000, Timmermans & Epstein 2010, Meyer 1977).
@@
 
-@RN5012 argues that the process of college access is structured by third-party recruiting products utilized by colleges.
-@RN5012 deconstructs the College Board *Student Search Service* product.
+Whereas most scholarship assumes that recruiting is something done by individual colleges, @RN4985 argues that structural inequality is embedded in third-party recruiting products utilized by colleges.
+@RN4985 deconstructs the College Board *Student Search Service* product.
 Student list products are "selection devices" [@RN4778] that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
-Drawing from @RN4786, @RN5012 argue that several search filters (e.g., AP test score, zip code, geodemographic segment) are "racialized inputs," defined as ostensibly race-neutral inputs that are correlated with race because disadvantaged racial groups have historically been excluded from the input.
-Using a national sample of high school students from the High School Longitudinal Study of 2009 (HSLS), they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
+Drawing from @RN4786, @RN4985 argue that several search filters (e.g., AP test score, zip code, geodemographic segment) are "racialized inputs," defined as ostensibly race-neutral inputs that are correlated with race because disadvangaged racial groups have historically been excluded from the input.
+Using a national sample of high school students, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
 Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
 
-[NEED CLEANER TRANSITION]Two geographic search filters -- geo-market and geodemographic segment -- that @RN5012 observe in lists purchased by public could not recreate using HSLS.
-Using geographic borders to determine selection for an intervention has been widely studied as a mechanism of structural inequality because of historic and contemporary racial segregation in American communities and schools [e.g., @RN4775; @RN4849; @RN4779].
-The geo-market and geodemographic segment search filters of the College Board *Student Search Service* are examples of creating new geographic borders for the purpose of targeting students for educational opportunities.
-The geo-market filter slices states and metropolitan areas into smaller, "local" markets.
-For example, CA08 refers to "Alameda Country excluding Oakdland" and a university might purchase prospect contact information by filtering on this geographic region, alongside additional filters.
-The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors.
-This manuscript focuses on geo-markets.
-The first question we asked ourselves was, "where did these geo-markets come from?" The answer is, market research.
+# OLD SHIT
 
-```{=html}
-<!--The enrollment management industry emphasizes market research.
-Market research is the process of producing information that informs decisions about product design or how to find customers.   -->
-```
-The enrollment management industry puts great stock in market research.
-@RN4986 define market research as all "activities intended to increase understanding of a market and consumer behavior in that market" (p.
-17).
-At the most basic level, (in-house) market research is a function of intentional record-keeping.
-@RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
-In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited.
-In the chapter entitled "Travel," @RN3519 [p.
-69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."
+We situate this scholarship broadly within the literature on social reproduction and narrowly within the literature on inequality in colege access, with a particuar focus on enrollment management behavior.
 
-The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals.
-In the late 1970s, College Board provided Carleton College with funding and data for a six-year project that analyzed college choice decisions by households in six metropolitan areas.[^2] This project was the basis for the landmark book *Applying Market Research in College Admissions* [@RN4986], published by College Board, and was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll.
-<!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.
--->
-
-[^2]: In the 1970s, Carleton College became a leader in applying market research to student recruiting, under the leadership of Larry Litten, the then director of institutional research.
-Litten would later become Associate Director of the Consortium of Financing Higher Education (COFHE).
-
-Similarly, in 1979 The College Board provided Robert Zemsky, a professor of education at UPenn, with data and funding to develop the *Comprehensive Undergraduate Enrollment Planning Project* [@RN4982].
-In 1983, The College Board published *The Structure of College Choice* by Robert Zemsky and Penney Oedel, developed the "market segment model" and defined local "geo-markets" based on a comprehensive analysis of SAT score-sending by prospective students to coleges and universities.
-In 1984/5, the market segment model and geo-markets became the basis for the College Board Enrollment Planning Service (EPS), the market research division of College Board.
-
-This manuscript analyzes the market segment model and geo-markets as a case study in order to investigate market research -- and its commodification into recruiting products -- as a mechanism for social reproduction.
-At root, third-party market research helps universities identify who are the desirable students, where are they located, which institutions are they considering, and how best to reach them.
-Today, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development \[e.g., CITE RNL\], but most only available to clients.
-EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
-Despite being central to enrollment management for at least 40 years [@RN4986], the role of third-party market research has been ignored by scholarship on college access, which views enrollment management behavior as the autonomous decisions of individual colleges, albeit structured by institutional logics [@RN4758; @RN4873; @RN3519].
-We argue that market research structures who colleges recruit, in ways that reinforce existing social stratification.
-
-# Quantification
-
-```{=html}
-<!--
-CUT GENERAL PARAGRAPH ON REPRODUCTION
-
-Within the sociology of education, @RN2377 describes education as a sieve, allocating people to opportunities in socially stratified, capitalist societies.
-This role highlights "Max Weber's core insight that education has a dual character" [@RN2377, p.
+@RN2377 describes education as a sieve, allocating people to opportunities in socially stratified, capitalist societies.
+This role highlights "Max Weber's core insight that education has a dual character" [\@ @RN2377, p.
 129], on one hand, serving social mobility by redistributing access to privileged positions and, on the other, serving social reproduction by granting privileged families disproportionate access to the educational institutions that confer pathways to privileged positions.
@@
 322\] argue, "the relevant question is not whether schools produce inequality; rather, the relevant questions are, when and where do schools produce inequality, \[and\] how do they do so." As @RN2377 show, whereas early research on social stratification sought to identify patterns of inequality [e.g., @RN4873; @RN290; @RN1268], the sociology of education is now more concerned with identifying the mechanisms underlying these patterns.
--->
-```
-We situate our study amidst scholarship on within scholarship on quantification as a mechanism that reproduces inequality.
-@RN5007 [p.
-402] define quantification as "the production and communication of numbers" (p.
-402) and introduce reacivity, discipline, and authority as three interrelated themes that describe the effects of quantification.
-Reactivity is the idea that salient quantitative measures cause people and organizations to change their behavior.
-Discipline is the idea that quantification causes actors to react in a particular way.
-Authority is the idea that quantification changes decision-making power, often weakening the discretion of local decision-makers.
+<!--  [NOTE: THIS SETS UP IDENTIFYING MARKET RESEARCH AS AN IMPORTANT **NEW** MECHANISM OF SOCIAL REPRODUCTION THAT SOCIOLOGY OF EDUCATION SHOULD BE STUDYING!!!] -->
 
-Scholarship about U.S.
-News & World Report (USNWR) Law School rankings demonstrate the reactivity, discipline, and authority effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
-Developed as a means of informing prospective students about quality, rankings affected which law schools applicants applied to, the intended effect.
-However, rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
-(clients evaluated firms based on the prestige of the law schools that firm lawyers attended) and
-
-USNWR developed the rankings as a means to inform prospective students about law school quality. However, rankings affected
-
-US News developed law schools rankings as a means of informing prospective students about law school quality.
-In turn, rankings affected how prospective students conceive law school quality and which schools they applied to.
-Rankings also affected the cognition and behavior of other actors.
-Looking downstream, entities that received law school students as an input -- for example, law firms -- used rankings as a hiring criterion because their clients evaluated the organization based on the the prestige of the schools their firm's lawyers attended.
-
-Drawing from Foucault \[CITE DISCIPLINE AND PUNISH\], discipline is the idea that surveilance -- attributes of a law school being quantified and aggregated into a public-facing metric -- compels people and organizations to react in a particular way.
-For example, the reputational assessment of peer leaders as an important component of most US News rankings.
-Bastedo and Bowman showed that once (undergraduate) USNWR rankings became popular, the reputational assessment of peers primarily became a function of previous USNWR rankings.
-Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots schools felt pressured to take them seriously" \[QUOTE FROM ESPELAND AND STEVENS\].
-For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid \[CITE\].
-
-The theme of authority is that quantification destabilizes and weakens the authority of local decision-makers.
-For example, law school rankings weakened the authority of law schools to make independent decisions about which applicants to admit, ideal class size, and curricular offerings \[CITE THE BOOK\].
-In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.
-
-The analysis of UK League Tables by @RN5005 show how quantification can be a mechanism of social reproduction.
-In 1992, as part of a the broader *New Public Management* effort to instill market competition amongst public providers [@RN161], the UK government started publishing school "league tables" that ranked schools based on student performance on the high-stakes General Certificate of Secondary Education (GCSE) exams that students take at age 16.
-@RN5005 argue that quantifying the quality of schools entrenches class-based inequalities between schools particularly when the measure of quality -- standardized test scores -- is substantially a function of the class composition of schools.
-School league tables discipline households and schools to conceive of school quality in terms of test scores and they facilitate making evaluative comparisons between schools that are geographically close or distant.
-Compared to working-clas parents, parents with professional/managerial occupations are more aware of school league tables and have financial resources necessary to move to neighborhoods near a "good" school.
-
-Using school-level data and data on census data on occupational class measured at the local authority level (similar to a U.S.
-county), @RN5005 find that, following the adoption of league tables, local areas with higher performing schools experience growth in the share of managerial/professional occupational classes and a decline in the share of working class families.
-Additionally, using longitudinal survey data on 1% of the population of England and Wales and a diference-in-difference estimator, the authors find that managerial/professional households are more likely to move to local areas with higher ranked schools following the introduction of school league tables.
-Overall, @RN5005 show that quantifying which schools are high-performing contributes to social reproduction because managerial/professional households respond to consumer-facing metrics by moving to more expensive neighborhoods, near higher performing schools.
-Schools that performed well on initial league tables tended to be schools with a high share of professional/managerial families.
-Therefore, quantification takes a snapshot of existing social stratification and that snapshot becomes a power unto itself, amplifying the effect of previous strctural inequality on subsequent structural inequality.
-
-In the broader quantification literature, critical data studies [e.g., @RN4975; @RN4849; @RN4772] and sociological analyses of technology and inequality [e.g., @RN4778; @RN4786; @RN4877; @RN4810] introduce concepts salient to the analysis of market research.
-Our focus is on "selection devices," which utilize some combination of inputs and discretion to categorize places, people, and organizations for the purpose of allocating opportunities or interventions [@RN4778; @RN5012].
-
-A multi-disciplinary literature examines structural inequality produced by predictive analytics that capture correlational relationships [@RN4975].
-Predictive analytics utilize actuarial methods, which were pioneered by the insurance industry and proceed in two steps [@RN4835; @RN4778].
-First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
-Second, apply these results to future cases in order to make predictions and to assign levels of risk to each case.
-Importantly, these models predict outcomes based on correlations without requiring knowledge about causal relationships.
-@RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
-
-The use of predictive analytics reproduces structural inequality.
-The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
-Although these correlations may b e a function of enduring structural inequality, the underlying causes are not considered by applications of predictive models.[^3].
-This algorithm recommends high purchase volume in zip codes that sent large number of students in previous years.
-@RN4794 \[p.
-224\] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities.
-@RN4851 refers to this phenomenon as the"ratchet effect," whereby disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies disproportionate targeting/exclusion.[HMM.
-PERNICIOUS FEEDBACK LOOPS WOULD SEEM TO FIT HERE]
-
-[^3]: For example, the College Board *Student Search Service* allows colleges to filter student list purchases based on zip code, amongst other filters.
-The consulting firm Ruffalo Noel Levitz offers an algorithm that recommends how many prospect profiles a college should purchase from each zip code, based on historic enrollment data from the college [@jmu_rnl]
-
-@RN4849 shows how predictive analytics create "weapons of math destruction" (WMD), which are defined as algorithms that have three characteristics: opacity, meaning that the algorithm is proprietary or unknown to the people it affects; scale, meaning that the algorithm can affect large numbers of people; and damage, meaning that the algorithm can reinforce past inequality.
-An exemplar WMD is the XXX model used to predict a prisoner's chances of re-arrest.
-How the algorithm uses inputs to predict the outcome is unclear (opacity).[^4] The algorithm is used by 24 states (scale).
-Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
-@RN5006 observe that such "pernicious feedback loops" [@RN4849] are similar to the "ratchet effect" [@RN4851] and are extreme versions of "reactivity" from @RN5003.
-
-[^4]: By contrast, U.S. law school rankings and UK leage tables are not WMDs because these algorithms are transparent [@RN5006].
-
-Due to data limitations, weapons of math destruction (WMD) -- and predictive analytics more broadly -- often utilize "proxy" variables, which are variables that are correlated to a variable of interest [@RN4849].
-FICO credit scores are based solely on individual repayment history and do not utilize proxies.
-By contrast, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of these others.
-@RN4849 [p.
-146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
-
-@RN4975 argues that ideas about homophily often underlie the use of correlation to predict future behavior.
-Homophily means that actors who share common characteristics are likely to form connections with one another.
-The ubiquitous colloquialism is "birds of a feather flock together." Homophily has become a core assumption for network science [@mcpherson2001birds], in which actors (e.g., high school students) are linked to one another through direct and indirect network ties (e.g., submit SAT scores to the same college).
-Because network science often draws from rational choice theory, it assumes that homophily is the result of voluntary action by individuals, which "erases historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
-95] that underlie behavior consistent with homophily.
-@RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.
-Thus, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together" \[p.
-82\].
-
-Before the emergence of commercial social networks, homophily was the stated motivation behind geodemography, a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
-Geodemography emerged in the 1970s from parallel efforts in the U.S.
-and U.K to classify consumers based on publicly available census data.
-Market segments are subgroups within a larger market that have similar consumer demand.
-In the U.S., the Claritas Corporation classified zip codes into market segments that would be useful for direct mail marketing campaigns [@RN4957].
-Clusters were given evocative names like "Blue Blood Estates," "\[Archie\] Bunker's Neighbors", and "Share Croppers." Because U.S.
-publicly available U.S.
-Census data disaggregated data only to the zip code-level, the Claritas Corporation had a strong incentive to argue that consumer behavior within zip codes was very similar [@RN4957].
-@RN4565 emulated the behavior of Claritas by developing the *Segment Analysis Service: An Educationally Relevant Geodemographic Tagging Service*, which was an "add-on" to the *Student Search Service* student list product.
-Describing its conceptual underpinnings, @RN4565 [p.
-1] states, "the basic tenet of geodemography is that people with similar cultural backgrounds, means, and perspectives naturally gravitate toward one another or form relatively homogeneous communities; in other words, birds of a feather flock together." The Segment product allocated each high school and each census tract into a distinct "educational neighborhoods" thought to reflect useful market segments for colleges looking for students.
-<!--
-Negative consequences of homophily
-Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.-->
-
-Related to the creation of market segments, @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Historically, classifications were binary.
-Consumers with "good" credit were offered loans and those with bad credit were not.
-The innovation of individual FICO scores -- and, later, other credit scores -- enabled merchants to classify consumers into many, fine-grained groups or along a continuum.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].
-People previously denied credit were now granted credit on different terms.
-For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-Thus, at the bottom of the continuum, contemporary classification situations produce "predatory inclusion" -- exemplified by payday loans and for-profit colleges -- which is the logic of "including marginalized consumer-citizens into ostensibly democratizing mobility schemes on extractive terms" [@RN4774, p.
-443].
-Whereas FICO scores assign individuals a credit-worthiness value based solely on their individual record of financial transactions [@RN4877], other innovations classify consumers by integrating credit scores with demographic and geographic information [@RN4849].
-<!--
-The innovation of the FICO score assigned individuals a numeric credit-worthiness value, based solely on their individual record of financial transactions [@RN4877].
-In turn, the FICO score -- and, later, other credit scores -- enabled merchants to classify consumers into many, fine-grained groups or along a continuum.
+Given the substantial social and economic benefits of graduating from college \[CITE\], and the greater benefits of graduating from a selective college \[CITE CHETTY AND STUFF\], sociology has paid particular attention to social stratification in the transition from high school to college.
+A significant strand of research focuses on cultural capital as a broad mechanism contributing to social reproduction in access to college, particularly selective colleges [@RN1228; @RN1241].
+@RN4991 observes that "even as even as colleges' definitions of merit have evolved, from the post-World War II growth of standardized testing,\ldots to a renewed focus on holistic review," affluent, college-educated families have maintained a competitive advantage college access \[CITE\].
+These families are socialized to place value on selective colleges, have financial and cultural resources to devote to this competition, and are concentrated in communities and schools that orient student achievement and activities towards the admissions criteria of selective institutions \[CITE\].
+<!--  CAN YOU SAY THIS MORE EFFICIENTLY; LIKE ON THE STUDENT DEMAND SIDE, SOCIOLOGY OF EDUCATION HAS FOCUSED ON CULTURAL CAPITAL, WHEREBY PRIVILEGED FAMILIES EMBODY THE CHARACTERISTICS OF DESIRABLE APPLICANTS; AND ON SUPPLY SIDE COLLEGES REALLY WANT THESE APPLICANTS.
 -->
 
-Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
-To our knowledge, prior scholarship on quantification within the sociology of education has not examined the role of third-party market research in classifying consumers.
-Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
-Furthermore, prior scholarship has not examined how market research becomes embedded in products that structure how consumers are matched to producers.
-This manuscript examines how the "market segment model" [@RN4982] categorized high school students into vertical market segments -- local, in-state, regional, and national -- and simultaneously created local "geo-markets" that were evaluated based on their composition of student market segments.
-These geo-markets are highly correlated with race and class.
-The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target, and geo-markets were later incorporated as a filter in the College Board *Student Search Service*.
-Prior scholarship in the sociology of education shows the effect of quantification on social reproduction [@RN5007; @RN5005].
-This study provides insight about the mechanism underlying this effect, by showing how geo-markets are highly correlated with race and class and by showing how geomarkets are utilized within student list purchases to reproduce historical class and racial inequality in college access.
+On the supply-side of college access, an interdisciplinary literature focuses on enrollment management.
+The term "enrollment management" (EM) refers to a profession, an administrative structure, and an industry.
+As a profession, EM integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p.
+xiv].
+As an administrative structure, the EM office typically controls the activities of admissions, financial aid, and recruiting [@RN2406].
+The EM industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers in the consulting, market research, and technology spaces (e.g., College Board, ACT, EAB, Huron, Slate, PowerSchool).
 
-PARAGRAPH ABOUT CONTRIBUTION THAT IS MOVED HERE. WORK THESE IDEAS IN SOMEWHERE
+Figure @fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
+The funnel begins with a large pool of "prospects" (i.e., prospective students) that the university would like to enroll.
+"Leads" are prospects whose contact information has been obtained.
+"Inquiries" are prospects that contact the institution and consist of two types: first, inquiries who respond to an initial solicitation (e.g., email) from the university; and second, "student-as-first-contact" inquiries who reach out to the university on their own (e.g., sending ACT scores).
+<!--  Applicants consist of inquiries who apply plus "stealth applicants" who do not contact the university before applying.
+-->The funnel narrows at each successive stage in order to convey the assumption of "melt" (e.g., a subset of "inquiries" will apply).
+Practically, the purpose of the enrollment funnel is to inform recruiting interventions that target one or more stages.
+These interventions seek to increase the probability of "conversion" across stages [@RN4322].
+At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz].
+Purchased leads are served emails, brochures, and targeted social media advertisements designed to solicit inquiries and applications [@RN4895].
 
--   NOTE: THE TEXT FROM THE CASE STUDY WHERE YOU SAY HOW MANY COLLEGES WERE USING EPS CAN GO HERE; IT SPEAKS TO SALIENCE; WHEN TALKING ABOUT US NEWS AND UK LEAGUE TABLES ARE CONSUMER FACING AND HIGHLY SAIENT; CAN SAY THAT EPS IS BUSINESS FACING BUT HIGHLY SALIENT
+Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel they focus on.
+The majority of scholarship focuses on the admissions stage, for example which admissions criteria are used [@RN4874; @RN4993] and which applicants are admitted in practice [@RN4131; @RN3522] \[CITE SOMETHING BY BASTEDO\].
 
-Reviewing recent scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Law school rankings and school league tables seem to exemplify this sort of salience.
-By contrast, many quantification projects -- for example, human rights indicators -- have "little oomph" [@RN5006, p.
-265].
-FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created." The widely known studies of quantification in education \[e.g., \] [@RN5003; @RN5005] tend to study consumer-facing quantification.
-By contrast, market research targets businesses rather than consumers.
-To our knowledge, prior research on quantification in education does not examine market research.
-Unlike studies like @RN5005 we cannot show the effect of quantification.
-However, by analyzing and simulating student list purchases we can show *how* the "geo-markets" reproduce inequality \[NOTE: THIS PARAGRAPH CONTAINS TWO IDEAS; MIGHT MOVE OR CUT ONE OF THEM LATER\].
+A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
+Scholarship on recruiting has conceptualized recruiting behavior as an indicator of college enrollment priorities [@RN4758] and identifies recruiting practices as a mechanism for social reproduction in college access.
 
-# The Market Segment Model
+Ethnographies by @RN3519 and by @RN4407 analyze the connections between privileged colleges and privileged high schools.
+@RN4407 shows how guidance counselors at elite private school guidance counselors lobby admissions counselors on behalf of marginal students.
+@RN3519 finds that college admissions officers visit high schools as a means of maintaining positive relationships with guidance counselors at affluent feeder schools.
+Consistent with this finding, quantitative case studies of off-campus recruiting visits show that selective private colleges visit affluent, predominantly white schools and communities, with a disproportionate number of visits to private high schools [@RN4733; @RN4330].
+Additionally, public research universities often make make more out-of-state recruiting visits than in-state visits, and these out-of-state visits focus on the same set sorts of schools targeted by selective private universities [@RN4759; @RN4758].
 
-The market segment model -- and its associated geo-markets -- is a case that incorporates several themes of the quantification literature and one that is simultaneously historical and contemporary.
-In 1978, Robert Zemsky, a University of Pennsylvania professor, began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE) -- a group of 30 selective private institutions -- to understand questions like, "'Who thinks about Penn?' \[and\] also, 'What other institutions do they think about when they think about us?'" [@RN4982, p.
-x].
-As the project project became more ambitious, the researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincientally the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in tehse analyses was the state' [@RN4982, p.
-x].
-In 1979, College Board began providing data and funding for what became the Comprehensive Undergraduate Enrollment Planning Project (CUEPP).
-In 1983, the College Board published *The Structure of College Choice* [@RN4982], which developed the market segment model, introduced geo-markets, and became the basis for the College Board Enrollment Planning Service (EPS).
-
-Considering the @RN5007 \[p.
-402\] definition of quantification as "the production and communication of numbers," the market segment model is an effort to quantify and communicate the knowledge that admissions officers have about student demand.
-@RN4982 [pp.
-9-10] write that,
-
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting \ldots we believe that the institutions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process."
-
-@RN4982 [p. 10] stress the importance of "knowing the territory":
-
-> A good recruiter knows where to look for prospective applicants, as seen in the students' willingness or eluctance to travel.
-It is necessary to identify not only the most promising communities or poools, but also the specific neighborhoods within those communities -- hence the recruiter's emphasis on feeder high schools.
-From the beginning, we have sought to mold our research to these concepts, to capture and quantify the phenomenon behind the folklore.
-Thus, it is the admissions officers' notion of admissions pools that dictates the geographic units of our analysis." \[p.
-11\]
-
-**Creating the market segment model.** The first step in developing the market segment model was creating "three types of boundaries" [@RN4982, p.
-11], region, state, and community.
-The three regions were New England, Middle States, and the South.
-Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" \[in New England, Middle States, and the South\].
-These enrollment markets, later called "geo-markets" by he College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
-@RN4982 write that,
-
-> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
-Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
-In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."
-
-Having defined these geographic boundaries, high school students were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
-Each institution a student sends SAT scores to can be defined as "local" (institution is in the same local market that the student lives in), "in-state" (same state but different local market), "regional" (same region but different state), or "national." In turn, a "local" student submits more SAT scores to local institutions than they do to in-state, regional, or national institutions.
-An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
-<!--
-A "regional" student submits more SAT scores to regional institutions than they do to local, in-state, or national institutions.
-A "national" student submits more SAT scores to national institutions than they do to local, in-state, or regional institutions.
+Other studies have investigated recruiting behavior by non-selective institutions [@RN4520; @cottom2017lower].
+Analyzing recruiting from the perspective of high school students, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
+@cottom2017lower and @dache2018dangling show that for-profit colleges found a niche in Black and Hispanic communities because traditional colleges ignored these communities.
+<!--  WHAT YOU SAID ABOUT POSECZNICK Posecznick (2017) studied the “Ravenwood College,” a private nonprofit college founded to serve working Black women.
+Consistent with analyses of tuition discounting by Lapovsky (2001), the survival of Ravenwood depends on the success of admissions officers in identifying prospects (e.g., through on- and off-campus recruiting events) and “converting” these prospects to enrolled students.
 -->
 
-The two primary outputs of the market segment model are the (1) Market Segment Profile and (2) the Institutional Profile.
-The Market Segment Profile, which was produced separately for each of the 143 geo-markets, provides information about the number and characteristics of students in each market segment.
-@tbl-market-segment-characteristics reproduces a partial, simplified version of the bottom panel of @RN4982 [Figure 2.1] which represents the Market Segment Profile for "Connecticut Market 3: Fairfield County." @tbl-market-segment-characteristics has separate columns for each market segment (local, in-state, regional, national) and rows show the number of test-takers and characteristics of test-takers.
-For example, there were 550 "local" students in Fairfield county and these students submitted SAT scores to 2.8 postsecondary institutions on average.
-By contrast, there were 1,664 "regional" students who submitted test scores to 4.8 institutions on average.
-For local students, 11.2% had family income greater than \$35,000 and 9% had both parents with a BA.
-For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA.
-Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.[^5]
+We make two observations about the recruiting literature.
+First, looking across studies, the recruiting literature suggests a hierarchical competition in which different "types" of postsecondary institutions target different customer niches.
+For example, for-profit institutions target adults in communities not being targeted by traditional colleges [@cottom2017lower].
+Non-selective private non-profits target prospects who might be considering community college or a local state colleges [@RN4324].
+Selective private colleges target private high schools, with a particular link between sectarian high schools and colleges of the same faith [@RN4733], but these schools are increasingly targeted by out-of-state public research universities, suggesting growing competition for a coveted market niche.
 
-[^5]: @tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market.
-For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution.
-Of the 1,199 in-state students, 757 (63.1) sent scores to institution #1, a public doctoral granting institution, and 515 (43.0%) sent scores to institution #2, a public master's granting institution.
+Second, the recruiting literature -- and scholarship on enrollment management more broadly -- assumes that recruiting is a function of individual organizational behavior \[alternatively, "assumes that recruiting is something done by individual colleges"\].
+Why then, does recruiting behavior look so similar across colleges of a particular "type"?
+Institutional theory offers a compelling explanation [@RN527].
+In highly developed organizational fields like higher education, forces of competitive and institutional isomorphism -- including institutional locics -- compel organizations adopt similar practices as their peers.
+Practices deemed advantageous spread quickly, for example, public research universities targeting nonresident prospects to compensate for declines in state appropriations.
+These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.
+<!--WHAT DOMINA (2017) SAID ABOUT LOGICS, P.
+322: Like other distinctively modern institutions, contemporary schools are governed by standardized and rationalized institutional logics (Baker 2014, Bowker & Star 2000, Timmermans & Epstein 2010, Meyer 1977).
+These logics have profound implications for educational category assignment processes, leading educators, policy makers, scholars, and others to debate the universality of selection-->
 
-INSERT @tbl-market-segment-characteristics ABOUT HERE
+Whereas most scholarship assumes that recruiting is something done by individual colleges, @RN4985 argues that structural inequality is embedded in third-party recruiting products utilized by colleges.
+@RN4985 deconstructs the College Board *Student Search Service* product.
+Student list products are "selection devices" [@RN4778] that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
+Drawing from @RN4786, @RN4985 argue that several search filters (e.g., AP test score, zip code, geodemographic segment) are "racialized inputs," defined as ostensibly race-neutral inputs that are correlated with race because disadvangaged racial groups have historically been excluded from the input.
+Using a national sample of high school students, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
+Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
 
-The *Institutional Profile* describes students who send scores to a particular institution and which majors these students are interested in.
-For each institution, institutional profiles are created separately for students from a particular local market, for all students in a state, or all students in a region.
-@tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the institutional profile of an anonymous institution for students from Fairfield County, CT.
-@tbl-sample-institutional-profile shows that 58 in-state students submitted SAT scores to the institution.
-These 58 students represented 4.8% of the total 1,199 in-state students from Fairfield County.
-69 regional students sent scores to the institution, representing 4.1% of all 1,664 regional test-takers.
-Of these 69 regional students, 35 expressed interest in majoring in the liberal arts.
-These 35 students represent 7.0% of all regional students from Fairfield County who expressed interest in the liberal arts.
+This manuscript builds on @RN4985 by conceptualizing market research as a mechanism of social reproduction in education.
+Market research is "the process of gathering, analyzing, and interpreting information about a market, including its consumers, competitors, and the overall industry" \[CHAT GPT, REPLACE\].
+Market research is oriented to the goals producers/sellers and is the foundation for marketing efforts and products that target desired prospects.
+This manuscript shows how College Board "geo-markets" emerged from market research by @RN4982.
+@RN4982 used SAT score-sending behavior from 19XX to develop the "market segment model," which shows how student demand for different kinds of postsecondary institutions varies across different geographies.
+Geo-markets carve large metropolitan areas into smaller "local" markets.
+For each geo-market, the market segment model categorizes students by the kind of postsecondary institutions they are likely to be interested in and calculates institution-specific measures of student demand.
+In turn, the market segment model is the basis for the College Board Enrollment Planning Service (EPS), which helps colleges plan and implement recruiting strategy.
 
-INSERT @tbl-sample-institutional-profile ABOUT HERE
+Empirically, we examine how socioeconomic and racial composition differs across geomarkets.
+We conceptualize the market segment model as a snapshot of social stratification in student demand for higher education.
+Additionally, because of segregation and structural racism in income/wealth, we expect local geomarkets to be correlated with race.
+Later, College Board geo-markets become a filter on the Student Search Service product that enables colleges to purchase the contact information of prospective students.
+Using simulations, we show how the how the racial and socioeconomic composition of student list purchases change as different geomarkets are selected.
+We also examine actual student list purchases that filter on geomarket.
 
-@RN4982 argue that the Institutional Profile and the Market Segment Profile enable admissions officers to know where to look for students and which institutions are competing for those students (p.
-25):
+This research makes a broad contribution to the sociology of education by conceptualizing market research as a mechanism of inequality.
+Market research helps producers identify and target potential customers by taking snapshots of existing social stratification in who are the desirable prospective customers and where do they live.
+These snapshots are a function of historic and contemporary structural inequalities.
+Market research uses these snapshots to make recommendations about which prospective customers to recruit and to create marketing products.
+These recommendations an products are then used by producers.
+Because the snapshots are reflections of historic structural inequality, using these snapshots to identify future customers is likely to amplify historic structural inequality.\[discussion section identifies areas of future research\]
 
-> The Institutional Profile and the Market Segment Profile quantify the admission officers' intuitive grasp of market structure.
-Structure here carries a dual meaning, connoting both the structure of student choice and the structure of institutional competition\ldots This two-sided interpretation furnishes the essential framework for planning by individual colleges and universities\ldots To draw effectively on its own natural constituency, a college not only must contact the "right" kind of students --- that is, students who are predisposed toward that type of institution -- but also must persuade them of its special character.
-This means knowing the competition as well as the clientele.
+NEXT SUBSECTION IS MARKET RESEARCH AS A MECHANISM OF INEQUALITY IN SOCIOLOGY WRIT LARGE
 
-**Understanding and using the market segment model**.
-Chapter 3 of @RN4982 -- "A Sense of Place: Students, Families, and Communities" -- identifies the student characteristics associated with being in the local, in-state, regional, or national market segment and examines the extent to which these relationships hold across states and local markets.
-The data analyzed by @RN4982 can be seen as training data.
-In actual recruiting applications, colleges would have historic data on the number of students in each segment in a local market, but they would not know a current high school student is a local, in-state, regional, or national prospect.
-Therefore, knowing which observable student characteristics are associated with score sending outcomes can be useful for knowing which prospects an institution should target.
+-   poon, FICO scores
+-   geodemography
+    -   from EEPA article: using geography as a predictive input was pioneered by geodemography, a branch of market research that estimates the behavior of consumers based on where they live (Burrows and Gane, 2006)
+-   fourcade and healey
+    -   classification situations
+-   key theme is third-party market research and related products that sorts customers for the benefit of merchants.
 
-Chapter 3 of @RN4982 is fundamentally about correlations.
-The authors identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that individually or in combination predict score-sending behavior.
-These four variables "reflect the basic social patterns of the nation"<!--.
-It would have been surprising if these were not the four social variables that best explained the patterns of college choice"--> [@RN4982, p.
-33].
-The text uses authoritative language to summarize the correlational analyses.[^6] @RN4982 state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" (p.
-33) and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations.
-Commenting on family income, @RN4982 write that"we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time" (p.
-33).
-Further analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
-In their analysis of quantifying school quality in England, @RN5005 [p.
-517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin \[CITE\], the Market Segment Model essentially argues that student demand for higher education is mostly a function of social origin.
+This study extends @RN4985 in two ways.
+First, critical scholarship on data science finds that geography is the most common racialized input, owing to centuries of residential segregation.
+College Board student list product enables colleges to filter based on known geographic units (state, zip code, MSA).
+But it also creates new academic units ...
+bla bla
 
-[^6]: For example, @RN4982, pp.
-34-35 state, "it is hard to overemphasize the importance of this statistical pattern.
-Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice."
+second, PARTX: Scholars have not analyzed the ways that market research is built into products that recruit students In other words, market research identifies student demand for different kinds of institutions This market research is then incorporated into recruiting projects We argue that this phenomenon may be an important mechanism by which historic racial and socioeconomic inequality in student demand is utilized to inform the subsequent recruiting behavior of universities.
 
-@RN4982 find that the SAT score, parental education, and family income also predict student score-sending behavior at the local geo-market level but geo-markets differ in the relative abundance of students with particular socioeconomic characteristics.
-These findings have practical implications for recruiting.
-@RN4982 recommend that colleges target geo-markets with desirable compositions of socioeconomic characteristics and student market segment.
-@RN4982 [p.
-44] describes a hypothetical college that wants to target regional and national students in New England and is considering whether to expend limited recruiting efforts in the Boston geo-market -- which was relatively low-income in 1980 -- or in more affluent geo-markets nearby:
+We Overall, scholarship from sociology tends to assume that recruiting -- and, in turn, structural inequality in recruiting -- is a function of individual college behavior.
 
-> Where would you concentrate your energies?
-Ideally, you would seek communities with a high proportion of students already predisposed toward institutions such as your own.
-The Market Segment Model would provide this information through segment percentages for the community in question.
-Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
-Your efforts would surely be better directed toward three of the four communities in the bottom band, Manchester, Hartford, and Fairfield County.
-Simply by knowing a little bit about the students' backgrounds and academic records you could quickly focus your attention on those most likely to consider your kind of institution.
-Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own."
+TIFFANY HUANG SAYS THIS IS ABOUT REPRODUCTION; MECHANISMS OF HOW REPRODUCTION OF INEQUALITY HAPPENS; THIS FEELS RIGHT; Tiffany and she helped me realize that the literature I am contributing to (broadly) is social reproduction.
+more specifically, how (third-party) supply-side actors take snapshots of existing social stratification and build those into the supply-side architecture.
 
-In 1984, College Board created the Enrollment Planning Service (EPS) based on the @RN4982 market segment model [@RN4987; @RN4980].
-Each institution purchasing EPS information services would obtain access to the Market Segment Report for each local market.
-They also received the Institutional Profile -- their own and that of competitors -- for each local market.
-Promotional material highlights the themes of "locate the students you want" [@RN5019_2003] and "focus your valuable time and resources on the right prospects" [@RN5019].
-@RN5019_2010 promotional material describes EPS as:
-
-> the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
-You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, Zip codes, high schools and international regions.
-EPS provides you with comprehensive reports on your markets, your position in those markets, and your copetition."
-
-```{=html}
-<!--
-@RN5019 [2005]
-> The marketing software that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
-With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
-Focus your valuable time and resources on the right prospects
-
-Software documentation by @RN5017 states that "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
--->
-```
-Whereas @RN4982 [Appendices X-Y] created 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S.
-states, with 304 geo-markets in total.
-@RN4988 shows the College Board geo-markets circa 2023.
-Interestingly, the geo-markets for New England, Middle States, and the South remain nearly identical to those developed by @RN4982.
-Concrete information about how EPS created the remaining geo-markets is elusive.
-Documentation and promotional material suggests that geo-market borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^7]
-
-[^7]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
-Students from California don't all share the same college-going behaviors.
-We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."
-
-Although the methods utilized to create geomarket borders remain unclear, we expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
-Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
-Additionally, @RN4982 created geomarkets in service of the Market Segment Model.
-The Market Segment Model argues that socioeconomic status drives whether students belong to the local, in-state, regional, or national market segment.
-In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status.
-The language of @RN4982 [pp.
-11-12] suggests that in large metropolitan areas, the borders of geomarkets may have been drawn in ways that separate affluent from less affluent communities: "each major metroplitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs." We present the following hypothesis:
-
-**H1**. In large metropolitan areas that contain multiple geomarkets, we expect significant socioeconomic inequality (e.g., income, parental education) between geomarkets.
-
-We also expect substantial racial inequality between geomarkets in large metropolitan areas.
-Interestingly, whereas the Market Segment Model is explicitly based on socioeconomic stratification, @RN4982 do not mention race once.
-Structural racism is "a form of systematic racial bias embedded in the 'normal' functions of laws and social relations" [@RN4760, p.
-1143], whereby processes viewed as neutral or common sense systematically advantage dominant groups and disadvantage marginalized groups.
-Because of the extent of residential racial segregation in the U.S.
-[@RN4779; @rothstein2017color], algorithms that utilize geographic borders as an input have been shown to reproduce racial inequality [e.g., @RN4775; @RN4849; @RN4779].
-Unless they intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
-
-Geomarket borders may be correlated with race for two additional reasons.
-First, geomarket borders may have been drawn along class divides and there is a strong correlation between race and wealth [@RN4884].
-Second, geo-market borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
-Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
-Although these arguments are speculative, the extent of racial inequality between geomarkets can be assessed empirically.
-We present the following hypothesis:
-
-**H2**. In large metropolitan areas that contain multiple geomarkets, we expect significant racial inequality (e.g., income, parental education) between geomarkets.
-
-**Reproducing Inequality**.
-Finding that geomarkets are correlated with class and race is neither surprising nor noteworthy.
-However, we argue that the Market Segment Model and geomarkets are mechanisms for reproducing class-based and race-based inequality that exemplify many themes of the broader quantification literature.
-<!--
-In both their usage within the EPS market research product and the Student Search Service student list product, 
--->
-
-the Market Segment Model and geo-markets exemplify concerns about the application of correlational analysis.
-The data on 1980 SAT score-sending behavior analyzed by @RN4982 can be conceived as training data.
-These training data were used to define geomarkets (e.g., how many geomarkets in a metropolitan area is partially a function of the number of test-takers) and to identify student characteristics associated with the local, in-state, regional, and national market segments.
-These analyses found that student demand for different kinds of colleges was largely a function of social origin, and that student demand within a geo-market was largely a function of the class composition of the geo-market.
-but did not interrogate the historical and contemporary structural inequalities that produced observed patterns of social stratification.
-Instead, the Market Segment Model encouraged colleges to use this snapshot of student demand as the basis for future decisions about which communities and prospective students should be targeted by college recruiting campaigns.
-This process exeemplifies the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849] whereby predicting the behavior of future individuals based on the behavior of past individuals "threatens to reify and reporduce existing inequalities" [@RN4794, p.
-224].
-This process also exemplifies concerns about proxy variables whereby selective colleges view privileged social class as a proxy for desirable students and this creates an incentive to focus recruiting efforts on communities that have these characteristics.
-@RN4982 [p.
-44] corroborates these concerns:
-
-> On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
-Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
-The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
-
-The utilization of the Market Segment Model as the basis for Enrollment Planning Service (EPS) raises the themes of reactivity, discipline, and authority introduced by @RN5007.
-Reactivity means that quantitative measures cause people and organizations to change behavior.
-Both @RN5003 (law school rankings) and @RN5005 (UK leage tables) study highly salient consumer-facing products.
-Although EPS targeted colleges rather than students, we suggest that EPS -- and geomarkets -- were highly salient within the enrollment management community.
-@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
-Discipline is the idea that quantitative measures compel people and organizations to react in a particular way.
-To the best of our knowledge, EPS originally provided colleges permutations of the Market Segment Model -- how many students in each market segment -- and the Institutional Profile -- which students in a market segment were interested in my college or a competitor.
-Both of these outputs were provided at the geomarket level.
-Promotional material from College Board -[@RN5019; @RN5019_2010] suggests that EPS encouraged colleges to first choose geo-markets of interest and then to plan recruiting efforts within these selected geo-markets.
-
-Usage of EPS also raises the theme of authority, whereby quantification weakens the authority of local decision-makers [@RN5007].
-The Market Segment model sought to develop a concrete, data-driven framework that replicated the aggregate wisdom of admissions officers [@RN4982].
-This model makes the local expertise of admissions officers less valuable.
-EPS transformed the model into a product stored on a CD-ROM.
-In turn, the EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
-Individual admissions officers are relegated to implementation of the plan rather than being valued for their local expertise about student demand.
-
-The incorporation of geo-markets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction.
-First, geo-markets are opaque.
-The borders of a geomarket may be unclear to someone purchasing student lists and very few people know how these borders were created \[COULD TALK ABOUT PRODUCT UTILIZATION HERE; NOT CLEAR HOW GEOMARKETS INTERACT WITH OTHER STUDENT LIST FILTERS\].
-Second, geo-markets scale.
-Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on geo-markets in concert with other filters (e.g., SAT score, GPA).
-Third, geo-markets can cause damage.
-When a college decides to purchase names from one geo-market but exclude prospects from an adjacent geo-market, this affects which college students have an opportunity to attend.
-@RN4982 suggests that selective colleges focus on geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-
-Chapter 4 of @RN4982 -- The Company We Keep: Colleges and Their Competition -- shows that the Market Segment Model is based on homophily.
-The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^8] @RN4982 [p.
-42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^9] Subsequent analyses investigate the tuition price and socioeconomic competition of institutions in competition with one another.
-Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-Rather than considering the structural inequality that gives rise to these patterns, @RN4982 describe a naturally occuring process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities.[^10] @RN4982 [p.
-72] state that
-
-[^8]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
-Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
-
-[^9]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
-54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
-
-[^10]: These patterns are similar to different offers for credit products in The Ordinal Society [@RN5015], except @RN5015 show that the matching of consumers to products is engineered by firms rather than naturally occurring.
-
-> We now know why students so seldom speak of their own social or family backgrounds in explaining how they go about choosing a college.
-They have no need to.
-Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
-The layering of collegiate competition is primarily a socioeconomic layering.
-The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
-Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
-
-Although this perspective is not entirely inconsistent with mid-20th Century scholarship on social mobility [@RN915; @RN1273], the Market Segment Model was the basis for the EPS market research product and, later, geomarkets were incorporated into the *Student Search Service* student list product.
-Drawing from @RN4975, commodification of the Market Segment Model "engineers homophily" [@RN4982] by taking a snapshot of existing stratification -- without considering the class-based or race-based structures that produced these patterns -- and then recommending that colleges utilize this snapshot as the basis for decisions about where to concentrate recruiting efforts.
-@RN4982 [p.
-58] states that "correlations, do not simply predict actions; they also form them." The Market Segment Model not only describes the structure of college choice, it influences the future structure of college choice by becoming the basis for practices and products.
-Once the Market Segment Model is inscribed in EPS and Student Search Service, these products are likely to amplify the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.
-
-UTILIZATION.
-Finally, within the Student Search Service product, we suggest that geomarket filters are likely to be used in ways that amplify class-based and race-based disparities in college opportunity.
-In contrast to standardized selection devices (e.g., the FICO score algorithm), student list products are discretionary selection devices in that colleges have discretion of which filters they elect to select prospects.
-Compared to standardized selection devices, scholars tend to find that administrative discretion over selection devices causes structural inequality to increase [@RN4883; @RN4881; @RN4795].
-Even in the absence of explicit or implicit bias, discretionary selection devices can increase structural inequality when decision-makers have incomplete knowledge about how the inputs they choose are correlated with race \[CAN MOVE O'NEIL TEXT ON OPACITY HERE\].
-For example, research shows that Americans dramatically underestimate the magnitude of racial income inequality [@RN4884].
-As such, discretionary selection devices are particularly sensitive to inputs that are correlated with race but are perceived to be race-neutral.
-
-Colorblind racism describes ideology and associated behaviors that ignore the significance of race and racial discrimination [@RN5020].
-The Market Segment Model can be seen as an example of colorblind racism.
-@RN4982 develop an analysis of the structure of college choice that ignores race but we contend that geo-markets are correlated with race.
-Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975].
-Enrollment management professionals utilizing EPS and *Student Search Service* may interpret geomarkets as a race-neutral recruiting tool.
-When purchasing student lists, @RN4982 [p.
-X] suggest that affluent colleges should consider targeting affluent geomarkets.
-Student list purchases do not show how the characteristics of targeted prospects compare to the characteristics of the surrounding community.
-From this perspective, racial inequality from the use of geomarkets in student list purchases may be the unintentional result of using a seemingly race-neutral product.
-Furthermore, purchases that specify multiple filters (e.g., test scores, grades, intended major) including geomarkets can yield unintended racial inequality because administrators have incomplete knowledge about how the intersection of these filters interact with local patterns of segregation.
-<!-- We suggests that geomarkets are a colorblind racist market device.
-Futhermore, geomarkets may interact with other student list filters (e.g., test score, GPA) in ways that exacerbate exclusion of under-represented student populations.
--->
-
-Below, we use simulations and actual student list purchases that filtered on geomarkets.
-In purchases that filtered on test-score and/or GPA thresholds, we can simluate who would be included and excluded had certain geo-markets were been selected.
-In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.
+But another mechanism of reproduction is market research and marketing.
+Market research takes a picture of "student demand" one that is a function of class and racial inequality and then uses that picture to create river beds of where colleges should allow recruiting to flow.
 
@@
 ```
-
-\clearpage
-
-<!-- TABLES -->
-
-```{r market-segment-characteristics, echo = F}
-#| label: tbl-market-segment-characteristics
-#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'
-
-market_segment <- data.frame(
-  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
-  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
-  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
-  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
-  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
-)
-
-kable(
-  x = market_segment,
-  col.names = c('Characteristic','Local','In-state','Regional','National'),
-  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```
-
-\clearpage
-
-```{r sample-institutional-profile, echo = F}
-#| label: tbl-sample-institutional-profile
-#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'
-
-sample_inst <- data.frame(
-  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
-  local = c(1,0.2,0,0.0,0,0.0),
-  in_state = c(58,4.8,25,8.4,2,3.0),
-  regional= c(69,4.1,35,7.0,5,5.8),
-  national = c(109,2.9,61,5.3,3,0.8),
-  total = c(237,3.3,121,5.9,10,1.8)
-)
-
-kable(
-  x = sample_inst,
-  col.names = c('','Local','In-state','Regional','National','Total'),
-  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```
-
-\clearpage
-
-```{r market-segment-institutions, echo = F}
-#| label: tbl-market-segment-institutions
-#| tbl-cap: 'Top 5 institutions in terms of number of scores sent by segment, Connecticut Market 3: Fairfield County'
-
-market_segment_inst <- data.frame(
-  rank = c(1,2,3,4,5),
-  local_num_scores_sent = c(277,261,183,103,100),
-  local_pct_scores_sent = c(50.4,47.5,33.3,18.7,18.2),
-  local_inst_type = c('priv ma','priv ma','priv ma','pub doct','pub ma'),
-  state_num_scores_sent = c(757,515,438,183,177),
-  state_pct_scores_sent = c(63.1,43.0,36.5,15.3,14.8),
-  state_inst_type = c('pub doct','pub ma','pub ma','pub ma','pub ma'),
-  regional_num_scores_sent = c(610,348,272,248,197),
-  regional_pct_scores_sent = c(36.7,20.9,16.3,14.9,11.8),
-  regional_inst_type = c('pub doct','priv doct','priv doct','pub doct','pub doct'),
-  national_num_scores_sent = c(1226,371,327,312,308),
-  national_pct_scores_sent = c(32.6,9.9,8.7,8.3,8.2),
-  national_inst_type = c('pub doc','priv doct','priv res','priv doct','priv doct')
-)
-
-kable(
-  x = market_segment_inst,
-  col.names = c('','Num','Pct', 'Type','Num','Pct','Type','Num','Pct','Type','Num','Pct','Type'),
-  #caption = 'Top 5 institutions in terms of number of scores sent, by segment \\\\Connecticut Market 3: Fairfield County'
-) %>% 
-  #kable_styling(full_width = F) %>% 
-  add_header_above(c(' '=1,'Local (N=550)' = 3, 'In-state (N=1,199)' = 3,'Regional (N=1,664)' = 3, 'National (N=3,766)' = 3))
-```

### 2024-07-08 text/old_spring2024/intro.qmd
initial commit, copying files from messed up repo

--- 2024-07-08 text/front_end.qmd
@@
 library(kableExtra)
-
-#hey ho This many words: {{< words-total >}}
-```
-
-```{=tex}
-\pagenumbering{roman}
-\newpage
-\pagenumbering{gobble}
-\pagenumbering{arabic}
 ```
@@
 
-In 1983, Zemsky and Oedel authored *The Structure of College Choice*, published by *The College Board.* This book develops the Market Segment Model and introduces local "community-based enrollment markets" (p.14) -- later called "geomarkets" -- which are the basis for the College Board market Enrollment Planning Service (EPS), a market research product.
-@RN4982[pp., 9-10] write that the Market Segment Model sought to quantify the knowledge that college admissions officers have about student demand:
+```{=html}
+<!-- CUT ORIGINS OF MARKET SEGMENT MODEL; TAKES TOO LONG TO GET TO THE NICE QUOTES, WHICH IS SUPPOSED TO BE THE VIGNETTE PART
+@RN4982 describes the origins of the Market Segment Model.
+In 1978, Robert Zemsky, a University of Pennsylvania professor, was asked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" (p.
+x) To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), an association of thirty-ish selective private universities.
+As the project project became more ambitious, the researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincientally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in tehse analyses was the state' [@RN4982, p.
+x].
+In 1979, College Board began providing data and funding for what became the Comprehensive Undergraduate Enrollment Planning Project (CUEPP).^[say that Pennsylvania Department of Education was also a partner; see page x of Zemsky] [MAYBE CUT SOME TEXT OF THIS PARAGRAPH]
+-->
+```
+In 1983, Zemsky and Oedel authored *The Structure of College Choice*, published by *The College Board.* This book develops the Market Segment Model and introduces local "community-based enrollment markets" (p.14) -- later called "geomarkets" -- which are the basis for the College Board Enrollment Planning Service (EPS), which provides market research to colleges about where to look for students.
+Whereas subsequent College Board products (e.g., Predictor Plus) can be read as the mimicry of corporate market logic, the Market Segment Model attempts to quantify the knowledge that college admissions officers have about student demand [@RN4982, pp.
+9-10]:
 
@@
 
-Scholarship in the sociology of education describes schools and colleges as "sorting machines" that reproduce inequality by incorporating, creating, and allocating students to different categories [@RN4873].
+[TRANSITION TO SCHOLARSHIP] @RN4873 describes schools and colleges as "sorting machines" that reproduce inequality by incorporating, creating, and allocating students to different categories.
 Given the consequences subsequent outcomes, the sociology of education is particularly concerned with the transition from high school to college.
@@
 The authors show that the utilization of particular search filters -- particularly geographic filters, such as zip code -- is associated with systematic racial exclusion that does not depend on which college purchased the list.
+However, prior research has not examined the role of third-party market research in reproducing socioeconomic and racial inequality
 
-However, prior research has not examined the role of third-party market research in reproducing socioeconomic and racial inequality.
 This manuscript analyzes the Market Segment Model and College Board Geomarkets as a case study of quantification.
-We develop arguments from the broader literature on quantification [e.g.,] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
+We develop arguments from the broader literature on how quantification reproduces inequality \[e.g.,\] [@RN5003; @RN5005].
+In particular, we draw from the discussions of correlation and homophily by @RN4975.
 Predictive analytics are based on correlation.
-When we use data on past correlations to make recommendations about the future, we amplify the effects of historic structural inequality.
+However, when we use data from past cases to make recommendations about future cases, we amplify the effects of historic structural inequality.
 The Market Segment Model analyzes SAT score-sending data from 1980 and concludes that student demand for higher education is primarily a function of social class= [@RN4982].
@@
 We address this question by spatially joining the geomarket shapefile to Census data about socioeconomic and racial characteristics.
-Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets, in concert with other commonly used search filters?
+Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets, in concert with other commonly used search filters.
 We address this question by analyzing student list purchases that do not filter geomarkets and simulating which prospects would have been included had they filtered on particular geomarkets.
@@
 
-# Background: Enrollment Management
-
-The term "enrollment management" (EM) refers to a profession, an administrative structure, and an industry.
-As a profession, EM integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p.
-xiv].
-As an administrative structure, the EM office typically controls the activities of admissions, financial aid, and recruiting [@RN2406].
-The EM industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers that produce market research, software, and and consulting services (e.g., College Board, ACT, EAB, Huron, Slate, PowerSchool).
-
-Figure @fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
-The funnel begins with a large pool of "prospects" (i.e., prospective students) that the university would like to enroll.
-"Leads" are prospects whose contact information has been obtained.
-"Inquiries" are prospects that contact the institution and consist of two types: first, inquiries who respond to an initial solicitation (e.g., email) from the university; and second, "student-as-first-contact" inquiries who reach out to the university on their own (e.g., sending ACT scores).
-<!--  Applicants consist of inquiries who apply plus "stealth applicants" who do not contact the university before applying.
--->The funnel narrows at each successive stage in order to convey the assumption of "melt" (e.g., a subset of "inquiries" will apply).
-Practically, the purpose of the enrollment funnel is to inform recruiting interventions that target one or more stages.
-These interventions seek to increase the probability of "conversion" across stages [@RN4322].
-At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz].
-Purchased leads are served emails, brochures, and targeted social media advertisements designed to solicit inquiries and applications [@RN4895].
-
-Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel they focus on.
-The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
-A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
-Scholarship on recruiting has conceptualized recruiting behavior as an indicator of college enrollment priorities [@RN4758] and identifies recruiting practices as a mechanism for social reproduction in college access [@RN4324; @RN3519].
-
-Ethnographies by @RN3519 and by @RN4407 analyze the connections between privileged colleges and privileged high schools.
-@RN4407 shows how guidance counselors at elite private school guidance counselors lobby admissions counselors on behalf of marginal students.
-@RN3519 finds that college admissions officers visit high schools as a means of maintaining positive relationships with guidance counselors at affluent feeder schools.
-Consistent with this finding, quantitative case studies of off-campus recruiting visits show that selective private colleges visit affluent, predominantly white schools and communities, with a disproportionate number of visits to private high schools [@RN5014; @RN4330].
-Additionally, public research universities often make make more out-of-state recruiting visits than in-state visits, and these out-of-state visits focus on the same set sorts of schools targeted by selective private universities [@RN4759; @RN4758].
-
-Other studies investigate recruiting behavior by non-selective institutions [@RN4520; @cottom2017lower; @dache2018dangling].
-Analyzing recruiting from the perspective of high school students, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
-@cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.
-<!--  WHAT YOU SAID ABOUT POSECZNICK Posecznick (2017) studied the “Ravenwood College,” a private nonprofit college founded to serve working Black women.
-Consistent with analyses of tuition discounting by Lapovsky (2001), the survival of Ravenwood depends on the success of admissions officers in identifying prospects (e.g., through on- and off-campus recruiting events) and “converting” these prospects to enrolled students.
--->
-
-We make two observations about the recruiting literature.
-First, consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015] and industrial organization [@RN1549; @RN2133], the industry is characterized by a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches, a process that exemplifies the idea that higher education is mechanism of social reproduction [@RN2377; @RN4873].<!--  For example, for-profit institutions target adults in communities not being targeted by traditional colleges [@cottom2017lower].
-Non-selective private non-profits target prospects who might be considering community college or a local state colleges [@RN4324].
-Selective private colleges target private high schools, with a particular link between sectarian high schools and colleges of the same faith [@RN5014], but these schools are increasingly targeted by out-of-state public research universities, suggesting growing competition for a coveted market niche.
---> Second, by focusing on the behaviors of individual colleges, the literature implicitly states that recruiting is something done by individual colleges.
-<!--  
-Why then, does recruiting behavior look so similar across colleges of a particular "type"?[^1]
-[^1]: New institutional theory offers one explanation [@RN527].
-In highly developed organizational fields like higher education, forces of competitive and institutional isomorphism -- including institutional locics -- compel organizations adopt similar practices as their peers.
-Practices deemed advantageous spread quickly, for example, public research universities targeting nonresident prospects to compensate for declines in state appropriations.
-These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.
- -->
-
-<!-- make or buy  
-The process of recruiting students involves many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor.
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
--->
-
-However, the recruiting behavior of colleges is structured by third-party actors.
-Drawing from scholarship on organizational theory [@RN4738], enrollment management processes involve many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-<!-- As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
---> The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-
-Moreover, process of college access is structured by third-party recruiting products utilized by colleges.
-@RN5012 deconstruct the College Board *Student Search Service* product.
-Selection devices are procedures or routines for making categorical decisions based on input values [@RN4778].
-Student list products are selection devices that enable colleges to choose which prospective students they target or ignore based on search filters that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
-"Racialized inputs" are ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
-@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) are racialized inputs.
-Using a national sample of high school students, and also using data from actual student lists purchased by public universities, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
- <!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
--->
-
-Student list products offer many geographic filters.
-Geographic borders are perhaps the most commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887], owing to extreme racial segregation in the U.S.
-In addition to offering geographic filters based on known geographic borders (e.g., zip code, county, CBSA, state), the College Board created new geographic borders that subsequently became filters in the *Student Search Service*.
-The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
-Geodemographic search filters can be read as the mimicry of corporate market logic [@RN4004] in that they apply the logic and methods of commercial geodemographic market segmentation products, such as PRIZM by the Claritas Corporation [@RN4957], to the case of college recruiting.
-
-Geomarket search filters, the focus of this manuscript, slice states and metropolitan areas into smaller, local recruiting markets.
-For example, geomarket "CA08" refers to "Alameda Country excluding Oakland." A university might purchase prospect contact information by filtering on this geographic region in combination with additional filters, such as year of high school graduation and SAT test score.
-When we initiated this manuscript, we asked ourselves, "where did geomarket filters come from?" The answer surprised us.
-We had imagined ourselves studying the application of the "data science revolution" to college recruiting, but it turned out we were studying the 1970s and 1980s!
-The Market Segment Model and geomarkets can be read as the application of social science research methods to "capture and quantify the phenomenon behind the folklore" [@RN4982, p.
-11] of college admissions officers.
-This work, comissioned by College Board, was subsequently turned into a market research tool -- The College Board Enrollment Planning Service -- that helps colleges decide where to look for prospective students.
-
-```{=html}
-<!--The enrollment management industry emphasizes market research.
-Market research is the process of producing information that informs decisions about product design or how to find customers.   -->
-```
-
-The enrollment management industry puts great stock in market research.
-@RN4986[p.
-17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].^[@RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
-In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited.
-In the chapter entitled "Travel," @RN3519 [p.
-69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."] However, most market research in the contemporary enrollment management industry is produced by third-party organizations.
-For example, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development [e.g., @RN4741; @RN4728, but most only available to clients.
-EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
-
-Despite being central to enrollment management for at least 40 years [@RN4986], the role of third-party market research has been ignored by scholarship on college access.
-Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935;@RN1814].
-However, market research informs where colleges look for students and which prospective students to target.
-These recruiting interventions intervene on student demand [@RN4324; @cottom2017lower], thereby structuring college access pathways in ways that could amplify or diminish historic inequality in college access.
-The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals [@RN4986].
-This article offers one story about the quiet revolution that transformed college access into a sophisticated, supply-led enterprise.
-We argue that market research structures who colleges recruit, in ways that reinforce existing social stratification.
-The case of Market Segment Model and College Board geomarkets shows how market research -- and its commodification into recruiting products -- is a mechanism for social reproduction.
-[LAST TWO SENTENCES ARE REDUNDANT BUT BUT ARE TRYING TO DO SLIGHTLY DIFFERENT THINGS AND I LIKE THEM BOTH]
-
-# Quantification
-
-```{=html}
-<!--
-CUT GENERAL PARAGRAPH ON REPRODUCTION
-
-Within the sociology of education, @RN2377 describes education as a sieve, allocating people to opportunities in socially stratified, capitalist societies.
-This role highlights "Max Weber's core insight that education has a dual character" [@RN2377, p.
-129], on one hand, serving social mobility by redistributing access to privileged positions and, on the other, serving social reproduction by granting privileged families disproportionate access to the educational institutions that confer pathways to privileged positions.
-Similarly, @RN4873 describe schools and colleges as "sorting machines" that incorporate external categories (e.g., race, class), create categories (e.g., instructional tracks), and allocate students to categories.
-Initial category allocation affects subsequent allocation, such that school assignments and instructional track assignments affect college access, which affects employment assignments.
-"In evaluating schools as inequality regimes" @RN4873 \[p.
-322\] argue, "the relevant question is not whether schools produce inequality; rather, the relevant questions are, when and where do schools produce inequality, \[and\] how do they do so." As @RN2377 show, whereas early research on social stratification sought to identify patterns of inequality [e.g., @RN4873; @RN290; @RN1268], the sociology of education is now more concerned with identifying the mechanisms underlying these patterns.
--->
-```
-We situate our study amidst scholarship on within scholarship on quantification as a mechanism that reproduces inequality.
-@RN5007 [p.
-402] define quantification as "the production and communication of numbers" (p.
-402) and introduce reacivity, discipline, and authority as three interrelated themes that describe the effects of quantification.
-Reactivity is the idea that salient quantitative measures cause people and organizations to change their behavior.
-Discipline is the idea that quantification causes actors to react in a particular way.
-Authority is the idea that quantification changes decision-making power, often weakening the discretion of local decision-makers.
-
-Scholarship about U.S.
-News & World Report (USNWR) Law School rankings demonstrate the reactivity, discipline, and authority effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
-Developed as a means of informing prospective students about quality, rankings had the intended effect of affecting.
-Rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
-Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p.
-415].
-Rankings *disciplined* the behavior of law schools.
-For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
-Rankings also weakened the *authority* of admissions directors and admissions officers to make independent decisions about which applicants to admit, ideal class size, and curricular offerings [@RN5003].
-<!--
-In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.    -->
-
-The analysis of UK League Tables by @RN5005 show how quantification can be a mechanism of social reproduction.
-As part of a broader *New Public Management* effort to infuse market competition within public providers [@RN161], the UK government began publishing school "league tables" in 1992.
-League tables ranked schools based on student performance on high-stakes national exams taken at age 16.
-Using Census data on occupational class measured at the local authority level (similar to a U.S.
-county), @RN5005 find that localities with higher performing schools experienced growth in the share of managerial/professional households following the adoption of league tables and a decline in the share of working-class households.
-Additionally, using longitudinal household survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables.
-With respect to mechanisms, league tables facilitate making evaluative comparisons between schools regardless of geographic proximity.
-They also discipline households and schools to conceive of school quality in terms of test scores, which are substantially a function of the class composition of schools.
-Compared to working-class parents, parents with professional/managerial occupations are more aware of school league tables and have financial resources necessary to move to neighborhoods near a "good" school.
-Therefore, league tables contributed to social reproduction because managerial/professional households respond to consumer-facing metrics by moving to more expensive neighborhoods, near higher performing schools.
-<!-- THIS IS RATCHET EFFECT STUFF.
-I CUT IT BECAUSE YOU HAVEN'T INTRODUCED THE IDEA OF RATCHET EFFECT
-Schools that performed well on initial league tables tended to be schools with a high share of professional/managerial families.
-Therefore, quantification takes a snapshot of existing social stratification and that snapshot becomes a power unto itself, amplifying the effect of previous strctural inequality on subsequent structural inequality.
--->
-
-```{=html}
-<!-- THIS IS THE CUT TRANSITION PARAGRAPH
-In the broader quantification literature, critical data studies [e.g., @RN4975; @RN4849; @RN4772] and sociological analyses of technology and inequality [e.g., @RN4778; @RN4786; @RN4877; @RN4810] introduce concepts salient to the analysis of market research.
-Our focus is on "selection devices," which utilize some combination of inputs and discretion to categorize places, people, and organizations for the purpose of allocating opportunities or interventions [@RN4778; @RN5012].
--->
-```
-In the interdisciplinary literature on quantification [@RN5008], the discussions of correlation and homophily by @RN4975 introduce concepts that are salient to the analysis of market research.
-Chapter 1 of @RN4975 is *Correlating Eugenics*.
-Correlation measures the extent to which two or more variables move together.
-The Pearson Correlation Coefficient was developed by the 19th Century statisticians and "biometric eugenicists" Francis Galton and his student Karl Pearson [@RN4975].
-@RN4975 [p.
-50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared in his 2008 editorial 'The End of Theory,' big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'" Predictive analytics are developed in two steps.
-First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
-Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and to assign levels of risk to each case.
-These models predict outcomes based on correlations without requiring knowledge about underlying causal relationships.
-@RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
-
-Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated to a variable of interest.
-For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of these others.
-@RN4849 [p.
-146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
-
-Many studies show that using correlational models make predictions reproduces structural inequality.
-The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
-These correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
-@RN4794 [p.
-224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Therefore, disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies disproportionate targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].
-An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
-Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
-Note that predictive models such as the LSI-R model are not merely used for social science purposes.
-Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.
-
-Chapter 2 of @RN4975, *Homophily, or the Swarming of the Segregated Neighborhood*, argues that homophily often underlies the use of correlation to predict future behavior.
-Homophily means that actors who share common characteristics are likely to form connections with one another.
-The ubiquitous colloquialism is "birds of a feather flock together." Homophily is a core assumption for network science [@mcpherson2001birds], in which actors (e.g., high school students) are linked to one another through direct and indirect network ties (e.g., submit SAT scores to the same college).
-Because network science often draws from rational choice theory, it assumes that homophily is the result of voluntary action by individuals, thereby "eras\[ing\] historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
-95] that underlie behavior consistent with homophily.
-@RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.[^4]
-
-[^4]: Thus, @RN4975 [p.
-82] writes, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together."
-
-Within market research, homophily is central to geodemography and the creation of market segments.
-Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
-Market segments are subgroups within a larger market that have similar consumer demand.
-Early geodemographic classifications of consumers (e.g., PRIZM by Claritas Corporation) were derived from publicly available Census data, which disaggregated data to the zip code level [@RN4957].
-The Claritas Corporation argued -- and had a financial incentive to do so -- that people living near one another share similar consumer preferences.
-Therefore, geographic localities could be categorized into market segments that would be useful for direct mail marketing campaigns.
-Later, the development of individual credit scores (e.g., FICO score) enabled merchants to classify consumers into many, fine-grained groups [@RN4877].
-@RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].
-For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
-<!--
-ORIGINAL TEXT ABOUT GEODEMOGRAPHIC CLASSIFICATION SYSTEMS
-
-Clusters were given evocative names like "Blue Blood Estates," "\[Archie\] Bunker's Neighbors", and "Share Croppers." Because U.S.
-publicly available U.S.
-Census data disaggregated data only to the zip code-level, the Claritas Corporation had a strong incentive to argue that consumer behavior within zip codes was very similar [@RN4957].
-@RN4565 emulated the behavior of Claritas by developing the *Segment Analysis Service: An Educationally Relevant Geodemographic Tagging Service*, which was an "add-on" to the *Student Search Service* student list product.
-Describing its conceptual underpinnings, @RN4565 [p.
-1] states, "the basic tenet of geodemography is that people with similar cultural backgrounds, means, and perspectives naturally gravitate toward one another or form relatively homogeneous communities; in other words, birds of a feather flock together." The Segment product allocated each high school and each census tract into a distinct "educational neighborhoods" thought to reflect useful market segments for colleges looking for students.
-
-PREVIOUS TEXT ABOUT CLASSIFICATION SITUATIONS
-Related to the creation of market segments, @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Historically, classifications were binary.
-Consumers with "good" credit were offered loans and those with bad credit were not.
-The innovation of individual FICO scores -- and, later, other credit scores -- enabled merchants to classify consumers into many, fine-grained groups or along a continuum.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].
-People previously denied credit were now granted credit on different terms.
-For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-Thus, at the bottom of the continuum, contemporary classification situations produce "predatory inclusion" -- exemplified by payday loans and for-profit colleges -- which is the logic of "including marginalized consumer-citizens into ostensibly democratizing mobility schemes on extractive terms" [@RN4774, p.
-443].
-
-Negative consequences of homophily
-Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
--->
-
-Reviewing recent scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Law school rankings [e.g., @RN5003] and school league tables [e.g., @RN5005] seem to exemplify this sort of salience.
-Widely known studies of quantification in education tend to study consumer-facing quantification.
-To our knowledge, prior research on quantification in education has not examined how third-party market research classifies consumers.
-This manuscript analyzes "market segment model" [@RN4982], which categorized high school students into vertical market segments and simultaneously created local "geo-markets" that were evaluated based on their composition of student market segments.
-The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target.
-@RN5021 reports that in 1995, 37% of 4-year public universities and 49% of 4-year private universities used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
-Geo-markets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities.
-<!--
-CUT COMMENTS ON SALIENCE
-By contrast, many quantification projects -- for example, human rights indicators -- have "little oomph" [@RN5006, p.
-265].
-FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created."
---> Scholarship on predictive analytics [@RN4975; @RN4849] suggests the mechanisms by which market research reproduces structural inequality.
-Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
-Unlike studies like @RN5005 we cannot show the effect of quantification.
-However, this study provides inisight about the mechanism underlying the effect.
-By analyzing and simulating *Student Search Service* purchases that filter on geo-markets we can show *how* the geo-markets reproduce historical class-based and race-based inequality in college access.
-
-# The Market Segment Model and College Board Geomarkets
-
- <!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.-->
-As a member organization, several College Board products developed from joint collaborations with administrators or professors from member institutions.
-During the 1970s, under the leadership of Larry Litten, then director of institutional research, Carleton College pionered applying market research to student recruiting.
-In the late 1970s, College Board provided Carleton College with funding and data for a six-year project that analyzed college choice decisions by households in six metropolitan areas.
-This project was the basis for the landmark book *Applying Market Research in College Admissions* [@RN4986], published by College Board, and was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll.
-Litten would later become Associate Director of the Consortium of Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
-
-In 1978, Robert Zemsky, a University of Pennsylvania professor, was asked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
-x].
-To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE).
-As the project project became more ambitious, the researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in tehse analyses was the state' [@RN4982, p.
-x].
-In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.^[say that Pennsylvania Department of Education was also a partner; see page x of Zemsky] @RN4982[p.
-4] write that,
-
-> For our analysis, we sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
-The Market Segment model that we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
-The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways.
-
-**Creating the Market Segment Model.** Considering the @RN5007 \[p.
-402\] definition of quantification as "the production and communication of numbers," the Market Segment Model is an effort to quantify and communicate the knowledge that admissions officers have about student demand.
-@RN4982[p.
-4] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." They created  "three types of boundaries" [@RN4982, p.
-11], region, state, and community.
-The three regions were New England, Middle States, and the South.
-Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" \[in New England, Middle States, and the South\].
-
-These enrollment markets, later called "geomarkets" by the College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
-We know little about how geomarkets were created.
-@RN4982 write that,
-
-> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
-Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
-In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."
-
-Having defined these geographic boundaries, high school students were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
-Each institution a student sends SAT scores to can be defined as "local" (institution is in the same local market that the student lives in), "in-state" (same state but different local market), "regional" (same region but different state), or "national." In turn, a "local" student submits more SAT scores to local institutions than they do to in-state, regional, or national institutions.
-An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
-<!--
-A "regional" student submits more SAT scores to regional institutions than they do to local, in-state, or national institutions.
-A "national" student submits more SAT scores to national institutions than they do to local, in-state, or regional institutions.
--->
-
-The two primary outputs of the market segment model are the (1) Market Segment Profile and (2) the Institutional Profile.
-The Market Segment Profile, which was produced separately for each of the 143 geo-markets, provides information about the number and characteristics of students in each market segment.
-@tbl-market-segment-characteristics reproduces a partial, simplified version of the bottom panel of @RN4982 [Figure 2.1] which represents the Market Segment Profile for "Connecticut Market 3: Fairfield County." @tbl-market-segment-characteristics has separate columns for each market segment (local, in-state, regional, national) and rows show the number of test-takers and characteristics of test-takers.
-For example, there were 550 "local" students in Fairfield county and these students submitted SAT scores to 2.8 postsecondary institutions on average.
-By contrast, there were 1,664 "regional" students who submitted test scores to 4.8 institutions on average.
-For local students, 11.2% had family income greater than \$35,000 and 9% had both parents with a BA.
-For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA.
-Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.[^5]
-
-[^5]: @tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market.
-For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution.
-Of the 1,199 in-state students, 757 (63.1) sent scores to institution #1, a public doctoral granting institution, and 515 (43.0%) sent scores to institution #2, a public master's granting institution.
-
-INSERT @tbl-market-segment-characteristics ABOUT HERE
-
-The *Institutional Profile* describes students who send scores to a particular institution and which majors these students are interested in.
-For each institution, institutional profiles are created separately for students from a particular local market, for all students in a state, or all students in a region.
-@tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the institutional profile of an anonymous institution for students from Fairfield County, CT.
-@tbl-sample-institutional-profile shows that 58 in-state students submitted SAT scores to the institution.
-These 58 students represented 4.8% of the total 1,199 in-state students from Fairfield County.
-69 regional students sent scores to the institution, representing 4.1% of all 1,664 regional test-takers.
-Of these 69 regional students, 35 expressed interest in majoring in the liberal arts.
-These 35 students represent 7.0% of all regional students from Fairfield County who expressed interest in the liberal arts.
-
-INSERT @tbl-sample-institutional-profile ABOUT HERE
-
-@RN4982 argue that the Institutional Profile and the Market Segment Profile enable admissions officers to know where to look for students and which institutions are competing for those students (p.
-25):
-
-> The Institutional Profile and the Market Segment Profile quantify the admission officers' intuitive grasp of market structure.
-Structure here carries a dual meaning, connoting both the structure of student choice and the structure of institutional competition\ldots This two-sided interpretation furnishes the essential framework for planning by individual colleges and universities\ldots To draw effectively on its own natural constituency, a college not only must contact the "right" kind of students --- that is, students who are predisposed toward that type of institution -- but also must persuade them of its special character.
-This means knowing the competition as well as the clientele.
-
-**Correlations**.
-Chapter 3 of @RN4982 -- "A Sense of Place: Students, Families, and Communities" -- identifies the student characteristics associated with being in the local, in-state, regional, or national market segment and examines the extent to which these relationships hold across states and local markets.
-The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
-[@RN4982, p.
-33] state that these four variables "reflect the basic social patterns of the nation.
-It would have been surprising if these were not the four social variables that best explained the patterns of college choice."
-
-@RN4982 summarize their correlational analyses using authoritative language.
-@RN4982[pp.
-34-35] asserts, "it is hard to overemphasize the importance of this statistical pattern.
-Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice." With respect to family income, @RN4982[p.
-33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982[p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time." Further analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
-Considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model implicitly and explicitly states that student demand for higher education is mostly a function of social origin.
-@RN4982 also find that the SAT score, parental education, and family income predict student score-sending behavior at the local geo-market level.
-
-However, geomarkets differ in the relative abundance of students with particular socioeconomic characteristics and this has practical implications for recruiting.
-@RN4982 recommend that colleges target geo-markets with desirable compositions of socioeconomic characteristics and student market segment.
-@RN4982[p.
-44] describes a hypothetical college that wants to target regional and national students in New England and is considering whether to expend limited recruiting efforts in the Boston geo-market -- which was relatively low-income in 1980 -- or in more affluent geo-markets nearby:
-
-> Where would you concentrate your energies?
-Ideally, you would seek communities with a high proportion of students already predisposed toward institutions such as your own.
-The Market Segment Model would provide this information through segment percentages for the community in question.
-Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
-Your efforts would surely be better directed toward three of the four communities in the bottom band, Manchester, Hartford, and Fairfield County.
-Simply by knowing a little bit about the students' backgrounds and academic records you could quickly focus your attention on those most likely to consider your kind of institution.
-Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own."
-
-**EPS and College Board Geomarkets**.
-In 1984, College Board created the Enrollment Planning Service (EPS) based on the Market Segment Model [@RN4987; @RN4980].
-Each institution purchasing EPS information services would obtain access to the Market Segment Report for each local market.
-They also received the Institutional Profile -- their own and that of competitors -- for each local market.
-Promotional material highlights the themes of "locate the students you want" [@RN5019_2003] and "focus your valuable time and resources on the right prospects" [@RN5019].
-@RN5019_2010 promotional material describes EPS as:
-
-> the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
-You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, Zip codes, high schools and international regions.
-EPS provides you with comprehensive reports on your markets, your position in those markets, and your copetition."
-
-```{=html}
-<!--
-@RN5019 [2005]
-> The marketing software that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
-With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
-Focus your valuable time and resources on the right prospects
-
-Software documentation by @RN5017 states that "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
--->
-```
-Whereas @RN4982[Appendices B1-B3] identified 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S.
-states, with 304 geo-markets in total.
-@RN4988 shows the College Board geo-markets circa 2023.
-Interestingly, the geo-markets for New England, Middle States, and the South remain nearly identical to those developed by @RN4982.
-Concrete information about how EPS created the remaining geomarkets is elusive.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^7]
-
-[^7]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
-Students from California don't all share the same college-going behaviors.
-We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."
-
-Although how we lack clarity about how geomarkets were created, we expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
-Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
-Additionally, @RN4982 created geomarkets in service of the Market Segment Model.
-The Market Segment Model argues that socioeconomic status drives whether students belong to the local, in-state, regional, or national market segment.
-In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status.
-The language of @RN4982[pp.
-11-12, quoted above] suggests that geomarket borders within large metropolitan areas may have been drawn in ways that separate affluent from less affluent communities.
-We present the following hypothesis:
-
-**H1**. In large metropolitan areas that contain multiple geomarkets, we expect significant socioeconomic inequality (e.g., income, parental education) between geomarkets.
-
-We expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons.
-Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race once.
-However, U.S.
-cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779; @rothstein2017color].
-Structures built upon racialized structures are racialized structures.
-@RN4975 argues that unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity.
-Second, geomarket borders may have been drawn along class divides.
-There is a strong correlation between race and wealth [@RN4884].
-Third, geo-market borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
-Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
-Although these arguments are speculative, the extent of racial inequality between geomarkets can be assessed empirically.
-We present the following hypothesis:
-
-**H2**. In large metropolitan areas that contain multiple geomarkets, we expect significant racial inequality (e.g., income, parental education) between geomarkets.
-
-We argue that the Market Segment Model, geomarkets, and their commodification within EPS exemplify concerns about the application of correlational analysis.
-The data on 1980 SAT score-sending behavior analyzed by @RN4982 can be conceived as training data.
-These training data were used to define geomarkets (e.g., how many geomarkets in a metropolitan area is partially a function of the number of test-takers) and to identify student characteristics associated with the local, in-state, regional, and national market segments.
-These analyses found that student demand for different kinds of colleges was largely a function of social origin, and that student demand within a geo-market was largely a function of the class composition of the geo-market.
-However, @RN4982 did not interrogate the historical and contemporary structural inequalities that produced observed patterns of social stratification.
-Instead, the Market Segment Model encouraged colleges to use this snapshot of student demand as the basis for future decisions about which communities and prospective students should be targeted by college recruiting campaigns.
-This process exeemplifies the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849] whereby predicting the behavior of future individuals based on the behavior of past individuals "threatens to reify and reporduce existing inequalities" [@RN4794, p.
-224].
-This process also exemplifies concerns about proxy variables whereby selective colleges view privileged social class as a proxy for desirable students and this creates an incentive to focus recruiting efforts on communities that have these characteristics.
-@RN4982 [p.
-44] corroborates these concerns:
-
-> On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
-Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
-The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
-
-**Homophily and Utilization**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] shows that homophily is a core tenet of the Market Segment Model.
-The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^8] @RN4982[p.
-42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^9] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
-Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-@RN4982[p.
-72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
-
-[^8]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
-Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
-
-[^9]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
-54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
-
-<!--[^10]: These patterns are similar to different offers for credit products in The Ordinal Society [@RN5015], except @RN5015 show that the matching of consumers to products is engineered by firms rather than naturally occurring.-->
-
-> We now know why students so seldom speak of their own social or family backgrounds in explaining how they go about choosing a college.
-They have no need to.
-Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
-The layering of collegiate competition is primarily a socioeconomic layering.
-The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
-Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
-
-Although this perspective is broadly consistent with mid-20th Century scholarship on social mobility [@RN915; @RN1273], the Market Segment Model was the basis for the EPS market research product.
-Later, geomarkets were incorporated into the *Student Search Service* student list product.
-Commodification of the Market Segment Model "engineers homophily" [@RN4975] by taking a snapshot of existing stratification and then recommending that colleges utilize this snapshot as the basis for decisions about where to concentrate recruiting efforts.
-@RN4982 [p.
-58] states that "correlations, do not simply predict actions; they also form them." In itself, the Market Segment Model is simply a stylized depiction [HIRSCHMAN 2016 STYLIZED FACTS] of the structure of college choice, one that does not consider class-based and race-based structures that produced these patterns.
-Once the College Board inscribed the Market Segment Model into the EPS and Student Search Service Products, these products are likely to amplify the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.
-
-Utilization of the Market Segment Model and geomarkets within the Enrollment Planning Service (EPS) raises concerns related to the quantification themes of reactivity, discipline, and authority introduced [@RN5007].
-@RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Although EPS is a business-facing rather than a consumer-facing product, it has been a highly salient one.
-@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
-We suggest that EPS disciplines colleges to approach recruiting in a manner consistent with the Market Segment Model.
-Colleges are encouraged to, first, choose geomarkets with a high number of desirable prospects and, second, plan recruiting efforts within selected geomarkets.
-Under the principle of homophily, selective colleges should target geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-Consistent with this perspective, EPS software documentation from @RN5017 describes how geomarkets -- here, called "EPS market codes" -- can be utlized:
-
-> EPS\textsuperscript{\texttrademark} (Enrollment Planning Service) is a geographic and demographic data service offered annually by the College Board to Colleges and Universities.
-EPS provides information to subscribing institutions about competitors, feeder schools, and demographic strengths and weaknesses.
-EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful.
-
-EPS may also weaken the authority of local decision-makers.
-The Market Segment model sought to develop a concrete, data-driven framework that replicated the aggregate knowledge of local admissions officers [@RN4982].
-Once this knowledge is quantified and commodified into a product stored on a CD-ROM, the local expertise of admissions officers becomes less valuable.
-The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
-In turn, admissions officers are relegated to a foot-soldier implementation role rather than being sought out for their local expertise about student demand.
-
-Finally, we suggest that utilization of geomarket filters within the Student Search Service product is likely to amplify class-based and race-based disparities in college opportunity.
-In contrast to standardized selection devices (e.g., the FICO score algorithm), student list products are discretionary selection devices in that colleges have discretion of which filters they elect to select prospects.
-Compared to standardized selection devices, scholars tend to find that administrative discretion over selection devices causes structural inequality to increase [@RN4883; @RN4881; @RN4795].
-Even in the absence of explicit or implicit bias, discretionary selection devices can increase racial inequality when decision-makers have incomplete knowledge about how the inputs they choose are correlated with race.
-For example, research shows that Americans dramatically underestimate the magnitude of racial income inequality [@RN4884].
-As such, discretionary selection devices are particularly sensitive to inputs that are correlated with race but are perceived to be race-neutral.
-
-Geomarkets are opaque in that geomarket borders may be unclear to someone purchasing student lists and very few people know how these borders were created.
-@RN4849 argues that opacity in algorithmic products is intentional and is a core criterion of "weapons of math destruction." Enrollment management professionals utilizing EPS and *Student Search Service* may interpret geomarkets as a race-neutral recruiting tool.
-When purchasing student lists, @RN4982 [p.
-X] suggest that affluent colleges should consider targeting affluent geomarkets.
-Student list purchases do not show how the characteristics of targeted prospects compare to the characteristics of the surrounding community.
-From this perspective, racial inequality from the use of geomarkets in student list purchases may be the unintentional result of using a seemingly race-neutral product.
-Furthermore, purchases that specify multiple filters (e.g., test scores, grades, intended major) including geomarkets can yield unintended racial inequality because administrators have incomplete knowledge about how the intersection of these filters interact with local patterns of segregation.
-
-<!-- CUT TEXT DEFINING COLORBLIND RACISM AND RACE-NEUTRAL RACISM
-Colorblind racism describes ideology and associated behaviors that ignore the significance of race and racial discrimination [@RN5020].
-The Market Segment Model can be seen as an example of colorblind racism.
-@RN4982 develop an analysis of the structure of college choice that ignores race but we contend that geo-markets are correlated with race.
-Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975].
--->
-
- <!-- CUT PARAGRAPH: WHY GEOMARKETS ARE WEAPONS OF MATH DESTRUCTION
-The incorporation of geo-markets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction.
-First, geo-markets are opaque.
-The borders of a geomarket may be unclear to someone purchasing student lists and very few people know how these borders were created \[COULD TALK ABOUT PRODUCT UTILIZATION HERE; NOT CLEAR HOW GEOMARKETS INTERACT WITH OTHER STUDENT LIST FILTERS\].
-Second, geo-markets scale.
-Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on geo-markets in concert with other filters (e.g., SAT score, GPA).
-Third, geo-markets can cause damage.
-When a college decides to purchase names from one geo-market but exclude prospects from an adjacent geo-market, this affects which college students have an opportunity to attend.
-@RN4982 suggests that selective colleges focus on geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
--->
-
-# Methods
-
-Below, we use simulations and actual student list purchases that filtered on geomarkets.
-In purchases that filtered on test-score and/or GPA thresholds, we can simluate who would be included and excluded had certain geo-markets were been selected.
-In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.
-
-# Discussion
-
-In their analysis of quantifying school quality in England, @RN5005 [p.
-517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model argues that student demand for higher education is mostly a function of social origin.
-
 # References
@@
 :::
-
-\clearpage
-
-<!-- TABLES -->
-
-```{r fig-em-funnel, echo = FALSE, fig.align = 'center', fig.cap = "The Enrollment Funnel", out.width = "45%"}
-knitr::include_graphics('./../assets/images/enroll_funnelv2.png')
-#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
-#![The enrollment funnel](assets/images/enrollment_funnel.png)
-```
-
-\clearpage
-
-<!-- TABLES -->
-
-```{r market-segment-characteristics, echo = F}
-#| label: tbl-market-segment-characteristics
-#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'
-
-market_segment <- data.frame(
-  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
-  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
-  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
-  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
-  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
-)
-
-kable(
-  x = market_segment,
-  col.names = c('Characteristic','Local','In-state','Regional','National'),
-  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```
-
-\clearpage
-
-```{r sample-institutional-profile, echo = F}
-#| label: tbl-sample-institutional-profile
-#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'
-
-sample_inst <- data.frame(
-  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
-  local = c(1,0.2,0,0.0,0,0.0),
-  in_state = c(58,4.8,25,8.4,2,3.0),
-  regional= c(69,4.1,35,7.0,5,5.8),
-  national = c(109,2.9,61,5.3,3,0.8),
-  total = c(237,3.3,121,5.9,10,1.8)
-)
-
-kable(
-  x = sample_inst,
-  col.names = c('','Local','In-state','Regional','National','Total'),
-  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```
-
-\clearpage
-
-```{r market-segment-institutions, echo = F}
-#| label: tbl-market-segment-institutions
-#| tbl-cap: 'Top 5 institutions in terms of number of scores sent by segment, Connecticut Market 3: Fairfield County'
-
-market_segment_inst <- data.frame(
-  rank = c(1,2,3,4,5),
-  local_num_scores_sent = c(277,261,183,103,100),
-  local_pct_scores_sent = c(50.4,47.5,33.3,18.7,18.2),
-  local_inst_type = c('priv ma','priv ma','priv ma','pub doct','pub ma'),
-  state_num_scores_sent = c(757,515,438,183,177),
-  state_pct_scores_sent = c(63.1,43.0,36.5,15.3,14.8),
-  state_inst_type = c('pub doct','pub ma','pub ma','pub ma','pub ma'),
-  regional_num_scores_sent = c(610,348,272,248,197),
-  regional_pct_scores_sent = c(36.7,20.9,16.3,14.9,11.8),
-  regional_inst_type = c('pub doct','priv doct','priv doct','pub doct','pub doct'),
-  national_num_scores_sent = c(1226,371,327,312,308),
-  national_pct_scores_sent = c(32.6,9.9,8.7,8.3,8.2),
-  national_inst_type = c('pub doc','priv doct','priv res','priv doct','priv doct')
-)
-
-kable(
-  x = market_segment_inst,
-  col.names = c('','Num','Pct', 'Type','Num','Pct','Type','Num','Pct','Type','Num','Pct','Type'),
-  #caption = 'Top 5 institutions in terms of number of scores sent, by segment \\\\Connecticut Market 3: Fairfield County'
-) %>% 
-  #kable_styling(full_width = F) %>% 
-  add_header_above(c(' '=1,'Local (N=550)' = 3, 'In-state (N=1,199)' = 3,'Regional (N=1,664)' = 3, 'National (N=3,766)' = 3))
-```

### 2024-07-08 text/template.qmd
initial commit, copying files from messed up repo

---
title: "Some title"
author: "Some author"
date: last-modified
format: 
  html: 
    extensions: [andrewheiss/wordcount]
references:
- id: Lovelace1842
  author:
    - family: Lovelace
      given: Ada Augusta
  citation-key: Lovelace1842
  container-title: Taylor's Scientific Memoirs
  issued:
    - year: 1842
  language: en-GB
  page: 666–731
  title: >-
    Sketch of the analytical engine invented by Charles Babbage, by LF Menabrea,
    officer of the military engineers, with notes upon the memoir by the
    translator
  type: article-journal
  volume: 3

- id: Turing1936
  author:
    - family: Turing
      given: Alan Mathison
  citation-key: Turing1936
  container-title: Journal of Math
  issue: 345–363
  issued:
    - year: 1936
  page: 230–265
  title: On Computable Numbers, with an Application to the Entscheidungsproblem
  type: article-journal
  volume: 58
---

## Introduction

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua [@Lovelace1842]. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum [@Turing1936].

| Column A | Column B |
|----------|----------|
| 7        | Apple    |
| 19       | Banana   |

: Here's a table with a caption that doesn't get counted

::: {.no-count}

### No counting

You can use the `{.no-count}` class on a [fenced div](https://quarto.org/docs/authoring/markdown-basics.html#sec-divs-and-spans) to skip counting its contents, like this section.

:::

### Shortcodes

Word counts can be directly included in your document with shortcodes, like so:[^1]

- Use `{{{< words-total >}}}` to include a count of all words: **{{< words-total >}} words**

- Use `{{{< words-body >}}}` to include a count of the words in the text body only, omitting the references, notes, and appendix: **{{< words-body >}} words**

- Use `{{{< words-ref >}}}` to include a count of the words in the reference section: **{{< words-ref >}} words**

- Use `{{{< words-append >}}}` to include a count of the words in the appendix, which must be wrapped in a div with the `#appendix-count` id (see below for an example): **{{< words-append >}} words**

- Use `{{{< words-note >}}}` to include a count of the words in the notes: **{{< words-note >}} words**. You can also reliably count words in the footnotes which would have previously been included in whichever section they were.[^2]

- Use `{{{< words-sum ARG >}}}` where `ARG` is some concatenation of the four countable areas: `body`, `ref`, `append`, and `note`. For example, `{{{< words-sum body-note >}}}` includes a count of the words in the body and notes (**{{< words-sum body-note >}} words**); `{{{< words-sum ref-append >}}}` includes a count of the words in the references and appendix (**{{< words-sum ref-append >}} words**)

[^1]: **NOTE** shortcodes `{{{< words-body >}}}`, `{{{< words-ref >}}}`, `{{{< words-append >}}}`, and `{{{< words-note >}}}` do not overlap their counts.

[^2]: Count footnote words with `{{{< words-note >}}}`: {{< words-note >}}

### Minor issues with shortcode counting

This filter may not always produce accurate counts for documents that use shortcodes, specifically shortcodes that run after this filter. If a shortcode has multiple arguments but only returns one word, then it is possible that the word count is inflated,[^3] and conversely deflated if it returns more words.

[^3]: An example in this extension is `{{{< words-sum ARG >}}}`, where `ARG` is formatted as "ref-note" or "body-append", etc., to count specific sections together. Since `words-sum` has two words to construct the shortcode but only one word is returned, the region that uses this shortcode is inflated by one.

## Conclusion

Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.

## References [^4]

[^4]: Count reference words with `{{{< words-ref >}}}`: {{< words-ref >}}

::: {#refs}
:::

::: {#appendix-count}
## Appendix [^5] {.appendix}

Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
:::

[^5]: Count appendix words with `{{{< words-append >}}}`: {{< words-append >}}

### 2024-07-08 data/eps_market/build-eps-market-map.R

@@ -1,0 +1,157 @@
+### http://gis.stackexchange.com/questions/37370/merge-zip-codes-to-create-new-sales-markets
+
+
+## setup
+library(maptools)
+library(rgeos)
+library(gpclib) 
+gpclibPermit() ## needed to do union spatial polygons
+library(RODBC)
+library(stringr)
+
+## read in the zip file
+setwd("C:/Users/btibert/Documents/Dropbox/projects/eps-market-maps")
+myzip <- readShapePoly("US_Zip_-_Lower_48.shp")
+
+## read in the eps markets
+ch <- odbcConnectExcel("eps.xls")
+eps <- sqlFetch(ch, "Sheet 1", stringsAsFactors=F)
+close(ch)
+eps <- eps[,1:2]
+eps$zip <- as.character(eps$zip)
+eps <- subset(eps, !is.na(zip))
+eps$zip <- str_pad(eps$zip, 5, "left", "0")
+
+
+###############################################################################
+## make the EPS map
+## http://gis.stackexchange.com/questions/37503/rename-a-spatialpolygon-class-object-in-r
+## http://r-sig-geo.2731867.n2.nabble.com/merging-data-with-SpatialPolygonsDataFrame-td2764348.html
+############################################################################### 
+
+## merge on the data to the map, need to use match
+myzip@data=data.frame(myzip@data, eps[match(as.character(myzip@data$ZIP), eps$zip),])
+head(myzip@data); tail(myzip@data);
+
+## dissolve based on the merged eps region -- fixed eps file
+## some zips still are missing data, even after running epsGuess below
+eps.markets <- gUnionCascaded(myzip, id=myzip@data$eps)
+sdf <- data.frame(eps=row.names(eps.markets))
+row.names(sdf) <- row.names(eps.markets)
+eps.markets <-  SpatialPolygonsDataFrame(eps.markets, sdf)
+
+
+## TODO:  look into the issues below
+## using the merge above, these markets are still not included
+## US 7, GA03, FL06, HI 1, HI 2, OR04, AK 1, AK 2, TX11
+## AK and HI make sense, but a few bad codes and not sure about TX
+
+## plot the map, there are a few gaps to fill but not bad overall
+plot(eps.markets, col="red")
+
+## which zips are missing markets
+eps.missing <- subset(myzip@data, is.na(eps))
+
+
[…]

### 2024-07-08 scripts/chicken_scratch/build-eps-market-map_brock.R

@@ -1,0 +1,173 @@
+### http://gis.stackexchange.com/questions/37370/merge-zip-codes-to-create-new-sales-markets
+
+
+## setup
+library(maptools)
+library(rgeos)
+#library(gpclib) 
+#gpclibPermit() ## needed to do union spatial polygons
+library(RODBC)
+library(stringr)
+library(readxl)
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # 
+  list.files(path = data_dir)
+  
+eps_data_dir <- file.path(data_dir,'eps_market') # 
+  list.files(path = eps_data_dir)
+  
+scripts_dir <- file.path('.','scripts') # 
+  list.files(path = scripts_dir)
+
+
+## read in the zip file
+#setwd("C:/Users/btibert/Documents/Dropbox/projects/eps-market-maps")
+myzip <- readShapePoly(file.path(data_dir,'eps_market',"US_Zip_-_Lower_48.shp"))
+
+## read in the eps markets
+#ch <- odbcConnectExcel(file.path(eps_data_dir,'eps.xls'))
+#eps <- sqlFetch(ch, "Sheet 1", stringsAsFactors=F)
+#close(ch)
+#eps <- eps[,1:2]
+#eps$zip <- as.character(eps$zip)
+#eps <- subset(eps, !is.na(zip))
+#eps$zip <- str_pad(eps$zip, 5, "left", "0")
+eps <- read_excel(path = file.path(eps_data_dir,'eps.xls'), sheet = 'Sheet 1', col_names = TRUE)
+eps %>% str()
+
+###############################################################################
+## make the EPS map
+## http://gis.stackexchange.com/questions/37503/rename-a-spatialpolygon-class-object-in-r
+## http://r-sig-geo.2731867.n2.nabble.com/merging-data-with-SpatialPolygonsDataFrame-td2764348.html
+############################################################################### 
+
+## merge on the data to the map, need to use match
+myzip@data=data.frame(myzip@data, eps[match(as.character(myzip@data$ZIP), eps$zip),])
+head(myzip@data); tail(myzip@data);
+
+## dissolve based on the merged eps region -- fixed eps file
+## some zips still are missing data, even after running epsGuess below
+eps.markets <- gUnionCascaded(myzip, id=myzip@data$eps)
+sdf <- data.frame(eps=row.names(eps.markets))
[…]

### 2024-07-08 scripts/chicken_scratch/cb_geomarket_eda_oj.R

@@ -1,0 +1,221 @@
+################################################################################
+## [ PROJ ] < student list hsls >
+## [ FILE ] < make_data.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 5/8/2023
+## [ DESC ] < Create HSLS Dataset >
+################################################################################
+
+
+############## QUESTIONS
+  #
+  # should I be merging the geomarket geographies to zctas or tracts or zip codes?
+    # see: https://walker-data.com/census-r/mapping-census-data-with-r.html#understanding-and-working-with-zctas
+
+
+
+
+### SETTINGS
+rm(list = ls())
+options(max.print=1000)
+
+### LIBRARIES
+library(tidyverse)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+library(sf)
+library(tigris)
+#library(stars)
+#library(spatstat)
+
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # 
+  list.files(path = data_dir)
+  
+scripts_dir <- file.path('.','scripts') # 
+  list.files(path = scripts_dir)
+  
+### FUNCTIONS
+
+
+#### READ AND WRANGLE
+
+  file.path(data_dir,'eps_market','eps-map.Rdata')
+  load(file.path(data_dir,'eps_market','eps-map.Rdata'))
+
+  # read shape file
+  # https://cran.r-project.org/web/packages/sf/vignettes/sf2.html
+  file.path(data_dir,'eps_market',"US_Zip_-_Lower_48.shp")
+  myzip2 <- st_read(file.path(data_dir,'eps_market',"US_Zip_-_Lower_48.shp"))
+  
+#The Rdata file currently includes 3 objects. This will change as I finalize the map files.
+  # 1. eps.missing which is a data frame of zip codes that still need to be associated with an EPS territory
+  # 2. myzip which is a SpatialPolysDataFrame object. It is the map of the lower 48 by zip code. To plot, simply use the command plot(myzip) but note it will take a minute or so depending on your machine
+  # 3. eps.markets is the working draft of the eps markets map and is the same type as myzip  
+  
+  # eps.missing # data frame
+  
[…]

### 2024-07-08 scripts/chicken_scratch/create_geomarket_scratch.R

@@ -1,0 +1,543 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_geomarket.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 5/11/2023
+## [ DESC ] < Create geomarket shape files >
+################################################################################
+
+### SETTINGS
+  rm(list = ls())
+  options(max.print=1000)
+
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+library(sf)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # 
+  list.files(path = data_dir)
+  
+eps_data_dir <- file.path(data_dir,'eps_market') # 
+  list.files(path = eps_data_dir)
+  
+scripts_dir <- file.path('.','scripts') # 
+  list.files(path = scripts_dir)
+
+###########
+  
+  income_2015_zcta <- get_acs(
+    geography = "zcta",
+    variables = "B19013_001",
+    year = 2015,
+    geometry = TRUE,
+    survey = 'acs5'
+  )    
+  
+  income_2015_zcta %>% glimpse()
+  
+########### STEPS
+  # 1) Read in shape file that has polygons for every zip
+  # 2) read in .xls file that has name of geomarket associated with each zip
+  # 3) merge 
+  
+  # NEXT STEP, FIGURE OUT HOW TO DO THE UNION USING SF FUNCTIONS RATHER THAN RGEOS FUNCTIONS
+
+  # read eps.xls file which has geomarket for each zip code
+  eps_zip <- read_excel(path = file.path(eps_data_dir,'eps.xls'), sheet = 'Sheet 1', col_names = TRUE)
+  eps_zip %>% glimpse()
+  
+  # read zip code shape file
+  # https://cran.r-project.org/web/packages/sf/vignettes/sf2.html
+  zip_lower48 <- st_read(file.path(data_dir,'eps_market',"US_Zip_-_Lower_48.shp"))
+
+    # create sp version of object to compare w/ other script
+    #zip_sp <- as_Spatial(from = zip_df)
+    
+  # make sure zip codes are unique in each dataset
+    #zip_sp@data %>% group_by(ZIP) %>% summarise(n_per_group = n()) %>% ungroup %>% count(n_per_group) # one observation per zip code
[…]

### 2024-07-08 scripts/chicken_scratch/create_geomarket_scratch_take2.R

@@ -1,0 +1,640 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_geomarket.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 5/11/2023
+## [ DESC ] < Create geomarket shape files >
+################################################################################
+
+### SETTINGS
+  rm(list = ls())
+  options(max.print=1000)
+
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+library(sf)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # 
+  list.files(path = data_dir)
+  
+eps_data_dir <- file.path(data_dir,'eps_market') # 
+  list.files(path = eps_data_dir)
+  
+scripts_dir <- file.path('.','scripts') # 
+  list.files(path = scripts_dir)
+
+  
+########### STEPS
+  # Read in .xls file that has name of geomarket associated with each zip
+  # get ZCTA shape file from around 2010 from tidycensus function
+  # merge eps file to zcta file by zip code, to obtain dataset that has geomarket and shape-file for each zip-code
+  # create dataset that has one observation per geomarket by using aggregate to "dissolve" zip codes into geomarkets
+  # Using tidycensus, create dataset(s) with one observation per zip-code with data on desired characteristics (e.g., income, racial composition, population, education level, etc.)
+    # can use this dataset to do analyses of how characteristics vary within geomarket
+  # create dataset(s) with one obsr per geomarket by using group_by() and summarize()
+    # calculates descriptive statistics at geomarket level, to show how characteristics vary between geomarkets
+  # merge dataset w/ one obs per geomarket to dataset w/ geomarket shape file in order to create maps that graph geomarket characteristics
+      
+###########
+  
+# Read in .xls file that has name of geomarket associated with each zip
+  eps_zip <- read_excel(path = file.path(eps_data_dir,'eps.xls'), sheet = 'Sheet 1', col_names = TRUE)
+  eps_zip %>% glimpse()
[…]

### 2024-07-08 scripts/chicken_scratch/spencer2022_v2.R

@@ -1,0 +1,1132 @@
+# CLEAR MEMORY
+rm(list = ls())
+
+
+#LOAD LIBRARIES 
+    options(max.print=999999)
+    library(tidyverse) 
+    library(DataExplorer)
+    library(reshape2)
+    library(stringr)
+    #library(plyr) # consider getting rid of plyr; dplyr is replacement and is part of tidyverse
+    library(scales)
+    library(ggmap) #uses google maps API
+    library(maps)
+    library(mapdata)
+    library(ggplot2)
+    library(mapproj)
+    library(assertthat)
+    #library(zipcode)
+    library(rgeos)
+    library(rgdal)
+    library(maptools)
+    library(ggalt)
+    library(ggthemes)
+    library(ggrepel)
+    library(RColorBrewer)
+    library(leaflet)
+    library(leaflet.extras)
+    library(spdep)
+    library(formattable)
+    library(htmlwidgets)
+    library(sp)
+    library(tigris)
+               
+
+#############################################################################    
+#DATA
+
+#LOAD RECRUITING/ZIP/HS DATA, MERGE IN ZIP CODE DATA, CREATE EVENT TYPE CATEGORIZATION and ZIP VARS
+    
+  ##load in recruiting data
+    #setwd("~/Dropbox/recruiting-m1/analysis/scripts")
+    #all<-read.csv("../data/events_data_2.csv", colClasses=c(determined_zip="character"))
+    #all<-read.csv("data/events_data.csv", colClasses=c(determined_zip="character", event_state="character",  ipeds_id="factor"), na.strings=c("","NA"), encoding="UTF-8")
+    #all <- as.tibble(all)
+  ##fix state whitespace formatting
+    #all$event_state <- gsub('\\s+', '', all$event_state)
+    
+  #IMPORT REQ DATA   
+    all<-read.csv("data/events_data_2019-03-23_.csv", colClasses=c(determined_zip="character", event_state="character",  ipeds_id="factor"), na.strings=c("","NA"), encoding="UTF-8")
+    
+    table(all$univ_id)
+    all$univ_id <- as.character(all$univ_id)
+    all$univ_id[all$univ_id=="166629_req"]<-c("166629")
+    all$univ_id[all$univ_id=="110653_req"]<-c("110653")
+    all$univ_id[all$univ_id=="196097_req"]<-c("196097")
+    all$univ_id[all$univ_id=="186380_req"]<-c("186380")
[…]

### 2024-07-08 scripts/create_1980_decennial.R

@@ -1,0 +1,910 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_1980_decennial.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 6/27/2024
+## [ DESC ] < Get 1980 decennial census data, shape files, and (hopefully) merge w/ eps geomarkets >
+################################################################################
+
+# OTHER VARIABLES TO CREATE
+  # PERCENT LIVING IN POVERTY
+  # PERCENT RECEIVING PUBLIC ASSISTANCE
+
+### SETTINGS
+  rm(list = ls())
+  options(max.print=1000)
+  #options(width = 160)
+  #d1980_stf1f3_anal_eps %>% glimpse()
+  
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+  # get census api key
+    #http://api.census.gov/data/key_signup.html
+  # set census api key, and install for future use
+  #census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+  Sys.getenv("CENSUS_API_KEY") # retreive API key
+library(sf)
+  # Enable caching for tigris
+  options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # main data directory
+  list.files(path = data_dir)
+
+d1980_data_dir <-   file.path('.',data_dir,'1980_decennial') # main data directory
+  list.files(path = d1980_data_dir)
+
+# can't save shape files in the repo cuz they too big
+shape_dir <-   file.path('.','..','cb_geomarket_shape') # main data directory
+  list.files(path = shape_dir)
+
+eps_data_dir <- file.path(data_dir,'eps_market') # has eps geomarket data and spaical data files
+  list.files(path = eps_data_dir)
+  
+scripts_dir <- file.path('.','scripts') # 
+  list.files(path = scripts_dir)
+
+#################
+# CREATE EPS-LEVEL DATA THAT HAS SHAPEFILE FOR EACH GEOMARKET
+  
[…]

### 2024-07-08 scripts/create_2000_decennial.R

@@ -1,0 +1,124 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_2000_decennial.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 6/27/2024
+## [ DESC ] < Get 2000 decennial census data, shape files, and (hopefully) merge w/ eps geomarkets >
+################################################################################
+
+### SETTINGS
+rm(list = ls())
+options(max.print=1000)
+#options(width = 160)
+#d1980_stf1f3_anal_eps %>% glimpse()
+
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+# get census api key
+#http://api.census.gov/data/key_signup.html
+# set census api key, and install for future use
+#census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+Sys.getenv("CENSUS_API_KEY") # retreive API key
+library(sf)
+# Enable caching for tigris
+options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+
+### DIRECTORY PATHS
+
+##### GETTING STARTED FROM CHATGPT
+
+  # This code will retrieve census tract level data from the 2000 Decennial Census, including geometries for each tract, which you can use for further analysis or mapping. The comments provide explanations and additional notes for clarity.
+
+# Get 2000 Decennial Census data for a specific variable and geography (census tract level)
+census_2000_tract_data <- get_decennial(
+  geography = "tract",      # Set the geography parameter to "tract" to get data at the census tract level
+  variables = "P001001",    # Specify the variable code(s) you are interested in, such as "P001001" for total population
+  year = 2000,              # Set the year parameter to 2000 to get data from the 2000 Decennial Census
+  #state = "06",             # FIPS code for California; adjust as needed
+  #county = "*",             # Use "*" to get data for all counties within the state
+  geometry = TRUE           # Include geometries for spatial analysis
+)
+
+# View the data
+head(census_2000_tract_data)
+
+# Explanation:
+# 1. Census API Key: Ensure you have set your Census API key using census_api_key("your_api_key_here").
+# 2. Geography: Set the geography parameter to "tract" to get data at the census tract level.
[…]

### 2024-07-08 scripts/create_geomarket.R

@@ -1,0 +1,707 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_geomarket.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 5/11/2023
+## [ DESC ] < Create geomarket shape files >
+################################################################################
+
+### SETTINGS
+  rm(list = ls())
+  options(max.print=1000)
+
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+  # get census api key
+    #http://api.census.gov/data/key_signup.html
+  # set census api key, and install for future use
+  #census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+  Sys.getenv("CENSUS_API_KEY") # retreive API key
+library(sf)
+  # Enable caching for tigris
+  options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # 
+  list.files(path = data_dir)
+  
+eps_data_dir <- file.path(data_dir,'eps_market') # 
+  list.files(path = eps_data_dir)
+  
+scripts_dir <- file.path('.','scripts') # 
+  list.files(path = scripts_dir)
+
+  
+########### STEPS
+  
+  # Read in .xls file that has name of geomarket associated with each zip
+  # get ZCTA shape file from around 2010 from tidycensus function
+  # merge eps file to zcta file by zip code, to obtain dataset that has geomarket and shape-file for each zip-code
+  # create dataset that has one observation per geomarket by using aggregate to "dissolve" zip codes into geomarkets
+  # Using tidycensus, create dataset(s) with one observation per zip-code with data on desired characteristics (e.g., income, racial composition, population, education level, etc.)
+    # can use this dataset to do analyses of how characteristics vary within geomarket
+  # create dataset(s) with one obsr per geomarket by using group_by() and summarize()
[…]

### 2024-07-08 scripts/create_geomarket_decennial.R

@@ -1,0 +1,762 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_geomarket.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 5/11/2023
+## [ DESC ] < Create geomarket shape files >
+################################################################################
+
+### SETTINGS
+  rm(list = ls())
+  options(max.print=1000)
+
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+  # get census api key
+    #http://api.census.gov/data/key_signup.html
+  # set census api key, and install for future use
+  #census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+  Sys.getenv("CENSUS_API_KEY") # retreive API key
+library(sf)
+  # Enable caching for tigris
+  options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # 
+  list.files(path = data_dir)
+  
+eps_data_dir <- file.path(data_dir,'eps_market') # 
+  list.files(path = eps_data_dir)
+  
+scripts_dir <- file.path('.','scripts') # 
+  list.files(path = scripts_dir)
+
+  
+########### STEPS
+  
+  # Read in .xls file that has name of geomarket associated with each zip
+  # get ZCTA shape file from around 2010 from tidycensus function
+  # merge eps file to zcta file by zip code, to obtain dataset that has geomarket and shape-file for each zip-code
+  # create dataset that has one observation per geomarket by using aggregate to "dissolve" zip codes into geomarkets
+  # Using tidycensus, create dataset(s) with one observation per zip-code with data on desired characteristics (e.g., income, racial composition, population, education level, etc.)
+    # can use this dataset to do analyses of how characteristics vary within geomarket
+  # create dataset(s) with one obsr per geomarket by using group_by() and summarize()
+    # calculates descriptive statistics at geomarket level, to show how characteristics vary between geomarkets
[…]

### 2024-09-11 results/tables/socal_table.tex
worked on tables

\begin{table}[!h]
\centering
\begin{tabular}[t]{l|>{}l||r|r|r|r|r|r|r|r|r|r}
\hline
\multicolumn{2}{c|}{ } & \multicolumn{5}{c|}{2000} & \multicolumn{5}{c}{2020} \\
\cline{3-7} \cline{8-12}
\textbf{Variable} & \textbf{EPS Name} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75}\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{40.4} & \cellcolor{gray!6}{21.0} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{36.8} & \cellcolor{gray!6}{56.2} & \cellcolor{gray!6}{26.1} & \cellcolor{gray!6}{17.8} & \cellcolor{gray!6}{12.6} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{36.9}\\
\hline
\hline
\% White, non-Hispanic & Covina \& West Covina & 32.8 & 23.0 & 11.9 & 27.9 & 49.3 & 18.8 & 16.8 & 5.4 & 12.5 & 27.1\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{10.3} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{6.2} & \cellcolor{gray!6}{11.8} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{8.4}\\
\hline
\% White, non-Hispanic & Glendale and Pasadena & 39.9 & 24.0 & 16.0 & 40.4 & 61.3 & 35.3 & 24.0 & 12.8 & 34.1 & 57.2\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{26.5} & \cellcolor{gray!6}{28.5} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{40.8} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{25.7} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{22.6} & \cellcolor{gray!6}{53.5}\\
\hline
\% White, non-Hispanic & Long Beach & 35.1 & 28.9 & 10.0 & 21.8 & 64.0 & 25.8 & 23.0 & 6.6 & 16.4 & 45.5\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{45.2} & \cellcolor{gray!6}{21.8} & \cellcolor{gray!6}{28.1} & \cellcolor{gray!6}{43.6} & \cellcolor{gray!6}{63.7} & \cellcolor{gray!6}{27.7} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{12.5} & \cellcolor{gray!6}{23.5} & \cellcolor{gray!6}{40.7}\\
\hline
\% White, non-Hispanic & San Fernando Valley (East) & 45.1 & 28.1 & 21.4 & 40.8 & 74.2 & 43.3 & 25.7 & 22.5 & 39.6 & 67.0\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{58.3} & \cellcolor{gray!6}{25.6} & \cellcolor{gray!6}{40.4} & \cellcolor{gray!6}{67.0} & \cellcolor{gray!6}{78.8} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{24.2} & \cellcolor{gray!6}{22.8} & \cellcolor{gray!6}{46.6} & \cellcolor{gray!6}{64.7}\\
\hline
\% White, non-Hispanic & Santa Ana & 52.4 & 29.0 & 24.5 & 58.3 & 77.9 & 42.0 & 26.3 & 17.0 & 44.1 & 64.6\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{7.3} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{5.5} & \cellcolor{gray!6}{6.3} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{6.6}\\
\hline
\% White, non-Hispanic & South Bay & 41.3 & 29.4 & 12.4 & 41.3 & 68.3 & 29.1 & 24.5 & 7.1 & 22.1 & 51.4\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{66.0} & \cellcolor{gray!6}{18.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{70.4} & \cellcolor{gray!6}{78.1} & \cellcolor{gray!6}{52.9} & \cellcolor{gray!6}{18.1} & \cellcolor{gray!6}{40.7} & \cellcolor{gray!6}{55.0} & \cellcolor{gray!6}{66.3}\\
\hline
\% White, non-Hispanic & West Los Angeles \& West Beach & 63.9 & 24.4 & 43.0 & 72.9 & 84.0 & 60.2 & 20.2 & 45.9 & 62.9 & 75.5\\
\hline
\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{35.4} & \cellcolor{gray!6}{20.5} & \cellcolor{gray!6}{18.5} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{50.7} & \cellcolor{gray!6}{22.4} & \cellcolor{gray!6}{15.3} & \cellcolor{gray!6}{10.9} & \cellcolor{gray!6}{17.5} & \cellcolor{gray!6}{32.7}\\
\hline
\% Black, non-Hispanic & Anaheim & 3.9 & 5.3 & 1.3 & 2.3 & 4.4 & 4.4 & 5.6 & 0.8 & 2.3 & 6.0\\
\hline
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{5.2} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{4.7}\\
\hline
\% Black, non-Hispanic & East Los Angeles & 1.1 & 2.8 & 0.2 & 0.5 & 1.0 & 1.3 & 2.8 & 0.0 & 0.4 & 1.4\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{4.4} & \cellcolor{gray!6}{8.5} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{1.7} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{3.4} & \cellcolor{gray!6}{5.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{1.8} & \cellcolor{gray!6}{4.1}\\
\hline
\% Black, non-Hispanic & Hollywood \& Wilshire & 13.0 & 21.0 & 2.0 & 3.3 & 8.5 & 11.1 & 16.8 & 1.9 & 4.5 & 10.5\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{13.1} & \cellcolor{gray!6}{14.4} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{8.9} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{3.6} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{15.1}\\
\hline
\% Black, non-Hispanic & Riverside, San Bernardino, \& Ontario & 8.2 & 7.0 & 3.2 & 5.8 & 11.4 & 7.1 & 6.4 & 2.4 & 5.4 & 9.8\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{5.3} & \cellcolor{gray!6}{4.3} & \cellcolor{gray!6}{3.8} & \cellcolor{gray!6}{1.6} & \cellcolor{gray!6}{3.3} & \cellcolor{gray!6}{5.7}\\
\hline
\% Black, non-Hispanic & San Fernando Valley (West) & 3.2 & 3.1 & 1.2 & 2.5 & 3.8 & 3.8 & 3.9 & 1.1 & 2.7 & 5.4\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{1.8} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.9} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{1.3} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.7} & \cellcolor{gray!6}{1.6}\\
\hline
\% Black, non-Hispanic & South \& South Central Los Angeles & 24.7 & 24.4 & 2.6 & 17.7 & 39.6 & 17.5 & 18.8 & 2.7 & 11.6 & 25.3\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{13.2} & \cellcolor{gray!6}{20.2} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{15.2} & \cellcolor{gray!6}{12.1} & \cellcolor{gray!6}{16.9} & \cellcolor{gray!6}{1.7} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{14.3}\\
\hline
\% Black, non-Hispanic & South Orange Co & 1.4 & 1.2 & 0.7 & 1.2 & 1.8 & 1.5 & 1.8 & 0.3 & 1.0 & 2.3\\
\hline
\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{6.3} & \cellcolor{gray!6}{11.7} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{6.2}\\
\hline
\% Black, non-Hispanic & Whittier \& North Orange Co & 2.2 & 2.3 & 0.9 & 1.5 & 2.5 & 2.4 & 2.7 & 0.4 & 1.5 & 3.0\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{34.0} & \cellcolor{gray!6}{22.0} & \cellcolor{gray!6}{16.3} & \cellcolor{gray!6}{29.7} & \cellcolor{gray!6}{49.1} & \cellcolor{gray!6}{41.3} & \cellcolor{gray!6}{22.7} & \cellcolor{gray!6}{23.1} & \cellcolor{gray!6}{38.6} & \cellcolor{gray!6}{58.1}\\
\hline
\hline
\% Hispanic & Covina \& West Covina & 43.6 & 25.6 & 20.6 & 39.6 & 68.7 & 51.7 & 23.5 & 30.8 & 54.3 & 71.3\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{66.6} & \cellcolor{gray!6}{27.9} & \cellcolor{gray!6}{41.6} & \cellcolor{gray!6}{77.9} & \cellcolor{gray!6}{91.0} & \cellcolor{gray!6}{64.4} & \cellcolor{gray!6}{28.9} & \cellcolor{gray!6}{36.7} & \cellcolor{gray!6}{73.2} & \cellcolor{gray!6}{91.5}\\
\hline
\% Hispanic & Glendale and Pasadena & 34.4 & 25.8 & 13.8 & 26.3 & 53.5 & 36.8 & 26.9 & 16.3 & 26.3 & 53.4\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{43.7} & \cellcolor{gray!6}{28.3} & \cellcolor{gray!6}{15.8} & \cellcolor{gray!6}{44.4} & \cellcolor{gray!6}{68.5} & \cellcolor{gray!6}{38.2} & \cellcolor{gray!6}{24.6} & \cellcolor{gray!6}{15.7} & \cellcolor{gray!6}{34.7} & \cellcolor{gray!6}{57.1}\\
\hline
\% Hispanic & Long Beach & 37.1 & 24.3 & 16.3 & 32.1 & 53.1 & 45.8 & 23.3 & 26.6 & 43.5 & 63.4\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{38.3} & \cellcolor{gray!6}{20.5} & \cellcolor{gray!6}{21.7} & \cellcolor{gray!6}{34.2} & \cellcolor{gray!6}{52.9} & \cellcolor{gray!6}{53.8} & \cellcolor{gray!6}{22.0} & \cellcolor{gray!6}{36.3} & \cellcolor{gray!6}{54.1} & \cellcolor{gray!6}{72.4}\\
\hline
\% Hispanic & San Fernando Valley (East) & 40.2 & 27.5 & 11.3 & 35.8 & 62.5 & 40.6 & 27.2 & 14.8 & 37.3 & 60.9\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{27.4} & \cellcolor{gray!6}{25.0} & \cellcolor{gray!6}{9.4} & \cellcolor{gray!6}{16.4} & \cellcolor{gray!6}{38.8} & \cellcolor{gray!6}{36.8} & \cellcolor{gray!6}{25.9} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{28.8} & \cellcolor{gray!6}{56.9}\\
\hline
\% Hispanic & Santa Ana & 29.6 & 29.2 & 8.1 & 15.0 & 43.0 & 33.0 & 27.9 & 12.3 & 21.2 & 46.6\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{63.6} & \cellcolor{gray!6}{26.9} & \cellcolor{gray!6}{46.5} & \cellcolor{gray!6}{65.9} & \cellcolor{gray!6}{88.3} & \cellcolor{gray!6}{70.5} & \cellcolor{gray!6}{24.8} & \cellcolor{gray!6}{60.2} & \cellcolor{gray!6}{75.7} & \cellcolor{gray!6}{90.2}\\
\hline
\% Hispanic & South Bay & 25.6 & 20.6 & 8.8 & 19.8 & 39.0 & 33.2 & 22.5 & 12.9 & 29.8 & 51.8\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{18.0} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{7.6} & \cellcolor{gray!6}{11.6} & \cellcolor{gray!6}{21.3} & \cellcolor{gray!6}{22.2} & \cellcolor{gray!6}{17.0} & \cellcolor{gray!6}{10.4} & \cellcolor{gray!6}{16.8} & \cellcolor{gray!6}{27.2}\\
\hline
\% Hispanic & West Los Angeles \& West Beach & 16.2 & 16.6 & 4.6 & 9.4 & 21.1 & 15.8 & 13.4 & 6.3 & 12.2 & 20.4\\
\hline
\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{47.5} & \cellcolor{gray!6}{24.9} & \cellcolor{gray!6}{27.2} & \cellcolor{gray!6}{46.7} & \cellcolor{gray!6}{67.1} & \cellcolor{gray!6}{56.0} & \cellcolor{gray!6}{23.6} & \cellcolor{gray!6}{39.3} & \cellcolor{gray!6}{59.2} & \cellcolor{gray!6}{75.7}\\
\hline
\% API, non-Hispanic & Anaheim & 18.8 & 15.7 & 8.1 & 14.3 & 23.2 & 24.5 & 18.1 & 11.2 & 19.4 & 34.2\\
\hline
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{5.8} & \cellcolor{gray!6}{9.7} & \cellcolor{gray!6}{21.9} & \cellcolor{gray!6}{23.7} & \cellcolor{gray!6}{19.3} & \cellcolor{gray!6}{10.0} & \cellcolor{gray!6}{16.7} & \cellcolor{gray!6}{31.9}\\
\hline
\% API, non-Hispanic & East Los Angeles & 20.5 & 21.5 & 1.8 & 11.2 & 37.9 & 26.5 & 26.0 & 2.7 & 16.4 & 47.9\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{17.4} & \cellcolor{gray!6}{15.5} & \cellcolor{gray!6}{6.2} & \cellcolor{gray!6}{11.6} & \cellcolor{gray!6}{23.1} & \cellcolor{gray!6}{20.9} & \cellcolor{gray!6}{19.8} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{27.3}\\
\hline
\% API, non-Hispanic & Hollywood \& Wilshire & 13.8 & 12.9 & 4.6 & 9.5 & 18.9 & 15.7 & 14.0 & 5.7 & 11.4 & 20.1\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{10.1} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{15.3} & \cellcolor{gray!6}{13.2} & \cellcolor{gray!6}{11.3} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{18.2}\\
\hline
\% API, non-Hispanic & Riverside, San Bernardino, \& Ontario & 5.2 & 6.5 & 1.8 & 3.5 & 6.0 & 8.2 & 10.0 & 2.0 & 5.0 & 10.4\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{7.2} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{4.6} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{7.7} & \cellcolor{gray!6}{8.3} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{6.9} & \cellcolor{gray!6}{10.2}\\
\hline
\% API, non-Hispanic & San Fernando Valley (West) & 8.0 & 5.9 & 4.2 & 6.5 & 10.8 & 11.6 & 8.1 & 5.3 & 10.4 & 16.3\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{14.1} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{4.5} & \cellcolor{gray!6}{9.2} & \cellcolor{gray!6}{18.1} & \cellcolor{gray!6}{20.1} & \cellcolor{gray!6}{18.3} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{14.3} & \cellcolor{gray!6}{26.5}\\
\hline
\% API, non-Hispanic & South \& South Central Los Angeles & 2.9 & 5.3 & 0.3 & 0.8 & 2.7 & 3.9 & 6.9 & 0.2 & 1.2 & 4.4\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{16.6} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{6.1} & \cellcolor{gray!6}{11.0} & \cellcolor{gray!6}{26.8} & \cellcolor{gray!6}{21.2} & \cellcolor{gray!6}{15.9} & \cellcolor{gray!6}{8.2} & \cellcolor{gray!6}{17.8} & \cellcolor{gray!6}{30.3}\\
\hline
\% API, non-Hispanic & South Orange Co & 11.1 & 9.8 & 5.3 & 8.5 & 13.6 & 18.8 & 15.1 & 9.0 & 14.1 & 23.6\\
\hline
\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{9.9} & \cellcolor{gray!6}{7.9} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{7.5} & \cellcolor{gray!6}{11.7} & \cellcolor{gray!6}{13.1} & \cellcolor{gray!6}{9.5} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{17.6}\\
\hline
\% API, non-Hispanic & Whittier \& North Orange Co & 12.6 & 15.4 & 3.6 & 8.2 & 12.9 & 15.9 & 16.2 & 5.2 & 12.0 & 18.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.2}\\
\hline
\hline
\% Native, non-Hispanic & Covina \& West Covina & 0.3 & 0.2 & 0.2 & 0.3 & 0.4 & 0.2 & 0.7 & 0.0 & 0.0 & 0.2\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.2}\\
\hline
\% Native, non-Hispanic & Glendale and Pasadena & 0.3 & 0.2 & 0.1 & 0.2 & 0.3 & 0.2 & 0.5 & 0.0 & 0.0 & 0.1\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
\hline
\% Native, non-Hispanic & Long Beach & 0.4 & 0.3 & 0.2 & 0.3 & 0.5 & 0.6 & 6.9 & 0.0 & 0.0 & 0.1\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.3}\\
\hline
\% Native, non-Hispanic & San Fernando Valley (East) & 0.2 & 0.1 & 0.1 & 0.2 & 0.3 & 0.1 & 0.3 & 0.0 & 0.0 & 0.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.1}\\
\hline
\% Native, non-Hispanic & Santa Ana & 0.3 & 0.2 & 0.2 & 0.3 & 0.4 & 0.2 & 0.4 & 0.0 & 0.0 & 0.2\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
\hline
\% Native, non-Hispanic & South Bay & 0.2 & 0.1 & 0.1 & 0.2 & 0.3 & 0.2 & 0.5 & 0.0 & 0.0 & 0.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
\hline
\% Native, non-Hispanic & West Los Angeles \& West Beach & 0.2 & 0.2 & 0.1 & 0.2 & 0.3 & 0.2 & 0.4 & 0.0 & 0.0 & 0.0\\
\hline
\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{7.7} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.3}\\
\hline
pct\_nhisp\_multi & Anaheim & 2.4 & 0.9 & 1.8 & 2.4 & 3.0 & 2.9 & 2.3 & 1.1 & 2.1 & 4.5\\
\hline
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{2.9}\\
\hline
pct\_nhisp\_multi & East Los Angeles & 1.1 & 0.9 & 0.4 & 0.9 & 1.7 & 0.9 & 1.2 & 0.0 & 0.4 & 1.4\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{4.2}\\
\hline
pct\_nhisp\_multi & Hollywood \& Wilshire & 2.6 & 2.0 & 1.3 & 2.1 & 3.1 & 3.2 & 2.9 & 1.3 & 2.4 & 4.6\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{3.3} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{2.5} & \cellcolor{gray!6}{4.2}\\
\hline
pct\_nhisp\_multi & Riverside, San Bernardino, \& Ontario & 2.4 & 0.9 & 1.8 & 2.3 & 3.0 & 2.6 & 2.4 & 0.7 & 1.9 & 3.7\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{4.4}\\
\hline
pct\_nhisp\_multi & San Fernando Valley (West) & 2.6 & 1.5 & 1.7 & 2.4 & 3.2 & 3.1 & 2.6 & 1.1 & 2.7 & 4.1\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{2.2} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{4.9}\\
\hline
pct\_nhisp\_multi & South \& South Central Los Angeles & 1.1 & 1.1 & 0.4 & 0.8 & 1.5 & 1.3 & 2.0 & 0.0 & 0.5 & 1.8\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{3.8} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{5.6}\\
\hline
pct\_nhisp\_multi & South Orange Co & 3.0 & 6.2 & 1.8 & 2.4 & 3.2 & 4.1 & 2.7 & 2.1 & 3.7 & 5.5\\
\hline
\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{1.3} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{3.9} & \cellcolor{gray!6}{4.8} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{6.7}\\
\hline
pct\_nhisp\_multi & Whittier \& North Orange Co & 1.8 & 0.7 & 1.3 & 1.9 & 2.3 & 1.9 & 1.7 & 0.5 & 1.5 & 2.8\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{94.0} & \cellcolor{gray!6}{119.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{41.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{99.0} & \cellcolor{gray!6}{129.0}\\
\hline
\hline
Median income & Covina \& West Covina & 103.0 & 36.0 & 79.0 & 96.0 & 121.0 & 103.0 & 32.0 & 82.0 & 99.0 & 120.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{70.0} & \cellcolor{gray!6}{26.0} & \cellcolor{gray!6}{53.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{81.0} & \cellcolor{gray!6}{76.0} & \cellcolor{gray!6}{25.0} & \cellcolor{gray!6}{60.0} & \cellcolor{gray!6}{72.0} & \cellcolor{gray!6}{87.0}\\
\hline
Median income & Glendale and Pasadena & 96.0 & 40.0 & 69.0 & 86.0 & 110.0 & 102.0 & 40.0 & 73.0 & 96.0 & 121.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{41.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{86.0} & \cellcolor{gray!6}{84.0} & \cellcolor{gray!6}{43.0} & \cellcolor{gray!6}{52.0} & \cellcolor{gray!6}{75.0} & \cellcolor{gray!6}{105.0}\\
\hline
Median income & Long Beach & 84.0 & 41.0 & 54.0 & 79.0 & 108.0 & 90.0 & 36.0 & 62.0 & 85.0 & 111.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{36.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{81.0} & \cellcolor{gray!6}{111.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{37.0} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{88.0} & \cellcolor{gray!6}{119.0}\\
\hline
Median income & San Fernando Valley (East) & 94.0 & 62.0 & 60.0 & 75.0 & 97.0 & 99.0 & 56.0 & 61.0 & 83.0 & 116.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{52.0} & \cellcolor{gray!6}{87.0} & \cellcolor{gray!6}{113.0} & \cellcolor{gray!6}{156.0} & \cellcolor{gray!6}{121.0} & \cellcolor{gray!6}{46.0} & \cellcolor{gray!6}{87.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{146.0}\\
\hline
Median income & Santa Ana & 116.0 & 44.0 & 86.0 & 107.0 & 141.0 & 119.0 & 44.0 & 87.0 & 112.0 & 142.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{59.0} & \cellcolor{gray!6}{23.0} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{55.0} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{67.0} & \cellcolor{gray!6}{27.0} & \cellcolor{gray!6}{50.0} & \cellcolor{gray!6}{62.0} & \cellcolor{gray!6}{78.0}\\
\hline
Median income & South Bay & 111.0 & 52.0 & 73.0 & 99.0 & 134.0 & 115.0 & 51.0 & 78.0 & 103.0 & 141.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{140.0} & \cellcolor{gray!6}{49.0} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{137.0} & \cellcolor{gray!6}{167.0} & \cellcolor{gray!6}{142.0} & \cellcolor{gray!6}{46.0} & \cellcolor{gray!6}{109.0} & \cellcolor{gray!6}{140.0} & \cellcolor{gray!6}{171.0}\\
\hline
Median income & West Los Angeles \& West Beach & 132.0 & 78.0 & 80.0 & 104.0 & 168.0 & 137.0 & 61.0 & 98.0 & 122.0 & 165.0\\
\hline
\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{101.0} & \cellcolor{gray!6}{29.0} & \cellcolor{gray!6}{80.0} & \cellcolor{gray!6}{94.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{105.0} & \cellcolor{gray!6}{30.0} & \cellcolor{gray!6}{85.0} & \cellcolor{gray!6}{101.0} & \cellcolor{gray!6}{121.0}\\
\hline
Mean income & Anaheim & 127.0 & 61.0 & 97.0 & 111.0 & 135.0 & 131.0 & 54.0 & 99.0 & 118.0 & 146.0\\
\hline
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{126.0} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{144.0} & \cellcolor{gray!6}{125.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{100.0} & \cellcolor{gray!6}{115.0} & \cellcolor{gray!6}{143.0}\\
\hline
Mean income & East Los Angeles & 89.0 & 32.0 & 72.0 & 83.0 & 100.0 & 97.0 & 33.0 & 77.0 & 93.0 & 106.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{56.0} & \cellcolor{gray!6}{86.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{138.0} & \cellcolor{gray!6}{135.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{96.0} & \cellcolor{gray!6}{121.0} & \cellcolor{gray!6}{157.0}\\
\hline
Mean income & Hollywood \& Wilshire & 99.0 & 69.0 & 60.0 & 78.0 & 111.0 & 118.0 & 68.0 & 73.0 & 98.0 & 143.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{106.0} & \cellcolor{gray!6}{49.0} & \cellcolor{gray!6}{72.0} & \cellcolor{gray!6}{102.0} & \cellcolor{gray!6}{129.0} & \cellcolor{gray!6}{113.0} & \cellcolor{gray!6}{45.0} & \cellcolor{gray!6}{80.0} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{137.0}\\
\hline
Mean income & Riverside, San Bernardino, \& Ontario & 107.0 & 41.0 & 79.0 & 99.0 & 128.0 & 114.0 & 42.0 & 83.0 & 103.0 & 137.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{105.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{97.0} & \cellcolor{gray!6}{128.0} & \cellcolor{gray!6}{137.0} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{84.0} & \cellcolor{gray!6}{109.0} & \cellcolor{gray!6}{154.0}\\
\hline
Mean income & San Fernando Valley (West) & 154.0 & 69.0 & 106.0 & 134.0 & 183.0 & 152.0 & 65.0 & 105.0 & 138.0 & 180.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{145.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{102.0} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{167.0} & \cellcolor{gray!6}{152.0} & \cellcolor{gray!6}{62.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{139.0} & \cellcolor{gray!6}{174.0}\\
\hline
Mean income & South \& South Central Los Angeles & 77.0 & 26.0 & 62.0 & 72.0 & 87.0 & 86.0 & 32.0 & 66.0 & 78.0 & 95.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{141.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{118.0} & \cellcolor{gray!6}{164.0} & \cellcolor{gray!6}{149.0} & \cellcolor{gray!6}{77.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{180.0}\\
\hline
Mean income & South Orange Co & 173.0 & 66.0 & 125.0 & 163.0 & 203.0 & 179.0 & 62.0 & 138.0 & 168.0 & 213.0\\
\hline
\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{197.0} & \cellcolor{gray!6}{136.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{144.0} & \cellcolor{gray!6}{266.0} & \cellcolor{gray!6}{206.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{169.0} & \cellcolor{gray!6}{252.0}\\
\hline
Mean income & Whittier \& North Orange Co & 121.0 & 46.0 & 96.0 & 110.0 & 135.0 & 125.0 & 37.0 & 102.0 & 117.0 & 142.0\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{6.9} & \cellcolor{gray!6}{4.6} & \cellcolor{gray!6}{8.0} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{10.7} & \cellcolor{gray!6}{6.7} & \cellcolor{gray!6}{5.5} & \cellcolor{gray!6}{9.6} & \cellcolor{gray!6}{14.8}\\
\hline
\hline
\% in poverty & Covina \& West Covina & 10.3 & 6.7 & 5.2 & 8.8 & 13.5 & 10.3 & 6.4 & 5.8 & 9.0 & 12.8\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{19.2} & \cellcolor{gray!6}{9.5} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{25.4} & \cellcolor{gray!6}{16.0} & \cellcolor{gray!6}{7.9} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{15.1} & \cellcolor{gray!6}{21.1}\\
\hline
\% in poverty & Glendale and Pasadena & 12.2 & 7.4 & 6.5 & 10.7 & 16.9 & 12.6 & 7.7 & 6.4 & 11.3 & 16.8\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{22.1} & \cellcolor{gray!6}{12.9} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{20.8} & \cellcolor{gray!6}{31.0} & \cellcolor{gray!6}{17.7} & \cellcolor{gray!6}{11.8} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{14.7} & \cellcolor{gray!6}{23.9}\\
\hline
\% in poverty & Long Beach & 16.4 & 13.1 & 5.5 & 12.5 & 25.1 & 14.5 & 12.0 & 6.8 & 12.1 & 19.9\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{10.8} & \cellcolor{gray!6}{17.0} & \cellcolor{gray!6}{12.4} & \cellcolor{gray!6}{9.2} & \cellcolor{gray!6}{5.8} & \cellcolor{gray!6}{10.6} & \cellcolor{gray!6}{17.2}\\
\hline
\% in poverty & San Fernando Valley (East) & 14.6 & 9.0 & 7.4 & 12.7 & 19.5 & 14.6 & 8.9 & 8.2 & 13.3 & 19.0\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{8.2} & \cellcolor{gray!6}{7.0} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{10.8} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{7.8} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{13.3}\\
\hline
\% in poverty & Santa Ana & 8.7 & 7.1 & 3.6 & 6.3 & 12.2 & 9.8 & 6.4 & 5.0 & 8.3 & 13.0\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{26.7} & \cellcolor{gray!6}{12.7} & \cellcolor{gray!6}{17.9} & \cellcolor{gray!6}{26.1} & \cellcolor{gray!6}{35.7} & \cellcolor{gray!6}{21.7} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{20.6} & \cellcolor{gray!6}{28.5}\\
\hline
\% in poverty & South Bay & 10.2 & 7.9 & 3.9 & 7.2 & 16.4 & 9.6 & 6.9 & 4.9 & 7.5 & 13.3\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{5.9} & \cellcolor{gray!6}{2.5} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{6.4} & \cellcolor{gray!6}{7.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{6.1} & \cellcolor{gray!6}{9.6}\\
\hline
\% in poverty & West Los Angeles \& West Beach & 11.3 & 12.0 & 5.0 & 8.3 & 13.2 & 10.4 & 10.5 & 5.2 & 8.2 & 11.8\\
\hline
\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{5.1} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{12.3} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{5.9} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{8.3} & \cellcolor{gray!6}{12.8}\\
\hline
\% with BA+ & Anaheim & 22.2 & 15.5 & 11.9 & 16.5 & 26.3 & 31.1 & 17.4 & 19.2 & 25.3 & 40.8\\
\hline
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{16.0} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{35.3} & \cellcolor{gray!6}{30.5} & \cellcolor{gray!6}{16.6} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{28.5} & \cellcolor{gray!6}{44.0}\\
\hline
\% with BA+ & East Los Angeles & 13.6 & 12.8 & 4.6 & 8.3 & 20.5 & 21.2 & 13.5 & 11.5 & 16.2 & 30.1\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{29.4} & \cellcolor{gray!6}{17.6} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{27.5} & \cellcolor{gray!6}{40.5} & \cellcolor{gray!6}{39.1} & \cellcolor{gray!6}{19.3} & \cellcolor{gray!6}{24.2} & \cellcolor{gray!6}{39.2} & \cellcolor{gray!6}{53.7}\\
\hline
\% with BA+ & Hollywood \& Wilshire & 25.5 & 18.5 & 10.0 & 19.6 & 42.2 & 42.1 & 19.9 & 25.2 & 40.6 & 59.8\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{20.9} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{16.9} & \cellcolor{gray!6}{31.7} & \cellcolor{gray!6}{28.4} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{25.9} & \cellcolor{gray!6}{40.8}\\
\hline
\% with BA+ & Riverside, San Bernardino, \& Ontario & 16.6 & 12.7 & 7.0 & 12.4 & 23.0 & 22.5 & 14.2 & 10.8 & 19.4 & 31.3\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{28.2} & \cellcolor{gray!6}{18.6} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{23.3} & \cellcolor{gray!6}{40.3} & \cellcolor{gray!6}{39.3} & \cellcolor{gray!6}{21.4} & \cellcolor{gray!6}{21.3} & \cellcolor{gray!6}{34.4} & \cellcolor{gray!6}{59.1}\\
\hline
\% with BA+ & San Fernando Valley (West) & 30.1 & 16.3 & 18.2 & 27.7 & 43.2 & 38.0 & 17.8 & 24.7 & 37.1 & 52.0\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{30.6} & \cellcolor{gray!6}{19.4} & \cellcolor{gray!6}{12.9} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{46.1} & \cellcolor{gray!6}{40.6} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{40.8} & \cellcolor{gray!6}{56.3}\\
\hline
\% with BA+ & South \& South Central Los Angeles & 8.9 & 10.0 & 3.1 & 5.4 & 10.1 & 16.2 & 14.9 & 7.1 & 10.8 & 19.5\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{31.7} & \cellcolor{gray!6}{20.8} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{25.8} & \cellcolor{gray!6}{46.9} & \cellcolor{gray!6}{41.5} & \cellcolor{gray!6}{21.8} & \cellcolor{gray!6}{22.6} & \cellcolor{gray!6}{36.7} & \cellcolor{gray!6}{58.8}\\
\hline
\% with BA+ & South Orange Co & 40.4 & 16.3 & 29.2 & 40.0 & 53.4 & 52.3 & 15.7 & 42.9 & 51.6 & 63.8\\
\hline
\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{50.6} & \cellcolor{gray!6}{16.7} & \cellcolor{gray!6}{42.3} & \cellcolor{gray!6}{53.4} & \cellcolor{gray!6}{62.7} & \cellcolor{gray!6}{65.6} & \cellcolor{gray!6}{14.4} & \cellcolor{gray!6}{60.1} & \cellcolor{gray!6}{67.9} & \cellcolor{gray!6}{74.6}\\
\hline
\% with BA+ & Whittier \& North Orange Co & 21.4 & 14.0 & 10.2 & 16.3 & 32.0 & 29.0 & 14.9 & 17.9 & 23.9 & 40.9\\
\hline
\end{tabular}
\end{table}

### 2024-09-11 scripts/leaflet_maps.R

@@ -69,6 +69,10 @@ list.files(path = eps_data_dir)
 scripts_dir <- file.path('.','scripts') # 
 list.files(path = scripts_dir)
 
+tables_dir <- file.path('.','results','tables') # 
+list.files(path = tables_dir)
+
+
 ##### LOAD EPS-LEVEL DATA
 
 # load EPS data
@@ -200,8 +204,6 @@ cleveland_eps_codes
 
   allyr_anal_tract_sf %>% glimpse()
 
-# Define table_vars in the desired order
-table_vars <- c("pct_nhisp_white", "pct_nhisp_black", "pct_hisp_all", "pct_nhisp_api", "pct_nhisp_native", "med_inc_house", "mean_inc_house", "pct_pov_yes", "pct_edu_baplus_all")
 
 table_var_names <- list(
   "pct_hisp_all" = "% Hispanic",
@@ -216,69 +218,280 @@ table_var_names <- list(
   "pct_edu_baplus_all" = "% with BA+"
 )
 
-eps_table <- allyr_anal_tract_sf %>% 
-  as.data.frame() %>% 
-  filter(eps %in% chi_eps_codes) %>%
-  
-  # Divide income variables by 1,000 and round to nearest integer [don't round here]
-  mutate(across(contains("_inc"), ~ (.x / 1000))) %>%
-  
-  group_by(eps_name, year) %>%
-  summarize(
-    # Mean calculation
-    across(all_of(table_vars), ~ mean(.x, na.rm = TRUE), .names = "mean_{.col}"),
-    
-    # Standard deviation calculation
-    across(all_of(table_vars), ~ sd(.x, na.rm = TRUE), .names = "sd_{.col}"),
-    
-    # 25th percentile calculation
-    across(all_of(table_vars), ~ quantile(.x, probs = 0.25, na.rm = TRUE), .names = "p25_{.col}"),
-    
-    # 50th percentile (median) calculation
-    across(all_of(table_vars), ~ quantile(.x, probs = 0.50, na.rm = TRUE), .names = "p50_{.col}"),
-    
-    # 75th percentile calculation
-    across(all_of(table_vars), ~ quantile(.x, probs = 0.75, na.rm = TRUE), .names = "p75_{.col}")
-  ) %>%
-  
-  # Reshape variables from columns to rows
-  pivot_longer(
[…]

### 2024-10-18 aefp/aefp_geomarket.qmd
revising front end

---
title: "Structuring College Access: The Market Segment Model and College Board Geo-markets"
#author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
bibliography: ./../assets/bib/cb_geomarket_bib.bib
csl: ./../assets/bib/apa.csl
format:
  pdf:
    documentclass: article
    number-sections: true
    number-depth: 2
    #extensions: [andrewheiss/wordcount]
    fig_caption: true
    keep_tex: true  # Optional: Keeps the intermediate .tex file for debugging
editor: visual
geometry: "margin=1in"
fontsize: 11pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \cfoot{\thepage}
  - \fancypagestyle{plain}{
      \fancyhf{}
      \cfoot{\thepage}
    }
  - \usepackage{booktabs}
  - \usepackage{multirow}
  - \usepackage{array}
  - \usepackage{colortbl}
  - \usepackage{longtable}
  - \usepackage{caption}
  - \usepackage{pdflscape} 
  - \usepackage{rotating}
  - \usepackage{float}       # Required for [ht] specifier
  - \usepackage{graphicx}    # Required for \resizebox  
  - |
    \captionsetup{
      justification=raggedright,
      singlelinecheck=false
    }
  - \setlength{\aboverulesep}{0pt}
  - \setlength{\belowrulesep}{0pt}
  - \setlength{\extrarowheight}{0pt}
  - \setlength{\cmidrulekern}{0pt}
  - \renewcommand{\arraystretch}{1.0}
---

```{r setup, include=FALSE}
library(knitr)
library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)
library(tidyverse)
#hey ho This many words: {{< words-total >}}

```

```{r load_objects, include = FALSE}
getwd()
#load(file = file.path('.','..','..','cb_geomarket_shape','list_data','list_sf.RData'))

socal_eps_table <- readRDS(file.path('.','..','results','tables', "socal_eps_table.rds"))
philly_eps_table <- readRDS(file.path('.','..','results','tables', "philly_eps_table.rds"))
chi_eps_table <- readRDS(file.path('.','..','results','tables', "chi_eps_table.rds"))
```

```{r create_table_functions, include = FALSE}

format_eps_table <- function(df, 
                             years = c(1980, 2000, 2020), 
                             metro_area, 
                             format = "latex", 
                             label = NULL) {

  # Load necessary libraries
  library(dplyr)
  library(kableExtra)

  # Validate the provided years
  valid_years <- c(1980, 2000, 2020)
  if (!all(years %in% valid_years)) {
    stop("Invalid year(s) provided. Only 1980, 2000, and 2020 are allowed.")
  }

  # Define column names for the table (excluding "Variable" and "EPS Name")
  col_names <- c(
    "Geomarket name", 
    rep(c("Mean", "SD", "P25", "P50", "P75"), times = length(years))
  )

  # Build a vector of selected columns based on the years provided
  selected_columns <- c("variable", "eps_codename")
  for (year in years) {
    selected_columns <- c(
      selected_columns, 
      paste0("mean_", year), 
      paste0("sd_", year), 
      paste0("p25_", year), 
      paste0("p50_", year), 
      paste0("p75_", year)
    )
  }

  # Check if all selected columns exist in df
  missing_columns <- setdiff(selected_columns, colnames(df))
  if (length(missing_columns) > 0) {
    stop(paste("The following required columns are missing in the data frame:", 
               paste(missing_columns, collapse = ", ")))
  }

  # Select only the relevant columns from the dataframe and add row number for grouping
  df_selected <- df %>%
    select(all_of(selected_columns)) %>%
    mutate(
      eps_codename = gsub("([%#&_{}])", "\\\\\\1", eps_codename), # Escape special characters in Geomarket names
      row_num = row_number()
    )

  # Create headers for the years, merging cells appropriately
  year_header <- c(" " = 1)  # First column is "Geomarket name"
  for (year in years) {
    year_header <- c(year_header, setNames(5, as.character(year)))  # Each year spans 5 columns
  }

  # Create alignment vector, adding a vertical line after the first column
  align <- c("l|", rep("r", length(col_names) - 1))

  # Prepare group indices for grouping rows in the table
  group_indices <- df_selected %>%
    group_by(variable) %>%
    summarize(
      start_row = min(row_num),
      end_row = max(row_num)
    ) %>%
    ungroup()

  # Remove 'variable' and 'row_num' columns before creating the table
  df_kable <- df_selected %>%
    select(-variable, -row_num)

  # Construct the caption with the LaTeX label if provided
  if (!is.null(label)) {
    # Use double backslash to ensure proper LaTeX command
    caption_text <- paste0(
      "Racial/ethnic and socioeconomic characteristics of Geomarkets around ", 
      metro_area, 
      " \\label{tab:", label, "}"
    )
  } else {
    caption_text <- paste0(
      "Racial/ethnic and socioeconomic characteristics of Geomarkets around ", 
      metro_area
    )
  }

  # Generate the table using kable
  table <- df_kable %>%
    kbl(
      format = format,
      digits = 1,
      col.names = col_names,
      align = align,
      caption = caption_text,
      booktabs = TRUE,
      longtable = TRUE,
      escape = FALSE  # Prevent LaTeX from escaping the caption content
    ) %>%
    # Add a header row for the years
    add_header_above(year_header, bold = TRUE, escape = FALSE) %>%
    # Style the table with repeating header
    kable_styling(
      full_width = FALSE,
      position = "left",
      font_size = 10,
      latex_options = c("hold_position", "repeat_header", "striped")
    ) %>%
    # Specify that the column names row is bold
    row_spec(0, bold = TRUE)

  # Apply group_rows to add group labels and horizontal lines using kableExtra
  for (i in seq_len(nrow(group_indices))) {
    table <- table %>%
      kableExtra::group_rows(
        group_label = group_indices$variable[i],
        start = group_indices$start_row[i],
        end = group_indices$end_row[i]
      )
  }

  return(table)
}

```

\thispagestyle{empty} 
\newpage 
\setcounter{page}{1}

# PROPOSAL

**BACKGROUND, RESEARCH QUESTIONS.** In 1983, Zemsky and Oedel authored *The Structure of College Choice*. Based on an analysis of 1980 SAT score-sending behavior and published by *The College Board*, @RN4982 developed the Market Segment Model and created "Geomarkets." The Market Segment Model predicts how demand for a particular institution varies across geographic regions based on household/student characteristics of geographic regions.

Geomarkets are geographic borders that disaggregate states and large metropolitan areas into "community-based enrollment markets" [@RN4982, p. 14] for recruiting purposes. @fig-nyc-geomarkets shows New York City area Geomarkets. Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today. EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.^[In 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS [@RN5021].]

Geomarkets were incorporated into the _College Board_ "student list" product named Student Search Service. Student lists have been the primary source of lead generation in U.S. higher education for over 50 years [@belkin2019-studata; @list_biz].^[@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.] Lists contain contact information for prospective students. The Student Search Service database consists of College Board test-takers. Colleges pay a fee for each prospect (e.g., \$0.50). Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc. Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets. An analysis of 830 student lists purchased by 14 public universities found that 16% of purchases filtered on Geomarket [@RN5012].

We argue that Geomarket borders are substantially correlated with class and race. This correlation is partly intentional. @RN4982 viewed student demand as a function of class. @RN4982 sought to create Geomarkets that contained many high SES students willing to attend colleges far from home, for example "CT 3--Fairfield County." When colleges use the Market Segment Model, "the result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice" [@RN4982, p. 44]. Considering the nature of U.S. segregation [@rothstein2017color], Geomarkets may also be correlated with race.

<!--  
we argue that utilizing Geomarket filters to purchase student lists can contribute to socioeconomic and racial inequality on college opportunity.

Using data about past correlations -- here, Geomarket borders created with reference to 1980 SAT score sending behaviors -- to make future recommendations amplifies the effects of historical structural inequalities. 
-->

We draw from "quantification" scholarship in sociology [@RN5007; @RN5005] and data studies [@RN4849; @RN4975]. To the extent that Geomarket borders are correlated with class and race, student list purchases that filter on Geomarkets can yield systematic socioeconomic and racial exclusion in college opportunity [@RN4887]. We address two research questions:

1. What is the socioeconomic and racial variation between Geomarkets in metropolitan areas? How does this variation change over time? 
1. How does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?

Scholarship has not observed the mechanisms by which enrollment management products structure college access, funneling certain students to certain institutions. Showing how these products facilitate exclusion introduces new avenues for scholarship and policy discourse.

**DATA SOURCES**. To answer RQ1, we utilize census tract-level data from the 1980 Decennial Census, 2000 Decennial Census, and 2020 5-year American Community Survey (ACS). Variables of interest include: race/ethnicity; mean and median household income; percent of households below poverty line; and BA attainment of people aged 25+. We obtained shapefiles for College Board Geomarkets from a 2012 *R-bloggers* post. We assigned census tracts to Geomarkets by implementing a partial spatial join.^[Census tracts that intersected multiple Geomarkets were partially assigned based on the proportion of land area. Currently, our spatial join for the 2020 ACS is causing some invalid geometries that we will fix prior to AEFP.]

To answer RQ2, we issued public records requests about student lists purchased by 14 public universities from 2016 through 2020. For 414 College Board student list purchases, associated with about 2.6 million prospects, we received both (1) the order summary -- showing which search filters were utilized -- and (2) the de-identified prospect-level data, including race, ethnicity, zip code, and high school. @fig-asu-order-448922 shows the partial order summary, which we converted to tabular data, for a purchase by Arizona State University that filtered on the 2020 high school graduating class, PSAT scores of 1070 to 1180, and selected Geomarkets.^[Analyses for RQ2 additionally utilize data on public and private high schools from NCES.]

**METHOD**. The methods are simultaneously descriptive and spatial. We utilize a case study design in which metropolitan areas are cases. Like @RN4759, we envision focusing on 2-4 metropolitan areas, including the historical context of segregation and gentrification. An appendix will show results for a larger number of metropolitan areas.

To answer RQ1, we produce tables/graphs that show how Geomarkets around a metropolitan area vary on racial and socioeconomic characteristics and how they vary over time. We produce interactive maps at the census tract-level to show more granular variation within and between Geomarkets. We address RQ2 by analyzing actual student list purchases that encompass all Geomarkets around a given metropolitan area, showing which prospects would have been excluded/excluded had the purchase filtered on particular Geomarkets.

**PRELIMINARY FINDINGS**. We are currently writing functions that create tables and/or maps for the metropolitan area chosen by the function call.

For RQ1, Table \ref{tab:chi_eps_table_summary} shows racial and socioeconomic characteristics for Geomarkets around Chicago. @fig-socal-hispanic is a screenshot of an interactive choropleth map of Los Angeles, showing Geomarket borders and the tract-level %Hispanic in 2020. @fig-socal-med-inc shows tract-level median household income around Los Angeles.

Results for RQ2 show how many purchased prospects from each racial/ethnic group live in each Geomarket. These prospects would have been excluded had the purchase filtered on Geomarkets and not selected that Geomarket. @tbl-philly-sim-count, @tbl-philly-sim-col-pct, and @tbl-philly-sim-row-pct, respectively, show counts, column percentages, and row percentages for the Philadelphia metropolitan area, based on a purchase that filters on students from the 2020 high school class who scored between 1070 and 1180 on the PSAT. @tbl-philly-sim-col-pct shows that "PA5--Philadelphia County" contains 15% of all purchased prospect profiles but 44% of Black prospects, indicating that opportunity for Black students is sensitive to the inclusion of this Geomarket. @fig-philly-black-prospects is a screenshot of a (very) preliminary interactive map for this purchase, with marker size indicating the number of purchased prospect profiles who attended that high school and separate markers by race/ethnicity.
<!--  
The size of the high school marker is a function of the number of prospects purchased and map users wil be able to select the race/ethnicity of purchased prospects as well as census variables that determine the chloropleth map.
-->

\newpage

# References

::: {#refs}
:::

\newpage
\setcounter{page}{1}
\renewcommand{\thepage}{A-\arabic{page}}

# Appendix

```{r socal_eps_table_summary, echo=FALSE, results='asis', eval = FALSE}
# Generate the Los Angeles (socal) table with Bookdown label
socal_table <- format_eps_table(
  df = socal_eps_table, 
  years = c(2020), 
  metro_area = "Los Angeles", 
  format = "latex",
  label = "socal_eps_table_summary"  # This label matches the reference in your text
)

# Output the table directly
cat(socal_table)
```

```{r philly_eps_table_summary, echo=FALSE, results='asis', eval = FALSE}
# Wrap the table in the landscape environment
cat("\\begin{landscape}\n")

# Create the table for Philadelphia
philly_table <- format_eps_table(
  df = philly_eps_table, 
  years = c(2000, 2020), 
  metro_area = "Philadelphia metropolitan area", 
  format = "latex",
  label = "philly_eps_table_summary"  # This label matches the reference in your text  
)

# Output the tables directly
cat(philly_table)
cat("\n\\end{landscape}")
```

```{r chi_eps_table_summary, echo=FALSE, results='asis'}
# Wrap the table in the landscape environment
cat("\\begin{landscape}\n")

# Create the table for Philadelphia
chi_table <- format_eps_table(
  df = chi_eps_table, 
  years = c(2000, 2020), 
  metro_area = "Chicago-land", 
  format = "latex",
  label = "chi_eps_table_summary"  # This label matches the reference in your text  
)

# Output the tables directly
cat(chi_table)
cat("\n\\end{landscape}")
```

\newpage

```{r create_eps_sim_tables, echo = FALSE, results='asis', message = FALSE, warning = FALSE}
# Load required packages and input data
load(file = file.path('.', '..', 'results', 'tables', 'philly_asu_eps_lists.RData'))
# str(philly_order448440_psat1270_1520)
# str(philly_order448922_psat1070_1180)

library(dplyr)
library(knitr)
library(kableExtra)
library(scales)  # For adding commas to large numbers

# Function to format three separate tables
format_three_tables <- function(input_list) {
  # Extract individual data frames from the input list
  count_table <- input_list$count_table %>% 
    rename(all = stu_all, race_known = stu_race_known, white = stu_white, asian = stu_asian, black = stu_black, hispanic = stu_hispanic, two_races = stu_tworaces) %>%
    mutate(across(where(is.numeric), ~ if_else(. > 1000, comma(.), as.character(.))))  # Add commas to large numbers

  # Modify race_known to be a column percentage of the race_known total
  col_pct_table <- input_list$col_pct_table %>%
    rename(all = stu_all, race_known = stu_race_known, white = col_pct_white, asian = col_pct_asian, black = col_pct_black, hispanic = col_pct_hispanic, two_races = col_pct_tworaces) %>%
    select(-all) %>%
    mutate(
      race_known = (race_known / sum(race_known)) * 100,  # Convert race_known to a column percentage of the total
      across(-eps_codename, ~ round(., 1))  # Round percentage values to one decimal point
    )

  # Set race_known to always be 100% in row_pct_table
  row_pct_table <- input_list$row_pct_table %>%  
    rename(all = stu_all, race_known = stu_race_known, white = row_pct_white, asian = row_pct_asian, black = row_pct_black, hispanic = row_pct_hispanic, two_races = row_pct_tworaces) %>%
    select(-all) %>%
    mutate(
      race_known = 100.0,  # Set race_known to 100%
      across(-eps_codename, ~ round(., 1))  # Round percentage values to one decimal point
    )

  # Format and display count table with caption
  count_table_output <- count_table %>%
    kbl(format = "latex", booktabs = TRUE, align = c("l", rep("r", ncol(count_table) - 1))) %>%
    kable_styling(latex_options = c("striped", "hold_position"), position = "left", font_size = 10) %>%
    row_spec(0, bold = TRUE, background = "#D3D3D3")

  # Format and display column percentage table with caption
  col_pct_table_output <- col_pct_table %>%
    kbl(format = "latex", booktabs = TRUE, align = c("l", rep("r", ncol(col_pct_table) - 1))) %>%
    kable_styling(latex_options = c("striped", "hold_position"), position = "left", font_size = 10) %>%
    row_spec(0, bold = TRUE, background = "#D3D3D3")

  # Format and display row percentage table with caption
  row_pct_table_output <- row_pct_table %>%
    kbl(format = "latex", booktabs = TRUE, align = c("l", rep("r", ncol(row_pct_table) - 1))) %>%
    kable_styling(latex_options = c("striped", "hold_position"), position = "left", font_size = 10) %>%
    row_spec(0, bold = TRUE, background = "#D3D3D3")

  # Return formatted tables
  list(count_table_output, col_pct_table_output, row_pct_table_output)
}

# Example usage
formatted_tables <- format_three_tables(philly_order448922_psat1070_1180)

# Print formatted tables with landscape mode in LaTeX
#cat("\\begin{landscape}\n")
#cat(as.character(formatted_tables[[1]]))
#cat("\n\n")
#cat(as.character(formatted_tables[[2]]))
#cat("\n\n")
#cat(as.character(formatted_tables[[3]]))
#cat("\n\\end{landscape}")

```

\newpage

```{r philly-sim-count, echo = FALSE, results='asis'}
#| label: tbl-philly-sim-count
#| tbl-cap: 'Count of Philadelphia area students included in student list purchase that filtered on 2020 high school class and PSAT scores 1070 -- 1180 (Not shown: American Indian/Alaska Native (AI/AN); and Native Hawaiian and other Pacific Islander (NHPI))'

cat(as.character(formatted_tables[[1]]))
```

```{r philly-sim-col-pct, echo = FALSE, results='asis'}
#| label: tbl-philly-sim-col-pct
#| tbl-cap: 'Percent of total prospects purchased from each geomarket, Philadelphia area students included in student list purchase that filtered on 2020 high school class and PSAT scores 1070 -- 1180 (Not shown: AI/AN, NHPI)'

cat(as.character(formatted_tables[[2]]))
```

```{r philly-sim-row-pct, echo = FALSE, results='asis'}
#| label: tbl-philly-sim-row-pct
#| tbl-cap: 'Racial composition of purchased prospects from each Geomarket, Philadelphia area students included in student list purchase that filtered on 2020 high school class and PSAT scores 1070 -- 1180 (Not shown: AI/AN, NHPI)'

cat(as.character(formatted_tables[[3]]))
```

\newpage

```{r fig-nyc-geomarkets, echo=FALSE, fig.cap="Geomarkets in New York City Area, NY state", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
knitr::include_graphics(path = file.path('.','..','results','figures','nyc_geomarkets.png'))
```

\newpage
```{r fig-asu-order-448922, echo=FALSE, fig.cap="Partial order summary from a College Board student list purchase", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
knitr::include_graphics(path = file.path('.','..','results','figures','asu_order_448922_Redacted.pdf'))
```

\newpage
```{r fig-socal-hispanic, echo=FALSE, fig.cap="Screenshot of Los Angeles area Geomarkets, showing 2020 percent Hispanic by census tract", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
knitr::include_graphics(path = file.path('.','..','results','figures','socal_hispanic.png'))
```

\newpage

```{r fig-socal-med-inc, echo=FALSE, fig.cap="Screenshot of Los Angeles area Geomarkets, showing 2020 median household income by census tract", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
knitr::include_graphics(path = file.path('.','..','results','figures','socal_med_inc.png'))
```

\newpage

```{r fig-philly-black-prospects, echo=FALSE, fig.cap="Screenshot of Philadelphia area Geomarkets, markers are high schools and marker size a function of number of Black student profiles purchased (2020 HS class, PSAT scores 1070 -- 1180)", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
knitr::include_graphics(path = file.path('.','..','results','figures','philly_black_prospects.png'))
```

### 2024-10-18 aefp/results_test.qmd
revising front end

---
title: "Analysis Results"
output: 
  pdf_document:
    keep_tex: true
header-includes:
  - \usepackage{booktabs}
  - \usepackage{colortbl}
  - \usepackage[table]{xcolor}
  - \usepackage{array} # For custom column types if needed
---

## Results of the Analysis

Here we present the summarized data for EPS variables across different years.

```{r table-inclusion, echo=FALSE, results='asis'}
#cat(readLines("eps_table.tex"), sep = "\n")
getwd()
#cat(readLines(file.path('.','..','results','tables',"socal_table.tex")), sep = "\n")
#\input{../results/tables/socal_table.tex}
```
\input{../results/tables/socal_table.tex}

### 2024-10-18 results/tables/socal_table.tex
revising front end

@@
-\begin{table}[!h]
-\centering
-\begin{tabular}[t]{l|>{}l||r|r|r|r|r|r|r|r|r|r}
-\hline
-\multicolumn{2}{c|}{ } & \multicolumn{5}{c|}{2000} & \multicolumn{5}{c}{2020} \\
-\cline{3-7} \cline{8-12}
-\textbf{Variable} & \textbf{EPS Name} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75}\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{40.4} & \cellcolor{gray!6}{21.0} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{36.8} & \cellcolor{gray!6}{56.2} & \cellcolor{gray!6}{26.1} & \cellcolor{gray!6}{17.8} & \cellcolor{gray!6}{12.6} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{36.9}\\
-\hline
-\hline
-\% White, non-Hispanic & Covina \& West Covina & 32.8 & 23.0 & 11.9 & 27.9 & 49.3 & 18.8 & 16.8 & 5.4 & 12.5 & 27.1\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{10.3} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{6.2} & \cellcolor{gray!6}{11.8} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{8.4}\\
-\hline
-\% White, non-Hispanic & Glendale and Pasadena & 39.9 & 24.0 & 16.0 & 40.4 & 61.3 & 35.3 & 24.0 & 12.8 & 34.1 & 57.2\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{26.5} & \cellcolor{gray!6}{28.5} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{40.8} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{25.7} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{22.6} & \cellcolor{gray!6}{53.5}\\
-\hline
-\% White, non-Hispanic & Long Beach & 35.1 & 28.9 & 10.0 & 21.8 & 64.0 & 25.8 & 23.0 & 6.6 & 16.4 & 45.5\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{45.2} & \cellcolor{gray!6}{21.8} & \cellcolor{gray!6}{28.1} & \cellcolor{gray!6}{43.6} & \cellcolor{gray!6}{63.7} & \cellcolor{gray!6}{27.7} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{12.5} & \cellcolor{gray!6}{23.5} & \cellcolor{gray!6}{40.7}\\
-\hline
-\% White, non-Hispanic & San Fernando Valley (East) & 45.1 & 28.1 & 21.4 & 40.8 & 74.2 & 43.3 & 25.7 & 22.5 & 39.6 & 67.0\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{58.3} & \cellcolor{gray!6}{25.6} & \cellcolor{gray!6}{40.4} & \cellcolor{gray!6}{67.0} & \cellcolor{gray!6}{78.8} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{24.2} & \cellcolor{gray!6}{22.8} & \cellcolor{gray!6}{46.6} & \cellcolor{gray!6}{64.7}\\
-\hline
-\% White, non-Hispanic & Santa Ana & 52.4 & 29.0 & 24.5 & 58.3 & 77.9 & 42.0 & 26.3 & 17.0 & 44.1 & 64.6\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{7.3} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{5.5} & \cellcolor{gray!6}{6.3} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{6.6}\\
-\hline
-\% White, non-Hispanic & South Bay & 41.3 & 29.4 & 12.4 & 41.3 & 68.3 & 29.1 & 24.5 & 7.1 & 22.1 & 51.4\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{66.0} & \cellcolor{gray!6}{18.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{70.4} & \cellcolor{gray!6}{78.1} & \cellcolor{gray!6}{52.9} & \cellcolor{gray!6}{18.1} & \cellcolor{gray!6}{40.7} & \cellcolor{gray!6}{55.0} & \cellcolor{gray!6}{66.3}\\
-\hline
-\% White, non-Hispanic & West Los Angeles \& West Beach & 63.9 & 24.4 & 43.0 & 72.9 & 84.0 & 60.2 & 20.2 & 45.9 & 62.9 & 75.5\\
-\hline
-\cellcolor{gray!6}{\% White, non-Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{35.4} & \cellcolor{gray!6}{20.5} & \cellcolor{gray!6}{18.5} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{50.7} & \cellcolor{gray!6}{22.4} & \cellcolor{gray!6}{15.3} & \cellcolor{gray!6}{10.9} & \cellcolor{gray!6}{17.5} & \cellcolor{gray!6}{32.7}\\
-\hline
-\% Black, non-Hispanic & Anaheim & 3.9 & 5.3 & 1.3 & 2.3 & 4.4 & 4.4 & 5.6 & 0.8 & 2.3 & 6.0\\
-\hline
-\hline
-\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{5.2} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{4.7}\\
-\hline
-\% Black, non-Hispanic & East Los Angeles & 1.1 & 2.8 & 0.2 & 0.5 & 1.0 & 1.3 & 2.8 & 0.0 & 0.4 & 1.4\\
-\hline
-\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{4.4} & \cellcolor{gray!6}{8.5} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{1.7} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{3.4} & \cellcolor{gray!6}{5.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{1.8} & \cellcolor{gray!6}{4.1}\\
-\hline
-\% Black, non-Hispanic & Hollywood \& Wilshire & 13.0 & 21.0 & 2.0 & 3.3 & 8.5 & 11.1 & 16.8 & 1.9 & 4.5 & 10.5\\
-\hline
-\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{13.1} & \cellcolor{gray!6}{14.4} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{8.9} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{3.6} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{15.1}\\
-\hline
-\% Black, non-Hispanic & Riverside, San Bernardino, \& Ontario & 8.2 & 7.0 & 3.2 & 5.8 & 11.4 & 7.1 & 6.4 & 2.4 & 5.4 & 9.8\\
-\hline
-\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{5.3} & \cellcolor{gray!6}{4.3} & \cellcolor{gray!6}{3.8} & \cellcolor{gray!6}{1.6} & \cellcolor{gray!6}{3.3} & \cellcolor{gray!6}{5.7}\\
-\hline
-\% Black, non-Hispanic & San Fernando Valley (West) & 3.2 & 3.1 & 1.2 & 2.5 & 3.8 & 3.8 & 3.9 & 1.1 & 2.7 & 5.4\\
-\hline
-\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{1.8} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.9} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{1.3} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.7} & \cellcolor{gray!6}{1.6}\\
-\hline
-\% Black, non-Hispanic & South \& South Central Los Angeles & 24.7 & 24.4 & 2.6 & 17.7 & 39.6 & 17.5 & 18.8 & 2.7 & 11.6 & 25.3\\
-\hline
-\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{13.2} & \cellcolor{gray!6}{20.2} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{4.1} & \cellcolor{gray!6}{15.2} & \cellcolor{gray!6}{12.1} & \cellcolor{gray!6}{16.9} & \cellcolor{gray!6}{1.7} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{14.3}\\
-\hline
-\% Black, non-Hispanic & South Orange Co & 1.4 & 1.2 & 0.7 & 1.2 & 1.8 & 1.5 & 1.8 & 0.3 & 1.0 & 2.3\\
-\hline
-\cellcolor{gray!6}{\% Black, non-Hispanic} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{6.3} & \cellcolor{gray!6}{11.7} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{6.2}\\
-\hline
-\% Black, non-Hispanic & Whittier \& North Orange Co & 2.2 & 2.3 & 0.9 & 1.5 & 2.5 & 2.4 & 2.7 & 0.4 & 1.5 & 3.0\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{34.0} & \cellcolor{gray!6}{22.0} & \cellcolor{gray!6}{16.3} & \cellcolor{gray!6}{29.7} & \cellcolor{gray!6}{49.1} & \cellcolor{gray!6}{41.3} & \cellcolor{gray!6}{22.7} & \cellcolor{gray!6}{23.1} & \cellcolor{gray!6}{38.6} & \cellcolor{gray!6}{58.1}\\
-\hline
-\hline
-\% Hispanic & Covina \& West Covina & 43.6 & 25.6 & 20.6 & 39.6 & 68.7 & 51.7 & 23.5 & 30.8 & 54.3 & 71.3\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{66.6} & \cellcolor{gray!6}{27.9} & \cellcolor{gray!6}{41.6} & \cellcolor{gray!6}{77.9} & \cellcolor{gray!6}{91.0} & \cellcolor{gray!6}{64.4} & \cellcolor{gray!6}{28.9} & \cellcolor{gray!6}{36.7} & \cellcolor{gray!6}{73.2} & \cellcolor{gray!6}{91.5}\\
-\hline
-\% Hispanic & Glendale and Pasadena & 34.4 & 25.8 & 13.8 & 26.3 & 53.5 & 36.8 & 26.9 & 16.3 & 26.3 & 53.4\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{43.7} & \cellcolor{gray!6}{28.3} & \cellcolor{gray!6}{15.8} & \cellcolor{gray!6}{44.4} & \cellcolor{gray!6}{68.5} & \cellcolor{gray!6}{38.2} & \cellcolor{gray!6}{24.6} & \cellcolor{gray!6}{15.7} & \cellcolor{gray!6}{34.7} & \cellcolor{gray!6}{57.1}\\
-\hline
-\% Hispanic & Long Beach & 37.1 & 24.3 & 16.3 & 32.1 & 53.1 & 45.8 & 23.3 & 26.6 & 43.5 & 63.4\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{38.3} & \cellcolor{gray!6}{20.5} & \cellcolor{gray!6}{21.7} & \cellcolor{gray!6}{34.2} & \cellcolor{gray!6}{52.9} & \cellcolor{gray!6}{53.8} & \cellcolor{gray!6}{22.0} & \cellcolor{gray!6}{36.3} & \cellcolor{gray!6}{54.1} & \cellcolor{gray!6}{72.4}\\
-\hline
-\% Hispanic & San Fernando Valley (East) & 40.2 & 27.5 & 11.3 & 35.8 & 62.5 & 40.6 & 27.2 & 14.8 & 37.3 & 60.9\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{27.4} & \cellcolor{gray!6}{25.0} & \cellcolor{gray!6}{9.4} & \cellcolor{gray!6}{16.4} & \cellcolor{gray!6}{38.8} & \cellcolor{gray!6}{36.8} & \cellcolor{gray!6}{25.9} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{28.8} & \cellcolor{gray!6}{56.9}\\
-\hline
-\% Hispanic & Santa Ana & 29.6 & 29.2 & 8.1 & 15.0 & 43.0 & 33.0 & 27.9 & 12.3 & 21.2 & 46.6\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{63.6} & \cellcolor{gray!6}{26.9} & \cellcolor{gray!6}{46.5} & \cellcolor{gray!6}{65.9} & \cellcolor{gray!6}{88.3} & \cellcolor{gray!6}{70.5} & \cellcolor{gray!6}{24.8} & \cellcolor{gray!6}{60.2} & \cellcolor{gray!6}{75.7} & \cellcolor{gray!6}{90.2}\\
-\hline
-\% Hispanic & South Bay & 25.6 & 20.6 & 8.8 & 19.8 & 39.0 & 33.2 & 22.5 & 12.9 & 29.8 & 51.8\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{18.0} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{7.6} & \cellcolor{gray!6}{11.6} & \cellcolor{gray!6}{21.3} & \cellcolor{gray!6}{22.2} & \cellcolor{gray!6}{17.0} & \cellcolor{gray!6}{10.4} & \cellcolor{gray!6}{16.8} & \cellcolor{gray!6}{27.2}\\
-\hline
-\% Hispanic & West Los Angeles \& West Beach & 16.2 & 16.6 & 4.6 & 9.4 & 21.1 & 15.8 & 13.4 & 6.3 & 12.2 & 20.4\\
-\hline
-\cellcolor{gray!6}{\% Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{47.5} & \cellcolor{gray!6}{24.9} & \cellcolor{gray!6}{27.2} & \cellcolor{gray!6}{46.7} & \cellcolor{gray!6}{67.1} & \cellcolor{gray!6}{56.0} & \cellcolor{gray!6}{23.6} & \cellcolor{gray!6}{39.3} & \cellcolor{gray!6}{59.2} & \cellcolor{gray!6}{75.7}\\
-\hline
-\% API, non-Hispanic & Anaheim & 18.8 & 15.7 & 8.1 & 14.3 & 23.2 & 24.5 & 18.1 & 11.2 & 19.4 & 34.2\\
-\hline
-\hline
-\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{5.8} & \cellcolor{gray!6}{9.7} & \cellcolor{gray!6}{21.9} & \cellcolor{gray!6}{23.7} & \cellcolor{gray!6}{19.3} & \cellcolor{gray!6}{10.0} & \cellcolor{gray!6}{16.7} & \cellcolor{gray!6}{31.9}\\
-\hline
-\% API, non-Hispanic & East Los Angeles & 20.5 & 21.5 & 1.8 & 11.2 & 37.9 & 26.5 & 26.0 & 2.7 & 16.4 & 47.9\\
-\hline
-\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{17.4} & \cellcolor{gray!6}{15.5} & \cellcolor{gray!6}{6.2} & \cellcolor{gray!6}{11.6} & \cellcolor{gray!6}{23.1} & \cellcolor{gray!6}{20.9} & \cellcolor{gray!6}{19.8} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{27.3}\\
-\hline
-\% API, non-Hispanic & Hollywood \& Wilshire & 13.8 & 12.9 & 4.6 & 9.5 & 18.9 & 15.7 & 14.0 & 5.7 & 11.4 & 20.1\\
-\hline
-\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{10.1} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{15.3} & \cellcolor{gray!6}{13.2} & \cellcolor{gray!6}{11.3} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{18.2}\\
-\hline
-\% API, non-Hispanic & Riverside, San Bernardino, \& Ontario & 5.2 & 6.5 & 1.8 & 3.5 & 6.0 & 8.2 & 10.0 & 2.0 & 5.0 & 10.4\\
-\hline
-\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{7.2} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{4.6} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{7.7} & \cellcolor{gray!6}{8.3} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{4.7} & \cellcolor{gray!6}{6.9} & \cellcolor{gray!6}{10.2}\\
-\hline
-\% API, non-Hispanic & San Fernando Valley (West) & 8.0 & 5.9 & 4.2 & 6.5 & 10.8 & 11.6 & 8.1 & 5.3 & 10.4 & 16.3\\
-\hline
-\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{14.1} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{4.5} & \cellcolor{gray!6}{9.2} & \cellcolor{gray!6}{18.1} & \cellcolor{gray!6}{20.1} & \cellcolor{gray!6}{18.3} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{14.3} & \cellcolor{gray!6}{26.5}\\
-\hline
-\% API, non-Hispanic & South \& South Central Los Angeles & 2.9 & 5.3 & 0.3 & 0.8 & 2.7 & 3.9 & 6.9 & 0.2 & 1.2 & 4.4\\
-\hline
-\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{16.6} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{6.1} & \cellcolor{gray!6}{11.0} & \cellcolor{gray!6}{26.8} & \cellcolor{gray!6}{21.2} & \cellcolor{gray!6}{15.9} & \cellcolor{gray!6}{8.2} & \cellcolor{gray!6}{17.8} & \cellcolor{gray!6}{30.3}\\
-\hline
-\% API, non-Hispanic & South Orange Co & 11.1 & 9.8 & 5.3 & 8.5 & 13.6 & 18.8 & 15.1 & 9.0 & 14.1 & 23.6\\
-\hline
-\cellcolor{gray!6}{\% API, non-Hispanic} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{9.9} & \cellcolor{gray!6}{7.9} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{7.5} & \cellcolor{gray!6}{11.7} & \cellcolor{gray!6}{13.1} & \cellcolor{gray!6}{9.5} & \cellcolor{gray!6}{6.5} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{17.6}\\
-\hline
-\% API, non-Hispanic & Whittier \& North Orange Co & 12.6 & 15.4 & 3.6 & 8.2 & 12.9 & 15.9 & 16.2 & 5.2 & 12.0 & 18.0\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.2}\\
-\hline
-\hline
-\% Native, non-Hispanic & Covina \& West Covina & 0.3 & 0.2 & 0.2 & 0.3 & 0.4 & 0.2 & 0.7 & 0.0 & 0.0 & 0.2\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.2}\\
-\hline
-\% Native, non-Hispanic & Glendale and Pasadena & 0.3 & 0.2 & 0.1 & 0.2 & 0.3 & 0.2 & 0.5 & 0.0 & 0.0 & 0.1\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
-\hline
-\% Native, non-Hispanic & Long Beach & 0.4 & 0.3 & 0.2 & 0.3 & 0.5 & 0.6 & 6.9 & 0.0 & 0.0 & 0.1\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.6} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.3}\\
-\hline
-\% Native, non-Hispanic & San Fernando Valley (East) & 0.2 & 0.1 & 0.1 & 0.2 & 0.3 & 0.1 & 0.3 & 0.0 & 0.0 & 0.0\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.5} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.1}\\
-\hline
-\% Native, non-Hispanic & Santa Ana & 0.3 & 0.2 & 0.2 & 0.3 & 0.4 & 0.2 & 0.4 & 0.0 & 0.0 & 0.2\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.1} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
-\hline
-\% Native, non-Hispanic & South Bay & 0.2 & 0.1 & 0.1 & 0.2 & 0.3 & 0.2 & 0.5 & 0.0 & 0.0 & 0.0\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0}\\
-\hline
-\% Native, non-Hispanic & West Los Angeles \& West Beach & 0.2 & 0.2 & 0.1 & 0.2 & 0.3 & 0.2 & 0.4 & 0.0 & 0.0 & 0.0\\
-\hline
-\cellcolor{gray!6}{\% Native, non-Hispanic} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.2} & \cellcolor{gray!6}{0.3} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{7.7} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.0} & \cellcolor{gray!6}{0.3}\\
-\hline
-pct\_nhisp\_multi & Anaheim & 2.4 & 0.9 & 1.8 & 2.4 & 3.0 & 2.9 & 2.3 & 1.1 & 2.1 & 4.5\\
-\hline
-\hline
-\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{0.8} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{0.4} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{2.9}\\
-\hline
-pct\_nhisp\_multi & East Los Angeles & 1.1 & 0.9 & 0.4 & 0.9 & 1.7 & 0.9 & 1.2 & 0.0 & 0.4 & 1.4\\
-\hline
-\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{1.9} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{4.2}\\
-\hline
-pct\_nhisp\_multi & Hollywood \& Wilshire & 2.6 & 2.0 & 1.3 & 2.1 & 3.1 & 3.2 & 2.9 & 1.3 & 2.4 & 4.6\\
-\hline
-\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{3.3} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{2.3} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{2.5} & \cellcolor{gray!6}{4.2}\\
-\hline
-pct\_nhisp\_multi & Riverside, San Bernardino, \& Ontario & 2.4 & 0.9 & 1.8 & 2.3 & 3.0 & 2.6 & 2.4 & 0.7 & 1.9 & 3.7\\
-\hline
-\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{3.0} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{4.4}\\
-\hline
-pct\_nhisp\_multi & San Fernando Valley (West) & 2.6 & 1.5 & 1.7 & 2.4 & 3.2 & 3.1 & 2.6 & 1.1 & 2.7 & 4.1\\
-\hline
-\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{2.2} & \cellcolor{gray!6}{1.2} & \cellcolor{gray!6}{1.4} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.1} & \cellcolor{gray!6}{2.6} & \cellcolor{gray!6}{4.9}\\
-\hline
-pct\_nhisp\_multi & South \& South Central Los Angeles & 1.1 & 1.1 & 0.4 & 0.8 & 1.5 & 1.3 & 2.0 & 0.0 & 0.5 & 1.8\\
-\hline
-\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{1.0} & \cellcolor{gray!6}{2.1} & \cellcolor{gray!6}{2.8} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{3.8} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{1.5} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{5.6}\\
-\hline
-pct\_nhisp\_multi & South Orange Co & 3.0 & 6.2 & 1.8 & 2.4 & 3.2 & 4.1 & 2.7 & 2.1 & 3.7 & 5.5\\
-\hline
-\cellcolor{gray!6}{pct\_nhisp\_multi} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{3.2} & \cellcolor{gray!6}{1.3} & \cellcolor{gray!6}{2.4} & \cellcolor{gray!6}{3.1} & \cellcolor{gray!6}{3.9} & \cellcolor{gray!6}{4.8} & \cellcolor{gray!6}{2.9} & \cellcolor{gray!6}{2.7} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{6.7}\\
-\hline
-pct\_nhisp\_multi & Whittier \& North Orange Co & 1.8 & 0.7 & 1.3 & 1.9 & 2.3 & 1.9 & 1.7 & 0.5 & 1.5 & 2.8\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{94.0} & \cellcolor{gray!6}{119.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{41.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{99.0} & \cellcolor{gray!6}{129.0}\\
-\hline
-\hline
-Median income & Covina \& West Covina & 103.0 & 36.0 & 79.0 & 96.0 & 121.0 & 103.0 & 32.0 & 82.0 & 99.0 & 120.0\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{70.0} & \cellcolor{gray!6}{26.0} & \cellcolor{gray!6}{53.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{81.0} & \cellcolor{gray!6}{76.0} & \cellcolor{gray!6}{25.0} & \cellcolor{gray!6}{60.0} & \cellcolor{gray!6}{72.0} & \cellcolor{gray!6}{87.0}\\
-\hline
-Median income & Glendale and Pasadena & 96.0 & 40.0 & 69.0 & 86.0 & 110.0 & 102.0 & 40.0 & 73.0 & 96.0 & 121.0\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{41.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{86.0} & \cellcolor{gray!6}{84.0} & \cellcolor{gray!6}{43.0} & \cellcolor{gray!6}{52.0} & \cellcolor{gray!6}{75.0} & \cellcolor{gray!6}{105.0}\\
-\hline
-Median income & Long Beach & 84.0 & 41.0 & 54.0 & 79.0 & 108.0 & 90.0 & 36.0 & 62.0 & 85.0 & 111.0\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{36.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{81.0} & \cellcolor{gray!6}{111.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{37.0} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{88.0} & \cellcolor{gray!6}{119.0}\\
-\hline
-Median income & San Fernando Valley (East) & 94.0 & 62.0 & 60.0 & 75.0 & 97.0 & 99.0 & 56.0 & 61.0 & 83.0 & 116.0\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{52.0} & \cellcolor{gray!6}{87.0} & \cellcolor{gray!6}{113.0} & \cellcolor{gray!6}{156.0} & \cellcolor{gray!6}{121.0} & \cellcolor{gray!6}{46.0} & \cellcolor{gray!6}{87.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{146.0}\\
-\hline
-Median income & Santa Ana & 116.0 & 44.0 & 86.0 & 107.0 & 141.0 & 119.0 & 44.0 & 87.0 & 112.0 & 142.0\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{59.0} & \cellcolor{gray!6}{23.0} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{55.0} & \cellcolor{gray!6}{68.0} & \cellcolor{gray!6}{67.0} & \cellcolor{gray!6}{27.0} & \cellcolor{gray!6}{50.0} & \cellcolor{gray!6}{62.0} & \cellcolor{gray!6}{78.0}\\
-\hline
-Median income & South Bay & 111.0 & 52.0 & 73.0 & 99.0 & 134.0 & 115.0 & 51.0 & 78.0 & 103.0 & 141.0\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{140.0} & \cellcolor{gray!6}{49.0} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{137.0} & \cellcolor{gray!6}{167.0} & \cellcolor{gray!6}{142.0} & \cellcolor{gray!6}{46.0} & \cellcolor{gray!6}{109.0} & \cellcolor{gray!6}{140.0} & \cellcolor{gray!6}{171.0}\\
-\hline
-Median income & West Los Angeles \& West Beach & 132.0 & 78.0 & 80.0 & 104.0 & 168.0 & 137.0 & 61.0 & 98.0 & 122.0 & 165.0\\
-\hline
-\cellcolor{gray!6}{Median income} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{101.0} & \cellcolor{gray!6}{29.0} & \cellcolor{gray!6}{80.0} & \cellcolor{gray!6}{94.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{105.0} & \cellcolor{gray!6}{30.0} & \cellcolor{gray!6}{85.0} & \cellcolor{gray!6}{101.0} & \cellcolor{gray!6}{121.0}\\
-\hline
-Mean income & Anaheim & 127.0 & 61.0 & 97.0 & 111.0 & 135.0 & 131.0 & 54.0 & 99.0 & 118.0 & 146.0\\
-\hline
-\hline
-\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{126.0} & \cellcolor{gray!6}{44.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{116.0} & \cellcolor{gray!6}{144.0} & \cellcolor{gray!6}{125.0} & \cellcolor{gray!6}{42.0} & \cellcolor{gray!6}{100.0} & \cellcolor{gray!6}{115.0} & \cellcolor{gray!6}{143.0}\\
-\hline
-Mean income & East Los Angeles & 89.0 & 32.0 & 72.0 & 83.0 & 100.0 & 97.0 & 33.0 & 77.0 & 93.0 & 106.0\\
-\hline
-\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{56.0} & \cellcolor{gray!6}{86.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{138.0} & \cellcolor{gray!6}{135.0} & \cellcolor{gray!6}{57.0} & \cellcolor{gray!6}{96.0} & \cellcolor{gray!6}{121.0} & \cellcolor{gray!6}{157.0}\\
-\hline
-Mean income & Hollywood \& Wilshire & 99.0 & 69.0 & 60.0 & 78.0 & 111.0 & 118.0 & 68.0 & 73.0 & 98.0 & 143.0\\
-\hline
-\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{106.0} & \cellcolor{gray!6}{49.0} & \cellcolor{gray!6}{72.0} & \cellcolor{gray!6}{102.0} & \cellcolor{gray!6}{129.0} & \cellcolor{gray!6}{113.0} & \cellcolor{gray!6}{45.0} & \cellcolor{gray!6}{80.0} & \cellcolor{gray!6}{104.0} & \cellcolor{gray!6}{137.0}\\
-\hline
-Mean income & Riverside, San Bernardino, \& Ontario & 107.0 & 41.0 & 79.0 & 99.0 & 128.0 & 114.0 & 42.0 & 83.0 & 103.0 & 137.0\\
-\hline
-\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{105.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{97.0} & \cellcolor{gray!6}{128.0} & \cellcolor{gray!6}{137.0} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{84.0} & \cellcolor{gray!6}{109.0} & \cellcolor{gray!6}{154.0}\\
-\hline
-Mean income & San Fernando Valley (West) & 154.0 & 69.0 & 106.0 & 134.0 & 183.0 & 152.0 & 65.0 & 105.0 & 138.0 & 180.0\\
-\hline
-\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{145.0} & \cellcolor{gray!6}{64.0} & \cellcolor{gray!6}{102.0} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{167.0} & \cellcolor{gray!6}{152.0} & \cellcolor{gray!6}{62.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{139.0} & \cellcolor{gray!6}{174.0}\\
-\hline
-Mean income & South \& South Central Los Angeles & 77.0 & 26.0 & 62.0 & 72.0 & 87.0 & 86.0 & 32.0 & 66.0 & 78.0 & 95.0\\
-\hline
-\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{141.0} & \cellcolor{gray!6}{78.0} & \cellcolor{gray!6}{89.0} & \cellcolor{gray!6}{118.0} & \cellcolor{gray!6}{164.0} & \cellcolor{gray!6}{149.0} & \cellcolor{gray!6}{77.0} & \cellcolor{gray!6}{95.0} & \cellcolor{gray!6}{124.0} & \cellcolor{gray!6}{180.0}\\
-\hline
-Mean income & South Orange Co & 173.0 & 66.0 & 125.0 & 163.0 & 203.0 & 179.0 & 62.0 & 138.0 & 168.0 & 213.0\\
-\hline
-\cellcolor{gray!6}{Mean income} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{197.0} & \cellcolor{gray!6}{136.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{144.0} & \cellcolor{gray!6}{266.0} & \cellcolor{gray!6}{206.0} & \cellcolor{gray!6}{108.0} & \cellcolor{gray!6}{133.0} & \cellcolor{gray!6}{169.0} & \cellcolor{gray!6}{252.0}\\
-\hline
-Mean income & Whittier \& North Orange Co & 121.0 & 46.0 & 96.0 & 110.0 & 135.0 & 125.0 & 37.0 & 102.0 & 117.0 & 142.0\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Anaheim} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{6.9} & \cellcolor{gray!6}{4.6} & \cellcolor{gray!6}{8.0} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{10.7} & \cellcolor{gray!6}{6.7} & \cellcolor{gray!6}{5.5} & \cellcolor{gray!6}{9.6} & \cellcolor{gray!6}{14.8}\\
-\hline
-\hline
-\% in poverty & Covina \& West Covina & 10.3 & 6.7 & 5.2 & 8.8 & 13.5 & 10.3 & 6.4 & 5.8 & 9.0 & 12.8\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{East Los Angeles} & \cellcolor{gray!6}{19.2} & \cellcolor{gray!6}{9.5} & \cellcolor{gray!6}{11.4} & \cellcolor{gray!6}{18.8} & \cellcolor{gray!6}{25.4} & \cellcolor{gray!6}{16.0} & \cellcolor{gray!6}{7.9} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{15.1} & \cellcolor{gray!6}{21.1}\\
-\hline
-\% in poverty & Glendale and Pasadena & 12.2 & 7.4 & 6.5 & 10.7 & 16.9 & 12.6 & 7.7 & 6.4 & 11.3 & 16.8\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Hollywood \& Wilshire} & \cellcolor{gray!6}{22.1} & \cellcolor{gray!6}{12.9} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{20.8} & \cellcolor{gray!6}{31.0} & \cellcolor{gray!6}{17.7} & \cellcolor{gray!6}{11.8} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{14.7} & \cellcolor{gray!6}{23.9}\\
-\hline
-\% in poverty & Long Beach & 16.4 & 13.1 & 5.5 & 12.5 & 25.1 & 14.5 & 12.0 & 6.8 & 12.1 & 19.9\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Riverside, San Bernardino, \& Ontario} & \cellcolor{gray!6}{13.0} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{10.8} & \cellcolor{gray!6}{17.0} & \cellcolor{gray!6}{12.4} & \cellcolor{gray!6}{9.2} & \cellcolor{gray!6}{5.8} & \cellcolor{gray!6}{10.6} & \cellcolor{gray!6}{17.2}\\
-\hline
-\% in poverty & San Fernando Valley (East) & 14.6 & 9.0 & 7.4 & 12.7 & 19.5 & 14.6 & 8.9 & 8.2 & 13.3 & 19.0\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{San Fernando Valley (West)} & \cellcolor{gray!6}{8.2} & \cellcolor{gray!6}{7.0} & \cellcolor{gray!6}{3.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{10.8} & \cellcolor{gray!6}{9.8} & \cellcolor{gray!6}{7.8} & \cellcolor{gray!6}{4.2} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{13.3}\\
-\hline
-\% in poverty & Santa Ana & 8.7 & 7.1 & 3.6 & 6.3 & 12.2 & 9.8 & 6.4 & 5.0 & 8.3 & 13.0\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{South \& South Central Los Angeles} & \cellcolor{gray!6}{26.7} & \cellcolor{gray!6}{12.7} & \cellcolor{gray!6}{17.9} & \cellcolor{gray!6}{26.1} & \cellcolor{gray!6}{35.7} & \cellcolor{gray!6}{21.7} & \cellcolor{gray!6}{11.5} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{20.6} & \cellcolor{gray!6}{28.5}\\
-\hline
-\% in poverty & South Bay & 10.2 & 7.9 & 3.9 & 7.2 & 16.4 & 9.6 & 6.9 & 4.9 & 7.5 & 13.3\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{South Orange Co} & \cellcolor{gray!6}{5.4} & \cellcolor{gray!6}{5.9} & \cellcolor{gray!6}{2.5} & \cellcolor{gray!6}{4.0} & \cellcolor{gray!6}{6.4} & \cellcolor{gray!6}{7.5} & \cellcolor{gray!6}{6.0} & \cellcolor{gray!6}{3.7} & \cellcolor{gray!6}{6.1} & \cellcolor{gray!6}{9.6}\\
-\hline
-\% in poverty & West Los Angeles \& West Beach & 11.3 & 12.0 & 5.0 & 8.3 & 13.2 & 10.4 & 10.5 & 5.2 & 8.2 & 11.8\\
-\hline
-\cellcolor{gray!6}{\% in poverty} & \cellcolor{gray!6}{Whittier \& North Orange Co} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{5.7} & \cellcolor{gray!6}{5.1} & \cellcolor{gray!6}{8.1} & \cellcolor{gray!6}{12.3} & \cellcolor{gray!6}{9.3} & \cellcolor{gray!6}{5.9} & \cellcolor{gray!6}{4.9} & \cellcolor{gray!6}{8.3} & \cellcolor{gray!6}{12.8}\\
-\hline
-\% with BA+ & Anaheim & 22.2 & 15.5 & 11.9 & 16.5 & 26.3 & 31.1 & 17.4 & 19.2 & 25.3 & 40.8\\
-\hline
-\hline
-\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Covina \& West Covina} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{16.0} & \cellcolor{gray!6}{10.2} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{35.3} & \cellcolor{gray!6}{30.5} & \cellcolor{gray!6}{16.6} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{28.5} & \cellcolor{gray!6}{44.0}\\
-\hline
-\% with BA+ & East Los Angeles & 13.6 & 12.8 & 4.6 & 8.3 & 20.5 & 21.2 & 13.5 & 11.5 & 16.2 & 30.1\\
-\hline
-\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Glendale and Pasadena} & \cellcolor{gray!6}{29.4} & \cellcolor{gray!6}{17.6} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{27.5} & \cellcolor{gray!6}{40.5} & \cellcolor{gray!6}{39.1} & \cellcolor{gray!6}{19.3} & \cellcolor{gray!6}{24.2} & \cellcolor{gray!6}{39.2} & \cellcolor{gray!6}{53.7}\\
-\hline
-\% with BA+ & Hollywood \& Wilshire & 25.5 & 18.5 & 10.0 & 19.6 & 42.2 & 42.1 & 19.9 & 25.2 & 40.6 & 59.8\\
-\hline
-\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Long Beach} & \cellcolor{gray!6}{20.9} & \cellcolor{gray!6}{16.2} & \cellcolor{gray!6}{7.1} & \cellcolor{gray!6}{16.9} & \cellcolor{gray!6}{31.7} & \cellcolor{gray!6}{28.4} & \cellcolor{gray!6}{17.1} & \cellcolor{gray!6}{13.6} & \cellcolor{gray!6}{25.9} & \cellcolor{gray!6}{40.8}\\
-\hline
-\% with BA+ & Riverside, San Bernardino, \& Ontario & 16.6 & 12.7 & 7.0 & 12.4 & 23.0 & 22.5 & 14.2 & 10.8 & 19.4 & 31.3\\
-\hline
-\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{San Fernando Valley (East)} & \cellcolor{gray!6}{28.2} & \cellcolor{gray!6}{18.6} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{23.3} & \cellcolor{gray!6}{40.3} & \cellcolor{gray!6}{39.3} & \cellcolor{gray!6}{21.4} & \cellcolor{gray!6}{21.3} & \cellcolor{gray!6}{34.4} & \cellcolor{gray!6}{59.1}\\
-\hline
-\% with BA+ & San Fernando Valley (West) & 30.1 & 16.3 & 18.2 & 27.7 & 43.2 & 38.0 & 17.8 & 24.7 & 37.1 & 52.0\\
-\hline
-\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{Santa Ana} & \cellcolor{gray!6}{30.6} & \cellcolor{gray!6}{19.4} & \cellcolor{gray!6}{12.9} & \cellcolor{gray!6}{31.2} & \cellcolor{gray!6}{46.1} & \cellcolor{gray!6}{40.6} & \cellcolor{gray!6}{20.4} & \cellcolor{gray!6}{23.8} & \cellcolor{gray!6}{40.8} & \cellcolor{gray!6}{56.3}\\
-\hline
-\% with BA+ & South \& South Central Los Angeles & 8.9 & 10.0 & 3.1 & 5.4 & 10.1 & 16.2 & 14.9 & 7.1 & 10.8 & 19.5\\
-\hline
-\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{South Bay} & \cellcolor{gray!6}{31.7} & \cellcolor{gray!6}{20.8} & \cellcolor{gray!6}{13.7} & \cellcolor{gray!6}{25.8} & \cellcolor{gray!6}{46.9} & \cellcolor{gray!6}{41.5} & \cellcolor{gray!6}{21.8} & \cellcolor{gray!6}{22.6} & \cellcolor{gray!6}{36.7} & \cellcolor{gray!6}{58.8}\\
-\hline
-\% with BA+ & South Orange Co & 40.4 & 16.3 & 29.2 & 40.0 & 53.4 & 52.3 & 15.7 & 42.9 & 51.6 & 63.8\\
-\hline
-\cellcolor{gray!6}{\% with BA+} & \cellcolor{gray!6}{West Los Angeles \& West Beach} & \cellcolor{gray!6}{50.6} & \cellcolor{gray!6}{16.7} & \cellcolor{gray!6}{42.3} & \cellcolor{gray!6}{53.4} & \cellcolor{gray!6}{62.7} & \cellcolor{gray!6}{65.6} & \cellcolor{gray!6}{14.4} & \cellcolor{gray!6}{60.1} & \cellcolor{gray!6}{67.9} & \cellcolor{gray!6}{74.6}\\
-\hline
-\% with BA+ & Whittier \& North Orange Co & 21.4 & 14.0 & 10.2 & 16.3 & 32.0 & 29.0 & 14.9 & 17.9 & 23.9 & 40.9\\
-\hline
-\end{tabular}
-\end{table}
+\begingroup\fontsize{10}{12}\selectfont
+
+\begin{longtable}[l]{l|rrrrr}
+\caption{Characteristics of Geomarkets around Los Angeles {#tab:socal_eps_table_summary}}\\
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{5}{c}{2020} \\
+\textbf{Geomarket name} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75}\\
+\midrule
+\endfirsthead
+\caption[]{Characteristics of Geomarkets around Los Angeles {#tab:socal_eps_table_summary} \textit{(continued)}}\\
+\toprule
+\textbf{Geomarket name} & \textbf{Mean} & \textbf{SD} & \textbf{P25} & \textbf{P50} & \textbf{P75}\\
+\midrule
+\endhead
+
+\endfoot
+\bottomrule
+\endlastfoot
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% White, non-Hispanic}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 44.0 & 24.2 & 22.8 & 46.6 & 64.7\\
+\hspace{1em}CA15, San Fernando Valley (East) & 43.3 & 25.7 & 22.5 & 39.6 & 67.0\\
+\hspace{1em}CA16, Glendale and Pasadena & 35.3 & 24.0 & 12.8 & 34.1 & 57.2\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 60.2 & 20.2 & 45.9 & 62.9 & 75.5\\
+\hspace{1em}CA18, Hollywood & Wilshire & 31.2 & 25.7 & 8.1 & 22.6 & 53.5\\
+\hspace{1em}CA19, East Los Angeles & 6.5 & 7.1 & 2.0 & 4.1 & 8.4\\
+\hspace{1em}CA20, South Bay & 29.1 & 24.5 & 7.1 & 22.1 & 51.4\\
+\hspace{1em}CA21, South & South Central Los Angeles & 6.3 & 11.4 & 0.6 & 1.9 & 6.6\\
+\hspace{1em}CA22, Long Beach & 25.8 & 23.0 & 6.6 & 16.4 & 45.5\\
+\hspace{1em}CA23, Covina & West Covina & 18.8 & 16.8 & 5.4 & 12.5 & 27.1\\
+\hspace{1em}CA24, Whittier & North Orange Co & 22.4 & 15.3 & 10.9 & 17.5 & 32.7\\
+\hspace{1em}CA25, Anaheim & 26.1 & 17.8 & 12.6 & 20.4 & 36.9\\
+\hspace{1em}CA26, Santa Ana & 42.0 & 26.3 & 17.0 & 44.1 & 64.6\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 27.7 & 18.8 & 12.5 & 23.5 & 40.7\\
+\hspace{1em}CA28, South Orange Co & 52.9 & 18.1 & 40.7 & 55.0 & 66.3\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% Black, non-Hispanic}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 3.8 & 3.9 & 1.1 & 2.7 & 5.4\\
+\hspace{1em}CA15, San Fernando Valley (East) & 4.3 & 3.8 & 1.6 & 3.3 & 5.7\\
+\hspace{1em}CA16, Glendale and Pasadena & 3.4 & 5.2 & 0.5 & 1.8 & 4.1\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 5.4 & 8.1 & 1.1 & 2.9 & 6.2\\
+\hspace{1em}CA18, Hollywood & Wilshire & 11.1 & 16.8 & 1.9 & 4.5 & 10.5\\
+\hspace{1em}CA19, East Los Angeles & 1.3 & 2.8 & 0.0 & 0.4 & 1.4\\
+\hspace{1em}CA20, South Bay & 12.1 & 16.9 & 1.7 & 5.7 & 14.3\\
+\hspace{1em}CA21, South & South Central Los Angeles & 17.5 & 18.8 & 2.7 & 11.6 & 25.3\\
+\hspace{1em}CA22, Long Beach & 11.4 & 13.0 & 3.6 & 8.1 & 15.1\\
+\hspace{1em}CA23, Covina & West Covina & 3.2 & 3.1 & 0.8 & 2.3 & 4.7\\
+\hspace{1em}CA24, Whittier & North Orange Co & 2.4 & 2.7 & 0.4 & 1.5 & 3.0\\
+\hspace{1em}CA25, Anaheim & 4.4 & 5.6 & 0.8 & 2.3 & 6.0\\
+\hspace{1em}CA26, Santa Ana & 1.3 & 1.9 & 0.1 & 0.7 & 1.6\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 7.1 & 6.4 & 2.4 & 5.4 & 9.8\\
+\hspace{1em}CA28, South Orange Co & 1.5 & 1.8 & 0.3 & 1.0 & 2.3\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% Hispanic}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 36.8 & 25.9 & 16.2 & 28.8 & 56.9\\
+\hspace{1em}CA15, San Fernando Valley (East) & 40.6 & 27.2 & 14.8 & 37.3 & 60.9\\
+\hspace{1em}CA16, Glendale and Pasadena & 36.8 & 26.9 & 16.3 & 26.3 & 53.4\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 15.8 & 13.4 & 6.3 & 12.2 & 20.4\\
+\hspace{1em}CA18, Hollywood & Wilshire & 38.2 & 24.6 & 15.7 & 34.7 & 57.1\\
+\hspace{1em}CA19, East Los Angeles & 64.4 & 28.9 & 36.7 & 73.2 & 91.5\\
+\hspace{1em}CA20, South Bay & 33.2 & 22.5 & 12.9 & 29.8 & 51.8\\
+\hspace{1em}CA21, South & South Central Los Angeles & 70.5 & 24.8 & 60.2 & 75.7 & 90.2\\
+\hspace{1em}CA22, Long Beach & 45.8 & 23.3 & 26.6 & 43.5 & 63.4\\
+\hspace{1em}CA23, Covina & West Covina & 51.7 & 23.5 & 30.8 & 54.3 & 71.3\\
+\hspace{1em}CA24, Whittier & North Orange Co & 56.0 & 23.6 & 39.3 & 59.2 & 75.7\\
+\hspace{1em}CA25, Anaheim & 41.3 & 22.7 & 23.1 & 38.6 & 58.1\\
+\hspace{1em}CA26, Santa Ana & 33.0 & 27.9 & 12.3 & 21.2 & 46.6\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 53.8 & 22.0 & 36.3 & 54.1 & 72.4\\
+\hspace{1em}CA28, South Orange Co & 22.2 & 17.0 & 10.4 & 16.8 & 27.2\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% API, non-Hispanic}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 11.6 & 8.1 & 5.3 & 10.4 & 16.3\\
+\hspace{1em}CA15, San Fernando Valley (East) & 8.3 & 5.4 & 4.7 & 6.9 & 10.2\\
+\hspace{1em}CA16, Glendale and Pasadena & 20.9 & 19.8 & 7.1 & 13.0 & 27.3\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 13.1 & 9.5 & 6.5 & 9.8 & 17.6\\
+\hspace{1em}CA18, Hollywood & Wilshire & 15.7 & 14.0 & 5.7 & 11.4 & 20.1\\
+\hspace{1em}CA19, East Los Angeles & 26.5 & 26.0 & 2.7 & 16.4 & 47.9\\
+\hspace{1em}CA20, South Bay & 21.2 & 15.9 & 8.2 & 17.8 & 30.3\\
+\hspace{1em}CA21, South & South Central Los Angeles & 3.9 & 6.9 & 0.2 & 1.2 & 4.4\\
+\hspace{1em}CA22, Long Beach & 13.2 & 11.3 & 4.9 & 10.2 & 18.2\\
+\hspace{1em}CA23, Covina & West Covina & 23.7 & 19.3 & 10.0 & 16.7 & 31.9\\
+\hspace{1em}CA24, Whittier & North Orange Co & 15.9 & 16.2 & 5.2 & 12.0 & 18.0\\
+\hspace{1em}CA25, Anaheim & 24.5 & 18.1 & 11.2 & 19.4 & 34.2\\
+\hspace{1em}CA26, Santa Ana & 20.1 & 18.3 & 6.5 & 14.3 & 26.5\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 8.2 & 10.0 & 2.0 & 5.0 & 10.4\\
+\hspace{1em}CA28, South Orange Co & 18.8 & 15.1 & 9.0 & 14.1 & 23.6\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% Native, non-Hispanic}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 0.2 & 0.5 & 0.0 & 0.0 & 0.1\\
+\hspace{1em}CA15, San Fernando Valley (East) & 0.1 & 0.3 & 0.0 & 0.0 & 0.0\\
+\hspace{1em}CA16, Glendale and Pasadena & 0.2 & 0.5 & 0.0 & 0.0 & 0.1\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 0.2 & 0.4 & 0.0 & 0.0 & 0.0\\
+\hspace{1em}CA18, Hollywood & Wilshire & 0.2 & 0.5 & 0.0 & 0.0 & 0.0\\
+\hspace{1em}CA19, East Los Angeles & 0.2 & 0.6 & 0.0 & 0.0 & 0.2\\
+\hspace{1em}CA20, South Bay & 0.2 & 0.5 & 0.0 & 0.0 & 0.0\\
+\hspace{1em}CA21, South & South Central Los Angeles & 0.1 & 0.4 & 0.0 & 0.0 & 0.0\\
+\hspace{1em}CA22, Long Beach & 0.6 & 6.9 & 0.0 & 0.0 & 0.1\\
+\hspace{1em}CA23, Covina & West Covina & 0.2 & 0.7 & 0.0 & 0.0 & 0.2\\
+\hspace{1em}CA24, Whittier & North Orange Co & 0.8 & 7.7 & 0.0 & 0.0 & 0.3\\
+\hspace{1em}CA25, Anaheim & 0.5 & 6.0 & 0.0 & 0.0 & 0.2\\
+\hspace{1em}CA26, Santa Ana & 0.2 & 0.4 & 0.0 & 0.0 & 0.2\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 0.3 & 0.6 & 0.0 & 0.0 & 0.3\\
+\hspace{1em}CA28, South Orange Co & 0.2 & 0.4 & 0.0 & 0.0 & 0.0\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% two+ races, non-Hispanic}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 3.1 & 2.6 & 1.1 & 2.7 & 4.1\\
+\hspace{1em}CA15, San Fernando Valley (East) & 3.1 & 2.9 & 1.0 & 2.7 & 4.4\\
+\hspace{1em}CA16, Glendale and Pasadena & 3.0 & 2.7 & 1.0 & 2.4 & 4.2\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 4.8 & 2.9 & 2.7 & 4.2 & 6.7\\
+\hspace{1em}CA18, Hollywood & Wilshire & 3.2 & 2.9 & 1.3 & 2.4 & 4.6\\
+\hspace{1em}CA19, East Los Angeles & 0.9 & 1.2 & 0.0 & 0.4 & 1.4\\
+\hspace{1em}CA20, South Bay & 3.8 & 2.9 & 1.5 & 3.2 & 5.6\\
+\hspace{1em}CA21, South & South Central Los Angeles & 1.3 & 2.0 & 0.0 & 0.5 & 1.8\\
+\hspace{1em}CA22, Long Beach & 2.9 & 2.3 & 1.2 & 2.5 & 4.2\\
+\hspace{1em}CA23, Covina & West Covina & 2.0 & 2.1 & 0.4 & 1.5 & 2.9\\
+\hspace{1em}CA24, Whittier & North Orange Co & 1.9 & 1.7 & 0.5 & 1.5 & 2.8\\
+\hspace{1em}CA25, Anaheim & 2.9 & 2.3 & 1.1 & 2.1 & 4.5\\
+\hspace{1em}CA26, Santa Ana & 3.2 & 2.9 & 1.1 & 2.6 & 4.9\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 2.6 & 2.4 & 0.7 & 1.9 & 3.7\\
+\hspace{1em}CA28, South Orange Co & 4.1 & 2.7 & 2.1 & 3.7 & 5.5\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{Median income}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 121.0 & 46.0 & 87.0 & 116.0 & 146.0\\
+\hspace{1em}CA15, San Fernando Valley (East) & 99.0 & 56.0 & 61.0 & 83.0 & 116.0\\
+\hspace{1em}CA16, Glendale and Pasadena & 102.0 & 40.0 & 73.0 & 96.0 & 121.0\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 137.0 & 61.0 & 98.0 & 122.0 & 165.0\\
+\hspace{1em}CA18, Hollywood & Wilshire & 84.0 & 43.0 & 52.0 & 75.0 & 105.0\\
+\hspace{1em}CA19, East Los Angeles & 76.0 & 25.0 & 60.0 & 72.0 & 87.0\\
+\hspace{1em}CA20, South Bay & 115.0 & 51.0 & 78.0 & 103.0 & 141.0\\
+\hspace{1em}CA21, South & South Central Los Angeles & 67.0 & 27.0 & 50.0 & 62.0 & 78.0\\
+\hspace{1em}CA22, Long Beach & 90.0 & 36.0 & 62.0 & 85.0 & 111.0\\
+\hspace{1em}CA23, Covina & West Covina & 103.0 & 32.0 & 82.0 & 99.0 & 120.0\\
+\hspace{1em}CA24, Whittier & North Orange Co & 105.0 & 30.0 & 85.0 & 101.0 & 121.0\\
+\hspace{1em}CA25, Anaheim & 108.0 & 41.0 & 78.0 & 99.0 & 129.0\\
+\hspace{1em}CA26, Santa Ana & 119.0 & 44.0 & 87.0 & 112.0 & 142.0\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 95.0 & 37.0 & 68.0 & 88.0 & 119.0\\
+\hspace{1em}CA28, South Orange Co & 142.0 & 46.0 & 109.0 & 140.0 & 171.0\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{Mean income}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 152.0 & 65.0 & 105.0 & 138.0 & 180.0\\
+\hspace{1em}CA15, San Fernando Valley (East) & 137.0 & 89.0 & 84.0 & 109.0 & 154.0\\
+\hspace{1em}CA16, Glendale and Pasadena & 135.0 & 57.0 & 96.0 & 121.0 & 157.0\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 206.0 & 108.0 & 133.0 & 169.0 & 252.0\\
+\hspace{1em}CA18, Hollywood & Wilshire & 118.0 & 68.0 & 73.0 & 98.0 & 143.0\\
+\hspace{1em}CA19, East Los Angeles & 97.0 & 33.0 & 77.0 & 93.0 & 106.0\\
+\hspace{1em}CA20, South Bay & 149.0 & 77.0 & 95.0 & 124.0 & 180.0\\
+\hspace{1em}CA21, South & South Central Los Angeles & 86.0 & 32.0 & 66.0 & 78.0 & 95.0\\
+\hspace{1em}CA22, Long Beach & 113.0 & 45.0 & 80.0 & 104.0 & 137.0\\
+\hspace{1em}CA23, Covina & West Covina & 125.0 & 42.0 & 100.0 & 115.0 & 143.0\\
+\hspace{1em}CA24, Whittier & North Orange Co & 125.0 & 37.0 & 102.0 & 117.0 & 142.0\\
+\hspace{1em}CA25, Anaheim & 131.0 & 54.0 & 99.0 & 118.0 & 146.0\\
+\hspace{1em}CA26, Santa Ana & 152.0 & 62.0 & 108.0 & 139.0 & 174.0\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 114.0 & 42.0 & 83.0 & 103.0 & 137.0\\
+\hspace{1em}CA28, South Orange Co & 179.0 & 62.0 & 138.0 & 168.0 & 213.0\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% in poverty}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 9.8 & 7.8 & 4.2 & 8.1 & 13.3\\
+\hspace{1em}CA15, San Fernando Valley (East) & 14.6 & 8.9 & 8.2 & 13.3 & 19.0\\
+\hspace{1em}CA16, Glendale and Pasadena & 12.6 & 7.7 & 6.4 & 11.3 & 16.8\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 10.4 & 10.5 & 5.2 & 8.2 & 11.8\\
+\hspace{1em}CA18, Hollywood & Wilshire & 17.7 & 11.8 & 9.3 & 14.7 & 23.9\\
+\hspace{1em}CA19, East Los Angeles & 16.0 & 7.9 & 10.2 & 15.1 & 21.1\\
+\hspace{1em}CA20, South Bay & 9.6 & 6.9 & 4.9 & 7.5 & 13.3\\
+\hspace{1em}CA21, South & South Central Los Angeles & 21.7 & 11.5 & 13.7 & 20.6 & 28.5\\
+\hspace{1em}CA22, Long Beach & 14.5 & 12.0 & 6.8 & 12.1 & 19.9\\
+\hspace{1em}CA23, Covina & West Covina & 10.3 & 6.4 & 5.8 & 9.0 & 12.8\\
+\hspace{1em}CA24, Whittier & North Orange Co & 9.3 & 5.9 & 4.9 & 8.3 & 12.8\\
+\hspace{1em}CA25, Anaheim & 10.7 & 6.7 & 5.5 & 9.6 & 14.8\\
+\hspace{1em}CA26, Santa Ana & 9.8 & 6.4 & 5.0 & 8.3 & 13.0\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 12.4 & 9.2 & 5.8 & 10.6 & 17.2\\
+\hspace{1em}CA28, South Orange Co & 7.5 & 6.0 & 3.7 & 6.1 & 9.6\\
+\addlinespace[0.5em]
+\multicolumn{6}{l}{\textbf{\% with BA+}}\\
+\hspace{1em}CA14, San Fernando Valley (West) & 38.0 & 17.8 & 24.7 & 37.1 & 52.0\\
+\hspace{1em}CA15, San Fernando Valley (East) & 39.3 & 21.4 & 21.3 & 34.4 & 59.1\\
+\hspace{1em}CA16, Glendale and Pasadena & 39.1 & 19.3 & 24.2 & 39.2 & 53.7\\
+\hspace{1em}CA17, West Los Angeles & West Beach & 65.6 & 14.4 & 60.1 & 67.9 & 74.6\\
+\hspace{1em}CA18, Hollywood & Wilshire & 42.1 & 19.9 & 25.2 & 40.6 & 59.8\\
+\hspace{1em}CA19, East Los Angeles & 21.2 & 13.5 & 11.5 & 16.2 & 30.1\\
+\hspace{1em}CA20, South Bay & 41.5 & 21.8 & 22.6 & 36.7 & 58.8\\
+\hspace{1em}CA21, South & South Central Los Angeles & 16.2 & 14.9 & 7.1 & 10.8 & 19.5\\
+\hspace{1em}CA22, Long Beach & 28.4 & 17.1 & 13.6 & 25.9 & 40.8\\
+\hspace{1em}CA23, Covina & West Covina & 30.5 & 16.6 & 16.2 & 28.5 & 44.0\\
+\hspace{1em}CA24, Whittier & North Orange Co & 29.0 & 14.9 & 17.9 & 23.9 & 40.9\\
+\hspace{1em}CA25, Anaheim & 31.1 & 17.4 & 19.2 & 25.3 & 40.8\\
+\hspace{1em}CA26, Santa Ana & 40.6 & 20.4 & 23.8 & 40.8 & 56.3\\
+\hspace{1em}CA27, Riverside, San Bernardino, & Ontario & 22.5 & 14.2 & 10.8 & 19.4 & 31.3\\
+\hspace{1em}CA28, South Orange Co & 52.3 & 15.7 & 42.9 & 51.6 & 63.8\\*
+\end{longtable}
+\endgroup{}

### 2024-10-18 soc_of_ed/soc_of_ed_geomarket.qmd
revising front end

--- 2024-10-18 aefp/aefp_geomarket.qmd
@@
   - \usepackage{setspace}
-  - \doublespacing
+  #- \doublespacing
   - \usepackage{fancyhdr}
@@
 Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
-EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.^[In 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS [@RN5021].]
+EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.[^1]
 
-Geomarkets were incorporated into the _College Board_ "student list" product named Student Search Service.
+[^1]: In 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS [@RN5021].
+
+Geomarkets were incorporated into the *College Board* "student list" product named Student Search Service.
 Student lists have been the primary source of lead generation in U.S.
-higher education for over 50 years [@belkin2019-studata; @list_biz].^[@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.] Lists contain contact information for prospective students.
+higher education for over 50 years [@belkin2019-studata; @list_biz].[^2] Lists contain contact information for prospective students.
 The Student Search Service database consists of College Board test-takers.
@@
 An analysis of 830 student lists purchased by 14 public universities found that 16% of purchases filtered on Geomarket [@RN5012].
+
+[^2]: @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
 
@@
 1.  What is the socioeconomic and racial variation between Geomarkets in metropolitan areas? How does this variation change over time?
-1. How does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
+2.  How does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
 
@@
 We obtained shapefiles for College Board Geomarkets from a 2012 *R-bloggers* post.
-We assigned census tracts to Geomarkets by implementing a partial spatial join.^[Census tracts that intersected multiple Geomarkets were partially assigned based on the proportion of land area.
-Currently, our spatial join for the 2020 ACS is causing some invalid geometries that we will fix prior to AEFP.]
+We assigned census tracts to Geomarkets by implementing a partial spatial join.[^3]
+
+[^3]: Census tracts that intersected multiple Geomarkets were partially assigned based on the proportion of land area.
+Currently, our spatial join for the 2020 ACS is causing some invalid geometries that we will fix prior to AEFP.
 
@@
 For 414 College Board student list purchases, associated with about 2.6 million prospects, we received both (1) the order summary -- showing which search filters were utilized -- and (2) the de-identified prospect-level data, including race, ethnicity, zip code, and high school.
-@fig-asu-order-448922 shows the partial order summary, which we converted to tabular data, for a purchase by Arizona State University that filtered on the 2020 high school graduating class, PSAT scores of 1070 to 1180, and selected Geomarkets.^[Analyses for RQ2 additionally utilize data on public and private high schools from NCES.]
+@fig-asu-order-448922 shows the partial order summary, which we converted to tabular data, for a purchase by Arizona State University that filtered on the 2020 high school graduating class, PSAT scores of 1070 to 1180, and selected Geomarkets.[^4]
+
+[^4]: Analyses for RQ2 additionally utilize data on public and private high schools from NCES.
 
@@
 \newpage
+
 ```{r fig-asu-order-448922, echo=FALSE, fig.cap="Partial order summary from a College Board student list purchase", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
@@
 \newpage
+
 ```{r fig-socal-hispanic, echo=FALSE, fig.cap="Screenshot of Los Angeles area Geomarkets, showing 2020 percent Hispanic by census tract", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}

### 2024-10-18 text/front_end_v2.qmd
revising front end

--- 2024-07-08 text/front_end.qmd
@@
 
-```{=tex}
 \pagenumbering{roman}
@@
 \pagenumbering{arabic}
-```
 
@@
 
-In 1983, Zemsky and Oedel authored *The Structure of College Choice*, published by *The College Board.* This book develops the Market Segment Model and introduces local "community-based enrollment markets" (p.14) -- later called "geomarkets" -- which are the basis for the College Board market Enrollment Planning Service (EPS), a market research product.
-@RN4982[pp., 9-10] write that the Market Segment Model sought to quantify the knowledge that college admissions officers have about student demand:
-
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting\ldots We believe that the institutions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research, thus, is based on listening carefully to what admissions officers have to say."
-
-@RN4982 [p. 10] stress the importance of "knowing the territory":
+In 1983, Robert Zemsky and Penney Oedel authored *The Structure of College Choice*.
+Based on an analysis of 1980 SAT score-sending behavior and published by *The College Board*, @RN4982 developed the Market Segment Model to help colleges recruit students.
+The underlying idea of the Market Segment Model is that student demand for your institution is fundamentally a function of class.
+Therefore, the key to recruitment is identifying territories where large numbers of your class live.
+@RN4982[p.
+11] describe the Market Segment Model as an effort to "quantify" the knowledge that college admissions officers have about the geography of student demand:
 
@@
 
-The Market Segment Model was created by analyzing College Board data about which colleges SAT test-takers from 1980 sent scores to.
-High school students who sent scores mostly to colleges in their local geomarket were defined as "local" students.
-"In-state" students were those who sent most scores to colleges outside their local market but in their home state.
-"Regional" students were those who sent most scores to colleges outside their home state but in their region (e.g., New England, Midwest).
-"National" students were those who sent most scores to colleges outside their region.
-Having categorized students, each local geomarket could be evaluated by the number of students from each of the four market segments, yielding the Market Segment Profile (see @tbl-market-segment-characteristics).
-Next, for each local geomarket, the Institutional Profile (see @tbl-sample-institutional-profile) the number of students from each market segment that sent scores to a particular college, which could be your college or a competitor.
+@RN4982 also created local "community based markets" (p.
+14) -- later known as "Geomarkets." Geomarkets are geographic borders that disaggregate states and large metropolitan areas for recruiting purposes.
+@fig-nyc-geomarkets shows New York City area Geomarkets.
+In defining Geomarkets, @RN4982 did not draw new geographic borders, but rather made certain extant borders more salient to the process of recruiting.
 
-Collectively, the quantitative outputs of the Market Segment Model provided intelligence about where are the desirable students and which students might consider your college or a competitor.
-The Market Segment model and geomarkets became the basis or the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
-@RN5019_2010 says "Enrollment Planning Service (EPS\textsuperscript{\textregistered}) the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found." @RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
+Geomarkets have been incorporated into the supply-side structure of college access in ways that are simultaneusly mundane and fundamental.
+We posit three ways.
+First, Geomarkets became a principle for how many -- if -- admissions offices organized their labor.
+In states where a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory [CITE].
+Geomarkets were created to mirror the territories of admissions officers and, in turn, admissions offices structured their territories around Geomarkets.
+Second, The Market Segment Model and Geomarkets became the basis for Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
+EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
+@RN5019_2010 says "Enrollment Planning Service (EPS\textsuperscript{\textregistered}) the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found." @RN5021 reports that, in 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS.
 
-In 19XX, geomarkets were incorporated into the College Board Student Search Service student list product, which College Board created in 1972 [@belkin2019-studata].
-Student lists contain the contact information of prospective students, the basic input for recruiting campaigns that target individual prospects using mail, email, and targeted social media.
-RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+Third, Geomarkets were incorporated into the College Board "student list" product named Student Search Service.
+Student lists have been the primary source of lead generation in U.S.
+higher education since 1972, when College Board began selling "names" [@belkin2019-studata; @list_biz].
+@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+Lists contain contact information for prospective students.
 The Student Search Service database consists of College Board test-takers.
-Colleges pay a fee for each prospect (e.g., \$0.50 in 20XX) and control which prospect profiles through purchase by selecting search filters, such as high school graduation year, SAT score, AP test score, state, etc.
-Geomarket filters enable colleges to include/exclude prospects from particular geomarkets.
+Colleges pay a fee for each prospect (e.g., \$0.50 in 2021).
+Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc.
+Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets.
+An analysis of 830 student lists purchased by 14 public universities found that 16% of purchases filtered on Geomarket [@RN5012].
 
-Scholarship in the sociology of education describes schools and colleges as "sorting machines" that reproduce inequality by incorporating, creating, and allocating students to different categories [@RN4873].
-Given the consequences subsequent outcomes, the sociology of education is particularly concerned with the transition from high school to college.
-Over the last decade or so, scholarship on college access has begun to take more seriously the idea that colleges are not passive recipients of applications; rather they expend substantial resources soliciting demand from desirable applicants [@RN3519; @RN4774; @RN4324].
-Most scholarship on recruiting analyzes the behavior of individual colleges, thereby assuming that recruiting behavior is a function of individual colleges [e.g., @RN4758; @RN4759].
-@RN5012 analyze the College Board *Student Search Service* product, highlighting the role that third-party products play in structuring the recruiting behavior of colleges.
-The authors show that the utilization of particular search filters -- particularly geographic filters, such as zip code -- is associated with systematic racial exclusion that does not depend on which college purchased the list.
+Geomarkets have never been the subject of empirical research.
+This omission is surprising because the sociology of education is concerned with sorting [@RN4873], particularly sorting into college.
+Traditionally, most scholarship -- and certainly most scholarship in economics [HOXBY?
+CHETTY] -- views inequality in college access as a function households or k-12 schools, which represent the demand-side of higher education.
+More recently, scholarship in the sociology of education takes seriously the idea that colleges are not passive recipients of applications, but rather they expend substantial resources soliciting demand from desirable applicants [@RN3519; @RN4774; @RN4324].
+However, most scholarship on recruiting analyzes the behavior of individual colleges [e.g., @RN4758; @RN4759], thereby creating the assumption that recruiting is a function of individual colleges.
+What these studies have missed is the broader enrollment management industry -- third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access, funneling certain kinds of students to certain kinds of institutions.
 
-However, prior research has not examined the role of third-party market research in reproducing socioeconomic and racial inequality.
-This manuscript analyzes the Market Segment Model and College Board Geomarkets as a case study of quantification.
-We develop arguments from the broader literature on quantification [e.g.,] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
+This manuscript analyzes one of those mechanisms, The Market Segment Model and College Board Geomarkets, as a case study of quantification.
+We develop arguments from the broader literature on quantification \[e.g.,\] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
 Predictive analytics are based on correlation.
 When we use data on past correlations to make recommendations about the future, we amplify the effects of historic structural inequality.
-The Market Segment Model analyzes SAT score-sending data from 1980 and concludes that student demand for higher education is primarily a function of social class= [@RN4982].
-Homophily is the idea that actors that share common characteristics are likely to form connections with one another.
-The Market Segment Model argues that homophily is the organizing principle of student demand and competition between colleges, that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students" [@RN4982, p.
-72].
-Subsequently, the assumption of homophily is programmed into the EPS and Student Service Service products that colleges utilize to decide where to focus recruiting energy.
+The Market Segment Model analyzes SAT score-sending data from 1980 and concludes that student demand for higher education is primarily a function of social class [@RN4982].
+The Market Segment Model argues that homophily -- actors that share characteristics are likely to form connections --  is the organizing principle of student demand and competition between colleges.
+@RN4982[p.
+72] state that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students."  This snapshot of student demand in 1980, itself a consequence historic structural inequality, was programmed into the way admissions offices organize labor and programmed into the EPS and Student Service Service products that colleges utilize to identify and target prospective students.
+The result is a supply-side that reinforces structural inequalities observed on the demand-side.
 
-Empirically, our analyses address the following research questions, which speak to how geomarkets are utilized within the Enrollment Planning Service and within the Student Search Service student list product.
+Our analyses address the two research questions, which speak to how geomarkets are utilized within EPS and within the Student Search Service student list product.
 First, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
 We address this question by spatially joining the geomarket shapefile to Census data about socioeconomic and racial characteristics.
-Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets, in concert with other commonly used search filters?
-We address this question by analyzing student list purchases that do not filter geomarkets and simulating which prospects would have been included had they filtered on particular geomarkets.
-We obtained these data by issuing public records requests to public universities in four states.
-Third, in student list purchases that filtered on geomarkets, how does the socioeconomic and racial composition of included students compare the that of the geomarket and how does it compare to that of the broader metropolitan area?
+Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
+We address this question by analyzing actual student list purchases that utilize commonly used search filters (e.g., PSAT score, GPA) but do not filter on Geomarkets.
+We simulate which prospects in particular metropolitan areas would have been included/excluded had the student list purchase filtered on particular Geomarkets.
+We obtained these data by issuing public records requests to public universities.
+<!-- Third, in student list purchases that filtered on geomarkets, how does the socioeconomic and racial composition of included students compare the that of the geomarket and how does it compare to that of the broader metropolitan area?
+-->
 
+<!-- 
+THIS TEXT IS FINE, BUT TRYING TO SET UP THE CONTRIBUTION TO BROADER QUANTIFICATION LITERATURE, VIS-A-VIS NO-ONE HAS STUDIED MARKET RESEARCH; DON'T THINK THIS IS NECESSARY AND YOU ARE NOT ON FIRM GROUND HERE IN CALLING THIS "MARKET RESEARCH"
 Scholarship on quantification in education has focused on consumer-facing measures such as U.S.
@@
 However, we show the mechanisms of how geomarkets reproduce inequality by simulating who is included/excluded when geomarkets are utilized within student list purchases as recommended by @RN4982.
+-->
 
@@
 Fifth, we present results.
-Finally, we discuss implications for scholarship, policy, and practice.
+Finally, we discuss implications for scholarship and for policy.
 
-# Background: Enrollment Management
+# Enrollment Management and Quantification
 
-The term "enrollment management" (EM) refers to a profession, an administrative structure, and an industry.
-As a profession, EM integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p.
+**Enrollment Management**
+
+As a profession, enrollment management (EM) integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p.
 xiv].
 As an administrative structure, the EM office typically controls the activities of admissions, financial aid, and recruiting [@RN2406].
-The EM industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers that produce market research, software, and and consulting services (e.g., College Board, ACT, EAB, Huron, Slate, PowerSchool).
-
-Figure @fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
+@fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
 The funnel begins with a large pool of "prospects" (i.e., prospective students) that the university would like to enroll.
 "Leads" are prospects whose contact information has been obtained.
-"Inquiries" are prospects that contact the institution and consist of two types: first, inquiries who respond to an initial solicitation (e.g., email) from the university; and second, "student-as-first-contact" inquiries who reach out to the university on their own (e.g., sending ACT scores).
+"Inquiries" are prospects that contact the institution, consisting of those that respond to an initial soliciation (e.g., email) and those that reach out on their own (e.g., sending SAT scores).
 <!--  Applicants consist of inquiries who apply plus "stealth applicants" who do not contact the university before applying.
--->The funnel narrows at each successive stage in order to convey the assumption of "melt" (e.g., a subset of "inquiries" will apply).
-Practically, the purpose of the enrollment funnel is to inform recruiting interventions that target one or more stages.
+The funnel narrows at each successive stage in order to convey the assumption of "melt" (e.g., a subset of "inquiries" will apply).
+-->The purpose of the enrollment funnel is to inform recruiting interventions that target one or more stages.
 These interventions seek to increase the probability of "conversion" across stages [@RN4322].
@@
 
-Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel they focus on.
+Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel it speaks to.
 The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
-A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
-Scholarship on recruiting has conceptualized recruiting behavior as an indicator of college enrollment priorities [@RN4758] and identifies recruiting practices as a mechanism for social reproduction in college access [@RN4324; @RN3519].
-
-Ethnographies by @RN3519 and by @RN4407 analyze the connections between privileged colleges and privileged high schools.
-@RN4407 shows how guidance counselors at elite private school guidance counselors lobby admissions counselors on behalf of marginal students.
-@RN3519 finds that college admissions officers visit high schools as a means of maintaining positive relationships with guidance counselors at affluent feeder schools.
-Consistent with this finding, quantitative case studies of off-campus recruiting visits show that selective private colleges visit affluent, predominantly white schools and communities, with a disproportionate number of visits to private high schools [@RN5014; @RN4330].
-Additionally, public research universities often make make more out-of-state recruiting visits than in-state visits, and these out-of-state visits focus on the same set sorts of schools targeted by selective private universities [@RN4759; @RN4758].
-
-Other studies investigate recruiting behavior by non-selective institutions [@RN4520; @cottom2017lower; @dache2018dangling].
-Analyzing recruiting from the perspective of high school students, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
-@cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.
-<!--  WHAT YOU SAID ABOUT POSECZNICK Posecznick (2017) studied the “Ravenwood College,” a private nonprofit college founded to serve working Black women.
-Consistent with analyses of tuition discounting by Lapovsky (2001), the survival of Ravenwood depends on the success of admissions officers in identifying prospects (e.g., through on- and off-campus recruiting events) and “converting” these prospects to enrolled students.
+Economists often investigate financial aid leveraging, which seeks to convert admits to enrolled students [CITE].
+<!--  Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935; @RN1814].
 -->
 
-We make two observations about the recruiting literature.
-First, consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015] and industrial organization [@RN1549; @RN2133], the industry is characterized by a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches, a process that exemplifies the idea that higher education is mechanism of social reproduction [@RN2377; @RN4873].<!--  For example, for-profit institutions target adults in communities not being targeted by traditional colleges [@cottom2017lower].
-Non-selective private non-profits target prospects who might be considering community college or a local state colleges [@RN4324].
-Selective private colleges target private high schools, with a particular link between sectarian high schools and colleges of the same faith [@RN5014], but these schools are increasingly targeted by out-of-state public research universities, suggesting growing competition for a coveted market niche.
---> Second, by focusing on the behaviors of individual colleges, the literature implicitly states that recruiting is something done by individual colleges.
+A growing college access literature in sociology analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
+@RN4758 conceptualize recruiting behavior as an indicator of college enrollment priorities.
+Ethnographies by @RN3519 and @RN4407 identify connections between private school guidance counselors and college admissions officers as a mechanism for social reproduction.
+Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly white public schools [@RN4758; @RN4759; @RN5023].
+From the student perspective, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
+@cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.
+Consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015], this literature reveals a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches segmented by class and race, a process consistent with the notion of that higher education is mechanism of social reproduction [@RN2377; @RN4873].
+
+By focusing on the behaviors of individual colleges, scholarship on college access implicitly states that recruiting -- and enrollment management more broadly -- is something done by individual colleges.
+This statement raises the question, why does recruiting behavior look so similar across colleges of a particular "type." Institutional theory defines the organizational field as "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services and products" [@RN527, p.
+143].
+The broader enrollment management (EM) industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers that produce market research, software, and consulting services (e.g., College Board, ACT, EAB, Slate, PowerSchool).
+We argue college access is structured by third-party servicers and products that interact with direct-providers (colleges).
+Enrollment management is often described as an opaque industry because obtaining concrete information about EM practices is difficult.
+This is why the title of @RN5024 is *Lifting the Veil on Enrollment Management*.
+Although sociologists have hinted at the ways enrollment management contributes to inequality in colege access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
 <!--  
-Why then, does recruiting behavior look so similar across colleges of a particular "type"?[^1]
 [^1]: New institutional theory offers one explanation [@RN527].
@@
 
-<!-- make or buy  
-The process of recruiting students involves many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor.
+Drawing from scholarship on organizational theory [@RN4738], enrollment management processes involve many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
 EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
+More recently, the EM consulting business model is based on software-as-service products (e.g., for example, EAB's [Enroll 360](https://eab.com/solutions/enroll360/) product) [@list_biz].
 As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
 The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
+<!-- In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
 -->
 
-However, the recruiting behavior of colleges is structured by third-party actors.
-Drawing from scholarship on organizational theory [@RN4738], enrollment management processes involve many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-<!-- As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
---> The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
+The EM industry puts great stock in market research, in part because the EM profession emerged from marketing and market research [@RN4986; @RN2771].
+@RN4986 [p.
+17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].[^1] In recruiting, market research informs decisions about which prospective students to target with recruiting interventions.<!-- THIS IDEA HAS TO GO SOMEWHERE: IN THE HOOK AT END OF THIS SECTION.
+These recruiting interventions intervene on student demand [@RN4324; @cottom2017lower], structuring college access pathways in ways that could amplify or diminish historic inequality in college access.
+--> How to acquire market research is another make-or-buy decision.
+The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals.
+In 1983, College Board published *Applying Market Research in College Admissions* [@RN4986], which became the basis for their Admitted Student Questionnaire (ASQ) product, and published *The Structure of College Choice* [@RN4982], which became the basis for the Enrollment Planning Service (EPS) product/division.
+Today, EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
+Market research often distills large amounts of data into simpler indicators.
+Therefore, the quantification literature provides a useful lens to understand market research.
 
-Moreover, process of college access is structured by third-party recruiting products utilized by colleges.
-@RN5012 deconstruct the College Board *Student Search Service* product.
-Selection devices are procedures or routines for making categorical decisions based on input values [@RN4778].
-Student list products are selection devices that enable colleges to choose which prospective students they target or ignore based on search filters that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
-"Racialized inputs" are ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
-@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) are racialized inputs.
-Using a national sample of high school students, and also using data from actual student lists purchased by public universities, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
- <!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
--->
-
-Student list products offer many geographic filters.
-Geographic borders are perhaps the most commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887], owing to extreme racial segregation in the U.S.
-In addition to offering geographic filters based on known geographic borders (e.g., zip code, county, CBSA, state), the College Board created new geographic borders that subsequently became filters in the *Student Search Service*.
-The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
-Geodemographic search filters can be read as the mimicry of corporate market logic [@RN4004] in that they apply the logic and methods of commercial geodemographic market segmentation products, such as PRIZM by the Claritas Corporation [@RN4957], to the case of college recruiting.
-
-Geomarket search filters, the focus of this manuscript, slice states and metropolitan areas into smaller, local recruiting markets.
-For example, geomarket "CA08" refers to "Alameda Country excluding Oakland." A university might purchase prospect contact information by filtering on this geographic region in combination with additional filters, such as year of high school graduation and SAT test score.
-When we initiated this manuscript, we asked ourselves, "where did geomarket filters come from?" The answer surprised us.
-We had imagined ourselves studying the application of the "data science revolution" to college recruiting, but it turned out we were studying the 1970s and 1980s!
-The Market Segment Model and geomarkets can be read as the application of social science research methods to "capture and quantify the phenomenon behind the folklore" [@RN4982, p.
-11] of college admissions officers.
-This work, comissioned by College Board, was subsequently turned into a market research tool -- The College Board Enrollment Planning Service -- that helps colleges decide where to look for prospective students.
-
-```{=html}
-<!--The enrollment management industry emphasizes market research.
-Market research is the process of producing information that informs decisions about product design or how to find customers.   -->
-```
-
-The enrollment management industry puts great stock in market research.
-@RN4986[p.
-17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].^[@RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
+[^1]: @RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
 In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited.
 In the chapter entitled "Travel," @RN3519 [p.
-69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."] However, most market research in the contemporary enrollment management industry is produced by third-party organizations.
-For example, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development [e.g., @RN4741; @RN4728, but most only available to clients.
-EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
+69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."
 
-Despite being central to enrollment management for at least 40 years [@RN4986], the role of third-party market research has been ignored by scholarship on college access.
-Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935;@RN1814].
-However, market research informs where colleges look for students and which prospective students to target.
-These recruiting interventions intervene on student demand [@RN4324; @cottom2017lower], thereby structuring college access pathways in ways that could amplify or diminish historic inequality in college access.
-The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals [@RN4986].
-This article offers one story about the quiet revolution that transformed college access into a sophisticated, supply-led enterprise.
-We argue that market research structures who colleges recruit, in ways that reinforce existing social stratification.
-The case of Market Segment Model and College Board geomarkets shows how market research -- and its commodification into recruiting products -- is a mechanism for social reproduction.
-[LAST TWO SENTENCES ARE REDUNDANT BUT BUT ARE TRYING TO DO SLIGHTLY DIFFERENT THINGS AND I LIKE THEM BOTH]
+**Quantification**
 
-# Quantification
-
-```{=html}
-<!--
-CUT GENERAL PARAGRAPH ON REPRODUCTION
-
-Within the sociology of education, @RN2377 describes education as a sieve, allocating people to opportunities in socially stratified, capitalist societies.
-This role highlights "Max Weber's core insight that education has a dual character" [@RN2377, p.
-129], on one hand, serving social mobility by redistributing access to privileged positions and, on the other, serving social reproduction by granting privileged families disproportionate access to the educational institutions that confer pathways to privileged positions.
-Similarly, @RN4873 describe schools and colleges as "sorting machines" that incorporate external categories (e.g., race, class), create categories (e.g., instructional tracks), and allocate students to categories.
-Initial category allocation affects subsequent allocation, such that school assignments and instructional track assignments affect college access, which affects employment assignments.
-"In evaluating schools as inequality regimes" @RN4873 \[p.
-322\] argue, "the relevant question is not whether schools produce inequality; rather, the relevant questions are, when and where do schools produce inequality, \[and\] how do they do so." As @RN2377 show, whereas early research on social stratification sought to identify patterns of inequality [e.g., @RN4873; @RN290; @RN1268], the sociology of education is now more concerned with identifying the mechanisms underlying these patterns.
--->
-```
-We situate our study amidst scholarship on within scholarship on quantification as a mechanism that reproduces inequality.
 @RN5007 [p.
 402] define quantification as "the production and communication of numbers" (p.
-402) and introduce reacivity, discipline, and authority as three interrelated themes that describe the effects of quantification.
+402).
+They introduce reacivity, discipline, and authority as three interrelated themes that describe the effects of quantification.
 Reactivity is the idea that salient quantitative measures cause people and organizations to change their behavior.
@@
 News & World Report (USNWR) Law School rankings demonstrate the reactivity, discipline, and authority effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
-Developed as a means of informing prospective students about quality, rankings had the intended effect of affecting.
-Rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
+Law school rankings were developed as a means of informing prospective students about the relative quality of different programs.
+Speaking to *reactivity* effects, rankings had the intended effect of affecting where students apply and enroll [CITE].
+However, rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
 Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p.
@@
 For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
-Rankings also weakened the *authority* of admissions directors and admissions officers to make independent decisions about which applicants to admit, ideal class size, and curricular offerings [@RN5003].
+Rankings also weakened the *authority* of admissions directors and admissions officers to make independent decisions about which applicants to admit, ideal class size, and curricular offerinags [@RN5003].
 <!--
@@
 The analysis of UK League Tables by @RN5005 show how quantification can be a mechanism of social reproduction.
-As part of a broader *New Public Management* effort to infuse market competition within public providers [@RN161], the UK government began publishing school "league tables" in 1992.
-League tables ranked schools based on student performance on high-stakes national exams taken at age 16.
-Using Census data on occupational class measured at the local authority level (similar to a U.S.
+In 1992, the UK government began publishing school "league tables," which ranked schools based on student performance on high-stakes national exams taken at age 16.
+League tables facilitate making evaluative comparisons between schools regardless of geographic proximity.
+They also discipline households and schools to conceive of school quality in terms of test scores, which are substantially a function of the class composition of schools.
+Using Census data measured at the local authority level (similar to a U.S.
 county), @RN5005 find that localities with higher performing schools experienced growth in the share of managerial/professional households following the adoption of league tables and a decline in the share of working-class households.
-Additionally, using longitudinal household survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables.
-With respect to mechanisms, league tables facilitate making evaluative comparisons between schools regardless of geographic proximity.
-They also discipline households and schools to conceive of school quality in terms of test scores, which are substantially a function of the class composition of schools.
-Compared to working-class parents, parents with professional/managerial occupations are more aware of school league tables and have financial resources necessary to move to neighborhoods near a "good" school.
-Therefore, league tables contributed to social reproduction because managerial/professional households respond to consumer-facing metrics by moving to more expensive neighborhoods, near higher performing schools.
-<!-- THIS IS RATCHET EFFECT STUFF.
+Using longitudinal survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables.
+The introduction of league tables contributed to social reproduction because professional/managerial households were more aware of these consumer-facing metrics and had resources to respond by moving to more expensive neighbhorhoods, near higher performing schools.<!-- THIS IS RATCHET EFFECT STUFF.
 I CUT IT BECAUSE YOU HAVEN'T INTRODUCED THE IDEA OF RATCHET EFFECT
@@
 
-```{=html}
 <!-- THIS IS THE CUT TRANSITION PARAGRAPH
@@
 -->
-```
-In the interdisciplinary literature on quantification [@RN5008], the discussions of correlation and homophily by @RN4975 introduce concepts that are salient to the analysis of market research.
-Chapter 1 of @RN4975 is *Correlating Eugenics*.
-Correlation measures the extent to which two or more variables move together.
-The Pearson Correlation Coefficient was developed by the 19th Century statisticians and "biometric eugenicists" Francis Galton and his student Karl Pearson [@RN4975].
-@RN4975 [p.
+
+**Correlation**.
+<!-- In the literature on quantification from critical data studies, --> The discussions of correlation and homophily by @RN4975 introduce ideas salient to the analysis the Market Segment Model and College Board Geomarkets.
+Correlation measures the extent to which two or more variables move together.<!--The Pearson Correlation Coefficient was developed by the 19th Century statisticians and "biometric eugenicists" Francis Galton and his student Karl Pearson [@RN4975].
+--> @RN4975 [p.
 50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared in his 2008 editorial 'The End of Theory,' big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'" Predictive analytics are developed in two steps.
+As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'" Predictive analytics are based on correlation and are developed in two steps.
 First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
@@
 Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated to a variable of interest.
-For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of these others.
+For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
 @RN4849 [p.
@@
 
-Many studies show that using correlational models make predictions reproduces structural inequality.
+Several studies show that using correlational models make predictions reproduces structural inequality [@RN4794].
 The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
 These correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
-@RN4794 [p.
-224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Therefore, disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies disproportionate targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].
-An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
+Reviewing scholarship about algorithms, @RN4794 [p.
+224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
+This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].^[An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
 Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
 Note that predictive models such as the LSI-R model are not merely used for social science purposes.
-Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.
+Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.]
 
-Chapter 2 of @RN4975, *Homophily, or the Swarming of the Segregated Neighborhood*, argues that homophily often underlies the use of correlation to predict future behavior.
-Homophily means that actors who share common characteristics are likely to form connections with one another.
-The ubiquitous colloquialism is "birds of a feather flock together." Homophily is a core assumption for network science [@mcpherson2001birds], in which actors (e.g., high school students) are linked to one another through direct and indirect network ties (e.g., submit SAT scores to the same college).
-Because network science often draws from rational choice theory, it assumes that homophily is the result of voluntary action by individuals, thereby "eras\[ing\] historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
+**Homophily**.
+@RN4975 argues that homophily is central to models that use correlation to predict future behavior.
+Homophily means that actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together." Homophily is a core assumption for network science [@mcpherson2001birds], in which actors -- in this manuscript, high school students -- are linked to one another through direct and indirect network ties -- submitting SAT scores to the same college.
+Because network science models often draws from rational choice theory, they assume that homophily is the result of voluntary action by individuals, thereby "eras\[ing\] historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
 95] that underlie behavior consistent with homophily.
 @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.[^4]
+In commercial social networks (e.g., Faceboo, X, TikTok), homophily is more than an assumption; homophily is programmed into the algorithms that create connections between users.[^2]
 
-[^4]: Thus, @RN4975 [p.
+[^2]: Thus, @RN4975 [p.
 82] writes, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together."
 
-Within market research, homophily is central to geodemography and the creation of market segments.
+Within the market research profession, homophily is central to geodemography and the creation of market segments.
 Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
 Market segments are subgroups within a larger market that have similar consumer demand.
-Early geodemographic classifications of consumers (e.g., PRIZM by Claritas Corporation) were derived from publicly available Census data, which disaggregated data to the zip code level [@RN4957].
-The Claritas Corporation argued -- and had a financial incentive to do so -- that people living near one another share similar consumer preferences.
-Therefore, geographic localities could be categorized into market segments that would be useful for direct mail marketing campaigns.
+Early geodemographic classifications of consumers (e.g., PRIZM by Claritas Corporation) were derived from publicly available Census data, which disaggregated data to the zip code level.
+The Claritas Corporation had a financial incentive to argue that people living near one another share similar consumer preferences because geographic localities could then be categorized into market segments that would be useful for direct mail marketing campaigns [@RN4957].
 Later, the development of individual credit scores (e.g., FICO score) enabled merchants to classify consumers into many, fine-grained groups [@RN4877].
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].
-For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].^[For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.] Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
 <!--
@@
 
-Reviewing recent scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Law school rankings [e.g., @RN5003] and school league tables [e.g., @RN5005] seem to exemplify this sort of salience.
-Widely known studies of quantification in education tend to study consumer-facing quantification.
-To our knowledge, prior research on quantification in education has not examined how third-party market research classifies consumers.
-This manuscript analyzes "market segment model" [@RN4982], which categorized high school students into vertical market segments and simultaneously created local "geo-markets" that were evaluated based on their composition of student market segments.
+Reviewing scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
+U.S.
+law school rankings [@RN5003] and UK school league tables [@RN5005] exemplify this sort of salience.
+Studies of quantification in the sociology of education tend to study consumer-facing metrics.
+Law school rankings can be conceived as a business-to-consumer product because the goal is to inform consumers (prospective law school students) and consumers pay for the product.
+By contrast, market research helps businesses identify/target customers and can be conceived as a business-to-business (B2B) product because businesses pay for market research.
+To our knowledge, scholarship on quantification in education has not examined how third-party market research classifies consumers.
+Additionally, scholarship on college access has failed to analyze third-party products that colleges use to identify and target prospective students.
+This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
+The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target.
+Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
+
+Scholarship on quantification suggests the mechanisms by which market research reproduces structural inequality [CITE CHUN; SOME SOCIOLOGY].
+Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
+Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification.
+However, this study provides inisight about the mechanism underlying the effect.
+By analyzing and simulating *Student Search Service* purchases that filter on Geomarkets we show how Geomarkets reproduce historical race-based inequality in college access [THIS LAST BIT FEELS LIKE THE GENERAL CONTRIBUTION TO SOCIOLOGY].
+
+<!--
 The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target.
@@
 Geo-markets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities.
-<!--
-CUT COMMENTS ON SALIENCE
+
 By contrast, many quantification projects -- for example, human rights indicators -- have "little oomph" [@RN5006, p.
@@
 FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created."
---> Scholarship on predictive analytics [@RN4975; @RN4849] suggests the mechanisms by which market research reproduces structural inequality.
-Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
-Unlike studies like @RN5005 we cannot show the effect of quantification.
-However, this study provides inisight about the mechanism underlying the effect.
-By analyzing and simulating *Student Search Service* purchases that filter on geo-markets we can show *how* the geo-markets reproduce historical class-based and race-based inequality in college access.
+-->
+
+# REVIEW OF EEPA 2024 PROLLY GOES IN "THE CASE"
+
+@RN5012 deconstruct the College Board *Student Search Service* product.
+Selection devices are procedures or routines for making categorical decisions based on input values [@RN4778].
+Student list products are selection devices that enable colleges to choose which prospective students they target or ignore based on search filters that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
+"Racialized inputs" are ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
+@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) are racialized inputs.
+Using a national sample of high school students, and also using data from actual student lists purchased by public universities, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
+<!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
+-->
+
+Student list products offer many geographic filters.
+Geographic borders are perhaps the most commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887], owing to extreme racial segregation in the U.S.
+In addition to offering geographic filters based on known geographic borders (e.g., zip code, county, CBSA, state), the College Board created new geographic borders that subsequently became filters in the *Student Search Service*.
+The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
+Geodemographic search filters can be read as the mimicry of corporate market logic [@RN4004] in that they apply the logic and methods of commercial geodemographic market segmentation products, such as PRIZM by the Claritas Corporation [@RN4957], to the case of college recruiting.
+
+Geomarket search filters, the focus of this manuscript, slice states and metropolitan areas into smaller, local recruiting markets.
+For example, geomarket "CA08" refers to "Alameda Country excluding Oakland." A university might purchase prospect contact information by filtering on this geographic region in combination with additional filters, such as year of high school graduation and SAT test score.
+When we initiated this manuscript, we asked ourselves, "where did geomarket filters come from?" The answer surprised us.
+We had imagined ourselves studying the application of the "data science revolution" to college recruiting, but it turned out we were studying the 1970s and 1980s!
+The Market Segment Model and geomarkets can be read as the application of social science research methods to "capture and quantify the phenomenon behind the folklore" [@RN4982, p.
+11] of college admissions officers.
+This work, comissioned by College Board, was subsequently turned into a market research tool -- The College Board Enrollment Planning Service -- that helps colleges decide where to look for prospective students.
 
@@
 <!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.-->
+
 As a member organization, several College Board products developed from joint collaborations with administrators or professors from member institutions.
@@
 x].
-In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.^[say that Pennsylvania Department of Education was also a partner; see page x of Zemsky] @RN4982[p.
+In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.[^3] @RN4982 [p.
 4] write that,
+
+[^3]: say that Pennsylvania Department of Education was also a partner; see page x of Zemsky
 
@@
 402\] definition of quantification as "the production and communication of numbers," the Market Segment Model is an effort to quantify and communicate the knowledge that admissions officers have about student demand.
-@RN4982[p.
+@RN4982 [p.
 4] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." They created "three types of boundaries" [@RN4982, p.
@@
 For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA.
-Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.[^5]
+Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.[^4]
 
-[^5]: @tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market.
+[^4]: @tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market.
 For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution.
@@
 @RN4982 summarize their correlational analyses using authoritative language.
-@RN4982[pp.
-34-35] asserts, "it is hard to overemphasize the importance of this statistical pattern.
-Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice." With respect to family income, @RN4982[p.
-33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982[p.
+@RN4982 \[pp.
+34-35\] asserts, "it is hard to overemphasize the importance of this statistical pattern.
+Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice." With respect to family income, @RN4982 [p.
+33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
 33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time." Further analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
@@
 @RN4982 recommend that colleges target geo-markets with desirable compositions of socioeconomic characteristics and student market segment.
-@RN4982[p.
+@RN4982 [p.
 44] describes a hypothetical college that wants to target regional and national students in New England and is considering whether to expend limited recruiting efforts in the Boston geo-market -- which was relatively low-income in 1980 -- or in more affluent geo-markets nearby:
@@
 
-```{=html}
 <!--
@@
 -->
-```
-Whereas @RN4982[Appendices B1-B3] identified 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S.
+
+Whereas @RN4982 [Appendices B1-B3] identified 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S.
 states, with 304 geo-markets in total.
@@
 Concrete information about how EPS created the remaining geomarkets is elusive.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^7]
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^5]
 
-[^7]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
+[^5]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
 Students from California don't all share the same college-going behaviors.
@@
 In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status.
-The language of @RN4982[pp.
+The language of @RN4982 [pp.
 11-12, quoted above] suggests that geomarket borders within large metropolitan areas may have been drawn in ways that separate affluent from less affluent communities.
@@
 *The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] shows that homophily is a core tenet of the Market Segment Model.
-The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^8] @RN4982[p.
-42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^9] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
+The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^6] @RN4982 [p.
+42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^7] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
 Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-@RN4982[p.
+@RN4982 [p.
 72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
 
-[^8]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+[^6]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
 Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another.
@@
 
-[^9]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
+[^7]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
 54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
@@
 @RN4982 [p.
-58] states that "correlations, do not simply predict actions; they also form them." In itself, the Market Segment Model is simply a stylized depiction [HIRSCHMAN 2016 STYLIZED FACTS] of the structure of college choice, one that does not consider class-based and race-based structures that produced these patterns.
+58] states that "correlations, do not simply predict actions; they also form them." In itself, the Market Segment Model is simply a stylized depiction \[HIRSCHMAN 2016 STYLIZED FACTS\] of the structure of college choice, one that does not consider class-based and race-based structures that produced these patterns.
 Once the College Board inscribed the Market Segment Model into the EPS and Student Search Service Products, these products are likely to amplify the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.
@@
 -->
-
 <!-- CUT PARAGRAPH: WHY GEOMARKETS ARE WEAPONS OF MATH DESTRUCTION

### 2024-10-18 text/old_spring2024/front_end.qmd
revising front end

--- 2024-07-08 text/front_end.qmd
@@
 
-In 1983, Zemsky and Oedel authored *The Structure of College Choice*, published by *The College Board.* This book develops the Market Segment Model and introduces local "community-based enrollment markets" (p.14) -- later called "geomarkets" -- which are the basis for the College Board market Enrollment Planning Service (EPS), a market research product.
-@RN4982[pp., 9-10] write that the Market Segment Model sought to quantify the knowledge that college admissions officers have about student demand:
+In 1983, Robert Zemsky and Penney Oedel authored *The Structure of College Choice*, published by *The College Board.* This book develops the Market Segment Model and introduces local "community-based enrollment markets" (p.14) -- later called "geomarkets" -- which are the basis for the College Board market Enrollment Planning Service (EPS), a market research product.
+@RN4982 [pp., 9-10] write that the Market Segment Model sought to quantify the knowledge that college admissions officers have about the geography of student demand:
 
@@
 Student lists contain the contact information of prospective students, the basic input for recruiting campaigns that target individual prospects using mail, email, and targeted social media.
-RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
 The Student Search Service database consists of College Board test-takers.
-Colleges pay a fee for each prospect (e.g., \$0.50 in 20XX) and control which prospect profiles through purchase by selecting search filters, such as high school graduation year, SAT score, AP test score, state, etc.
+Colleges pay a fee for each prospect (e.g., \$0.50 in 2021) and control which prospect profiles through purchase by selecting search filters, such as high school graduation year, SAT score, AP test score, state, etc.
 Geomarket filters enable colleges to include/exclude prospects from particular geomarkets.
@@
 This manuscript analyzes the Market Segment Model and College Board Geomarkets as a case study of quantification.
-We develop arguments from the broader literature on quantification [e.g.,] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
+We develop arguments from the broader literature on quantification \[e.g.,\] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
 Predictive analytics are based on correlation.
@@
 
-```{=html}
 <!--The enrollment management industry emphasizes market research.
 Market research is the process of producing information that informs decisions about product design or how to find customers.   -->
-```
 
 The enrollment management industry puts great stock in market research.
-@RN4986[p.
-17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].^[@RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
+@RN4986 [p.
+17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].[^1] However, most market research in the contemporary enrollment management industry is produced by third-party organizations.
+For example, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development \[e.g., @RN4741; @RN4728, but most only available to clients.
+EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
+
+[^1]: @RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
 In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited.
 In the chapter entitled "Travel," @RN3519 [p.
-69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."] However, most market research in the contemporary enrollment management industry is produced by third-party organizations.
-For example, Ruffalo-Noel Levitz and EAB both produce prodigious amounts of market research, some publicly available as a means of business development [e.g., @RN4741; @RN4728, but most only available to clients.
-EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
+69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."
 
 Despite being central to enrollment management for at least 40 years [@RN4986], the role of third-party market research has been ignored by scholarship on college access.
-Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935;@RN1814].
+Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935; @RN1814].
 However, market research informs where colleges look for students and which prospective students to target.
@@
 The case of Market Segment Model and College Board geomarkets shows how market research -- and its commodification into recruiting products -- is a mechanism for social reproduction.
-[LAST TWO SENTENCES ARE REDUNDANT BUT BUT ARE TRYING TO DO SLIGHTLY DIFFERENT THINGS AND I LIKE THEM BOTH]
+\[LAST TWO SENTENCES ARE REDUNDANT BUT BUT ARE TRYING TO DO SLIGHTLY DIFFERENT THINGS AND I LIKE THEM BOTH\]
 
@@
 -->
-```
+
 We situate our study amidst scholarship on within scholarship on quantification as a mechanism that reproduces inequality.
@@
 -->
-```
+
 In the interdisciplinary literature on quantification [@RN5008], the discussions of correlation and homophily by @RN4975 introduce concepts that are salient to the analysis of market research.
@@
 @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.[^4]
+In commercial social networks, such as Facebook, Twitter/X, and TikTok, homophily is programmed into the algorithms that create connections between users.[^2]
 
-[^4]: Thus, @RN4975 [p.
+[^2]: Thus, @RN4975 [p.
 82] writes, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together."

### 2024-10-18 scripts/create_stu_list.R

@@ -1,0 +1,1029 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_stu_list.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 8/22/2024
+## [ DESC ] < Create analysis datasets to simulate who is included/excluded when student list purchases filter on geomarket >
+################################################################################
+
+### SETTINGS
+rm(list = ls()) # remove all objects
+options(max.print=1000)
+#options(width = 160)
+# Set the scipen option to a high value to avoid scientific notation
+options(scipen = 999)
+
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+# get census api key
+#http://api.census.gov/data/key_signup.html
+# set census api key, and install for future use
+#census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+Sys.getenv("CENSUS_API_KEY") # retreive API key
+
+
+library(sf)
+# Enable caching for tigris
+options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+library(lwgeom) # library has checks/vixes for valid geometries
+library(leaflet)
+library(shiny)
+library(kableExtra)
+
+######## CREATE VECTORS OF EPS CODES FOR PARTICULAR METRO AREAS
+
+
+socal_eps_codes <- paste0("CA", 14:28) # socal 
+bay_eps_codes <- c(paste0("CA ", 4:9), paste0("CA", 10:11)) # bay area
+bay_eps_codes 
+ca_eps_codes <- c(paste0("CA ", 1:9), paste0("CA", 10:34))
+chi_eps_codes <- c(paste0("IL ", 7:9), paste0("IL", 10:13))
+chi_eps_codes
+
+cleveland_eps_codes <- paste0("OH ", 2:6) #
+cleveland_eps_codes
+
+philly_eps_codes <- paste0("PA ", 1:5) #
+philly_eps_codes
+
+nj_eps_codes <- c(paste0("NJ ", 1:9), paste0("NJ", 10:12))
+nj_eps_codes
+
+nj_metro_eps_codes <- c('NJ 2','NJ 4','NJ 5',paste0('NJ ', 7:9),paste0("NJ", 10:11))
+nj_metro_eps_codes
+
[…]

### 2024-10-18 scripts/input_tex_file.R

@@ -1,0 +1,15 @@
+\newpage
+\newgeometry{landscape, left=.5in, right=.5in, top=1in, bottom=1in}
+
+\input{../results/tables/socal_table.tex}
+
+\restoregeometry
+\newpage
+
+\newpage
+\newgeometry{landscape, left=.5in, right=.5in, top=1in, bottom=1in}
+
+\input{../results/tables/philly_table.tex}
+
+\restoregeometry
+\newpage

### 2024-10-18 scripts/interactive_maps_crystal.R

@@ -1,0 +1,325 @@
+library(leaflet)
+library(rgdal)
+library(rgeos)
+library(formattable)
+library(tidyverse)
+library(htmlwidgets)
+
+
+data_dir <- file.path('.', 'data')
+scripts_dir <- file.path('.', 'scripts')
+outputs_dir <- file.path('.', 'outputs', 'maps')
+
+
+# Load data
+
+load(file.path(data_dir, 'map_data_final.RData'))
+
+zip_data <- acs_zip %>% 
+  select(
+    zip_code, state_code, cbsa_1, cbsa_1_ratio, cbsatitle_1,
+    pop_total, medincome_2564, ends_with('_15_19_pct')
+  ) %>% 
+  mutate(
+    inc_brks = cut(
+      medincome_2564,
+      breaks = c(-1, 50000, 75000, 100000, 150000, 200000, 10000000),
+      labels = c('<$50k', '$50k-74k', '$75k-99k',  '$100k-149k', '$150k-199k', '$200k+')
+    ),
+    # Note: pop_native_15_19 = pop_amerindian_15_19 + pop_nativehawaii_15_19
+    pop_poc_15_19_pct = pop_black_15_19_pct + pop_hispanic_15_19_pct + pop_native_15_19_pct,
+    race_brks = cut(
+      pop_poc_15_19_pct,
+      breaks = c(-1, 20, 40, 60, 80, 90, 101),
+      labels = c('0-19%', '20-39%', '40-59%', '60-79%', '80-89%', '90-100%')
+    )
+  )
+
+zip_shp <- readOGR(file.path(data_dir, 'cb_2018_us_zcta510_500k', 'cb_2018_us_zcta510_500k.shp')) %>% 
+  merge(zip_data, by.x = 'ZCTA5CE10', by.y = 'zip_code', all.x = T)
+
+ccd_data <- ccd %>% 
+  filter(g12 >= 10, is_virtual == 0, updated_status %in% c('1', '3', '8')) %>% 
+  select(cbsa_1, ncessch, sch_name, sch_type, latitude, longitude, total_students, total_white, total_asian, total_black, total_hispanic, total_amerindian, total_nativehawaii, total_tworaces, total_unknown) %>% 
+  rename(
+    name = sch_name,
+    school_type = sch_type
+  ) %>% 
+  mutate(
+    school_type = recode(
+      school_type,
+      `1` = 'Regular',
+      `2` = 'Special education',
[…]

### 2024-10-18 scripts/leaflet_maps.R

@@ -1,6 +1,6 @@
 ################################################################################
 ## [ PROJ ] < College Board Geomarket >
-## [ FILE ] < mapping_eps.R >
+## [ FILE ] < leaflet_maps.R >
 ## [ AUTH ] < Ozan Jaquette >
 ## [ INIT ] < 8/22/2024
 ## [ DESC ] < work with EPS-level sf objects and other objects to create interactive maps of characteristics associated with geomarkets >
@@ -196,6 +196,14 @@ chi_eps_codes
 cleveland_eps_codes <- paste0("OH ", 2:6) #
 cleveland_eps_codes
 
+philly_eps_codes <- paste0("PA ", 1:5) #
+philly_eps_codes
+
+nj_eps_codes <- c(paste0("NJ ", 1:9), paste0("NJ", 10:12))
+nj_eps_codes
+
+nj_metro_eps_codes <- c('NJ 2','NJ 4','NJ 5',paste0('NJ ', 7:9),paste0("NJ", 10:11))
+nj_metro_eps_codes
 
 ############## CREATING GEOMARKET LEVEL TABLES
 #######################
@@ -221,83 +229,6 @@ table_var_names <- list(
 # Define table_vars in the desired order
 table_varlist <- c("pct_nhisp_white", "pct_nhisp_black", "pct_hisp_all", "pct_nhisp_api", "pct_nhisp_native", "pct_nhisp_multi", "med_inc_house", "mean_inc_house", "pct_pov_yes", "pct_edu_baplus_all")
 
-create_eps_table <- function(eps_codes, table_vars) {
-  # Human-readable names for the variables (you can modify this list based on the variables passed in)
-  table_var_names <- list(
-    "pct_hisp_all" = "% Hispanic",
-    "mean_inc_house" = "Mean income",
-    "med_inc_house" = "Median income",
-    "pct_nhisp_asian" = "% Asian, non-Hispanic",
-    "pct_nhisp_white" = "% White, non-Hispanic",
-    "pct_nhisp_black" = "% Black, non-Hispanic",
-    "pct_nhisp_api" = "% API, non-Hispanic", 
-    "pct_nhisp_native" = "% Native, non-Hispanic",
-    "pct_pov_yes" = "% in poverty",
-    "pct_edu_baplus_all" = "% with BA+"
-  )
-  
-  # Main table creation pipeline
-  eps_table <- allyr_anal_tract_sf %>% 
-    as.data.frame() %>%
[…]

### 2024-10-18 scripts/sim-table that works.R

@@ -1,0 +1,85 @@
+
+format_eps_sim_table_simple <- function(df) {
+  library(dplyr)
+  library(kableExtra)
+  
+  # Extract the individual tables
+  counts <- df$count_table
+  col_pct <- df$col_pct_table
+  row_pct <- df$row_pct_table
+  
+  # Rename columns for consistency
+  counts_formatted <- counts %>%
+    rename(
+      `EPS Codename` = eps_codename,
+      `Total Students` = stu_all,
+      `Race Known` = stu_race_known,
+      White = stu_white,
+      Asian = stu_asian,
+      Black = stu_black,
+      Hispanic = stu_hispanic,
+      `Two Races` = stu_tworaces
+    ) %>%
+    mutate(Section = "1. Counts") %>%
+    select(Section, everything())
+  
+  col_pct_formatted <- col_pct %>%
+    rename(
+      `EPS Codename` = eps_codename,
+      `Total Students` = stu_all,
+      `Race Known` = stu_race_known,
+      White = col_pct_white,
+      Asian = col_pct_asian,
+      Black = col_pct_black,
+      Hispanic = col_pct_hispanic,
+      `Two Races` = col_pct_tworaces
+    ) %>%
+    mutate(Section = "2. Column Percentages") %>%
+    select(Section, everything())
+  
+  row_pct_formatted <- row_pct %>%
+    rename(
+      `EPS Codename` = eps_codename,
+      `Total Students` = stu_all,
+      `Race Known` = stu_race_known,
+      White = row_pct_white,
+      Asian = row_pct_asian,
+      Black = row_pct_black,
+      Hispanic = row_pct_hispanic,
+      `Two Races` = row_pct_tworaces
+    ) %>%
+    mutate(Section = "3. Row Percentages") %>%
+    select(Section, everything())
+  
+  # Combine all sections into one table
+  full_table <- bind_rows(counts_formatted, col_pct_formatted, row_pct_formatted)
+  
+  # Create a LaTeX-compatible table using kable
+  table_kable <- full_table %>%
+    kbl(
+      format = "latex",
+      booktabs = TRUE,
+      escape = FALSE,
+      col.names = c("Section", "EPS Codename", "Total Students", "Race Known", 
+                    "White", "Asian", "Black", "Hispanic", "Two Races"),
+      caption = "Characteristics of Geomarkets around Philadelphia \\label{tab:philly_eps_table_summary}"
+    ) %>%
+    kable_styling(
[…]

### 2025-01-15 soc_of_ed/soc_of_ed_geomarket.qmd
working through results, homie

@@
 ---
-title: "Structuring College Access: The Market Segment Model and College Board Geo-markets"
+#title: "Structuring College Access: The Market Segment Model and College Board Geomarkets"
 #author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
@@
     keep_tex: true  # Optional: Keeps the intermediate .tex file for debugging
+    titlepage: false
 editor: visual
-geometry: "margin=1in"
+geometry: "top=0.75in, bottom=0.5in, left=1in, right=1in"
 fontsize: 11pt
@@
     }
+  - \renewcommand{\headrulewidth}{0pt}  % Remove the top header line
+  #- \renewcommand{\footrulewidth}{0.4pt}  # Keeps the bottom line    
   - \usepackage{booktabs}
@@
 
+<!--  
 \thispagestyle{empty} 
@@
 \setcounter{page}{1}
+Please submit a detailed 2-page précis that includes a brief discussion of the topic, pertinent literature and
+theory, research methods, findings, and study significance
 
-# PROPOSAL
+EXPECTED OUTLINE
+- topic and RQs
+- literature and theory
+- methods
+- findings
+- significance
 
-**BACKGROUND, RESEARCH QUESTIONS.** In 1983, Zemsky and Oedel authored *The Structure of College Choice*.
-Based on an analysis of 1980 SAT score-sending behavior and published by *The College Board*, @RN4982 developed the Market Segment Model and created "Geomarkets." The Market Segment Model predicts how demand for a particular institution varies across geographic regions based on household/student characteristics of geographic regions.
+THEME
+- What can we do about it? Sociologists as change agents
 
-Geomarkets are geographic borders that disaggregate states and large metropolitan areas into "community-based enrollment markets" [@RN4982, p.
+HOW TO CONNECT WITH THE THEME
+- sociology of organizations knows that regulating an industry requires understanding the organizational field, which includes third-party vendors and products.
+these third party vendors and agents structure college access in ways that benefit historically privileged groups and contribute to accumulation of capital
+- we must understand the field before we can regulate it; CFPB has 
+Education policy has been dominated by economists. economics doesn't have the theoretical concepts necessary 
+-->
+
+\setlength{\headsep}{0pt}
+\setlength{\headheight}{0pt}
+\begin{center} Structuring College Access: The Market Segment Model and College Board Geomarkets \end{center}
+\vspace{-0.5cm}
+
+# PRECIS
+
+**BACKGROUND.** In 1983, Zemsky and Oedel authored *The Structure of College Choice*.
+Based on an analysis of 1980 SAT score-sending behavior and published by *The College Board*, @RN4982 developed the Market Segment Model and created "Geomarkets." The Market Segment Model predicts how student demand for a particular college varies across local geographic markets based on the characteristics of students in those local markets.
+
+Geomarkets are geographic areas that divide states and large metropolitan areas into "community-based enrollment markets" [@RN4982, p.
 14] for recruiting purposes.
 @fig-nyc-geomarkets shows New York City area Geomarkets.
-Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
-EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.[^1]
+The Market Segment Model and Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
+EPS software recommends which Geomarkets a college should recruit from and which high schools within targeted Geomarkets a college should visit.[^1]
 
@@
 Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc.
-Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets.
-An analysis of 830 student lists purchased by 14 public universities found that 16% of purchases filtered on Geomarket [@RN5012].
+Geomarket filters enable colleges to target prospects living in particular Geomarkets.^[An analysis of 830 student lists purchased by 14 public universities found that 16% of purchases filtered on Geomarket [@RN5012].]
 
-[^2]: @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+[^2]: @RN4895 reported that 87% of private and 86% of public four-year institutions purchase lists.
 
-We argue that Geomarket borders are substantially correlated with class and race.
-This correlation is partly intentional.
-@RN4982 viewed student demand as a function of class.
-@RN4982 sought to create Geomarkets that contained many high SES students willing to attend colleges far from home, for example "CT 3--Fairfield County." When colleges use the Market Segment Model, "the result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice" [@RN4982, p.
-44].
-Considering the nature of U.S.
-segregation [@rothstein2017color], Geomarkets may also be correlated with race.
+**THEORY & RQs.** We analyze the Market Segment Model and Geomarkets using scholarship on quantification [e.g., @RN5003], particularly the discussions of correlation and homophily from @RN4975.
+Predictive analytics are based on correlation.
+The 1980 SAT score-sending data analyzed by @RN4982 can be conceived as "training data," which was used to define four "market segments" of students: "local" students, who send most SAT scores to colleges in their Geomarket; "in-state" students, who primarily send scores to colleges in their state; "regional" students, who send most scores to colleges outside their state but in their region (e.g., New England); and "national" students.
+@RN4982 analyzed the student characteristics correlated with student market segment and concluded that student demand was primarily a function of class.
+The overarching thesis of the Market Segment Model is homophily, the idea that similar colleges compete with one another for similar students.
+@RN4982[p.
+42-45] states that selective colleges primarily draw from students in the "regional" and "national" segments and, therefore, their recruiting visits and student list purchases should focus on Geomarkets with large numbers of affluent, highly educated households.
 
-<!--  
-we argue that utilizing Geomarket filters to purchase student lists can contribute to socioeconomic and racial inequality on college opportunity.
-
-Using data about past correlations -- here, Geomarket borders created with reference to 1980 SAT score sending behaviors -- to make future recommendations amplifies the effects of historical structural inequalities.
--->
-
-We draw from "quantification" scholarship in sociology [@RN5007; @RN5005] and data studies [@RN4849; @RN4975].
-To the extent that Geomarket borders are correlated with class and race, student list purchases that filter on Geomarkets can yield systematic socioeconomic and racial exclusion in college opportunity [@RN4887].
+When we use data on past correlations to make recommendations about the future, we amplify the effects of historic structural inequality [@RN4794].
+A snapshot of student demand in 1980, itself a consequence of historical structural inequality, became inscribed into the EPS and Student Search Service products used by colleges to plan and implement recruiting strategy.
+The result is a supply-side that reinforces structural inequalities observed on the demand-side.
 We address two research questions:
@@
 
-Scholarship has not observed the mechanisms by which enrollment management products structure college access, funneling certain students to certain institutions.
-Showing how these products facilitate exclusion introduces new avenues for scholarship and policy discourse.
+**SIGNIFICANCE.** Scholarship on quantification in the sociology of education [e.g.,  @RN2356;@RN5005] has not analyzed third-party products that colleges use to classify and identify prospective students.
+More broadly, the concept of "organizational fields" includes all organizations that "constitute a recognized area of institutional life"  [@RN527, p.
+143].
+In the enrollment management industry, third-party vendors and their products structure college access in ways that funnel certain kinds of students to certain kinds of institutions.
+Higher education policy over the last 20 years -- dominated by economists -- has focused on regulating consumers (students) and direct providers (colleges).
+However, federal regulatory agencies (e.g., FTC, CFPB) are increasingly interested in regulating the entire organizational field because they know that third-party oligopolies can wield market power over consumers and providers [@philippon2019great].
+Therefore, scholarship from sociology should guide which problems and actors become the focus of future federal regulation.
 
@@
 We obtained shapefiles for College Board Geomarkets from a 2012 *R-bloggers* post.
-We assigned census tracts to Geomarkets by implementing a partial spatial join.[^3]
-
-[^3]: Census tracts that intersected multiple Geomarkets were partially assigned based on the proportion of land area.
-Currently, our spatial join for the 2020 ACS is causing some invalid geometries that we will fix prior to AEFP.
+We assigned census tracts to Geomarkets by implementing a partial spatial join.
 
@@
 For 414 College Board student list purchases, associated with about 2.6 million prospects, we received both (1) the order summary -- showing which search filters were utilized -- and (2) the de-identified prospect-level data, including race, ethnicity, zip code, and high school.
-@fig-asu-order-448922 shows the partial order summary, which we converted to tabular data, for a purchase by Arizona State University that filtered on the 2020 high school graduating class, PSAT scores of 1070 to 1180, and selected Geomarkets.[^4]
-
-[^4]: Analyses for RQ2 additionally utilize data on public and private high schools from NCES.
+@fig-asu-order-448922 shows the partial order summary, which we converted to tabular data, for a purchase by Arizona State University that filtered on the 2020 high school graduating class, PSAT scores of 1070 to 1180, and selected Geomarkets.
 
@@
 We produce interactive maps at the census tract-level to show more granular variation within and between Geomarkets.
-We address RQ2 by analyzing actual student list purchases that encompass all Geomarkets around a given metropolitan area, showing which prospects would have been excluded/excluded had the purchase filtered on particular Geomarkets.
+We address RQ2 by analyzing actual student list purchases that encompass all Geomarkets around a given metropolitan area, showing which prospects would have been excluded if had the purchase had filtered on particular Geomarkets.
 
-**PRELIMINARY FINDINGS**. We are currently writing functions that create tables and/or maps for the metropolitan area chosen by the function call.
-
+**PRELIMINARY FINDINGS**.
 For RQ1, Table \ref{tab:chi_eps_table_summary} shows racial and socioeconomic characteristics for Geomarkets around Chicago.
@@
 These prospects would have been excluded had the purchase filtered on Geomarkets and not selected that Geomarket.
-@tbl-philly-sim-count, @tbl-philly-sim-col-pct, and @tbl-philly-sim-row-pct, respectively, show counts, column percentages, and row percentages for the Philadelphia metropolitan area, based on a purchase that filters on students from the 2020 high school class who scored between 1070 and 1180 on the PSAT.
+@tbl-philly-sim-count, @tbl-philly-sim-col-pct, and @tbl-philly-sim-row-pct, respectively, show counts, column percentages, and row percentages for the Philadelphia metropolitan area, based on a purchase that filtered on students from the 2020 high school class who scored between 1070 and 1180 on the PSAT.
 @tbl-philly-sim-col-pct shows that "PA5--Philadelphia County" contains 15% of all purchased prospect profiles but 44% of Black prospects, indicating that opportunity for Black students is sensitive to the inclusion of this Geomarket.
 @fig-philly-black-prospects is a screenshot of a (very) preliminary interactive map for this purchase, with marker size indicating the number of purchased prospect profiles who attended that high school and separate markers by race/ethnicity.
-<!--
-The size of the high school marker is a function of the number of prospects purchased and map users wil be able to select the race/ethnicity of purchased prospects as well as census variables that determine the chloropleth map.
--->
 

### 2025-01-15 text/front_end_v2.qmd
working through results, homie

@@
 ---
-title: "Structuring College Access: The Market Segment Model and College Board Geo-markets"
+title: "Structuring College Access: The Market Segment Model and College Board Geomarkets"
 author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
@@
 Therefore, the key to recruitment is identifying territories where large numbers of your class live.
-@RN4982[p.
+@RN4982 [p.
 11] describe the Market Segment Model as an effort to "quantify" the knowledge that college admissions officers have about the geography of student demand:
@@
 First, Geomarkets became a principle for how many -- if -- admissions offices organized their labor.
-In states where a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory [CITE].
+In states where a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory \[CITE\].
 Geomarkets were created to mirror the territories of admissions officers and, in turn, admissions offices structured their territories around Geomarkets.
@@
 This omission is surprising because the sociology of education is concerned with sorting [@RN4873], particularly sorting into college.
-Traditionally, most scholarship -- and certainly most scholarship in economics [HOXBY?
-CHETTY] -- views inequality in college access as a function households or k-12 schools, which represent the demand-side of higher education.
+Traditionally, most scholarship -- and certainly most scholarship in economics \[HOXBY?
+CHETTY\] -- views inequality in college access as a function households or k-12 schools, which represent the demand-side of higher education.
 More recently, scholarship in the sociology of education takes seriously the idea that colleges are not passive recipients of applications, but rather they expend substantial resources soliciting demand from desirable applicants [@RN3519; @RN4774; @RN4324].
@@
 This manuscript analyzes one of those mechanisms, The Market Segment Model and College Board Geomarkets, as a case study of quantification.
-We develop arguments from the broader literature on quantification \[e.g.,\] [@RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
+We develop arguments from the broader literature on quantification [e.g., @RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
 Predictive analytics are based on correlation.
@@
 The Market Segment Model argues that homophily -- actors that share characteristics are likely to form connections -- is the organizing principle of student demand and competition between colleges.
-@RN4982[p.
+@RN4982 [p.
 72] state that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students." This snapshot of student demand in 1980, itself a consequence historic structural inequality, was programmed into the way admissions offices organize labor and programmed into the EPS and Student Service Service products that colleges utilize to identify and target prospective students.
 The result is a supply-side that reinforces structural inequalities observed on the demand-side.
+
+\[ADD TEXT SOMEWHERE AROUND HERE ABOUT WHY THIS TOPIC IS SIGNIFICANT AND IMPORTANT FOR CURRENT/FUTURE POLICY DEBATES: Why market segment model important.
+Software as service products dominate em industry and structure college access.
+Look at eab machine learning product But rarely do we see the recipe underneath these products.
+The book structuring college access is that recipe for EPS software\]
 
@@
 
+```{=html}
 <!-- 
@@
 -->
+```
 
@@
 Finally, we discuss implications for scholarship and for policy.
+
+A note from RESULTS that I texted to Bastedo on 12/19/2024
+
+-   But these geomarkets are pretty fucked up.
+They were designed to capture class.
+I see lots of orders that condition on test scores and rich geomarkets.
+But I know from similar purchases that don’t condition on geomarket, that the high scoring first gen Asian and Hispanic students tend to come from the poor geomarkets that get excluded from a lot of purchases.
+By contrast, The Asian and Hispanic students that do get included in purchases that target rich geomarkets are usually not first gen
 
@@
 The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
-Economists often investigate financial aid leveraging, which seeks to convert admits to enrolled students [CITE].
+Economists often investigate financial aid leveraging, which seeks to convert admits to enrolled students \[CITE\].
 <!--  Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935; @RN1814].
@@
 Law school rankings were developed as a means of informing prospective students about the relative quality of different programs.
-Speaking to *reactivity* effects, rankings had the intended effect of affecting where students apply and enroll [CITE].
+Speaking to *reactivity* effects, rankings had the intended effect of affecting where students apply and enroll \[CITE\].
 However, rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
@@
 
+```{=html}
 <!-- THIS IS THE CUT TRANSITION PARAGRAPH
@@
 224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].^[An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
+This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^2]
+
+[^2]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
 Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
 Note that predictive models such as the LSI-R model are not merely used for social science purposes.
-Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.]
+Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.
 
@@
 @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-In commercial social networks (e.g., Faceboo, X, TikTok), homophily is more than an assumption; homophily is programmed into the algorithms that create connections between users.[^2]
+In commercial social networks (e.g., Faceboo, X, TikTok), homophily is more than an assumption; homophily is programmed into the algorithms that create connections between users.[^3]
 
-[^2]: Thus, @RN4975 [p.
+[^3]: Thus, @RN4975 [p.
 82] writes, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together."
@@
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].^[For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.] Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
 <!--
@@
 
+[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
+
 Reviewing scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
@@
 This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
-The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target.
+The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
@@
 
+```{=html}
 <!--
-The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which geo-markets to target.
+The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
 @RN5021 reports that in 1995, 37% of 4-year public universities and 49% of 4-year private universities used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
-Geo-markets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities.
+Geomarkets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities.
 
@@
 -->
-
-# REVIEW OF EEPA 2024 PROLLY GOES IN "THE CASE"
-
-@RN5012 deconstruct the College Board *Student Search Service* product.
-Selection devices are procedures or routines for making categorical decisions based on input values [@RN4778].
-Student list products are selection devices that enable colleges to choose which prospective students they target or ignore based on search filters that incorporate search filters (e.g., high school graduating class, state) which enable colleges to select prospective students.
-"Racialized inputs" are ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
-@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) are racialized inputs.
-Using a national sample of high school students, and also using data from actual student lists purchased by public universities, they show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
-<!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
--->
-
-Student list products offer many geographic filters.
-Geographic borders are perhaps the most commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887], owing to extreme racial segregation in the U.S.
-In addition to offering geographic filters based on known geographic borders (e.g., zip code, county, CBSA, state), the College Board created new geographic borders that subsequently became filters in the *Student Search Service*.
-The geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
-Geodemographic search filters can be read as the mimicry of corporate market logic [@RN4004] in that they apply the logic and methods of commercial geodemographic market segmentation products, such as PRIZM by the Claritas Corporation [@RN4957], to the case of college recruiting.
-
-Geomarket search filters, the focus of this manuscript, slice states and metropolitan areas into smaller, local recruiting markets.
-For example, geomarket "CA08" refers to "Alameda Country excluding Oakland." A university might purchase prospect contact information by filtering on this geographic region in combination with additional filters, such as year of high school graduation and SAT test score.
-When we initiated this manuscript, we asked ourselves, "where did geomarket filters come from?" The answer surprised us.
-We had imagined ourselves studying the application of the "data science revolution" to college recruiting, but it turned out we were studying the 1970s and 1980s!
-The Market Segment Model and geomarkets can be read as the application of social science research methods to "capture and quantify the phenomenon behind the folklore" [@RN4982, p.
-11] of college admissions officers.
-This work, comissioned by College Board, was subsequently turned into a market research tool -- The College Board Enrollment Planning Service -- that helps colleges decide where to look for prospective students.
+```
 
@@
 
-<!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.-->
+```{=html}
+<!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.
 
@@
 
-In 1978, Robert Zemsky, a University of Pennsylvania professor, was asked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
+The Market Segment Model was an effort to quantify and communicate the knowledge that admissions officers have about student demand. 
+-->
+```
+
+**Creating the Market Segment Model.** In 1978, Robert Zemsky, a University of Pennsylvania professor, was asked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
 x].
-To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE).
-As the project project became more ambitious, the researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in tehse analyses was the state' [@RN4982, p.
+To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
+The researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in these analyses was the state' [@RN4982, p.
 x].
-In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.[^3] @RN4982 [p.
+In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.
+@RN4982 [p.
 4] write that,
-
-[^3]: say that Pennsylvania Department of Education was also a partner; see page x of Zemsky
 
@@
 
-**Creating the Market Segment Model.** Considering the @RN5007 \[p.
-402\] definition of quantification as "the production and communication of numbers," the Market Segment Model is an effort to quantify and communicate the knowledge that admissions officers have about student demand.
-@RN4982 [p.
-4] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." They created "three types of boundaries" [@RN4982, p.
+@RN4982 \[p.
+4\] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." Quoting an admissions officer, @RN4982 [p.
+11] write, "'There are only three kinds of colege-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, [@RN4982] created "three types of boundaries" [@RN4982, p.
 11], region, state, and community.
 The three regions were New England, Middle States, and the South.
-Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" \[in New England, Middle States, and the South\].
+Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets."
 
-These enrollment markets, later called "geomarkets" by the College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
+These enrollment markets, later called "Geomarkets" by the College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors [@RN4982].
 We know little about how geomarkets were created.
@@
 The two primary outputs of the market segment model are the (1) Market Segment Profile and (2) the Institutional Profile.
-The Market Segment Profile, which was produced separately for each of the 143 geo-markets, provides information about the number and characteristics of students in each market segment.
-@tbl-market-segment-characteristics reproduces a partial, simplified version of the bottom panel of @RN4982 [Figure 2.1] which represents the Market Segment Profile for "Connecticut Market 3: Fairfield County." @tbl-market-segment-characteristics has separate columns for each market segment (local, in-state, regional, national) and rows show the number of test-takers and characteristics of test-takers.
-For example, there were 550 "local" students in Fairfield county and these students submitted SAT scores to 2.8 postsecondary institutions on average.
-By contrast, there were 1,664 "regional" students who submitted test scores to 4.8 institutions on average.
-For local students, 11.2% had family income greater than \$35,000 and 9% had both parents with a BA.
-For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA.
-Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.[^4]
+Appendix A describes these outputs in more detail.
+The Market Segment Profile provides information about the number and characteristics of students in each market segment.
+This information is produced separately for each of the 143 Geomarkets.
+For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
+The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+A university could obtain this for their own institution or for a competitor.
+Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
 
-[^4]: @tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market.
-For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution.
-Of the 1,199 in-state students, 757 (63.1) sent scores to institution #1, a public doctoral granting institution, and 515 (43.0%) sent scores to institution #2, a public master's granting institution.
+**Enrollment Planning Service**.
+In 1984, College Board created the Enrollment Planning Service (EPS) based on the Market Segment Model [@RN4987; @RN4980].
+EPS was a simple software-as-service platform that informed college decisions about recruiting by essentially recreating the analyses of @RN4982.
+Institutions purchasing EPS could obtain the Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors -- for each local market.
+Typical marketing material describes EPS as,
 
-INSERT @tbl-market-segment-characteristics ABOUT HERE
+> The marketing software that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
+With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
+Focus your valuable time and resources on the right prospects [@RN5019].
 
-The *Institutional Profile* describes students who send scores to a particular institution and which majors these students are interested in.
-For each institution, institutional profiles are created separately for students from a particular local market, for all students in a state, or all students in a region.
-@tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the institutional profile of an anonymous institution for students from Fairfield County, CT.
-@tbl-sample-institutional-profile shows that 58 in-state students submitted SAT scores to the institution.
-These 58 students represented 4.8% of the total 1,199 in-state students from Fairfield County.
-69 regional students sent scores to the institution, representing 4.1% of all 1,664 regional test-takers.
-Of these 69 regional students, 35 expressed interest in majoring in the liberal arts.
-These 35 students represent 7.0% of all regional students from Fairfield County who expressed interest in the liberal arts.
-
-INSERT @tbl-sample-institutional-profile ABOUT HERE
-
-@RN4982 argue that the Institutional Profile and the Market Segment Profile enable admissions officers to know where to look for students and which institutions are competing for those students (p.
-25):
-
-> The Institutional Profile and the Market Segment Profile quantify the admission officers' intuitive grasp of market structure.
-Structure here carries a dual meaning, connoting both the structure of student choice and the structure of institutional competition\ldots This two-sided interpretation furnishes the essential framework for planning by individual colleges and universities\ldots To draw effectively on its own natural constituency, a college not only must contact the "right" kind of students --- that is, students who are predisposed toward that type of institution -- but also must persuade them of its special character.
-This means knowing the competition as well as the clientele.
-
-**Correlations**.
-Chapter 3 of @RN4982 -- "A Sense of Place: Students, Families, and Communities" -- identifies the student characteristics associated with being in the local, in-state, regional, or national market segment and examines the extent to which these relationships hold across states and local markets.
-The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
-[@RN4982, p.
-33] state that these four variables "reflect the basic social patterns of the nation.
-It would have been surprising if these were not the four social variables that best explained the patterns of college choice."
-
-@RN4982 summarize their correlational analyses using authoritative language.
-@RN4982 \[pp.
-34-35\] asserts, "it is hard to overemphasize the importance of this statistical pattern.
-Social data seldom line up as we expect \ldots The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice." With respect to family income, @RN4982 [p.
-33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time." Further analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
-Considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model implicitly and explicitly states that student demand for higher education is mostly a function of social origin.
-@RN4982 also find that the SAT score, parental education, and family income predict student score-sending behavior at the local geo-market level.
-
-However, geomarkets differ in the relative abundance of students with particular socioeconomic characteristics and this has practical implications for recruiting.
-@RN4982 recommend that colleges target geo-markets with desirable compositions of socioeconomic characteristics and student market segment.
-@RN4982 [p.
-44] describes a hypothetical college that wants to target regional and national students in New England and is considering whether to expend limited recruiting efforts in the Boston geo-market -- which was relatively low-income in 1980 -- or in more affluent geo-markets nearby:
-
-> Where would you concentrate your energies?
-Ideally, you would seek communities with a high proportion of students already predisposed toward institutions such as your own.
-The Market Segment Model would provide this information through segment percentages for the community in question.
-Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
-Your efforts would surely be better directed toward three of the four communities in the bottom band, Manchester, Hartford, and Fairfield County.
-Simply by knowing a little bit about the students' backgrounds and academic records you could quickly focus your attention on those most likely to consider your kind of institution.
-Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own."
-
-**EPS and College Board Geomarkets**.
-In 1984, College Board created the Enrollment Planning Service (EPS) based on the Market Segment Model [@RN4987; @RN4980].
-Each institution purchasing EPS information services would obtain access to the Market Segment Report for each local market.
-They also received the Institutional Profile -- their own and that of competitors -- for each local market.
+```{=html}
+<!--
 Promotional material highlights the themes of "locate the students you want" [@RN5019_2003] and "focus your valuable time and resources on the right prospects" [@RN5019].
@@
 > the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
-You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, Zip codes, high schools and international regions.
-EPS provides you with comprehensive reports on your markets, your position in those markets, and your copetition."
-
-<!--
-@RN5019 [2005]
-> The marketing software that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
-With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
-Focus your valuable time and resources on the right prospects
+You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, zip codes, high schools and international regions.
+EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition."
 
@@
 -->
+```
 
-Whereas @RN4982 [Appendices B1-B3] identified 143 geo-markets covering the New England, Middle States, and South region, EPS created geo-markets for the remaining U.S.
-states, with 304 geo-markets in total.
-@RN4988 shows the College Board geo-markets circa 2023.
-Interestingly, the geo-markets for New England, Middle States, and the South remain nearly identical to those developed by @RN4982.
-Concrete information about how EPS created the remaining geomarkets is elusive.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^5]
+Whereas @RN4982 identified 143 Geomarkets covering the New England, Middle States, and South region, EPS created Geomarkets for the remaining U.S.
+states, with 304 Geomarkets in total.
+@RN4988 shows the contemporary Geomarkets.
+Interestingly, the Geomarkets for New England, Middle States, and the South remain nearly identical to those developed by @RN4982.
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^5].
+Furthermore, this material encourages colleges to, first, identify which Geomarkets they will target.
+For example, Software documentation by @RN5017 states that
 
@@
 
-Although how we lack clarity about how geomarkets were created, we expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
+> EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful.
+
+**Correlations**.
+Chapter 3 of @RN4982 identifies the student characteristics associated with being in the local, in-state, regional, or national market segment.
+The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
+[@RN4982] state that these four variables "reflect the basic social patterns of the nation.
+It would have been surprising if these were not the four social variables that best explained the patterns of college choice" (p.
+33) and "The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice" (pp.
+34-35).
+\[QUOTES REDUNDANT\]
+
+The primary takeaway of Chapter 3 is that student demand for higher education is mostly a function of social origin.
+For example, @RN4982 \[p.
+33\] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
+33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^6]
+
+[^6]: The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
+Additionally, analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
+However, prior research shows that SAT scores are substantially a function of social origin [@RN1273].
+
+However, geomarkets differ in the relative abundance of students with particular socioeconomic characteristics and this has practical implications for recruiting.
+@RN4982 recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics and student market segment.
+
+**Homophily**.
+Chapter 4 of @RN4982 -- "The Company We Keep: Colleges and Their Competition" -- shows that homophily is a core assumption of the Market Segment Model.
+The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^7] @RN4982 [p.
+42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^8] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
+Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
+@RN4982 [p.
+72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+
+[^7]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another.
+These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
+
+[^8]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
+54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
+For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
+
+> The layering of collegiate competition is primarily a socioeconomic layering.
+The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
+Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
+
+Research question 1 asks, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
+We expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
 Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
@@
 Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race once.
-However, U.S.
-cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779; @rothstein2017color].
+U.S.
+cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779].
 Structures built upon racialized structures are racialized structures.
-@RN4975 argues that unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity.
+Unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
 Second, geomarket borders may have been drawn along class divides.
-There is a strong correlation between race and wealth [@RN4884].
-Third, geo-market borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
+A strong correlation exists between race and wealth [@RN4884].
+Third, Geomarket borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
 Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
@@
 
-We argue that the Market Segment Model, geomarkets, and their commodification within EPS exemplify concerns about the application of correlational analysis.
+The Market Segment Model and Geomarkets raise concerns common to the application of correlations.
 The data on 1980 SAT score-sending behavior analyzed by @RN4982 can be conceived as training data.
-These training data were used to define geomarkets (e.g., how many geomarkets in a metropolitan area is partially a function of the number of test-takers) and to identify student characteristics associated with the local, in-state, regional, and national market segments.
-These analyses found that student demand for different kinds of colleges was largely a function of social origin, and that student demand within a geo-market was largely a function of the class composition of the geo-market.
+These training data defined the Geomarkets in that the number of geomarkets created in a metropolitan area was based on the number of test-takers [@RN4982].
+The training data identified student characteristics correlated with the local, in-state, regional, and national market segments.
+These analyses found that student demand for different kinds of colleges was largely a function of social origin, and that student demand within a Geomarket was largely a function of the class composition of the Geomarket.
 However, @RN4982 did not interrogate the historical and contemporary structural inequalities that produced observed patterns of social stratification.
 Instead, the Market Segment Model encouraged colleges to use this snapshot of student demand as the basis for future decisions about which communities and prospective students should be targeted by college recruiting campaigns.
-This process exeemplifies the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849] whereby predicting the behavior of future individuals based on the behavior of past individuals "threatens to reify and reporduce existing inequalities" [@RN4794, p.
+This process exeemplifies the "ratchet effect" [@RN4851], whereby predicting the behavior of future individuals based on the behavior of past individuals "threatens to reify and reporduce existing inequalities" [@RN4794, p.
 224].
-This process also exemplifies concerns about proxy variables whereby selective colleges view privileged social class as a proxy for desirable students and this creates an incentive to focus recruiting efforts on communities that have these characteristics.
+This process also exemplifies concerns about proxy variables, whereby selective colleges view privileged social class as a proxy for desirable students.
+This creates an incentive to focus recruiting efforts on communities that have these characteristics.
 @RN4982 [p.
-44] corroborates these concerns:
+44] discuss these concerns:
 
@@
 
-**Homophily and Utilization**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] shows that homophily is a core tenet of the Market Segment Model.
-The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^6] @RN4982 [p.
-42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^7] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
-Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-@RN4982 [p.
-72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+Commodification of the Market Segment Model and Geomarkets "engineers homophily" [@RN4975].
+In itself, the Market Segment model is a "stylized fact" [@RN5025] about the structure of student demand, one that does not consider historic structural inequalities that produced observed patterns.
+Once the College Board inscribed the Market Segment Model and Geomarkets into the EPS product, this product may amplify the effect of historic structural inequalities on future opportunity structure; a snapshot of extant stratification in student demand becomes the basis for college decisions about which local markets to target.
 
-[^6]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
-Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
-
-[^7]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
-54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
-
-<!--[^10]: These patterns are similar to different offers for credit products in The Ordinal Society [@RN5015], except @RN5015 show that the matching of consumers to products is engineered by firms rather than naturally occurring.-->
-
-> We now know why students so seldom speak of their own social or family backgrounds in explaining how they go about choosing a college.
-They have no need to.
-Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
-The layering of collegiate competition is primarily a socioeconomic layering.
-The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
-Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
-
-Although this perspective is broadly consistent with mid-20th Century scholarship on social mobility [@RN915; @RN1273], the Market Segment Model was the basis for the EPS market research product.
-Later, geomarkets were incorporated into the *Student Search Service* student list product.
-Commodification of the Market Segment Model "engineers homophily" [@RN4975] by taking a snapshot of existing stratification and then recommending that colleges utilize this snapshot as the basis for decisions about where to concentrate recruiting efforts.
-@RN4982 [p.
-58] states that "correlations, do not simply predict actions; they also form them." In itself, the Market Segment Model is simply a stylized depiction \[HIRSCHMAN 2016 STYLIZED FACTS\] of the structure of college choice, one that does not consider class-based and race-based structures that produced these patterns.
-Once the College Board inscribed the Market Segment Model into the EPS and Student Search Service Products, these products are likely to amplify the effect of historic structural inequalities on future opportunity structure by encouraging colleges to pursue local markets -- and student segments within these local markets -- in ways that are consistent with the patterns observed by @RN4982.
-
-Utilization of the Market Segment Model and geomarkets within the Enrollment Planning Service (EPS) raises concerns related to the quantification themes of reactivity, discipline, and authority introduced [@RN5007].
+Utilization of the EPS product raises concerns about the quantification themes of reactivity, discipline, and authority [@RN5007].
 @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Although EPS is a business-facing rather than a consumer-facing product, it has been a highly salient one.
+Market research suggests that EPS was highly salient.
 @RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
 We suggest that EPS disciplines colleges to approach recruiting in a manner consistent with the Market Segment Model.
-Colleges are encouraged to, first, choose geomarkets with a high number of desirable prospects and, second, plan recruiting efforts within selected geomarkets.
-Under the principle of homophily, selective colleges should target geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-Consistent with this perspective, EPS software documentation from @RN5017 describes how geomarkets -- here, called "EPS market codes" -- can be utlized:
-
-> EPS\textsuperscript{\texttrademark} (Enrollment Planning Service) is a geographic and demographic data service offered annually by the College Board to Colleges and Universities.
-EPS provides information to subscribing institutions about competitors, feeder schools, and demographic strengths and weaknesses.
-EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful.
+Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters by, first, choosing which Geomarkets to recruit from -- those with a large number of prospects from the desired market segment -- and, second, informing which high schools to visit within these Geomarkets.
+Under the principle of homophily, selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
 
 EPS may also weaken the authority of local decision-makers.
-The Market Segment model sought to develop a concrete, data-driven framework that replicated the aggregate knowledge of local admissions officers [@RN4982].
-Once this knowledge is quantified and commodified into a product stored on a CD-ROM, the local expertise of admissions officers becomes less valuable.
+@RN4982 sought to develop a concrete, data-driving framework -- The Market Segment Model -- replicated the aggregate knowledge of local admissions officers.
+Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
 The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
-In turn, admissions officers are relegated to a foot-soldier implementation role rather than being sought out for their local expertise about student demand.
+In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
+However, use of EPS contributed to most colleges visiting the same sets of affluent, high-achieving high schools.[^9]
 
+[^9]: We were told that savvier EPS users visited high schools that EPS recommended not visiting because there would be less competition for the good students that attended these high schools.
+
+## Geomarket Filter in Student Search Service
+
+In XXXX, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service.
+A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.
+For public universities that purchased lists, 80% purchase more than 50,000 names annually.
+@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.
+Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+
+Student lists are positively associated with student outcomes.
+A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.
+Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).[^10] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^11]
+
+[^10]: Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
+
+[^11]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
+Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
+The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).
+
+However, @RN5012 argue that student list products exacerbate racial inequality in college access.
+@RN5012 conceptualize student list products as "selection devices" [@RN4778] that enable colleges to select which prospective students they target by incorporating search filters (e.g., high school graduating class, state).
+@RN4786 defines "racialized inputs" as ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
+@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) meet the criteria of racialized inputs.
+Using a national sample of high school students and using data from actual student lists purchased by public universities, @RN5012 show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
+<!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
+-->
+
+Because of the extent of residential segregation in the U.S., geographic borders are a commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887].
+Products built upon structural racism are structurally racist products.
+Student list products offer many geographic filters.
+Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
+Additionally, the College Board created geographic borders that subsequently became filters in the *Student Search Service*.[^12] Geomarket search filters, slice states and metropolitan areas into smaller, local recruiting markets.
+Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
+First, they are ostensibly race-neutral inputs in that neither @RN4982, nor subsequent promotional material [e.g., @RN1621], mention race.
+Second, Geomarkets may be systematically correlated with race.
+For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda Country, excluding Oakland." A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.
+This article extends @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
+
+[^12]: For example, the geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
+
+@RN4982 [pp.
+42-44] provides recommendations about how The Market Segment Model and Geomarkets should be utilized in the context of purchasing student lists.
+The Market Segment model "simply demonstrates what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations" (p.
+42).
+The authors describe a college that "draws most of its students from regional or national segments" (p.
+42).
+Focusing on New England -- including the Boston Geomarket, which was relatively low-income in 1980 -- the authors ask, "where would you concentrate your energies" (p.
+42),
+
+> Ideally, you would seek communities \[Geomarkets\] with a high proportion of students already predisposed toward institutions such as your own.
+The Market Segment Model would provide this information through segment percentages for the community in question.
+Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
+Your efforts would surely be better directed toward three of the four communities in the bottom band, Manchester, Hartford, and Fairfield County.
+Simply by knowing a little bit about the students' backgrounds and academic records you could quickly focus your attention on those most likely to consider your kind of institution.
+Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own" (p.
+42-44)
+
+In both mailings and off-campus recruiting visits, @RN4982 recommend focusing on Geomarkets with large populations of students who are deciding between "institutions like your own" (p.
+44).
+Selective colleges primarily draw from students in the "regional" and "national" segments.
+@RN4982 suggest that the student list purchases of selective colleges should focus on Geomarkets with large numbers of affluent, highly educated households.
+By contrast, the student list purchases of local state colleges and community colleges, which rely on enrollment from students in the "local" and "in-state," should focus on nearby Geomarkets with large numbers of low-SES students.
+Our analyses attempt to recreate the strategy recommended by @RN4982.
+Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access.
+Furthermore, we anticipate that the students of color targeted by this strategy are disproportionately enrolled in affluent, predominantly white schools.
+
+```{=html}
+<!-- THIS TEXT IS ABOUT UTILIZATION. CUT FOR NOW.
 Finally, we suggest that utilization of geomarket filters within the Student Search Service product is likely to amplify class-based and race-based disparities in college opportunity.
@@
 Furthermore, purchases that specify multiple filters (e.g., test scores, grades, intended major) including geomarkets can yield unintended racial inequality because administrators have incomplete knowledge about how the intersection of these filters interact with local patterns of segregation.
+-->
+```
 
+```{=html}
 <!-- CUT TEXT DEFINING COLORBLIND RACISM AND RACE-NEUTRAL RACISM
@@
 The Market Segment Model can be seen as an example of colorblind racism.
-@RN4982 develop an analysis of the structure of college choice that ignores race but we contend that geo-markets are correlated with race.
+@RN4982 develop an analysis of the structure of college choice that ignores race but we contend that Geomarkets are correlated with race.
 Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975].
 -->
+```
+
+```{=html}
 <!-- CUT PARAGRAPH: WHY GEOMARKETS ARE WEAPONS OF MATH DESTRUCTION
-The incorporation of geo-markets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction.
-First, geo-markets are opaque.
+The incorporation of Geomarkets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction.
+First, Geomarkets are opaque.
 The borders of a geomarket may be unclear to someone purchasing student lists and very few people know how these borders were created \[COULD TALK ABOUT PRODUCT UTILIZATION HERE; NOT CLEAR HOW GEOMARKETS INTERACT WITH OTHER STUDENT LIST FILTERS\].
-Second, geo-markets scale.
-Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on geo-markets in concert with other filters (e.g., SAT score, GPA).
-Third, geo-markets can cause damage.
-When a college decides to purchase names from one geo-market but exclude prospects from an adjacent geo-market, this affects which college students have an opportunity to attend.
+Second, Geomarkets scale.
+Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on Geomarkets in concert with other filters (e.g., SAT score, GPA).
+Third, Geomarkets can cause damage.
+When a college decides to purchase names from one Geomarket but exclude prospects from an adjacent Geomarket, this affects which college students have an opportunity to attend.
 @RN4982 suggests that selective colleges focus on geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
 -->
+```
 
-# Methods
+# Data and Methods
+
+## Data and Variables
+
+**Research question 1, data**.
+RQ1 asks, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas?
+And, how does this variation change over time?
+To answer this question, we use census tract-level data from the US Census Bureau.
+Census tract is the smallest geographic area for which measures of both race/ethnicity and socioeconomic characteristics are available.
+Utilizing a small geographic area is important for implementing a spatial merge between Census data and Geomarket shape files, described below.
+
+We utilize data from the 1980 Decennial Census, specifically variables from Summary Tape File 1 (STF1), which contains data from the about race and hispanic origin from the "short form" questions answered by all households and data from Summary Tape File 3 (STF3), which contains data about socioeconomic characteristics -- including income, education, and poverty -- derived from the "long form" questionnaire completed by a sample of households.
+The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
+In order to examine how the characteristics of Geomarkets changed over time, we utilize tract-level data from the 2000 Decennial Census, similarly using data from Summary File 1A (SF1A) to obtain measures of race and Hispanic origin and using data from Summary File 3A (SF3A) to obtain measures of income, education, and poverty.[^13] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+These questions were replaced by the American Community Survey (ACS).
+We also utilized data from the 2020 5-year ACS, which includes data collected from 2016-2020.[^14] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+
+[^13]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
+
+[^14]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
+
+We assigned each census tract to College Board Geomarkets by implementing a partial spatial merge.
+Spatial manipulations utilized functions from the *sf* R package.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^15] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Using these zip code-level data, we utilized the `aggregate()` function to create shapefile for each Geomarket.
+
+[^15]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
+
+Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated shapefiles.
+Census tract shapefiles that intersected with multiple Geomarkets were broken into smaller shapefiles wholly contained within a single Geomarket.
+The variables in these shapefiles were then replaced with the original variable value multiplied by the proportion of land area of the census tract that was contained within the Geomarket.
+For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 1,000 people who identify as Hispanic non-white.
+The partial spatial merge splits this census tract into two observations.
+The observation assigned to Geomarket A reports $.60*1000 = 600$ Hispanic non-white people and the observation assigned to Geomarket B reports $.40*1000 = 400$ Hispanic non-white people.
+
+**Research question 1, variables**.
+RQ1 asks, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas?
+And, how does this variation change over time?
+We created measures of race and ethnicity, specifically the number and proportion of people (all ages) in the following categories: White, non-Hispanic; Black, non-Hispanic; Asian, non-Hispanic; Pacific Islander, non-Hispanic; American Indian and Alaskan Native, non-Hispanic; and two+ races, non-Hispanic.
+Measures of XXX were not available for 1980 Decennial Census.
+We also created socioeconomic measures that approxmiate the predictors of market segment identified by @RN4982.
+These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^16] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
+
+[^16]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
+
+**Research question 2, data**.
+Geomarket is a search filter available in the College Board Student Search Service product.
+RQ2 asks, how does the racial and socioeconomic composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
+We answer this question using actual student lists purchased by public universities.
+
+In February 2020, we began issuing public records requests to public universities about student lists purchased from 2016 through 2020.
+We narrowed the scope of our request to student lists purchased from College Board, ACT, and the National Research Center for College and University Admissions (NRCCUA), the three largest student list vendors at the time.
+For each student list purchased from 2016 through 2020, we requested two related pieces of data: (1) the order summary, which specifies search criteria for the student list purchase; and (2) the de-identified prospect-level list produced from the search criteria.
+
+The data collection sample for this project was all public universities in CA, IL, MN, and TX.
+Utilizing public records requests to obtain public records is a painstaking process.
+Initially, the majority of universities did not respond to our request or denied our request.
+Subsequently, we obtained pro bono representation from four law firms, which substantially increased the success of data collection.
+We collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
+However, we were unable to obtain legal representation for TX.
+
+Even with firm representation, data collection remained difficult.
+Some universities provided records that were not usable for quantitative analyses (e.g., summary statistics across multiple orders; or data did not contain important fields).
+Some universities did not provide records based on legitimate grounds (e.g., data not in university possession; not required to create records that do not currently exist).
+We learned that many universities outsourced student list purchases to a third-party consulting firm.
+Unfortunately, we were rarely able to obtain usable data from these universities.
+A small number of universities denied requests based on potentially questionable legal rationale, but we lacked the resources to litigate.
+
+This article analyzes student lists purchased from College Board.
+We exclude student list purchases by MN public universities from this report because Minnesota is predominantly an "ACT state." 14 public universities provided PDF order summaries containing the search criteria for 830 student list purchases.
+We utilized the XXXX Python/R package to convert these order summaries into tabular data.
+About 16% of student list purchases utilized the Geomarket filter.
+
+We received both (a) the order summary data and (b) the de-identified prospect-level student list data for 414 orders associated with 2,549,085 prospects.
+We draw from these data to answer RQ2.
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^17].
+Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
+
+[^17]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
+
+In particular, we answer RQ2 by analyzing particular student list purchases that did *not* utilize the Geomarket search filter, but did purchase prospects from all Geomarkets in a particular metropolitan area.
+For each prospect we know their home zip code and, therefore, which Geomarket they belong to.
+We answer RQ2 by simulating which prospective students would have been included or excluded from the student list purchase had the purchase filtered on partiulcar Geomarkets.
+
+**Research question 2, variables**.
+The variables of interest for RQ2 the characteristics of prospective students who would have been included/excluded by our simulations and the characteristics of their schools and neighborhoods.
+Prospect characteristics are derived from the pretest questionnaire administered to College Board test-takers.
+The de-identified student lists we received contain a standard subset of variables (e.g., HS code, HS GPA range, intended major).
+We create detailed measures of race and ethnicity.
+Information about student socioeconomic status is limited.
+However, we create a measure of whether the prospect is first-generation college student.
+We can also measure the socioeconomic characteristics of the students' home zip code and characteristics of their high school \[EXPAND OR DELETE THIS SENTENCE DEPENDING ON WHETHER WE USE THESE DATA\].
+
+## Methods
+
+**Research design**.
+We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
+Following XXX \[CITE\], we focus on a small number \[HOW MANY\] of metropolitan areas in order to provide sufficient analytic depth while also situating each case within the context of historic segregation and contemporary gentrification.
+Online Appendix XXX presents results for a larger number of metropolitan areas.
+
+Choice of metropolitan areas is informed by several factors.
+Based on our read of @RN4982, we are interested in "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
+11-12).
+Building on our conceptual framework, we are interested in racial and socioeconomic inequality between geomarkets within a metropolitan area (RQ1) and how this inequality contributes to inequality in which prospective students are targeted by student list purchases (RQ2).
+\[DISCUSSION OF WHICH CASES TO SELECT.
+KARINA\].
+Choice of metropolitan areas is informed by data availability.
+Although Census data are available for all metropolitan areas (RQ1) we do not have good candidate student list purchases for all metropolitan areas (RQ2).
+\[SAY SOMETHING ABOUT WHAT IS A GOOD CANDIDATE STUDENT LIST PURCHASE -- PURCHASE WITH NOT TOO MANY CRITERIA THAT INCLUDE ALL GEOMARKETS IN A METRO\].
+
+**Analyses**.
+Analyses are simultaneously descriptive and spatial.
+We answer RQ1 by producing Geomarket-level tables and graphs that show how the Geomarkets in a selected metropolitan area vary on racial and socioeconomic characteristics and how they vary over time.
+For example, in the San Francisco Bay Area, how does the City of Oakland Geomarket (CA07) differ from the Alameda County Exluding Oakland Geomarket (CA08).
+We also produce interactive maps at the census tract-level to show more granular variation within and between Geomarkets.
+
+We answer RQ2 by analyzing student list purchases that encompassed all Geomarkets in a selected metropolitan areas.
+Descriptive tables and graphs describe the racial and socioeconomic characteristics of prospects that would have been included/excluded had the purchase filtered on particular Geomarkets.
+Interactive maps provide a more granual visualization of where prospects lived and went to high school.
+For example, for purchased profiles who lived in the Chester County Geomarket (PA02) \[OR PHILADELPHIA PA05\] and identified as Black, where in the county did these students live and which high schools did they attend?
+
+NEXT STEPS: CREATE TABLE FOR PHILLY STUDENT LIST PURCHASES THAT FOCUSES ON FIRST GEN STATUS.
+
+Our conceptual framework suggests that racial and socioeconomic inequality is Geomarket regions is associated with the number of geomarkets in the metropolitan area.
+and is associated with the patterns of racial segregation in the metropolitan area.
+comment more on this.
 
 Below, we use simulations and actual student list purchases that filtered on geomarkets.
-In purchases that filtered on test-score and/or GPA thresholds, we can simluate who would be included and excluded had certain geo-markets were been selected.
+In purchases that filtered on test-score and/or GPA thresholds, we can simluate who would be included and excluded had certain Geomarkets were been selected.
 In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.
@@
 
+\appendix
+\renewcommand{\thetable}{A\arabic{table}}
+
+# Appendix A
+
+@tbl-market-segment-characteristics reproduces a partial, simplified version of the bottom panel of @RN4982 [Figure 2.1] which represents the Market Segment Profile for "Connecticut Market 3: Fairfield County." @tbl-market-segment-characteristics has separate columns for each market segment (local, in-state, regional, national) and rows show the number of test-takers and characteristics of test-takers.
+For example, there were 550 "local" students in Fairfield county and these students submitted SAT scores to 2.8 postsecondary institutions on average.
+By contrast, there were 1,664 "regional" students who submitted test scores to 4.8 institutions on average.
+For local students, 11.2% had family income greater than \$35,000 and 9% had both parents with a BA.
+For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA.
+Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.
+
+@tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market.
+For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution.
+Of the 1,199 in-state students, 757 (63.1) sent scores to institution #1, a public doctoral granting institution, and 515 (43.0%) sent scores to institution #2, a public master's granting institution.
+
+The *Institutional Profile* describes students who send scores to a particular institution and which majors these students are interested in.
+For each institution, institutional profiles are created separately for students from a particular local market, for all students in a state, or all students in a region.
+@tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the institutional profile of an anonymous institution for students from Fairfield County, CT.
+@tbl-sample-institutional-profile shows that 58 in-state students submitted SAT scores to the institution.
+These 58 students represented 4.8% of the total 1,199 in-state students from Fairfield County.
+69 regional students sent scores to the institution, representing 4.1% of all 1,664 regional test-takers.
+Of these 69 regional students, 35 expressed interest in majoring in the liberal arts.
+These 35 students represent 7.0% of all regional students from Fairfield County who expressed interest in the liberal arts.
+
+@RN4982 argue that the Institutional Profile and the Market Segment Profile enable admissions officers to know where to look for students and which institutions are competing for those students (p.
+25):
+
+> The Institutional Profile and the Market Segment Profile quantify the admission officers' intuitive grasp of market structure.
+Structure here carries a dual meaning, connoting both the structure of student choice and the structure of institutional competition\ldots This two-sided interpretation furnishes the essential framework for planning by individual colleges and universities\ldots To draw effectively on its own natural constituency, a college not only must contact the "right" kind of students --- that is, students who are predisposed toward that type of institution -- but also must persuade them of its special character.
+This means knowing the competition as well as the clientele.
+
 ```{r market-segment-characteristics, echo = F}
@@
   #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```
-
-\clearpage
-
-```{r sample-institutional-profile, echo = F}
-#| label: tbl-sample-institutional-profile
-#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'
-
-sample_inst <- data.frame(
-  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
-  local = c(1,0.2,0,0.0,0,0.0),
-  in_state = c(58,4.8,25,8.4,2,3.0),
-  regional= c(69,4.1,35,7.0,5,5.8),
-  national = c(109,2.9,61,5.3,3,0.8),
-  total = c(237,3.3,121,5.9,10,1.8)
-)
-
-kable(
-  x = sample_inst,
-  col.names = c('','Local','In-state','Regional','National','Total'),
-  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
 )
@@
 ```
+
+\clearpage
+
+```{r sample-institutional-profile, echo = F}
+#| label: tbl-sample-institutional-profile
+#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'
+
+sample_inst <- data.frame(
+  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
+  local = c(1,0.2,0,0.0,0,0.0),
+  in_state = c(58,4.8,25,8.4,2,3.0),
+  regional= c(69,4.1,35,7.0,5,5.8),
+  national = c(109,2.9,61,5.3,3,0.8),
+  total = c(237,3.3,121,5.9,10,1.8)
+)
+
+kable(
+  x = sample_inst,
+  col.names = c('','Local','In-state','Regional','National','Total'),
+  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
+)
+```

### 2025-01-15 scripts/Untitled.R

@@ -1,161 +1,0 @@
-
-###################
-################### implementing a modular approach
-###################
-###################
-
-create_layer <- function(map, data, var_name, yr, palette) {
-  group_name <- paste(var_name, yr)
-  
-  map <- map %>%
-    addPolygons(
-      data = data,
-      fillColor = ~palette(data[[var_name]]),
-      weight = 1,
-      opacity = 1,
-      color = '#808080',  # Line color
-      dashArray = '3',
-      fillOpacity = 0.7,
-      highlightOptions = highlightOptions(
-        weight = 4,
-        color = "#666",
-        dashArray = "",
-        fillOpacity = 0.7,
-        bringToFront = TRUE
-      ),
-      label = sprintf(
-        "<strong>%s</strong> - %s<br/>%g",
-        data$eps, data$eps_name, data[[var_name]]
-      ) %>% lapply(htmltools::HTML),
-      group = group_name
-    )
-  
-  return(map)
-}
-
-# 2. create_legend()
-  #This function will create a legend for a given variable.
-
-create_legend <- function(map, data, var_name, palette) {
-  map <- map %>%
-    addLegend(
-      position = "topright",
-      pal = palette,
-      values = data[[var_name]],
-      title = var_name,
-      opacity = 0.7,
-      className = paste0("legend-", var_name)
-    )
-  
-  return(map)
-}
-
-#3. create_layers_control()
-  #This function will create the layers control based on the provided group names.
-
-create_layers_control <- function(map, group_names) {
-  map <- map %>%
-    addLayersControl(
-      baseGroups = group_names,
-      position = 'bottomleft',
-      options = layersControlOptions(collapsed = FALSE)
-    )
-  
-  return(map)
-}
-
-# 4. create_eps_maps()
-  #This is the main function that coordinates the map creation by calling the above modular functions.
-
-create_eps_maps <- function(eps_codes, vars) {
-  
-  # Create the leaflet map
-  map <- leaflet() %>%
-    addProviderTiles(provider = providers$CartoDB.Positron)
-  
-  # Store group names for layers control
-  group_names <- c()
-  
-  # Initialize a list to store palettes for each variable
-  palettes <- list()
-  
-  # Add layers for each year and variable
-  for (var in vars) {
-    var_group_names <- c()  # Initialize the group name for each variable
-    
-    for (yr in c(1980, 2000, 2020)) {
[…]

### 2025-01-15 scripts/append_census.R

@@ -1,0 +1,188 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < leaflet_maps.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 1/13/2025
+## [ DESC ] < append tract-level and EPS-level census data from 1980, 2000, and 2016-2020 ACS (for RQ1) >
+################################################################################
+
+
+### SETTINGS
+#rm(list = ls()) # remove all objects
+options(max.print=1000)
+#options(width = 160)
+# Set the scipen option to a high value to avoid scientific notation
+options(scipen = 999)
+
+################################# LIBRARIES
+  library(tidyverse)
+  library(readxl)
+  library(lubridate)
+  library(haven)
+  library(labelled)
+  library(tidycensus)
+  # get census api key
+  #http://api.census.gov/data/key_signup.html
+  # set census api key, and install for future use
+  #census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+  Sys.getenv("CENSUS_API_KEY") # retreive API key
+
+
+  library(sf)
+  # Enable caching for tigris
+  options(tigris_use_cache = TRUE)
+  library(tigris)
+  #library(stars)
+  #library(spatstat)
+  #library(rgeos)
+  library(lwgeom) # library has checks/vixes for valid geometries
+  library(leaflet)
+  library(shiny)
+  library(kableExtra)
+
+################################# DIRECTORY PATHS
+
+  getwd()
+  
+  data_dir <- file.path('.','data') # main data directory
+  list.files(path = data_dir)
+  
+  d1980_data_dir <-   file.path('.',data_dir,'1980_decennial') # main data directory
+  list.files(path = d1980_data_dir)
+  
+  d2000_data_dir <-   file.path('.',data_dir,'2000_decennial') # 
+  list.files(path = d2000_data_dir)
+  
+  acs2020_data_dir <-   file.path('.',data_dir,'2020_acs') # 
+  list.files(path = acs2020_data_dir)
+  
+  # can't save shape files in the repo cuz they too big
[…]

### 2025-01-15 scripts/chicken_scratch/experimenting_with_leaflet.R

@@ -1,0 +1,589 @@
+# EXPERIMENTING WITH LEAFLET
+
+### SETTINGS
+rm(list = ls()) # remove all objects
+options(max.print=1000)
+#options(width = 160)
+# Set the scipen option to a high value to avoid scientific notation
+options(scipen = 999)
+
+### LIBRARIES
+library(tidyverse)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+# get census api key
+#http://api.census.gov/data/key_signup.html
+# set census api key, and install for future use
+#census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+Sys.getenv("CENSUS_API_KEY") # retreive API key
+# mapbox token: pk.eyJ1Ijoib3phbmoiLCJhIjoiY20wYjdwYnZxMDY1NzJrcHZuNHpkZjk4ZCJ9.QTfW1xSq0MgKrXBEe3fSUg
+# default public token: pk.eyJ1Ijoib3phbmoiLCJhIjoiY20wYjhzd2c0MDZmOTJqcHRmcXFmaW0zaCJ9.HlonUhSIdWIuWJXUsqh-eA
+library(sf)
+# Enable caching for tigris
+options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+library(lwgeom) # library has checks/vixes for valid geometries
+library(leaflet)
+
+##############
+  # from https://walker-data.com/census-r/mapping-census-data-with-r.html#interactive-mapping
+
+dallas_bachelors <- get_acs(
+  geography = "tract",
+  variables = "DP02_0068P", # percentage of 25+ with a BA
+  year = 2020,
+  state = "TX",
+  county = "Dallas",
+  geometry = TRUE
+)
+
+# get an error when mapping in leaflet unless we use different CRS
+  # Reproject dallas_bachelors to WGS84
+  dallas_bachelors <- st_transform(dallas_bachelors, crs = 4326)
+
+dallas_bachelors %>% class()
+dallas_bachelors %>% glimpse()
+
+dallas_bachelors %>% select(GEOID) %>% as.data.frame() %>% group_by(GEOID) %>% summarise(n_per_group = n()) %>% ungroup %>% count(n_per_group) # uniquely identifies
+
+
+# mapview library
+
+library(mapview)
[…]

### 2025-01-15 scripts/create_2020_acs.R

@@ -124,6 +124,7 @@ variables <- c(
 variables <- c(variables,edu_var_names)
 variables
 
+
 acs2020_vars %>% filter(name %in% variables) %>% print()
 
 # READ IN TRACT-LEVEL 2016-2020 ACS DATA FROM TIDYCENSUS

### 2025-01-15 scripts/create_stu_list.R

@@ -44,7 +44,6 @@ library(kableExtra)
 
 socal_eps_codes <- paste0("CA", 14:28) # socal 
 bay_eps_codes <- c(paste0("CA ", 4:9), paste0("CA", 10:11)) # bay area
-bay_eps_codes 
 ca_eps_codes <- c(paste0("CA ", 1:9), paste0("CA", 10:34))
 chi_eps_codes <- c(paste0("IL ", 7:9), paste0("IL", 10:13))
 chi_eps_codes
@@ -400,12 +399,19 @@ orders_df$race_filter[orders_df$race_filter==''] <- NA
   # 12 University of Illinois at Urbana-Champaign 11470 [145637]
   # 10 University of Illinois at Chicago          5015  [id = 145600]
   
-    
+  lists_orders_zip_hs_df %>% glimpse()
+  
+  lists_orders_zip_hs_df %>% count() 
+  
+  lists_orders_zip_hs_df %>% count(stu_first_gen) 
+  
+  
+  
   lists_orders_zip_hs_df %>% filter(hs_state_code == 'PA' & univ_id ==104151) %>% count(ord_num) %>% arrange(desc(n)) %>% print(n=20) # ASU
   lists_orders_zip_hs_df %>% filter(hs_state_code == 'PA' & univ_id ==110680) %>% count(ord_num) %>% arrange(desc(n)) %>% print(n=10) # UC san diego
   lists_orders_zip_hs_df %>% filter(hs_state_code == 'PA' & univ_id ==145600) %>% count(ord_num) %>% arrange(desc(n)) %>% print(n=10) # U Illinois Chicago
   lists_orders_zip_hs_df %>% filter(hs_state_code == 'PA' & univ_id ==145637) %>% count(ord_num) %>% arrange(desc(n)) %>% print(n=10) # U Illinois Urbana-Champaign
-  
+
   orders_df %>% filter(order_num == '395793')
   
   orders_df %>% glimpse()
@@ -417,6 +423,7 @@ orders_df$race_filter[orders_df$race_filter==''] <- NA
   ##########################
   # order that was just native american indian/alaskan native
   lists_orders_zip_hs_df %>% glimpse()
+  lists_orders_zip_hs_df %>% filter(ord_num == '487927') %>% count(stu_first_gen)
   lists_orders_zip_hs_df %>% filter(ord_num == '487927') %>% count(hs_state_code) %>% arrange(desc(n)) %>% print(n=50)
[…]

### 2025-01-15 scripts/experimenting_with_leaflet.R

@@ -1,589 +1,0 @@
-# EXPERIMENTING WITH LEAFLET
-
-### SETTINGS
-rm(list = ls()) # remove all objects
-options(max.print=1000)
-#options(width = 160)
-# Set the scipen option to a high value to avoid scientific notation
-options(scipen = 999)
-
-### LIBRARIES
-library(tidyverse)
-library(readxl)
-library(lubridate)
-library(haven)
-library(labelled)
-library(tidycensus)
-# get census api key
-#http://api.census.gov/data/key_signup.html
-# set census api key, and install for future use
-#census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
-Sys.getenv("CENSUS_API_KEY") # retreive API key
-# mapbox token: pk.eyJ1Ijoib3phbmoiLCJhIjoiY20wYjdwYnZxMDY1NzJrcHZuNHpkZjk4ZCJ9.QTfW1xSq0MgKrXBEe3fSUg
-# default public token: pk.eyJ1Ijoib3phbmoiLCJhIjoiY20wYjhzd2c0MDZmOTJqcHRmcXFmaW0zaCJ9.HlonUhSIdWIuWJXUsqh-eA
-library(sf)
-# Enable caching for tigris
-options(tigris_use_cache = TRUE)
-library(tigris)
-#library(stars)
-#library(spatstat)
-#library(rgeos)
-library(lwgeom) # library has checks/vixes for valid geometries
-library(leaflet)
-
-##############
-  # from https://walker-data.com/census-r/mapping-census-data-with-r.html#interactive-mapping
-
-dallas_bachelors <- get_acs(
-  geography = "tract",
-  variables = "DP02_0068P", # percentage of 25+ with a BA
-  year = 2020,
-  state = "TX",
-  county = "Dallas",
-  geometry = TRUE
-)
-
-# get an error when mapping in leaflet unless we use different CRS
-  # Reproject dallas_bachelors to WGS84
-  dallas_bachelors <- st_transform(dallas_bachelors, crs = 4326)
-
-dallas_bachelors %>% class()
-dallas_bachelors %>% glimpse()
-
-dallas_bachelors %>% select(GEOID) %>% as.data.frame() %>% group_by(GEOID) %>% summarise(n_per_group = n()) %>% ungroup %>% count(n_per_group) # uniquely identifies
-
-
-# mapview library
-
-library(mapview)
[…]

### 2025-01-15 scripts/geometries_tutorial.R

@@ -1,138 +1,0 @@
-# investigating different kinds of geometries
-
-# Load the necessary library
-library(sf)
-
-# Create a simple POLYGON geometry
-# The st_polygon function takes a list of coordinates that form the exterior ring of the polygon.
-# Here we define a square with coordinates (0,0), (1,0), (1,1), (0,1), and close it by returning to (0,0)
-polygon_geom <- st_sfc(st_polygon(list(rbind(
-  c(0, 0),  # First corner (x, y)
-  c(1, 0),  # Second corner (x, y)
-  c(1, 1),  # Third corner (x, y)
-  c(0, 1),  # Fourth corner (x, y)
-  c(0, 0)   # Closing the polygon (back to first corner)
-))))
-
-polygon_geom %>% class()
-polygon_geom %>% str()
-polygon_geom %>% glimpse()
-
-# View the numerical representation of the POLYGON geometry
-print(polygon_geom)
-
-# Plot the POLYGON
-# We use the plot() function to visualize the polygon.
-# The col argument sets the fill color, and border sets the color of the polygon's outline.
-plot(polygon_geom, col = 'lightblue', border = 'darkblue', main = "Simple Polygon")
-
-###############
-############### MULTIPOLYGON
-###############
-
-# Load the necessary library
-library(sf)
-
-# Create a MULTIPOLYGON geometry
-# A MULTIPOLYGON consists of multiple POLYGONs, each defined by a set of coordinates.
-# Below we define two squares (two separate POLYGONs) as part of the MULTIPOLYGON:
-# - The first square has corners (0,0), (1,0), (1,1), (0,1)
-# - The second square has corners (2,2), (3,2), (3,3), (2,3)
-multi_polygon_geom <- st_sfc(st_multipolygon(list(
-  list(rbind(
-    c(0, 0),  # First square's first corner
-    c(1, 0),  # First square's second corner
-    c(1, 1),  # First square's third corner
-    c(0, 1),  # First square's fourth corner
-    c(0, 0)   # Closing the first square
-  )),
-  list(rbind(
-    c(2, 2),  # Second square's first corner
-    c(3, 2),  # Second square's second corner
-    c(3, 3),  # Second square's third corner
-    c(2, 3),  # Second square's fourth corner
-    c(2, 2)   # Closing the second square
-  ))
-)))
-
[…]

### 2025-01-15 scripts/leaflet_maps.R

@@ -24,7 +24,7 @@ library(tidycensus)
 #http://api.census.gov/data/key_signup.html
 # set census api key, and install for future use
 #census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
-Sys.getenv("CENSUS_API_KEY") # retreive API key
+#Sys.getenv("CENSUS_API_KEY") # retreive API key
 
 
 library(sf)
@@ -284,7 +284,7 @@ create_eps_table <- function(eps_codes, table_vars) {
     
     # Set the order of the 'variable' column based on the provided table_vars
     mutate(variable = factor(variable, levels = recode(table_vars, !!!table_var_names))) %>%
-    
+
     # Reshape statistics and years from rows to columns
     pivot_wider(
       names_from = c(year, statistic),  # Pivot both year and statistic to columns
@@ -310,7 +310,7 @@ create_eps_table <- function(eps_codes, table_vars) {
 
 # Example usage:
 chi_eps_table <- create_eps_table(chi_eps_codes, table_varlist)
-chi_eps_table
+chi_eps_table %>% filter(year == 2020) %>%  print(n=100)
 
 # Example usage:
 chi_eps_table <- create_eps_table(chi_eps_codes, table_varlist)
@@ -333,7 +333,291 @@ chi_eps_table
     philly_eps_table %>% print(n=100)  
     
     saveRDS(socal_eps_table, file.path(tables_dir, "socal_eps_table.rds"))
-    saveRDS(philly_eps_table, file.path(tables_dir, "philly_eps_table.rds"))    
+    saveRDS(philly_eps_table, file.path(tables_dir, "philly_eps_table.rds"))
+    
+########
+######## EXPERIMENT W/ CREATING GRAPHS FROM TABLE 
+########
+    
+chi_eps_table <- create_eps_table(chi_eps_codes, table_varlist)
+chi_eps_table %>% glimpse()
+
+# uniquely identify rows
+chi_eps_table %>% # start with data frame object
+  group_by(eps_codename, variable) %>% # group by unitid
+  summarise(n_per_group=n()) %>% # create measure of number of obs per group
+  ungroup %>% # ungroup (otherwise frequency table [next step] created) separately for each group (i.e., separate frequency table for each value of unitid)
[…]

### 2025-01-15 scripts/leaflet_working.R

@@ -1,682 +1,0 @@
-
-# This script contains code that is not my current version but is good enough that I don't want to delete it forever
-
-
-
-
-##### this solution doesn't use any javascript; the downside is that legends are shown for all variables, regardless of which base-layer is shown
-
-create_eps_maps <- function(eps_codes, vars) {
-  
-  # Create the leaflet map
-  map <- leaflet() %>%
-    addProviderTiles(provider = providers$CartoDB.Positron)
-  
-  # Store group names for layers control
-  group_names <- c()
-  palettes <- list()  # Store palette functions
-  domains <- list()   # Store domains (data ranges)
-  
-  # Iterate through each variable
-  for (var in vars) {
-    all_years_data <- NULL
-    
-    # Aggregate data across all years for this variable
-    for (yr in c(1980, 2000, 2020)) {
-      filtered_data <- allyr_anal_eps_sf %>% filter(year == yr, eps %in% eps_codes)
-      if (nrow(filtered_data) > 0 && !all(is.na(filtered_data[[var]]))) {
-        all_years_data <- c(all_years_data, filtered_data[[var]])
-        
-        # Create and add the layer for this year
-        max_value <- max(filtered_data[[var]], na.rm = TRUE)
-        min_value <- min(filtered_data[[var]], na.rm = TRUE)
-        bins <- if (grepl("inc", var, ignore.case = TRUE)) {
-          c(0, 50000, 75000, 100000, 125000, 150000, 175000, 200000, 225000, max_value)
-        } else {
-          pretty(c(min_value, max_value), n = 5)
-        }
-        
-        palette <- colorBin(palette = 'YlGnBu', domain = filtered_data[[var]], bins = bins)
-        map <- create_layer(map, filtered_data, var, yr, palette)
-        group_names <- c(group_names, paste(var, yr))
-      }
-    }
-    
-    # Create a single palette for this variable across all years
-    if (!is.null(all_years_data)) {
-      combined_palette <- colorBin(palette = 'YlGnBu', domain = all_years_data, bins = bins)
-      palettes[[var]] <- combined_palette
-      domains[[var]] <- all_years_data
-    }
-  }
-  
-  # Add a single legend for each variable
-  for (var in vars) {
-    if (var %in% names(palettes)) {
-      map <- map %>%
-        addLegend(
-          pal = palettes[[var]], 
-          values = domains[[var]],  # Use the combined domain for the variable
[…]

### 2025-01-15 scripts/metro_eps_codes.R

@@ -1,0 +1,39 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < metro_eps_codes.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 1/13/2025
+## [ DESC ] < creates character vectors of EPS codes for particular metros (or states), for use across multiple scripts >
+################################################################################
+
+library(tidyverse)
+
+socal_eps_codes <- paste0("CA", 14:28) # socal 
+bay_eps_codes <- c(paste0("CA ", 4:9), paste0("CA", 10:11)) # bay area
+bay_eps_codes 
+
+ca_eps_codes <- c(paste0("CA ", 1:9), paste0("CA", 10:34))
+chi_eps_codes <- c(paste0("IL ", 7:9), paste0("IL", 10:13))
+chi_eps_codes
+
+cleveland_eps_codes <- paste0("OH ", 2:6) #
+cleveland_eps_codes
+
+philly_eps_codes <- paste0("PA ", 1:5) #
+philly_eps_codes
+
+nj_eps_codes <- c(paste0("NJ ", 1:9), paste0("NJ", 10:12))
+nj_eps_codes
+
+nj_metro_eps_codes <- c('NJ 2','NJ 4','NJ 5',paste0('NJ ', 7:9),paste0("NJ", 10:11))
+nj_metro_eps_codes
+
+long_island_eps_codes <- c(paste0('NY',16:21))
+
+atl_eps_codes <- paste0("GA ", 1:4) #
+atl_eps_codes
+
+htown_eps_codes <- str_c("TX",15:18)
+htown_eps_codes
+
+dallas_eps_codes <- c('TX19','TX20','TX21','TX22','TX23','TX24')

### 2025-01-15 scripts/rq1_tables_graphs.R

@@ -1,0 +1,490 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < rq1_tables_graphs.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 1/13/2025
+## [ DESC ] < functions that create tables and graphs for RQ1 (differences between geomarkets, based on census data) >
+################################################################################
+
+### SETTINGS
+rm(list = ls()) # remove all objects
+options(max.print=1000)
+#options(width = 160)
+# Set the scipen option to a high value to avoid scientific notation
+options(scipen = 999)
+
+### LIBRARIES
+library(tidyverse)
+library(ggh4x) # additional graphing capabilities
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+# get census api key
+#http://api.census.gov/data/key_signup.html
+# set census api key, and install for future use
+#census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+#Sys.getenv("CENSUS_API_KEY") # retreive API key
+
+library(sf)
+# Enable caching for tigris
+options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+library(lwgeom) # library has checks/vixes for valid geometries
+library(leaflet)
+library(shiny)
+library(kableExtra)
+
+### DIRECTORY PATHS
+
+getwd()
+
+data_dir <- file.path('.','data') # main data directory
+list.files(path = data_dir)
+
+d1980_data_dir <-   file.path('.',data_dir,'1980_decennial') # main data directory
+list.files(path = d1980_data_dir)
+
+d2000_data_dir <-   file.path('.',data_dir,'2000_decennial') # 
+list.files(path = d2000_data_dir)
+
+acs2020_data_dir <-   file.path('.',data_dir,'2020_acs') # 
+list.files(path = acs2020_data_dir)
+
+# can't save shape files in the repo cuz they too big
+shape_dir <-   file.path('.','..','cb_geomarket_shape') # main data directory
+list.files(path = shape_dir)
+
+# had to move analysis data files into shape folder because too big for git
[…]

### 2025-01-15 scripts/sim-table that works.R

@@ -1,85 +1,0 @@
-
-format_eps_sim_table_simple <- function(df) {
-  library(dplyr)
-  library(kableExtra)
-  
-  # Extract the individual tables
-  counts <- df$count_table
-  col_pct <- df$col_pct_table
-  row_pct <- df$row_pct_table
-  
-  # Rename columns for consistency
-  counts_formatted <- counts %>%
-    rename(
-      `EPS Codename` = eps_codename,
-      `Total Students` = stu_all,
-      `Race Known` = stu_race_known,
-      White = stu_white,
-      Asian = stu_asian,
-      Black = stu_black,
-      Hispanic = stu_hispanic,
-      `Two Races` = stu_tworaces
-    ) %>%
-    mutate(Section = "1. Counts") %>%
-    select(Section, everything())
-  
-  col_pct_formatted <- col_pct %>%
-    rename(
-      `EPS Codename` = eps_codename,
-      `Total Students` = stu_all,
-      `Race Known` = stu_race_known,
-      White = col_pct_white,
-      Asian = col_pct_asian,
-      Black = col_pct_black,
-      Hispanic = col_pct_hispanic,
-      `Two Races` = col_pct_tworaces
-    ) %>%
-    mutate(Section = "2. Column Percentages") %>%
-    select(Section, everything())
-  
-  row_pct_formatted <- row_pct %>%
-    rename(
-      `EPS Codename` = eps_codename,
-      `Total Students` = stu_all,
-      `Race Known` = stu_race_known,
-      White = row_pct_white,
-      Asian = row_pct_asian,
-      Black = row_pct_black,
-      Hispanic = row_pct_hispanic,
-      `Two Races` = row_pct_tworaces
-    ) %>%
-    mutate(Section = "3. Row Percentages") %>%
-    select(Section, everything())
-  
-  # Combine all sections into one table
-  full_table <- bind_rows(counts_formatted, col_pct_formatted, row_pct_formatted)
-  
-  # Create a LaTeX-compatible table using kable
-  table_kable <- full_table %>%
-    kbl(
-      format = "latex",
-      booktabs = TRUE,
-      escape = FALSE,
-      col.names = c("Section", "EPS Codename", "Total Students", "Race Known", 
-                    "White", "Asian", "Black", "Hispanic", "Two Races"),
-      caption = "Characteristics of Geomarkets around Philadelphia \\label{tab:philly_eps_table_summary}"
-    ) %>%
-    kable_styling(
[…]

### 2025-01-30 text/cb_geomarket_online_appendix.qmd
tweaker

---
title: "Online appendix"
format:
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
---

# header 1

```{r}
getwd()
```

```{r}
# You can load libraries, set options, define your list of metros, etc.
library(tools)  # for toTitleCase() if desired

# Example lists
metros <- c("chicago", "denver")
RQs <- c("RQ1", "RQ2")

# Define the function to insert a figure, given a metro + RQ.
# 'base_path' is where your .png and .txt files live.
insert_figure <- function(metro, rq, base_path = "../results/graphs") {

  # Construct file paths
  fig_file    <- file.path(base_path, paste0(metro, "_", rq, ".png"))
  cap_file    <- file.path(base_path, paste0(metro, "_", rq, "_caption.txt"))
  note_file   <- file.path(base_path, paste0(metro, "_", rq, "_note.txt"))

  # Read the caption (if missing, use a default)
  if (file.exists(cap_file)) {
    caption_text <- paste(readLines(cap_file, warn = FALSE), collapse = " ")
  } else {
    caption_text <- paste("No caption found for", metro, rq)
  }

  # Read the note (if missing, it remains empty)
  if (file.exists(note_file)) {
    note_text <- paste(readLines(note_file, warn = FALSE), collapse = " ")
    # Format the note to appear after the figure
    note_string <- paste0("\n\n**Note**: ", note_text)
  } else {
    note_string <- ""
  }

  # Build the Quarto figure block as a string
  # - We use fig-cap= for the caption
  # - We can specify width, alignment, etc.
  # - Insert the note at the bottom if present
  fig_block <- paste0(
    "::: {float='figure' fig-align='center' fig-cap='", caption_text, "'}\n\n",
    "![](", fig_file, "){width='70%'}\n",
    note_string, "\n",
    ":::\n"
  )

  return(fig_block)
}

```

### 2025-01-30 text/front_end_v2.qmd
tweaker

@@
 
+**ABSTRACT**
+
+In 1984, the College Board created "Geomarkets," which carve states and metropolitan areas into smaller geographic units meant to define local recruiting markets.
+Geomarkets are an input in two College Board products.
+First, the Enrollment Planning Service (EPS) software makes recommendations about which Geomarkets and high schools admissions recruiters should recruit from.
+Second, Geomarkets became a search filter within the Student Search Service product, which enables colleges to buy "student lists" containing the contact information of prospective students.
+The concept of Geomarkets was introduced by *The Structure of College Choice* [@RN4982].
+This book develops the "Market Segment Model," which argues that student demand for a particular college is a function of class and geography.
+We conceptualize Geomarkets and the Market Segment Model by drawing from scholarship on quantification, particularly the discussions of correlation and homophily by @RN4975.
+We address two research questions: First, what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
+Second, how does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
+We answer RQ1 by analyzing Census data from 1980, 2000, and 2020.
+We answer RQ2 by analyzing data from actual student lists purchased by public universities.
+Metropolitan areas are cases.
+Analyses consist of graphs of descriptive statistics and interactive maps.
+
 # Introduction
@@
 11] describe the Market Segment Model as an effort to "quantify" the knowledge that college admissions officers have about the geography of student demand:
+
+@list_empirics say this:
 

### 2025-01-30 scripts/rq1_tables_graphs.R

@@ -371,14 +371,36 @@ create_eps_graph <- function(eps_codes,
       )
   }
   
+  file_prefix_underscore <- str_replace_all(string = file_prefix, pattern = ' ', replacement = '_')
+  writeLines(file_prefix)
+  
+  # create strings for file names
+  plot_name <- str_c('rq1',file_prefix_underscore,graph_type, sep = '_')
+  writeLines(plot_name)
+  
   # Save plot to file
   ggsave(
-    filename = file.path(graphs_dir, str_c(file_prefix, '_', graph_type, '.png')),
+    filename = file.path(graphs_dir, str_c(plot_name, '.png')),
     plot = plot,
-    width = 10,
+    width = 14,
     height = 8,
     bg = 'white'
   )
+  
+  # Define the figure title
+  if (graph_type == "race") {
+    which_characteristics <- 'Racial/ethnic composition' 
+  } else if (graph_type == "ses") {
+    which_characteristics <- 'Socioeconomic characteristics' 
+  }
+  figure_title <- str_c(
+    which_characteristics, "of", str_to_title(file_prefix), "area Geomarkets", sep = ' '
+  )
+  
+  writeLines(figure_title)
+  
+  # Save the title to a .txt file
+  writeLines(figure_title, file.path(graphs_dir, str_c(plot_name, 'title.txt', sep = '_')))    
   
   # 2) Create the text you want to store [REVISE NOTE TEXT LATER!]
   note_text <- c(
@@ -390,35 +412,44 @@ create_eps_graph <- function(eps_codes,
   )
   
   # 3) Write that text to a file
-  writeLines(note_text, file.path(graphs_dir, str_c(file_prefix, '_', graph_type, '.txt')))
+  writeLines(note_text, file.path(graphs_dir, str_c(plot_name, 'note.txt', sep = '_')))    
              
   # Return the plot
   return(plot)
 }
 
+create_eps_graph(
+  eps_codes = dallas_eps_codes, 
+  prefix = 'dallas',
+  graph_type = "ses", 
+  note = "additional text for my note!"
+)
+
+
 all_codes <- list(
-  philly = philly_eps_codes,
-  dallas = dallas_eps_codes,
-  atl = atl_eps_codes,
-  chicago = chi_eps_codes,
-  cleveland = cleveland_eps_codes,
[…]

### 2025-01-31 text/cb_geomarket_online_appendix.qmd
creating appendix doc with all graphs

@@
     number-sections: true
+    documentclass: article
+    classoption: landscape
+    geometry: margin=0.1in
+    include-in-header:
+      text: |
+        \pagenumbering{gobble}    
+header-includes:
+  - |
+    \usepackage{ragged2e}
+    \usepackage{caption}
+    \captionsetup[figure]{justification=RaggedRight, singlelinecheck=false}    
 ---
 
-# header 1
+```{r, include = FALSE, eval=FALSE}
+getwd()
+file.exists("../results/graphs/rq1_chicago_ses.png")
 
-```{r}
-getwd()
 ```
+<!-- this code run in markdown would include the graph: ![](../results/graphs/rq1_chicago_ses.png){width=70%} -->
 
-```{r}
+```{r, include = FALSE}
 # You can load libraries, set options, define your list of metros, etc.
 library(tools)  # for toTitleCase() if desired
+library(tidyverse)
 
 # Example lists
-metros <- c("chicago", "denver")
-RQs <- c("RQ1", "RQ2")
+metros <- c(
+  "atlanta",
+  "bay area",
+  "boston",
+  "chicago",
+  "cleveland",
+  "dallas",
+  "dc maryland virginia",
+  "detroit",
+  "houston",
+  "long island",
+  "los angeles",
+  "miami",
+  "new york city",
+  "northern new jersey",
+  "orange county",
+  "philadelphia",
+  "san diego"
+)
+
+metros <- c(
+  "atlanta",
+  "chicago",
+  "philadelphia"
+)
+
+rqs <- c("rq1",'rq2')
+graph_types = c('race','ses')
 
@@
 # 'base_path' is where your .png and .txt files live.
-insert_figure <- function(metro, rq, base_path = "../results/graphs") {
+insert_figure <- function(rq, metro, graph_type,base_path = "../results/graphs") {
 
+  metro <- str_replace_all(string = metro, pattern = ' ', replacement = '_')
   # Construct file paths
-  fig_file    <- file.path(base_path, paste0(metro, "_", rq, ".png"))
-  cap_file    <- file.path(base_path, paste0(metro, "_", rq, "_caption.txt"))
-  note_file   <- file.path(base_path, paste0(metro, "_", rq, "_note.txt"))
+  fig_file    <- file.path(base_path, str_c(rq,'_',metro,'_',graph_type, ".png"))
+  cap_file    <- file.path(base_path, str_c(rq,'_',metro,'_',graph_type,"_title.txt"))
+  note_file    <- file.path(base_path, str_c(rq,'_',metro,'_',graph_type,"_note.txt"))
 
@@
     # Format the note to appear after the figure
-    note_string <- paste0("\n\n**Note**: ", note_text)
+    note_string <- paste0("\n**Note**: ", note_text)
   } else {
@@
   # - Insert the note at the bottom if present
-  fig_block <- paste0(
-    "::: {float='figure' fig-align='center' fig-cap='", caption_text, "'}\n\n",
-    "![](", fig_file, "){width='70%'}\n",
+fig_block <- str_c(
+  "::: {.figure}\n\n",
+  "![", caption_text, "](", fig_file, "){fig-align='left' width='101%'}\n",
   note_string, "\n",
@@
 
+#rq2_philadelphia_order_448922_race_by_firstgen
+#rq1_philadelphia_ses
+#rq2_philadelphia_firstgen_row_plot
+
+#insert_figure(rq = 'rq2', metro = 'philadelphia', graph_type = 'firstgen_row_plot',base_path = "../results/graphs")
+#insert_figure(rq = 'rq2', metro = 'philadelphia', graph_type = 'order_448922_race_by_firstgen',base_path = "../results/graphs")
+#insert_figure(rq = 'rq1', metro = 'dallas', graph_type = 'ses',base_path = "../results/graphs")
 ```
+
+```{r, results = 'asis', echo = FALSE}
+# results='asis' so that the text we "cat()" is interpreted as Markdown.
+
+for (m in metros) {
+
+  # Print a top-level heading for the metro area
+  cat("\n# ", toTitleCase(m), " area\n", sep="")
+
+  for (rq in rqs) {
+
+    rq_num <- str_replace(string = rq, pattern = "rq", replacement = "")
+
+    if (rq == 'rq1') { # rq1 run for every metro area
+
+      # Print a second-level heading for the RQ
+      cat("## Research question ", rq_num, "\n", sep="")
+
+      for (g in graph_types) {
+
+        # Call our figure-inserting function
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = g)
+
+        # Output the figure block to the doc
+        cat(figure_md, "\n")
+      } # for (g in graph_types) {
+
+    } else if (rq == 'rq2' & (m %in% c('chicago','philadelphia'))) { # rq2 run only for specific metro areas and orders
+
+      # Print a second-level heading for the RQ
+      cat("## Research question ", rq_num, "\n", sep="")
+
+      if (m == 'chicago') {
+
+        # Call our figure-inserting function
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_487984_race_by_firstgen')
+
+        # Output the figure block to the doc
+        cat(figure_md, "\n")
+
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_488035_488053_race_by_firstgen')
+
+        # Output the figure block to the doc
+        cat(figure_md, "\n")
+
+      } else if (m == 'philadelphia') {
+
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'race_row_plot')
+        cat(figure_md, "\n")
+
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'race_col_plot')
+        cat(figure_md, "\n")
+
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'firstgen_row_plot')
+        cat(figure_md, "\n")
+
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'firstgen_col_plot')
+        cat(figure_md, "\n")
+
+        # Call our figure-inserting function
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_448922_race_by_firstgen')
+
+        # Output the figure block to the doc
+        cat(figure_md, "\n")
+
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'order_448427_448440_race_by_firstgen')
+
+        # Output the figure block to the doc
+        cat(figure_md, "\n")
+
+      } #else if (condition3) {
+        # Code block for condition3
+      #}
+
+    }
+
+  } # for (rq in rqs) {
+
+}
+
+```

### 2025-01-31 scripts/rq2_tables_graphs_create_functions.R

@@ -1,0 +1,1067 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_stu_list.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 1/16/2025
+## [ DESC ] < Create tables and graphs for RQ2 of Geomarket paper >
+################################################################################
+
+### SETTINGS
+#rm(list = ls()) # remove all objects
+options(max.print=1000)
+#options(width = 160)
+# Set the scipen option to a high value to avoid scientific notation
+options(scipen = 999)
+
+### LIBRARIES
+library(tidyverse)
+library(formattable)
+library(patchwork)
+library(scales)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+# get census api key
+#http://api.census.gov/data/key_signup.html
+# set census api key, and install for future use
+#census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+#Sys.getenv("CENSUS_API_KEY") # retreive API key
+
+
+library(sf)
+# Enable caching for tigris
+options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+library(lwgeom) # library has checks/vixes for valid geometries
+library(leaflet)
+library(shiny)
+library(kableExtra)
+
+######### CREATE ANALYSIS DATASET(S) FOR RQ2 TABLES N' GRAPHS
+
+#getwd()
+#source(file = file.path('scripts', 'create_stu_list.R'))
+#getwd()
+
+
+# RACE TABLE FUNCTION FUNCTION. CREATES TABLE W/ PERCENT OF PEOPLE IN EACH RACIAL GROUP X GEOMARKET FOR AN ORDER(S)
+
+# Define the function
+
+create_sim_eps_race_table <- function(data, ord_nums, eps_codes) {
+  df <- data %>%
+    as.data.frame() %>%
+    filter(ord_num %in% ord_nums, eps %in% eps_codes) %>%
+    mutate(eps_codename = str_c(eps, eps_name, sep = ", ")) %>% 
+    group_by(eps_codename) %>%
+    summarize(
+      stu_all = n(),
[…]

### 2025-01-31 scripts/rq2_tables_graphs_run_functions.R

@@ -1,0 +1,2031 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < create_stu_list.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 1/16/2025
+## [ DESC ] < Create tables and graphs for RQ2 of Geomarket paper >
+################################################################################
+
+### SETTINGS
+#rm(list = ls()) # remove all objects
+options(max.print=1000)
+#options(width = 160)
+# Set the scipen option to a high value to avoid scientific notation
+options(scipen = 999)
+
+### LIBRARIES
+library(tidyverse)
+library(formattable)
+library(patchwork)
+library(scales)
+library(readxl)
+library(lubridate)
+library(haven)
+library(labelled)
+library(tidycensus)
+# get census api key
+#http://api.census.gov/data/key_signup.html
+# set census api key, and install for future use
+#census_api_key('aff360d1fe8a919619776f48e975f03b8bb1379e', install = TRUE)
+#Sys.getenv("CENSUS_API_KEY") # retreive API key
+
+
+library(sf)
+# Enable caching for tigris
+options(tigris_use_cache = TRUE)
+library(tigris)
+#library(stars)
+#library(spatstat)
+#library(rgeos)
+library(lwgeom) # library has checks/vixes for valid geometries
+library(leaflet)
+library(shiny)
+library(kableExtra)
+
+
+######### CREATE ANALYSIS DATASET(S) FOR RQ2 TABLES N' GRAPHS
+
+getwd()
+source(file = file.path('scripts', 'create_stu_list.R'))
+getwd()
+
+# script that creates functions to create tables and graphs
+source(file = file.path(scripts_dir, 'rq2_tables_graphs_create_functions.R'))
+
+###########################
+############################ RUN FUNCTION create_sim_eps_graph TO CREATE GRAPH OF RACE AND GRAPH OF FIRSTGEN
+###########################
+
+all_orders <- list(
+  chicago = c('chicago', 'chi_eps_codes', 'SAT score 1020 - 1150', '487984'),
+  chicago = c('chicago', 'chi_eps_codes', 'SAT score 1160 - 1600', '488035', '488053'),
[…]

### 2025-02-01 text/cb_geomarket_online_appendix.qmd
more progress creating appendix with all results

@@
 
-metros <- c(
-  "atlanta",
-  "chicago",
-  "philadelphia"
-)
+#metros <- c(
+#  "atlanta",
+#  "chicago",
+#  "philadelphia"
+#)
 
@@
 
+      # separate rq2 graphs for race and for firstgen status; they apply to all above metro areas
+      for (suffix in c('race_row_plot','race_col_plot','firstgen_row_plot','firstgen_col_plot')) {
+
+        figure_md <- insert_figure(metro = m, rq = rq, graph_type = suffix)
+        cat(figure_md, "\n")
+
+        }
+
       if (m == 'chicago') {
@@
       } else if (m == 'philadelphia') {
-
-        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'race_row_plot')
-        cat(figure_md, "\n")
-
-        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'race_col_plot')
-        cat(figure_md, "\n")
-
-        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'firstgen_row_plot')
-        cat(figure_md, "\n")
-
-        figure_md <- insert_figure(metro = m, rq = rq, graph_type = 'firstgen_col_plot')
-        cat(figure_md, "\n")
 

### 2025-02-01 text/front_end_v2.qmd
more progress creating appendix with all results

@@
 
-In 1984, the College Board created "Geomarkets," which carve states and metropolitan areas into smaller geographic units meant to define local recruiting markets.
+College Board "Geomarkets" carve states and metropolitan areas into smaller geographic units meant to define local recruiting markets.
+Geomarkets were initially introduced in *The Structure of College Choice* [@RN4982].
+This book develops the "Market Segment Model," which argues that student demand for a particular college is a function of class and geography.
 Geomarkets are an input in two College Board products.
@@
 Second, Geomarkets became a search filter within the Student Search Service product, which enables colleges to buy "student lists" containing the contact information of prospective students.
-The concept of Geomarkets was introduced by *The Structure of College Choice* [@RN4982].
-This book develops the "Market Segment Model," which argues that student demand for a particular college is a function of class and geography.
 We conceptualize Geomarkets and the Market Segment Model by drawing from scholarship on quantification, particularly the discussions of correlation and homophily by @RN4975.

### 2025-02-01 scripts/rq2_tables_graphs_run_functions.R

@@ -109,9 +109,7 @@ rq2_philadelphia_firstgen_plot_order_448427_448440$plot_r
 rq2_philadelphia_firstgen_plot_order_448427_448440$plot_c
 
 
-
-
-# combined graphs
+# Create combined graphs for Philadelphia
 # race row pct. for lower SAT score and higher SAT score
 rq2_philadelphia_race_row_plot <- rq2_philadelphia_race_plot_order_448922$plot_r + rq2_philadelphia_race_plot_order_448427_448440$plot_r + plot_layout(ncol = 1)
 # race column pct. for lower SAT score and higher SAT score
@@ -122,16 +120,53 @@ rq2_philadelphia_firstgen_row_plot <- rq2_philadelphia_firstgen_plot_order_44892
 # firstgen column pct. for lower SAT score and higher SAT score
 rq2_philadelphia_firstgen_col_plot <- rq2_philadelphia_firstgen_plot_order_448922$plot_c + rq2_philadelphia_firstgen_plot_order_448427_448440$plot_c + plot_layout(ncol = 1)
 
+# create combined graphs for Chicago
+
+# race row pct. for lower SAT score and higher SAT score
+rq2_chicago_race_row_plot <- rq2_chicago_race_plot_order_487984$plot_r + rq2_chicago_race_plot_order_488035_488053$plot_r + plot_layout(ncol = 1)
+# race column pct. for lower SAT score and higher SAT score
+rq2_chicago_race_col_plot <- rq2_chicago_race_plot_order_487984$plot_c + rq2_chicago_race_plot_order_488035_488053$plot_c + plot_layout(ncol = 1)
+
+# firstgen row pct. for lower SAT score and higher SAT score
+rq2_chicago_firstgen_row_plot <- rq2_chicago_firstgen_plot_order_487984$plot_r + rq2_chicago_firstgen_plot_order_488035_488053$plot_r + plot_layout(ncol = 1)
+# firstgen column pct. for lower SAT score and higher SAT score
+rq2_chicago_firstgen_col_plot <- rq2_chicago_firstgen_plot_order_487984$plot_c + rq2_chicago_firstgen_plot_order_488035_488053$plot_c + plot_layout(ncol = 1)
+
+
 # loop that saves combined graphs as .png and creates title.txt and note.txt
-for (m in c('philadelphia')) {
+for (m in c('chicago','philadelphia')) {
[…]

### 2025-02-02 index.qmd
add initial draft of slides

---
title: 'Title'
subtitle: 'Subtitle'
author: ['Author', 'Author']
bibliography: ./assets/bib/cb_geomarket_bib.bib
format:
  revealjs: 
    theme: [default, ./assets/css/custom.scss]
    controls: true
    controls-layout: bottom-right
    controls-tutorial: true
    transition: slide
    background-transition: fade
    auto-stretch: false
    slide-level: 3
    template-partials:
      - ./assets/html/title-slide.html
include-in-header: ./assets/html/header.html
---

```{r setup, include = FALSE}
library(tidyverse)
library(kableExtra)  # For collapse_rows() to work: https://stackoverflow.com/a/67803915/6373540
library(tools)

knitr::opts_chunk$set(echo = F, message = F)

results_dir <- file.path('.', 'results')
graphs_dir <- file.path(results_dir, 'graphs')

metros <- c(
  'atlanta',
  'bay area',
  'boston',
  'chicago',
  'cleveland',
  'dallas',
  'dc maryland virginia',
  'detroit',
  'houston',
  'long island',
  'los angeles',
  'miami',
  'new york city',
  'northern new jersey',
  'orange county',
  'philadelphia',
  'san diego'
)

rqs <- c('rq1', 'rq2')
graph_types <- c('race', 'ses')
```

# Level-1 heading

## Level-2 heading

### Level-3 heading

### Level-3 heading
#### Level-4 subheading

### Test map

<iframe src="https://comforting-kashata-600ce0.netlify.app/" width=100% height=100% allowtransparency="true"></iframe>

# Results

```{r, echo=FALSE, results='asis'}
insert_figure <- function(rq, metro, graph_type, base_path = graphs_dir) {
  metro <- str_replace_all(string = metro, pattern = ' ', replacement = '_')

  fig_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '.png'))
  cap_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '_title.txt'))
  note_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '_note.txt'))

  # Print caption
  writeLines(paste0('#### ', readLines(cap_file, warn = F)))

  # Output the figure block to the doc
  writeLines(paste0('![](', fig_file, ')\n'))

  # Print footnotes
  writeLines(paste0('<div class="footnote">', paste0(readLines(note_file, warn = F), collapse = '</br>'), '</div>'))
}

for (m in metros) {

  # Print a top-level heading for the metro area
  writeLines(paste0('## ', toTitleCase(m), ' area'))

  for (rq in rqs) {
    rq_num <- str_replace(string = rq, pattern = 'rq', replacement = '')

    if (rq == 'rq1') { # rq1 run for every metro area

      for (g in graph_types) {

        # Print a heading for the RQ
        writeLines(paste0('### Research question ', rq_num))

        insert_figure(metro = m, rq = rq, graph_type = g)
      }

    } else if (rq == 'rq2' & (m %in% c('chicago','philadelphia'))) { # rq2 run only for specific metro areas and orders

      # Separate rq2 graphs for race and for firstgen status; they apply to all above metro areas
      for (suffix in c('race_row_plot', 'race_col_plot', 'firstgen_row_plot', 'firstgen_col_plot')) {

        # Print a heading for the RQ
        writeLines(paste0('### Research question ', rq_num))

        insert_figure(metro = m, rq = rq, graph_type = suffix)

      }

      if (m == 'chicago') {

        # Print a heading for the RQ
        writeLines(paste0('### Research question ', rq_num))

        insert_figure(metro = m, rq = rq, graph_type = 'order_487984_race_by_firstgen')

        # Print a heading for the RQ
        writeLines(paste0('### Research question ', rq_num))

        insert_figure(metro = m, rq = rq, graph_type = 'order_488035_488053_race_by_firstgen')

      } else if (m == 'philadelphia') {

        # Print a heading for the RQ
        writeLines(paste0('### Research question ', rq_num))

        insert_figure(metro = m, rq = rq, graph_type = 'order_448922_race_by_firstgen')

        # Print a heading for the RQ
        writeLines(paste0('### Research question ', rq_num))

        insert_figure(metro = m, rq = rq, graph_type = 'order_448427_448440_race_by_firstgen')

      }
    }
  }
}
```

### References

::: {#refs}
:::

### 2025-02-02 assets/js/script.js

@@ -1,0 +1,32 @@
+$(function() {
+
+  const modal = '<div class="modal"><span class="close">&times;</span><img /></div>';
+  $('.slides section').prepend(modal);
+  
+  const link = '<a class="link" href="https://ozanj.github.io/cb_geo/">ozanj.github.io/cb_geo</a>';
+  $('.slides section.level3').prepend(link);
+  
+  $('.slides section img').on('click', function() {
+    let src = $(this).attr('src');
+    
+    var $slide = $(this).closest('.slide');
+    $slide.find('.modal img').attr('src', src);
+    
+    $(this).closest('.slide').find('.modal').fadeIn(600);
+    $(this).closest('.slide').find('p img').addClass('disabled');
+  });
+  
+  $('.close').on('click', function() {
+    $(this).closest('.slide').find('.modal').fadeOut();
+    $(this).closest('.slide').find('p img').removeClass('disabled');
+  });
+  
+  $('<hr>').insertAfter('.reveal .slide h4');
+  $('<hr>').insertAfter('.reveal .slide:not(.caption) h3:not(:has(+ h4))');
+  
+  Reveal.addEventListener('slidechanged', function() {
+    $('.modal').fadeOut();
+    $('img').removeClass('disabled');
+  });
+  
+});

### 2025-02-04 index.qmd
working towards a more perfect union

@@
 insert_figure <- function(rq, metro, graph_type, base_path = graphs_dir) {
-  metro <- str_replace_all(string = metro, pattern = ' ', replacement = '_')
+  # Convert spaces -> underscores for filenames
+  metro <- str_replace_all(string = metro, pattern = " ", replacement = "_")
 
-  fig_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '.png'))
-  cap_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '_title.txt'))
-  note_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '_note.txt'))
+  fig_file  <- file.path(base_path, str_c(rq, "_", metro, "_", graph_type, ".png"))
+  cap_file  <- file.path(base_path, str_c(rq, "_", metro, "_", graph_type, "_title.txt"))
+  note_file <- file.path(base_path, str_c(rq, "_", metro, "_", graph_type, "_note.txt"))
 
   # Print caption
-  writeLines(paste0('#### ', readLines(cap_file, warn = F)))
+  writeLines(paste0("#### ", readLines(cap_file, warn = FALSE)))
 
   # Output the figure block to the doc
-  writeLines(paste0('![](', fig_file, ')\n'))
+  writeLines(str_c("![](", fig_file, ")\n"))
 
   # Print footnotes
-  writeLines(paste0('<div class="footnote">', paste0(readLines(note_file, warn = F), collapse = '</br>'), '</div>'))
+  writeLines(
+    str_c("<div class=\"footnote\">",
+          paste0(readLines(note_file, warn = FALSE), collapse = "</br>"),
+          "</div>")
+  )
 }
@@
   # Print a top-level heading for the metro area
-  writeLines(paste0('## ', toTitleCase(m), ' area'))
+  writeLines(str_c("## ", toTitleCase(m), " area"))
 
   for (rq in rqs) {
-    rq_num <- str_replace(string = rq, pattern = 'rq', replacement = '')
+    rq_num <- str_replace(string = rq, pattern = "rq", replacement = "")
 
-    if (rq == 'rq1') { # rq1 run for every metro area
-
+    if (rq == "rq1") { 
+      # ---------------------------------------
+      #  rq1 => run for every metro area 
+      # ---------------------------------------
       for (g in graph_types) {
-
         # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
+        writeLines(str_c("### Research question ", rq_num))
 
@@
 
-    } else if (rq == 'rq2' & (m %in% c('chicago','philadelphia'))) { # rq2 run only for specific metro areas and orders
+    } else if (rq == "rq2" & (m %in% c("chicago","philadelphia","los angeles","orange county","san diego","bay area"))) {
+      # -------------------------------------------------
+      # rq2 => run only for the specified metro areas
+      # -------------------------------------------------
 
-      # Separate rq2 graphs for race and for firstgen status; they apply to all above metro areas
-      for (suffix in c('race_row_plot', 'race_col_plot', 'firstgen_row_plot', 'firstgen_col_plot')) {
+      # 1) For LA & Bay Area, we have SEPARATE row/col plots per order (2 files each).
+      # 2) For all other metros in this subset, we still do the old "combined" row/col plots.
 
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
+      # If LA or Bay Area => do the "separate" approach
+      if (m %in% c("los angeles","bay area")) {
 
+        # We'll insert each of the 8 new plots: race_row_plot_..., race_col_plot_...,
+        # and firstgen_row_plot_..., firstgen_col_plot_...
+        # Each suffix also ends with the order ID.
+        for (suffix in c(
+          "race_row_plot_448375", "race_row_plot_448374_448420",
+          "race_col_plot_448375", "race_col_plot_448374_448420",
+          "firstgen_row_plot_448375", "firstgen_row_plot_448374_448420",
+          "firstgen_col_plot_448375", "firstgen_col_plot_448374_448420"
+        )) {
+          writeLines(str_c("### Research question ", rq_num))
           insert_figure(metro = m, rq = rq, graph_type = suffix)
-
         }
 
-      if (m == 'chicago') {
+      } else {
+        # For the other metros in the same set
+        # (chicago, philadelphia, orange county, san diego):
+        # we still show the combined row/col figure:
 
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
-
-        insert_figure(metro = m, rq = rq, graph_type = 'order_487984_race_by_firstgen')
-
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
-
-        insert_figure(metro = m, rq = rq, graph_type = 'order_488035_488053_race_by_firstgen')
-
-      } else if (m == 'philadelphia') {
-
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
-
-        insert_figure(metro = m, rq = rq, graph_type = 'order_448922_race_by_firstgen')
-
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
-
-        insert_figure(metro = m, rq = rq, graph_type = 'order_448427_448440_race_by_firstgen')
-
+        for (suffix in c("race_row_plot", "race_col_plot",
+                         "firstgen_row_plot", "firstgen_col_plot")) {
+          writeLines(str_c("### Research question ", rq_num))
+          insert_figure(metro = m, rq = rq, graph_type = suffix)
         }
       }
+
+      # -------------------------------------------------
+      # Then your custom calls by city:
+      # -------------------------------------------------
+      if (m == "chicago") {
+        # two separate calls for 'chicago' order IDs
+        writeLines(str_c("### Research question ", rq_num))
+        insert_figure(metro = m, rq = rq, graph_type = "order_487984_race_by_firstgen")
+
+        writeLines(str_c("### Research question ", rq_num))
+        insert_figure(metro = m, rq = rq, graph_type = "order_488035_488053_race_by_firstgen")
+
+      } else if (m == "philadelphia") {
+        # likewise for 'philadelphia'
+        writeLines(str_c("### Research question ", rq_num))
+        insert_figure(metro = m, rq = rq, graph_type = "order_448922_race_by_firstgen")
+
+        writeLines(str_c("### Research question ", rq_num))
+        insert_figure(metro = m, rq = rq, graph_type = "order_448427_448440_race_by_firstgen")
+
+      } else if (m %in% c("los angeles","orange county","san diego","bay area")) {
+        # likewise for these 4 metros
+        writeLines(str_c("### Research question ", rq_num))
+        insert_figure(metro = m, rq = rq, graph_type = "order_448375_race_by_firstgen")
+
+        writeLines(str_c("### Research question ", rq_num))
+        insert_figure(metro = m, rq = rq, graph_type = "order_448374_448420_race_by_firstgen")
       }
-}
+
+    } # end of rq==rq2 section
+
+  } # end of rqs
+
+} # end of metros
+
 ```

### 2025-02-04 scripts/metro_eps_codes.R

@@ -10,8 +10,8 @@ library(tidyverse)
 
 
 
-bay_eps_codes <- c(paste0("CA ", 4:9), paste0("CA", 10:11)) # bay area
-bay_eps_codes 
+bay_area_eps_codes <- c(paste0("CA ", 4:9), paste0("CA", 10:11)) # bay area
+bay_area_eps_codes
 
 ca_eps_codes <- c(paste0("CA ", 1:9), paste0("CA", 10:34))
 chi_eps_codes <- c(paste0("IL ", 7:9), paste0("IL", 10:13))

### 2025-02-04 scripts/rq2_tables_graphs_run_functions.R

@@ -60,9 +60,50 @@ all_orders <- list(
   chicago = c('chicago', 'chi_eps_codes', 'SAT score 1020 - 1150', '487984'),
   chicago = c('chicago', 'chi_eps_codes', 'SAT score 1160 - 1600', '488035', '488053'),
   philadelphia = c('philadelphia', 'philly_eps_codes', 'PSAT score 1070 - 1180', '448922'),
-  philadelphia = c('philadelphia', 'philly_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440')
+  philadelphia = c('philadelphia', 'philly_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440'),
+  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1110 - 1210', '366935'), # ordered jan 2018; don't like these cutpoints as much
+  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1220 - 1520', '366934', '366932'),
+  los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1070 - 1180', '448375'), # ordered jan 2019
+  los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1220 - 1520', '448374', '448420'),
+  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1070 - 1180', '546954'), # ordered jan 2020; much smaller sample size than orders from jan 2019
+  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1220 - 1520', '546946', '546945'),
+  #orange_county = c('orange_county', 'orange_county_eps_codes', 'PSAT score 1110 - 1210', '366935'), # ordered jan 2018; don't like these cutpoints as much
+  #orange_county = c('orange_county', 'orange_county_eps_codes', 'PSAT score 1220 - 1520', '366934', '366932'),
+  orange_county = c('orange_county', 'orange_county_eps_codes', 'PSAT score 1070 - 1180', '448375'), # ordered jan 2019
+  orange_county = c('orange_county', 'orange_county_eps_codes', 'PSAT score 1220 - 1520', '448374', '448420'),
[…]

### 2025-02-05 index.qmd
changed how results organized

--- 2025-02-02 index.qmd
@@
 
-```{r setup, include = FALSE}
-library(tidyverse)
-library(kableExtra)  # For collapse_rows() to work: https://stackoverflow.com/a/67803915/6373540
-library(tools)
-
-knitr::opts_chunk$set(echo = F, message = F)
-
-results_dir <- file.path('.', 'results')
-graphs_dir <- file.path(results_dir, 'graphs')
-
-metros <- c(
-  'atlanta',
-  'bay area',
-  'boston',
-  'chicago',
-  'cleveland',
-  'dallas',
-  'dc maryland virginia',
-  'detroit',
-  'houston',
-  'long island',
-  'los angeles',
-  'miami',
-  'new york city',
-  'northern new jersey',
-  'orange county',
-  'philadelphia',
-  'san diego'
-)
-
-rqs <- c('rq1', 'rq2')
-graph_types <- c('race', 'ses')
-```
-
 # Level-1 heading
@@
 
+```{r setup, echo=FALSE, results = 'asis'}
 ```{r, echo=FALSE, results='asis'}
+library(tidyverse)
+library(kableExtra)  # For collapse_rows() to work
+library(tools)
+
+knitr::opts_chunk$set(echo = F, message = F)
+
+# Example directories
+graphs_dir <- file.path(".", "results", "graphs")
+
+# 1) Define the `insert_figure` helper function
 insert_figure <- function(rq, metro, graph_type, base_path = graphs_dir) {
-  metro <- str_replace_all(string = metro, pattern = ' ', replacement = '_')
 
-  fig_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '.png'))
-  cap_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '_title.txt'))
-  note_file <- file.path(base_path, str_c(rq, '_', metro, '_', graph_type, '_note.txt'))
+  # Convert spaces -> underscores for filenames
+  metro_underscored <- str_replace_all(metro, " ", "_")
 
-  # Print caption
-  writeLines(paste0('#### ', readLines(cap_file, warn = F)))
+  # Build the three file paths (PNG, title TXT, note TXT)
+  fig_file  <- file.path(base_path, rq, str_c(rq, "_", metro_underscored, "_", graph_type, ".png"))
+  cap_file  <- file.path(base_path, rq, str_c(rq, "_", metro_underscored, "_", graph_type, "_title.txt"))
+  note_file <- file.path(base_path, rq, str_c(rq, "_", metro_underscored, "_", graph_type, "_note.txt"))
 
-  # Output the figure block to the doc
-  writeLines(paste0('![](', fig_file, ')\n'))
+  # Print a heading / caption
+  caption_text <- if (file.exists(cap_file)) readLines(cap_file, warn = FALSE) else "No Title Found"
+  writeLines(paste0("#### ", caption_text))
 
-  # Print footnotes
-  writeLines(paste0('<div class="footnote">', paste0(readLines(note_file, warn = F), collapse = '</br>'), '</div>'))
+  # Output the figure (if it exists)
+  if (file.exists(fig_file)) {
+    writeLines(str_c("![](", fig_file, ")\n"))
+  } else {
+    writeLines(paste("*(Missing figure file:", fig_file, ")*\n"))
   }
 
+  # Print footnotes (if the .txt file exists)
+  if (file.exists(note_file)) {
+    notes_txt <- readLines(note_file, warn = FALSE)
+    writeLines(
+      str_c(
+        "<div class=\"footnote\">",
+        paste0(notes_txt, collapse = "</br>"),
+        "</div>"
+      )
+    )
+  }
+}
+
+# 2) Define your metros / RQs
+metros <- c(
+  "atlanta",
+  "bay area",
+  "boston",
+  "chicago",
+  "cleveland",
+  "dallas",
+  "dc maryland virginia",
+  "detroit",
+  "houston",
+  "long island",
+  "los angeles",
+  "miami",
+  "new york city",
+  "northern new jersey",
+  "orange county",
+  "philadelphia",
+  "san diego"
+)
+
+rqs <- c("rq1", "rq2")
+
+# For RQ1, we have two plot types: "race" and "ses"
+graph_types_rq1 <- c("race", "ses")
+
+# 3) Subset of metros with RQ2 data (2 orders each)
+rq2_metros <- c(
+  "chicago", "philadelphia", "los angeles", 
+  "orange county", "san diego", "bay area",
+  "northern new jersey"
+)
+
+# We'll define a "metro_orders" list for those RQ2 metros:
+metro_orders <- list(
+  "chicago"             = c("487984", "488035_488053"),
+  "philadelphia"        = c("448922", "448427_448440"),
+  "northern_new_jersey" = c("448922", "448427_448440"),
+  "los_angeles"         = c("448375", "448374_448420"),
+  "orange_county"       = c("448375", "448374_448420"),
+  "san_diego"           = c("448375", "448374_448420"),
+  "bay_area"            = c("448375", "448374_448420")
+)
+
+# 4) Main loop
 for (m in metros) {
@@
   # Print a top-level heading for the metro area
-  writeLines(paste0('## ', toTitleCase(m), ' area'))
+  writeLines(str_c("## ", tools::toTitleCase(m), " area"))
 
   for (rq in rqs) {
-    rq_num <- str_replace(string = rq, pattern = 'rq', replacement = '')
 
-    if (rq == 'rq1') { # rq1 run for every metro area
+    # For clarity, parse out the numeric portion
+    rq_num <- str_replace(string = rq, pattern = "rq", replacement = "")
 
-      for (g in graph_types) {
+    # ---------------------------------------------
+    # RQ1 Logic
+    # ---------------------------------------------
+    if (rq == "rq1") {
 
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
-
-        insert_figure(metro = m, rq = rq, graph_type = g)
+      # Each metro has 2 plots: "race" & "ses"
+      for (g in graph_types_rq1) {
+        writeLines(str_c("### Research question ", rq_num))
+        insert_figure(rq = rq, metro = m, graph_type = g)
       }
 
-    } else if (rq == 'rq2' & (m %in% c('chicago','philadelphia'))) { # rq2 run only for specific metro areas and orders
+    # ---------------------------------------------
+    # RQ2 Logic (row vs. col approach)
+    # ---------------------------------------------
+    } else if (rq == "rq2") {
 
-      # Separate rq2 graphs for race and for firstgen status; they apply to all above metro areas
-      for (suffix in c('race_row_plot', 'race_col_plot', 'firstgen_row_plot', 'firstgen_col_plot')) {
+      # Only run RQ2 code if the metro is in our RQ2 list
+      if (!(m %in% rq2_metros)) next
 
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
+      # Convert name -> underscores for matching list keys
+      m_underscored <- str_replace_all(m, " ", "_")
+      orders_for_m  <- metro_orders[[m_underscored]]
 
-        insert_figure(metro = m, rq = rq, graph_type = suffix)
+      # If a metro isn't in the list or no orders, skip
+      if (is.null(orders_for_m)) next
 
+      # We'll do 2 group types: race & firstgen
+      for (g in c("race", "firstgen")) {
+
+        # We'll do row and col for each group type
+        for (rc in c("row", "col")) {
+
+          if (rc == "row") {
+            #
+            # ROW => single combined figure on one slide
+            #
+            writeLines(str_c("### Research question ", rq_num))
+            graph_type_row <- str_c(g, "_", rc, "_plot")
+            insert_figure(rq = rq, metro = m, graph_type = graph_type_row)
+
+          } else {
+            #
+            # COL => separate file(s) by order ID => each on its own slide
+            #
+            for (order_id in orders_for_m) {
+
+              # Each order ID gets its own "###" => new Reveal.js slide
+              writeLines(str_c("### Research question ", rq_num))
+
+              # e.g. "rq2_chicago_race_col_plot_487984.png"
+              graph_type_col <- str_c(g, "_", rc, "_plot_", order_id)
+              insert_figure(rq = rq, metro = m, graph_type = graph_type_col)
             }
 
-      if (m == 'chicago') {
+          } # end if row vs col
 
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
+        } # end rc loop
 
-        insert_figure(metro = m, rq = rq, graph_type = 'order_487984_race_by_firstgen')
+      } # end g loop
 
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
+    } # end if (rq == "rq2")
 
-        insert_figure(metro = m, rq = rq, graph_type = 'order_488035_488053_race_by_firstgen')
+  } # end rq loop
 
-      } else if (m == 'philadelphia') {
+} # end metros loop
 
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
-
-        insert_figure(metro = m, rq = rq, graph_type = 'order_448922_race_by_firstgen')
-
-        # Print a heading for the RQ
-        writeLines(paste0('### Research question ', rq_num))
-
-        insert_figure(metro = m, rq = rq, graph_type = 'order_448427_448440_race_by_firstgen')
-
-      }
-    }
-  }
-}
 ```

### 2025-02-05 index_old.qmd
changed how results organized

---
title: 'Title'
subtitle: 'Subtitle'
author: ['Author', 'Author']
bibliography: ./assets/bib/cb_geomarket_bib.bib
format:
  revealjs: 
    theme: [default, ./assets/css/custom.scss]
    controls: true
    controls-layout: bottom-right
    controls-tutorial: true
    transition: slide
    background-transition: fade
    auto-stretch: false
    slide-level: 3
    template-partials:
      - ./assets/html/title-slide.html
include-in-header: ./assets/html/header.html
---

```{r setup, include = FALSE}
library(tidyverse)
library(kableExtra)  # For collapse_rows() to work: https://stackoverflow.com/a/67803915/6373540
library(tools)

knitr::opts_chunk$set(echo = F, message = F)

results_dir <- file.path('.', 'results')
graphs_dir <- file.path(results_dir, 'graphs')

metros <- c(
  'atlanta',
  'bay area',
  'boston',
  'chicago',
  'cleveland',
  'dallas',
  'dc maryland virginia',
  'detroit',
  'houston',
  'long island',
  'los angeles',
  'miami',
  'new york city',
  'northern new jersey',
  'orange county',
  'philadelphia',
  'san diego'
)

rqs <- c('rq1', 'rq2')
graph_types <- c('race', 'ses')
```

# Level-1 heading

## Level-2 heading

### Level-3 heading

### Level-3 heading
#### Level-4 subheading

### Test map

<iframe src="https://comforting-kashata-600ce0.netlify.app/" width=100% height=100% allowtransparency="true"></iframe>

# Results

```{r, echo=FALSE, results='asis'}
insert_figure <- function(rq, metro, graph_type, base_path = graphs_dir) {
  # Convert spaces -> underscores for filenames
  metro <- str_replace_all(string = metro, pattern = " ", replacement = "_")

  fig_file  <- file.path(base_path,rq, str_c(rq, "_", metro, "_", graph_type, ".png"))
  cap_file  <- file.path(base_path,rq, str_c(rq, "_", metro, "_", graph_type, "_title.txt"))
  note_file <- file.path(base_path,rq, str_c(rq, "_", metro, "_", graph_type, "_note.txt"))

  # Print caption
  writeLines(paste0("#### ", readLines(cap_file, warn = FALSE)))

  # Output the figure block to the doc
  writeLines(str_c("![](", fig_file, ")\n"))

  # Print footnotes
  writeLines(
    str_c("<div class=\"footnote\">",
          paste0(readLines(note_file, warn = FALSE), collapse = "</br>"),
          "</div>")
  )
}

for (m in metros) {

  # Print a top-level heading for the metro area
  writeLines(str_c("## ", toTitleCase(m), " area"))

  for (rq in rqs) {
    rq_num <- str_replace(string = rq, pattern = "rq", replacement = "")

    if (rq == "rq1") { 
      # ---------------------------------------
      #  rq1 => run for every metro area 
      # ---------------------------------------
      for (g in graph_types) {
        # Print a heading for the RQ
        writeLines(str_c("### Research question ", rq_num))

        insert_figure(metro = m, rq = rq, graph_type = g)
      }

    } else if (rq == "rq2" & (m %in% c("chicago","philadelphia","los angeles","orange county","san diego","bay area","northern new jersey"))) {
      # -------------------------------------------------
      # rq2 => run only for the specified metro areas
      # -------------------------------------------------

      # 1) For LA & Bay Area, we have SEPARATE row/col plots per order (2 files each).
      # 2) For all other metros in this subset, we still do the old "combined" row/col plots.

      # If LA or Bay Area => do the "separate" approach
      if (m %in% c("los angeles","bay area")) {

        # We'll insert each of the 8 new plots: race_row_plot_..., race_col_plot_...,
        # and firstgen_row_plot_..., firstgen_col_plot_...
        # Each suffix also ends with the order ID.
        for (suffix in c(
          "race_row_plot_448375", "race_row_plot_448374_448420",
          "race_col_plot_448375", "race_col_plot_448374_448420",
          "firstgen_row_plot_448375", "firstgen_row_plot_448374_448420",
          "firstgen_col_plot_448375", "firstgen_col_plot_448374_448420"
        )) {
          writeLines(str_c("### Research question ", rq_num))
          insert_figure(metro = m, rq = rq, graph_type = suffix)
        }

      } else {
        # For the other metros in the same set
        # (chicago, philadelphia, orange county, san diego):
        # we still show the combined row/col figure:

        for (suffix in c("race_row_plot", "race_col_plot",
                         "firstgen_row_plot", "firstgen_col_plot")) {
          writeLines(str_c("### Research question ", rq_num))
          insert_figure(metro = m, rq = rq, graph_type = suffix)
        }
      }

      # -------------------------------------------------
      # Then your custom calls by city:
      # -------------------------------------------------
      if (m == "chicago") {
        # two separate calls for 'chicago' order IDs
        writeLines(str_c("### Research question ", rq_num))
        insert_figure(metro = m, rq = rq, graph_type = "order_487984_race_by_firstgen")

        writeLines(str_c("### Research question ", rq_num))
        insert_figure(metro = m, rq = rq, graph_type = "order_488035_488053_race_by_firstgen")

      } else if (m %in% c("philadelphia",'northern new jersey')) {
        # likewise for 'philadelphia'
        writeLines(str_c("### Research question ", rq_num))
        insert_figure(metro = m, rq = rq, graph_type = "order_448922_race_by_firstgen")

        writeLines(str_c("### Research question ", rq_num))
        insert_figure(metro = m, rq = rq, graph_type = "order_448427_448440_race_by_firstgen")

      } else if (m %in% c("los angeles","orange county","san diego","bay area")) {
        # likewise for these 4 metros
        writeLines(str_c("### Research question ", rq_num))
        insert_figure(metro = m, rq = rq, graph_type = "order_448375_race_by_firstgen")

        writeLines(str_c("### Research question ", rq_num))
        insert_figure(metro = m, rq = rq, graph_type = "order_448374_448420_race_by_firstgen")
      }

    } # end of rq==rq2 section

  } # end of rqs

} # end of metros

```

### References

::: {#refs}
:::

### 2025-02-05 scripts/leaflet_maps.R

@@ -7,7 +7,7 @@
 ################################################################################
 
 ### SETTINGS
-rm(list = ls()) # remove all objects
+#rm(list = ls()) # remove all objects
 options(max.print=1000)
 #options(width = 160)
 # Set the scipen option to a high value to avoid scientific notation
@@ -233,6 +233,14 @@ display_names <- list(
   "pct_nhisp_black" = "% Black, non-Hispanic"
 )
 
+# new jersey metro area
+nj_north_metro_eps_codes
+create_eps_maps(nj_north_metro_eps_codes, c("pct_nhisp_white","pct_nhisp_black","pct_hisp_all", "med_inc_house"), mapping_level = "eps", display_names = display_names, eps_border_weight = 4)
+
+create_eps_maps(nj_eps_codes, c("pct_nhisp_white","pct_nhisp_black","pct_hisp_all", "med_inc_house"), mapping_level = "eps", display_names = display_names, eps_border_weight = 4)
+
+
+
 # figuring out socal
 create_eps_maps(socal_eps_codes, c("med_inc_house","pct_nhisp_white","pct_nhisp_black","pct_hisp_all","pct_nhisp_asian"), mapping_level = "eps", display_names = display_names, eps_border_weight = 4)
 
@@ -277,11 +285,6 @@ create_eps_maps(long_island_eps_codes, c("pct_nhisp_white","pct_nhisp_black","pc
 create_eps_maps(c('NY13','NY15'), c("pct_nhisp_white","pct_nhisp_black","pct_hisp_all", "mean_inc_house"), mapping_level = "eps", display_names = display_names, eps_border_weight = 4)
 
 
-# new jersey metro area
-nj_metro_eps_codes
-create_eps_maps(nj_metro_eps_codes, c("pct_nhisp_white","pct_nhisp_black","pct_hisp_all", "med_inc_house"), mapping_level = "eps", display_names = display_names, eps_border_weight = 4)
-
-create_eps_maps(nj_eps_codes, c("pct_nhisp_white","pct_nhisp_black","pct_hisp_all", "med_inc_house"), mapping_level = "eps", display_names = display_names, eps_border_weight = 4)
 
 # DALLAS
 create_eps_maps(c('TX19','TX20','TX21','TX22','TX23','TX24'), c("pct_nhisp_white","pct_nhisp_black","pct_hisp_all", "med_inc_house"), mapping_level = "eps", display_names = display_names, eps_border_weight = 4)

### 2025-02-05 scripts/metro_eps_codes.R

@@ -42,7 +42,8 @@ dallas_eps_codes <- c('TX19','TX20','TX21','TX22','TX23','TX24')
 nj_eps_codes <- c(paste0("NJ ", 1:9), paste0("NJ", 10:12))
 nj_eps_codes
 
-nj_north_metro_eps_codes <- c(paste0('NJ ', 4:9),paste0("NJ", 10:11))
+#nj_north_metro_eps_codes <- c(paste0('NJ ', 4:9),paste0("NJ", 10:11))
+nj_north_metro_eps_codes <- c(paste0('NJ ',c(4,6,7,8,9)),paste0("NJ", 10:11)) # taking monmouth out!
 nj_north_metro_eps_codes
 
 ######    # GO THROUGH NEW JERSEY

### 2025-02-05 scripts/rq1_tables_graphs.R

@@ -392,7 +392,7 @@ create_eps_graph <- function(eps_codes,
   
   # Save plot to file
   ggsave(
-    filename = file.path(graphs_dir, str_c(plot_name, '.png')),
+    filename = file.path(graphs_dir,'rq1', str_c(plot_name, '.png')),
     plot = plot,
     width = 14,
     height = 8,
@@ -412,7 +412,7 @@ create_eps_graph <- function(eps_codes,
   writeLines(figure_title)
   
   # Save the title to a .txt file
-  writeLines(figure_title, file.path(graphs_dir, str_c(plot_name, 'title.txt', sep = '_')))    
+  writeLines(figure_title, file.path(graphs_dir,'rq1', str_c(plot_name, 'title.txt', sep = '_')))    
   
   # 2) Create the text you want to store [REVISE NOTE TEXT LATER!]
   note_text <- c(
@@ -424,7 +424,7 @@ create_eps_graph <- function(eps_codes,
   )
   
   # 3) Write that text to a file
-  writeLines(note_text, file.path(graphs_dir, str_c(plot_name, 'note.txt', sep = '_')))    
+  writeLines(note_text, file.path(graphs_dir,'rq1', str_c(plot_name, 'note.txt', sep = '_')))    
              
   # Return the plot
   return(plot)
@@ -476,41 +476,3 @@ for (metro_name in names(all_codes)) {
   }
 }
 
-philly_eps_codes
-dallas_eps_codes
-
-
-
-  for (g in 1:length(graph_types)) {
-    
-    # call function
-    create_eps_graph(
-      eps_codes = philly_eps_codes, 
-      graph_type = graph_types[g], 
-      note = "additional text for my note!"
-    )  
-  }
-
-for (g in 1:length(graph_types)) {
-  
-  # call function
-  create_eps_graph(
-    eps_codes = dallas_eps_codes, 
-    graph_type = graph_types[g], 
-    note = "additional text for my note!"
-  )  
-}
-
-
-
-
-vec <- c(a = 5, b = -10, c = 30)
-vec
-length(vec)  # View length of atomic vector object
-1:length(vec)  # Create sequence from `1` to `length(vec)`
-
-
-for (i in 1:length(vec)) {
-  writeLines(str_c("\nvalue of object i=", i, "; type=", typeof(i)))
-  #str(vec[[i]])  # Access element contents using [[]]
-}

### 2025-02-05 scripts/rq2_tables_graphs_create_functions.R

@@ -1036,11 +1036,11 @@ create_race_by_firstgen_graph <- function(data_graph,
   writeLines(figure_title)
   
   # Save the title to a .txt file
-  writeLines(figure_title, file.path(graphs_dir, str_c(plot_name, '_title.txt')))  
+  writeLines(figure_title, file.path(graphs_dir,'rq2', str_c(plot_name, '_title.txt')))  
   
   # Save plot to file
   ggsave(
-    filename = file.path(graphs_dir, str_c(plot_name, '.png', sep = '')),
+    filename = file.path(graphs_dir,'rq2', str_c(plot_name, '.png', sep = '')),
     plot = plot,
     width = 14,
     height = 8,
@@ -1056,7 +1056,7 @@ create_race_by_firstgen_graph <- function(data_graph,
   )
   
   # 3) Write that text to a file
-  writeLines(note_text, file.path(graphs_dir, str_c(plot_name, '_note.txt')))  
+  writeLines(note_text, file.path(graphs_dir,'rq2', str_c(plot_name, '_note.txt')))  
   
   # Return the plot
   return(plot)

### 2025-02-05 scripts/rq2_tables_graphs_run_functions.R

@@ -82,10 +82,19 @@ all_orders <- list(
   #bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1110 - 1210', '366935'), # ordered jan 2018; don't like these cutpoints as much
   #bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1220 - 1520', '366934', '366932'),
   bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1070 - 1180', '448375'), # ordered jan 2019
-  bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1220 - 1520', '448374', '448420')
+  bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1220 - 1520', '448374', '448420'),
   #bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1070 - 1180', '546954'), # ordered jan 2020; much smaller sample size than orders from jan 2019
-  #bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1220 - 1520', '546946', '546945')  
+  #bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1220 - 1520', '546946', '546945'),
+  northern_new_jersey = c('northern_new_jersey', 'nj_north_metro_eps_codes', 'PSAT score 1070 - 1180', '448922'),
+  northern_new_jersey = c('northern_new_jersey', 'nj_north_metro_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440')
+  
 )
+
+
+#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448922'), eps_codes = nj_metro_eps_codes)
+#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448427'), eps_codes = nj_metro_eps_codes) # black students concentrated in nj08
+#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448440'), eps_codes = nj_metro_eps_codes) # black students stil concentrated in nj08
+
 
 all_orders %>% str()
 
@@ -147,28 +156,29 @@ get_order_ids <- function(metro) {
   if (metro == "chicago") {
     result$first_order  <- 487984
     result$second_order <- "488035_488053"
-  
-  # Philadelphia
-  } else if (metro == "philadelphia") {
+    
+    # Philadelphia or Northern New Jersey
[…]

### 2025-02-06 index.qmd
adding metros to rq2

@@
 ```{r setup, echo=FALSE, results = 'asis'}
-```{r, echo=FALSE, results='asis'}
+
 library(tidyverse)
@@
   "orange county", "san diego", "bay area",
-  "northern new jersey"
+  "northern new jersey","long island",'detroit','dallas'
 )
@@
   "northern_new_jersey" = c("448922", "448427_448440"),
+  "long_island"         = c("448922", "448427_448440"),
+  "detroit"             = c("448922", "448427_448440"),
+  "dallas"              = c("448922", "448427_448440"),
   "los_angeles"         = c("448375", "448374_448420"),
@@
     # ---------------------------------------------
-    # RQ2 Logic (row vs. col approach)
+    # RQ2 Logic (row vs. col approach + race_by_firstgen)
     # ---------------------------------------------
@@
 
-      # If a metro isn't in the list or no orders, skip
+      # If a metro isn't in the list or has no orders, skip
       if (is.null(orders_for_m)) next
@@
 
-          } # end if row vs col
+          } # end if (row vs col)
 
@@
       } # end g loop
+
+      # -------------------------------------------
+      # ADDITIONAL: race_by_firstgen for each order
+      # -------------------------------------------
+      # Like the column approach, one slide per order_id
+      for (order_id in orders_for_m) {
+        writeLines(str_c("### Research question ", rq_num))
+        # Filenames look like:
+        #   rq2_northern_new_jersey_order_448922_race_by_firstgen.png
+        graph_type_rbf <- str_c("order_", order_id, "_race_by_firstgen")
+        insert_figure(rq = rq, metro = m, graph_type = graph_type_rbf)
+      }
 

### 2025-02-06 scripts/rq1_maps.R

@@ -112,5 +112,7 @@ names(all_codes)
 create_rq1_map(c('bay_area'))
 create_rq1_map(c('bay_area'), shared_legend = T)
 
+create_rq1_map(c('long_island'), shared_legend = T)
+
 create_rq1_map(c('bay_area', 'philly', 'chicago'))
 saveWidget(create_rq1_map(c('bay_area', 'philly', 'chicago')), file.path('.', 'results', 'maps', 'rq1_map_bayarea_philly_chicago.html'), background = 'transparent', selfcontained = T)

### 2025-02-06 scripts/rq2_tables_graphs_create_functions.R

@@ -17,6 +17,7 @@ options(scipen = 999)
 library(tidyverse)
 library(formattable)
 library(patchwork)
+library(ggbreak) # custom breaks
 library(scales)
 library(readxl)
 library(lubridate)
@@ -676,7 +677,7 @@ create_sim_eps_graph <- function(data_graph,
           )
         )
       ) %>%
-      dplyr::filter(!group %in% c("AIAN, non-Hispanic", "NHPI, non-Hispanic"))
+      dplyr::filter(!group %in% c("AIAN, non-Hispanic", "NHPI, non-Hispanic",'Two+, non-Hispanic'))
     
     # Titles and legend labels
     row_plot_title   <- "Race Distribution Within Each EPS Code (Row %)"

### 2025-02-06 scripts/rq2_tables_graphs_run_functions.R

@@ -86,14 +86,18 @@ all_orders <- list(
   #bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1070 - 1180', '546954'), # ordered jan 2020; much smaller sample size than orders from jan 2019
   #bay_area = c('bay_area', 'bay_area_eps_codes', 'PSAT score 1220 - 1520', '546946', '546945'),
   northern_new_jersey = c('northern_new_jersey', 'nj_north_metro_eps_codes', 'PSAT score 1070 - 1180', '448922'),
-  northern_new_jersey = c('northern_new_jersey', 'nj_north_metro_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440')
-  
+  northern_new_jersey = c('northern_new_jersey', 'nj_north_metro_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440'),
+  long_island = c('long_island', 'long_island_eps_codes', 'PSAT score 1070 - 1180', '448922'),
+  long_island = c('long_island', 'long_island_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440'),
+  detroit = c('detroit', 'detroit_eps_codes', 'PSAT score 1070 - 1180', '448922'),
+  detroit = c('detroit', 'detroit_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440'),
+  dallas = c('dallas', 'dallas_eps_codes', 'PSAT score 1070 - 1180', '448922'),
+  dallas = c('dallas', 'dallas_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440')  
 )
 
-
-#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448922'), eps_codes = nj_metro_eps_codes)
-#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448427'), eps_codes = nj_metro_eps_codes) # black students concentrated in nj08
-#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448440'), eps_codes = nj_metro_eps_codes) # black students stil concentrated in nj08
+#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448922'), eps_codes = c('TX19','TX20','TX21','TX22','TX23','TX24')) # 
+#create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448427'), eps_codes = c('TX19','TX20','TX21','TX22','TX23','TX24')) 
[…]

### 2025-02-07 index.qmd
adding rq2 for houston

@@
   "orange county", "san diego", "bay area",
-  "northern new jersey","long island",'detroit','dallas'
+  "northern new jersey","long island",'detroit','dallas','houston'
 )
@@
   "san_diego"           = c("448375", "448374_448420"),
-  "bay_area"            = c("448375", "448374_448420")
+  "bay_area"            = c("448375", "448374_448420"),
+  'houston'             = c('329702') #   # houston = c('houston', 'houston_eps_codes', 'PSAT score 1010 - 1520', '329702')
+
 )

### 2025-02-07 scripts/metro_eps_codes.R

@@ -28,8 +28,8 @@ long_island_eps_codes <- c(paste0('NY',16:21))
 atl_eps_codes <- paste0("GA ", 1:4) #
 atl_eps_codes
 
-htown_eps_codes <- str_c("TX",15:18)
-htown_eps_codes
+houston_eps_codes <- str_c("TX",15:18)
+houston_eps_codes
 
 # stuff to add
   # some metro for los angeles

### 2025-02-07 scripts/rq2_tables_graphs_run_functions.R

@@ -92,8 +92,11 @@ all_orders <- list(
   detroit = c('detroit', 'detroit_eps_codes', 'PSAT score 1070 - 1180', '448922'),
   detroit = c('detroit', 'detroit_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440'),
   dallas = c('dallas', 'dallas_eps_codes', 'PSAT score 1070 - 1180', '448922'),
-  dallas = c('dallas', 'dallas_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440')  
+  dallas = c('dallas', 'dallas_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440'),
+  houston = c('houston', 'houston_eps_codes', 'PSAT score 1010 - 1520', '329702')
 )
+
+# 2 Stephen F Austin State University, univ_id=228431 ord_num=329702          PSAT 1010-1520; grades C+ to A+, 2019 HS grads, all of TX, ordered 8/31/2017 [but seems to have TX 17....]
 
 #create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448922'), eps_codes = c('TX19','TX20','TX21','TX22','TX23','TX24')) # 
 #create_sim_eps_race_table(data = fa20_oos_psat_sf, ord_nums = c('448427'), eps_codes = c('TX19','TX20','TX21','TX22','TX23','TX24')) 
@@ -151,6 +154,7 @@ for (orders in all_orders) {
 
 
 ################
+# 1) Utility function to fetch the order IDs
 get_order_ids <- function(metro) {
   
   # Initialize a result list
@@ -161,7 +165,11 @@ get_order_ids <- function(metro) {
     result$first_order  <- 487984
     result$second_order <- "488035_488053"
     
-    # Philadelphia or Northern New Jersey
+    # Houston
+  } else if (metro %in% c("houston")) {
+    result$first_order  <- 329702
+    
+    # Philadelphia, Northern NJ, etc.
   } else if (metro %in% c("philadelphia", "northern new jersey",'long island','detroit','dallas')) {
     result$first_order  <- 448922
     result$second_order <- "448427_448440"
@@ -171,7 +179,6 @@ get_order_ids <- function(metro) {
     result$first_order  <- 448375
     result$second_order <- "448374_448420"
     
-    # If we ever pass in a metro not covered above, throw error
   } else {
     stop("No matching order logic for metro: ", metro)
   }
[…]

### 2025-02-07 scripts/map_functions.R
nut n' honey

@@ -12,7 +12,7 @@ all_codes <- list(
   chicago = list(name = 'Chicago', eps = chi_eps_codes),
   cleveland = list(name = 'Cleveland', eps = cleveland_eps_codes),
   northern_nj = list(name = 'Northern New Jersey', eps = nj_north_metro_eps_codes),
-  houston = list(name = 'Houston', eps = htown_eps_codes),
+  houston = list(name = 'Houston', eps = houston_eps_codes),
   bay_area = list(name = 'Bay Area', eps = bay_area_eps_codes),
   long_island = list(name = 'Long Island', eps = long_island_eps_codes),
   ny_ny = list(name = 'New York, New York', eps = nyny_metro_eps_codes),

### 2025-02-07 scripts/rq2_maps.R

@@ -2,7 +2,7 @@ library(leaflet)
 library(tidyverse)
 library(formattable)
 library(htmlwidgets)
-
+library(haven)
 
 # Load data
 

### 2025-02-07 index.qmd
started presentation

@@
 ---
-title: 'Title'
-subtitle: 'Subtitle'
-author: ['Author', 'Author']
+title: 'Structuring College Access'
+subtitle: 'The Market Segment Model and College Board Geomarkets'
+author:
+  - name: Ozan Jaquette
+    affiliation: UCLA
+  - name: Karina Salazar
+    affiliation: University of Arizona
+  - name: Crystal Han
+    affiliation: Stanford University
 bibliography: ./assets/bib/cb_geomarket_bib.bib
@@
 
-# Level-1 heading
+# Introduction
+
+### Introduction
+#### Talent is everywhere, but structure allocates talent to unequal opportunities
+
+College Board "Geomarkets"
+
+- Carve states and metropolitan areas into smaller geographic regions meant to define local recruiting markets
+
+POP OUT OF GEOMARKETS
+
+- Chicago
+- New York
+
+Geomarkets are an input to two College Board products
+
+- Enrollment Planning Service:
+- Student Search Service:
+
+**Research questions** focus on metropolitan areas
+
+1. What is SES and racial variation between, within geomarkets? How does this change over time?
+1. What is SES and racial composition of included vs. excluded prospects, if student list purchases filter on particular Geomarkets?
+
+### New York and Chicago Geomarkets
+
+::: columns
+::: {.column width="40%"}
+
+![](./assets/images/nyc_geomarkets.png)
+:::
+
+::: {.column width="60%"}
+
+![](./assets/images/chicago_geomarkets.png)
+:::
+:::
 
@@
 
+# Empirical Context
+
+# Theory
 ### Test map
@@
 
+# Discussion
+
 ### References

### 2025-02-08 index.qmd
update TOC

@@
     slide-level: 3
+    menu:
+      titleSelector: 'h1, h2, .slide:not(.hide) h3'
+      useTextContentForMissingTitles: false
+      hideMissingTitles: true
     template-partials:
@@
 
-### Level-3 heading
+### Level-3 heading {.hide}
 #### Level-4 subheading
@@
 
+# For RQ2, we have two plot types: "race" and "firstgen"
+graph_types_rq2 <- c("race", "firstgen")
+
 # 3) Subset of metros with RQ2 data (2 orders each)
@@
       # Each metro has 2 plots: "race" & "ses"
-      for (g in graph_types_rq1) {
-        writeLines(str_c("### Research question ", rq_num))
+      for (i in seq(graph_types_rq1)) {
+        g <- graph_types_rq1[i]
+
+        writeLines(str_c("### Research question ", rq_num, if_else(i == 1, "", " {.hide}")))
         insert_figure(rq = rq, metro = m, graph_type = g)
@@
       # We'll do 2 group types: race & firstgen
-      for (g in c("race", "firstgen")) {
+      for (i in seq(graph_types_rq2)) {
+        g <- graph_types_rq2[i]
 
@@
             #
-            writeLines(str_c("### Research question ", rq_num))
+            writeLines(str_c("### Research question ", rq_num, if_else(i == 1, "", " {.hide}")))
             graph_type_row <- str_c(g, "_", rc, "_plot")
@@
               # Each order ID gets its own "###" => new Reveal.js slide
-              writeLines(str_c("### Research question ", rq_num))
+              writeLines(str_c("### Research question ", rq_num, " {.hide}"))
 
@@
       for (order_id in orders_for_m) {
-        writeLines(str_c("### Research question ", rq_num))
+        writeLines(str_c("### Research question ", rq_num, rq_num, " {.hide}"))
         # Filenames look like:

### 2025-02-08 assets/js/script.js

@@ -29,4 +29,14 @@ $(function() {
     $('img').removeClass('disabled');
   });
   
+  const h3 = $('.slide:not(.hide) h3').map(function() {
+    return $(this).text();
+  }).get();
+
+  $('.slide-menu-item-vertical').filter(function() {
+    if (h3.includes($(this).text())) {
+      $(this).css('padding-left', '42px');
+    }
+  });
+
 });

### 2025-02-08 index.qmd
update affiliation

@@
 subtitle: 'The Market Segment Model and College Board Geomarkets'
-author:
-  - name: Ozan Jaquette
-    affiliation: UCLA
-  - name: Karina Salazar
-    affiliation: University of Arizona
-  - name: Crystal Han
-    affiliation: Stanford University
+author: ['Ozan Jaquette', 'Karina Salazar', 'Crystal Han']
+institute: ['UCLA', 'University of Arizona', 'Stanford University']
 bibliography: ./assets/bib/cb_geomarket_bib.bib

### 2025-02-10 index.qmd
test hiding slides from toc
starting presentation in earnest

@@
 
-### Introduction
+### Introduction {.hide}
 #### Talent is everywhere, but structure allocates talent to unequal opportunities
@@
 
-POP OUT OF GEOMARKETS
-
-- Chicago
-- New York
-
 Geomarkets are an input to two College Board products
 
-- Enrollment Planning Service:
-- Student Search Service:
+- Enrollment Planning Service (EPS): software recommends which Geomarkets to recruit and which high schools to visit
+- Student Search Service: purchase the contact information of prospects that meet "search filter" criteria
 
-**Research questions** focus on metropolitan areas
+**Research questions** (focus on metropolitan areas):
 
-1. What is SES and racial variation between, within geomarkets? How does this change over time?
+1. What is SES and racial variation between, within Geomarkets? How does this change over time?
 1. What is SES and racial composition of included vs. excluded prospects, if student list purchases filter on particular Geomarkets?
 
-### New York and Chicago Geomarkets
+### New York and Chicago Geomarkets {.hide}
 
@@
 
-## Level-2 heading
+# Background
 
-### Level-3 heading
+### Background{.hide}
+#### Creating Market Segment Model and Geomarkets
 
-### Level-3 heading {.hide}
-#### Level-4 subheading
+Project origins
 
-# Empirical Context
+Market Segment Model
+
+- Allocates students into four market segments based on score-sending behavior of 1980 SAT test-takers
+- "local" student submits most SAT scores to institutions in their own Geomarket
+- "An “in-state” student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions
+
+Geomarkets
+
+> Zemsky & Oedel (1983) [p.
+4] write that,  [Often] the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
+Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
 
@@
 
-### References
+### Discussion{.hide}
+#### Summary of results
+
+THIS IS TEXT
+
+### Discussion{.hide}
+#### Contributin to scholarship
+
+THIS IS TEXT
+
+# References
 

### 2025-02-12 index.qmd
fleshing out presentation

--- 2025-02-08 index.qmd
@@
 
-POP OUT OF GEOMARKETS
-
-- Chicago
-- New York
-
 Geomarkets are an input to two College Board products
 
-- Enrollment Planning Service:
-- Student Search Service:
+- Enrollment Planning Service (est. 1984): software recommends which Geomarkets to recruit and which high schools to visit
+- Student Search Service (est. 1972): purchase the contact information of prospects that meet "search filter" criteria
 
-**Research questions** focus on metropolitan areas
+**Research questions** (focus on metropolitan areas):
 
-1. What is SES and racial variation between, within geomarkets? How does this change over time?
+1. What is SES and racial variation between, within Geomarkets? How does this change over time?
 1. What is SES and racial composition of included vs. excluded prospects, if student list purchases filter on particular Geomarkets?
 
-### New York and Chicago Geomarkets
+### New York and Chicago Geomarkets {.hide}
 
@@
 
-## Level-2 heading
+# Background and Theory
 
-### Level-3 heading
+### Background and Theory {.hide}
+#### Geomarkets and Market Segment Model created by @RN4982
 
-### Level-3 heading {.hide}
-#### Level-4 subheading
+**Geomarkets** [@RN4982, p. 4]:
 
-# Empirical Context
+> Market boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
+Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
 
-# Theory
+**Market Segment Model**, based on student SAT score-sending behavior
+
+- Send scores to colleges: local (in Geomarket); in-state (in state, diff Geomarket); regional (in region, diff state); or national (diff region)
+- Students sorted into 4 market segments: local, in-state, regional, national
+  - "local" students submit most SAT scores to local institutions
+- *(Geo)Market Segment Profile*: number of students by market segment
+- *Institutional Profile*: number of students from each market segment the send scores to institution *i*, separately by Geomarket
+
+### Background and Theory {.hide}
+#### Draw from scholarship on quantification, particularly @RN4975
+
+**Correlation**, extent to which variables move together
+
+<div style="margin-left: 1em;">
+<details>
+  <summary>@RN4982: SAT score, parental education, family income, educational aspirations predict student market segment</summary>
+> Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p.
+42].
+</details>
+</div>
+
+  - Observed correlations are a function of structural inequality
+    - Predicting future based on past correlations reproduces inequality
+
+\medskip
+
+**Homophily**, birds of a feather flock together
+
+<div style="margin-left: 1em;">
+<details>
+  <summary>Market Segment Model based on idea of (voluntary) homophily</summary>
+> The layering of collegiate competition is primarily a socioeconomic layering.
+The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
+Competition among colleges, as admissions officers have told us for so long, is a matter of keeping company with one’s peers [@RN4982, p.
+72].
+</details>
+</div>
+
+- Comercial social networks engineer homophily [@RN4975]
+- Commodifying the Market Segment Model engineers homophily
+    - Software tells colleges which Geomarkets, high schools to target
+
+### Background and Theory {.hide}
+#### Research questions and expectations
+
+**Research questions**
+
+1. What is the SES and racial variation between, within Geomarkets in a metro area? How does this change over time?
+
+    - Expect socioeconomic inequality between Geomarkets because Geomarkets/Market Segment Model based on social stratification
+    - Expect racial inequality between Geomarkets because racial segregation, correlation between class and race
+
+1. What is the SES and racial composition of included vs. excluded prospects, if "student list" purchases filter on particular Geomarkets?
+
+    - Excluding Geomarkets that are poor and non-White associated with:
+      - disproportionate exclusion of first-generation students
+      - disproportionate exclusion of non-white students
+      - disproportionate exclusion of first-generation, non-white students
+
+# Methods
+
+### Methods
+#### Data
+
+RQ1: SES and racial variation over time between, within Geomarkets
+
+- Census tract-level data
+  - 1980, 2000 Census; 2016-2020 American Community Survey (ACS)
+  - Variables: race/Hispanic origin; median income, % poverty, % BA+
+- Geomarket shapefiles obtained from 2012 blog post
+  - Each zip code assigned to one Geomarket
+  - Partial spatial assigns each Census tract to one or more Geomarkets
+
+RQ2: SES and racial composition of student list purchases, by Geomarket
+
+- 414 student list purchases (associated with 2.5 million prospects) made by 14 public universities from 2016-2020
+  - Each prospect has a home zip code and a HS zip code (point geoms)
+- Analyze purchases that include all Geomarkets in a metro area
+  - Show who would be excluded had search criteria exluded a Geomarket
+
+### Methods
+#### Research design and analyses
+
+Research design
+
+- Metropolitan areas associated with multiple Geomarkets are cases
+
+Case selection
+
+- Initial case selection based on availability of student list purchases (RQ2)
+  - Want purchases that filter on test score, high school class, GPA, but not on student preference filters (e.g., college size)
+  - For some metro areas (e.g., DC/Maryland/Virgina) we don't have purchases that include all Geomarkets
+- Final case selection (3 metro areas)
+  - TBD, but conduct analyses for larger sample of metros
+
+Analyses (for both RQ1 and RQ2)
+
+- Descriptive statistics for between Geomarket variation
+- Interactive maps for within Geomarket variation
+
 ### Test map

### 2025-02-12 index.qmd
tweaks to quote size in slides

@@
   <summary>@RN4982: SAT score, parental education, family income, educational aspirations predict student market segment</summary>
+<div style="font-size: 0.85em;">
 > Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p.
 42].
+</div>
 </details>
@@
 - Observed correlations are a function of structural inequality
-    - Predicting future based on past correlations reproduces inequality
-
+  - "Predicting the future on the basis of the past threatens to reify and reproduce exist
+ing inequalities" [@RN4794, p. 224]
+<!--Predicting future based on past correlations reproduces inequality  -->
 \medskip
@@
   <summary>Market Segment Model based on idea of (voluntary) homophily</summary>
+<div style="font-size: 0.85em;">  
 > The layering of collegiate competition is primarily a socioeconomic layering.
-The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
+The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of communities from which colleges draw their students.
 Competition among colleges, as admissions officers have told us for so long, is a matter of keeping company with one’s peers [@RN4982, p.
 72].
+</div>
 </details>

### 2025-02-13 index.qmd
working on prez

--- 2025-02-10 index.qmd
@@
 ### Introduction {.hide}
-#### Talent is everywhere, but structure allocates talent to unequal opportunities
 
@@
 
-- Enrollment Planning Service (EPS): software recommends which Geomarkets to recruit and which high schools to visit
-- Student Search Service: purchase the contact information of prospects that meet "search filter" criteria
+- Enrollment Planning Service (est. 1984): software recommends which Geomarkets to recruit and which high schools to visit
+- Student Search Service (est. 1972): purchase the contact information of prospects that meet "search filter" criteria
 
-**Research questions** (focus on metropolitan areas):
+Research questions
 
-1. What is SES and racial variation between, within Geomarkets? How does this change over time?
+1. What is SES, racial variation between, within Geomarkets? Over time?
 1. What is SES and racial composition of included vs. excluded prospects, if student list purchases filter on particular Geomarkets?
+
+Research interest: how quantification reproduces social stratification
 
@@
 ### Background{.hide}
-#### Creating Market Segment Model and Geomarkets
+#### Geomarkets and Market Segment Model created by @RN4982
 
-Project origins
+**Market Segment Model**, based on student SAT score-sending behavior
 
-Market Segment Model
+- Each student-to-college score assigned to one of four categories:
+  - Local (in Geomarket); in-state (diff Geomarket); regional; national
+- Students sorted into 4 market segments: local, in-state, regional, national
+  - "Local" students submit most SAT scores to local institutions
+- *Market Segment Profile* (produced by Geomarket): 
+  - How many students in each market segment
+- *Institutional Profile*  (produced by Geomarket): 
+  - How many students by market segement send scores to college *i*
 
-- Allocates students into four market segments based on score-sending behavior of 1980 SAT test-takers
-- "local" student submits most SAT scores to institutions in their own Geomarket
-- "An “in-state” student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions
+<style>
+/* Remove the triangle marker from summary elements */
+details summary {
+  list-style: none; /* In some browsers, this alone may suffice */
+}
 
-Geomarkets
+/* For WebKit-based browsers (Chrome, Safari, etc.) */
+details summary::-webkit-details-marker {
+  display: none;
+}
+</style>
 
-> Zemsky & Oedel (1983) [p.
-4] write that,  [Often] the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
-Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
+<div style="margin: 0.5em 0 0 0;">
+<details style="margin: 0; padding: 0;">
+  <summary>**Geomarkets**</summary>
+<div style="font-size: 0.80em; margin-bottom: 0;">
+> Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
+Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs [@RN4982, p.
+4].
+</div>
+</details>
+</div>
+
+- Meant to reproduce the recruiting territories of admissions officers
+- @RN4982 create 143 Geomarkets in three regions
+- Enrollment Planning Service (EPS) uses 304 Geomarkets
 
 # Theory
-### Test map
+
+### Scholarship on quantification {.hide}
+#### Discussion of correlation by @RN4975[chapter 1]
+
+**Correlation**, extent to which variables move together
+
+- Predictive analytics in two steps: 
+    1. Measure correlations in previous cases (training data)
+    1. Apply training data correlations to predict future cases
+- But correlations from training data are point-in-time snapshot
+  - Observed correlations caused by enduring structural inequality
+  - "Predicting the future on the basis of the past threatens to reify and reproduce existing inequalities" [@RN4794, p. 224]
+
+<div style="margin: 0.5em 0 0 0;">
+<details style="margin: 0; padding: 0;">
+  <summary>**@RN4982[chapter 3]**</summary>
+<div style="font-size: 0.85em; margin-bottom: 0;">
+> Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p.
+42].
+</div>
+</details>
+</div>
+
+- Which variables correlated with student market segment (training data)?
+  - Parental education, family income, SAT score, educational aspirations
+  - Conclusion: demand for higher education is a function of social origin
+- Recommendation: selective colleges target high-SES Geomarkets
+
+### Scholarship on quantification {.hide}
+#### Discussion of homophily by @RN4975[chapter 2]
+
+**Homophily**, birds of a feather flock together
+
+- Network science assumes homophily result of voluntary action by actors
+- But homophily programmed into algorithms that create connections
+- Homophily in market research/geodemography
+  - Based on snapshot of existing social stratification, match vertically categorized consumers to vertically categorized producers
+  - Amplifies the effect of initial stratificaiton on subsequent stratification
+
+<div style="margin: 0.5em 0 0 0;">
+<details style="margin: 0; padding: 0;">
+  <summary>**@RN4982[chapter 4]**</summary>
+<div style="font-size: 0.85em; margin-bottom: 0;">
+> The layering of collegiate competition is primarily a socioeconomic layering.
+The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of communities from which colleges draw their students.
+Competition among colleges, as admissions officers have told us for so long, is a matter of keeping company with one’s peers [@RN4982, p.
+72].
+</div>
+</details>
+</div>
+
+- Conducted a social network analysis (training data):
+  - 2-mode network: students (1) tie to colleges (2) by sending scores
+  - Colleges *i* and *j* compete if many students send scores to both
+  - Conclusion: Like colleges compete for like (social origin) students
+
+<div style="margin: 0.5em 0 0 0;">
+<details style="margin: 0; padding: 0;">
+  <summary> Recommendation: target Geomarkets, schools targeted by peer colleges</summary>
+<div style="font-size: 0.85em; margin-bottom: 0;">
+> On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
+Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
+The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice [@RN4982, p.
+44].
+</div>
+</details>
+</div>
+
+### Research questions and expectations{.hide}
+
+**Research goals**
+
+- Examine extent of social inequaltiy between Geomarkets (RQ1)
+- Show how application of Geomarkets produces inequality (RQ2)
+
+**Research questions**
+
+1. SES and racial variation between, within Geomarkets? Change over time?
+
+    - Geomarkets/Market Segment Model based on social stratification
+    - US Racial segregation, correlation between class and race
+
+1. SES and racial composition of included vs. excluded prospects, if "student list" purchases filter on particular Geomarkets?
+
+    - Excluding Geomarkets that are poor and non-White associated with:
+      - Disproportionate exclusion of first-generation students
+      - Disproportionate exclusion of non-white students
+      - Disproportionate exclusion of first-generation, non-white students
+
+# Methods
+
+### Methods{.hide}
+#### Data
+
+**RQ1: SES and racial variation over time between, within Geomarkets**
+
+- Census tract-level data
+  - 1980, 2000 Census; 2016-2020 American Community Survey (ACS)
+  - Variables: race/Hispanic origin; median income, % poverty, % BA+
+- Geomarket shapefiles obtained from 2012 blog post
+  - Each zip code assigned to one Geomarket
+  - Partial spatial assigns each Census tract to one or more Geomarkets
+
+**RQ2: SES and racial composition of student list purchases**
+
+- 414 student lists -- associated with 2.5 million prospects -- purchaseds by 14 public universities from 2016-2020
+  - Each prospect has a home zip code and a HS zip code (point geoms)
+- Analyze purchases that include all Geomarkets in a metro area
+  - Show who would be excluded had search criteria exluded a Geomarket
+
+### Methods{.hide}
+#### Research design: quantitative case study
+
+Cases are metro areas associated with multiple Geomarkets (n = 17)
+
+<div style="margin-top: 0.8em;"></div>
+
+Case selection
+
+- Initial case selection based on availability of student list purchases (RQ2)
+  - Prefer purchases that filter on test score, HS class, GPA, but not on student preferences (e.g., college size)
+    - Ideal: purchases differ on test score range, but same on other filters
+  - For some metros, don't have purchases that include all Geomarkets
+- Final set of cases for manuscript main text (3 metro areas)
+  - TBD; unsure about criteria
+  - Conducted analyses for larger sample of metros
+
+Analyses (RQ1 and RQ2)
+
+- Descriptive statistics for between Geomarket variation
+- Interactive maps for within Geomarket variation
+
+### Test map{.hide}
 
@@
 
-THIS IS TEXT
+**RQ1: SES and racial variation over time between, within Geomarkets**
+
+- Substantial inequality between and within Geomarkets
+- In 1980, Black people are usually concentrated in poorest Geomarket
+  - In some metro areas, this persists (e.g., Cleveland)
+  - In some metro areas,  (e.g., Atlanta)
+
+**RQ2: SES and racial composition of student list purchases**
+
+- Poor, non-white Geomarkets produce disproportionate share of Black, Hispanic students
+- Poor, non-white Geomarkets produce disproportionate share of first-generation studnets
+- Purchased student profiles by race X parental education
+  - Asian students from low-income Geomarkets tend to be first-gen compared to Asain students from rich Geomarkets
+  - Same result for Hispanic students
 

### 2025-02-13 index.qmd
pop out quotes

@@
 
-<style>
-/* Remove the triangle marker from summary elements */
-details summary {
-  list-style: none; /* In some browsers, this alone may suffice */
-}
-
-/* For WebKit-based browsers (Chrome, Safari, etc.) */
-details summary::-webkit-details-marker {
-  display: none;
-}
-</style>
-
-<div style="margin: 0.5em 0 0 0;">
-<details style="margin: 0; padding: 0;">
-  <summary>**Geomarkets**</summary>
-<div style="font-size: 0.80em; margin-bottom: 0;">
-> Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
+<p class="quote">Geomarkets</p>
+<p style="display:none;">Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
 Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs [@RN4982, p.
-4].
-</div>
-</details>
-</div>
+4].</p>
 
@@
 
-<div style="margin: 0.5em 0 0 0;">
-<details style="margin: 0; padding: 0;">
-  <summary>**@RN4982[chapter 3]**</summary>
-<div style="font-size: 0.85em; margin-bottom: 0;">
-> Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p.
-42].
-</div>
-</details>
-</div>
+<p class="quote">@RN4982[chapter 3]</p>
+<p style="display:none;">Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p.
+42].</p>
 
@@
 
-<div style="margin: 0.5em 0 0 0;">
-<details style="margin: 0; padding: 0;">
-  <summary>**@RN4982[chapter 4]**</summary>
-<div style="font-size: 0.85em; margin-bottom: 0;">
-> The layering of collegiate competition is primarily a socioeconomic layering.
+<p class="quote">@RN4982[chapter 4]</p>
+<p style="display:none;">The layering of collegiate competition is primarily a socioeconomic layering.
 The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of communities from which colleges draw their students.
 Competition among colleges, as admissions officers have told us for so long, is a matter of keeping company with one’s peers [@RN4982, p.
-72].
-</div>
-</details>
-</div>
+72].</p>
 
@@
 
-<div style="margin: 0.5em 0 0 0;">
-<details style="margin: 0; padding: 0;">
-  <summary> Recommendation: target Geomarkets, schools targeted by peer colleges</summary>
-<div style="font-size: 0.85em; margin-bottom: 0;">
-> On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
+<p class="quote">Recommendation: target Geomarkets, schools targeted by peer colleges</p>
+<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
 Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
 The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice [@RN4982, p.
-44].
-</div>
-</details>
-</div>
+44].</p>
 

### 2025-02-13 assets/js/script.js

@@ -7,25 +7,45 @@ $(function() {
   $('.slides section.level3').prepend(link);
   
   $('.slides section img').on('click', function() {
-    let src = $(this).attr('src');
+    const src = $(this).attr('src');
     
-    var $slide = $(this).closest('.slide');
+    const $slide = $(this).closest('.slide');
     $slide.find('.modal img').attr('src', src);
     
-    $(this).closest('.slide').find('.modal').fadeIn(600);
-    $(this).closest('.slide').find('p img').addClass('disabled');
+    $slide.find('.modal').fadeIn(600);
+    $slide.find('p img').addClass('disabled');
   });
   
+  $('.slides section .quote').on('click', function() {
+    const $slide = $(this).closest('.slide');
+    $slide.find('.quote-container').remove();
+    
+    const quote = $('<div class="quote-container"></div>').append($(this).next().html());
+    
+    $slide.find('.modal').append(quote);
+    $slide.find('.modal').css('height', 'auto');
+    
+    $slide.find('.modal').fadeIn(600);
+  })
+  
   $('.close').on('click', function() {
-    $(this).closest('.slide').find('.modal').fadeOut();
-    $(this).closest('.slide').find('p img').removeClass('disabled');
+    const $slide = $(this).closest('.slide');
+    
+    $slide.find('.modal').fadeOut(400, function() {
+      $(this).css('height', '');
+      $(this).find('.quote-container').remove();
+    });
+    $slide.find('p img').removeClass('disabled');
   });
   
   $('<hr>').insertAfter('.reveal .slide h4');
   $('<hr>').insertAfter('.reveal .slide:not(.caption) h3:not(:has(+ h4))');
   
   Reveal.addEventListener('slidechanged', function() {
-    $('.modal').fadeOut();
+    $('.modal').fadeOut(400, function() {
+      $(this).css('height', '');
+      $(this).find('.quote-container').remove();
+    });
     $('img').removeClass('disabled');
   });
   

### 2025-02-13 assets/js/script.js
allow custom width for pop up images

@@ -6,11 +6,19 @@ $(function() {
   const link = '<a class="link" href="https://ozanj.github.io/cb_geo/">ozanj.github.io/cb_geo</a>';
   $('.slides section.level3').prepend(link);
   
-  $('.slides section img').on('click', function() {
-    const src = $(this).attr('src');
-    
+  $('.slides section p img').on('click', function() {
+    const src = $(this).attr('src'),
+      width = $(this).attr('data-width');
+
     const $slide = $(this).closest('.slide');
     $slide.find('.modal img').attr('src', src);
+    
+    // reset width from any previous image
+    $slide.find('.modal img').css('width', '');
+    
+    if (width) {
+      $slide.find('.modal img').css('width', width);
+    }
     
     $slide.find('.modal').fadeIn(600);
     $slide.find('p img').addClass('disabled');

### 2025-02-13 index.qmd
allow custom width for pop up images
discussion slide

@@
 
-![](./assets/images/nyc_geomarkets.png)
+<img src="./assets/images/nyc_geomarkets.png" data-width="60%">
 :::
@@
 
-<p class="quote">Recommendation: target Geomarkets, schools targeted by peer colleges</p>
+<p class="quote">Recommendation: target Geomarkets, HS targeted by peer colleges</p>
 <p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
@@
 - In 1980, Black people are usually concentrated in poorest Geomarket
-  - In some metro areas, this persists (e.g., Cleveland)
-  - In some metro areas,  (e.g., Atlanta)
+  - In some metros, this pattern persists (e.g., Cleveland)
+  - In some metros, these Geomarkets experience gentrification
 
@@
 
-- Poor, non-white Geomarkets produce disproportionate share of Black, Hispanic students
-- Poor, non-white Geomarkets produce disproportionate share of first-generation studnets
+- Poor, non-white Geomarkets produce disproportionate share of Black, Hispanic students, (and often Asian)
+- Poor, non-white Geomarkets produce disproportionate share of first-generation students
 - Purchased student profiles by race X parental education
-  - Asian students from low-income Geomarkets tend to be first-gen compared to Asain students from rich Geomarkets
+  - Asian students from low-income Geomarkets tend to be first-gen compared to Asain students from affluent Geomarkets
   - Same result for Hispanic students
@@
 ### Discussion{.hide}
-#### Contributin to scholarship
+#### Contribution to scholarship
 
-THIS IS TEXT
+Scholarship on quantification (in education)
+
+- Scholarship has not examined "b2b" quantification that classifies consumers for the sake of producers
+- @RN5005 show effect of quantification, but not mechanisms of *how* quantification produces inequality
+
+Scholarship on postsecondary "organizational field"
+
+- Scholarship focuses on schools and policies as "sorting machines" [@RN4873]
+  - But EPS, Student Search Service products structure college access
+  - College access structured by third-party vendors and products that amplify structural inequalities observed on demand-side
+- Org field is "totality of relevant actors" [@RN527, p. 148]
+  - The organizational field should be focus regulation
+  - Sociology should identify high-leverage actors/products in org field
 

### 2025-02-13 scripts/rq1_maps.R
remove test slide

@@ -107,5 +107,7 @@ create_rq1_map(c('bay_area'), shared_legend = T)
 
 create_rq1_map(c('long_island'), shared_legend = F)
 
+create_rq1_map(c('philly', 'chicago'))
+
 create_rq1_map(c('bay_area', 'philly', 'chicago'))
 saveWidget(create_rq1_map(c('bay_area', 'philly', 'chicago')), file.path('.', 'results', 'maps', 'rq1_map_bayarea_philly_chicago.html'), background = 'transparent', selfcontained = T)

### 2025-02-15 index.qmd
remove test slide
start revising presentation

@@
 
-College Board "Geomarkets"
+Historically, College Board offered two recruiting products to colleges
 
-- Carve states and metropolitan areas into smaller geographic regions meant to define local recruiting markets
+- Student Search Service (est. 1972): purchase contact information of prospects that meet "search filter" criteria
+- Enrollment Planning Service (est. 1984): software-as-service, says which "Geomarkets" to target and which high schools to visit
 
-Geomarkets are an input to two College Board products
+College Board "Geomarkets" are an input in both products
 
-- Enrollment Planning Service (est. 1984): software recommends which Geomarkets to recruit and which high schools to visit
-- Student Search Service (est. 1972): purchase the contact information of prospects that meet "search filter" criteria
+- Geomarkets carve states and metro areas into smaller geographic regions meant to define local recruiting markets [@RN4982]
 
-Research questions
+Concerns: Geomarket borders correlated with race/class; likely to be used in ways that undermine equality of opportunity
 
 1. What is SES, racial variation between, within Geomarkets? Over time?
-1. What is SES and racial composition of included vs. excluded prospects, if student list purchases filter on particular Geomarkets?
+1. What is SES and racial composition of included vs. excluded prospects when student list purchases filter on particular Geomarkets?
 
-Research interest: how quantification reproduces social stratification
+<!-- Research interest: how quantification reproduces social stratification -->
 
@@
 ::: columns
-::: {.column width="40%"}
+::: {.column width="50%"}
 
-<img src="./assets/images/nyc_geomarkets.png" data-width="60%">
+![](./assets/images/chicago_geomarkets.png)
+
+<!-- <img src="./assets/images/nyc_geomarkets.png" data-width="60%"> -->
 :::
 
-::: {.column width="60%"}
-
-![](./assets/images/chicago_geomarkets.png)
+::: {.column width="50%"}
+**[University of Chicago admissions counselors, Chicago-land](https://collegeadmissions.uchicago.edu/contact)**
+![](./assets/images/chicagoland_territories_u_chicago_redacted.png)
 :::
@@
 
-**Market Segment Model**, based on student SAT score-sending behavior
+<p class="quote">Geomarkets</p>
+<p style="display:none;">Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
+Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs [@RN4982, p.
+4].</p>
+
+- Meant to reproduce the recruiting territories of admissions officers
+- @RN4982 create 143 Geomarkets in three regions
+- Enrollment Planning Service (EPS) uses 304 Geomarkets
+
+**Market Segment Model**, based on student SAT score-sending behavior [SIMPLIFY]
 
@@
   - How many students by market segement send scores to college *i*
-
-<p class="quote">Geomarkets</p>
-<p style="display:none;">Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
-Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs [@RN4982, p.
-4].</p>
-
-- Meant to reproduce the recruiting territories of admissions officers
-- @RN4982 create 143 Geomarkets in three regions
-- Enrollment Planning Service (EPS) uses 304 Geomarkets
 
@@
 - Interactive maps for within Geomarket variation
-
-### Test map{.hide}
-
-<iframe src="https://comforting-kashata-600ce0.netlify.app/" width=100% height=100% allowtransparency="true"></iframe>
 

### 2025-02-17 index.qmd
add maps to slides

@@
 
+      # Add RQ1 map
+      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
+      writeLines("#### Map")
+
+      rq1_map <- if_else(
+        m == 'los angeles',
+        'https://rq1-map-los-angeles.netlify.app',
+        str_c('./results/maps/rq1_map_', str_replace_all(m, " ", "_"), '.html')
+      )
+
+      writeLines(str_c('<iframe data-src="', rq1_map, '" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'))
+
     # ---------------------------------------------
@@
       for (order_id in orders_for_m) {
-        writeLines(str_c("### Research question ", rq_num, rq_num, " {.hide}"))
+        writeLines(str_c("### Research question ", rq_num, " {.hide}"))
         # Filenames look like:
@@
       }
+
+      # Add RQ2 map
+      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
+      writeLines("#### Map")
+
+      writeLines(str_c('<iframe data-src="./results/maps/rq2_map_', str_replace_all(m, " ", "_"), '.html" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'))
 

### 2025-02-17 assets/js/script.js

@@ -57,6 +57,12 @@ $(function() {
     $('img').removeClass('disabled');
   });
   
+  Reveal.on('ready', (event) => {
+    $('iframe').each(function() {
+        $(this).attr('src', $(this).data('src'));
+    });
+  });
+  
   const h3 = $('.slide:not(.hide) h3').map(function() {
     return $(this).text();
   }).get();

### 2025-02-17 scripts/map_functions.R

@@ -6,24 +6,26 @@ source(file.path(scripts_dir, 'append_census.R'))
 source(file.path(scripts_dir, 'metro_eps_codes.R'))
 
 all_codes <- list(
-  philly = list(name = 'Philadelphia', eps = philly_eps_codes),
+  philadelphia = list(name = 'Philadelphia', eps = philly_eps_codes),
   dallas = list(name = 'Dallas', eps = dallas_eps_codes),
-  atl = list(name = 'Atlanta', eps = atl_eps_codes),
+  atlanta = list(name = 'Atlanta', eps = atl_eps_codes),
   chicago = list(name = 'Chicago', eps = chi_eps_codes),
   cleveland = list(name = 'Cleveland', eps = cleveland_eps_codes),
-  northern_nj = list(name = 'Northern New Jersey', eps = nj_north_metro_eps_codes),
+  northern_new_jersey = list(name = 'Northern New Jersey', eps = nj_north_metro_eps_codes),
   houston = list(name = 'Houston', eps = houston_eps_codes),
   bay_area = list(name = 'Bay Area', eps = bay_area_eps_codes),
   long_island = list(name = 'Long Island', eps = long_island_eps_codes),
-  ny_ny = list(name = 'New York, New York', eps = nyny_metro_eps_codes),
+  new_york_city = list(name = 'New York, New York', eps = nyny_metro_eps_codes),
   detroit = list(name = 'Detroit', eps = detroit_eps_codes),
   boston = list(name = 'Boston', eps = boston_eps_codes),
   miami = list(name = 'Miami', eps = miami_eps_codes),
-  dmv = list(name = 'DMV', eps = dmv_eps_codes),
+  dc_maryland_virginia = list(name = 'DMV', eps = dmv_eps_codes),
   orange_county = list(name = 'Orange County', eps = orange_county_eps_codes),
   san_diego = list(name = 'San Diego', eps = san_diego_eps_codes),
   los_angeles = list(name = 'Los Angeles', eps = los_angeles_eps_codes)
 )
+
+sf_use_s2(T)
 
 regions_data <- data.frame(
   region = flatten_chr(map(names(all_codes), \(x) rep(x, length(all_codes[[x]]$eps)))),
@@ -71,14 +73,20 @@ format_vars <- function(data_df) {
     )
 }
 
[…]

### 2025-02-17 scripts/rq1_maps.R

@@ -21,11 +21,11 @@ create_rq1_map <- function(metros, shared_legend = F) {
   yrs <- c(1980, 2000, 2020)
   
   m <- leaflet() %>%
-    addProviderTiles(providers$CartoDB.Positron) %>%
+    addProviderTiles(providers$CartoDB.Positron) # %>%
     
-    addEasyButton(easyButton(
-      icon = 'fa-globe', title = 'Select Metro Area',
-      onClick = JS("function(btn, map){ $('.custom-control').not('#metro-control').slideUp(); $('#metro-control').slideToggle(); }")))
+    # addEasyButton(easyButton(
+    #   icon = 'fa-globe', title = 'Select Metro Area',
+    #   onClick = JS("function(btn, map){ $('.custom-control').not('#metro-control').slideUp(); $('#metro-control').slideToggle(); }")))
   
   for (metro in metros) {
     print(metro)
@@ -111,3 +111,8 @@ create_rq1_map(c('philly', 'chicago'))
 
 create_rq1_map(c('bay_area', 'philly', 'chicago'))
 saveWidget(create_rq1_map(c('bay_area', 'philly', 'chicago')), file.path('.', 'results', 'maps', 'rq1_map_bayarea_philly_chicago.html'), background = 'transparent', selfcontained = T)
+
+
+for (region in regions_data$region) {
+  saveWidget(create_rq1_map(region), file.path('.', 'results', 'maps', paste0('rq1_map_', region, '.html')), background = 'transparent', selfcontained = T)
+}

### 2025-02-17 scripts/rq1_maps.js

@@ -16,19 +16,19 @@ function(el, x, choices) {
   $('.easy-button-button').css('width', 'auto');
   $('.easy-button-button .button-state .fa').css({'float': 'left', 'margin-top': '8px'});
   
-  $('button[title="Select Metro Area"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Metro Area</span>');
+  // $('button[title="Select Metro Area"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Metro Area</span>');
   
   // metro selection options
   
-  let metroControlHTML = '<div id="metro-control" class="leaflet-control-layers leaflet-control leaflet-control-layers-expanded custom-control" style="width: auto; height: auto;">';
+  // let metroControlHTML = '<div id="metro-control" class="leaflet-control-layers leaflet-control leaflet-control-layers-expanded custom-control" style="width: auto; height: auto;">';
   
-  region_name.forEach(function(curr, idx) {
-    metroControlHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="metro-choice" data-region="' + region[idx] + '" data-lat="' + latitude[idx] + '" data-lng="' + longitude[idx] + '"><span> ' + curr + '</span></div>';
-  });
+  // region_name.forEach(function(curr, idx) {
+  //   metroControlHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="metro-choice" data-region="' + region[idx] + '" data-lat="' + latitude[idx] + '" data-lng="' + longitude[idx] + '"><span> ' + curr + '</span></div>';
+  // });
     
-  metroControlHTML += '</div>';
+  // metroControlHTML += '</div>';
   
-  $('button[title="Select Metro Area"]').parent().after(metroControlHTML);
+  // $('button[title="Select Metro Area"]').parent().after(metroControlHTML);
   
   // year selection options
   
@@ -64,7 +64,7 @@ function(el, x, choices) {
   
   $('.leaflet-control-layers-base label:nth-child(5)').after(raceControlHTML);
   
-  $('#race-container').append(raceOptions).slideUp();
+  $('#race-container').append(raceOptions).slideUp(0);
   
   $('#race-label').on('click', function(e) {
     $(this).toggleClass('active');
@@ -79,21 +79,21 @@ function(el, x, choices) {
   
   // handle selections
     
-  $('input[name="metro-choice"]').on('change', function(e) {
[…]

### 2025-02-17 scripts/rq2_maps.R

@@ -181,7 +181,17 @@ student_data <- purchased_data %>%
     other_somecollege = 0,
     other_nosomecollege = other_nocollege + other_somecollege,
     all_nosomecollege = all_nocollege + all_somecollege
-  ) %>% 
+  )
+
+student_data_multi_orders <- lapply(c('488035_488053', '448427_448440', '448420_448374'), function(orders) {
+  student_data %>% 
+    filter(ord_num %in% str_split(orders, '_')[[1]]) %>%
+    mutate(ord_num = orders) %>% 
+    group_by(univ_id, ord_num, hs_ncessch) %>% 
+    summarise(across(aian_all:all_nosomecollege, sum), .groups = 'drop')
+})
+
+student_data <- do.call(bind_rows, c(list(student_data), student_data_multi_orders)) %>% 
   rename('ncessch' = 'hs_ncessch') %>%
   inner_join(hs_data, by = 'ncessch') %>%
   mutate(
@@ -200,10 +210,31 @@ student_data <- purchased_data %>%
     )
   )
 
+lists_orders_zip_hs_df %>% 
+  select(ord_num, ord_sat_score_min, ord_sat_score_max, ord_psat_score_min, ord_psat_score_max) %>% 
+  filter(ord_num %in% c('487984', '488035', '488053', '448922', '448427', '448440', '448375', '448374', '448420', '329702')) %>% 
+  distinct() %>% 
+  View()
+
 orders_data <- data.frame(
-  region = c(rep('chicago', 3), rep('philly', 3), rep('long_island', 3), rep('bay_area', 3)),
-  order_num = c('487984', '488035', '488053', rep(c('448922', '448427', '448440'), 2), '448375', '448420', '448374'),
-  order_title = c('SAT 1020-1150', 'SAT 1160-1300', 'SAT 1310-1600', rep(c('PSAT 1070-1180', 'PSAT 1190-1260', 'PSAT 1270-1520'), 2), 'PSAT 1070-1180', 'PSAT 1190-1260', 'PSAT 1270-1520')
+  region = c(
+    rep('chicago', 4),
+    rep('philadelphia', 4), rep('northern_new_jersey', 4), rep('long_island', 4), rep('detroit', 4), rep('dallas', 4),
+    rep('los_angeles', 4), rep('orange_county', 4), rep('san_diego', 4), rep('bay_area', 4),
+    'houston'
+  ),
+  order_num = c(
[…]

### 2025-02-17 scripts/rq2_maps.js

@@ -26,20 +26,20 @@ function(el, x, choices) {
   $('.easy-button-button').css('width', 'auto');
   $('.easy-button-button .button-state .fa').css({'float': 'left', 'margin-top': '8px'});
   
-  $('button[title="Select Metro Area"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Metro Area</span>');
+  // $('button[title="Select Metro Area"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Metro Area</span>');
   $('button[title="Select Order Number"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Order Number</span>');
   
   // metro selection options
   
-  let metroControlHTML = '<div id="metro-control" class="leaflet-control-layers leaflet-control leaflet-control-layers-expanded custom-control" style="width: auto; height: auto;">';
+  // let metroControlHTML = '<div id="metro-control" class="leaflet-control-layers leaflet-control leaflet-control-layers-expanded custom-control" style="width: auto; height: auto;">';
   
-  region_name.forEach(function(curr, idx) {
-    metroControlHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="metro-choice" data-region="' + region[idx] + '" data-lat="' + latitude[idx] + '" data-lng="' + longitude[idx] + '"><span> ' + curr + '</span></div>';
-  });
+  // region_name.forEach(function(curr, idx) {
+  //   metroControlHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="metro-choice" data-region="' + region[idx] + '" data-lat="' + latitude[idx] + '" data-lng="' + longitude[idx] + '"><span> ' + curr + '</span></div>';
+  // });
     
-  metroControlHTML += '</div>';
+  // metroControlHTML += '</div>';
   
-  $('button[title="Select Metro Area"]').parent().after(metroControlHTML);
+  // $('button[title="Select Metro Area"]').parent().after(metroControlHTML);
   
   // order selection options
   
@@ -104,7 +104,7 @@ function(el, x, choices) {
     orderOptionsHTML = '';
     order_region.forEach(function(curr, idx) {
       if (curr === active_attr.active_metro) {
[…]

### 2025-02-18 index.qmd
tweakin

@@
 ::: {.column width="50%"}
-**[University of Chicago admissions counselors, Chicago-land](https://collegeadmissions.uchicago.edu/contact)**
+**[University of Chicago admissions counselors, IL territories](https://collegeadmissions.uchicago.edu/contact)**
 ![](./assets/images/chicagoland_territories_u_chicago_redacted.png)

### 2025-02-18 scripts/rq2_tables_graphs_run_functions.R

@@ -55,6 +55,12 @@ source(file = file.path(scripts_dir, 'rq2_tables_graphs_create_functions.R'))
 ###########################
 ############################ RUN FUNCTION create_sim_eps_graph TO CREATE GRAPH OF RACE AND GRAPH OF FIRSTGEN
 ###########################
+#CA ORDER NUMBERS
+
+# 2 448375  33893 FA20 - CA PSAT AD {JAN19); ordered 1/8/2019; 2020 HS grad class; CA; PSAT 1070 - 1180
+# 6 448420  19437; FA20 - CA PSAT SE (JAN19); ordered 1/8/2019; 2020/21 hs grad class; CA; PSAT 1190-1260
+# 5 448374  19775; FA20 - CA PSAT BE (JAN19); ordered 1/8/2019; 2020/21 hs grad class; CA; PSAT 1270-1520
+
 
 all_orders <- list(
   chicago = c('chicago', 'chi_eps_codes', 'SAT score 1020 - 1150', '487984'),
@@ -64,27 +70,27 @@ all_orders <- list(
   #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1110 - 1210', '366935'), # ordered jan 2018; don't like these cutpoints as much
   #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1220 - 1520', '366934', '366932'),
   los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1070 - 1180', '448375'), # ordered jan 2019
-  los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1220 - 1520', '448374', '448420'),
+  los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1190 - 1520', '448374', '448420'),
   #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1070 - 1180', '546954'), # ordered jan 2020; much smaller sample size than orders from jan 2019
-  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1220 - 1520', '546946', '546945'),
+  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1190 - 1520', '546946', '546945'),
   #orange_county = c('orange_county', 'orange_county_eps_codes', 'PSAT score 1110 - 1210', '366935'), # ordered jan 2018; don't like these cutpoints as much
[…]

### 2025-02-19 index.qmd
make maps fullscreen

@@
       writeLines(str_c('<iframe data-src="', rq1_map, '" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'))
+      writeLines('<p class="fullscreen">Fullscreen</p>')
 
@@
       writeLines(str_c('<iframe data-src="./results/maps/rq2_map_', str_replace_all(m, " ", "_"), '.html" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'))
+      writeLines('<p class="fullscreen">Fullscreen</p>')
 

### 2025-02-19 assets/js/script.js

@@ -63,6 +63,10 @@ $(function() {
     });
   });
   
+  $('.fullscreen').on('click', function() {
+    $(this).prev('iframe').get(0).requestFullscreen();
+  });
+  
   const h3 = $('.slide:not(.hide) h3').map(function() {
     return $(this).text();
   }).get();

### 2025-02-19 index.qmd
think i finished presentation

@@
 
-Historically, College Board offered two recruiting products to colleges
+Two recruiting products sold by College Board
 
-- Student Search Service (est. 1972): purchase contact information of prospects that meet "search filter" criteria
-- Enrollment Planning Service (est. 1984): software-as-service, says which "Geomarkets" to target and which high schools to visit
+- Student Search Service (est. 1972): purchase contact information of prospects that meet "search filter" criteria [@RN5012]
+- Enrollment Planning Service (EPS, est. 1984): software recommends which "Geomarkets" to target and which high schools to visit
 
@@
 
-# Background
+# Background and Theory
 
@@
 - Meant to reproduce the recruiting territories of admissions officers
-- @RN4982 create 143 Geomarkets in three regions
-- Enrollment Planning Service (EPS) uses 304 Geomarkets
+- @RN4982 has 143 Geomarkets in 3 regions; EPS has 304
 
-**Market Segment Model**, based on student SAT score-sending behavior [SIMPLIFY]
+**Market Segment Model**
 
-- Each student-to-college score assigned to one of four categories:
-  - Local (in Geomarket); in-state (diff Geomarket); regional; national
-- Students sorted into 4 market segments: local, in-state, regional, national
-  - "Local" students submit most SAT scores to local institutions
-- *Market Segment Profile* (produced by Geomarket): 
-  - How many students in each market segment
-- *Institutional Profile*  (produced by Geomarket): 
-  - How many students by market segement send scores to college *i*
+- Predicts how demand for a college varies by Geomarket, based on characteristics of households in Geomarket
+- Categorize students into 4 market segments -- local, in-state, regional, national -- based on SAT score-sending behavior
+  - local: send most scores to colleges within Geomarket
+  - regional: send most scores out-of-state but in-region (e.g., Midwest)
+- "Good" colleges should look for regional and national students
+  - Like @RN2247, except College Board creates product from it
+  - @RN4982 create outputs that become EPS software
 
-# Theory
+### Scholarship on Quantification {.hide}
+#### Discussions of correlation and homophily by @RN4975[chapter 1]
+
+<p class="quote">**Correlation**: extent to which variables move together</p>
+<p style="display:none;">Our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations” [@RN4982, p.
+42].</p>
+
+- Correlations from training data caused by structural inequality
+  - "Predicting the future on the basis of the past threatens to reify and reproduce existing inequalities" [@RN4794, p. 224]
+- @RN4982 find 4 predictors of student market segment
+  - Parental education, family income, SAT score, educational aspirations
+  - Recommend selective colleges target high-SES Geomarkets
+
+<p class="quote">**Homophily**: birds of a feather flock together</p>
+<p style="display:none;">The layering of collegiate competition is primarily a socioeconomic layering.
+The hierarchical structure of collegiate competition reflects the stratified social and economic dimensions of communities from which colleges draw their students.
+Competition among colleges, as admissions officers have told us for so long, is a matter of keeping company with one’s peers [@RN4982, p.
+72].</p>
+
+- Homophily not the result of voluntary action; homophily programmed into algorithms that create connections
+  - Match vertically ranked consumers to vertically ranked producers
+- @RN4982: like colleges compete for like students (SES)
+<p class="quote" style="font-size: 0.85em;">Recommend colleges target Geomarkets, schools targeted by peer colleges</p>
+<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
+Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
+The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice [@RN4982, p.
+44].</p>
+
+<!--
 
@@
 44].</p>
+ -->
 
-### Research questions and expectations{.hide}
+### Research questions{.hide}
 
-**Research goals**
+<br>
 
-- Examine extent of social inequaltiy between Geomarkets (RQ1)
-- Show how application of Geomarkets produces inequality (RQ2)
+**Research interests**
+
+- How unequal are these Geomarkets?
+- Who is included/excluded if we purchase student lists and filter on Geomarkets as recommended by @RN4982
+
+<br>
 
@@
 
-1. SES and racial variation between, within Geomarkets? Change over time?
-
-    - Geomarkets/Market Segment Model based on social stratification
-    - US Racial segregation, correlation between class and race
-
-1. SES and racial composition of included vs. excluded prospects, if "student list" purchases filter on particular Geomarkets?
-
-    - Excluding Geomarkets that are poor and non-White associated with:
-      - Disproportionate exclusion of first-generation students
-      - Disproportionate exclusion of non-white students
-      - Disproportionate exclusion of first-generation, non-white students
+1. What is the socioeconomic and racial variation between and within Geomarkets? How does this change over time?
+1. What is the socioeconomic and and racial composition of included vs. excluded prospects, if "student list" purchases filter on particular Geomarkets?
 
@@
   - 1980, 2000 Census; 2016-2020 American Community Survey (ACS)
-  - Variables: race/Hispanic origin; median income, % poverty, % BA+
-- Geomarket shapefiles obtained from 2012 blog post
+- Geomarket shapefiles obtained from a 2012 blog post
   - Each zip code assigned to one Geomarket
@@
 
-- 414 student lists -- associated with 2.5 million prospects -- purchaseds by 14 public universities from 2016-2020
+- Student lists purchased by 14 public universities from 2016-2020
+  - 414 student lists associated with 2.5 million prospects
   - Each prospect has a home zip code and a HS zip code (point geoms)
 - Analyze purchases that include all Geomarkets in a metro area
-  - Show who would be excluded had search criteria exluded a Geomarket
+  - Show who would be included/excluded had search criteria filtered on particular Geomarket(s)
 
 ### Methods{.hide}
-#### Research design: quantitative case study
-
-Cases are metro areas associated with multiple Geomarkets (n = 17)
+#### Research design is a quantitative case study
 
@@
 
-Case selection
+Cases
 
+- Cases are metro areas associated with multiple Geomarkets (n ~ 17)
 - Initial case selection based on availability of student list purchases (RQ2)
@@
 
-- Poor, non-white Geomarkets produce disproportionate share of Black, Hispanic students, (and often Asian)
+- Poor, non-white Geomarkets produce disproportionate share of Black, Hispanic (and often Asian) students
 - Poor, non-white Geomarkets produce disproportionate share of first-generation students
-- Purchased student profiles by race X parental education
+- Intersection of race and parental education
   - Asian students from low-income Geomarkets tend to be first-gen compared to Asain students from affluent Geomarkets
@@
   - But EPS, Student Search Service products structure college access
-  - College access structured by third-party vendors and products that amplify structural inequalities observed on demand-side
+  - College access structured by third-party vendors, products that amplify structural inequalities observed on demand-side
 - Org field is "totality of relevant actors" [@RN527, p. 148]
-  - The organizational field should be focus regulation
+  - Organizational field should be focus of regulation
   - Sociology should identify high-leverage actors/products in org field

### 2025-02-19 scripts/rq1_tables_graphs.R

@@ -417,9 +417,7 @@ create_eps_graph <- function(eps_codes,
   # 2) Create the text you want to store [REVISE NOTE TEXT LATER!]
   note_text <- c(
     "Figure Notes:",
-    "- Median income is shown in thousands ($k).",
-    "- '% in poverty' is the share of individuals below the federal poverty line.",
-    "- Data sources: US Census, etc.",
+    "- Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic",
     note
   )
   
@@ -445,7 +443,7 @@ all_codes <- list(
   'chicago' = chi_eps_codes,
   'cleveland' = cleveland_eps_codes,
   'northern new jersey' = nj_north_metro_eps_codes,
-  'houston' = htown_eps_codes,
+  'houston' = houston_eps_codes,
   'bay area' = bay_area_eps_codes,
   'long island' = long_island_eps_codes,
   'new york city' = nyny_metro_eps_codes,
@@ -471,7 +469,7 @@ for (metro_name in names(all_codes)) {
       eps_codes  = metro_codes,
       prefix     = metro_name,        # <--- pass "philly" or "dallas"
       graph_type = graph_types[g],
-      note       = "some note"
+      note       = NULL
     )
   }
 }

### 2025-02-19 scripts/rq2_tables_graphs_create_functions.R

@@ -1051,9 +1051,7 @@ create_race_by_firstgen_graph <- function(data_graph,
   # 2) Create the text you want to store [REVISE NOTE TEXT LATER!]
   note_text <- c(
     "Figure Notes:",
-    "- Median income is shown in thousands ($k).",
-    "- '% in poverty' is the share of individuals below the federal poverty line.",
-    "- Data sources: US Census, etc."
+    "- Excludes students who have missing values for race or parental education"
   )
   
   # 3) Write that text to a file

### 2025-02-19 scripts/rq2_tables_graphs_run_functions.R

@@ -289,12 +289,9 @@ for (m in c("chicago", "philadelphia", "los angeles",
           )
           
           # Write out a .txt file with the footnotes
-          note_text <- c(
-            "Figure Notes:",
-            "- Median income is shown in thousands ($k).",
-            "- '% in poverty' is the share of individuals below the federal poverty line.",
-            "- Data sources: US Census, etc."
-          )
+          note_text <- c('Figure Notes:',str_c(
+            "- Excludes students with missing values for", g, sep = ' '
+          ))
           writeLines(
             note_text, 
             file.path(graphs_dir, "rq2", str_c(plot_name, "_note.txt"))
@@ -335,12 +332,9 @@ for (m in c("chicago", "philadelphia", "los angeles",
             )
             
             # Write out a .txt file with the footnotes
-            note_text <- c(
-              "Figure Notes:",
-              "- Median income is shown in thousands ($k).",
-              "- '% in poverty' is the share of individuals below the federal poverty line.",
-              "- Data sources: US Census, etc."
-            )
+            note_text <- c('Figure Notes:',str_c(
+              "- Excludes students with missing values for", g, sep = ' '
+            ))
             writeLines(
               note_text, 
               file.path(graphs_dir, "rq2", str_c(plot_name, "_note.txt"))
@@ -385,12 +379,9 @@ for (m in c("chicago", "philadelphia", "los angeles",
         )
         
         # Footnotes
-        note_text <- c(
-          "Figure Notes:",
-          "- Median income is shown in thousands ($k).",
-          "- '% in poverty' is the share of individuals below the federal poverty line.",
-          "- Data sources: US Census, etc."
-        )
+        note_text <- c('Figure Notes:',str_c(
+          "- Excludes students with missing values for", g, sep = ' '
+        ))
[…]

### 2025-02-19 index.qmd
la la

@@
 
-- Correlations from training data caused by structural inequality
+- Correlations observed in training data caused by structural inequality
   - "Predicting the future on the basis of the past threatens to reify and reproduce existing inequalities" [@RN4794, p. 224]

### 2025-02-21 index.qmd
tweaks to presentation

@@
   - 414 student lists associated with 2.5 million prospects
-  - Each prospect has a home zip code and a HS zip code (point geoms)
+    - Data: search filter criteria; and de-identified student-level data
+    - Each prospect has a home and a HS zip code (point geoms)
 - Analyze purchases that include all Geomarkets in a metro area
-  - Show who would be included/excluded had search criteria filtered on particular Geomarket(s)
+  - Show who included/excluded if we filter on particular Geomarkets
 
@@
 
-- Substantial inequality between and within Geomarkets
+- Geomarkets highly correllated with race and class in many metros
 - In 1980, Black people are usually concentrated in poorest Geomarket
   - In some metros, this pattern persists (e.g., Cleveland)
-  - In some metros, these Geomarkets experience gentrification
 
@@
 
-- Poor, non-white Geomarkets produce disproportionate share of Black, Hispanic (and often Asian) students
+- Poor, non-white Geomarkets produce disproportionate share of Black and Hispanic students
 - Poor, non-white Geomarkets produce disproportionate share of first-generation students
 - Intersection of race and parental education
-  - Asian students from low-income Geomarkets tend to be first-gen compared to Asain students from affluent Geomarkets
-  - Same result for Hispanic students
+  - Compared to students from affluent Geomarkets, Asian and Hispanic students from low-income Geomarkets tend to be first-gen
+  - Targeting non-white students in affluent Geomarkets yields mostly non-first-gen
 
@@
 
-- Scholarship has not examined "b2b" quantification that classifies consumers for the sake of producers
+- Hasn't examined quantification that classifies consumers for producers
+  - Structural inequality $\rightarrow$ unequal student demand for college
+  - College access structured by third-party products that encourage supply-side to amplify structural inequalities observed on demand-side
 - @RN5005 show effect of quantification, but not mechanisms of *how* quantification produces inequality
@@
 
-- Scholarship focuses on schools and policies as "sorting machines" [@RN4873]
-  - But EPS, Student Search Service products structure college access
-  - College access structured by third-party vendors, products that amplify structural inequalities observed on demand-side
+- Soc-of-ed focuses on schools (and policies) as "sorting machines"
+  - But third-party vendors have structured college access for decades
 - Org field is "totality of relevant actors" [@RN527, p. 148]

### 2025-02-21 index.qmd
final sea presentationgit add .

@@
 
-### New York and Chicago Geomarkets {.hide}
+### Geomarkets around Chicago-land {.hide}
 
@@
 <p style="display:none;">Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
-Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs [@RN4982, p.
+**Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.** In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful.
+[@RN4982, p.
 4].</p>
@@
 - Census tract-level data
-  - 1980, 2000 Census; 2016-2020 American Community Survey (ACS)
+  - 1980, 2000 Census, 2016-2020 American Community Survey (ACS)
 - Geomarket shapefiles obtained from a 2012 blog post
@@
   - Structural inequality $\rightarrow$ unequal student demand for college
-  - College access structured by third-party products that encourage supply-side to amplify structural inequalities observed on demand-side
+  - Third-party products take snapshot student demand $\rightarrow$ encourage supply-side to amplify structural inequalities observed on demand-side
 - @RN5005 show effect of quantification, but not mechanisms of *how* quantification produces inequality

### 2025-02-26 index.qmd
changing metro orders and all figures

@@
 
-knitr::opts_chunk$set(echo = F, message = F)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE)
 
@@
 
-  # Convert spaces -> underscores for filenames
-  metro_underscored <- str_replace_all(metro, " ", "_")
+  # 'metro' is already underscored (e.g. "los_angeles").
+  # We'll use it directly for file paths, but if you want
+  # to be extra safe, you can do:
+  # metro_underscored <- str_replace_all(metro, " ", "_")
+  # ... (assuming it has no spaces).
 
-  # Build the three file paths (PNG, title TXT, note TXT)
-  fig_file  <- file.path(base_path, rq, str_c(rq, "_", metro_underscored, "_", graph_type, ".png"))
-  cap_file  <- file.path(base_path, rq, str_c(rq, "_", metro_underscored, "_", graph_type, "_title.txt"))
-  note_file <- file.path(base_path, rq, str_c(rq, "_", metro_underscored, "_", graph_type, "_note.txt"))
+  fig_file  <- file.path(base_path, rq, str_c(rq, "_", metro, "_", graph_type, ".png"))
+  cap_file  <- file.path(base_path, rq, str_c(rq, "_", metro, "_", graph_type, "_title.txt"))
+  note_file <- file.path(base_path, rq, str_c(rq, "_", metro, "_", graph_type, "_note.txt"))
 
@@
 
-# 2) Define your metros / RQs
+# 2) Define your *underscored* metros / RQs
 metros <- c(
   "atlanta",
-  "bay area",
+  "bay_area",
   "boston",
@@
   "dallas",
-  "dc maryland virginia",
+  "dc_maryland_virginia",
   "detroit",
   "houston",
-  "long island",
-  "los angeles",
+  "long_island",
+  "los_angeles",
   "miami",
-  "new york city",
-  "northern new jersey",
-  "orange county",
+  "new_york_city",
+  "northern_new_jersey",
+  "orange_county",
   "philadelphia",
-  "san diego"
+  "san_diego"
 )
@@
 
-# 3) Subset of metros with RQ2 data (2 orders each)
-rq2_metros <- c(
-  "chicago", "philadelphia", "los angeles", 
-  "orange county", "san diego", "bay area",
-  "northern new jersey","long island",'detroit','dallas','houston'
-)
+# 3) Read the CSV (underscored metro column)
+orders_df <- read_csv("scripts/metro_orders.csv")
 
-# We'll define a "metro_orders" list for those RQ2 metros:
-metro_orders <- list(
-  "chicago"             = c("487984", "488035_488053"),
-  "philadelphia"        = c("448922", "448427_448440"),
-  "northern_new_jersey" = c("448922", "448427_448440"),
-  "long_island"         = c("448922", "448427_448440"),
-  "detroit"             = c("448922", "448427_448440"),
-  "dallas"              = c("448922", "448427_448440"),
-  "los_angeles"         = c("448375", "448374_448420"),
-  "orange_county"       = c("448375", "448374_448420"),
-  "san_diego"           = c("448375", "448374_448420"),
-  "bay_area"            = c("448375", "448374_448420"),
-  'houston'             = c('329702') #   # houston = c('houston', 'houston_eps_codes', 'PSAT score 1010 - 1520', '329702')
-
-)
+# Derive the RQ2 metros from the CSV
+rq2_metros <- unique(orders_df$metro)
 
@@
 
-  # Print a top-level heading for the metro area
-  writeLines(str_c("## ", tools::toTitleCase(m), " area"))
+  # Convert "los_angeles" -> "Los Angeles" for display in slides
+  heading_text <- gsub("_", " ", m, fixed = TRUE)
+  heading_text <- tools::toTitleCase(heading_text)
+
+  # Print a top-level heading for the metro area in user-friendly form
+  writeLines(str_c("## ", heading_text, " area"))
 
@@
       # Each metro has 2 plots: "race" & "ses"
-      for (i in seq(graph_types_rq1)) {
+      for (i in seq_along(graph_types_rq1)) {
         g <- graph_types_rq1[i]
 
+        # The first RQ1 slide for a metro has a real heading;
+        # subsequent ones are hidden slides (".hide")
         writeLines(str_c("### Research question ", rq_num, if_else(i == 1, "", " {.hide}")))
@@
 
-      rq1_map <- if_else(
-        m == 'los angeles',
-        'https://rq1-map-los-angeles.netlify.app',
-        str_c('./results/maps/rq1_map_', str_replace_all(m, " ", "_"), '.html')
-      )
+      # Use 'm' for the file path to the map (underscored version).
+      # If you had a special case for "los_angeles", you could handle that here.
+      rq1_map <- str_c('./results/maps/rq1_map_', m, '.html')
+
+      # But let's illustrate one special example for "los_angeles":
+      if (m == "los_angeles") {
+        rq1_map <- 'https://rq1-map-los-angeles.netlify.app'
+      }
 
@@
 
-      # Convert name -> underscores for matching list keys
-      m_underscored <- str_replace_all(m, " ", "_")
-      orders_for_m  <- metro_orders[[m_underscored]]
+      # Filter orders_df to just rows for this underscored metro
+      df_metro <- filter(orders_df, metro == m)
 
-      # If a metro isn't in the list or has no orders, skip
-      if (is.null(orders_for_m)) next
+      # If none, skip
+      if (nrow(df_metro) == 0) next
+
+      # Gather all the underscore-delimited order_ids
+      orders_for_m <- df_metro$order_ids
 
       # We'll do 2 group types: race & firstgen
-      for (i in seq(graph_types_rq2)) {
+      for (i in seq_along(graph_types_rq2)) {
         g <- graph_types_rq2[i]
@@
 
-              # e.g. "rq2_chicago_race_col_plot_487984.png"
+              # e.g. "race_col_plot_448375_546954"
+              # => "rq2_los_angeles_race_col_plot_448375_546954.png"
               graph_type_col <- str_c(g, "_", rc, "_plot_", order_id)
@@
       # -------------------------------------------
-      # ADDITIONAL: race_by_firstgen for each order
+      # ADDITIONAL: "race_by_firstgen" => goes to RQ2B
       # -------------------------------------------
-      # Like the column approach, one slide per order_id
       for (order_id in orders_for_m) {
         writeLines(str_c("### Research question ", rq_num, " {.hide}"))
-        # Filenames look like:
-        #   rq2_northern_new_jersey_order_448922_race_by_firstgen.png
+        # e.g. "rq2b_los_angeles_order_448375_546954_race_by_firstgen.png"
         graph_type_rbf <- str_c("order_", order_id, "_race_by_firstgen")
-        insert_figure(rq = rq, metro = m, graph_type = graph_type_rbf)
+        insert_figure(rq = "rq2b", metro = m, graph_type = graph_type_rbf)
       }
@@
 
-      writeLines(str_c('<iframe data-src="./results/maps/rq2_map_', str_replace_all(m, " ", "_"), '.html" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'))
+      writeLines(str_c(
+        '<iframe data-src="./results/maps/rq2_map_',
+        m,
+        '.html" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'
+      ))
       writeLines('<p class="fullscreen">Fullscreen</p>')

### 2025-02-26 scripts/metro_eps_codes.R

@@ -24,6 +24,7 @@ philly_eps_codes <- paste0("PA ", 1:5) #
 philly_eps_codes
 
 long_island_eps_codes <- c(paste0('NY',16:21))
+long_island_eps_codes
 
 atl_eps_codes <- paste0("GA ", 1:4) #
 atl_eps_codes
@@ -38,6 +39,7 @@ houston_eps_codes
   # 
 
 dallas_eps_codes <- c('TX19','TX20','TX21','TX22','TX23','TX24')
+dallas_eps_codes
 
 nj_eps_codes <- c(paste0("NJ ", 1:9), paste0("NJ", 10:12))
 nj_eps_codes

### 2025-02-26 scripts/rq2_metro_orders.R

@@ -1,0 +1,398 @@
+################################################################################
+## [ PROJ ] < College Board Geomarket >
+## [ FILE ] < rq2_metro_orders.R >
+## [ AUTH ] < Ozan Jaquette >
+## [ INIT ] < 2/25/2025
+## [ DESC ] < Create a csv file that has info about the orders by metro >
+################################################################################
+
+# Load necessary packages
+library(dplyr)
+library(readr)
+
+# -------------------------------------------------------------------------
+# Purpose: Convert revised 'all_orders' list into a CSV containing
+#   (1) metro
+#   (2) eps_codes
+#   (3) order_ids
+#   (4) test_range
+#   (5) figure_note
+#
+# Note: We'll write to "metro_orders.csv" in the working directory.
+# -------------------------------------------------------------------------
+
+library(dplyr)
+library(readr)
+
+# 1) Your revised all_orders list (example below):
+all_orders <- list(
+  # CHICAGO
+  chicago = c(
+    "chicago",
+    "chi_eps_codes",
+    "392833_487984",
+    "SAT score 1020 - 1150",
+    "Illinois standard 2019 and 2020; some notes here"
+  ),
+  chicago = c(
+    "chicago",
+    "chi_eps_codes",
+    "392834_488035",
+    "SAT score 1160 - 1300",
+    "Illinois Honors 2019 and 2020; some notes here"
+  ),
+  chicago = c(
+    "chicago",
+    "chi_eps_codes",
+    "392835_488053",
+    "SAT score 1310 - 1600",
+    "Illinois GPPA 2019 and 2020; some notes here"
+  ),
+  
+  # PHILADELPHIA
+  philadelphia = c(
+    "philadelphia",
+    "philly_eps_codes",
+    "448922",
+    "PSAT score 1070 - 1180",
+    "Some notes here"
+  ),
+  philadelphia = c(
+    "philadelphia",
+    "philly_eps_codes",
+    "448427",
+    "PSAT score 1190 - 1260",
+    "Some notes here"
+  ),
+  philadelphia = c(
+    "philadelphia",
+    "philly_eps_codes",
+    "448440",
+    "PSAT score 1270 - 1520",
[…]

### 2025-02-26 scripts/rq2_tables_graphs_create_functions.R

@@ -904,14 +904,16 @@ custom_race_labels <- c(
 )
 
 
-create_race_by_firstgen_graph <- function(data_graph,
-                                          ord_nums_graph,
-                                          eps_codes_graph,
-                                          metro_name = '',
-                                          exclude_race_graph = c(1,8,12),
-                                          title_suf = '') {
-
-  # 3) Create & pivot the table
+create_race_by_firstgen_graph <- function(
+    data_graph,
+    ord_nums_graph,
+    eps_codes_graph,
+    metro_name = '',
+    exclude_race_graph = c(1,8,12),
+    title_suf = ''
+) {
+  
+  # 1) Build the main data frame
   df <- create_sim_eps_race_firstgen_table(
     data      = data_graph, 
     ord_nums  = ord_nums_graph, 
@@ -924,143 +926,144 @@ create_race_by_firstgen_graph <- function(data_graph,
       values_to = "value"
     ) %>%
     mutate(
-      # Relabel race
       stu_race_cb = factor(
         stu_race_cb, 
         levels = names(custom_race_labels), 
         labels = custom_race_labels
-      ),      
-      # Relabel group
+      ),
       group = factor(
         group, 
         levels = custom_group_order, 
         labels = custom_group_labels
       ),
-      # Reverse the order of eps_codename globally
       eps_codename = factor(
         eps_codename,
         levels = rev(unique(eps_codename))  # reverse the unique order
       )
     )
   
-  df %>% print(n=250)
-  
-  # 4) Split out the subgroup vs. total N
-  df_plot <- df %>%
-    filter(group != "All")
-  
+  # 2) Separate the "All" group (totals) from the subgroups
+  df_plot <- df %>% filter(group != "All")
   df_totals <- df %>%
     filter(group == "All") %>%
-    distinct(stu_race_cb, eps_codename, value) %>%  # <--- ensure no duplicates
+    distinct(stu_race_cb, eps_codename, value) %>%
     rename(total_n = value)
   
-  # 5) Join the total N onto the plotting data (and remove any exact duplicates)
+  # 3) Attach total_n to each row in df_plot
   df_plot <- df_plot %>%
     left_join(df_totals, by = c("stu_race_cb", "eps_codename")) %>%
-    distinct()
-  
-  # 6) Create a new label (N=xxxx) & factor it once
[…]

### 2025-02-26 scripts/rq2_tables_graphs_run_functions.R

@@ -52,184 +52,85 @@ getwd()
 # script that creates functions to create tables and graphs
 source(file = file.path(scripts_dir, 'rq2_tables_graphs_create_functions.R'))
 
-###########################
-############################ RUN FUNCTION create_sim_eps_graph TO CREATE GRAPH OF RACE AND GRAPH OF FIRSTGEN
-###########################
-#CA ORDER NUMBERS
 
-# 2 448375  33893 FA20 - CA PSAT AD {JAN19); ordered 1/8/2019; 2020 HS grad class; CA; PSAT 1070 - 1180
-# 6 448420  19437; FA20 - CA PSAT SE (JAN19); ordered 1/8/2019; 2020/21 hs grad class; CA; PSAT 1190-1260
-# 5 448374  19775; FA20 - CA PSAT BE (JAN19); ordered 1/8/2019; 2020/21 hs grad class; CA; PSAT 1270-1520
+# 1) Load the CSV into a data frame
+orders_df <- read_csv(file.path(scripts_dir,"metro_orders.csv"))
 
 
-all_orders <- list(
-  chicago = c('chicago', 'chi_eps_codes', 'SAT score 1020 - 1150', '487984'),
-  chicago = c('chicago', 'chi_eps_codes', 'SAT score 1160 - 1600', '488035', '488053'),
-  philadelphia = c('philadelphia', 'philly_eps_codes', 'PSAT score 1070 - 1180', '448922'),
-  philadelphia = c('philadelphia', 'philly_eps_codes', 'PSAT score 1190 - 1520', '448427', '448440'),
-  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1110 - 1210', '366935'), # ordered jan 2018; don't like these cutpoints as much
-  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1220 - 1520', '366934', '366932'),
-  los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1070 - 1180', '448375'), # ordered jan 2019
-  los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1190 - 1520', '448374', '448420'),
-  #los_angeles = c('los_angeles', 'los_angeles_eps_codes', 'PSAT score 1070 - 1180', '546954'), # ordered jan 2020; much smaller sample size than orders from jan 2019
[…]

### 2025-02-26 index.qmd
revisions to the way figures and presentation created

@@
 
-  # 'metro' is already underscored (e.g. "los_angeles").
-  # We'll use it directly for file paths, but if you want
-  # to be extra safe, you can do:
-  # metro_underscored <- str_replace_all(metro, " ", "_")
-  # ... (assuming it has no spaces).
-
+  # We'll use 'metro' directly since it's already underscored (e.g. "los_angeles").
   fig_file  <- file.path(base_path, rq, str_c(rq, "_", metro, "_", graph_type, ".png"))
@@
 
-  # Convert "los_angeles" -> "Los Angeles" for display in slides
+  # Convert underscores -> spaces, then title case
+  # But add special checks for "dc_maryland_virginia" and "bay_area"
+  if (m == "dc_maryland_virginia") {
+    heading_text <- "D.C., Maryland, and Virginia"
+  } else if (m == "bay_area") {
+    heading_text <- "Bay Area"
+  } else {
+    # Normal logic
     heading_text <- gsub("_", " ", m, fixed = TRUE)
     heading_text <- tools::toTitleCase(heading_text)
+    heading_text <- str_c(heading_text, " area")
+  }
 
-  # Print a top-level heading for the metro area in user-friendly form
-  writeLines(str_c("## ", heading_text, " area"))
+  # Print a top-level heading for the metro area
+  writeLines(str_c("## ", heading_text))
 
@@
 
-      # Use 'm' for the file path to the map (underscored version).
-      # If you had a special case for "los_angeles", you could handle that here.
       rq1_map <- str_c('./results/maps/rq1_map_', m, '.html')
 
-      # But let's illustrate one special example for "los_angeles":
+      # Example special case for "los_angeles"
       if (m == "los_angeles") {
@@
 
-              # e.g. "race_col_plot_448375_546954"
-              # => "rq2_los_angeles_race_col_plot_448375_546954.png"
               graph_type_col <- str_c(g, "_", rc, "_plot_", order_id)

### 2025-02-26 scripts/rq1_tables_graphs.R

@@ -321,16 +321,11 @@ create_eps_graph <- function(eps_codes,
       scale_y_continuous(labels = scales::percent_format()) +
       facet_wrap(~ year, ncol = 1, scales = "free_y") +
       theme_custom + 
-      # Add/modify this theme() call:
       theme(
-        # Increase legend text/title size
         legend.text  = element_text(size = 14),
         legend.title = element_text(size = 14),
-        # Increase size of geomarket names (on y-axis after coord_flip)
         axis.text.y  = element_text(size = 12),
-        # Increase size of percentages (on x-axis after coord_flip)
         axis.text.x  = element_text(size = 12),
-        # Increase size of facet labels (i.e., the year labels)
         strip.text   = element_text(size = 14)
       )
     
@@ -371,11 +366,9 @@ create_eps_graph <- function(eps_codes,
       ) +
       facetted_pos_scales(
         y = list(
-          # For Median income
           (variable == "Median income") ~ scale_y_continuous(
             labels = scales::label_number(suffix = "k")
           ),
-          # For the percentage variables
           (variable %in% c("% in poverty", "% with BA+")) ~ scale_y_continuous(
             labels = scales::label_number(suffix = "%")
           )
@@ -387,7 +380,7 @@ create_eps_graph <- function(eps_codes,
   writeLines(file_prefix)
   
   # create strings for file names
-  plot_name <- str_c('rq1',file_prefix_underscore,graph_type, sep = '_')
+  plot_name <- str_c('rq1', file_prefix_underscore, graph_type, sep = '_')
   writeLines(plot_name)
   
   # Save plot to file
@@ -399,22 +392,62 @@ create_eps_graph <- function(eps_codes,
     bg = 'white'
   )
   
-  # Define the figure title
+  # Define which_characteristics for figure title
   if (graph_type == "race") {
     which_characteristics <- 'Racial/ethnic composition' 
   } else if (graph_type == "ses") {
     which_characteristics <- 'Socioeconomic characteristics' 
   }
+  
+  # -------------------------------------------------------------
+  # (A) Build friendly_prefix and do if-else
+  # -------------------------------------------------------------
+  friendly_prefix <- str_to_title(file_prefix)
+  
[…]

### 2025-02-26 scripts/rq2_metro_orders.R

@@ -32,239 +32,230 @@ all_orders <- list(
     "chi_eps_codes",
     "392833_487984",
     "SAT score 1020 - 1150",
-    "Illinois standard 2019 and 2020; some notes here"
+    "Order number 392833 (ordered 7/10/2018): HS class 2019 OR 2020; state = IL; SAT 1020 - 1150; GPA B- to A+. Order number 487984 (ordered 7/19/2019): HS class 2020 OR 2021; state = IL; SAT 1020 - 1150; GPA B- to A+ "
   ),
   chicago = c(
     "chicago",
     "chi_eps_codes",
     "392834_488035",
     "SAT score 1160 - 1300",
-    "Illinois Honors 2019 and 2020; some notes here"
+    "Order number 392834 (ordered 7/10/2018): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300. Order number 488035 (ordered 7/19/2019): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300; GPA B- to A+"
   ),
   chicago = c(
     "chicago",
     "chi_eps_codes",
     "392835_488053",
     "SAT score 1310 - 1600",
-    "Illinois GPPA 2019 and 2020; some notes here"
+    "Order number 392835 (ordered 7/9/2018): HS class 2019 OR 2020; state = IL; SAT 1310 - 1600; GPA B- to A+. Order number 488053 (ordered 7/19/2019): HS class 2020 OR 2021; state = IL; SAT 1310 - 1600; GPA B- to A+ "
   ),
-  
   # PHILADELPHIA
   philadelphia = c(
     "philadelphia",
     "philly_eps_codes",
     "448922",
     "PSAT score 1070 - 1180",
-    "Some notes here"
+    "Order number 448922 (ordered 1/9/2019): HS class 2020; selected Geomarkets; PSAT 1070 - 1180"
   ),
   philadelphia = c(
     "philadelphia",
     "philly_eps_codes",
     "448427",
     "PSAT score 1190 - 1260",
-    "Some notes here"
+    "Order number 448427 (ordered 1/8/2019): HS class 2020; selected Geomarkets; PSAT 1190 - 1260"
   ),
   philadelphia = c(
     "philadelphia",
     "philly_eps_codes",
     "448440",
     "PSAT score 1270 - 1520",
-    "Some notes here"
[…]

### 2025-02-26 scripts/rq2_tables_graphs_create_functions.R

@@ -677,9 +677,8 @@ create_sim_eps_graph <- function(data_graph,
           )
         )
       ) %>%
-      dplyr::filter(!group %in% c("AIAN, non-Hispanic", "NHPI, non-Hispanic",'Two+, non-Hispanic'))
+      dplyr::filter(!group %in% c("AIAN, non-Hispanic", "NHPI, non-Hispanic", "Two+, non-Hispanic"))
     
-    # Titles and legend labels
     row_plot_title   <- "Race Distribution Within Each EPS Code (Row %)"
     col_plot_title   <- "Distribution of Each Race Across EPS Codes (Column %)"
     fill_legend      <- "Geomarket"        # for the column plot
@@ -767,28 +766,29 @@ create_sim_eps_graph <- function(data_graph,
     row_legend_title <- "Parental education" # for the row plot
   }
   
-  # (A) Append the user-supplied title to each plot title if provided
+  # (A) Append the user-supplied title if provided
   if (!is.null(title) && nzchar(title)) {
     row_plot_title <- paste0(title)
     col_plot_title <- paste0(title)
   }
   
-  # Determine which column contains the overall sample size.
+  # (B) For the row plot, build the "eps_label" with sample size
   known_col <- if (variable == "race") "race_known" else "stu_known"
   
-  # ----------------------------------------------------------------
-  # (B) Create a new label that combines the EPS code and its sample size
-  # ----------------------------------------------------------------
   df_r <- df_r %>%
     dplyr::group_by(eps_codename) %>%
     dplyr::mutate(
-      eps_label = paste0(as.character(eps_codename),
-                         " (n=", formattable::comma(first(.data[[known_col]]), digits = 0), ")")
+      eps_label = paste0(
+        as.character(eps_codename),
+        " (n=",
+        formattable::comma(dplyr::first(.data[[known_col]]), digits = 0),
+        ")"
+      )
     ) %>%
     dplyr::ungroup()
   
   # ----------------------------------------------------------------
-  # (C) Build the Row-Percent Plot using the new eps_label
+  # (C) Build the Row-Percent Plot (plot_r)
   # ----------------------------------------------------------------
   plot_r <- df_r %>%
     ggplot2::ggplot(
@@ -798,8 +798,7 @@ create_sim_eps_graph <- function(data_graph,
         fill = group
       )
     ) +
-    ggplot2::geom_bar(stat = "identity", 
[…]

### 2025-02-26 scripts/rq2_tables_graphs_run_functions.R

@@ -56,23 +56,45 @@ source(file = file.path(scripts_dir, 'rq2_tables_graphs_create_functions.R'))
 # 1) Load the CSV into a data frame
 orders_df <- read_csv(file.path(scripts_dir,"metro_orders.csv"))
 
-
+orders_df
 # Part #1) Creating "rq2_{metro}_{g}_plot_order_{ids}" objects
 # via create_sim_eps_graph()
 
 for (i in seq_len(nrow(orders_df))) {
   
-  # Extract row i
   row_i <- orders_df[i, ]
   
-  # Underscored name
-  metro_name_graph <- row_i$metro  # e.g. "long_island"
+  # E.g. "long_island"
+  metro_name_graph <- row_i$metro
   
-  eps_codes_graph  <- get(row_i$eps_codes, envir = .GlobalEnv)
-  graph_title      <- row_i$test_range
+  eps_codes_graph <- get(row_i$eps_codes, envir = .GlobalEnv)
   
-  # For the object name, we keep underscores
+  # Split into a character vector
   ord_nums_graph <- str_split(row_i$order_ids, "_")[[1]]
+  n_orders <- length(ord_nums_graph)
+  
+  # Construct the "orders" portion
+  if (n_orders == 1) {
+    # Single order
+    orders_str <- ord_nums_graph[1]
+    orders_prefix <- "Order "
+  } else if (n_orders == 2) {
+    # Two orders
+    orders_str <- str_c(ord_nums_graph[1], " and ", ord_nums_graph[2])
+    orders_prefix <- "Orders "
+  } else {
+    # Three or more orders
+    all_but_last <- ord_nums_graph[1:(n_orders - 1)]
+    last_one     <- ord_nums_graph[n_orders]
+    # Join the first n-1 with commas
+    joined       <- str_c(all_but_last, collapse = ", ")
+    # Add the final one with ", and ..."
+    orders_str   <- str_c(joined, ", and ", last_one)
+    orders_prefix <- "Orders "
+  }
+  
+  # E.g. "Orders 392833, 487984, and 555444, SAT score 1020 - 1150"
+  graph_title <- str_c(orders_prefix, orders_str, ", ", row_i$test_range)
   
   for (g in c("race", "firstgen")) {
     
@@ -81,6 +103,7 @@ for (i in seq_len(nrow(orders_df))) {
       metro_name_graph,
       g,
       "plot_order",
+      # Keep underscores in the object name
[…]

### 2025-02-27 index.qmd
added aian

@@
 <p class="quote" style="font-size: 0.85em;">Recommend colleges target Geomarkets, schools targeted by peer colleges</p>
-<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
+<p style="display:none;">On occasion, senior spokespersons for the profession worry that students outside the main [Geo]market areas remain forgotten and hence, unchallenged.
 Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
@@
 
+############################################################################
+# 5) AI/AN results loop
+############################################################################
+
+writeLines("## AI/AN students")
+
+# The five metros (with 'nyny' last) and the single Native American order
+native_metros <- c("chicago", "dallas", "houston", "los_angeles", "nyny")
+native_order  <- "487927"
+
+# We want exactly 3 slides per metro, in this order:
+var_rc_pairs <- list(
+  c("firstgen", "row"),
+  c("firstgen", "col"),
+  c("race",     "row")
+)
+
+for (m in native_metros) {
+
+  # Determine the heading text for each metro:
+  # If m == 'nyny', use "New York City".
+  # Otherwise, convert underscores -> spaces, then "Title Case + area".
+  heading_text <- if (m == "nyny") {
+    "New York City"
+  } else {
+    # Normal logic
+    metro_no_underscores <- gsub("_", " ", m, fixed = TRUE)
+    title_case           <- tools::toTitleCase(metro_no_underscores)
+    paste0(title_case, " area")
+  }
+
+  # For each metro, we produce 3 slides
+  for (i in seq_along(var_rc_pairs)) {
+
+    # If this is the 1st slide => show in TOC (no .hide).
+    # Otherwise => hide from TOC with {.hide}.
+    if (i == 1) {
+      writeLines(str_c("### ", heading_text))
+    } else {
+      writeLines(str_c("### ", heading_text, " {.hide}"))
+    }
+
+    # Extract the variable (firstgen/race) and approach (row/col)
+    g  <- var_rc_pairs[[i]][1]
+    rc <- var_rc_pairs[[i]][2]
+
+    # Build the figure name: e.g. "race_row_plot_487927"
+    graph_type_native <- str_c(g, "_", rc, "_plot_", native_order)
+
+    # Call your helper function; it prints "#### <title from _title.txt>"
+    # and displays the figure + footnotes.
+    insert_figure(
+      rq         = "rq2",
+      metro      = m,
+      graph_type = graph_type_native
+    )
+  }
+}
+
 ```

### 2025-02-27 text/front_end_v2.qmd
added aian

--- 2025-01-15 text/front_end_v2.qmd
@@
 
+**ABSTRACT**
+
+*The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Model." Geomarkets carve states and metropolitan areas into smaller geographic units, meant to define local recruiting markets.
+The Market Segment Model predicts how student demand for a particular college varies by Geomarket, based on the socioeconomic characteristics of households.
+Geomarkets became an input for two College Board products that help colleges recruit students.
+First, the Enrollment Planning Service (EPS) software recommends specific Geomarkets and high schools from which colleges should recruit.
+Second, the Student Search Service sells the contact information of prospective students -- referred to as "student lists" -- and colleges can filter by Geomarket to determine which prospect profiles they purchase.
+We draw from scholarship on quantification, particularly the discussions of correlation and homophily by @RN4975, to conceptualize how recruiting products incorporate Geomarkets.
+We address two research questions: What is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
+How does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
+We answer RQ1 by analyzing Census data from 1980, 2000, and 2020.
+We answer RQ2 using data on student lists purchased by public universities, which we collected by issuing public records requests.
+We utilize a quantitative case study design.
+Metropolitan areas are cases.
+Analyses consist of descriptive statistics and interactive maps.
+
 # Introduction

### 2025-02-27 scripts/rq2_metro_orders.R

@@ -21,9 +21,6 @@ library(readr)
 # Note: We'll write to "metro_orders.csv" in the working directory.
 # -------------------------------------------------------------------------
 
-library(dplyr)
-library(readr)
-
 # 1) Your revised all_orders list (example below):
 all_orders <- list(
   # CHICAGO
@@ -253,9 +250,58 @@ all_orders <- list(
     "329702",
     "PSAT score 1010 - 1520",
     "Order number 329702 (ordered 8/31/2017): HS class 2019; State = TX; PSAT 1010-1520"
+  ),
+  ##### NATIVE AMERICAN ORDERS
+  # NYC
+  nyny = c(
+    "nyny",
+    "nyny_metro_eps_codes",
+    "487927",
+    "SAT score 1040 - 1600",
+    "Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+"
+  ),
+  # CHICAGO
+  chicago = c(
+    "chicago",
+    "chi_eps_codes",
+    "487927",
+    "SAT score 1040 - 1600",
+    "Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+"
+  ),
+  # HOUSTON
+  houston = c(
+    "houston",
+    "houston_eps_codes",
+    "487927",
+    "SAT score 1040 - 1600",
+    "Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+"
+  ),
+  # LOS ANGELES
+  los_angeles = c(
+    "los_angeles",
+    "los_angeles_eps_codes",
+    "487927",
+    "SAT score 1040 - 1600",
+    "Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+"
+  ),
+  # DALLAS
+  dallas = c(
+    "dallas",
+    "dallas_eps_codes",
+    "487927",
+    "SAT score 1040 - 1600",
+    "Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+"
   )
 )
 
+
[…]

### 2025-02-27 scripts/rq2_tables_graphs_create_functions.R

@@ -447,7 +447,7 @@ create_sim_eps_race_firstgen_table <- function(data, ord_nums, eps_codes, exclud
 # philly metro area
 # order 448922: PSAT 1070 - 1180; order 448427: PSAT 1190 - 1260; order 448440: PSAT 1270 - 1520
 
-create_sim_eps_race_firstgen_table(data = lists_orders_zip_hs_df_sf, ord_nums = c('487984'), eps_codes = chi_eps_codes) %>% print(n=70) # 
+#create_sim_eps_race_firstgen_table(data = lists_orders_zip_hs_df_sf, ord_nums = c('487984'), eps_codes = chi_eps_codes) %>% print(n=70) # 
 #create_sim_eps_race_firstgen_table(data = lists_orders_zip_hs_df_sf, ord_nums = c('487984'), eps_codes = chi_eps_codes, exclude_race = '') %>% print(n=70) # 
 #create_sim_eps_race_firstgen_table(data = lists_orders_zip_hs_df_sf, ord_nums = c('487984'), eps_codes = chi_eps_codes, exclude_race = c(1,8,12)) %>% print(n=70) # 
 #lists_orders_zip_hs_df_sf %>% as_tibble() %>% count(stu_race_cb)
@@ -1116,7 +1116,7 @@ create_race_by_firstgen_graph <- function(
 
 #create_race_by_firstgen_graph(data_graph,ord_nums_graph,eps_codes_graph,metro_name = '',exclude_race_graph = c(1,8,12),title_suf = '')
   
-create_race_by_firstgen_graph(lists_orders_zip_hs_df_sf, c('487984'), chi_eps_codes, 'chicago', exclude_race = c(1,8,12), title_suf = ', SAT score 1020 - 1150')
+#create_race_by_firstgen_graph(lists_orders_zip_hs_df_sf, c('487984'), chi_eps_codes, 'chicago', exclude_race = c(1,8,12), title_suf = ', SAT score 1020 - 1150')
 
 #stu_race_cb                                 n
 #1  0 [no response]                       135083

### 2025-02-27 scripts/rq2_tables_graphs_run_functions.R

@@ -126,7 +126,13 @@ for (i in seq_len(nrow(orders_df))) {
   }
 }
 
+# -------------------------------------------------------------------------
 # Part #2) Generating row/col percent figures (patchwork logic)
+# -------------------------------------------------------------------------
+
+# List the special AI/AN order(s) that should not be combined with others
+special_orders <- c("487927")
+
 unique_metros <- unique(orders_df$metro)
 
 for (m in unique_metros) {
@@ -143,6 +149,7 @@ for (m in unique_metros) {
   
   for (g in c("race", "firstgen")) {
     
+    # For non‐special orders, we build a default prefix
     g_prefix <- if (g == "race") "Racial/ethnic" else "First-generation status"
     
     for (rc in c("row", "col")) {
@@ -152,19 +159,18 @@ for (m in unique_metros) {
       plot_name_base <- str_c("rq2_", m, "_", g, "_", rc, "_plot")
       
       # ---------------------------------------------------------
-      # Build figure title (conditional on "Bay Area")
+      # Build the *default* figure title for non‐special orders
       # ---------------------------------------------------------
       if (rc == "row") {
         # e.g. "Racial/ethnic composition of purchased student profiles by Geomarket, Los Angeles area"
         if (friendly_metro_name == "Bay Area") {
-          figure_title <- str_c(
+          default_figure_title <- str_c(
             g_prefix,
             " composition of purchased student profiles by Geomarket, ",
             friendly_metro_name
-            # no " area" appended
           )
         } else {
-          figure_title <- str_c(
+          default_figure_title <- str_c(
             g_prefix,
             " composition of purchased student profiles by Geomarket, ",
             friendly_metro_name,
@@ -174,15 +180,14 @@ for (m in unique_metros) {
       } else {
         # e.g. "Geomarket contribution to purchased student profiles by Racial/ethnic group, Los Angeles area"
         if (friendly_metro_name == "Bay Area") {
-          figure_title <- str_c(
+          default_figure_title <- str_c(
             "Geomarket contribution to purchased student profiles by ",
             g_prefix,
             " group, ",
             friendly_metro_name
-            # no " area" appended
           )
         } else {
-          figure_title <- str_c(
+          default_figure_title <- str_c(
             "Geomarket contribution to purchased student profiles by ",
             g_prefix,
             " group, ",
@@ -193,17 +198,16 @@ for (m in unique_metros) {
       }
       
[…]

### 2025-03-01 index.qmd
ugh. AIAN loop messed

@@
 
-writeLines("## AI/AN students")
+# Make AI/AN a top-level heading (level 1)
+writeLines("# AI/AN students")
 
@@
     if (i == 1) {
-      writeLines(str_c("### ", heading_text))
+      writeLines(str_c("## ", heading_text))
     } else {
-      writeLines(str_c("### ", heading_text, " {.hide}"))
+      writeLines(str_c("## ", heading_text, " {.hide}"))
     }
@@
 
-    # Call your helper function; it prints "#### <title from _title.txt>"
+    # Call your helper function; it prints a heading/caption
     # and displays the figure + footnotes.

### 2025-03-01 scripts/rq2_metro_orders.R

@@ -251,6 +251,22 @@ all_orders <- list(
     "PSAT score 1010 - 1520",
     "Order number 329702 (ordered 8/31/2017): HS class 2019; State = TX; PSAT 1010-1520"
   ),
+  houston = c(
+    "houston",
+    "houston_eps_codes",
+    "404873",
+    "PSAT score 1120 - 1300",
+    "LA LA"
+  ),
+  # don't have firstgen status for these students
+  #houston = c(
+  #  "houston",
+  #  "houston_eps_codes",
+  #  "448007_548238",
+  #  "PSAT score 1470 - 1520",
+  #  "Order number 448007 (ordered 1/8/2019): HS class 2020; State = TX; PSAT 1470 - 1520; GPA B- to A+"
+  #),
+  # Order number 548238 (ordered 1/7/2020): HS class 2021; Selected states; 1470 - 1520; GPA B- to A+
   ##### NATIVE AMERICAN ORDERS
   # NYC
   nyny = c(
@@ -478,3 +494,10 @@ write_csv(orders_df, file.path(scripts_dir,"metro_orders.csv"))
 # Don’t have tx19,tx20,tx21,tx22
 # 546978 (ordered 1/6/2020); 2021 HS grad class; PSAT 1270 - 1520 [N]
 # Don’t have tx19,tx20,tx21,tx22
+
+lists_orders_zip_hs_df %>% filter(hs_state_code == 'TX') %>% 
+  mutate(univ_ord = str_c(univ_name,univ_id,ord_num,sep = '_')) %>% count(univ_ord) %>% arrange(desc(n)) %>% print(n=50) 
+
+
+lists_orders_zip_hs_df %>% filter(hs_state_code == 'TX' & univ_id =='228723') %>% 
+  mutate(univ_ord = str_c(univ_name,univ_id,ord_num,sep = '_')) %>% count(univ_ord) %>% arrange(desc(n)) %>% print(n=50) 

### 2025-03-03 index.qmd
tweaking presentation and AI/AN slides

--- 2025-02-27 index.qmd
@@
 
-# 1) Define the `insert_figure` helper function
+# Revised insert_figure() that does NOT print a heading
 insert_figure <- function(rq, metro, graph_type, base_path = graphs_dir) {
@@
   # We'll use 'metro' directly since it's already underscored (e.g. "los_angeles").
-  fig_file  <- file.path(base_path, rq, str_c(rq, "_", metro, "_", graph_type, ".png"))
-  cap_file  <- file.path(base_path, rq, str_c(rq, "_", metro, "_", graph_type, "_title.txt"))
-  note_file <- file.path(base_path, rq, str_c(rq, "_", metro, "_", graph_type, "_note.txt"))
+  fig_file  <- file.path(base_path, rq, stringr::str_c(rq, "_", metro, "_", graph_type, ".png"))
+  cap_file  <- file.path(base_path, rq, stringr::str_c(rq, "_", metro, "_", graph_type, "_title.txt"))
+  note_file <- file.path(base_path, rq, stringr::str_c(rq, "_", metro, "_", graph_type, "_note.txt"))
 
-  # Print a heading / caption
+  # Print a heading/caption
   caption_text <- if (file.exists(cap_file)) readLines(cap_file, warn = FALSE) else "No Title Found"
@@
   if (file.exists(fig_file)) {
-    writeLines(str_c("![](", fig_file, ")\n"))
+    writeLines(stringr::str_c("![](", fig_file, ")\n"))
   } else {
@@
     writeLines(
-      str_c(
+      stringr::str_c(
         "<div class=\"footnote\">",
@@
 
+      # -- HERE is the new line to remove "487927" --
+      orders_for_m <- orders_for_m[ orders_for_m != "487927" ]
+
       # We'll do 2 group types: race & firstgen
@@
 
+# A special version insert_figure_aian for AI/AN that does NOT print a heading
+insert_figure_aian <- function(rq, metro, graph_type, base_path = graphs_dir) {
+
+  fig_file  <- file.path(base_path, rq, paste0(rq, "_", metro, "_", graph_type, ".png"))
+  cap_file  <- file.path(base_path, rq, paste0(rq, "_", metro, "_", graph_type, "_title.txt"))
+  note_file <- file.path(base_path, rq, paste0(rq, "_", metro, "_", graph_type, "_note.txt"))
+
+  # Instead of printing a heading, let the AI/AN loop do it (#### subtitle).
+  # So we skip caption_text or print it differently if you want.
+
+  if (file.exists(fig_file)) {
+    writeLines(paste0("![](", fig_file, ")\n"))
+  } else {
+    writeLines(paste("*(Missing figure file:", fig_file, ")*\n"))
+  }
+
+  # Print footnotes (if the .txt file exists)
+  if (file.exists(note_file)) {
+    notes_txt <- readLines(note_file, warn = FALSE)
+    writeLines(
+      paste0(
+        "<div class=\"footnote\">",
+        paste0(notes_txt, collapse = "</br>"),
+        "</div>"
+      )
+    )
+  }
+}
+
 ############################################################################
@@
 
-writeLines("## AI/AN students")
+# (Level-1) "AI/AN students"
+writeLines("# AI/AN students\n")
 
@@
 
-  # Determine the heading text for each metro:
-  # If m == 'nyny', use "New York City".
-  # Otherwise, convert underscores -> spaces, then "Title Case + area".
+  # (Level-2) heading for each metro (not a slide at slide-level=3)
   heading_text <- if (m == "nyny") {
@@
   } else {
-    # Normal logic
-    metro_no_underscores <- gsub("_", " ", m, fixed = TRUE)
-    title_case           <- tools::toTitleCase(metro_no_underscores)
-    paste0(title_case, " area")
+    m_no_underscores <- gsub("_", " ", m, fixed = TRUE)
+    m_title          <- tools::toTitleCase(m_no_underscores)
+    paste0(m_title, " area")
   }
+  writeLines(paste0("## ", heading_text, "\n"))
 
-  # For each metro, we produce 3 slides
+  # For each var/rc pair, create a new slide at level-3 (always hidden in TOC)
   for (i in seq_along(var_rc_pairs)) {
 
-    # If this is the 1st slide => show in TOC (no .hide).
-    # Otherwise => hide from TOC with {.hide}.
-    if (i == 1) {
-      writeLines(str_c("### ", heading_text))
-    } else {
-      writeLines(str_c("### ", heading_text, " {.hide}"))
-    }
-
-    # Extract the variable (firstgen/race) and approach (row/col)
     g  <- var_rc_pairs[[i]][1]
@@
 
-    # Build the figure name: e.g. "race_row_plot_487927"
-    graph_type_native <- str_c(g, "_", rc, "_plot_", native_order)
+    # Always hidden from TOC => add "{.hide}" every time
+    writeLines(paste0("### ", heading_text, " (AI/AN) {.hide}\n"))
 
-    # Call your helper function; it prints "#### <title from _title.txt>"
-    # and displays the figure + footnotes.
-    insert_figure(
+    # (Level-4) subtitle on the same slide
+    subtitle <- if (g == "firstgen") {
+      "First-generation status composition of purchased AI/AN student profiles by geomarket"
+    } else if (g == "race") {
+      "Racial composition of purchased AI/AN student profiles by geomarket"
+    } else {
+      "Other composition of purchased AI/AN student profiles"
+    }
+    writeLines(paste0("#### ", subtitle, "\n"))
+
+    # Build figure name and call your "no-heading" figure function
+    graph_type <- paste0(g, "_", rc, "_plot_", native_order)
+    insert_figure_aian(
       rq         = "rq2",
       metro      = m,
-      graph_type = graph_type_native
+      graph_type = graph_type
     )
@@
 }
-
 ```

### 2025-03-03 scripts/rq1_tables_graphs.R

@@ -233,15 +233,15 @@ create_eps_graph <- function(eps_codes,
     stop("Please provide a valid `eps_codes` vector.")
   }
   
-  # create string for pathname
-  # If `prefix` isn't provided, fall back to the old logic:
+  # 1) Determine prefix for filenames
   if (is.null(prefix)) {
+    # Fall back to old logic: use the variable name for eps_codes
     file_prefix <- str_replace(deparse(substitute(eps_codes)), '_eps_codes', '')
   } else {
     file_prefix <- prefix
   }
   
-  # Main table creation pipeline
+  # 2) Build the main table to plot
   eps_table <- allyr_anal_tract_sf %>% 
     as.data.frame() %>%
     filter(eps %in% eps_codes) %>%
@@ -290,15 +290,15 @@ create_eps_graph <- function(eps_codes,
     mutate(eps_codename = fct_rev(factor(eps_codename))) %>%
     mutate(year = factor(year, levels = c(2020, 2000, 1980)))
   
-  # Debugging: Print intermediate table if debug is TRUE
+  # If debug=TRUE, show intermediate table
   if (debug) {
-    print(paste("Using statistic:", stat))
+    message("Using statistic: ", stat)
     print(head(eps_table))
   }
   
-  # Filter and create plots based on graph type
+  # 3) Filter table & build plot
   if (graph_type == "race") {
-    
+    # Race/ethnicity approach
     eps_table <- eps_table %>%
       filter(
         statistic == stat,
@@ -320,7 +320,7 @@ create_eps_graph <- function(eps_codes,
       ) +
       scale_y_continuous(labels = scales::percent_format()) +
       facet_wrap(~ year, ncol = 1, scales = "free_y") +
-      theme_custom + 
+      theme_custom +
       theme(
         legend.text  = element_text(size = 14),
         legend.title = element_text(size = 14),
@@ -333,8 +333,8 @@ create_eps_graph <- function(eps_codes,
       plot <- plot + scale_fill_manual(values = fill_palette)
     }
     
-  } else if (graph_type == "ses") {
-    
+  } else {
+    # SES approach
     eps_table <- eps_table %>%
       filter(
         statistic == stat,
@@ -359,8 +359,8 @@ create_eps_graph <- function(eps_codes,
       theme(
         strip.text.x = element_text(size = 12, face = "bold"),
[…]

### 2025-03-06 text/front_end_v3.qmd
working through front end. again.

--- 2025-02-27 text/front_end_v2.qmd
@@
 
+<!-- la la -->
+Enrollment management is simultaneously a profession, an administrative structure, and an industry.
 As a profession, enrollment management (EM) integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p.
@@
 As an administrative structure, the EM office typically controls the activities of admissions, financial aid, and recruiting [@RN2406].
+<!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
+-->
+
 @fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
@@
 
-A growing college access literature in sociology analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
+A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
 @RN4758 conceptualize recruiting behavior as an indicator of college enrollment priorities.
 Ethnographies by @RN3519 and @RN4407 identify connections between private school guidance counselors and college admissions officers as a mechanism for social reproduction.
+Recruiting visits to high schools are a means of maintaining ties with guidance counselors at feeder schools and establishing relationships with prospective students [@RN3519; @RN4895].
 Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly white public schools [@RN4758; @RN4759; @RN5023].
@@
 @cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.
-Consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015], this literature reveals a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches segmented by class and race, a process consistent with the notion of that higher education is mechanism of social reproduction [@RN2377; @RN4873].
+<!--  Consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015], this literature reveals a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches segmented by class and race, a process consistent with the notion of that higher education is mechanism of social reproduction [@RN2377; @RN4873].
+-->
 
-By focusing on the behaviors of individual colleges, scholarship on college access implicitly states that recruiting -- and enrollment management more broadly -- is something done by individual colleges.
-This statement raises the question, why does recruiting behavior look so similar across colleges of a particular "type." Institutional theory defines the organizational field as "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services and products" [@RN527, p.
-143].
-The broader enrollment management (EM) industry consists of university personnel (e.g., admissions counselors, VP for enrollment management), professional associations (e.g., National Association for College Admission Counseling), and third-party servicers that produce market research, software, and consulting services (e.g., College Board, ACT, EAB, Slate, PowerSchool).
+By focusing on the behaviors of individual colleges, scholarship on EM implicitly states that recruiting is something done by individual colleges. 
+In addition to university personnel  (e.g., admissions counselors, VP for enrollment management), the EM industry incorporates external stakeholders in the organizational field, including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
 We argue college access is structured by third-party servicers and products that interact with direct-providers (colleges).
-Enrollment management is often described as an opaque industry because obtaining concrete information about EM practices is difficult.
-This is why the title of @RN5024 is *Lifting the Veil on Enrollment Management*.
-Although sociologists have hinted at the ways enrollment management contributes to inequality in colege access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
+Although sociologists have hinted at the ways enrollment management contributes to inequality in college access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
+<!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
+-->
 <!--  
@@
 EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-More recently, the EM consulting business model is based on software-as-service products (e.g., for example, EAB's [Enroll 360](https://eab.com/solutions/enroll360/) product) [@list_biz].
 As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
 The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-<!-- In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
--->
+In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
+
+Although the enrollment management industry is increasingly characterized by software-as-service products sold by private equity backed firms (e.g., for example, EAB's [Enroll 360](https://eab.com/solutions/enroll360/) product) [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
+The Student Search Service, created in 1972, became the ubiquitous means of identifying prospects BLAH BLAH.
+Second, EPS, launcched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
+Both products are applications of quantification.
+Student Search Service includes or excludes prospects based on selected attributes while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
+THEREFORE WE REVIEW SCHOLARSHIP ON QUANTIFICATION.
+
+<!--
 
@@
 @RN4986 [p.
-17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].[^1] In recruiting, market research informs decisions about which prospective students to target with recruiting interventions.<!-- THIS IDEA HAS TO GO SOMEWHERE: IN THE HOOK AT END OF THIS SECTION.
-These recruiting interventions intervene on student demand [@RN4324; @cottom2017lower], structuring college access pathways in ways that could amplify or diminish historic inequality in college access.
---> How to acquire market research is another make-or-buy decision.
+17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].[^1] In recruiting, market research informs decisions about which prospective students to target with recruiting interventions.
+How to acquire market research is another make-or-buy decision.
 The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals.
@@
 69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."
+-->
 
@@
 402).
-They introduce reacivity, discipline, and authority as three interrelated themes that describe the effects of quantification.
-Reactivity is the idea that salient quantitative measures cause people and organizations to change their behavior.
-Discipline is the idea that quantification causes actors to react in a particular way.
-Authority is the idea that quantification changes decision-making power, often weakening the discretion of local decision-makers.
-
+Reacivity, discipline, and authority are three interrelated themes that describe the effects of quantification.
+*Reactivity* is the idea that salient quantitative measures cause people and organizations to change their behavior.
+Quantification *disciplines* actors to react in particular ways.
+Quantification changes decision-making power *authority*, often weakening the discretion of local decision-makers.
 Scholarship about U.S.
-News & World Report (USNWR) Law School rankings demonstrate the reactivity, discipline, and authority effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
+News & World Report (USNWR) Law School rankings demonstrate the effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
 Law school rankings were developed as a means of informing prospective students about the relative quality of different programs.
-Speaking to *reactivity* effects, rankings had the intended effect of affecting where students apply and enroll \[CITE\].
+Rankings affect (reactivity) where students apply and enroll \[CITE\].
 However, rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
@@
 
-The analysis of UK League Tables by @RN5005 show how quantification can be a mechanism of social reproduction.
-In 1992, the UK government began publishing school "league tables," which ranked schools based on student performance on high-stakes national exams taken at age 16.
+The analysis of UK school "league tables" by @RN5005 shows how quantification can be a mechanism of social reproduction.
+In 1992, the UK government began publishing school league tables, which ranked schools based on student performance on national exams taken at age 16.
 League tables facilitate making evaluative comparisons between schools regardless of geographic proximity.
@@
 **Correlation**.
-<!-- In the literature on quantification from critical data studies, --> The discussions of correlation and homophily by @RN4975 introduce ideas salient to the analysis the Market Segment Model and College Board Geomarkets.
-Correlation measures the extent to which two or more variables move together.<!--The Pearson Correlation Coefficient was developed by the 19th Century statisticians and "biometric eugenicists" Francis Galton and his student Karl Pearson [@RN4975].
---> @RN4975 [p.
-50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'" Predictive analytics are based on correlation and are developed in two steps.
+The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4849; @RN4772; ].
+In particular, the discussions of correlation (chapter 1) and homophily (chapter 2) by @RN4975 introduce ideas salient to the analysis the Market Segment Model and College Board Geomarkets.
+Correlation measures the extent to which two or more variables move together.
+Predictive analytics are based on correlation and are developed in two steps.
 First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
 Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and to assign levels of risk to each case.
+@RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
 These models predict outcomes based on correlations without requiring knowledge about underlying causal relationships.
-@RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
-
-Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated to a variable of interest.
+@RN4975 [p.
+50] writes that "correlation grounds big data's so-called revolutionary potential.
+As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"^[Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
 For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
 @RN4849 [p.
-146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
+146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?']
 
-Several studies show that using correlational models make predictions reproduces structural inequality [@RN4794].
+A bevy of studies show that predictions based on correlations reproduce structural inequality [for a review see @RN4794].
 The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
-These correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
+The observed correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
 Reviewing scholarship about algorithms, @RN4794 [p.
 224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^2]
+This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^2] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
+They then recommended that selective colleges recruit from localities that have a high share of affluent, college-educated households.
 
@@
 **Homophily**.
-@RN4975 argues that homophily is central to models that use correlation to predict future behavior.
-Homophily means that actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together." Homophily is a core assumption for network science [@mcpherson2001birds], in which actors -- in this manuscript, high school students -- are linked to one another through direct and indirect network ties -- submitting SAT scores to the same college.
-Because network science models often draws from rational choice theory, they assume that homophily is the result of voluntary action by individuals, thereby "eras\[ing\] historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
-95] that underlie behavior consistent with homophily.
+Homophily is the idea actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together" [@mcpherson2001birds].
+Homophily is a core concept of network science (writ large), in which actors (nodes) are connected to one another directly and indirectly via network ties (edges).
+As discussed below, @RN4982[chapter 4] *The company We Keep: Colleges and Their Competition* conducts a network analysis based on 1980 SAT score sending behavior whereby two colleges are defined as competitors if a large number of students send SAT scores to both colleges.
+Because network science models often draws from rational choice theory, @RN4975 argues, it assumes that homophily is the result of voluntary action by individuals.
+
 @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-In commercial social networks (e.g., Faceboo, X, TikTok), homophily is more than an assumption; homophily is programmed into the algorithms that create connections between users.[^3]
+Like correlational models discussed above, network science models "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
+95] that cause behavior consistent with homophily.
+<!--"Homophily assumes and creates segregation" [@RN4975, p.
+95] writes.
+"It transforms individuals into 'neighbors' who naturally want to live with people 'like them'." --> In commercial social networks, @RN4975 argues, homopphily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
+Thus, "social networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@RN4975, p.
+YY].
+We observe similar processes in college access because higher education is an industry in which providers are evaluated based on the characteristics of their customers [@RN1549; @RN1237].
+Based on a snapshot of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
+In turn, @RN4982 reasons that colleges should target the Geomarkets and high schools targeted by peer colleges.
+This logic is subsequently programmed into EPS software that recommends which Geomarkets and high schools to recruit from.
+Thus, homophily observed on the demand side -- which is a function of historic structural inequality –- is programmed into the supply-side.
 
@@
 
-Within the market research profession, homophily is central to geodemography and the creation of market segments.
+Homophily is also central to products that categorize customers.
 Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
@@
 Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
+We observe similar processes in our case study, whereby @RN4982 categorizes students into four market segments -- local, in-state, regional, and national -- based on their score sending behavior and then evaluates the attractiveness of each Geomarket based on how many "regional" and "national" students live there.
 <!--
@@
 Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
--->
-
-[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
 
@@
 Law school rankings can be conceived as a business-to-consumer product because the goal is to inform consumers (prospective law school students) and consumers pay for the product.
-By contrast, market research helps businesses identify/target customers and can be conceived as a business-to-business (B2B) product because businesses pay for market research.
-To our knowledge, scholarship on quantification in education has not examined how third-party market research classifies consumers.
-Additionally, scholarship on college access has failed to analyze third-party products that colleges use to identify and target prospective students.
+
+-->
+
+[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
+
+**Contribution**.
+As @RN3519 demonstrates in the chapter *Numbers*, enrollment management is fundamentally concerned with quantification.
+In the case of college rankings, sociology made visible how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges who are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
+Beyond rankings, sociological analyses of enrollment management focus on the behavior of colleges and their agents, yielding the implicit assumption that inequalities creates by enrollment management are a function of individual college behavior.
+We argue that third-party vendors and products structure the enrollment management behavior of colleges.
+These products take a snapshot of student demand -- without considering structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
+Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
+These products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociologists.
+
 This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
@@
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
-
-Scholarship on quantification suggests the mechanisms by which market research reproduces structural inequality [CITE CHUN; SOME SOCIOLOGY].
-Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
 Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification.
@@
 
-**Creating the Market Segment Model.** In 1978, Robert Zemsky, a University of Pennsylvania professor, was asked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
+**Creating Geomarkets and the Market Segment Model.** In 1978, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
+x].^[Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.] Zemsky reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential" [@RN4982, p.
 x].
-To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
-The researchers reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest demographic unit used in these analyses was the state' [@RN4982, p.
-x].
-In 1979, College Board began providing data and funding for what became the *Comprehensive Undergraduate Enrollment Planning Project (CUEPP)*.
+
+The geographic concepts inscribed in the Market Segment Model were based on conversations with college admissions officers.
+@RN4982 \[p.
+4\] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." Quoting an admissions officer, @RN4982 [p.
+11] write, "'There are only three kinds of colege-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, [@RN4982] created three geographic [@RN4982, p.
+11], region, state, and community.
+The three regions were New England, Middle States, and the South.
+Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets." These enrollment markets, later called Geomarkets, were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
+@RN4982[pp.
+11-12] briefly describes the creation of Geomarket borders:
+
+> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
+Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
+In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."
+
+The goal of the Market Segment Model was to quantify the knowledge that admissions counselors have about student demand.
+In 1979, College Board began providing funding and data on the score-sending behavior of SAT test-takers.
 @RN4982 [p.
@@
 
-@RN4982 \[p.
-4\] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." Quoting an admissions officer, @RN4982 [p.
-11] write, "'There are only three kinds of colege-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, [@RN4982] created "three types of boundaries" [@RN4982, p.
-11], region, state, and community.
-The three regions were New England, Middle States, and the South.
-Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets."
-
-These enrollment markets, later called "Geomarkets" by the College Board Enrollment Planning Service were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors [@RN4982].
-We know little about how geomarkets were created.
-@RN4982 write that,
-
-> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
-Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
-In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."
-
-Having defined these geographic boundaries, high school students were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
-Each institution a student sends SAT scores to can be defined as "local" (institution is in the same local market that the student lives in), "in-state" (same state but different local market), "regional" (same region but different state), or "national." In turn, a "local" student submits more SAT scores to local institutions than they do to in-state, regional, or national institutions.
+SAT test-takers were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
+For a given student, each college that receives a score from the student can be defined as "local" (college located in the the same Geomarket as the student), "in-state" (same state but different Geomarket), "regional" (same region but different state), or "national" (different region).
+In turn, a test-taker is categorized in the "local" market segment if they submit more SAT scores to local institutions than they do to in-state, regional, or national institutions.
 An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
@@
 
-The two primary outputs of the market segment model are the (1) Market Segment Profile and (2) the Institutional Profile.
+The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
 Appendix A describes these outputs in more detail.
-The Market Segment Profile provides information about the number and characteristics of students in each market segment.
-This information is produced separately for each of the 143 Geomarkets.
-For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
-The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
-A university could obtain this for their own institution or for a competitor.
-Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
-This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
+Both outputs are created separately for each Geomarket.
+For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than $35,000, percent with both parents having a BA, etc.^[For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+A university could obtain this for their own institution or for a competitor.^[Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.]
+
+Taken together, the Market Segment Model show a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+These outputs became the basis for the Enrollment Planning Service (EPS) software.
+
+**Correlations**.
+@RN4982[Chapter 3], *A Sense of Place: Students, Families, and Communities*, identifies the student characteristics correlated with being in the local, in-state, regional, or national market segment.
+The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
+These four variables "reflect the basic social patterns of the nation.
+It would have been surprising if these were not the four social variables that best explained the patterns of college choice" [@RN4982, p.
+33].<!--
+and "The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice" (pp.
+34-35).
+\[QUOTES REDUNDANT\]
+-->
+
+The primary takeaway of Chapter 3 is that student demand for higher education is mostly a function of social origin.
+For example, @RN4982[p.
+33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
+33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^6] @RN4982[p.
+42] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
+Therefore, @RN4982[p.
+44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:
+
+> On occasion, senior spokespersons for the profession worry that students outside the main [Geo]market areas remain forgotten and hence, unchallenged.
+Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
+The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
+
+[^6]: The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
+Additionally, analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
+However, prior research shows that SAT scores are substantially a function of social origin [@RN1273].
+
+**Homophily**.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.
+This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
+Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
+These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.[^8]
+
+Based on these analyses, @RN4982 [p.
+46] state that homophily describes the nature of competition between colleges: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
+Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
+The authors argue that like-colleges compete for like-students as defined by socioeconomic characteristics.
+@RN4982 [p.
+72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+
+> Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
+The layering of collegiate competition is primarily a socioeconomic layering.
+The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
+Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
+
+[^8]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
+54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
+For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
+
+The discussion of competition by @RN4982 exemplifies the concerns about correlation and homophily described by [@RN4975].
+A correlational analysis of 1980 SAT score-sending patterns finds competition between colleges is defined by socioeconomic homophily.
+This homophily is presented as a naturally occurring phenomenon.
+Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer-coleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
+In itself, @RN4982's Market Segment Model is merely a social science depiction of the structure of student demand, one that does not consider historic, structural inequalities that cause observed patterns.
+However, by inscribing the Market Segment Model and Geomarkets into the EPS software, College Board amplified  structural inequalities that contributed to homophily observed in 1980 SAT score-sending beahivor.
+In this way, commodification of the Market Segment Model and Geomarkets "engineers homophily" [@RN4975].
 
 **Enrollment Planning Service**.
-In 1984, College Board created the Enrollment Planning Service (EPS) based on the Market Segment Model [@RN4987; @RN4980].
-EPS was a simple software-as-service platform that informed college decisions about recruiting by essentially recreating the analyses of @RN4982.
-Institutions purchasing EPS could obtain the Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors -- for each local market.
+In 1984, College Board created the Enrollment Planning Service (EPS), based on the Market Segment Model [@RN4987; @RN4980].
+EPS was an early software-as-service platform that recreated the analyses of @RN4982.
+For each Geomarket, colleges could obtain Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors.
+Based on conversations with enrollment management professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket.
+Therefore, colleges used EPS software to decide which Geomarkets to recruit from and which high schools to visit within targeted Geomarkets.
 Typical marketing material describes EPS as,
 
-> The marketing software that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
+> The marketing software that pinpoints the schools and Geomarkets where your best prospects are most likely to be found.
 With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
@@
 @RN4988 shows the contemporary Geomarkets.
-Interestingly, the Geomarkets for New England, Middle States, and the South remain nearly identical to those developed by @RN4982.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^5].
-Furthermore, this material encourages colleges to, first, identify which Geomarkets they will target.
-For example, Software documentation by @RN5017 states that
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^5] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
 
@@
 
-> EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful.
+@RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
+Market research suggests that EPS was highly salient.
+@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
 
-**Correlations**.
-Chapter 3 of @RN4982 identifies the student characteristics associated with being in the local, in-state, regional, or national market segment.
-The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
-[@RN4982] state that these four variables "reflect the basic social patterns of the nation.
-It would have been surprising if these were not the four social variables that best explained the patterns of college choice" (p.
-33) and "The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice" (pp.
-34-35).
-\[QUOTES REDUNDANT\]
+EPS software may discipline colleges to approach recruiting in a manner consistent with the Market Segment Model.
+Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
+EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.^[Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."].
+Second, EPS users decide which high schools they will visit within selected Geomarkets.
+Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
 
-The primary takeaway of Chapter 3 is that student demand for higher education is mostly a function of social origin.
-For example, @RN4982 \[p.
-33\] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^6]
+START HERE FRIDAY!!!!!
 
-[^6]: The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
-Additionally, analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
-However, prior research shows that SAT scores are substantially a function of social origin [@RN1273].
+EPS may also weaken the authority of local decision-makers.
+@RN4982 sought to develop a concrete, data-driving framework -- The Market Segment Model -- replicated the aggregate knowledge of local admissions officers.
+Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
+The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
+In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
+However, use of EPS contributed to most colleges visiting the same sets of affluent, high-achieving high schools.[^9]
 
-However, geomarkets differ in the relative abundance of students with particular socioeconomic characteristics and this has practical implications for recruiting.
-@RN4982 recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics and student market segment.
-
-**Homophily**.
-Chapter 4 of @RN4982 -- "The Company We Keep: Colleges and Their Competition" -- shows that homophily is a core assumption of the Market Segment Model.
-The authors conduct a social network analysis that defines two institutions as being in competition with one another if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.[^7] @RN4982 [p.
-42] sate, "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions."[^8] Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
-Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-@RN4982 [p.
-72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
-
-[^7]: We can think of students sending SAT scores to colleges as a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
-Next, the authors develop "Tinker Toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
-
-[^8]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
-54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
-
-> The layering of collegiate competition is primarily a socioeconomic layering.
-The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
-Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
+[^9]: We were told that savvier EPS users visited high schools that EPS recommended not visiting because there would be less competition for the good students that attended these high schools.
 
@@
 Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
+
 Additionally, @RN4982 created geomarkets in service of the Market Segment Model.
@@
 We present the following hypothesis:
-
-**H1**. In large metropolitan areas that contain multiple geomarkets, we expect significant socioeconomic inequality (e.g., income, parental education) between geomarkets.
 
@@
 We present the following hypothesis:
-
-**H2**. In large metropolitan areas that contain multiple geomarkets, we expect significant racial inequality (e.g., income, parental education) between geomarkets.
-
-The Market Segment Model and Geomarkets raise concerns common to the application of correlations.
-The data on 1980 SAT score-sending behavior analyzed by @RN4982 can be conceived as training data.
-These training data defined the Geomarkets in that the number of geomarkets created in a metropolitan area was based on the number of test-takers [@RN4982].
-The training data identified student characteristics correlated with the local, in-state, regional, and national market segments.
-These analyses found that student demand for different kinds of colleges was largely a function of social origin, and that student demand within a Geomarket was largely a function of the class composition of the Geomarket.
-However, @RN4982 did not interrogate the historical and contemporary structural inequalities that produced observed patterns of social stratification.
-Instead, the Market Segment Model encouraged colleges to use this snapshot of student demand as the basis for future decisions about which communities and prospective students should be targeted by college recruiting campaigns.
-This process exeemplifies the "ratchet effect" [@RN4851], whereby predicting the behavior of future individuals based on the behavior of past individuals "threatens to reify and reporduce existing inequalities" [@RN4794, p.
-224].
-This process also exemplifies concerns about proxy variables, whereby selective colleges view privileged social class as a proxy for desirable students.
-This creates an incentive to focus recruiting efforts on communities that have these characteristics.
-@RN4982 [p.
-44] discuss these concerns:
-
-> On occasion, senior spokespersons for the profession worry that students outside the main market areas remain forgotten and hence, unchallenged.
-Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
-The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
-
-Commodification of the Market Segment Model and Geomarkets "engineers homophily" [@RN4975].
-In itself, the Market Segment model is a "stylized fact" [@RN5025] about the structure of student demand, one that does not consider historic structural inequalities that produced observed patterns.
-Once the College Board inscribed the Market Segment Model and Geomarkets into the EPS product, this product may amplify the effect of historic structural inequalities on future opportunity structure; a snapshot of extant stratification in student demand becomes the basis for college decisions about which local markets to target.
-
-Utilization of the EPS product raises concerns about the quantification themes of reactivity, discipline, and authority [@RN5007].
-@RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Market research suggests that EPS was highly salient.
-@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
-We suggest that EPS disciplines colleges to approach recruiting in a manner consistent with the Market Segment Model.
-Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters by, first, choosing which Geomarkets to recruit from -- those with a large number of prospects from the desired market segment -- and, second, informing which high schools to visit within these Geomarkets.
-Under the principle of homophily, selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-
-EPS may also weaken the authority of local decision-makers.
-@RN4982 sought to develop a concrete, data-driving framework -- The Market Segment Model -- replicated the aggregate knowledge of local admissions officers.
-Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
-The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
-In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
-However, use of EPS contributed to most colleges visiting the same sets of affluent, high-achieving high schools.[^9]
-
-[^9]: We were told that savvier EPS users visited high schools that EPS recommended not visiting because there would be less competition for the good students that attended these high schools.
 
@@
 
+Institutional theory defines the organizational field as "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services and products" [@RN527, p.
+143].
+
 # References

### 2025-03-08 index.qmd
starting to revise front end
revising front end v3 continued

@@
 **[University of Chicago admissions counselors, IL territories](https://collegeadmissions.uchicago.edu/contact)**
-![](./assets/images/chicagoland_territories_u_chicago_redacted.png)
+![](./assets/images/territories/chicagoland_territories_u_chicago_redacted.png)
 :::

### 2025-03-08 text/front_end_v3.qmd
revising front end v3 continued

@@
 \newpage
-\pagenumbering{gobble}
-\pagenumbering{arabic}
 
@@
 
+\newpage
+\pagenumbering{gobble}
+\pagenumbering{arabic}
+
 # Introduction
+
+On January 9, 2019 a public research university purchased three "student lists" from College Board in order to obtain the contact information of out-of-state prospects, who would subsequently be sent email, mail, and targeted social media.
+Each purchase utilized three "search filters," thereby controlling which prospects profiles were purchased.
+First, all three student list purchases filtered for the 2020 high school class.
+Second, under the heading "College Board Exams," each purchase filtered on a different PSAT score range: 1070 - 1180 for order #448922; 1190 - 1260 for order #448427; and 1270 - 1520 for order #448440.
+Third, under the heading "Geography", each purchase filtered on the same 67 "Geomarkets," which had titles like "IL07 - Chain of Lakes", "IL08 - Northwest Suburbs", "IL09 - North Shore."
+
+Geomarkets carve states and metroplitan areas into smaller geographic regions meant to define local recruiting markets.
+A closer look at the Geomarkets selected for orders #448922, #448427, and #448440 shows that each order selected all Geomarkets in the Chicago-land area, except "IL10 - City of Chicago", which is significantly poorer and less White than surrounding Geomarkets.
+Similarly, in the DC/Maryland/Virginia metro area, each order included the "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but not the "MD05 - Prince Georges Metropolitan" "MD07 - Baltimore (Urban)" or "DC01 - District of Columbia" Geomarkets which are poorer and more Black.
+Recent research suggests that student lists are surprisingly important for college access outcomes and college graduation outcomes, particularly for first-generation and underrepresented minority students [@RN4739; @RN4894].
+However, the above Geomarket selections raise the concern that Geomarkets -- and other geographic search filters -- may be means of excluding working class and non-white students from educational opportunities.
+
+This manuscript does blah blah.
 
@@
 11] describe the Market Segment Model as an effort to "quantify" the knowledge that college admissions officers have about the geography of student demand:
-
-@list_empirics say this:
 
@@
 higher education since 1972, when College Board began selling "names" [@belkin2019-studata; @list_biz].
-@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
-Lists contain contact information for prospective students.
-The Student Search Service database consists of College Board test-takers.
+<!-- @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+Lists contain contact information for prospective students.--> The Student Search Service database consists of College Board test-takers.
 Colleges pay a fee for each prospect (e.g., \$0.50 in 2021).
@@
 Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets.
-An analysis of 830 student lists purchased by 14 public universities found that 16% of purchases filtered on Geomarket [@RN5012].
+An analysis of 830 student lists purchased by 14 public universities found that X% of purchases filtered on Geomarket [@RN5012].
 
@@
 A university could obtain this for their own institution or for a competitor.^[Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
-This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.]
-
-Taken together, the Market Segment Model show a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
 These outputs became the basis for the Enrollment Planning Service (EPS) software.
@@
 Therefore, colleges used EPS software to decide which Geomarkets to recruit from and which high schools to visit within targeted Geomarkets.
-Typical marketing material describes EPS as,
+Typical @RN5019 marketing material describes EPS as,
 
@@
 With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
-Focus your valuable time and resources on the right prospects [@RN5019].
+Focus your valuable time and resources on the right prospects.
 
@@
 EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.^[Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."].
-Second, EPS users decide which high schools they will visit within selected Geomarkets.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."] Second, EPS users decide which high schools they will visit within selected Geomarkets.
 Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
 
-START HERE FRIDAY!!!!!
-
 EPS may also weaken the authority of local decision-makers.
-@RN4982 sought to develop a concrete, data-driving framework -- The Market Segment Model -- replicated the aggregate knowledge of local admissions officers.
+@RN4982 sought to develop a concrete, data-driven framework -- The Market Segment Model -- that replicates the aggregate knowledge of local admissions officers.
 Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
@@
 In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
-However, use of EPS contributed to most colleges visiting the same sets of affluent, high-achieving high schools.[^9]
+However, EPS software recommended visiting the same sets of affluent, high-achieving high schools that were receiving visits from other colleges.
+We were told that savvier, quantitatively-adept admissions offices used EPS to visit schools explicitly not recommended by EPS because there would be less competition for the good students that attended these high schools.
 
-[^9]: We were told that savvier EPS users visited high schools that EPS recommended not visiting because there would be less competition for the good students that attended these high schools.
+Finally, we suspect that Geomarkets affect how admissions offices organize the recruiting territories assigned to each admissions officer.
+Geomarkets were created to mirror the territories of admissions officers [@RN4982].
+In turn, as Geomarkets became more salient, admissions offices often structured their territories around Geomarkets.
+On background, admissions officers and enrollment consultants told us it was common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket.
+In those states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
+This can be seen in the Chicago-land recruiting territories of University of Chicago [ADD TO APPENDIX], of Lake Forest College,  of Saint Mary's College [ADD TO APPENDIX], [ADD TO APPENDIX].
+<!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
 
-Research question 1 asks, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
+Although we cannot gain access to EPS software, Geomarkets are fundamental to EPS software and (arguably) to the organization of recruiting territories for many colleges.
+Therefore, we are interested in the extent to which this building block is associated with race and class.
+Research question 1 asks, what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
+We focus on metropolitan areas that are associated with three or more Geomarkets.
+
 We expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
 Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
+Moreover, @RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards distinguishing geographic areas that differed from one another in terms of class composition and college-going behavior.
 
+<!-- 
 Additionally, @RN4982 created geomarkets in service of the Market Segment Model.
@@
 We present the following hypothesis:
+-->
 
-We expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons.
+We also expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons.
 Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race once.
@@
 cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779].
-Structures built upon racialized structures are racialized structures.
+Structures built upon racialized structures are racialized structures [@RN4786].
 Unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
@@
 Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
-Although these arguments are speculative, the extent of racial inequality between geomarkets can be assessed empirically.
-We present the following hypothesis:
 
-## Geomarket Filter in Student Search Service
-
-In XXXX, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service.
+**Geomarket Filter in Student Search Service**.
+By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service [CITE].
 A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.
-For public universities that purchased lists, 80% purchase more than 50,000 names annually.
-@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.
-Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.^[For public universities that purchased lists, 80% purchase more than 50,000 names annually.
+@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
 
@@
 A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
-After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.
-Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).[^10] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^11]
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^10] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^11]
 
-[^10]: Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
+[^10]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
+Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
 
@@
 Using a national sample of high school students and using data from actual student lists purchased by public universities, @RN5012 show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
-<!-- Using data from actual student lists purchased by public universities, they show that raciaized search filters -- used in conjunction with other search filters -- yield dramatic racial inequality in which prospective students are targeted.
--->
 
 Because of the extent of residential segregation in the U.S., geographic borders are a commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887].
-Products built upon structural racism are structurally racist products.
 Student list products offer many geographic filters.
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-Additionally, the College Board created geographic borders that subsequently became filters in the *Student Search Service*.[^12] Geomarket search filters, slice states and metropolitan areas into smaller, local recruiting markets.
+College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^12] This manuscript focuses on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 
-@RN4982 [pp.
-42-44] provides recommendations about how The Market Segment Model and Geomarkets should be utilized in the context of purchasing student lists.
-The Market Segment model "simply demonstrates what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations" (p.
-42).
-The authors describe a college that "draws most of its students from regional or national segments" (p.
-42).
-Focusing on New England -- including the Boston Geomarket, which was relatively low-income in 1980 -- the authors ask, "where would you concentrate your energies" (p.
-42),
+@RN4982 [pp. 42-44] discuss how the Market Segment Model and Geomarkets should be utilized in the context of purchasing student lists. The authors ask,
 
-> Ideally, you would seek communities \[Geomarkets\] with a high proportion of students already predisposed toward institutions such as your own.
+> Imagine for the moment that you recruit for a college that draws most of its students from regional or national segments.
+Where would you concentrate your energies?
+Ideally, you would seek communities \[Geomarkets\] with a high proportion of students already predisposed toward institutions such as your own.
 The Market Segment Model would provide this information through segment percentages for the community in question.
 Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
-Your efforts would surely be better directed toward three of the four communities in the bottom band, Manchester, Hartford, and Fairfield County.
-Simply by knowing a little bit about the students' backgrounds and academic records you could quickly focus your attention on those most likely to consider your kind of institution.
-Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own" (p.
-42-44)
+Your efforts would surely be better directed toward \ldots Manchester, Hartford, and Fairfield County \ldots Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own.
 
@@
 By contrast, the student list purchases of local state colleges and community colleges, which rely on enrollment from students in the "local" and "in-state," should focus on nearby Geomarkets with large numbers of low-SES students.
-Our analyses attempt to recreate the strategy recommended by @RN4982.
+Our analyses attempt to recreate the strategy recommended by @RN4982 by showing who is included/excluded when student list purchases filter on particular Geomarkets.
 Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access.
-Furthermore, we anticipate that the students of color targeted by this strategy are disproportionately enrolled in affluent, predominantly white schools.
-
-```{=html}
-<!-- THIS TEXT IS ABOUT UTILIZATION. CUT FOR NOW.
-Finally, we suggest that utilization of geomarket filters within the Student Search Service product is likely to amplify class-based and race-based disparities in college opportunity.
-In contrast to standardized selection devices (e.g., the FICO score algorithm), student list products are discretionary selection devices in that colleges have discretion of which filters they elect to select prospects.
-Compared to standardized selection devices, scholars tend to find that administrative discretion over selection devices causes structural inequality to increase [@RN4883; @RN4881; @RN4795].
-Even in the absence of explicit or implicit bias, discretionary selection devices can increase racial inequality when decision-makers have incomplete knowledge about how the inputs they choose are correlated with race.
-For example, research shows that Americans dramatically underestimate the magnitude of racial income inequality [@RN4884].
-As such, discretionary selection devices are particularly sensitive to inputs that are correlated with race but are perceived to be race-neutral.
-
-Geomarkets are opaque in that geomarket borders may be unclear to someone purchasing student lists and very few people know how these borders were created.
-@RN4849 argues that opacity in algorithmic products is intentional and is a core criterion of "weapons of math destruction." Enrollment management professionals utilizing EPS and *Student Search Service* may interpret geomarkets as a race-neutral recruiting tool.
-When purchasing student lists, @RN4982 [p.
-X] suggest that affluent colleges should consider targeting affluent geomarkets.
-Student list purchases do not show how the characteristics of targeted prospects compare to the characteristics of the surrounding community.
-From this perspective, racial inequality from the use of geomarkets in student list purchases may be the unintentional result of using a seemingly race-neutral product.
-Furthermore, purchases that specify multiple filters (e.g., test scores, grades, intended major) including geomarkets can yield unintended racial inequality because administrators have incomplete knowledge about how the intersection of these filters interact with local patterns of segregation.
--->
-```
-
-```{=html}
-<!-- CUT TEXT DEFINING COLORBLIND RACISM AND RACE-NEUTRAL RACISM
-Colorblind racism describes ideology and associated behaviors that ignore the significance of race and racial discrimination [@RN5020].
-The Market Segment Model can be seen as an example of colorblind racism.
-@RN4982 develop an analysis of the structure of college choice that ignores race but we contend that Geomarkets are correlated with race.
-Race-neutral racism is racial inequality that results from using ostensibly race-neutral inputs that are systematically correlated with race [CITE] [e.g., @RN4775; @RN4786; @RN4975].
--->
-```
-
-```{=html}
-<!-- CUT PARAGRAPH: WHY GEOMARKETS ARE WEAPONS OF MATH DESTRUCTION
-The incorporation of Geomarkets within the Student Search Service student list product raises concerns about the three characteristics of weapons of math destruction.
-First, Geomarkets are opaque.
-The borders of a geomarket may be unclear to someone purchasing student lists and very few people know how these borders were created \[COULD TALK ABOUT PRODUCT UTILIZATION HERE; NOT CLEAR HOW GEOMARKETS INTERACT WITH OTHER STUDENT LIST FILTERS\].
-Second, Geomarkets scale.
-Each year, thousands of colleges purchase student lists and it is easy to filter purchases based on Geomarkets in concert with other filters (e.g., SAT score, GPA).
-Third, Geomarkets can cause damage.
-When a college decides to purchase names from one Geomarket but exclude prospects from an adjacent Geomarket, this affects which college students have an opportunity to attend.
-@RN4982 suggests that selective colleges focus on geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
--->
-```
+Furthermore, we anticipate that the students of color from affluent Geomarkets will tend to have college-educated parents and be enrolled in affluent, predominantly white schools.
 

### 2025-03-11 text/front_end_v3.qmd
revised front end

@@
 
-On January 9, 2019 a public research university purchased three "student lists" from College Board in order to obtain the contact information of out-of-state prospects, who would subsequently be sent email, mail, and targeted social media.
-Each purchase utilized three "search filters," thereby controlling which prospects profiles were purchased.
-First, all three student list purchases filtered for the 2020 high school class.
-Second, under the heading "College Board Exams," each purchase filtered on a different PSAT score range: 1070 - 1180 for order #448922; 1190 - 1260 for order #448427; and 1270 - 1520 for order #448440.
-Third, under the heading "Geography", each purchase filtered on the same 67 "Geomarkets," which had titles like "IL07 - Chain of Lakes", "IL08 - Northwest Suburbs", "IL09 - North Shore."
+On January 9, 2019 a public research university purchased College Board *Student Search Service* order no.
+$448922$, thereby obtaining the contact information of out-of-state prospective students, who would subsequently receive brochures, postcards, email, and targeted social media.
+This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
+The order specified three "search filters" to control which names were purchased.
+First, the order filtered on 2020 high school graduating class.
+Second, under the heading "College Board Exams," the order filtered on PSAT score range of 1070 - 1180.
+Third, under the heading "Geography", the order filtered on 67 "Geomarkets." Geomarkets carve states and metroplitan areas into smaller geographic units and have names like, IL08 - Northwest Suburbs" and "IL09 - North Shore."
 
-Geomarkets carve states and metroplitan areas into smaller geographic regions meant to define local recruiting markets.
-A closer look at the Geomarkets selected for orders #448922, #448427, and #448440 shows that each order selected all Geomarkets in the Chicago-land area, except "IL10 - City of Chicago", which is significantly poorer and less White than surrounding Geomarkets.
-Similarly, in the DC/Maryland/Virginia metro area, each order included the "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but not the "MD05 - Prince Georges Metropolitan" "MD07 - Baltimore (Urban)" or "DC01 - District of Columbia" Geomarkets which are poorer and more Black.
+On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
+On closer inspection, the orders filtered on all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago", which is significantly poorer and less White than surrounding Geomarkets.
+In the DC/Maryland/Virginia metro area, the orders included the "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets, which are poorer and have a higher share of Black residents.
+An analysis of 500 [check] student lists purchased by 9 public research universities found that 11% of purchases filtered on Geomarket [@RN5012].
 Recent research suggests that student lists are surprisingly important for college access outcomes and college graduation outcomes, particularly for first-generation and underrepresented minority students [@RN4739; @RN4894].
-However, the above Geomarket selections raise the concern that Geomarkets -- and other geographic search filters -- may be means of excluding working class and non-white students from educational opportunities.
+To the extent that Geomarket borders are correlated with race and class, the Geomarket selections observed here raise concern that Geomarkets -- and other geographic search filters -- may systematically filter working class and non-white students from educational opportunities.
 
-This manuscript does blah blah.
+The origins of Geomarkets, and how they were incorporated into recruiting products, turns out to be an interesting story.
+In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*.
+College Board published the book, provided data, and financed the project, as part of their "efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p.
+x].
+@RN4982 sought to "capture and quantify" (p.
+11) the knowledge of admissions officers,
 
-In 1983, Robert Zemsky and Penney Oedel authored *The Structure of College Choice*.
-Based on an analysis of 1980 SAT score-sending behavior and published by *The College Board*, @RN4982 developed the Market Segment Model to help colleges recruit students.
-The underlying idea of the Market Segment Model is that student demand for your institution is fundamentally a function of class.
-Therefore, the key to recruitment is identifying territories where large numbers of your class live.
-@RN4982 [p.
-11] describe the Market Segment Model as an effort to "quantify" the knowledge that college admissions officers have about the geography of student demand:
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots [because] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
+9-10).
 
-> A good recruiter knows where to look for prospective applicants, as seen in the students' willingness or eluctance to travel.
-It is necessary to identify not only the most promising communities or poools, but also the specific neighborhoods within those communities -- hence the recruiter's emphasis on feeder high schools.
-From the beginning, we have sought to mold our research to these concepts, to capture and quantify the phenomenon behind the folklore.
-Thus, it is the admissions officers' notion of admissions pools that dictates the geographic units of our analysis." \[p.
-11\]
+@RN4982 developed Geomarkets and the Market Segment Model based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel." Created by analyzing the SAT score-sending behavior of 1980 high school seniors, the thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
+Therefore, colleges should recruit from territories that contain large populations of your target social class.
+Geomarkets are geographic borders meant to define local recruiting markets, a territory of an admissions recruiter.
+Figure X, panel A shows Geomarkets in the Chicago-land area.
+The Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
 
-@RN4982 also created local "community based markets" (p.
-14) -- later known as "Geomarkets." Geomarkets are geographic borders that disaggregate states and large metropolitan areas for recruiting purposes.
-@fig-nyc-geomarkets shows New York City area Geomarkets.
-In defining Geomarkets, @RN4982 did not draw new geographic borders, but rather made certain extant borders more salient to the process of recruiting.
+Geomarkets have been incorporated into the supply-side structure of college access in several ways.
+Albeit anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
+Figure X panel B shows that Chicago-land Geomarkets define recruiting territories for University of Chicago admissions recruiters (Appendix Figures BLAH BLAH).
+Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS).
+Founded in 1984 and still active today, EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
+According to @RN5021, in 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis product, which was based on EPS.
 
-Geomarkets have been incorporated into the supply-side structure of college access in ways that are simultaneusly mundane and fundamental.
-We posit three ways.
-First, Geomarkets became a principle for how many -- if -- admissions offices organized their labor.
-In states where a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory \[CITE\].
-Geomarkets were created to mirror the territories of admissions officers and, in turn, admissions offices structured their territories around Geomarkets.
-Second, The Market Segment Model and Geomarkets became the basis for Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
-EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
-@RN5019_2010 says "Enrollment Planning Service (EPS\textsuperscript{\textregistered}) the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found." @RN5021 reports that, in 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS.
-
-Third, Geomarkets were incorporated into the College Board "student list" product named Student Search Service.
-Student lists have been the primary source of lead generation in U.S.
-higher education since 1972, when College Board began selling "names" [@belkin2019-studata; @list_biz].
-<!-- @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
-Lists contain contact information for prospective students.--> The Student Search Service database consists of College Board test-takers.
+Third, Geomarkets were incorporated into College Board's student list product, Student Search Service.
+Student lists contain the contact information of prospective students and have been the primary source of lead generation in U.S.
+higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz].
+@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+The Student Search Service database consists of College Board test-takers.
 Colleges pay a fee for each prospect (e.g., \$0.50 in 2021).
@@
 Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets.
-An analysis of 830 student lists purchased by 14 public universities found that X% of purchases filtered on Geomarket [@RN5012].
 
-Geomarkets have never been the subject of empirical research.
-This omission is surprising because the sociology of education is concerned with sorting [@RN4873], particularly sorting into college.
-Traditionally, most scholarship -- and certainly most scholarship in economics \[HOXBY?
-CHETTY\] -- views inequality in college access as a function households or k-12 schools, which represent the demand-side of higher education.
-More recently, scholarship in the sociology of education takes seriously the idea that colleges are not passive recipients of applications, but rather they expend substantial resources soliciting demand from desirable applicants [@RN3519; @RN4774; @RN4324].
-However, most scholarship on recruiting analyzes the behavior of individual colleges [e.g., @RN4758; @RN4759], thereby creating the assumption that recruiting is a function of individual colleges.
-What these studies have missed is the broader enrollment management industry -- third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access, funneling certain kinds of students to certain kinds of institutions.
+Scholarship at the nexus of college access and recruiting has focused on the behavior of colleges [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
+Scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access.
+These mechanisms, which lie upstream of the behavior of individual colleges, are incorporated by individual colleges in ways that funnel certain kinds of students to certain kinds of institutions.
+This omission is surprising because sociology is concerned with sorting [@RN4873], particularly sorting into college.
 
-This manuscript analyzes one of those mechanisms, The Market Segment Model and College Board Geomarkets, as a case study of quantification.
-We develop arguments from the broader literature on quantification [e.g., @RN5003; @RN5005; @RN5007], particularly the discussions of correlation and homophily from @RN4975.
-Predictive analytics are based on correlation.
-When we use data on past correlations to make recommendations about the future, we amplify the effects of historic structural inequality.
-The Market Segment Model analyzes SAT score-sending data from 1980 and concludes that student demand for higher education is primarily a function of social class [@RN4982].
-The Market Segment Model argues that homophily -- actors that share characteristics are likely to form connections -- is the organizing principle of student demand and competition between colleges.
-@RN4982 [p.
-72] state that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students." This snapshot of student demand in 1980, itself a consequence historic structural inequality, was programmed into the way admissions offices organize labor and programmed into the EPS and Student Service Service products that colleges utilize to identify and target prospective students.
-The result is a supply-side that reinforces structural inequalities observed on the demand-side.
+This manuscript analyzes College Board Geomarkets as a case study of quantification.
+In particular, we draw from the discussions of correlation and homophily from @RN4975.
+@RN4982 identified the correlates of 1980 SAT score-sending behavior and concluded that student demand for higher education is primarily a function of social class.
+The Market Segment Model argues that homophily -- actors that share characteristics form connections -- is the organizing principle of competition and college choice, stating that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students" [@RN4982, p.
+72].
+Scholarship on quantification demonstrates that making recommendations based on past correlations amplifies the effects of historic structural inequality [@RN4794].
+The snapshot of student demand in 1980 -- itself a consequence historic structural inequality -- was programmed into recruiting products that colleges utilize to identify and target prospective students.
+The result is a supply-side that amplifies structural inequalities observed on the demand-side.
 
-\[ADD TEXT SOMEWHERE AROUND HERE ABOUT WHY THIS TOPIC IS SIGNIFICANT AND IMPORTANT FOR CURRENT/FUTURE POLICY DEBATES: Why market segment model important.
-Software as service products dominate em industry and structure college access.
-Look at eab machine learning product But rarely do we see the recipe underneath these products.
-The book structuring college access is that recipe for EPS software\]
-
-Our analyses address the two research questions, which speak to how geomarkets are utilized within EPS and within the Student Search Service student list product.
+Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS and within the Student Search Service student list product.
 First, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
@@
 We obtained these data by issuing public records requests to public universities.
-<!-- Third, in student list purchases that filtered on geomarkets, how does the socioeconomic and racial composition of included students compare the that of the geomarket and how does it compare to that of the broader metropolitan area?
--->
+
+The following section provides background about enrollment management and introduces salient concepts from scholarship on quantification.
+Second, we introduce the case, describing how @RN4982 developed Geomarkets and the Market Segment Model and how these concepts were incorporated into College Board recruiting products.
+Third, we describe data and methods.
+Fourth, we present results.
+[We find that\ldots ?PARAGRAPH ON RESULTS?] Finally, we discuss implications for scholarship and for policy.
+The sociology of education should move beyond the fixation on schools and policies as the primary sorting machines.
+@RN527[p.
+148] remind us that the organizational field includes the totality of relevant actors." Third-party vendors have structured college access for decades and their influence is growing.
+The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [CITE].
+By investigating high-leverage actors and products in the organizational field, sociology can develop important insights that shape future federal and state regulatory policy.
 
 ```{=html}
+\[ADD TEXT SOMEWHERE AROUND HERE ABOUT WHY THIS TOPIC IS SIGNIFICANT AND IMPORTANT FOR CURRENT/FUTURE POLICY DEBATES: Why market segment model important.
+Software as service products dominate em industry and structure college access.
+Look at eab machine learning product But rarely do we see the recipe underneath these products.
+The book structuring college access is that recipe for EPS software\]
 <!-- 
@@
 ```
-
-In the following section we provide background information about enrollment management and scholarship on recruiting.
-Second, we introduce core ideas and concepts from scholarship on quantification.
-Third, we apply these ideas to a close read of @RN4982 and motivate hypotheses.
-Four, we describe data and methods.
-Fifth, we present results.
-Finally, we discuss implications for scholarship and for policy.
-
-A note from RESULTS that I texted to Bastedo on 12/19/2024
-
--   But these geomarkets are pretty fucked up.
-They were designed to capture class.
-I see lots of orders that condition on test scores and rich geomarkets.
-But I know from similar purchases that don’t condition on geomarket, that the high scoring first gen Asian and Hispanic students tend to come from the poor geomarkets that get excluded from a lot of purchases.
-By contrast, The Asian and Hispanic students that do get included in purchases that target rich geomarkets are usually not first gen
 
@@
 "Leads" are prospects whose contact information has been obtained.
-"Inquiries" are prospects that contact the institution, consisting of those that respond to an initial soliciation (e.g., email) and those that reach out on their own (e.g., sending SAT scores).
+"Inquiries" are prospects that contact the institution, including those that respond to an initial solicitation (e.g., email) and those that reach out on their own (e.g., sending SAT scores).
 <!--  Applicants consist of inquiries who apply plus "stealth applicants" who do not contact the university before applying.
@@
 At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz].
-Purchased leads are served emails, brochures, and targeted social media advertisements designed to solicit inquiries and applications [@RN4895].
+Purchased leads are served emails, brochures, and targeted social media designed to solicit inquiries and applications [@RN4895].
 
@@
 
-Drawing from scholarship on organizational theory [@RN4738], enrollment management processes involve many "make or buy" decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
+Drawing from scholarship on organizational theory, enrollment management processes involve many "make or buy" [@RN4738] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
 EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
 As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest enrollment consulting firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
+The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
 In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
@@
 **Correlation**.
-The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4849; @RN4772; ].
-In particular, the discussions of correlation (chapter 1) and homophily (chapter 2) by @RN4975 introduce ideas salient to the analysis the Market Segment Model and College Board Geomarkets.
+The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4772; @RN4849].
+In particular, the discussions of correlation (chapter 1) and homophily (chapter 2) by @RN4975 introduce ideas salient to the analysis the Market Segment Model and Geomarkets.
 Correlation measures the extent to which two or more variables move together.
@@
 @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-Like correlational models discussed above, network science models "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
+Like correlational models discussed above, because most network science models do not observe how historical structures affect relationships, they "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
 95] that cause behavior consistent with homophily.
-<!--"Homophily assumes and creates segregation" [@RN4975, p.
-95] writes.
-"It transforms individuals into 'neighbors' who naturally want to live with people 'like them'." --> In commercial social networks, @RN4975 argues, homopphily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
+Second, in commercial social networks, homopphily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
 Thus, "social networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@RN4975, p.
 YY].
-We observe similar processes in college access because higher education is an industry in which providers are evaluated based on the characteristics of their customers [@RN1549; @RN1237].
+We observe similar processes in college access.
 Based on a snapshot of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
@@
 
-Homophily is also central to products that categorize customers.
+Homophily is central to market research products that categorize customers.
 Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
@@
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are targeted to a vertical hierarchy of consumers.
-We observe similar processes in our case study, whereby @RN4982 categorizes students into four market segments -- local, in-state, regional, and national -- based on their score sending behavior and then evaluates the attractiveness of each Geomarket based on how many "regional" and "national" students live there.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
+We observe similar processes in our case study.
+@RN4982 categorizes students into four market segments -- local, in-state, regional, and national -- based on their score sending behavior and then evaluates the attractiveness of each Geomarket based on how many "regional" and "national" students live there.
 <!--
@@
 However, this study provides inisight about the mechanism underlying the effect.
-By analyzing and simulating *Student Search Service* purchases that filter on Geomarkets we show how Geomarkets reproduce historical race-based inequality in college access [THIS LAST BIT FEELS LIKE THE GENERAL CONTRIBUTION TO SOCIOLOGY].
+By analyzing and simulating *Student Search Service* purchases that filter on Geomarkets we show how Geomarkets reproduce historical race-based inequality in college access.
 
@@
 **Creating Geomarkets and the Market Segment Model.** In 1978, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
-x].^[Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.] Zemsky reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential" [@RN4982, p.
 x].
-
-The geographic concepts inscribed in the Market Segment Model were based on conversations with college admissions officers.
-@RN4982 \[p.
-4\] state that "our initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools." Quoting an admissions officer, @RN4982 [p.
-11] write, "'There are only three kinds of colege-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, [@RN4982] created three geographic [@RN4982, p.
+Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
+<!--The geographic concepts inscribed in the Market Segment Model were based on conversations with college admissions officers.
+-->  The "initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982,p.
+4].
+Quoting an admissions officer, @RN4982 [p.
+11] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, [@RN4982] created three geographic [@RN4982, p.
 11], region, state, and community.
 The three regions were New England, Middle States, and the South.
-Next, "we divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets." These enrollment markets, later called Geomarkets, were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
+They "divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" (p.
+11).
+These enrollment markets, later called Geomarkets, were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
 @RN4982[pp.
@@
 
-The goal of the Market Segment Model was to quantify the knowledge that admissions counselors have about student demand.
-In 1979, College Board began providing funding and data on the score-sending behavior of SAT test-takers.
-@RN4982 [p.
-4] write that,
+The goal of the Market Segment Model was to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households in the Geomarket.
+Zemsky reached out to College Board because "we needed a database that described most institutions and most students" [@RN4982, p.
+x].
+<!--Zemsky reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential" [@RN4982, p.
+x].
+-->In 1979, College Board began providing funding and data on the score-sending behavior of SAT test-takers.
 
@@
 The Market Segment model that we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
-The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways.
+The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways [@RN4982, p.
+4].
 
@@
 
-The primary takeaway of Chapter 3 is that student demand for higher education is mostly a function of social origin.
+The thesis of Chapter 3 is that student demand for higher education is a function of social origin.
 For example, @RN4982[p.
@@
 **Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.
-This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
-The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
-Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.[^8]
-
-Based on these analyses, @RN4982 [p.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^8] Based on these analyses, @RN4982 [p.
 46] state that homophily describes the nature of competition between colleges: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
@@
 
-[^8]: For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
+[^8]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
+Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
+These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
+For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
 54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.

### 2025-03-12 text/front_end_v3.qmd
intro

@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased.
-First, the order filtered on 2020 high school graduating class.
-Second, under the heading "College Board Exams," the order filtered on PSAT score range of 1070 - 1180.
-Third, under the heading "Geography", the order filtered on 67 "Geomarkets." Geomarkets carve states and metroplitan areas into smaller geographic units and have names like, IL08 - Northwest Suburbs" and "IL09 - North Shore."
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets." Geomarkets carve states and metroplitan areas into smaller geographic units and have names like, IL08 - Northwest Suburbs" and "IL09 - North Shore." On the same day, the university purchased two additional orders that were identical, except they filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$.
+<!--On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
+-->
 
-On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
-On closer inspection, the orders filtered on all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago", which is significantly poorer and less White than surrounding Geomarkets.
-In the DC/Maryland/Virginia metro area, the orders included the "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets, which are poorer and have a higher share of Black residents.
-An analysis of 500 [check] student lists purchased by 9 public research universities found that 11% of purchases filtered on Geomarket [@RN5012].
+On closer inspection, the three orders selected on all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago", which is significantly poorer and less White.
+In the DC/Maryland/Virginia metro area, the orders selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets, which are poorer and have a higher share of Black residents.
+An analysis of 500 \[check\] student lists purchased by 9 public research universities found that 11% of purchases filtered on Geomarket [@RN5012].
 Recent research suggests that student lists are surprisingly important for college access outcomes and college graduation outcomes, particularly for first-generation and underrepresented minority students [@RN4739; @RN4894].
@@
 
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots [because] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
 9-10).
@@
 The sociology of education should move beyond the fixation on schools and policies as the primary sorting machines.
-@RN527[p.
+@RN527 [p.
 148] remind us that the organizational field includes the totality of relevant actors." Third-party vendors have structured college access for decades and their influence is growing.
-The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [CITE].
+The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product \[CITE\].
 By investigating high-leverage actors and products in the organizational field, sociology can develop important insights that shape future federal and state regulatory policy.
@@
 <!-- la la -->
+
 Enrollment management is simultaneously a profession, an administrative structure, and an industry.
@@
 <!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
--->
-<!--  
+--> <!--
 [^1]: New institutional theory offers one explanation [@RN527].
@@
 
+```{=html}
 <!--
@@
 50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"^[Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
+As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"[^1]
+
+[^1]: Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
 For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
 @RN4849 [p.
-146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?']
+146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
 
@@
 Homophily is a core concept of network science (writ large), in which actors (nodes) are connected to one another directly and indirectly via network ties (edges).
-As discussed below, @RN4982[chapter 4] *The company We Keep: Colleges and Their Competition* conducts a network analysis based on 1980 SAT score sending behavior whereby two colleges are defined as competitors if a large number of students send SAT scores to both colleges.
+As discussed below, @RN4982 [chapter 4] *The company We Keep: Colleges and Their Competition* conducts a network analysis based on 1980 SAT score sending behavior whereby two colleges are defined as competitors if a large number of students send SAT scores to both colleges.
 Because network science models often draws from rational choice theory, @RN4975 argues, it assumes that homophily is the result of voluntary action by individuals.
@@
 
-[^3]: Thus, @RN4975 [p.
-82] writes, "echo chambers are not unfortunate errors, but deliberate goals" because "homophily is used to create agitated clusters of individuals whose angry similarity and overwhelming attrication to their common object of hatred both repel them from one another and glue them together."
-
 Homophily is central to market research products that categorize customers.
@@
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^3] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
 We observe similar processes in our case study.
@@
 
-[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
+[^3]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
 
@@
 These enrollment markets, later called Geomarkets, were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
-@RN4982[pp.
+@RN4982 [pp.
 11-12] briefly describes the creation of Geomarket borders:
@@
 Both outputs are created separately for each Geomarket.
-For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than $35,000, percent with both parents having a BA, etc.^[For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
-A university could obtain this for their own institution or for a competitor.^[Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
-This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^4] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+A university could obtain this for their own institution or for a competitor.[^5] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
 These outputs became the basis for the Enrollment Planning Service (EPS) software.
 
+[^4]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
+
+[^5]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
+
 **Correlations**.
-@RN4982[Chapter 3], *A Sense of Place: Students, Families, and Communities*, identifies the student characteristics correlated with being in the local, in-state, regional, or national market segment.
+@RN4982 [Chapter 3], *A Sense of Place: Students, Families, and Communities*, identifies the student characteristics correlated with being in the local, in-state, regional, or national market segment.
 The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
@@
 The thesis of Chapter 3 is that student demand for higher education is a function of social origin.
-For example, @RN4982[p.
-33] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^6] @RN4982[p.
-42] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
-Therefore, @RN4982[p.
+For example, @RN4982 \[p.
+33\] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
+33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^6] @RN4982 \[p.
+42\] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
+Therefore, @RN4982 [p.
 44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:
-
-> On occasion, senior spokespersons for the profession worry that students outside the main [Geo]market areas remain forgotten and hence, unchallenged.
-Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
-The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
 
@@
 
+> On occasion, senior spokespersons for the profession worry that students outside the main \[Geo\]market areas remain forgotten and hence, unchallenged.
+Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
+The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
+
 **Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^8] Based on these analyses, @RN4982 [p.
-46] state that homophily describes the nature of competition between colleges: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^7] Based on these analyses, @RN4982 \[p.
+46\] state that homophily describes the nature of competition between colleges: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
 Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
@@
 
-> Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
-The layering of collegiate competition is primarily a socioeconomic layering.
-The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
-Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
-
-[^8]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+[^7]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
 The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
@@
 For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
+
+> Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
+The layering of collegiate competition is primarily a socioeconomic layering.
+The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
+Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
 
@@
 @RN4988 shows the contemporary Geomarkets.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^5] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^8] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
 
-[^5]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
+[^8]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
 Students from California don't all share the same college-going behaviors.
@@
 Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
-EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.^[Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."] Second, EPS users decide which high schools they will visit within selected Geomarkets.
+EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
 Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
+
+[^9]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
 
@@
 In those states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
-This can be seen in the Chicago-land recruiting territories of University of Chicago [ADD TO APPENDIX], of Lake Forest College,  of Saint Mary's College [ADD TO APPENDIX], [ADD TO APPENDIX].
+This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX\], \[ADD TO APPENDIX\].
 <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
@@
 **Geomarket Filter in Student Search Service**.
-By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service [CITE].
+By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service \[CITE\].
 A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.^[For public universities that purchased lists, 80% purchase more than 50,000 names annually.
-@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^10] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+
+[^10]: For public universities that purchased lists, 80% purchase more than 50,000 names annually.
+@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.
 
@@
 A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
-After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^10] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^11]
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^11] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^12]
 
-[^10]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
+[^11]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
 Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
 
-[^11]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
+[^12]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
 Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
@@
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^12] This manuscript focuses on Geomarkets.
+College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^13] This manuscript focuses on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 
-[^12]: For example, the geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
+[^13]: For example, the geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
 
@@
 The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-In order to examine how the characteristics of Geomarkets changed over time, we utilize tract-level data from the 2000 Decennial Census, similarly using data from Summary File 1A (SF1A) to obtain measures of race and Hispanic origin and using data from Summary File 3A (SF3A) to obtain measures of income, education, and poverty.[^13] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+In order to examine how the characteristics of Geomarkets changed over time, we utilize tract-level data from the 2000 Decennial Census, similarly using data from Summary File 1A (SF1A) to obtain measures of race and Hispanic origin and using data from Summary File 3A (SF3A) to obtain measures of income, education, and poverty.[^14] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
 These questions were replaced by the American Community Survey (ACS).
-We also utilized data from the 2020 5-year ACS, which includes data collected from 2016-2020.[^14] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+We also utilized data from the 2020 5-year ACS, which includes data collected from 2016-2020.[^15] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
 
-[^13]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
+[^14]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
 
-[^14]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
+[^15]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
 
@@
 Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^15] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^16] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
 Using these zip code-level data, we utilized the `aggregate()` function to create shapefile for each Geomarket.
 
-[^15]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
+[^16]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
 
@@
 We also created socioeconomic measures that approxmiate the predictors of market segment identified by @RN4982.
-These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^16] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
+These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^17] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
 
-[^16]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
+[^17]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
 
@@
 We draw from these data to answer RQ2.
-Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^17].
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^18].
 Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
 
-[^17]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
+[^18]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
 

### 2025-03-13 text/front_end_v3.qmd
man i'm a psycho

@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets." Geomarkets carve states and metroplitan areas into smaller geographic units and have names like, IL08 - Northwest Suburbs" and "IL09 - North Shore." On the same day, the university purchased two additional orders that were identical, except they filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$.
-<!--On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."  On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
+
+<!--Geomarkets carve states and metroplitan areas into smaller geographic units and have names like,  On the same day, the university purchased two additional orders that were identical, except they filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$.
+On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
 -->
 
-On closer inspection, the three orders selected on all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago", which is significantly poorer and less White.
+Geomarkets carve states and metroplitan areas into smaller geographic units.
+Upon closer inspection, the three orders selected on all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago", which is significantly poorer and less White.
 In the DC/Maryland/Virginia metro area, the orders selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets, which are poorer and have a higher share of Black residents.

### 2025-03-13 index.qmd
update maps + add qr code

@@
       writeLines(str_c("### Research question ", rq_num, " {.hide}"))
-      writeLines("#### Map")
+      writeLines(str_c("#### Racial/ethnic and socioeconomic characteristics of ", heading_text, " Geomarkets over time"))
 
@@
       writeLines('<p class="fullscreen">Fullscreen</p>')
+      writeLines('<div class="footnote">Figure Notes:<br>- Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic<br>- Household income measured using 2024 CPI</div>')
 
@@
       writeLines(str_c("### Research question ", rq_num, " {.hide}"))
-      writeLines("#### Map")
+      writeLines(str_c("#### Purchased student profiles by racial/ethnic composition and first-generation status, ", heading_text))
 
@@
         m,
-        '.html" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'
+        '.html" src="about:blank" width=100% height=100% style="max-height: 75%" allowtransparency="true"></iframe>'
       ))
       writeLines('<p class="fullscreen">Fullscreen</p>')
+      notes_txt <- readLines(file.path(graphs_dir, rq, str_c('rq2_', m, '_race_row_plot_note.txt')), warn = FALSE)
+      writeLines(
+        str_replace(
+          str_c('<div class="footnote">', paste0(notes_txt, collapse = "</br>"), '</div>'),
+          'Excludes students with missing values for race\\. ', ''
+        )
+      )
 
@@
   }
+
+  # Add RQ2 map
+  writeLines(paste0("### ", heading_text, " (AI/AN) {.hide}\n"))
+  writeLines(str_c("#### Purchased AI/AN student profiles by racial/ethnic composition and first-generation status"))
+
+  writeLines(str_c(
+    '<iframe data-src="./results/maps/rq2_aian_map_',
+    m,
+    '.html" src="about:blank" width=100% height=100% allowtransparency="true"></iframe>'
+  ))
+  writeLines('<p class="fullscreen">Fullscreen</p>')
+  writeLines('<div class="footnote">Figure Notes:<br>- Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+.</div>')
 }
@@
 
+# Thank you!{.align-center}
+
+<div><img class="qr" src="./assets/images/qr_code.png" /></div>
+
 # References

### 2025-03-13 scripts/map_functions.R

@@ -1,4 +1,5 @@
 library(tidyverse)
+library(rmapshaper)
 
 
 source(file.path('scripts', 'directories.R'))

### 2025-03-13 scripts/rq2_maps.R

@@ -183,7 +183,27 @@ student_data <- purchased_data %>%
     all_nosomecollege = all_nocollege + all_somecollege
   )
 
-student_data_multi_orders <- lapply(c('488035_488053', '448427_448440', '448420_448374'), function(orders) {
+orders_data_full <- read_csv(file.path('scripts', 'metro_orders.csv')) %>% 
+  select(-eps_codes, -figure_note) %>% 
+  add_row(metro = 'chicago', order_ids = '392834_488035_392835_488053', test_range = 'SAT score 1160 - 1600') %>% 
+  add_row(metro = 'philadelphia', order_ids = '448427_448440', test_range = 'PSAT score 1190 - 1520') %>%  
+  add_row(metro = 'northern_new_jersey', order_ids = '448427_448440', test_range = 'PSAT score 1190 - 1520') %>%  
+  add_row(metro = 'long_island', order_ids = '448427_448440', test_range = 'PSAT score 1190 - 1520') %>%  
+  add_row(metro = 'dallas', order_ids = '448427_448440', test_range = 'PSAT score 1190 - 1520') %>%  
+  add_row(metro = 'detroit', order_ids = '448427_547005_448440_546978', test_range = 'PSAT score 1190 - 1520') %>% 
+  add_row(metro = 'los_angeles', order_ids = '448420_546946_448374_546945', test_range = 'PSAT score 1190 - 1520') %>%
+  add_row(metro = 'orange_county', order_ids = '448420_546946_448374_546945', test_range = 'PSAT score 1190 - 1520') %>% 
+  add_row(metro = 'san_diego', order_ids = '448420_546946_448374_546945', test_range = 'PSAT score 1190 - 1520') %>% 
+  add_row(metro = 'bay_area', order_ids = '448420_546946_448374_546945', test_range = 'PSAT score 1190 - 1520') %>%
+  mutate(
+    type = c(rep('rq2', 32), rep('aian', 5), rep('rq2', 10))
+  )
+
+names(orders_data_full) <- c('region', 'order_num', 'order_title', 'type')
+
+unique_orders <- unique(orders_data_full$order_num)
+
+student_data_multi_orders <- lapply(unique_orders[str_detect(unique_orders, '_')], function(orders) {
   student_data %>% 
     filter(ord_num %in% str_split(orders, '_')[[1]]) %>%
     mutate(ord_num = orders) %>% 
[…]

### 2025-03-13 scripts/rq2_maps.js

@@ -104,7 +104,8 @@ function(el, x, choices) {
     orderOptionsHTML = '';
     order_region.forEach(function(curr, idx) {
       if (curr === active_attr.active_metro) {
-        orderOptionsHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="order-choice" data-order="' + order_num[idx] + '"><span> ' + order_num[idx].split('_').join(' & ') + ': ' + order_title[idx] + '</span></div>';
+        let t = 0;
+        orderOptionsHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="order-choice" data-order="' + order_num[idx] + '"><span> ' + order_num[idx].split('_').join(' & ').replace(/ & /g, match => ++t === 2 ? ', ' : match) + ': ' + order_title[idx] + '</span></div>';
       }
     });
     

### 2025-03-14 index.qmd
add rq1 tables

@@
 graphs_dir <- file.path(".", "results", "graphs")
+tables_dir <- file.path(".", "results", "tables")
 
@@
   }
+}
+
+insert_rq1_table <- function(metro) {
+  table_df <- readRDS(file.path(tables_dir, str_c('rq1_table_', metro, '.rds')))
+  table_df[is.na(table_df)] <- NA_real_
+
+  group_headings <- unique(table_df$variable)
+  nrows <- nrow(table_df) / length(group_headings)
+  i <- 1
+
+  k <- kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3)))
+
+  for (g in group_headings) {
+    k <- k %>% pack_rows(g, i, i + nrows - 1, colnum = 1, label_row_css = '')
+    i <- i + nrows
+  }
+
+  k %>% 
+    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
 }
@@
       }
+
+      # Add RQ1 table
+      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
+      writeLines(str_c("#### Racial/ethnic and socioeconomic characteristics of ", heading_text, " Geomarkets over time"))
+
+      writeLines('<div class="table-wrapper freeze-pane">')
+      writeLines(insert_rq1_table(m))
+      writeLines('</div>')
+      writeLines('<div class="footnote">Table Notes:<br>- Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic<br>- Household income measured using 2024 CPI</div>')
 

### 2025-03-14 scripts/rq1_tables_graphs.R

@@ -179,6 +179,33 @@ philly_eps_table %>% print(n=100)
 saveRDS(socal_eps_table, file.path(tables_dir, "socal_eps_table.rds"))
 saveRDS(philly_eps_table, file.path(tables_dir, "philly_eps_table.rds"))
 
+# metros <- c(
+#   "atlanta",
+#   "bay_area",
+#   "boston",
+#   "chicago",
+#   "cleveland",
+#   "dallas",
+#   "dc_maryland_virginia",
+#   "detroit",
+#   "houston",
+#   "long_island",
+#   "los_angeles",
+#   "miami",
+#   "new_york_city",
+#   "northern_new_jersey",
+#   "orange_county",
+#   "philadelphia",
+#   "san_diego"
+# )
+# 
+# for (metro in metros) {
+#   saveRDS(
+#     create_eps_table(all_codes[[metro]]$eps, table_varlist),  # all_codes is from map_functions.R
+#     file.path(tables_dir, str_c('rq1_table_', metro, '.rds'))
+#   )
+# }
+
 ########## FUNCTION TO CREATE (INPUT DATA FRAME FOR) GRAPHS RELEVANT TO RQ1
 
 # Human-readable names for the variables (you can modify this list based on the variables passed in)

### 2025-03-15 index.qmd
small changes to comd

--- 2025-03-13 index.qmd
@@
 
-- Geomarkets carve states and metro areas into smaller geographic regions meant to define local recruiting markets [@RN4982]
+- Geomarkets carve states and metro areas into smaller geographic regions
 
@@
   - regional: send most scores out-of-state but in-region (e.g., Midwest)
-- "Good" colleges should look for regional and national students
-  - Like @RN2247, except College Board creates product from it
+- Selective colleges should look for regional and national students
 - @RN4982 create outputs that become EPS software
 
+<!--   - Like @RN2247, except College Board creates product from it -->
+
 ### Scholarship on Quantification {.hide}
-#### Discussions of correlation and homophily by @RN4975[chapter 1]
+#### Discussions of correlation and homophily by @RN4975
 
@@
     - Data: search filter criteria; and de-identified student-level data
-    - Each prospect has a home and a HS zip code (point geoms)
+    - Each prospect has home and HS zip code (point geometry)
 - Analyze purchases that include all Geomarkets in a metro area

### 2025-03-24 index.qmd
add rq2 tables

@@
 
-  group_headings <- unique(table_df$variable)
-  nrows <- nrow(table_df) / length(group_headings)
-  i <- 1
+  headings <- unique(table_df$variable)
+  headings_group <- rep(nrow(table_df) / length(headings), length(headings))
+  names(headings_group) <- headings
 
-  k <- kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3)))
-
-  for (g in group_headings) {
-    k <- k %>% pack_rows(g, i, i + nrows - 1, colnum = 1, label_row_css = '')
-    i <- i + nrows
+  kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3))) %>% 
+    pack_rows(index = headings_group, colnum = 1, label_row_css = '') %>% 
+    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
 }
 
-  k %>% 
-    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
+insert_rq2_table <- function(metro, table_type, rq_name = 'rq2') {
+  table_df <- readRDS(file.path(tables_dir, str_c(rq_name, '_table_', metro, '.rds')))[[table_type]]
+  table_df[is.na(table_df)] <- NA_real_
+  table_df$order_ids <- as_factor(table_df$order_ids)
+
+  orders <- str_c(if_else(str_detect(unique(table_df$order_ids), '_'), 'Orders ', 'Order '), sub('_', ' & ', unique(table_df$order_ids)), ', ', unique(table_df$test_range))
+  orders_group <- table(table_df$order_ids)
+  names(orders_group) <- orders
+
+  if (table_type == 'race') {
+    tables_group <- rep(nrow(table_df) / length(orders) / length(unique(table_df$table)), length(unique(table_df$table)))
+    names(tables_group) <- c('Count of purchased students', 'Purchased students by geomarket', 'Purchased students by racial/ethnic group')
+
+    return(
+      kbl(
+        table_df %>%
+          mutate(
+            across(all, \(x) format(round(x, 0), big.mark = ',')),
+            across(race_known, ~ if_else(table == 'col_pct_table' & eps_codename != 'All', sprintf('%.1f', .), format(round(., 0), big.mark = ','))),
+            across(white:nhpi, ~ if_else(table == 'count_table' | (table == 'col_pct_table' & eps_codename == 'All'), format(round(., 0), big.mark = ','), sprintf('%.1f', .)))
+          ) %>%
+          select(-table, -order_ids, -test_range),
+        align = 'lrrrrrrrrr', escape = F,
+        col.names = c('', 'All', 'Race known', 'White', 'Asian', 'Black', 'Hispanic', '2+ races', 'AIAN', 'NHPI')
+      ) %>% 
+        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
+        pack_rows(index = rep(tables_group, length(orders)), colnum = 1, label_row_css = '')
+    )
+  } else if (table_type == 'firstgen') {
+    tables_group <- rep(nrow(table_df) / length(orders) / length(unique(table_df$table)), length(unique(table_df$table)))
+    names(tables_group) <- c('Count of purchased students', 'Purchased students by geomarket', 'Purchased students by first-generation status')
+
+    return(
+      kbl(
+        table_df %>%
+          mutate(
+            across(all, \(x) format(round(x, 0), big.mark = ',')),
+            across(known, ~ if_else(table == 'col_pct_table' & eps_codename != 'All', sprintf('%.1f', .), format(round(., 0), big.mark = ','))),
+            across(no_col:not_first, ~ if_else(table == 'count_table' | (table == 'col_pct_table' & eps_codename == 'All'), format(round(., 0), big.mark = ','), sprintf('%.1f', .)))
+          ) %>%
+          select(-unknown, -table, -order_ids, -test_range),
+        align = 'lrrrrr', escape = F,
+        col.names = c('', 'All', 'First-gen known', 'No college', 'Some college', 'Not first-gen')
+      ) %>% 
+        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
+        pack_rows(index = rep(tables_group, length(orders)), colnum = 1, label_row_css = '')
+    )
+  } else if (table_type == 'race_firstgen') {
+    group_df <- table_df %>% group_by(order_ids, stu_race_cb) %>% summarise(n = n())
+    race_group <- group_df$n
+    names(race_group) <- str_replace(str_to_title(group_df$stu_race_cb), 'Or', 'or')
+
+    return(
+      kbl(
+        table_df %>% 
+          mutate(
+            across(all, \(x) format(round(x, 0), big.mark = ',')),
+            across(row_no_col:row_not_first_gen, \(x) sprintf('%.1f', x))
+          ) %>% 
+          select(-stu_race_cb, -order_ids, -test_range),
+        align = 'lrrrr', escape = F,
+        col.names = c('', 'All', 'No college', 'Some college', 'Not first-gen')
+      ) %>% 
+        pack_rows(index = orders_group, colnum = 1, label_row_css = 'background-color: #eaeaea; padding: 10px 7.5px;') %>% 
+        pack_rows(index = race_group, colnum = 1, label_row_css = '')
+    )
+  }
 }
@@
 
+      get_notes <- function(missing_txt = '') {
+        str_replace(
+          str_c('<div class="footnote">', paste0(readLines(file.path(graphs_dir, rq, str_c('rq2_', m, '_race_row_plot_note.txt')), warn = FALSE), collapse = "</br>"), '</div>'),
+          'Excludes students with missing values for race\\. ', missing_txt
+        )
+      }
+
       # We'll do 2 group types: race & firstgen
@@
 
+        # Add RQ2 table
+        writeLines(str_c("### Research question ", rq_num, " {.hide}"))
+        writeLines(if_else(
+          g == 'race',
+          str_c('#### Racial/ethnic composition of purchased student profiles by Geomarket, ', heading_text),
+          str_c('#### First-generation status of purchased student profiles by Geomarket, ', heading_text)
+        ))
+
+        writeLines('<div class="table-wrapper freeze-pane">')
+        writeLines(insert_rq2_table(m, g))
+        writeLines('</div>')
+        writeLines(get_notes())
+
       } # end g loop
@@
 
+      # Add RQ2B table
+      writeLines(str_c("### Research question ", rq_num, " {.hide}"))
+      writeLines(str_c("#### First-generation status by race for ", heading_text, " Geomarkets"))
+
+      writeLines('<div class="table-wrapper freeze-pane">')
+      writeLines(insert_rq2_table(m, 'race_firstgen'))
+      writeLines('</div>')
+      writeLines(get_notes('Excludes students who have missing values for race or parental education. '))
+
       # Add RQ2 map
@@
       writeLines('<p class="fullscreen">Fullscreen</p>')
-      notes_txt <- readLines(file.path(graphs_dir, rq, str_c('rq2_', m, '_race_row_plot_note.txt')), warn = FALSE)
-      writeLines(
-        str_replace(
-          str_c('<div class="footnote">', paste0(notes_txt, collapse = "</br>"), '</div>'),
-          'Excludes students with missing values for race\\. ', ''
-        )
-      )
+      writeLines(get_notes())
 
@@
     )
+
+    # Add RQ2 table
+    if ((g == 'firstgen' && rc == 'col') | (g == 'race' && rc == 'row')) {
+      writeLines(paste0("### ", heading_text, " (AI/AN) {.hide}\n"))
+      writeLines(if_else(
+        g == 'race',
+        str_c('#### Racial/ethnic composition of purchased student profiles by Geomarket, ', heading_text),
+        str_c('#### First-generation status of purchased student profiles by Geomarket, ', heading_text)
+      ))
+
+      writeLines('<div class="table-wrapper freeze-pane">')
+      writeLines(insert_rq2_table(m, g, 'rq2_aian'))
+      writeLines('</div>')
+      writeLines('<div class="footnote">Figure Notes:<br>- Order number 487927 (ordered 7/19/2019): HS class 2020 OR 2021; Ethnicity/Race = American Indian or Alaska Native; SAT 1040-1600; GPA B- to A+.</div>')
+    }
   }

### 2025-03-24 scripts/rq2_tables_graphs_run_functions.R

@@ -56,6 +56,61 @@ source(file = file.path(scripts_dir, 'rq2_tables_graphs_create_functions.R'))
 # 1) Load the CSV into a data frame
 orders_df <- read_csv(file.path(scripts_dir,"metro_orders.csv"))
 
+# get_table_data <- function(metro_df) {
+#   list(
+#     race = do.call(bind_rows, lapply(1:nrow(metro_df), function(i) {
+#       res <- create_sim_eps_race_table(data = lists_orders_zip_hs_df_sf, ord_nums = str_split(metro_df[[i, 'order_ids']], '_')[[1]], eps_codes = get(metro_df[[i, 'eps_codes']]))
+#       
+#       for (r in names(res)) {
+#         names(res[[r]]) <- sub('^stu_|^r_|^c_', '', names(res[[r]]))
+#         if ('known' %in% names(res[[r]])) {
+#           res[[r]] <- res[[r]] %>% select(-race_known) %>% rename('race_known' = 'known')
+#         }
+#         res[[r]]$table <- r
+#       }
+#       
+#       do.call(bind_rows, res) %>% mutate(order_ids = metro_df[[i, 'order_ids']], test_range = metro_df[[i, 'test_range']])
+#     })),
+#     firstgen = do.call(bind_rows, lapply(1:nrow(metro_df), function(i) {
+#       res <- create_sim_eps_firstgen_table(data = lists_orders_zip_hs_df_sf, ord_nums = str_split(metro_df[[i, 'order_ids']], '_')[[1]], eps_codes = get(metro_df[[i, 'eps_codes']]))
+#       
+#       for (r in names(res)) {
+#         if (r %in% c('row_pct_table', 'col_pct_table')) {
+#           res[[r]] <- res[[r]] %>% select(-stu_known)
+#         }
+#         names(res[[r]]) <- sub('^stu_|^r_|^c_', '', names(res[[r]]))
+#         res[[r]]$table <- r
+#       }
+#       
+#       do.call(bind_rows, res) %>% mutate(order_ids = metro_df[[i, 'order_ids']], test_range = metro_df[[i, 'test_range']])
+#     })),
+#     race_firstgen = do.call(bind_rows, lapply(1:nrow(metro_df), function(i) {
[…]

### 2025-03-26 text/front_end_v3.qmd
revising intro, again

@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."  On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."^[On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
+Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and less White.
 
@@
 On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
--->
 
@@
 To the extent that Geomarket borders are correlated with race and class, the Geomarket selections observed here raise concern that Geomarkets -- and other geographic search filters -- may systematically filter working class and non-white students from educational opportunities.
+-->
+Sociology offers several perspectives to describe the process of sorting students into colleges.
+The status attainment model argues that student postsecondary destination is a function of family socioeconomic background, particularly parental education [e.g., @RN915; @RN1273; @RN1268; @RN5033; @RN1300].
+The cultural capital model argues that bourgeois households bestow the pedigree, social networks, and information necessary for their progeny to claim spots at selective colleges [@RN1228; @RN1241; @RN4991] [ADD CITES].
+@RN4873 describes schools as "sorting machines" that create, incorporate, and assign students to categories that determine educational opportunities, often reinforcing reinforcing racial and class inequality.
+Finally, scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
+We argue that sociology has failed to consider how college access is structured by third-party products that identify and sort students on behalf of colleges.
+While third-party recruiting products have become increasingly powerful, granular, and ubiquitous in the era of platform capitalism, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.
 
-The origins of Geomarkets, and how they were incorporated into recruiting products, turns out to be an interesting story.
 In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*.
@@
 x].
-@RN4982 sought to "capture and quantify" (p.
-11) the knowledge of admissions officers,
-
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
-9-10).
-
-@RN4982 developed Geomarkets and the Market Segment Model based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel." Created by analyzing the SAT score-sending behavior of 1980 high school seniors, the thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
+@RN4982 developed Geomarkets and the Market Segment Model as an effort to "capture and quantify" (p.
+11) the knowledge of admissions officers, based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel" (p.
+11).
+Created by analyzing the SAT score-sending behavior of 1980 high school seniors, the thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
 Therefore, colleges should recruit from territories that contain large populations of your target social class.
@@
 
-**Creating Geomarkets and the Market Segment Model.** In 1978, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
+**Creating Geomarkets and the Market Segment Model.**
+
+PUT THIS SOMEWHERE:
+
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
+9-10).
+
+In 1978, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
 x].

### 2025-03-27 text/front_end_v3.qmd
thought of a sentence for the hook paragraph

@@
 Finally, scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
-We argue that sociology has failed to consider how college access is structured by third-party products that identify and sort students on behalf of colleges.
+All of these perspectives share the view that transitions from school to college tend towards social reproduction.
+We argue that these perspectives have ignored an important mechanism of social reproduction: third-party vendors create products that sort students on behalf of college, often utilizing historical snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
 While third-party recruiting products have become increasingly powerful, granular, and ubiquitous in the era of platform capitalism, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.

### 2025-04-04 index.qmd
modifying slides

@@
 ### Introduction {.hide}
+
+- [College Board Student Search Service, order \#448922](./assets/images/geomarket_orders_redacted.pdf){target="_blank"}
+
+Sorting students into colleges, perspectives from sociology
+
+- Status attainment model [e.g., @RN915]
+- Cultural capital model [e.g., @RN1228]
+<!-- - "Schools as sorting machines" [@RN4873] -->
+- Enrollment management behavior of colleges [e.g., @RN3519]
+
+Sociology misses an important mechanism of social reproduction:
+
+- Third-party vendors/products sort students on behalf of colleges
+- Recommendations based on historical snapshots of student demand
+
+We investigate "Geomarkets," an input in two College Board recruiting products
+
+1. What is SES, racial variation between, within Geomarkets? Over time?
+1. What is SES and racial composition of included vs. excluded prospects when student list purchases filter on particular Geomarkets?
+
+<!-- Research interest: how quantification reproduces social stratification -->
+
+### Introduction og {.hide}
 
@@
 :::
+:::
+
+# Scholarship on Enrollment Management
+
+### __The U.S. Market for Higher Education__
+#### _A National Voucher System_
+
+- Federal government
+  - Takes a "steer, don't row" approach
+  - Subsidizes higher education by providing Title IV grants and loans to students
+  - Department of Education regulates Title IV institutions
+- Students
+  - Attend college to improve career, social, and civic outcomes
+  - Pay for college using: 
+    - Household savings; grants and loans from federal, state, private sources
+- Direct providers (Title IV institutions)
+  - A mix of public, private nonprofit, for-profit providers
+  - Funding follows students to whichever institution they enroll in
+    - A national voucher system
+  - Colleges have a financial incentive to enroll students, so competition for students
+
+### __Student Lists are a Match-Making Intermediary__
+#### _Connecting Colleges to Prospective Students_
+
+@RN2247 on emergence of a national higher education system
+
+- Information costs were primary barrier to efficient matches
+  - Students want to attend best college, but don't know options or costs
+  - Colleges want to enroll best students, but don't know who or where they are
+- SAT causes "dramatic fall" in cost of "information about students" (p. 102)
+
+In 1972, College Board began selling lists of prospects to colleges [@belkin2019-studata]
+
+- Student list contains contract info of prospective students who meet "search filter" criteria (e.g., test score GPA) specified by the college
+
+"Lead generation" is process of connecting merchants to potential consumers
+
+- "List-based" lead generation
+  - Based on direct mail business model, but has evolved into "database arketing"
+- "Behavioral-based" targeting
+  - Target users of a platform while they use that platform or a partner platform
+  - Contemporary lead generation in higher education moving here
+
+### __Enrollment Management (EM)__ {.columns-2}
+#### _And the Enrollment Funnel_
+
+::: columns
+
+:::: {.column width=50%}
+
+- EM profession integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]
+- EM office subsumes activities of recruiting, admissions, financial aid
+
+__Interventions along the funnel__
+
+- Convert prospects to leads 
+  - Purchase student lists
+- Convert leads/inquiries to applicants
+  - Email, mail, school visits
+- Convert applicants to admits
+  - Decide who to admit
+- Convert admits to enrollees
+  - Financial aid offers [e.g., @RN2241]
+::::
+
+:::: {.column width=50%}
+```{r, echo=FALSE, out.width="60%", fig.cap=""}
+knitr::include_graphics("./assets/images/enroll_funnelv2.png")
+```
+::::
+
 :::
@@
 
-```{r setup, echo=FALSE, results = 'asis'}
+```{r setup, echo=FALSE, results = 'asis', eval = FALSE}
 

### 2025-04-04 index.qmd
adding background and lit section

--- 2025-03-24 index.qmd
@@
 
-### Introduction {.hide}
+### [__College Board _Student Search Service_, order \#448922__](./assets/images/geomarket_orders_redacted.pdf){target="_blank"} {.hide}
+
+__Sorting students into colleges, perspectives from sociology:__
+
+- Status attainment model [e.g., @RN915]
+- Cultural capital model [e.g., @RN1228]
+<!-- - "Schools as sorting machines" [@RN4873] -->
+- Enrollment management behavior of colleges [e.g., @RN3519]
+
+__Sociology misses an important mechanism of social reproduction:__
+
+- Third-party vendors/products sort students on behalf of colleges
+- Recommendations based on historical snapshots of student demand
+
+__"Geomarkets" are an input in two College Board recruiting products:__
+
+- Student Search Service (est. 1972); Enrollment Planning Service (est. 1984)
+
+__Research questions:__
+
+1. What is SES, racial variation between, within Geomarkets? Over time?
+1. What is SES and racial composition of included vs. excluded prospects when student list purchases filter on particular Geomarkets?
+
+<!--
+
+### Introduction og {.hide}
 
@@
 
-<!-- Research interest: how quantification reproduces social stratification -->
+ -->
 
@@
 
-# Background and Theory
+### Presentation Outline {.hide}
+
+1. Enrollment management: Background \& scholarship
+1. Empirical case and theory
+1. Methods
+1. Results
+1. Discussion
+    - Creating a "knowledge infrastructure" [@RN5025] to study the role of private firms and private equity funding in education
+
+# Enrollment Management: Background \& Scholarship
+
+### __Enrollment Management (EM)__ {.hide .columns-2}
+
+::: columns
+
+:::: {.column width=65%}
+
+__Enrollment management__
+
+- EM profession integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]
+- EM office controls recruiting, admissions, fin aid
+
+__Interventions along enrollment funnel__
+
+- Convert prospects to leads 
+  - Purchase student lists
+- Convert leads/inquiries to applicants
+  - Email, mail, school visits
+- Convert applicants to admits
+  - Holistic review [e.g., @RN4146]
+- Convert admits to enrollees
+  - Financial aid offers [e.g., @RN2241]
+::::
+
+:::: {.column width=35%}
+
+__Enrollment Funnel__
+```{r, echo=FALSE, out.width="85%", fig.cap=""}
+knitr::include_graphics("./assets/images/enroll_funnelv2.png")
+```
+::::
+
+:::
+
+### __Scholarship on Recruiting "in the Wild"__{.hide}
+
+__Sociological ethnographies of enrollment management__
+
+- Private Colleges: visit affluent "feeder schools" to maintain relationships with HS guidance counselors [@RN3519; @RN4407]
+- For-profits: door-to-door visits to local, low-income communities of color [@cottom2017lower]
+
+__Web-scraped the travel schedules of public research and selective private univs__
+
+- Visit rich, mostly White public schools [@RN4758]
+- Disproportionate visits to private schools [@RN5023]
+- "Recruitment redlining": circuitous avoidance of Black and Latinx communities along visit paths [@RN4759] [INSERT CLICKABLE IMAGE OF LA MAP]
+
+___Recruiting not a practice wholly done by colleges...___
+
+- Data vendors and consultancies help colleges find students
+
+### <span style="font-size:0.9em;">__U.S. Higher Education Market: A National Voucher System__</span> {.hide}
+
+__Federal government__
+
+- Takes a "steer, don't row" approach
+- Subsidizes higher education by providing Title IV grants and loans to students
+- Department of Education regulates Title IV institutions
+
+__Students__
+
+- Attend college to improve career, social, and civic outcomes
+- Pay for college using: 
+  - Household savings; grants and loans from federal, state, private sources
+
+__Direct providers (Title IV institutions)__
+
+- A mix of public, private nonprofit, for-profit providers
+- Funding follows students to whichever institution they enroll in
+  - A national voucher system
+- Colleges have a financial incentive to enroll students, so competition for students
+
+### __Student Lists are a Match-Making Intermediary__{.hide}
+
+__@RN2247 on emergence of a national higher education system:__
+
+- Information costs were primary barrier to efficient matches
+  - Students want to attend best college, but don't know options or costs
+  - Colleges want to enroll best students, but don't know who or where they are
+- SAT causes "dramatic fall" in cost of "information about students" (p. 102)
+
+__"Lead generation" is process of connecting merchants to potential consumers__
+
+- "List-based" lead generation based on direct-mail model (legacy)
+  - In 1972, College Board began selling lists of prospects to colleges
+- "Behavioral-based" targeting (contemporary)
+  - Target users of a platform while they use that platform or a partner platform
+
+__Students who opt into College Board _Search_ are [@RN4739]:__
+
+  - 25% more likely to enroll in a four-year university after high school
+  - 31% more likely to graduate college within four years 
+- Magnitude of effects largest for Black, Latinx, and first-gen college students
+
+### __College Board Lists and Student Outcomes__{.hide}
+#### _Howell, Hurwitz, Mabel et al. (2021)_
+
+```{r, fig.align="left"}
+knitr::include_graphics("./assets/images/cb-fig-1.png")
+```
+
+### __The Student List Project__{.hide}
+
+[THIS SLIDE NEEDS LOTS OF WORK; NEED TO SHOW THE SEARCH FILTERS AS PROBLEMATIC]
+Public records requests to public universities in CA, IL, MN, TX, AZ from 2016-2020
+
+- "Order summary" specifying search filter criteria [(Sample)](https://drive.google.com/drive/u/1/folders/18CHMlUIa5OO0zRkET7eSIzYnvk4S9_U1)
+- De-identified prospect-level student list [(Sample)](https://drive.google.com/drive/u/1/folders/1rDIAuHLsh3IfglSsqhNuCmYsfnbtc0tV)
+
+__What is the relationship between using academic and geographic filters and the demographic composition of resulting student lists?__ [@RN5012]
+
+- Analyze descriptive relationships observed in list purchases from College Board and HSLS:09 simulated relationships
+- Findings: Racial (and socioeconomic) inequity results from:
+  - College Board student list databases exclude non-test-takers
+  - Test scores and geographic filters compound disparities in excluded prospects
+
+MOTIVATE THIS STUDY
+
+# Emprical Case and Theory
 
@@
 
+- Geomarkets carve states and metro areas into smaller geographic regions
 - Meant to reproduce the recruiting territories of admissions officers

### 2025-04-10 index.qmd
un-comment out the empirics from the presentation
adding capabilities to slides

@@
   revealjs:
+    revealjs-plugins: [multimodal]   # ← correct key
     theme: [default, ./assets/css/custom.scss]
@@
 
-__Sorting students into colleges, perspectives from sociology:__
+__"Geomarkets" carve states and metro areas into smaller geographic regions__
+
+- Geomarkets are an input in two College Board recruiting products:
+  - Student Search Service (est. 1972); Enrollment Planning Service (est. 1984)
+
+__Perspectives from sociology on sorting students into colleges:__
 
@@
 - Cultural capital model [e.g., @RN1228]
-<!-- - "Schools as sorting machines" [@RN4873] -->
 - Enrollment management behavior of colleges [e.g., @RN3519]
+- Sociology ignores third-party products that sort students on behalf of colleges
+  - Designed to amplify historic inequality in student demand
 
+<!-- 
 __Sociology misses an important mechanism of social reproduction:__
@@
 - Third-party vendors/products sort students on behalf of colleges
-- Recommendations based on historical snapshots of student demand
-
-__"Geomarkets" are an input in two College Board recruiting products:__
-
-- Student Search Service (est. 1972); Enrollment Planning Service (est. 1984)
+  - Amplify past structural inequality in student demand
+ -->
 
@@
 
-1. Enrollment management: Background \& scholarship
-1. Empirical case and theory
-1. Methods
-1. Results
-1. Discussion
-    - Creating a "knowledge infrastructure" [@RN5025] to study the role of private firms and private equity funding in education
+<br>
+
+1. <span style="color:blue">Enrollment management: Background & scholarship</span>  
+1. <span style="color:red">Empirical case and theory</span>  
+1. <span style="color:red">Methods</span>  
+1. <span style="color:red">Results</span>  
+1. <span style="color:blue">Discussion<br>– Creating a "knowledge infrastructure" [@RN5025] to study the role of private firms and private equity funding in education</span>
 
@@
 - Disproportionate visits to private schools [@RN5023]
-- "Recruitment redlining": circuitous avoidance of Black and Latinx communities along visit paths [@RN4759] [INSERT CLICKABLE IMAGE OF LA MAP]
+- "Recruitment redlining": circuitous avoidance of Black and Latinx communities along visit paths [@RN4759] ([LA map](#){data-modal-type="image" data-modal-url="assets/images/recruitment_redlining_map.jpg"})
 
-___Recruiting not a practice wholly done by colleges...___
+___But recruiting process not wholly done by colleges...___
 
-- Data vendors and consultancies help colleges find students
+- Third-party vendors and consultancies help colleges find and recruit students
+<!-- [LA map](#){data-modal-type="image" data-modal-url="assets/images/chicago_geomarkets.png"} -->
 
@@
 MOTIVATE THIS STUDY
+
+- Filters used by public research universities
+- [Link](#){data-modal-type="image" data-modal-url="assets/images/research_univ_list_filters.png"}
 

### 2025-04-11 text/back_end_KS.qmd
RQ1 Chicago

--- 2025-03-27 text/front_end_v3.qmd
@@
 
+```{=tex}
 \pagenumbering{roman}
 \newpage
-
+```
 **ABSTRACT**
@@
 
+```{=tex}
 \newpage
@@
 \pagenumbering{arabic}
-
+```
 # Introduction
@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."^[On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
 Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and less White.
 
+[^1]: On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
+
+```{=html}
 <!--Geomarkets carve states and metroplitan areas into smaller geographic units and have names like,  On the same day, the university purchased two additional orders that were identical, except they filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$.
@@
 -->
+```
 Sociology offers several perspectives to describe the process of sorting students into colleges.
@@
 ```
-
 # Enrollment Management and Quantification
@@
 ```
-
 **Quantification**
@@
 ```
-
 **Correlation**.
@@
 50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"[^1]
+As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"[^2]
 
-[^1]: Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
+[^2]: Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
 For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
@@
 224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^2] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
+This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^3] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
 They then recommended that selective colleges recruit from localities that have a high share of affluent, college-educated households.
 
-[^2]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
+[^3]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
 Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
@@
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^3] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
 We observe similar processes in our case study.
@@
 
-[^3]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
+[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
 
@@
 ```
-
 # The Market Segment Model and College Board Geomarkets
@@
 ```
-
 **Creating Geomarkets and the Market Segment Model.**
@@
 Both outputs are created separately for each Geomarket.
-For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^4] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
-A university could obtain this for their own institution or for a competitor.[^5] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
 These outputs became the basis for the Enrollment Planning Service (EPS) software.
 
-[^4]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
+[^5]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
 
-[^5]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+[^6]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
 This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
@@
 33\] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^6] @RN4982 \[p.
+33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^7] @RN4982 \[p.
 42\] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
@@
 
-[^6]: The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
+[^7]: The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
 Additionally, analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
@@
 **Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^7] Based on these analyses, @RN4982 \[p.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^8] Based on these analyses, @RN4982 \[p.
 46\] state that homophily describes the nature of competition between colleges: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
@@
 
-[^7]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+[^8]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
 The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
@@
 ```
-
 Whereas @RN4982 identified 143 Geomarkets covering the New England, Middle States, and South region, EPS created Geomarkets for the remaining U.S.
@@
 @RN4988 shows the contemporary Geomarkets.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^8] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^9] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
 
-[^8]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
+[^9]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
 Students from California don't all share the same college-going behaviors.
@@
 Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
-EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
+EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^10] Second, EPS users decide which high schools they will visit within selected Geomarkets.
 Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
 
-[^9]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+[^10]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
 Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
@@
 ```
-
 We also expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons.
@@
 A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^10] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
 
-[^10]: For public universities that purchased lists, 80% purchase more than 50,000 names annually.
+[^11]: For public universities that purchased lists, 80% purchase more than 50,000 names annually.
 @RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.
@@
 A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
-After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^11] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^12]
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^12] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^13]
 
-[^11]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
+[^12]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
 Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
 
-[^12]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
+[^13]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
 Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
@@
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^13] This manuscript focuses on Geomarkets.
+College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^14] This manuscript focuses on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 
-[^13]: For example, the geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
+[^14]: For example, the geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
 
@@
 The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-In order to examine how the characteristics of Geomarkets changed over time, we utilize tract-level data from the 2000 Decennial Census, similarly using data from Summary File 1A (SF1A) to obtain measures of race and Hispanic origin and using data from Summary File 3A (SF3A) to obtain measures of income, education, and poverty.[^14] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+In order to examine how the characteristics of Geomarkets changed over time, we utilize tract-level data from the 2000 Decennial Census, similarly using data from Summary File 1A (SF1A) to obtain measures of race and Hispanic origin and using data from Summary File 3A (SF3A) to obtain measures of income, education, and poverty.[^15] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
 These questions were replaced by the American Community Survey (ACS).
-We also utilized data from the 2020 5-year ACS, which includes data collected from 2016-2020.[^15] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+We also utilized data from the 2020 5-year ACS, which includes data collected from 2016-2020.[^16] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
 
-[^14]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
+[^15]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
 
-[^15]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
+[^16]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
 
@@
 Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^16] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^17] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
 Using these zip code-level data, we utilized the `aggregate()` function to create shapefile for each Geomarket.
 
-[^16]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
+[^17]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
 
@@
 We also created socioeconomic measures that approxmiate the predictors of market segment identified by @RN4982.
-These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^17] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
+These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^18] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
 
-[^17]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
+[^18]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
 
@@
 We draw from these data to answer RQ2.
-Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^18].
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^19].
 Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
 
-[^18]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
+[^19]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
 
@@
 
-**Research design**.
-We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
+**Research design**. We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
+
 Following XXX \[CITE\], we focus on a small number \[HOW MANY\] of metropolitan areas in order to provide sufficient analytic depth while also situating each case within the context of historic segregation and contemporary gentrification.
@@
 
+# Findings
+
+## Chicago
+
+Figure X presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
+Chicago Geomarkets follow a classic geographical structure of an inner city core surrounded by a ring of suburbs, as described by @RN4982 in the development of The Market Segment Model.
+Northern suburb Geomarkets-Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie- experienced relatively small population growth from 1980 through 2020.
+The greatest population growth across the metropolitan area from 1980 through 2020 occurred in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
+The City of Chicago, which encapsulates the center core of the metropolitan area, was the only Geomarket to experience population declines.
+Despite remaining the most densely populated Geomarket, The City of Chicago population decreased from 3.1 Million in 1980 to 2.8 Million in 2020.
+
+These population changes have resulted in shifting racial and ethnic demographics within and across Geomarkets over time.
+Figure X illustrates the racial and ethnic composition of each of the seven Chicago Geomarkets across three census periods.
+In 1980, all Geomarkets were predominantly White, ranging from 83.1% in the South and Southwest Suburbs to approximately 93.3% in the North Shore—except for the City of Chicago, where the White population made up only 45.5%.
+By 2020, many Geomarkets that were previously predominantly White shifted to a nearly balanced distribution between White residents and Residents of Color, including Chain of Lakes (58.7% White, 7.6% Black, 23.5% Hispanic; 7.4% Asian Pacific Islander), Northwest Suburbs (61.5% White, 3% Black, 18.2% Hispanic; 14.8% Asian Pacific Islander), Evanston and Skokie (63.7% White, 6.7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (55.6% White, 9.3% Black, 23.4% Hispanic; 9.2% Asian Pacific Islander), and South and Southwest Suburbs (53.4% White, 23.8% Black, 17.9% Hispanic; 2.8% Asian Pacific Islander).
+For many Geomarkets, this balance in racial/ethnic demographics was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7.4 percentage point increase for the Evanston and Skokie Geomarket to a 19.2 percentage point increase for Western Suburbs.
+Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
+
+Figure X also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
+For instance, North Shore experienced some increases in Hispanic (2.2.% in 1980 to 7.9% in 2020), Asian Pacific Islander (6.6% in 2000 to 9.9% in 2020), and Black (1.8.% in 1980 to 2.6% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
+City of Chicago also experienced relatively modest declines in its Black population (36.6% in 1980 to 32.9% in 2020), alongside increases in its Hispanic (14.8% in 1980 to 25.8% in 2020) and Asian Pacific Islander populations (4.0% in 2000 to 6.1% in 2020), further solidifying it as a majority resident of color Geomarket.
+
+We also explore socioeconomic demographics - median income, percent living below poverty, and educational attainment - within and across Geomarkets over time in Figure X.
+Overall, median income both within and across Geomarkets in the Chicago metropolitan area remained relatively stable across the three census periods.
+For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$158,000 in the North Shore Geomarket to a low of \$66,000 in the City of Chicago Geomarket.
+However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence—\$120,000 in the Northwest Suburbs, \$119,000 in Evanston and Skokie, \$108,000 in Chain of Lakes, \$105,000 in the Western Suburbs, and \$99,000 in the South and Southwest Suburbs.
+While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$79,000) and decrease in the South and Southwest Suburbs Geomarket (\$93,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$14,000.
+
+Figure X also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
+Similar patterns are evident in educational attainment across Geomarkets.
+In 1980, the City of Chicago had the lowest percentage of residents with a bachelor's degree or higher (12.5%).
+However, by 2020, it experienced the largest percentage point increase, reaching 37.6%.
+In comparison, all other Geomarkets experienced smaller gains in the share of residents with a bachelor’s degree or higher over the same period.
+These trends contributed to a narrowing of the educational attainment gap across Geomarkets from 1980 to 2020.
+
+## Dallas
+
+## Los Angeles
+
 # Discussion
@@
 
+```{=tex}
 \appendix
 \renewcommand{\thetable}{A\arabic{table}}
-
+```
 # Appendix A

### 2025-04-13 text/back_end_KS.qmd
Add figs to findings

@@
 
-Figure X presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
+@fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
 Chicago Geomarkets follow a classic geographical structure of an inner city core surrounded by a ring of suburbs, as described by @RN4982 in the development of The Market Segment Model.
@@
 These population changes have resulted in shifting racial and ethnic demographics within and across Geomarkets over time.
-Figure X illustrates the racial and ethnic composition of each of the seven Chicago Geomarkets across three census periods.
+@fig-chicago-rq1-race illustrates the racial and ethnic composition of each of the seven Chicago Geomarkets across three census periods.
 In 1980, all Geomarkets were predominantly White, ranging from 83.1% in the South and Southwest Suburbs to approximately 93.3% in the North Shore—except for the City of Chicago, where the White population made up only 45.5%.
@@
 
-Figure X also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
+@fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
 For instance, North Shore experienced some increases in Hispanic (2.2.% in 1980 to 7.9% in 2020), Asian Pacific Islander (6.6% in 2000 to 9.9% in 2020), and Black (1.8.% in 1980 to 2.6% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
@@
 
-We also explore socioeconomic demographics - median income, percent living below poverty, and educational attainment - within and across Geomarkets over time in Figure X.
+We also explore socioeconomic demographics - median income, percent living below poverty, and educational attainment - within and across Geomarkets over time in @fig-chicago-rq1-ses.
 Overall, median income both within and across Geomarkets in the Chicago metropolitan area remained relatively stable across the three census periods.
@@
 
-Figure X also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
+@fig-chicago-rq1-ses also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
 Similar patterns are evident in educational attainment across Geomarkets.
@@
 
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospecs across the entire state of Illinois.
+Two of these orders applied the same filters: relatively lower SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
+The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
+Because all six orders included an Illinois state filter, we can use prospects within these orders coming from the Chicago metropolitan area and simulate student lists if Geomarket filters were applied across low, mid, and high range score thresholds.
+
+For instance, @fig-chicago-rq2-lowSAT shows the contributions of each Geomarket to the pool of purchased prospect profiles from the two orders targeting lower SAT score ranges (1020-1150) by race/ethnicity.
+The top panel presents the full distribution of the 15,981 total prospects included in these orders by Geomarkets: 11% reside in Chain of Lakes, 10% in Northwest Suburbs, 2% in North Shore, 4% in Evanston and Skokie, 20% in City of Chicago, 25% in Western Suburbs, and 26% in South and Southwest Suburbs.
+These overall distributions are compared to those in the lower panels of Figure X, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
+
+For instance, nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White.
+With the exception of the City of Chicago, @fig-chicago-rq2-lowSAT shows all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
+The City of Chicago Geomarket, however, contributed only 5% of White prospects, despite accounting for 20% of all included prospects.
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, exept for Evanston and Skokie (12% versus 4% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (10% versus 26% for all prospects) contributing smaller shares relative to their overall representation.
+
+In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT.
+Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all prospects whereas all other Geomarkets contributed substantially smaller proportions.
+The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions than their overall representation among all included prospects.
+
 ## Dallas
@@
 knitr::include_graphics('./../assets/images/enroll_funnelv2.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-maps, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarkets, Total Population 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/chicago_maps.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Chicago Area Geomarkets, 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq1_chicago_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Chicago Area Geomarkets, 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq1_chicago_ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-lowSAT, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, low SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT.png')
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-04-14 text/back_end_KS.qmd
rq2 ses results

@@
 
+### Geomarket Changes Over Time \[Need better RQ1 heading\]
+
 @fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
@@
 
+### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]
+
 To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospecs across the entire state of Illinois.
@@
 
-For instance, @fig-chicago-rq2-lowSAT shows the contributions of each Geomarket to the pool of purchased prospect profiles from the two orders targeting lower SAT score ranges (1020-1150) by race/ethnicity.
-The top panel presents the full distribution of the 15,981 total prospects included in these orders by Geomarkets: 11% reside in Chain of Lakes, 10% in Northwest Suburbs, 2% in North Shore, 4% in Evanston and Skokie, 20% in City of Chicago, 25% in Western Suburbs, and 26% in South and Southwest Suburbs.
-These overall distributions are compared to those in the lower panels of Figure X, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
-
-For instance, nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White.
-With the exception of the City of Chicago, @fig-chicago-rq2-lowSAT shows all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
+For instance, @fig-chicago-rq2-lowSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting lower SAT score ranges (1020-1150) by race/ethnicity.
+The top panel presents the full distribution of the 15,981 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% reside in Chain of Lakes, 10% in Northwest Suburbs, 2% in North Shore, 4% in Evanston and Skokie, 20% in City of Chicago, 25% in Western Suburbs, and 26% in South and Southwest Suburbs.
+These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-lowSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
+
+Nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White.
+With the exception of the City of Chicago, @fig-chicago-rq2-lowSAT-race shows all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
 The City of Chicago Geomarket, however, contributed only 5% of White prospects, despite accounting for 20% of all included prospects.
@@
 
-In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT.
+In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race.
 Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all prospects whereas all other Geomarkets contributed substantially smaller proportions.
 The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions than their overall representation among all included prospects.
+
+@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores.
+Both figures demonstrate how representational Geomarket patterns remain relatively consistent SAT scores increase.
+For instance, @fig-chicago-rq2-midSAT-race shows that the Chain of Lakes, North Shore, and Western Suburbs Geomarkets contributed similar proportions of White and Asian students, yet smaller proportions of Black and Hispanic students, relative to their overall representation among all included prospects.
+The Northwest Suburbs and Evanston and Skokie similarly contributed a disproportionately higher share of Asian prospects, but lower shares of Black and Hispanic students.
+Conversely, the South and Southwest Suburbs Geomarket contributed a substantially larger proportion of Black students, while the City of Chicago yielded larger proportions of both Black and Hispanic students- again, relative to their overall representation among all included prospects targeted at middle range SAT scores.
+Despite declines in the overall number of Black (n=240) and Hispanic (n=950) prospects included in orders filtering for high-ranges of SAT scores (1310-1600), @fig-chicago-rq2-highSAT-race shows overall Geomarket contribution patterns persist.
+
+We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
+Similar to race/ethnicity figures above, @fig-chicago-rq2-lowSAT-ses presents the Geomarket distribution for the 16,060 total prospects (whose first-generations status is known) included in low-range SAT score orders (1020-1150).
+Among the 3,479 prospects whose parents did not attend college, the City of Chicago stands out as the only Geomarket that contributing a disproportionately larger share of first-generation students - 34% in comparison to 20% for all prospects in this SAT range.
+In contrast, all other Geomarkets contributed disproportionately smaller shares, ranging from a one-percentage-point difference (Chain of Lakes, Northwest Suburbs, North Shore) to a five-percentage-point difference (Western Suburbs).
+For prospects whose parents attended some college but did not complete their degrees, @fig-chicago-rq2-lowSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
+These representational patterns are reversed among prospects that are not first-generationcollege students.
+While all other Geomarkets contributed either comparable (Northwest Suburbs, South and Southwest Suburbs) or disproportionately larger shares (Chain of Lakes, North Sore, Evanston and Skokie, Western Suburbs), the City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (13%) relative to all prospects (20%) in this SAT range.
+
+@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status become increasingly pronounced in orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores, respectively.
+At the middle SAT score range, the City of Chicago Geomarket contributes a substantially larger share of first-generation students whose parents did not attend college—38%, compared to 17% of all prospects in this range, representing a 21-percentage-point disparity.
+This disproportionality increases to a 25-percentage-point difference at the highest SAT score thresholds, where the City of Chicago accounts for 40% of such first-generation students, despite representing only 15% of all prospects in this score range.
 
@@
 
-```{r fig-chicago-rq2-lowSAT, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, low SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT.png')
+```{r fig-chicago-rq2-lowSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Low-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-highSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, High-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-lowSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Low-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-highSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, High-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_ses.png')
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-04-14 text/back_end_KS.qmd
sesXrace Chicago

@@
 
+Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity.
+@fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150).
+For example, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% had parents who attended some college but did not complete their degree, and 67% were not first-generation college students.
+However, @fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed their degrees (29% and 28%, respectively).
+In contrast, Geomarkets such as Chain of Lakes (72%), North Shore (88%), Evanston and Skokie (73%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
+
+@fig-chicago-rq2-lowSAT-combo indicates these Geomarket patterns persist across all other racial/ethnic categories.
+The City of Chicago consistently contributed a disproportionately larger share of first-generation college students who parents did not attend college across all racial/ethnic categories.
+However, the magnitude of this disproportionality varied- from a 3-percentage-point difference for Black prospects to a 21 percentage-point difference for Asian prospects.
+While Asian, Black, and Hispanic prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.
+
 ## Dallas
@@
 knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-lowSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Low-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_combo.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_combo.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-chicago-rq2-highSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, High-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_combo.png')
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-04-14 index.qmd
remove .ds store files

@@
 
-[THIS SLIDE NEEDS LOTS OF WORK; NEED TO SHOW THE SEARCH FILTERS AS PROBLEMATIC]
-Public records requests to public universities in CA, IL, MN, TX, AZ from 2016-2020
+__Public records requests to public univs in CA, IL, MN, TX, AZ, 2016-2020__
 
-- "Order summary" specifying search filter criteria [(Sample)](https://drive.google.com/drive/u/1/folders/18CHMlUIa5OO0zRkET7eSIzYnvk4S9_U1)
-- De-identified prospect-level student list [(Sample)](https://drive.google.com/drive/u/1/folders/1rDIAuHLsh3IfglSsqhNuCmYsfnbtc0tV)
+- for each purchase, request: (1) [order summary](https://drive.google.com/drive/u/1/folders/18CHMlUIa5OO0zRkET7eSIzYnvk4S9_U1); (2) [prospect-level data](https://drive.google.com/drive/u/1/folders/1rDIAuHLsh3IfglSsqhNuCmYsfnbtc0tV)
 
@@
 
-- Analyze descriptive relationships observed in list purchases from College Board and HSLS:09 simulated relationships
-- Findings: Racial (and socioeconomic) inequity results from:
+- Student list products are "selection devices" [@RN4778]
+- Utilize "racialized inputs"[@RN4786], disadvantage non-white communities because they have been historically excluded from input (e.g., zip codes, AP)
+
+__Analyze who included in purchases: (1) simulated, [HSLS:09](https://nces.ed.gov/surveys/hsls09/); (2) College Board__
+
+- Racial inequity results from:
   - College Board student list databases exclude non-test-takers
   - Test scores and geographic filters compound disparities in excluded prospects
-
-MOTIVATE THIS STUDY
-
-- Filters used by public research universities
-- [Link](#){data-modal-type="image" data-modal-url="assets/images/research_univ_list_filters.png"}
+    - [SAT scores + Segment map, LA](#){data-modal-type="image" data-modal-url="assets/images/la_segment_map.png"}
+    - CRYSTAL -- REPLACE W/ STATIC MAP THAT JUST HAS LA
+- Collge Board creates new geographic filters that include/exclude ([FIGURE](#){data-modal-type="image" data-modal-url="assets/images/research_univ_list_filters.png"})
 
@@
 
-### Background{.hide}
-#### Geomarkets and Market Segment Model created by @RN4982
+### *The Structure of College Choice* [@RN4982] {.hide}
 
-<p class="quote">Geomarkets</p>
+> Admissions officers are tellers of stories – about the colleges they represent, about the colleges they attended, about each other, and about the vagabond life of college recruiting...We begin with this celebration of sotrytelling...[because] the intuitions of admissions officers comprise a remarkably systematic body of knowledge about the college selection process (p.
+9-10).
+
+Goal to "capture and quantify" (p.
+11) knowledge of admissions officers, based on idea that "a good recruiter knows where to look for prospective applicants, as seen in the students’ willingness or reluctance to travel" (p.
+11)
+
+Creates three geographic levels: region, state, community (Geomarket)
+
+<p class="quote">"initial task was to define enrollment markets in a manner consistent with admissions officers’ intuitive understanding of student pools" (p. 4)</p>
+<p style="display:none;">'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can’t visit without calling first.'</p>
+
+<p class="quote">Geomarkets carve states and metro areas into smaller geographic regions</p>
 <p style="display:none;">Boundaries [often] match formal political and educational divisions, reflecting natural channels of communication.
@@
 
-- Geomarkets carve states and metro areas into smaller geographic regions
 - Meant to reproduce the recruiting territories of admissions officers
-- @RN4982 has 143 Geomarkets in 3 regions; EPS has 304
 
-**Market Segment Model**
+<!--   College Board published the book, provided data, and financed the project, as part of their “efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state” (Zemsky & Oedel, 1983, p.
+x).
+-->
 
-- Predicts how demand for a college varies by Geomarket, based on characteristics of households in Geomarket
+### __Creating the Market Segment Model__{.hide}
+
+College Board provides data, funding, and publishes book
+
+-  as part their "efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state (Zemsky & Oedel, 1983, p.
+x).
+
+**Market Segment Model predicts how demand for a college varies by Geomarket**
+
+> We sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
+The Market Segment Model we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups (p.
+4).
+
 - Categorize students into 4 market segments -- local, in-state, regional, national -- based on SAT score-sending behavior
@@
 
-<!--   - Like @RN2247, except College Board creates product from it -->
+<!--   
+The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways
+
+- Like @RN2247, except College Board creates product from it 
+  - *Market Segment Profile*; *Institutional Profile*-->
 

### 2025-04-15 index.qmd
worked on front end of presentation

@@
   - Test scores and geographic filters compound disparities in excluded prospects
-    - [SAT scores + Segment map, LA](#){data-modal-type="image" data-modal-url="assets/images/la_segment_map.png"}
+    - [SAT scores + Segment, LA map](#){data-modal-type="image" data-modal-url="assets/images/la_segment_map.png"}
     - CRYSTAL -- REPLACE W/ STATIC MAP THAT JUST HAS LA
@@
 
+__Project goal: "capture and quantify" (p. 11) knowledge of admissions officers__
+
 > Admissions officers are tellers of stories – about the colleges they represent, about the colleges they attended, about each other, and about the vagabond life of college recruiting...We begin with this celebration of sotrytelling...[because] the intuitions of admissions officers comprise a remarkably systematic body of knowledge about the college selection process (p.
-9-10).
+9-10).<br><br>A good recruiter knows where to look for prospective applicants, as seen in the students’ willingness or reluctance to travel (p.
+11).
 
-Goal to "capture and quantify" (p.
-11) knowledge of admissions officers, based on idea that "a good recruiter knows where to look for prospective applicants, as seen in the students’ willingness or reluctance to travel" (p.
-11)
+<p class="quote">Initial task "was to define enrollment markets in a manner consistent with admissions officers’ intuitive understanding of student pools" (p. 4)</p>
+<p style="display:none;">'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can’t visit without calling first.' (p.
+11)</p>
 
-Creates three geographic levels: region, state, community (Geomarket)
-
-<p class="quote">"initial task was to define enrollment markets in a manner consistent with admissions officers’ intuitive understanding of student pools" (p. 4)</p>
-<p style="display:none;">'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can’t visit without calling first.'</p>
+- Creates three geographic levels: region, state, community (Geomarket)
 
@@
 **Each major metro area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.** In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful.
-[@RN4982, p.
-4].</p>
+[@RN4982, pp.
+11-12].</p>
 
@@
 
-<!--   College Board published the book, provided data, and financed the project, as part of their “efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state” (Zemsky & Oedel, 1983, p.
-x).
--->
-
 ### __Creating the Market Segment Model__{.hide}
 
-College Board provides data, funding, and publishes book
-
--  as part their "efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state (Zemsky & Oedel, 1983, p.
-x).
+> We sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
+The Market Segment Model we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
+The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways (p.
+4).
 
 **Market Segment Model predicts how demand for a college varies by Geomarket**
-
-> We sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
-The Market Segment Model we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups (p.
-4).
 
@@
   - Selective colleges should look for regional and national students
-- @RN4982 create outputs that become EPS software
 
-<!--   
-The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways
+<p class="quote">College Board underwrites project to enter enrollment management space</p>
+<p style="display:none;">efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state [@RN4982, p.
+x]</p>
 
-- Like @RN2247, except College Board creates product from it 
-  - *Market Segment Profile*; *Institutional Profile*-->
+- @RN4982 outputs are basis for College Board EPS software
 

### 2025-04-15 text/back_end_KS.qmd
results chicago

@@
 Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity.
-@fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150).
-For example, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% had parents who attended some college but did not complete their degree, and 67% were not first-generation college students.
-However, @fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed their degrees (29% and 28%, respectively).
+@fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150) orders.
+For instance, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% had parents who attended some college but did not complete their degree, and 67% were not first-generation college students.
+@fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed degrees (29% and 28%, respectively).
 In contrast, Geomarkets such as Chain of Lakes (72%), North Shore (88%), Evanston and Skokie (73%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
 
-@fig-chicago-rq2-lowSAT-combo indicates these Geomarket patterns persist across all other racial/ethnic categories.
-The City of Chicago consistently contributed a disproportionately larger share of first-generation college students who parents did not attend college across all racial/ethnic categories.
+@fig-chicago-rq2-lowSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
+The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
 However, the magnitude of this disproportionality varied- from a 3-percentage-point difference for Black prospects to a 21 percentage-point difference for Asian prospects.
-While Asian, Black, and Hispanic prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.
+Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares.
+Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.
+
+These Geomarket contribution patterns become more pronounced as SAT score thresholds increase.
+For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders.
+The City of Chicago Geomarket still contributes a disproportionately larger share of first -generation college students - both those whose parents did not attend college and those whose parents attended some college but did not complete a degree - across most racial and ethnic groups.
+However, the magnitude of this disproportionately for City of Chicago becomes even greater for Asian students, with a nearly 30-percentage-point difference collectively for all first-generation college students (35% no college and 22% some college) relative to their overall Geomarket representation (28%).
+Similar to patterns observed at lower SAT score thresholds, the Geomarket contributions to first-generation Hispanic college students remain relatively balanced.
+All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation college students across both parental education categories to the overall prospect pool.
+
+@fig-chicago-rq2-highSAT-combo shows Geomarket patterns become even more pronounced at the highest SAT score range (1310-1600), largely driven by the relatively fewer number of first-generation college students included within these orders.
+Again, the City of Chicago Geomarket disproportionately contributes a disproportionate share of first-generation Asian college students (43% across both parental education categories) and first-generation Hispanic college students (52% across both parental education categories) in comparison to the overall distribution of first-generation college students within each racial/ethnic category (13% for Asian and 40% for Hispanic, respectively).
+Conversely, nearly all other Geomarkets contribute a disproportionately larger share of non-first generation college students across all racial/ethnic categories.
 

### 2025-04-15 text/back_end_KS.qmd
chicago results
smooth rq1 results chi

@@
 @fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
-Chicago Geomarkets follow a classic geographical structure of an inner city core surrounded by a ring of suburbs, as described by @RN4982 in the development of The Market Segment Model.
-Northern suburb Geomarkets-Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie- experienced relatively small population growth from 1980 through 2020.
-The greatest population growth across the metropolitan area from 1980 through 2020 occurred in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
-The City of Chicago, which encapsulates the center core of the metropolitan area, was the only Geomarket to experience population declines.
-Despite remaining the most densely populated Geomarket, The City of Chicago population decreased from 3.1 Million in 1980 to 2.8 Million in 2020.
-
-These population changes have resulted in shifting racial and ethnic demographics within and across Geomarkets over time.
-@fig-chicago-rq1-race illustrates the racial and ethnic composition of each of the seven Chicago Geomarkets across three census periods.
-In 1980, all Geomarkets were predominantly White, ranging from 83.1% in the South and Southwest Suburbs to approximately 93.3% in the North Shore—except for the City of Chicago, where the White population made up only 45.5%.
-By 2020, many Geomarkets that were previously predominantly White shifted to a nearly balanced distribution between White residents and Residents of Color, including Chain of Lakes (58.7% White, 7.6% Black, 23.5% Hispanic; 7.4% Asian Pacific Islander), Northwest Suburbs (61.5% White, 3% Black, 18.2% Hispanic; 14.8% Asian Pacific Islander), Evanston and Skokie (63.7% White, 6.7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (55.6% White, 9.3% Black, 23.4% Hispanic; 9.2% Asian Pacific Islander), and South and Southwest Suburbs (53.4% White, 23.8% Black, 17.9% Hispanic; 2.8% Asian Pacific Islander).
-For many Geomarkets, this balance in racial/ethnic demographics was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7.4 percentage point increase for the Evanston and Skokie Geomarket to a 19.2 percentage point increase for Western Suburbs.
+The maps in the figure show how Chicago Geomarkets reflect a classic urban geographical structure - an inner-city core surrounded by a ring of suburbs - as described by @RN4982 in the development of The Market Segment Model.
+Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area -Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
+In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
+The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most densely populated Geomarket.
+
+These population shifts have contributed to changes in the racial/ethnic composition within and across Geomarkets over time.
+@fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarkets across the three census periods.
+In 1980, all Geomarkets —except for the City of Chicago - were predominantly White, with proportions ranging from 83.1% in the South and Southwest Suburbs to approximately 93.3% in North Shore.
+In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 45.5% of its total population.
+By 2020, several Geomarkets that were previously predominantly White shifted to a more balanced racial/ethnic distribution, including Chain of Lakes (58.7% White, 7.6% Black, 23.5% Hispanic; 7.4% Asian Pacific Islander), Northwest Suburbs (61.5% White, 3% Black, 18.2% Hispanic; 14.8% Asian Pacific Islander), Evanston and Skokie (63.7% White, 6.7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (55.6% White, 9.3% Black, 23.4% Hispanic; 9.2% Asian Pacific Islander), and South and Southwest Suburbs (53.4% White, 23.8% Black, 17.9% Hispanic; 2.8% Asian Pacific Islander).
+For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7.4 percentage point increase for the Evanston and Skokie Geomarket to a 19.2 percentage point increase for Western Suburbs.
 Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
@@
 For instance, North Shore experienced some increases in Hispanic (2.2.% in 1980 to 7.9% in 2020), Asian Pacific Islander (6.6% in 2000 to 9.9% in 2020), and Black (1.8.% in 1980 to 2.6% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
-City of Chicago also experienced relatively modest declines in its Black population (36.6% in 1980 to 32.9% in 2020), alongside increases in its Hispanic (14.8% in 1980 to 25.8% in 2020) and Asian Pacific Islander populations (4.0% in 2000 to 6.1% in 2020), further solidifying it as a majority resident of color Geomarket.
+City of Chicago also experienced relatively modest declines in its Black population (36.6% in 1980 to 32.9% in 2020), alongside increases in its Hispanic (14.8% in 1980 to 25.8% in 2020) and Asian Pacific Islander populations (4.0% in 2000 to 6.1% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
 
@@
 In 1980, the City of Chicago had the lowest percentage of residents with a bachelor's degree or higher (12.5%).
-However, by 2020, it experienced the largest percentage point increase, reaching 37.6%.
+However, by 2020, the Geomarket experienced the largest percentage point increase in college degree attainment, reaching 37.6%.
 In comparison, all other Geomarkets experienced smaller gains in the share of residents with a bachelor’s degree or higher over the same period.
@@
 
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospecs across the entire state of Illinois.
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospects across the entire state of Illinois.
 Two of these orders applied the same filters: relatively lower SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
 The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
-Because all six orders included an Illinois state filter, we can use prospects within these orders coming from the Chicago metropolitan area and simulate student lists if Geomarket filters were applied across low, mid, and high range score thresholds.
+Because all six orders included an Illinois state filter, we can utilize prospects from the Chicago metropolitan area within the resulting student lists to examine Geomarket contributions by simulating the application of a Geomarket filter across low, mid, and high range score threshold orders.
 
@@
 
-These Geomarket contribution patterns become more pronounced as SAT score thresholds increase.
+Geomarket contribution patterns become more pronounced as SAT score thresholds increase.
 For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders.

### 2025-04-15 text/back_end_KS.qmd
smooth rq2 results chi

@@
 To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospects across the entire state of Illinois.
-Two of these orders applied the same filters: relatively lower SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
-The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
-Because all six orders included an Illinois state filter, we can utilize prospects from the Chicago metropolitan area within the resulting student lists to examine Geomarket contributions by simulating the application of a Geomarket filter across low, mid, and high range score threshold orders.
+Two of these orders applied the same filters: SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
+The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with relatively middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
+Because all six orders included an Illinois state filter, we can utilize prospects from the Chicago metropolitan area within the resulting student lists to examine Geomarket contributions by simulating the application of a Geomarket filter across low-, middle-, and high-range SAT threshold orders.
 
@@
 In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race.
-Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all prospects whereas all other Geomarkets contributed substantially smaller proportions.
-The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions than their overall representation among all included prospects.
+Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
+The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
 
 @fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores.
-Both figures demonstrate how representational Geomarket patterns remain relatively consistent SAT scores increase.
+Both figures demonstrate how representational Geomarket patterns across race/ethnicity remain relatively consistent even as SAT scores increase.
 For instance, @fig-chicago-rq2-midSAT-race shows that the Chain of Lakes, North Shore, and Western Suburbs Geomarkets contributed similar proportions of White and Asian students, yet smaller proportions of Black and Hispanic students, relative to their overall representation among all included prospects.
@@
 Similar to race/ethnicity figures above, @fig-chicago-rq2-lowSAT-ses presents the Geomarket distribution for the 16,060 total prospects (whose first-generations status is known) included in low-range SAT score orders (1020-1150).
-Among the 3,479 prospects whose parents did not attend college, the City of Chicago stands out as the only Geomarket that contributing a disproportionately larger share of first-generation students - 34% in comparison to 20% for all prospects in this SAT range.
-In contrast, all other Geomarkets contributed disproportionately smaller shares, ranging from a one-percentage-point difference (Chain of Lakes, Northwest Suburbs, North Shore) to a five-percentage-point difference (Western Suburbs).
+Among the 3,479 total prospects whose parents did not attend college, the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 34% in comparison to 20% for all included prospects in this SAT range.
+In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students relative to all included prospects, with such disparity ranging from a one-percentage-point difference (Chain of Lakes, Northwest Suburbs, North Shore) to a five-percentage-point difference (Western Suburbs).
 For prospects whose parents attended some college but did not complete their degrees, @fig-chicago-rq2-lowSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
-These representational patterns are reversed among prospects that are not first-generationcollege students.
-While all other Geomarkets contributed either comparable (Northwest Suburbs, South and Southwest Suburbs) or disproportionately larger shares (Chain of Lakes, North Sore, Evanston and Skokie, Western Suburbs), the City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (13%) relative to all prospects (20%) in this SAT range.
+These representational patterns are reversed among prospects that are not first-generation college students.
+The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (13%) relative to all prospects (20%) in this SAT range, whereas all other Geomarkets contributed either comparable (Northwest Suburbs, South and Southwest Suburbs) or disproportionately larger shares (Chain of Lakes, North Sore, Evanston and Skokie, Western Suburbs).
 
 @fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status become increasingly pronounced in orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores, respectively.
-At the middle SAT score range, the City of Chicago Geomarket contributes a substantially larger share of first-generation students whose parents did not attend college—38%, compared to 17% of all prospects in this range, representing a 21-percentage-point disparity.
-This disproportionality increases to a 25-percentage-point difference at the highest SAT score thresholds, where the City of Chicago accounts for 40% of such first-generation students, despite representing only 15% of all prospects in this score range.
-
-Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity.
+At the middle SAT score range in @fig-chicago-rq2-midSAT-ses, the City of Chicago Geomarket contributes a substantially larger share of first-generation students whose parents did not attend college—38%, compared to 17% of all included prospects in this range, representing a 21-percentage-point disparity.
+This disproportionality increases to a 25-percentage-point difference at the highest SAT score thresholds in @fig-chicago-rq2-highSAT-ses, where the City of Chicago accounts for 40% of such first-generation students, despite representing only 15% of all included prospects in this score range.
+
+Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
 @fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150) orders.
-For instance, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% had parents who attended some college but did not complete their degree, and 67% were not first-generation college students.
+For instance, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% were first-generation college students whose parents attended some college but did not complete their degree, and 67% were not first-generation college students.
 @fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed degrees (29% and 28%, respectively).
@@
 However, the magnitude of this disproportionality varied- from a 3-percentage-point difference for Black prospects to a 21 percentage-point difference for Asian prospects.
-Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares.
+Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
 Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.
 
-Geomarket contribution patterns become more pronounced as SAT score thresholds increase.
+Geomarket contribution patterns to prospect pools by both first-generation college student status and race/ethnicity become more pronounced as SAT score thresholds increase.
 For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders.
@@
 Similar to patterns observed at lower SAT score thresholds, the Geomarket contributions to first-generation Hispanic college students remain relatively balanced.
-All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation college students across both parental education categories to the overall prospect pool.
-
-@fig-chicago-rq2-highSAT-combo shows Geomarket patterns become even more pronounced at the highest SAT score range (1310-1600), largely driven by the relatively fewer number of first-generation college students included within these orders.
+All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation Hispanic college students across both parental education categories to the overall prospect pool.
+
+@fig-chicago-rq2-highSAT-combo shows the magnitude of Geomarket contribution disparities increases at the highest SAT score range (1310-1600), partly driven by the relatively fewer number of first-generation college students included within these orders.
 Again, the City of Chicago Geomarket disproportionately contributes a disproportionate share of first-generation Asian college students (43% across both parental education categories) and first-generation Hispanic college students (52% across both parental education categories) in comparison to the overall distribution of first-generation college students within each racial/ethnic category (13% for Asian and 40% for Hispanic, respectively).

### 2025-04-15 text/back_end_KS.qmd
rq1 dallas maps

@@
 
-## Dallas
+## Dallas-Fort Worth
+
+### Geomarket Changes Over Time \[Need better RQ1 heading\]
+
+The six Geomarkets within the Dallas-Fort Worth metropolitan area are illustrated in @fig-dallas-maps by total population across the three census periods.
+The Dallas-Fort Worth region represents a unique metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982.
+As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
+These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas Co, Collin & Rockwall Co, and West of Dallas/Ft.
+Worth Metroplex).
+
+From 1980 to 2020, the population of the overall Dallas-Fort Worth metropolitan area increased substantially from approximately 2.7 Million to 6.8 Million.
+@fig-dallas-maps showcases all six Geomarkets experienced population growth over the three census periods.
+However, the most pronounced increases occured in the West of Dallas/Ft.
+Worth Metroplex and Collin and Rockwall Co Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
+These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers - City of Forth Worth and City of Dallas - into the surrounding suburban and rural areas of the metropolitan areas.
+This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket.
 
@@
 knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_combo.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png
+```
+
+```{r fig-dallas-maps, echo = FALSE, fig.align = 'center', fig.cap = "Dallas-Forth Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/dallas_maps.png')
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-04-16 index.qmd
working on presentation

@@
 
-- EM profession integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]
+- EM profession integrates techniques from marketing and economics to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p. xiv]
 - EM office controls recruiting, admissions, fin aid
@@
 ### <span style="font-size:0.9em;">__U.S. Higher Education Market: A National Voucher System__</span> {.hide}
+#### Motivating the role of student lists in college access
 
@@
 
-### Discussion{.hide}
-#### Contribution to scholarship
+## EM Industry Dynamics
 
-Scholarship on quantification (in education)
+### __Dynamics in the Market for Student List Data__{.hide}
 
-- Hasn't examined quantification that classifies consumers for producers
+[Jaquette, O., Salazar, K., & Martin, P. M. (2022). *The student list business*.](https://ticas.org/wp-content/uploads/2022/09/The-Student-List-Business_-Primer-and-Market-Dynamics.pdf)
+
+1. __Centrality of consulting firms to EM industry__
+    - In our data collection, ~50\% of univs outsourced student list purchases
+1. __Technology and new data sources to identify+serve prospects__
+    - "Free" college search engines (e.g., Cappex)
+    - College planning software purchased by k-12 districts (e.g., Naviance, Scoir)
+1. __Test optional and moves by incumbents__
+    - Test-optional reduces coverage advantage of testing companies
+    - ACT becomes an edtech firm (epic fail); College Board has AP+savvy
+1. __From owner-operated to private-equity owned firms__
+    - 1990s - 2000s = firm entry; 2010s = vertical+horizontal acquisitions
+      - [Vista Equity Partners portfolio](https://www.vistaequitypartners.com/companies/portfolio/)
+      - [EAB](https://eab.com/) + [Ruffalo Noel Levitz](https://www.ruffalonl.com/) claim to serve more than 3,000 colleges
+    - Consulting firms sell software-as-service "solutions"
+      - Build subscription recruiting platforms around database of prospects
+
+<!-- This is a comment in the Quarto document body
+
+### __Enrollment Management (EM) Consulting Firms__{.hide}
+#### _From Owner-Operated to Private Equity_
+
+- Domains of EM consulting along the student lifecycle
+  - Marketing/recruiting, pricing and financial aid, student success, advancement
+  - Colleges hire EM firms to choose tuition price and "discounting" strategy
+    - Typically based on modeling behavior of past applicants
+
+Evolution of EM consultancies mirrors evolution of for-profit colleges [@list_biz; @RN5027; @RN5022]
+
+- 1980s and 1990s: rise of owner-operated firms
+  - In 1983, Bill Royall founds Royall & Co., direct marketing for political campaigns
+- 2010s: acquisitions and market concentration from private equity investment
+  - In 2015, Advisory Board Company purchases Royall & Co. for $850 million (2015 CPI), Royall renamed EAB
+  - In 2017, Vista Equity Partners buys EAB for $1.5 billion
+- By 2022:
+  - EAB claims to serve more than 1,100 colleges [@eabcomm]
+  - Ruffalo Noel Levitz (RNL) claims to serve 1,900 orgs [@rnlabout]
+
+ -->
+
+### __The EAB Story: Acquire and Integrate__{.hide}
+#### _Acquistiions in lead generation [@list_biz; @RN5027]_
+
+Goal of acquisitions is to make colleges depend on EAB (software) products
+
+- [YouVisit](https://www.youvisit.com/), 2019 -- virtual tours
+- Cappex (now [appily](https://www.appily.com/)), 2020 -- college search engine
+- Hobsons, 2021 -- edtech/consulting firm
+  - PowerSchool buys Naviance; [EAB becomes exclusive reseller of Intersect](#){data-modal-type="image" data-modal-url="assets/images/intersect_connection_subscription_renewal_notice_university_utah_1_7_2022_redacted.png"}
+    - (PowerSchool IPO in 2021; acquired by Bain Capital in 2024)
+- [Concourse](https://discover.concourseglobal.net/student/start), 2022 -- direct admissions
+
+EAB creates [Enroll360](https://eab.com/solutions/enroll360/) in 2021:
+
+<blockquote style="font-size: 0.9em;">
+We spent the last couple of years creating a connected recruitment ecosystem....
+This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit....Individually, each solution can solve important challenges at various stages of the enrollment funnel...By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021].
+</blockquote>
+
+### __The Private Equity Business Model__ {.hide .columns-2}
+#### _Software Subscriptions and Upselling_
+
+::: columns
+
+:::: {.column width=50%}
+
+[Vista Equity Partners.
+(2021).
+_The next frontier of software investment and the private markets_](https://info.vistaequitypartners.com/rs/vistaequitypartnersiiillc/images/vista-equity-partners-whitepaper-uk-next-frontier-software-investment-private-markets.PDF)
+```{r, echo=FALSE, out.width="80%", fig.cap=""}
+knitr::include_graphics("./assets/images/vista_software_value_proposition.png")
+```
+
+::::
+
+:::: {.column width=50%}
+
+[Securities and Exchange Commission.
+(2022).
+_Form 10-K: PowerSchool Holdings, Inc.
+for fiscal year ended Dec.
+31, 2021_](https://s27.q4cdn.com/190453437/files/doc_financials/2021/q4/78e2074c-196d-4e96-b3e6-e471d3ac452e.pdf), p.
+43:
+
+<blockquote style="font-size: 0.9em;">
+Many customers begin their journey with us by using only two of our 15 products...
+As customers begin to appreciate the benefits of an integrated software platform..., they increase the number of solutions they buy from us...
+Our future revenue growth is dependent upon our ability to expand our customers’ use of our platform, and our go-to-market efforts are designed to drive cross-sell growth (p.
+43).
+</blockquote>
+::::
+
+:::
+
+## Organizational Field of College Access
+
+### __Organizational Fields [@RN527]__{.hide}
+
+<p class="quote">"Organizational field" salient to college access, according to prior research</p>
+<p style="display:none;">Those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products.
+The virtue of this unit of analysis is that it directs our attention...to the totality of relevant actors [@RN527, p.
+148].</p>
+
+- Economics $\rightarrow$ effects of schools and policies (district/state/federal) on students
+- Sociology $\rightarrow$ how schools and households sort students into colleges
+  - Status attainment model [e.g., @RN915]
+  - Cultural capital model [e.g., @RN1228; @RN4991]
+  - Enrollment management behavior of colleges [e.g., @RN3519]
+
+__Research ignores third-party vendors that classify consumers for producers__
+
 - Structural inequality $\rightarrow$ unequal student demand for college
   - Third-party products take snapshot student demand $\rightarrow$ encourage supply-side to amplify structural inequalities observed on demand-side
-- @RN5005 show effect of quantification, but not mechanisms of *how* quantification produces inequality
+- Vendors have structured college access for decades [@RN5045]
+  - Private equity + data science + software-as-service $\rightarrow$ financiers controlling process of sorting students to colleges
+- Organizational field should be focus of regulation
+
+## Building a Knowledge Infrastructure
+
+### The "Knowledge Infrastructure" of Education Research{.hide}
+
+__Knowledge infrastructures (KIs) [@RN5025]__
+
+- "Systems of observation and measurement" [@RN5025, p. 743] that provide the foundation for scalable research
+- Collect, process, and distribute data in ways that "enable certain kinds of knowledge production while channeling researchers away from questions not readily answerable within...that infrastructure” [@RN5025, p.
+742]
+
+__The KI of education policy research__
+
+- Data on students and schools from district, state, and federal sources
+- Researchers investigated for-profit colleges because for-profit colleges (direct providers) and students included in ongoing data collections
+
+__Expanding the KI to consider private firms and private equity__
+
+- Case-study research
+  - Granular data; critical analyses; gnarly data collection; difficult to scale
+- Secondary data subscriptions
+  - Can these data be leveraged to
+
+### ya ya ya {.hide}
 

### 2025-04-16 text/back_end_KS.qmd
dallas maps rq1

@@
 @fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarkets across the three census periods.
-In 1980, all Geomarkets —except for the City of Chicago - were predominantly White, with proportions ranging from 83.1% in the South and Southwest Suburbs to approximately 93.3% in North Shore.
-In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 45.5% of its total population.
-By 2020, several Geomarkets that were previously predominantly White shifted to a more balanced racial/ethnic distribution, including Chain of Lakes (58.7% White, 7.6% Black, 23.5% Hispanic; 7.4% Asian Pacific Islander), Northwest Suburbs (61.5% White, 3% Black, 18.2% Hispanic; 14.8% Asian Pacific Islander), Evanston and Skokie (63.7% White, 6.7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (55.6% White, 9.3% Black, 23.4% Hispanic; 9.2% Asian Pacific Islander), and South and Southwest Suburbs (53.4% White, 23.8% Black, 17.9% Hispanic; 2.8% Asian Pacific Islander).
-For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7.4 percentage point increase for the Evanston and Skokie Geomarket to a 19.2 percentage point increase for Western Suburbs.
+In 1980, all Geomarkets —except for the City of Chicago - were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
+In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
+By 2020, several Geomarkets that were previously predominantly White shifted to a more balanced racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
+For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
 Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
@@
 @fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
-For instance, North Shore experienced some increases in Hispanic (2.2.% in 1980 to 7.9% in 2020), Asian Pacific Islander (6.6% in 2000 to 9.9% in 2020), and Black (1.8.% in 1980 to 2.6% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
-City of Chicago also experienced relatively modest declines in its Black population (36.6% in 1980 to 32.9% in 2020), alongside increases in its Hispanic (14.8% in 1980 to 25.8% in 2020) and Asian Pacific Islander populations (4.0% in 2000 to 6.1% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
+For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
+City of Chicago also experienced relatively modest declines in its Black population (37% in 1980 to 33% in 2020), alongside increases in its Hispanic (15% in 1980 to 26% in 2020) and Asian Pacific Islander populations (4% in 2000 to 6% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
 
@@
 Similar patterns are evident in educational attainment across Geomarkets.
-In 1980, the City of Chicago had the lowest percentage of residents with a bachelor's degree or higher (12.5%).
-However, by 2020, the Geomarket experienced the largest percentage point increase in college degree attainment, reaching 37.6%.
+In 1980, the City of Chicago had the lowest percentage of residents with a bachelor's degree or higher (13%).
+However, by 2020, the Geomarket experienced the largest percentage point increase in college degree attainment, reaching 38%.
 In comparison, all other Geomarkets experienced smaller gains in the share of residents with a bachelor’s degree or higher over the same period.
@@
 This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket.
+
+For Dallas-Fort Worth, Figure X illustrates the racial and ethnic composition of each Geomarket across the three census periods.
+In 1980, all Geomarkets were predominantly White, with proportions ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Ft.
+Worth Metroplex.
+However, by 2000, Geomarkets located in the central part of the metropolitan area shifted to a more balanced racial/ethnic distribution, including Dallas Co (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
+For the City of Dallas Geomarket,
 

### 2025-04-16 text/back_end_KS.qmd
rq1 dallas

@@
 As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
-These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas Co, Collin & Rockwall Co, and West of Dallas/Ft.
+These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
 Worth Metroplex).
@@
 However, the most pronounced increases occured in the West of Dallas/Ft.
-Worth Metroplex and Collin and Rockwall Co Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
+Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
 These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers - City of Forth Worth and City of Dallas - into the surrounding suburban and rural areas of the metropolitan areas.
@@
 
-For Dallas-Fort Worth, Figure X illustrates the racial and ethnic composition of each Geomarket across the three census periods.
-In 1980, all Geomarkets were predominantly White, with proportions ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Ft.
+Figure X presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
+In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Ft.
 Worth Metroplex.
-However, by 2000, Geomarkets located in the central part of the metropolitan area shifted to a more balanced racial/ethnic distribution, including Dallas Co (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
-For the City of Dallas Geomarket,
+By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
+In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 - 78% and 80%, respectively.
+Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.
+
+These shifts in racial and ethnic composition within and across Geomarkets became more pronounced between 2000 and 2020.
+The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents.
+However, both Geomarkets remained predominantly White (58% and 63%, respectively) relative to all other Geomarkets.
+In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas CO (33% White, 23% Black, 32% Hispanic; 9% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian Pacific Islander), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian Pacific Islander).
+Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
+
+Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in Figure X for the Dallas-Fort Worth metropolitan area.
+Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
+Between 1980 to 2000, affluence levels across Geomarkets were consistent, ranging from a low of \$82,000 in City of Fort Worth to a high of \$135,00 in the Collin & Rockwall Counties Geomarket.
+However, the substantial increase in median income for the Collin & Rockwall Counties Geomarket during this time period plateaued by 2020, reshuffling relatively affluence levels among Geomarkets.
+By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Ft.
+Worth Metroplex (\$116,000), and Dallas County (\$90,000).
+In contrast, median household income for City of Dallas and City of Fort Worth remain relatively stable and continued to rank at the lower levels of affluence within the metropolitan area.
+\[KS: NEED TO CHECK THIS; I THINK CITY OF DALLAS AND CITY OF FORT WORTH ARE MIXED UP BETWEEN GRAPH VERSUS TABLE\]
+
+Figure X also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
+In 1980, poverty rates were comparably low in both for Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Ft.
+Worth Metroplex Geomarkets, and higherst for City of Dallas and City of Fort Worth Geomarkets.
+Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Ft.
+Worth Metroplex, while increasing for all other Geomarkets.
+These divergent trends resulted in a widening gap in poverty levels across the Dallas-Fort Worth metropolitan area.
+A similar widening disparity is evident evident in educational attainment.
+Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23.1% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
 

### 2025-04-16 index.qmd
modest prez

@@
 <p class="quote">College Board underwrites project to enter enrollment management space</p>
-<p style="display:none;">efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state [@RN4982, p.
+<p style="display:none;">We asked the College Board if it might provide the database we required.
+Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state [@RN4982, p.
 x]</p>
@@
 
-### Discussion{.hide}
-#### Summary of results
+### Summary of results{.hide}
 
@@
 
-- Case-study research
+- Case-study research (e.g., this study)
   - Granular data; critical analyses; gnarly data collection; difficult to scale
-- Secondary data subscriptions
-  - Can these data be leveraged to
-
-### ya ya ya {.hide}
-
-Scholarship on postsecondary "organizational field"
-
-- Soc-of-ed focuses on schools (and policies) as "sorting machines"
-  - But third-party vendors have structured college access for decades
-- Org field is "totality of relevant actors" [@RN527, p. 148]
-  - Organizational field should be focus of regulation
-  - Sociology should identify high-leverage actors/products in org field
+- Data subscriptions created for investment community (e.g., PitchBook, Orbis)
+  - Can these data be leveraged to expand the domain of education research?
 

### 2025-04-16 index.qmd
update la_segment_map.png

@@
     - [SAT scores + Segment, LA map](#){data-modal-type="image" data-modal-url="assets/images/la_segment_map.png"}
-    - CRYSTAL -- REPLACE W/ STATIC MAP THAT JUST HAS LA
 - Collge Board creates new geographic filters that include/exclude ([FIGURE](#){data-modal-type="image" data-modal-url="assets/images/research_univ_list_filters.png"})

### 2025-04-18 index.qmd
included the results in the presentation
workin on discussion

@@
 
-```{r setup, echo=FALSE, results = 'asis', eval = FALSE}
+```{r setup, echo=FALSE, results = 'asis', eval = TRUE}
 
@@
 
+## Organizational Field of College Access
+
+### __Organizational Fields [@RN527]__{.hide}
+
+<p class="quote">"Organizational field" salient to college access, according to prior research</p>
+<p style="display:none;">Those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products.
+The virtue of this unit of analysis is that it directs our attention...to the totality of relevant actors [@RN527, p.
+148].</p>
+
+<!-- - Economics $\rightarrow$ effects of schools and policies (district/state/federal) on students -->
+
+__Sociology $\rightarrow$ how schools and households sort students into colleges__
+
+- Demand-side explanations
+  - Status attainment model [e.g., @RN915]
+  - Cultural capital model [e.g., @RN1228; @RN4991]
+- Supply-side explanations
+  - The "credentialism" literature [e.g., @RN865; @RN861]
+  - Enrollment management behavior of colleges [e.g., @RN3519]
+
+__Research ignores third-party vendors that classify consumers for producers__
+
+- Structural inequality $\rightarrow$ unequal student demand for college
+  - Third-party products take snapshot student demand $\rightarrow$ encourage supply-side to amplify structural inequalities observed on demand-side
+- Vendors have structured college access for decades [@RN5045]
+  - Private equity + data science + software-as-service $\rightarrow$ financiers control which students sorted to which colleges
+
 ## EM Industry Dynamics
@@
 - [YouVisit](https://www.youvisit.com/), 2019 -- virtual tours
-- Cappex (now [appily](https://www.appily.com/)), 2020 -- college search engine
+- [Cappex](https://www.appily.com/), 2020 -- college search engine
 - Hobsons, 2021 -- edtech/consulting firm
@@
 
-## Organizational Field of College Access
-
-### __Organizational Fields [@RN527]__{.hide}
-
-<p class="quote">"Organizational field" salient to college access, according to prior research</p>
-<p style="display:none;">Those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products.
-The virtue of this unit of analysis is that it directs our attention...to the totality of relevant actors [@RN527, p.
-148].</p>
-
-- Economics $\rightarrow$ effects of schools and policies (district/state/federal) on students
-- Sociology $\rightarrow$ how schools and households sort students into colleges
-  - Status attainment model [e.g., @RN915]
-  - Cultural capital model [e.g., @RN1228; @RN4991]
-  - Enrollment management behavior of colleges [e.g., @RN3519]
-
-__Research ignores third-party vendors that classify consumers for producers__
-
-- Structural inequality $\rightarrow$ unequal student demand for college
-  - Third-party products take snapshot student demand $\rightarrow$ encourage supply-side to amplify structural inequalities observed on demand-side
-- Vendors have structured college access for decades [@RN5045]
-  - Private equity + data science + software-as-service $\rightarrow$ financiers controlling process of sorting students to colleges
-- Organizational field should be focus of regulation
-
 ## Building a Knowledge Infrastructure

### 2025-04-18 text/discussion.qmd
workin on discussion

---
title: "Structuring College Access: The Market Segment Model and College Board Geomarkets"
author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
bibliography: ./../assets/bib/cb_geomarket_bib.bib
csl: ./../assets/bib/apa.csl
format:
  pdf:
    documentclass: article
    #papersize: letter
    number-sections: true
    number-depth: 2
    #toc: FALSE
    #pandoc_args: !expr rmdfiltr::add_wordcount_filter(rmdfiltr::add_citeproc_filter(args = NULL))
    #keep_tex: true
    extensions: [andrewheiss/wordcount]
editor: visual
geometry:
  margin=1in
fontsize: 11pt
header-includes:
  - \usepackage{setspace}
  - \doublespacing
---

```{r setup, include=FALSE}
library(knitr)
#library(bookdown)

# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
library(kableExtra)

#hey ho This many words: {{< words-total >}}
```

# Discussion

The discussions of correlation and homophily by @RN4975 describe the logic of The Market Segment model and also how the enrollment management industry thinks about the process of recruiting students. The thesis of the Market Segment Model is that of the status attainment model, that student demand for college is a function of parental education and parental income. @RN4982 demonstrated this thesis by showing the correlation between student SAT score-sending behavior and measures of class. @RN4982 argued that households of particular class and college aspirations are likely to live in particular parts of a metropolitan area and that these geographic territories can be meaningfully captured by Geomarkets. This argument relies on the assumption of homophily, that birds of a feather flock together. Drawing from this logic, colleges should identify their core student market segment -- local, in-state, regional, or national -- identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets. This information is contained in the Market Segment Profile (Appendix A), which is a standard output of EPS software. When considering new Geomarkets to target, colleges should consider which Geomarkets have large numbers of households that are interested in colleges that are direct competitors. This information is contained in the Institutional Profile (Appendix A). Thus, the recommendations of the Market Segment Model can be boiled down to finding those Geomarkets with your desired class of student and targeting those high schools and communities that are popular with colleges like yours.
[WHAT IS MISSING HERE: YOU DON'T TALK ABOUT HOW A SNAPSHOT OF THE PAST IS USED TO MAKE DECISIONS ABOUT FUTURE RECRUITING]

Interestingly, @RN4982 claim that the Market Segment Model merely formalizes the logic that admissions officers from selective colleges had been utilizing for decades. The idea is that good students from good families want to go to good colleges that are far from home. These good families are found in particular parts -- Geomarkets -- of metropolitan areas and students from other parts of the metro have the achievement and disposition to attend their local college. The Market Segment Model formalized this logic and commodified it in Enrollment Planning Service (EPS) software that helped colleges identify desirable Geomarkets and desirable high schools within those Geomarkets. As suggested by @RN5007, once commodified and quantified, the local knowledge of admissions officers becomes redundant.

Technologies that target target customers based on geography utilize racial and class segregation [@RN4775; @RN4849; @RN4975]. Geographic segmentation has become increasingly granular over time. Geomarkets were a landmark innovation in Geographic segmentation in the context of education. College Board provided data for @RN4982 because "the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" (p. x).

Research question 1 investigates racial and socioeconomic variation between and within Geomarkets over time. We find that Geomarkets are highly correlated with race and class. For examle, KARINA GIVE A COUPLE OF FINDINGS.

These findings are not surprising given the extent of race and class segregation in the US. Furthermore, the correlation between Geomarkets and class is expcted because geomarket borders were created in service of the Market Segment Model, a model which views demand for higher education as a function of class. Interestingly, @RN4982 literally have nothing to say about race and demand for higher education. @RN4982 is written around New England, with a particular focus on Geomarkets in Connecticut and Massachusetts. Their silence is about race is conspicuous given that Busing was a political contentious issue in Boston and nearby metropolitan areas.

We show that in 1980, when Geomarkets were being created, that Black people tended to be highly concentrated in the poorest Geomarket in the metropolitan area. Examples include, CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), TX19 -- City of Dallas, CA21 -- South & South Central Los Angeles, and PA5 -- Philadelphia County. In a handful of metropolitan areas (e.g., Cleveland), Black people remain conentrated in the poorest Geomarkets. In Chicago-Land,

Some metro areas follow a pattern whereby Black residents [KARINA -- ONE OR TWO SENTENCES]

Research question 2 analyzes the socioeconomic and racial composition of included versus excluded prospects under the hypothetical scenario that a student list purchase

Research question 2 analyzes the socioeconomic and racial composition of student list purchases that include all Geomarkets in a metro area. We examine which Geomarkets these prospects live in as a means of assessing the consequences of student list purchases that exclude particular Geomarkets. This question is motivated by the @RN4982 recommendation that selective colleges that enroll students from the regional and national market segment should target affluent Geomarkets, while community colleges and non-selective 4-year colleges should focus on middle- and working-class Geomarkets [OTHER RATIONALE FOR THIS RQ FROM BOBBY?].

KARINA -- ADD ONE OR TWO PARAGRAPHS SUMMARIZING RESULTS.

These analyses contribute to scholarship about how students are sorted into colleges. The status attainment model argues that college destination is a function of parental education and occupation [@RN1273; @RN1271; @RN1272]. Indeed, the Market Segment Model can be seen as a market research application of the status attainment model. @RN5033 finds that Asian American students -- particularly children of Chinese, Indian, Korean, and Vietnamese immigrants -- tend to have high levels of educational achievement even if their parents do not, while the educational achievement of later-generation Asian Americans conforms to status attainment theory. In our analyses of prospects included in student list purchases, Asian American prospects from poor Geomarkets tended to be first-generation students while Asian American prospects from affluent Geomarkets tended to have parents with a BA. Drawing from @RN5033, one explanation for this finding is that Asian students living in low-income Geomarkets tend to have parents who immigrated to the U.S. whereas Asian students living in affluent Geomarkets are more likely to have parents who were born in the U.S.

The cultural capital model explains a process by which upper and upper-middle class families sort themselves into selective colleges by providing their children with the pedigree (academic, extracurricular) valued by selective colleges, information about how to navigate the admissions gauntlet, and social networks that provide an inside track [@RN1228; @RN1241; @RN4991; @RN1814]. The cultural capital model explains how affluent households maintain a disproportionate enrollment share at selective institutions in an era of holistic admissions that has the purported mission of increasing racial and class diversity. <!--Amidst concern about discrimination against Asian students at Ivy Plus institutions, @RN4991 shows how private admissions consultants coach Asian American students to......NEED TO REVISE THIS .-->

Both the status attainment model and the cultural capital model are demand-side explanations for why education is a "social sieve" [@RN2377; @RN5046] that allows for a modicum of social mobility while maintaining a much larger flow of intergenerational class transmission [@RN861]. The relative sizes of these flows can be observed in @RN4722, which obtained parental income from federal income tax returns for every college in the U.S. The disproportionate enrollment share of high income families at selective private colleges -- and even at most public research universities -- is staggering. @RN1297[p. 241-242] observes that educational credentials are property that "support their holders' claims for intermarriages with notable families\ldots, claims to monopolize socially and economically advantageous positions" and the claims made by educational credentials "are a setback for talent (charisma) in favor of property" that comes from "the desire for restricting the supply for these positions and their monopolization by the owners of educational certificates."

Complementing the cultural capital explanation of how students are sorted into colleges are the supply-side explanations of the credentialism literature and the more recent literature on enrollment management. The credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required for socially and economically advantageous positions [@RN865; @RN861; @RN880; @RN869]. The enrollment management literature describes which prospects colleges want to enroll and what colleges do to get these students, at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758; @RN3544; @RN3547; @cottom2017lower]. @RN3519 and @RN4321 describe a tacit arrangement between high school guidance counselors on the demand side and college admissions officers on the supply side. High school guidance counselors at well-resourced schools, especially private schools, are motivated to give their students a competitive advantage in admissions. Meanwhile, college admissions officers are motivated to enroll students who can afford full tuition price and are likely to donate in the future. These mutually beneficial desires are consummated by recognzing that upper and upper-middle class applicants satisfy the extracurricular needs of the college. The band needs oboists. The lacrosse team needs players. Quantitative analyses of visits from colleges to high schools are consistent with @RN3519; selective private colleges and public research universities devote most of their recruiting resources on courting students from privileged schools and communities [@RN5023; @RN4758]. However, the extant enrollment management literature is flawed in that it views recruiting as something that is done by individual colleges.

We argue that existing explanations how students are sorted into colleges miss an important supply-side mechanism, third-party vendors that sort students on behalf of colleges. Based on a snapshot of student demand from 1980, @RN4982 created the Market Segment Model and Geomarkets. @RN4982 argues that demand for higher education is correlated with class, ingoring the historical structural barriers that produced class- and race-based inequality in student demand. In itself, this explanation is basically the same as the status attainment model. However, the College Board goes further by commodifying the Market Segment Model into products that tell colleges which Geomarkets to target. Therefore, third-party products encourage supply-side enrollment management behaviors to amplify historical structural inequalities observed on the demand-side. Geomarkets were incorporated into the Student Search Service product. We show that that excluding low-income, non-white Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-white students with strong test scores.

## Enrollment Management Industry

The growing salience of third-party vendors warrants a reconfiguration of the organizational field that is salient to college access. @RN527[p. 148] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities. @RN3547 shows how local external stakeholders intervene on admissions standards. @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side. College access scholarship from economics analyzes the effects of policies (district/state/federal) and the effects of information interventions [CITE HOXBY TYPE]. None of these explanations seriously consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.

@list_biz describes four key dynamics in the enrollment management (EM) industry, with a focus on the market for student list data. First, EM consulting firms are central to the creation and implementation of recruiting campaigns [@RN5027]. When we issued records requests to public universities about their student list purchases, at least 50\% of universities indicated that they outsourced student list purchases to an EM consultancy []. In many cases, the university was buying names each year but no university employee had knowledge about which student list vendors they were buying names from. This reliance on EM consultancies is partially explained by the high level of burnout and employee turnover in the EM profession [@hoover_em].

Second, technological advances created means of identifying and serving prospects. College Board began selling the contact information of test-takers in 1972 [@belkin2019-studata]. ACT followed suit and the two testing companies enjoyed a near duopoly for several decades. This business can be described as list-based lead generation based on the direct-mail model. "Free" college search engines (e.g., Cappex, Niche) yielded news sources of student list data. Another source of student list data is college planning software purchased by high school districts and utilized by high school students and guidance counselors. The most widely-known product is Naviance, which claims to be used by more than used by more than 10 million K-12 students and by 40% ofUS high schools [@powerschool]. *Naviance* college planning software feeds into the *Intersect* recruiting platform, which allows colleges to target Naviance users while they are on the platform. *Intersect* is an example of behavioral-based -- as opposed to list-based -- targeting. However, key to both approaches is the extent to which the product is built around a large, proprietary database of prospects. This can be observed in the justification for the purchase of Intersect from a University of Utah procurement document [@utah_procurement]:

> The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to. There is unique group of prospective students who are only in the PowerSchool Naviance platform. We have spoken with several vendors where we can reach out to prospective students. Each vendor has their own proprietary website or database where select students can be reached.

The third dynamic in the EM industry has been the growth of test-optional movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT. ACT responded to uncertainty in their core testing business by attempting to become an edtech company, failing miserably, and being acquired by Nexus Capital.

The fourth dynamic is the transformation from owner-operated to private-equity owned firms and the blurring between consultancy and software vendors. The 1980s and 1990s was a period of market entry in EM, mostly consistent of college admissions professionals or professors who decided to hang up a shingle [@RN5027]. The 2010s were a period of acquisitions and concentration. After regulatory scrutiny pushed private-equity interests out of for-profit college market [@RN5022], private equity investors saw value in acquiring firms that provided services to direct providers. Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions grow market share and vertical acquisitions designed to increase customer reliance.

EAB is the exemplar private equity darling. START HERE ON MONDAY. THE SLIDE IS

The EAB Story: Acquire and Integrate
Acquistiions in lead generation (Jaquette, Salazar, and Martin 2022; Marcus 2024)
Consulting firms sell software-as-service “solutions”
Build subscription recruiting platforms around database of prospects
In their analysis of quantifying school quality in England, @RN5005 [p. 517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model argues that student demand for higher education is mostly a function of social origin.

Institutional theory defines the organizational field as "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services and products" [@RN527, p. 143].

# References

::: {#refs}
:::

### 2025-04-21 text/back_end_KS.qmd
adding figs

@@
 In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race.
-Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares - 48% and 34%, respectively- compared to their overall contributions of 20% and 26% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
+Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 34%, respectively, compared to their overall contributions of 20% and 26% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
 The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
@@
 For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders.
-The City of Chicago Geomarket still contributes a disproportionately larger share of first -generation college students - both those whose parents did not attend college and those whose parents attended some college but did not complete a degree - across most racial and ethnic groups.
+The City of Chicago Geomarket still contributes a disproportionately larger share of first-generation college students for both those whose parents did not attend college and those whose parents attended some college but did not complete a degree across most racial and ethnic groups.
 However, the magnitude of this disproportionately for City of Chicago becomes even greater for Asian students, with a nearly 30-percentage-point difference collectively for all first-generation college students (35% no college and 22% some college) relative to their overall Geomarket representation (28%).
@@
 The Dallas-Fort Worth region represents a unique metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982.
-As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
+As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
 These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
@@
 By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
-In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 - 78% and 80%, respectively.
+In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
 Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.
@@
 
+### Geomarket Changes Over Time \[Need better RQ1 heading\]
+
+The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in Figure X by total population across the three census periods.
+The Los Angeles region is one of the only multi-nuclei cases metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982.
+As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
+These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
+Worth Metroplex).
+
 # Discussion
@@
 
-```{r fig-chicago-rq2-lowSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Low-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_race.png')
+```{r fig-chicago-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Chicago Area", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_race.png')
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 
-```{r fig-chicago-rq2-highSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, High-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_race.png')
-#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
-#![The enrollment funnel](assets/images/enrollment_funnel.png)
-```
-
-```{r fig-chicago-rq2-lowSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Low-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_ses.png')
-#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
-#![The enrollment funnel](assets/images/enrollment_funnel.png)
-```
-
 ```{r fig-chicago-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
@@
 
-```{r fig-chicago-rq2-highSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, High-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_ses.png')
-#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
-#![The enrollment funnel](assets/images/enrollment_funnel.png)
-```
-
-```{r fig-chicago-rq2-lowSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Low-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_lowSAT_combo.png')
-#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
-#![The enrollment funnel](assets/images/enrollment_funnel.png)
-```
-
 ```{r fig-chicago-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
@@
 
-```{r fig-chicago-rq2-highSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, High-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_highSAT_combo.png')
-#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
-#![The enrollment funnel](assets/images/enrollment_funnel.png
-```
-
 ```{r fig-dallas-maps, echo = FALSE, fig.align = 'center', fig.cap = "Dallas-Forth Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
 knitr::include_graphics('./../assets/images/dallas_maps.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-dallas-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Dallas Area Geomarkets, 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq1-dallas-race.png')
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-04-21 text/back_end_KS.qmd
dallas figs

@@
 
+```{r fig-dallas-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Dallas Area Geomarkets, 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq1-dallas-ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-dallas-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Dallas Area", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2-dallas-race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-dallas-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_dallas_midSAT_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+
+```
+
+```{r fig-dallas-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-dallas-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_dallas_midSAT_combo.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
 \clearpage

### 2025-04-21 text/back_end_KS.qmd
LA figs

@@
 
+```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/la-maps.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-la-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Los Angeles Area Geomarkets, 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq1_la_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-la-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Los Angeles Area Geomarkets, 1980-2020", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq1_la_ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-la-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Los Angeles Area", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_la_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-la-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_la_midSAT_race.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+
+```
+
+```{r fig-la-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+  knitr::include_graphics('./../assets/images/rq2_la_midSAT_ses.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{r fig-ls-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
 \clearpage

### 2025-04-21 text/discussion.qmd
tweaking presentation

@@
 *Intersect* is an example of behavioral-based -- as opposed to list-based -- targeting.
-However, key to both approaches is the extent to which the product is built around a large, proprietary database of prospects.
-This can be observed in the justification for the purchase of Intersect from a University of Utah procurement document [@utah_procurement]:
-
-> The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to.
-There is unique group of prospective students who are only in the PowerSchool Naviance platform.
-We have spoken with several vendors where we can reach out to prospective students.
-Each vendor has their own proprietary website or database where select students can be reached.
 
@@
 ACT responded to uncertainty in their core testing business by attempting to become an edtech company, failing miserably, and being acquired by Nexus Capital.
+[ADD A SENTENCE ABOUT COLLEGE BOARD]
 
@@
 
-EAB is the exemplar private equity darling. START HERE ON MONDAY. THE SLIDE IS
+EAB exemplifies the the trend towards private equity ownership and industry concentration.
+EAB is the preeminent enrollment management (EM) consulting firm, serving more than 2,100 colleges and offering software-as-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
+The origins of EAB are traced to Bill Royall, who founded Royall & Company in 1983 to provide direct marketing and fundraising for Republican political campaigns [@jump_2020].
+Royall did not sign its first colege client for several years, but EM became the primary focus by 1995.
+Royall was acquired by the health tech firm Advisory Board Company for \$850 million in 2015 and then by Vista Equity Partners for \$1.5 billion in 2017, renaming it EAB [@drew_hansen_2017].
 
-The EAB Story: Acquire and Integrate
-Acquistiions in lead generation (Jaquette, Salazar, and Martin 2022; Marcus 2024)
-Consulting firms sell software-as-service “solutions”
-Build subscription recruiting platforms around database of prospects
-In their analysis of quantifying school quality in England, @RN5005 [p.
-517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model argues that student demand for higher education is mostly a function of social origin.
+As an owner-operated EM consultancy, Royall was the market leader in purchasing student lists on behalf of clients to help realize enrollment goals.
+Under Vista, EAB tranformed the market for student list data by completing a series of acquisitions that targeted the "leads" and "inquiries" stages of the enrollment funnel.
+In 2019, EAB acquired YouVisit, a leading provider of virtual tours, which can be conceptualized as an engine for "inquiries" in that taking a virtual tour shows interest in a college and leaves a trail of contact information for subsequent recruiting interventions.
+In 2020, EAB acquired Cappex, a leading provider of free college search engine, which is an engine for leads.
+In 2021, EAB and PowerSchool -- the leading provider of K12 student information systems and a subsidiary of Vista Eqity Partners -- jointly acquired the EM consultancy/software vendor Hobsons.
+EAB acquired Hobson's student success software, Starfish, while PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect.
+At the same time, EAB paid PowerSchool to become the exlusive reseller of Intersect.
+Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.
 
-Institutional theory defines the organizational field as "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services and products" [@RN527, p.
-143].
+EAB combined all of these acquisitions into Enroll360:
+
+> We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students as they pursue these increasingly digital journeys to college.
+This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]
+
+Enroll360 exemplifies a new kind of student list product that leverages proprietary data to sell software.
+Historically, College Board and ACT sold prospect contact information at a price-per-prospect (e.g., \$0.50).
+Thus, the goal of lists was to generate revenue from the sale of names.
+By contrast, Enroll360 wraps several proprietary databases of names (e.g., Cappex, Intersect) into a software-as-service platform that recruits those prospects at different stages of the enrollment funnel.
+Here, the goal of lists is to create demand for software, which is more expensive than names.
+Colleges that are interested in recruiting these prospects must buy an Enroll360 subscription.
+Like College Board *Student Seardh Service* contemporary software-as-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
+For example, Investigative reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
+Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and wheather students used Naviance to "research competitor institutions" [@markup_naviance].
+A document from the University of Utah to procure an Intersect subscription explains that the Intersect subscription is necessary to recruit the proprietary database of Naviance users: [@utah_procurement]:
+
+> The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to.
+There is unique group of prospective students who are only in the PowerSchool Naviance platform.
+We have spoken with several vendors where we can reach out to prospective students.
+Each vendor has their own proprietary website or database where select students can be reached.
+
+These dynamics of vertical and horizontal acquisitions in the enrollment management market yield policy concerns:
+
+- START HERE TUESDAY: higher prices; funneling prospects only to colleges that are willing/able to pay for subscriptions.
+
+to become a new kind of student list vendor.
 

### 2025-04-22 text/back_end_KS.qmd
LA rq1

@@
 
-The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in Figure X by total population across the three census periods.
-The Los Angeles region is one of the only multi-nuclei cases metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982.
-As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets - City of Dallas and City of Fort Worth- are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
-These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
-Worth Metroplex).
+The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
+The Los Angeles metropolitan area is one of the only multi-nuclei models in the country that is formed when urban growth occurs around multiple centers that serve different functions rather than expanding from a single central business district core (CITE- Harris and Ullman 1945).
+As depicted in @fig-la-maps, Geomarkets closely follow this overall metropolitan model.
+Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring of made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
+The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland.
+
+The population growth in the Los Angeles metropolitan area increased from approximately 8 Million in 1980 to more than 12 Million in 2020.
+@fig-dallas-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model.
+The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernadino, and Ontario.
+The South and South Central Los Angeles Geomarket grew from approximately 1.2 to 1.6 Million from 1980 to 2020.
+In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets grew from approximately 700,000 to 1.2 Million and from 1.1 Million to nearly 3 Million during this time, respectively.
 

### 2025-04-22 text/discussion.qmd
discussion. ugh.

@@
 ---
-title: "Structuring College Access: The Market Segment Model and College Board Geomarkets"
-author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
+#title: "Structuring College Access: The Market Segment Model and College Board Geomarkets"
+#author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
 bibliography: ./../assets/bib/cb_geomarket_bib.bib
@@
 We find that Geomarkets are highly correlated with race and class.
-For examle, KARINA GIVE A COUPLE OF FINDINGS.
+[KARINA SUMMARIZE FINDINGS] We show that in 1980, when Geomarkets were being created, that Black people tended to be highly concentrated in the poorest Geomarket in the metropolitan area.
+Examples include, CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), TX19 -- City of Dallas, CA21 -- South & South Central Los Angeles, and PA5 -- Philadelphia County.
+In a handful of metropolitan areas (e.g., Cleveland), Black people remain conentrated in the poorest Geomarkets.
+In Chicago-Land,  [KARINA -- ONE OR TWO SENTENCES]
 
@@
 
-We show that in 1980, when Geomarkets were being created, that Black people tended to be highly concentrated in the poorest Geomarket in the metropolitan area.
-Examples include, CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), TX19 -- City of Dallas, CA21 -- South & South Central Los Angeles, and PA5 -- Philadelphia County.
-In a handful of metropolitan areas (e.g., Cleveland), Black people remain conentrated in the poorest Geomarkets.
-In Chicago-Land,
-
-Some metro areas follow a pattern whereby Black residents [KARINA -- ONE OR TWO SENTENCES]
-
-Research question 2 analyzes the socioeconomic and racial composition of included versus excluded prospects under the hypothetical scenario that a student list purchase
-
-Research question 2 analyzes the socioeconomic and racial composition of student list purchases that include all Geomarkets in a metro area.
-We examine which Geomarkets these prospects live in as a means of assessing the consequences of student list purchases that exclude particular Geomarkets.
-This question is motivated by the @RN4982 recommendation that selective colleges that enroll students from the regional and national market segment should target affluent Geomarkets, while community colleges and non-selective 4-year colleges should focus on middle- and working-class Geomarkets [OTHER RATIONALE FOR THIS RQ FROM BOBBY?].
-
-KARINA -- ADD ONE OR TWO PARAGRAPHS SUMMARIZING RESULTS.
+Research question 2 analyzes the racial and socioeconomic composition of student list purchases.
+We examine purchases that include all Geomarkets around a metropolitan area in order to assess the potential consequences if including/excluding particular Geomarkets from the purchase.
+This question is motivated by the @RN4982 recommendation that selective colleges that enroll students from the regional and national market segment should target affluent Geomarkets, while community colleges and non-selective 4-year colleges should focus on middle- and working-class Geomarkets.
+KARINA -- ADD TEXT SUMMARIZING RESULTS.
 
@@
 The disproportionate enrollment share of high income families at selective private colleges -- and even at most public research universities -- is staggering.
-@RN1297[p.
-241-242] observes that educational credentials are property that "support their holders' claims for intermarriages with notable families\ldots, claims to monopolize socially and economically advantageous positions" and the claims made by educational credentials "are a setback for talent (charisma) in favor of property" that comes from "the desire for restricting the supply for these positions and their monopolization by the owners of educational certificates."
+These patterns are consistent with the observation by @RN1297[p.
+241-242] that educational credentials are property that "support their holders'\ldots claims to monopolize socially and economically advantageous positions."
 
@@
 ACT responded to uncertainty in their core testing business by attempting to become an edtech company, failing miserably, and being acquired by Nexus Capital.
-[ADD A SENTENCE ABOUT COLLEGE BOARD]
+[ADD A COUPLE SENTENCES ABOUT COLLEGE BOARD; MAYBE GEODEMOGRAPHIC SEGMENT]
 
@@
 Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.
-
 EAB combined all of these acquisitions into Enroll360:
@@
 
-These dynamics of vertical and horizontal acquisitions in the enrollment management market yield policy concerns:
+We argue that an under-studied and under-regulated third-party EM industry has negative consequences for colleges and for students.
+Accross multiple student list products, we see a growing number of filters that recommend which prospects to target based on the beheavior of previous cohorts of students.
+The ACT/Encoura "Enrollment Predictor" filter enables colleges to filter prospects based on the predicted probability of enrollment, which is created based on analyzing the enrollment decisions of previous cohorts of students [schmidt_2019].
+Based on the Claritas/PRIZM market segmentation system that categorized zip codes into groups useful for merchants [@RN4957], The College Board Geodemographic Segment filters categorizes each Census Tract into one of about two dozen clusters based on the college-going behavior of previous cohorts [@RN4565].
+A @vista_eab_ml video promoting EAB's "Pipeline Analytics" product to investors states, "what we are doing is using artificial intelligence and machine learning to help schools build list of students may be more more likely to enroll and persist in their university." The product "identifies students that would be a good match for a school based on their similarity to students who have succeeded there before and then provides them [the students] information about those institutions so that they choose to apply and ultimately enroll." Each of these products are based on the underlying logic of correlations and homophily discussed by @RN4975 and applied by @RN4982.
+These products identify correlations in a historic snapshot of students and apply those relationships to make recommendations about which prospects to target in the present.
+As such, all of these technologies amplify the effect of previous inequality in student demand, which is itself a consequence of structural inequality in educational opportunity.
 
-- START HERE TUESDAY: higher prices; funneling prospects only to colleges that are willing/able to pay for subscriptions.
+Market dynamics in the EM industry may also affect competition, prices, and choice.
+To the extent that colleges need to buy expensive software to recruit prospects contained in an associated database, those costs will ultimately be passed on to students in the form of higher prices are reduced expenditure on education.
+Unfettered horizontal and vertical acquisitions in the EM industry causes the EM industry to become more concentrated.
+Smaller consultancies that don't have vertical inputs -- e.g., control over proprietary pools of prospects -- cannot compete.
+Fewer EM consultancies means less competition and higher consulting fees, which will be passed on to consumers.
+Finally, the blurring of the lines between consultancy and student list vendor suggests that large consultancies will funnel prospects only to those colleges that are willing and able to pay for subscriptions, which raises policy concerns about student choice.
 
-to become a new kind of student list vendor.
+## Developing a Knowledge Infrastructure
+
+Regulation of the higher education industry should be informed by research that investigates the "totality of relevant actors" [@RN527, p.
+148] in the organizational field.
+The paucity of research about third-party vendos that structure college access is substantially due to difficulty obtaining data for empirical analyses.
+
+Knowledge infrastructures are "systems of observation and measurement" [@RN5025, p.
+743] that provide the foundation for scalable research.
+They collect, process, and distribute data in ways that "enable certain kinds of knowledge production while simultaneously channeling researchers away from questions not readily answerable within\ldots that infrastructure" [@RN5025, p.
+742].
+The knowledge infrastructure of education research largely consists of data about students, schools, and colleges.
+It has been shaped by the ascendance of economics [@RN5036], which incorporates district, state, and federal administrative data to evaluate the effects of schools and policies on student outcomes [@RN5043].
+The research community mobilized to investigate for-profit colleges because for-profit colleges and students were included in ongoing data collection [@RN5022].
+By contrast, researchers ignored the growth of private, for-profit, third-party vendors (e.g., PowerSchool, Parchment) because these organizations are excluded from data collections known to education researchers.
+
+We argue that sociologists and adjacent scholars should expand the domain of the education research knowledge infrastructure to include third-party vendors in enrollment management, and edtech more broadly.
+This is a difficult task.
+Student unit record data owned by district, state, and federal agencies is simultaneously granular -- can evaluate a particular policy -- and macro -- can speak to broad trends in the education sector.
+By contrast, obtaining granular data about third-party products in education is difficult because these data are owned by private interests that do not want researchers to interrogate their products.
+
+We see two broad avenues for research: case studies involving primary data collection -- including this study -- which tend to have a granular focus on particular products or vendors; and secondary data analysis, which may have a more macro focus.
+Case study research traditionally relies on interviews, ethnographies, and document analysis.
+Another approach to data collection is to issue records requests to public entities that contract with third-party entities.
+For example, @RN5039 analyze public university contracts with online program managers (OPMs).
+We issued public records requests to obtain tabular data about student list purchases.
+However, this process was labor intensive, requiring the pro bono efforts of several law firms.
+Public records requests seem better suited to obtaining contracts.
+Text analysis of publicly available websites and social media provides another avenue for data collection.
+Here, data collection is efficient but data processing is often labor intensive.
+Data collected for case study research typically does not meet the knowledge infrastructure standard of repeated, ongoing data collection that can be the foundation for scalable research.
+Nevertheless, case study research is important for granular analyses of particular products -- like Intersect, Enroll360,  that yield important insights about underlying mechanisms about how vendors structure college access.
+
+A growing set of "market intelligence" data providers serve the information needs of the investment community by providing data on private firms and deals involving private firms.
+These databases may provide the foundation for scalable research on the role of private firms and private equity investors in education.
+These databases tend to include three "types" of data: (1) *firm-level data* (e.g., location, employees, financials); (2) *acquisitions*, encompassing who acquired whom and resulting ownership structure); and (3) *investments*, which include amounts/valuations/stake percentages of private equity investments.
+Some providers provide all three types of data (e.g., [*S&P Capital IQ*](https://www.spglobal.com/market-intelligence/en), [*LSEG Workspace*](https://www.lseg.com/en)).
+Other providers specialize.
+For example, [*PitchBook*](https://pitchbook.com/) is a leading provider of data about investments in private companies by private equity firms, while [*Orbis*](https://login.bvdinfo.com/R1/Orbis) is the leading provider of firm-level data for non-U.S.
+private firms [@RN5038].
+Interestingly, the database subscriptions operate similar to contemporary student list products; an interactive, user-facing platform is wrapped around several sources of proprietary data.
+University libraries provides full or partial access to some subscriptions, while others must be purchased.
+These databases are often utilized for articles in management journals that focus on private equity markets [CITE], or patterns of acquisitions [CITE].
+Considering the growing share of economic activity that  [HOW THESE DATABASES HAVE BEEN USED; HOW THEY MAY BE USED]
 

### 2025-04-23 text/back_end_KS.qmd
LA rq1

@@
 A similar widening disparity is evident evident in educational attainment.
-Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23.1% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
+Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
 
@@
 In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets grew from approximately 700,000 to 1.2 Million and from 1.1 Million to nearly 3 Million during this time, respectively.
+
+\@fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020.
+Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
+In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
+By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%).
+Remanining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (44% White, 4% Black, 39% Hispanic; 11% Asian Pacific Islander), San Fernando Valley -East (43% White, 4% Black, 41% Hispanic; 8% Asian Pacific Islander), Glendale and Pasadena (35% White, 3% Black, 37% Hispanic; 21% Asian Pacific Islander), South Bay (29% White, 12% Black, 33% Hispanic; 21% Asian Pacific Islander), Long Beach (26% White, 11% Black, 46% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 3% Black, 52% Hispanic; 24% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 54% Hispanic; 6% Asian Pacific Islander).
+However, the South and South Central Los Angeles and Hollywood and Wilshire Geomarkets experienced substantial declines in the share of Black residents from 1980 to 2020, dropping from 44% to 18% and 18% to 9%, respectively.
+
+@fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020.
+While median income increased across all of the metropolitan area over the three census periods, some shifts occurred in the relative socioeconomic levels across Geomarkets.
+Many Geomarkets kept their relative levels of affluence from 1980 to 2020.
+For instance, the Covina and West Covina, South Bay, and San Fernando Valley remained among the most affluent, whereas the San Fernando Valley East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario Geomarkets remained at middle ranges of affluence and the Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets remained at the lowest levels of affluence across the metropolitan area during this period.
+The West Los Angeles and West Beach Geomarket demonstrated the most substantial increase in median household income, rising from \$112,000 in 1980 to \$132,000 in 2000, and reaching \$137,000 by 2020.
+This sustained growth positioned it as the most affluent Geomarket in the Los Angeles metropolitan area.
+
+@fig-la-rq1-ses also illustrates shifting levels of poverty across Geomarkets in the Los Angeles metropolitan area over time.
+In 1980, poverty rates were relatively low (less than 12%) for all Geomarkets except for East Los Angeles (17%), South and South Central Los Angeles (24%), and Hollywood and Wilshire (18%).
+However, poverty rates declined in South and South Central Los Angeles (22%), remained stable in East Los Angeles and Hollywood and Wilshire, and increased in all other Geomarkets.
+Such shifts and West of Dallas/Ft.
+Worth Metroplex, while increasing for all other Geomarkets.
+These divergent trends resulted in more staggered poverty rates across the metropolitan area by 2020.
+This includes four Geomarkets with poverty rates less than 10% (San Fernando Valley- West, West Los Angeles and West Beach, South Bay, and Covina and West Covina), four Geomarkets with poverty rates ranging from 11%-15% (San Fernando Valley- East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario), and three Geomarkets with poverty rates ranging from 16%-22% (Hollywood and Wilshire, East Los Angeles, South and South Central Los Angeles).
+
+Lastly, @fig-la-rq1-ses shows a widening disparity in educational attainment across the Los Angeles metropolitan area.
+Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the West Los Angeles and West Beach Geomarket experienced a substantially greater increase - from 35% in 1980 to 66% in 2020- widening the gap with other Geomarkets in terms of educational attainment.
 

### 2025-04-24 text/back_end_KS.qmd
Dallas rq2

@@
 As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
-These central Geomarkets are surrounded by the more conventional outre ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
+These central Geomarkets are surrounded by the more conventional outer ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
 Worth Metroplex).
@@
 @fig-dallas-maps showcases all six Geomarkets experienced population growth over the three census periods.
-However, the most pronounced increases occured in the West of Dallas/Ft.
-Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
+However, the most pronounced increases occured in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
 These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers - City of Forth Worth and City of Dallas - into the surrounding suburban and rural areas of the metropolitan areas.
@@
 
-Figure X presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
+@fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
 In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Ft.
 Worth Metroplex.
-By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
+By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
 In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
@@
 The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents.
-However, both Geomarkets remained predominantly White (58% and 63%, respectively) relative to all other Geomarkets.
-In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas CO (33% White, 23% Black, 32% Hispanic; 9% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian Pacific Islander), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian Pacific Islander).
+However, both Geomarkets remained predominantly White (58% and 63%, respectively).
+In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian Pacific Islander), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian Pacific Islander).
 Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
 
-Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in Figure X for the Dallas-Fort Worth metropolitan area.
+Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
 Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
@@
 However, the substantial increase in median income for the Collin & Rockwall Counties Geomarket during this time period plateaued by 2020, reshuffling relatively affluence levels among Geomarkets.
-By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Ft.
-Worth Metroplex (\$116,000), and Dallas County (\$90,000).
-In contrast, median household income for City of Dallas and City of Fort Worth remain relatively stable and continued to rank at the lower levels of affluence within the metropolitan area.
+By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000).
+In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable and continued to rank at the lower levels of affluence within the metropolitan area.
 \[KS: NEED TO CHECK THIS; I THINK CITY OF DALLAS AND CITY OF FORT WORTH ARE MIXED UP BETWEEN GRAPH VERSUS TABLE\]
 
-Figure X also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
-In 1980, poverty rates were comparably low in both for Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Ft.
-Worth Metroplex Geomarkets, and higherst for City of Dallas and City of Fort Worth Geomarkets.
+@fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
+In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets.
 Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Ft.
@@
 Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
+
+### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]
+
+We also analyze student list orders collected through public requests to assess how the racial/ethnic and socioeconomic composition of included versus excluded prospects varies when filtering on particular Geomarkets within the Dallas-Fort Worth metropolitan area.
+Specifically, we analyze three list orders placed by a research university targeting prospects across 87 different Geomarkets across the country, including all six Geomarkets in the Dallas-Fort Worth Metropolitan area.
+Each order also filtered for prospects in the 2019-2020 graduating class and based on PSAT scores across three ranges: 1070-1180, 1190-1260, and 1270-1520.
+Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260) from the Dallas-Fort Worth metropolitan area to simulate Geomarket contributions and assess patterns of inclusion versus exclusion.
 

### 2025-04-25 text/back_end_KS.qmd
Dallas rq2 results

@@
 We also analyze student list orders collected through public requests to assess how the racial/ethnic and socioeconomic composition of included versus excluded prospects varies when filtering on particular Geomarkets within the Dallas-Fort Worth metropolitan area.
-Specifically, we analyze three list orders placed by a research university targeting prospects across 87 different Geomarkets across the country, including all six Geomarkets in the Dallas-Fort Worth Metropolitan area.
+Specifically, we analyze three list orders placed by a research university targeting prospects across 87 different Geomarkets nationwide, including all six Geomarkets in the Dallas-Fort Worth Metropolitan area.
 Each order also filtered for prospects in the 2019-2020 graduating class and based on PSAT scores across three ranges: 1070-1180, 1190-1260, and 1270-1520.
 Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260) from the Dallas-Fort Worth metropolitan area to simulate Geomarket contributions and assess patterns of inclusion versus exclusion.
+
+@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area.
+At the lowest PSAT score thresholds, notable differences emerge in the racial/ethnic composition of prospects within Geomarkets when compared to the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
+For example, the largest number of prospects in the 1070-1180 PSAT range were from the Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian), both of which yield predominantly White prospect pools.
+On the other hand, the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarket pools have the greatest proportions of non-White prospects (although yielding fewer overall students).
+Interestingly, while the City of Dallas is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-dallas-rq2-race includes larger proportion of Hispanic (41%) prospects but lower proportions of Asian (4%) and Black (9%) students comparatively.
+
+These racial/ethnic composition patterns shift as PSAT scores increase.
+Shares of White prospects within Geomarkets remain relatively proportional to the overall 2020 metropolitan distribution in @fig-dallas-rq1-race.
+Howeover, prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+
+@fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
+@fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).
+Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
+The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+
+Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-dallas-rq2-midSAT-race.
+All Geomarkets besides Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex contributed equal or larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Dallas-Fort Woth metropolitan area.
+However, the City of Dallas Geomarket contributed the largest disproportionate share of Hispanic prospects (19% versus 11% for all prospects), whereas the Dallas Counties Geomarket contributed the largest disproportionate share of Black prospects (21% versus 13% for all prospects).
+
+Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses.
+Only 303 prospects of the 3,928 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
+Again, all Geomarkets besides Collin and Rockwall Counties (14% versus 36% for all prospects) and West of Dallas/Fort Worth Metroplex (15% versus 24% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool.
+Disproportioantely larger contributions range from six percentage points by the Dallas Counties Geomarket (19% versus 13% for all prospects) to 11 percentage points by the City of Dallas Geomarket (22% versus 11% for all prospects).
+Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree.
+Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportinate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
+
+Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Dallas-Fort Worth metropolitan area.
+Among all White prospects (n=2,094) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
+@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool.
+The only exception is the Irving, Arlington, & Grand Prairie Geomarket contributing a relatively larger share of White first-generation college students whose parents attended some college but did not complete their degree (21% versus 11% for all prospects).
+
+Similar patterns of one or two Geomarkets disproportionately contributing larger shares of first-generation college students was evident across other racial/ethnic categories in @fig-dallas-rq2-midSAT-combo.
+For instance, City of Dallas (18%) and City of Fort Worth (24%) contributed larger proportions of first-generation college students whose parents did not attend college relative to the overall pool of all prospects (8%).
+On the other hand, the Irving, Arlington, & Grand Prairie (23%) and Dallas Counties (26%) Geomarkets contributed disproportionately larger proportions of Black students whose parents attended some college but did not complete their degree relative to the overall pool of all prospects (17%).
+For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
+City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
 

### 2025-04-25 text/back_end_KS.qmd
rq2 LA results

@@
 
+### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]
+
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university filtered for prospects across the entire state of California, in the 2019-2020 high school graduating class, and by PSAT scores.
+Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.
+
+The racial/ethnic composition of purchased student profiles within Geomarkets in the Los Angeles metropolitan area is presented in @fig-la-rq2-race.
+The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
+However, the racial/ethnic composition of prospects become less proportional to overall metropolitan patterns by Geomarket as PSAT scores increase.
+For instance, prospect pools at middle range PSAT scores (1190-1260) within Geomarkets become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+This pattern becomes most pronounced at the highest PSAT score prospect pools.
+This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
+@fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of residents within Geomarkets.
+Similar underrepresentation patterns are evident for Hispanic porspects are evident across all Geomarkets execept for a nearly equal representation in South and South Central Los Angeles (73% prospects versus 71% Hispanic residents).
+
 # Discussion

### 2025-04-25 text/back_end_KS.qmd
rq2 LA results

--- 2025-04-25 text/back_end_KS.qmd
@@
 
+### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]
+
+We assess how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas by analyzing six orders placed by a research university.
+These six orders filtered for prospects across the entire state of California, in the 2019-2020 high school graduating class, and by PSAT scores.
+Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.
+
+The racial/ethnic composition of prospects in the Los Angeles metropolitan area whose student profiles were purchased across these orders is presented in @fig-la-rq2-race by Geomarket.
+The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect a very similar racial/ethnic composition within Geomarkets than the overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
+However, the racial/ethnic composition of prospects become less proportional to the overall metropolitan by Geomarket as PSAT scores increase.
+For instance, prospect pools at middle range PSAT scores (1190-1260) within Geomarkets become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+This pattern becomes most pronounced at the highest PSAT score prospect pools.
+This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
+@fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of all residents within Geomarkets.
+Similar underrepresentation patterns are evident for Hispanic prospects across all Geomarkets execept for a nearly equal representation in South and South Central Los Angeles (73% prospects versus 71% Hispanic residents).
+
+@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area.
+For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.
+Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
+The Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
+
+Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race.
+South and South Central Los Angeles and Riverside, San Bernadino, and Ontario Geomarkets contributed larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Los Angeles metropolitan area.
+On the other hand, the San Fernando Valley- West and the Glendale and Pasadena Geomarket contributed smaller proportions of Black and Hispanic Prospects.
+However, the West Los Angeles and West Beach (10% versus 7% for all prospects), Hollywood Wilshire (7% versus 5% for all prospects), and the Long Beach (8% versus 5% for all prospects) Geomarkets contributed a larger share of Black prospects, whereas Covina and West Covina contributed a larger share of Hispanic students.
+All other Geomarkets contributed nearly equal shares of Black and Hispanic prospects.
+
+Contributions of each Geomarket to the pool of prospects from the Los Angeles metropolitan area based on first-generation college student status are presented in @fig-la-rq2-midSAT-ses.
+Only 889 of the more than 8,000 total prospects (11%) in this middle PSAT score range (1190-1260) pool were first-generation college students whose parents did not attend college.
+Only three of the 11 Geomarkets contributed substantially larger shares of these first-generation prospects: East Los Angeles (35%), South and South Central Los Angeles (47%), and Riverside, San Bernardino, and Ontario (15%).
+However, disproportionate contributions become more spread across Geomarkets for the nearly 1,436 of the more than 8,000 total prospects (18%) who were first-generation college students whose parents attended but did not complete their degree.
+Overall, six of the 11 Geomarkets contributed larger shares of these prospects: Hollywood & Wilshire (25%), East Los Angeles (26%), South and South Central Los Angeles (28%), Long Beach (25%), Covina and West Covina (22%), and Riverside, San Bernardino, and Ontario (22%).
+This leads to the remaining Geomarkets - San Fernando Valley- West (80%), San Fernando Valley- East (78%), Glendale and Pasadena (77%), West Los Angeles & West Beach (87%), South Bay (81%) - contributing disproportionately larger shares of not first-generation college students relative to the overall prospect pool (71%).
+
+Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
+Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
+@fig-dla-rq2-midSAT-combo shows all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students (across both no college and some college parents) relative to the overall pool.
+The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%).
+Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).
+
+Geomarkets contributions to first-generation college students was mixed across other racial/ethnic categories in @fig-la-rq2-midSAT-combo.
+For Asian prospects (11% first-generation no college, 21% first-generation some college, and 68% not first-generation), the Hollywood & Wilshire (12%) and East Los Angeles (36%) Geomarkets contributed larger proportions of first-generation college students whose parents did not attend college.
+The San Fernando Valley- East (28%), Hollywood & Wilshire (35%), East Los Angeles (25%), South and South Central Los Angeles (44%), Long Beach (27%), and Covina and West Covina (25%) Geomarkets also contributed larger proportions of first-generation college students whose parents attended college but did not complete their their degree.
+
+For Black prospects (5% first-generation no college, 20% first-generation some college, and 74% not first-generation), Geomarkets that disproportionately larger proportions of first-generation college students whose parents did not attend college include South Bay (11%), South and South Central Los Angeles (21%), Long Beach (7%), and Covina and West Covina (9%).
+The West Los Angeles and West Beach (26%), Hollywood & Wilshire (36%), East Los Angeles (33%), Long Beach (40%), and Covina and West Covina (27%) Geomarkets also contributed larger proportions of Black first-generation college students whose parents attended college but did not complete their their degree.
+
+Finally, Hispanic prospects in this middle range PSAT order were on average 25% first-generation college students whose parents did not attend college, 25% first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
+The Hollywood & Wilshire (30%), East Los Angeles (39%), South and South Central Los Angeles (54%), and Riverside, San Bernardino, and Ontario (31%) contributed disproportionate shares of Hispanic first-generation college students whose parents did not attend college.
+The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
+
 # Discussion
@@
 
-```{r fig-ls-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
 knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')

### 2025-04-25 text/front_end_v4.qmd
created front end v4

--- 2025-03-26 text/front_end_v3.qmd
@@
 The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."^[On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
-Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and less White.
+Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and has a higher share of Black residents.
+Similarly, in the DC/Maryland/Virginia metro area, the order selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets.
 
@@
 -->
-Sociology offers several perspectives to describe the process of sorting students into colleges.
-The status attainment model argues that student postsecondary destination is a function of family socioeconomic background, particularly parental education [e.g., @RN915; @RN1273; @RN1268; @RN5033; @RN1300].
+Sociology offers several perspectives to describe the process of sorting students into colleges as a process that tends toward social reproduction [@RN4873; @RN2377].
+On the student demand side, the status attainment model argues that student postsecondary destination is a function of family socioeconomic background, particularly parental education [e.g., @RN915; @RN1273; @RN1268; @RN5033; @RN1300].
 The cultural capital model argues that bourgeois households bestow the pedigree, social networks, and information necessary for their progeny to claim spots at selective colleges [@RN1228; @RN1241; @RN4991] [ADD CITES].
 @RN4873 describes schools as "sorting machines" that create, incorporate, and assign students to categories that determine educational opportunities, often reinforcing reinforcing racial and class inequality.
-Finally, scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
-We argue that sociology has failed to consider how college access is structured by third-party products that identify and sort students on behalf of colleges.
-While third-party recruiting products have become increasingly powerful, granular, and ubiquitous in the era of platform capitalism, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.
+On the supply-side, the credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required to claim advantageous social and economic positions [@RN1297], and this race for credentials benefits affluent households [@RN861] [cite others].
+Scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
+
+We argue that these perspectives ignore an important mechansim of social reproduction: third-party vendors create products that sort students on behalf of college.
+These products often amplify historic inequality in educational opportunity by using past snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
+In the contemporary era of private equity funded platform capitalism, third-party recruiting products have become increasingly powerful, granular, and ubiquitous.
+However, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.
 
 In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*.
-College Board published the book, provided data, and financed the project, as part of their "efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p.
+College Board underwrote the project -- providing data and funding -- and published the book as part of their nascent efforts "to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p.
 x].
-@RN4982 developed Geomarkets and the Market Segment Model as an effort to "capture and quantify" (p.
+The broader project was conceptualized as an effort to to "capture and quantify" (p.
 11) the knowledge of admissions officers, based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel" (p.
 11).
-Created by analyzing the SAT score-sending behavior of 1980 high school seniors, the thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
+@RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of 1980 high school seniors.
+The thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
 Therefore, colleges should recruit from territories that contain large populations of your target social class.
@@
 Figure X, panel A shows Geomarkets in the Chicago-land area.
-The Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
+As a planning tool for colleges, the Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
 
-Geomarkets have been incorporated into the supply-side structure of college access in several ways.
-Albeit anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
+Geomarkets have been incorporated into the supply-side structure of college access in three ways.
+Anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
 Figure X panel B shows that Chicago-land Geomarkets define recruiting territories for University of Chicago admissions recruiters (Appendix Figures BLAH BLAH).
-Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS).
-Founded in 1984 and still active today, EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
+Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), which was in 1984 and remains active today.
+EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
 According to @RN5021, in 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis product, which was based on EPS.
@@
 
-Scholarship at the nexus of college access and recruiting has focused on the behavior of colleges [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
+Scholarship on enrollment management has focused on college behavior [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
 Scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access.
@@
 Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
-We address this question by analyzing actual student list purchases that utilize commonly used search filters (e.g., PSAT score, GPA) but do not filter on Geomarkets.
-We simulate which prospects in particular metropolitan areas would have been included/excluded had the student list purchase filtered on particular Geomarkets.
-We obtained these data by issuing public records requests to public universities.
+This research question is motivated by @RN4982, which recommends that selective colleges target affluent Geomarkets.
+Additionally, prior research on enrollment management behavior finds that selective colleges focus recruiting visits on high schools in affluent, predominantly white communities [@RN3519; @RN4758].
+We analyze this research question using data from actual student list purchases, which were collected using public records requests.
+We analyze student list purchases that include all Geomarkets in a metropolitan area in only to assess which prospects would be included/excluded by had the student list purchase filtered on particular Geomarkets.
 

### 2025-04-28 text/back_end_KS.qmd
Chi Results Smoothing

--- 2025-04-25 text/back_end_KS.qmd
@@
 @fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
-The maps in the figure show how Chicago Geomarkets reflect a classic urban geographical structure - an inner-city core surrounded by a ring of suburbs - as described by @RN4982 in the development of The Market Segment Model.
-Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area -Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
+The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 in the development of The Market Segment Model.
+Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
 In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
@@
 @fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarkets across the three census periods.
-In 1980, all Geomarkets —except for the City of Chicago - were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
+In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
 In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
@@
 For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$158,000 in the North Shore Geomarket to a low of \$66,000 in the City of Chicago Geomarket.
-However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence—\$120,000 in the Northwest Suburbs, \$119,000 in Evanston and Skokie, \$108,000 in Chain of Lakes, \$105,000 in the Western Suburbs, and \$99,000 in the South and Southwest Suburbs.
+However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence, such as \$120,000 in the Northwest Suburbs, \$119,000 in Evanston and Skokie, \$108,000 in Chain of Lakes, \$105,000 in the Western Suburbs, and \$99,000 in the South and Southwest Suburbs.
 While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$79,000) and decrease in the South and Southwest Suburbs Geomarket (\$93,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$14,000.
@@
 
+From 1980 to 2020, these demographic shifts have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by racial/ethnic and socioeconomic characteristics.
+The original Geomarket geographical boundaries delineated significant racial and income segregation in the 1980s.
+This included Black, Hispanic, and Asian residents primarily in the City of Chicago and White residents in the suburban ring of the metropolitan area.
+Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs.
+However, declining White populations, Hispanic and Asian population growth, and Black outmigration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining.
+The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents) while the the Evanston and North Shore Geomarkets remain predominantly White.
+While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains across Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
+
 ### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]
@@
 
-For instance, @fig-chicago-rq2-lowSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting lower SAT score ranges (1020-1150) by race/ethnicity.
-The top panel presents the full distribution of the 15,981 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% reside in Chain of Lakes, 10% in Northwest Suburbs, 2% in North Shore, 4% in Evanston and Skokie, 20% in City of Chicago, 25% in Western Suburbs, and 26% in South and Southwest Suburbs.
-These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-lowSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
-
+@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area.
+At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets generally reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
+For example, more than 8,000 of the 16,485 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
+Similar to the larger metropolitan area, the Northshore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students.
+Interestingly, while the City of Chicago is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-chicago-rq2-race includes nearly double the proportion of Hispanic (50%) prospects relative to the larger Geomarket resident population (26% Hispanic) in @fig-chicago-rq1-race with a lower proportion of White prospects (14%).
+
+However, the racial/ethnic composition of prospect pools become less similar to the overall metropolitan area as PSAT scores increase.
+Notably, the shares of White and Asian prospects within Geomarkets increase, while Black and Hispanic prospects decrease.
+For instance, Asian prospects over representation in Geomarkets relative to the overall resident population at the 1310-1600 SAT score threshold ranges from a minimum of 4 percentage points in the North Shore Geomarket (17% of prospects versus 13% of residents) to a maximum of 21 percentage points in the Chain of Lakes (28% of prospects versus 7% residents) Geomarket.
+The City of Chicago Geomarket exhibits the largest underrepresentation of Black prospects (7% of prospects versus 33% of residents) while Chain of Lakes exhibits the largest underrepresentation of Hispanic prospects (7% of prospects versus 24% of residents).
+
+We also explore contributions across Geomarkets to overall prospect pools by race/ethnicity.
+For instance, @fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle range SAT scores (1160-1300) by race/ethnicity.
+The top panel presents the full distribution of the 13,643 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% reside in Chain of Lakes, 13% in Northwest Suburbs, 5% in North Shore, 5% in Evanston and Skokie, 17% in City of Chicago, 28% in Western Suburbs, and 20% in South and Southwest Suburbs.
+These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-midSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
+
+```{=html}
+<!-- OLD RESULTS BASED ON LOW VERSUS HIGH SAT ORDERS
 Nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White.
@@
 Conversely, nearly all other Geomarkets contribute a disproportionately larger share of non-first generation college students across all racial/ethnic categories.
+-->
+```
+More than half (n=8,256) of all prospects whose contact information was purchased in these middle SAT score orders identified as White.
+@fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
+One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects.
+On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively.
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+
+In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race.
+Among the nearly 1,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
+The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
+
+Appendix figures \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting lower-range (1020-1150) and high-range (1310-1600) SAT scores.
+Both figures demonstrate how representational Geomarket contributionsacross race/ethnicity, such as City of Chicago disproportionately contributing larger shares of Black and Hispanic prospects, remain consistent even at lower and higher SAT score thresholds.
+
+We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
+Similar to race/ethnicity figures above, @fig-chicago-rq2-midSAT-ses presents the Geomarket distribution for the 13,749 total prospects (whose first-generations status is known) included in middle-range SAT score orders (1160-1300).
+Among the 1,743 prospects whose parents did not attend college ("no college"), the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 38% in comparison to 17% for all included prospects in this SAT range.
+In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students whose parents did not attend college relative to all included prospects, with such disparity ranging from a one-percentage-point difference (South and Southwest Suburbs) to a nine-percentage-point difference (Western Suburbs).
+For prospects whose parents attended some college but did not complete their degrees ("some college"), @fig-chicago-rq2-midSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
+These representational patterns are reversed among prospects that are not first-generation college students.
+The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (12%) relative to all prospects (17%) in this SAT range, whereas all other Geomarkets contributed comparable shares.
+
+Relative to @fig-chicago-rq2-midSAT-ses, appendix figures \@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status are less pronounced in low-SAT range orders (1020-1150) and more pronounced in high-range SAT orders (1310-1600), respectively.
+
+Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
+@fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 12,981 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders.
+For instance, among all White prospects (n=8,155) whose contact information was purchased in this SAT score range, approximately 6% were first-generation college students whose parents did not attend college, 15% were first-generation college students whose parents attended some college but did not complete their degree, and 79% were not first-generation college students.
+@fig-chicago-rq2-midSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (10% and 9%, respectively).
+For first-generation college students whose parents attended some college but never completed degrees, only the South and Southwest Suburbs (25%) Geomarket contributed a substantially larger share of White prospects.
+In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
+
+@fig-chicago-rq2-midSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
+The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
+However, the magnitude of this disproportionality varied- from a 1-percentage-point difference for Black prospects to a 23 percentage-point difference for Asian prospects.
+Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
+Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs Geomarkets.
+Appendix figures @fig-chicago-rq2-lowSAT-combo and \@fig-chicago-rq2-highSAT-combo show the magnitude of these Geomarket contribution disparities become more pronounced at as SAT score thresholds increase.
 

### 2025-04-28 text/back_end_KS.qmd
Dallas results smoothing

@@
 
-### Geomarket Changes Over Time \[Need better RQ1 heading\]
+### Shifting Demographics, Persistent Segregation in Geomarket Landscape
 
@@
 From 1980 to 2020, these demographic shifts have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by racial/ethnic and socioeconomic characteristics.
-The original Geomarket geographical boundaries delineated significant racial and income segregation in the 1980s.
-This included Black, Hispanic, and Asian residents primarily in the City of Chicago and White residents in the suburban ring of the metropolitan area.
+The original Geomarket boundaries delineated significant racial and income segregation in the 1980s.
+This included Black, Hispanic, and Asian residents primarily in the City of Chicago Geomarket and White residents in Geomarkets forming the suburban ring of the metropolitan area.
 Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs.
@@
 
-### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]
+### Composition of Students Lists by Geomarket Contributions
 
@@
 
+The enduring segregated structure of the Chicago metropolitan area from 1980 to 2020 analyzed above plays a significant role in how Geomarkets shape the racial/ethnic and socioeconomic composition of prospect pools.
+Reflecting the larger metropolitan area dynamics, more central Geomarkets such as the City of Chicago and South and Southwest Suburbs consistently contribute disproportionately larger shares of Black, Hispanic, and first-generation college prospects, especially at lower SAT score thresholds.
+On the other hand, Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, particularly as SAT scores increase.
+However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+
 ## Dallas-Fort Worth
 
-### Geomarket Changes Over Time \[Need better RQ1 heading\]
+### Suburban Geomarket Growth Eases Racial Disparities, Concentrated Wealth Persists
 
 The six Geomarkets within the Dallas-Fort Worth metropolitan area are illustrated in @fig-dallas-maps by total population across the three census periods.
-The Dallas-Fort Worth region represents a unique metroplex case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic model originally proposed by @RN4982.
-As depicted in @fig-dallas-maps, two distinct inner-city core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
+The Dallas-Fort Worth region represents a unique "metroplex" case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic urban core and suburban ring model showcased by Chicago.
+As depicted in @fig-dallas-maps, two distinct inner-core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
 These central Geomarkets are surrounded by the more conventional outer ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.

### 2025-04-29 text/back_end_KS.qmd
DFW smoothing results

@@
 These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers - City of Forth Worth and City of Dallas - into the surrounding suburban and rural areas of the metropolitan areas.
-This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket.
+This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket from 1980 to 2000.
 
 @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
-In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Ft.
-Worth Metroplex.
+In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
 By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
@@
 In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets.
-Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Ft.
-Worth Metroplex, while increasing for all other Geomarkets.
+Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex, while increasing for all other Geomarkets.
 These divergent trends resulted in a widening gap in poverty levels across the Dallas-Fort Worth metropolitan area.
-A similar widening disparity is evident evident in educational attainment.
+A similar widening disparity is evident in educational attainment.
 Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
+
+The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s through 2020.
+The metropolitan area's unique urban structure of a dual central cities surrounded by concentric suburbs has shaped patterns of population growth across racial/ethnic, economic, and spatial shifts in the Geomarket lanscape.
+While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
+However, all Geomakets have experienced substantial growth in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex.
+On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment across Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
+Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
 

### 2025-04-29 text/front_end_v3.qmd
working through front end, again

--- 2025-03-13 text/front_end_v3.qmd
@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."  On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
+Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and has a higher share of Black residents.
+Similarly, in the DC/Maryland/Virginia metro area, the order selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets.
 
+[^1]: On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
+
+```{=html}
 <!--Geomarkets carve states and metroplitan areas into smaller geographic units and have names like,  On the same day, the university purchased two additional orders that were identical, except they filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$.
 On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
--->
 
@@
 To the extent that Geomarket borders are correlated with race and class, the Geomarket selections observed here raise concern that Geomarkets -- and other geographic search filters -- may systematically filter working class and non-white students from educational opportunities.
+-->
+```
 
-The origins of Geomarkets, and how they were incorporated into recruiting products, turns out to be an interesting story.
+Sociology offers several perspectives to describe the process of sorting students into colleges as a process that tends toward social reproduction [@RN4873; @RN2377].
+On the student demand side, the status attainment model argues that student postsecondary destination is a function of family socioeconomic background, particularly parental education [e.g., @RN915; @RN1273; @RN1268; @RN5033; @RN1300].
+The cultural capital model argues that bourgeois households bestow the pedigree, social networks, and information necessary for their progeny to claim spots at selective colleges [@RN1228; @RN1241; @RN4991] [ADD CITES].
+@RN4873 describes schools as "sorting machines" that create, incorporate, and assign students to categories that determine educational opportunities, often reinforcing reinforcing racial and class inequality.
+On the supply-side, the credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required to claim advantageous social and economic positions [@RN1297], and this race for credentials benefits affluent households [@RN861] \[cite others\].
+Scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
+
+We argue that these perspectives ignore an important mechansim of social reproduction: third-party vendors create products that sort students on behalf of college.
+These products often amplify historic inequality in educational opportunity by using past snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
+In the contemporary era of private equity funded platform capitalism, third-party recruiting products have become increasingly powerful, granular, and ubiquitous.
+However, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.
+
 In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*.
-College Board published the book, provided data, and financed the project, as part of their "efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p.
+College Board underwrote the project -- providing data and funding -- and published the book as part of their nascent efforts "to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p.
 x].
-@RN4982 sought to "capture and quantify" (p.
-11) the knowledge of admissions officers,
-
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
-9-10).
-
-@RN4982 developed Geomarkets and the Market Segment Model based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel." Created by analyzing the SAT score-sending behavior of 1980 high school seniors, the thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
+The broader project was conceptualized as an effort to to "capture and quantify" (p.
+11) the knowledge of admissions officers, based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel" (p.
+11).
+@RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of 1980 high school seniors.
+The thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
 Therefore, colleges should recruit from territories that contain large populations of your target social class.
@@
 Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
-We address this question by analyzing actual student list purchases that utilize commonly used search filters (e.g., PSAT score, GPA) but do not filter on Geomarkets.
-We simulate which prospects in particular metropolitan areas would have been included/excluded had the student list purchase filtered on particular Geomarkets.
-We obtained these data by issuing public records requests to public universities.
+This research question is motivated by @RN4982, which recommends that selective colleges target affluent Geomarkets.
+Additionally, prior research on enrollment management behavior finds that selective colleges focus recruiting visits on high schools in affluent, predominantly white communities [@RN3519; @RN4758].
+We analyze this research question using data from actual student list purchases, which were collected using public records requests.
+We analyze student list purchases that include all Geomarkets in a metropolitan area in only to assess which prospects would be included/excluded by had the student list purchase filtered on particular Geomarkets.
 
-The following section provides background about enrollment management and introduces salient concepts from scholarship on quantification.
+The following section introduces background and scholarship about enrollment management and introduces salient concepts from scholarship on quantification, particularly correlation and homophily.
 Second, we introduce the case, describing how @RN4982 developed Geomarkets and the Market Segment Model and how these concepts were incorporated into College Board recruiting products.
@@
 By investigating high-leverage actors and products in the organizational field, sociology can develop important insights that shape future federal and state regulatory policy.
+[REVISE THIS PARAGRAPH]
 
@@
 The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
-Economists often investigate financial aid leveraging, which seeks to convert admits to enrolled students \[CITE\].
+Scholarship from the economics of education often investigate financial aid leveraging, which seeks to convert admits to enrolled students [e.g., @RN3762; @RN3133].
 <!--  Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935; @RN1814].
@@
 Ethnographies by @RN3519 and @RN4407 identify connections between private school guidance counselors and college admissions officers as a mechanism for social reproduction.
-Recruiting visits to high schools are a means of maintaining ties with guidance counselors at feeder schools and establishing relationships with prospective students [@RN3519; @RN4895].
+In turn, recruiting visits to high schools -- and subsequent phone calls -- are a means of maintaining ties with guidance counselors at feeder schools and establishing relationships with prospective students [@RN3519; @RN4895].
 Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly white public schools [@RN4758; @RN4759; @RN5023].
@@
 
-By focusing on the behaviors of individual colleges, scholarship on EM implicitly states that recruiting is something done by individual colleges.
-In addition to university personnel (e.g., admissions counselors, VP for enrollment management), the EM industry incorporates external stakeholders in the organizational field, including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
+By focusing on the behavior of colleges, scholarship on EM implicitly states that recruiting is something done by individual colleges.
+The broader EM industry includes relevant stakeholders in the organizational field [@RN527], including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
 We argue college access is structured by third-party servicers and products that interact with direct-providers (colleges).
@@
 
-Drawing from scholarship on organizational theory, enrollment management processes involve many "make or buy" [@RN4738] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
+Drawing from scholarship on organizational theory, enrollment management processes require colleges to make many "make or buy" [@RN4738; @RN980] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
+Unfortunately, few empirical studies make EM vendors or products the object of empirical analysis.
+The most widely-known class of EM vendor is EM consulting firms, which provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
 As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
+The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].^[In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].]
 
-Although the enrollment management industry is increasingly characterized by software-as-service products sold by private equity backed firms (e.g., for example, EAB's [Enroll 360](https://eab.com/solutions/enroll360/) product) [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
-The Student Search Service, created in 1972, became the ubiquitous means of identifying prospects BLAH BLAH.
-Second, EPS, launcched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
+While the contemporary EM industry is characterized by software-as-service products sold to colleges by private equity backed firms [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
+Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit.
+The Student Search Service -- created in 1972 and copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
+Second, the Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
 Both products are applications of quantification.
-Student Search Service includes or excludes prospects based on selected attributes while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
-THEREFORE WE REVIEW SCHOLARSHIP ON QUANTIFICATION.
+Student Search Service includes or excludes prospects from a student list purchase based on value of selected variables while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
+The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on equality of opportunity.
 
@@
 Law school rankings were developed as a means of informing prospective students about the relative quality of different programs.
-Rankings affect (reactivity) where students apply and enroll \[CITE\].
-However, rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
+Rankings affect (reactivity) where students apply and enroll [@RN5003].
+Rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
 Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p.
@@
 Rankings *disciplined* the behavior of law schools.
-For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
-Rankings also weakened the *authority* of admissions directors and admissions officers to make independent decisions about which applicants to admit, ideal class size, and curricular offerinags [@RN5003].
+For example, characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
+Rankings also weakened the *authority* of admissions personnel to make independent decisions class size and which applicants to admit [@RN5003].
 <!--
@@
 First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
-Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and to assign levels of risk to each case.
+Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and/or to assign levels of risk to each case.
 @RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
@@
 @RN4975 [p.
-50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"[^1]
+50] writes that
 
-[^1]: Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
+> correlation grounds big data's so-called revolutionary potential.
+As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'
+
+Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
 For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
@@
 
-A bevy of studies show that predictions based on correlations reproduce structural inequality [for a review see @RN4794].
-The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
-The observed correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
+Many studies show that predictions based on correlations reproduce structural inequality [for a review see @RN4794].
+The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point in time.
+Observed correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
 Reviewing scholarship about algorithms, @RN4794 [p.
 224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^2] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
+This phenomenon has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^3] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
 They then recommended that selective colleges recruit from localities that have a high share of affluent, college-educated households.
 
-[^2]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
+[^3]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
 Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
@@
 Homophily is the idea actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together" [@mcpherson2001birds].
-Homophily is a core concept of network science (writ large), in which actors (nodes) are connected to one another directly and indirectly via network ties (edges).
-As discussed below, @RN4982 [chapter 4] *The company We Keep: Colleges and Their Competition* conducts a network analysis based on 1980 SAT score sending behavior whereby two colleges are defined as competitors if a large number of students send SAT scores to both colleges.
-Because network science models often draws from rational choice theory, @RN4975 argues, it assumes that homophily is the result of voluntary action by individuals.
+Homophily is a core concept of network science, in which actors (nodes) are connected to one another directly and indirectly via network ties (edges).
+In *The company We Keep: Colleges and Their Competition*, @RN4982 [chapter 4] conduct a network analysis whereby two colleges are defined as competitors if a large number of students submit SAT scores to both colleges.
+@RN4975 argues that because network science models often draw from rational choice theory, they assume that homophily is the result of voluntary action by individuals.
 
 @RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-Like correlational models discussed above, because most network science models do not observe how historical structures affect relationships, they "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
+Because network science models -- like correlational models above -- typically do not observe how historical structures affect relationships, they "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
 95] that cause behavior consistent with homophily.
@@
 YY].
-We observe similar processes in college access.
+We observe similar phenomena in our case.
 Based on a snapshot of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
@@
 
-[^3]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
+[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
 
-**Contribution**.
-As @RN3519 demonstrates in the chapter *Numbers*, enrollment management is fundamentally concerned with quantification.
-In the case of college rankings, sociology made visible how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges who are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
-Beyond rankings, sociological analyses of enrollment management focus on the behavior of colleges and their agents, yielding the implicit assumption that inequalities creates by enrollment management are a function of individual college behavior.
-We argue that third-party vendors and products structure the enrollment management behavior of colleges.
-These products take a snapshot of student demand -- without considering structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
+We make targeted contributions to scholarship on quantification in the sociology of education.
+First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
+As @RN3519 demonstrates in the chapter *Numbers*, EM is fundamentally concerned with quantification.
+However, sociological analyses of enrollment management focus on the behavior of colleges and their agents, yielding the implicit belief that inequalities created by EM are a function of individual colleges reacting to rankings and other macro structures in their environment.
+To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of producers.
+Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027; ].
+Drawing from scholarship from critical data studies [e.g., @RN4775; @RN4975; @RN4849], recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
 Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
-These products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociologists.
+Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociologists.
+We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
 
-This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
+This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
 The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
-Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification.
-However, this study provides inisight about the mechanism underlying the effect.
-By analyzing and simulating *Student Search Service* purchases that filter on Geomarkets we show how Geomarkets reproduce historical race-based inequality in college access.
+Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification on social reproduction.
+However, by showing which students are included/excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in educational opportunity.
 
@@
 
-**Creating Geomarkets and the Market Segment Model.** In 1978, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
+**Creating Geomarkets and the Market Segment Model.**
+
+In 1978, set against the backdrop of impending college-age demographic decline, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
 x].
-Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
-<!--The geographic concepts inscribed in the Market Segment Model were based on conversations with college admissions officers.
---> The "initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982,p.
+To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
+The project goal became the creation of the Market Segment Model, which had the lofty ambition of predicting student demand for any college from any locality.
+However, "to gain a truly comprehensive view of the collegiate enrollment market, we needed a database that described most institutions and most students" [@RN4982, p.
+x].
+The group approached College Board.
+"Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest deomgraphic unit used in these analyses was the state" [@RN4982, p.
+x].
+
+Although the analyses of @RN4982 are based on the score-sending behavior of SAT test-takers, the authors had great respect for admissions officers and conceived of the project as an effort to "capture and quantify" (p.
+11) their knowledge:
+
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
+9-10).
+
+The "initial task" was to define geographic boundaries "in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982,p.
 4].
 Quoting an admissions officer, @RN4982 [p.
-11] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, [@RN4982] created three geographic [@RN4982, p.
-11], region, state, and community.
-The three regions were New England, Middle States, and the South.
-They "divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" (p.
+11] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, @RN4982[p.
+11] created "three types of boundaries -- region, state, and community.
+The initial regions in the project were New England, Middle States, and the South.
+Next, they "divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" (p.
 11).
@@
 @RN4982 [pp.
-11-12] briefly describes the creation of Geomarket borders:
+11-12] described the creation of Geomarket borders only briefly:
 
@@
 
-The goal of the Market Segment Model was to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households in the Geomarket.
-Zemsky reached out to College Board because "we needed a database that described most institutions and most students" [@RN4982, p.
-x].
-<!--Zemsky reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential" [@RN4982, p.
-x].
--->In 1979, College Board began providing funding and data on the score-sending behavior of SAT test-takers.
+Having defined geography, the Market Segment Model sought to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households in the Geomarket:
 
@@
 The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
-These four variables "reflect the basic social patterns of the nation.
-It would have been surprising if these were not the four social variables that best explained the patterns of college choice" [@RN4982, p.
+Because these four variables "reflect the basic social patterns of the nation, it would have been surprising if these were not the four social variables that best explained the patterns of college choice" [@RN4982, p.
 33].<!--
@@
 The thesis of Chapter 3 is that student demand for higher education is a function of social origin.
-For example, @RN4982 \[p.
-33\] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^6] @RN4982 \[p.
-42\] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
+For the authors (p.
+33) state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and that "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
+33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^7] @RN4982 \[p.
+42\] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
+However, Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
 Therefore, @RN4982 [p.
 44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:
-
-[^6]: The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
-Additionally, analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
-However, prior research shows that SAT scores are substantially a function of social origin [@RN1273].
 
@@
 **Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^7] Based on these analyses, @RN4982 \[p.
-46\] state that homophily describes the nature of competition between colleges: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^8] Based on these analyses, @RN4982[p.
+46] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
 Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
@@
 @RN4982 [p.
-72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+72] describe observed patterns as a natural process in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
 
-[^7]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+[^8]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
 The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
@@
 
-The discussion of competition by @RN4982 exemplifies the concerns about correlation and homophily described by [@RN4975].
+The discussion of competition by @RN4982 exemplifies the concerns about correlation and homophily described by @RN4975.
 A correlational analysis of 1980 SAT score-sending patterns finds competition between colleges is defined by socioeconomic homophily.
@@
 Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer-coleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
-In itself, @RN4982's Market Segment Model is merely a social science depiction of the structure of student demand, one that does not consider historic, structural inequalities that cause observed patterns.
+In itself, @RN4982's Market Segment Model is a social science depiction of student demand -- akin to the status attainment model -- that does not consider historic, structural inequalities that cause observed patterns.
 However, by inscribing the Market Segment Model and Geomarkets into the EPS software, College Board amplified structural inequalities that contributed to homophily observed in 1980 SAT score-sending beahivor.
@@
 For each Geomarket, colleges could obtain Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors.
-Based on conversations with enrollment management professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket.
+Based on background conversations with enrollment management professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket.
 Therefore, colleges used EPS software to decide which Geomarkets to recruit from and which high schools to visit within targeted Geomarkets.
@@
 
-EPS software may discipline colleges to approach recruiting in a manner consistent with the Market Segment Model.
+EPS software may "discipline" [@RN5007] colleges to approach recruiting in a manner consistent with the Market Segment Model.
 Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
-EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
+EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^10] Second, EPS users decide which high schools they will visit within selected Geomarkets.
 Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
 
-[^9]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+[^10]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
 Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
@@
 In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
-However, EPS software recommended visiting the same sets of affluent, high-achieving high schools that were receiving visits from other colleges.
-We were told that savvier, quantitatively-adept admissions offices used EPS to visit schools explicitly not recommended by EPS because there would be less competition for the good students that attended these high schools.
+However, EPS software often recommended visiting the same sets of affluent, high-achieving high schools that were receiving visits from other colleges.
+We were told that savvier, quantitatively-adept admissions offices used EPS to visit schools that EPS recommended ignoring because there would be less competition for the good students who attended these high schools.
 
@@
 In turn, as Geomarkets became more salient, admissions offices often structured their territories around Geomarkets.
-On background, admissions officers and enrollment consultants told us it was common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket.
-In those states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
-This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX\], \[ADD TO APPENDIX\].
-<!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
+On background, admissions officers and enrollment consultants told us it remains common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket.
+In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
+This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX CRYSTAL\] <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
 
@@
 Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
-Moreover, @RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards distinguishing geographic areas that differed from one another in terms of class composition and college-going behavior.
+Moreover, @RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards identifying geographic areas that differed from one another in terms of class composition.
 
@@
 
-We also expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons.
-Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race once.
+We also expect substantial racial inequality between geomarkets in large metropolitan areas.
+Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race.
 U.S.
@@
 Unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
-Second, geomarket borders may have been drawn along class divides.
-A strong correlation exists between race and wealth [@RN4884].
+Second, although geomarket borders may have been drawn along class divides, a strong correlation exists between race and wealth [@RN4884].
 Third, Geomarket borders may have been drawn in a way that follows the countours of racial segregation in residential housing.

### 2025-04-30 text/back_end_KS.qmd
readded deleted LA rq2 wtf

@@
 The enduring segregated structure of the Chicago metropolitan area from 1980 to 2020 analyzed above plays a significant role in how Geomarkets shape the racial/ethnic and socioeconomic composition of prospect pools.
-Reflecting the larger metropolitan area dynamics, more central Geomarkets such as the City of Chicago and South and Southwest Suburbs consistently contribute disproportionately larger shares of Black, Hispanic, and first-generation college prospects, especially at lower SAT score thresholds.
-On the other hand, Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, particularly as SAT scores increase.
+Reflecting the larger metropolitan area dynamics, central Geomarkets that are predominantly comprised of low-income communities of color such as the City of Chicago and South and Southwest Suburbs consistently contribute disproportionately larger shares of Black, Hispanic, and first-generation college prospects, especially at lower SAT score thresholds.
+On the other hand, predominantly White and affluent Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, as well as non-first generation college students.
 However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
@@
 
-The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s through 2020.
-The metropolitan area's unique urban structure of a dual central cities surrounded by concentric suburbs has shaped patterns of population growth across racial/ethnic, economic, and spatial shifts in the Geomarket lanscape.
+The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s.
 While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
-However, all Geomakets have experienced substantial growth in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex.
+Such growth is largely attributed to increases in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex.
 On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment across Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
@@
 
-### Composition of Students Lists by Geomarkets \[Need Better RQ2 Heading\]
+### Composition of Students Lists by Geomarkets
 
@@
 
+Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions.
+As Dallas-Fort Worth became increasingly diverse from 1980 to 2020, a more balanced racial/ethnic distribution emerged across the metropolitan area's Geomarket landscape.
+However, these patterns did not translate to racial/ethnic composition of student prospect pools.
+Suburban Geomarkets like Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex yielded a disproportionately larger share of Asian prospects, whereas central Geomarkets such as the City of Dallas, Dallas County, and Irving, Arlington, and Grand Prairie contributed a disproportionately larger share of Hispanic and Black prospects.
+These racial/ethnic disparities become significantly more pronounced, particularly in the disproportionate share of Asian prospects, as PSAT score thresholds increase (see Appendix @fig-dallas-rq2-lowSAT-race and @fig-dallas-rq2-highSAT-race).
+Similar Geomarket contribution patterns are also evident by first-generation college student status, with Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex providing larger shares of non-first generation college student prospects, and central Geomarkets (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie) providing disproportionately larger shares of first-generation college student prospects.
+Moreover, central Geomarkets are particularly important for including Black (City of Dallas and City of Fort Worth), Hispanic (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie), and Asian (City of Fort Worth and Irving, Arlington, and Grand Prairie) first-generation college students.
+While Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex Geomarkets still contribute a notable overall number of Black, Hispanic, and Asian prospects, such students are predominantly not first-generation college students.
+
 ## Los Angeles
 
-### Geomarket Changes Over Time \[Need better RQ1 heading\]
+### Decentralized Diversity and Income in a Multi-Nucleo in the Los Angeles Geomarket Landscape
 
@@
 Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the West Los Angeles and West Beach Geomarket experienced a substantially greater increase - from 35% in 1980 to 66% in 2020- widening the gap with other Geomarkets in terms of educational attainment.
+
+The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
+From 1980 to 2020, the region’s population expanded significantly leading to overall declines in the White population and increases in Hispanic and Asian populations across Geomarkets.
+West Los Angeles and West Beach is the only majority White Geomarket in 2020.
+Socioeconomic indicators reveal enduring disparities, especially in historically lower-income Geomarkets like South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire.
+Meanwhile, Geomarkets such as West Los Angeles and West Beach became increasingly affluent and highly educated, highlighting growing socioeconomic and educational divides.
+Given these demographic shifts within the context of a multi-nuclei urban model, the Los Angeles metropolitan area presents a distinctive Geomarket landscape case in which racial/ethnic enclaves across a range of various socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.
 
@@
 
+@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area.
+For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.
+Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
+The Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
+
+Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race.
+South and South Central Los Angeles and Riverside, San Bernadino, and Ontario Geomarkets contributed larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Los Angeles metropolitan area.
+On the other hand, the San Fernando Valley- West and the Glendale and Pasadena Geomarket contributed smaller proportions of Black and Hispanic Prospects.
+However, the West Los Angeles and West Beach (10% versus 7% for all prospects), Hollywood Wilshire (7% versus 5% for all prospects), and the Long Beach (8% versus 5% for all prospects) Geomarkets contributed a larger share of Black prospects, whereas Covina and West Covina contributed a larger share of Hispanic students.
+All other Geomarkets contributed nearly equal shares of Black and Hispanic prospects.
+
+Contributions of each Geomarket to the pool of prospects from the Los Angeles metropolitan area based on first-generation college student status are presented in @fig-la-rq2-midSAT-ses.
+Only 889 of the more than 8,000 total prospects (11%) in this middle PSAT score range (1190-1260) pool were first-generation college students whose parents did not attend college.
+Only three of the 11 Geomarkets contributed substantially larger shares of these first-generation prospects: East Los Angeles (35%), South and South Central Los Angeles (47%), and Riverside, San Bernardino, and Ontario (15%).
+However, disproportionate contributions become more spread across Geomarkets for the nearly 1,436 of the more than 8,000 total prospects (18%) who were first-generation college students whose parents attended but did not complete their degree.
+Overall, six of the 11 Geomarkets contributed larger shares of these prospects: Hollywood & Wilshire (25%), East Los Angeles (26%), South and South Central Los Angeles (28%), Long Beach (25%), Covina and West Covina (22%), and Riverside, San Bernardino, and Ontario (22%).
+This leads to the remaining Geomarkets - San Fernando Valley- West (80%), San Fernando Valley- East (78%), Glendale and Pasadena (77%), West Los Angeles & West Beach (87%), South Bay (81%) - contributing disproportionately larger shares of not first-generation college students relative to the overall prospect pool (71%).
+
+Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
+Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
+@fig-la-rq2-midSAT-combo shows all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students (across both no college and some college parents) relative to the overall pool.
+The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%).
+Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).
+
+Geomarkets contributions to first-generation college students was mixed across other racial/ethnic categories in @fig-la-rq2-midSAT-combo.
+For Asian prospects (11% first-generation no college, 21% first-generation some college, and 68% not first-generation), the Hollywood & Wilshire (12%) and East Los Angeles (36%) Geomarkets contributed larger proportions of first-generation college students whose parents did not attend college.
+The San Fernando Valley- East (28%), Hollywood & Wilshire (35%), East Los Angeles (25%), South and South Central Los Angeles (44%), Long Beach (27%), and Covina and West Covina (25%) Geomarkets also contributed larger proportions of first-generation college students whose parents attended college but did not complete their their degree.
+
+For Black prospects (5% first-generation no college, 20% first-generation some college, and 74% not first-generation), Geomarkets that disproportionately larger proportions of first-generation college students whose parents did not attend college include South Bay (11%), South and South Central Los Angeles (21%), Long Beach (7%), and Covina and West Covina (9%).
+The West Los Angeles and West Beach (26%), Hollywood & Wilshire (36%), East Los Angeles (33%), Long Beach (40%), and Covina and West Covina (27%) Geomarkets also contributed larger proportions of Black first-generation college students whose parents attended college but did not complete their their degree.
+
+Finally, Hispanic prospects in this middle range PSAT order were on average 25% first-generation college students whose parents did not attend college, 25% first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
+The Hollywood & Wilshire (30%), East Los Angeles (39%), South and South Central Los Angeles (54%), and Riverside, San Bernardino, and Ontario (31%) contributed disproportionate shares of Hispanic first-generation college students whose parents did not attend college.
+The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
+
 # Discussion

### 2025-04-30 text/back_end_KS.qmd
smoothing LA rq2

@@
 The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
+
+Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
+Affluent Geomarkets like San Fernando Valley–West and West Los Angeles contribute an overrepresentation of White, Asian, and not-first generation college student prospects, the magnitude of which increases substantially as score ranges increase (see Appendix @fig-la-rq2-lowSAT-race and @fig-la-rq2-highSAT-race).
+On the other hand, more diverse and historically low-income Geomarkets, like South and South Central Los Angeles and Riverside, San Bernardino, and Ontario, contribute disproportionately more Black, Hispanic, and first-generation college student prospects.
+Similar to Chicago and Dallas-Fort Worth, Los Angeles further solidifies the pattern of some of the lowest-income Geomarkets consistently playing crucial roles in fostering college access for first-generation Asian (East Los Angeles), Black (South and South Central Los Angeles), and Hispanic college students (South and South Central Los Angeles, East Los Angeles, Hollywood and Wilshire, and Riverside, San Berdardino, and Ontario).
 

### 2025-04-30 text/back_end_KS.qmd
methods revisions

@@
 ```
-# Introduction
-
-On January 9, 2019 a public research university purchased College Board *Student Search Service* order no.
-$448922$, thereby obtaining the contact information of out-of-state prospective students, who would subsequently receive brochures, postcards, email, and targeted social media.
-This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
-Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and less White.
-
-[^1]: On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
-
-```{=html}
-<!--Geomarkets carve states and metroplitan areas into smaller geographic units and have names like,  On the same day, the university purchased two additional orders that were identical, except they filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$.
-On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
-
-Geomarkets carve states and metroplitan areas into smaller geographic units.
-Upon closer inspection, the three orders selected on all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago", which is significantly poorer and less White.
-In the DC/Maryland/Virginia metro area, the orders selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets, which are poorer and have a higher share of Black residents.
-An analysis of 500 \[check\] student lists purchased by 9 public research universities found that 11% of purchases filtered on Geomarket [@RN5012].
-Recent research suggests that student lists are surprisingly important for college access outcomes and college graduation outcomes, particularly for first-generation and underrepresented minority students [@RN4739; @RN4894].
-To the extent that Geomarket borders are correlated with race and class, the Geomarket selections observed here raise concern that Geomarkets -- and other geographic search filters -- may systematically filter working class and non-white students from educational opportunities.
--->
-```
-Sociology offers several perspectives to describe the process of sorting students into colleges.
-The status attainment model argues that student postsecondary destination is a function of family socioeconomic background, particularly parental education [e.g., @RN915; @RN1273; @RN1268; @RN5033; @RN1300].
-The cultural capital model argues that bourgeois households bestow the pedigree, social networks, and information necessary for their progeny to claim spots at selective colleges [@RN1228; @RN1241; @RN4991] [ADD CITES].
-@RN4873 describes schools as "sorting machines" that create, incorporate, and assign students to categories that determine educational opportunities, often reinforcing reinforcing racial and class inequality.
-Finally, scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
-All of these perspectives share the view that transitions from school to college tend towards social reproduction.
-We argue that these perspectives have ignored an important mechanism of social reproduction: third-party vendors create products that sort students on behalf of college, often utilizing historical snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
-While third-party recruiting products have become increasingly powerful, granular, and ubiquitous in the era of platform capitalism, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.
-
-In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*.
-College Board published the book, provided data, and financed the project, as part of their "efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p.
-x].
-@RN4982 developed Geomarkets and the Market Segment Model as an effort to "capture and quantify" (p.
-11) the knowledge of admissions officers, based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel" (p.
-11).
-Created by analyzing the SAT score-sending behavior of 1980 high school seniors, the thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
-Therefore, colleges should recruit from territories that contain large populations of your target social class.
-Geomarkets are geographic borders meant to define local recruiting markets, a territory of an admissions recruiter.
-Figure X, panel A shows Geomarkets in the Chicago-land area.
-The Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
-
-Geomarkets have been incorporated into the supply-side structure of college access in several ways.
-Albeit anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
-Figure X panel B shows that Chicago-land Geomarkets define recruiting territories for University of Chicago admissions recruiters (Appendix Figures BLAH BLAH).
-Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS).
-Founded in 1984 and still active today, EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
-According to @RN5021, in 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis product, which was based on EPS.
-
-Third, Geomarkets were incorporated into College Board's student list product, Student Search Service.
-Student lists contain the contact information of prospective students and have been the primary source of lead generation in U.S.
-higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz].
-@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
-The Student Search Service database consists of College Board test-takers.
-Colleges pay a fee for each prospect (e.g., \$0.50 in 2021).
-Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc.
-Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets.
-
-Scholarship at the nexus of college access and recruiting has focused on the behavior of colleges [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
-Scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access.
-These mechanisms, which lie upstream of the behavior of individual colleges, are incorporated by individual colleges in ways that funnel certain kinds of students to certain kinds of institutions.
-This omission is surprising because sociology is concerned with sorting [@RN4873], particularly sorting into college.
-
-This manuscript analyzes College Board Geomarkets as a case study of quantification.
-In particular, we draw from the discussions of correlation and homophily from @RN4975.
-@RN4982 identified the correlates of 1980 SAT score-sending behavior and concluded that student demand for higher education is primarily a function of social class.
-The Market Segment Model argues that homophily -- actors that share characteristics form connections -- is the organizing principle of competition and college choice, stating that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students" [@RN4982, p.
-72].
-Scholarship on quantification demonstrates that making recommendations based on past correlations amplifies the effects of historic structural inequality [@RN4794].
-The snapshot of student demand in 1980 -- itself a consequence historic structural inequality -- was programmed into recruiting products that colleges utilize to identify and target prospective students.
-The result is a supply-side that amplifies structural inequalities observed on the demand-side.
-
-Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS and within the Student Search Service student list product.
-First, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
-We address this question by spatially joining the geomarket shapefile to Census data about socioeconomic and racial characteristics.
-Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
-We address this question by analyzing actual student list purchases that utilize commonly used search filters (e.g., PSAT score, GPA) but do not filter on Geomarkets.
-We simulate which prospects in particular metropolitan areas would have been included/excluded had the student list purchase filtered on particular Geomarkets.
-We obtained these data by issuing public records requests to public universities.
-
-The following section provides background about enrollment management and introduces salient concepts from scholarship on quantification.
-Second, we introduce the case, describing how @RN4982 developed Geomarkets and the Market Segment Model and how these concepts were incorporated into College Board recruiting products.
-Third, we describe data and methods.
-Fourth, we present results.
-[We find that\ldots ?PARAGRAPH ON RESULTS?] Finally, we discuss implications for scholarship and for policy.
-The sociology of education should move beyond the fixation on schools and policies as the primary sorting machines.
-@RN527 [p.
-148] remind us that the organizational field includes the totality of relevant actors." Third-party vendors have structured college access for decades and their influence is growing.
-The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product \[CITE\].
-By investigating high-leverage actors and products in the organizational field, sociology can develop important insights that shape future federal and state regulatory policy.
-
-```{=html}
-\[ADD TEXT SOMEWHERE AROUND HERE ABOUT WHY THIS TOPIC IS SIGNIFICANT AND IMPORTANT FOR CURRENT/FUTURE POLICY DEBATES: Why market segment model important.
-Software as service products dominate em industry and structure college access.
-Look at eab machine learning product But rarely do we see the recipe underneath these products.
-The book structuring college access is that recipe for EPS software\]
-<!-- 
-THIS TEXT IS FINE, BUT TRYING TO SET UP THE CONTRIBUTION TO BROADER QUANTIFICATION LITERATURE, VIS-A-VIS NO-ONE HAS STUDIED MARKET RESEARCH; DON'T THINK THIS IS NECESSARY AND YOU ARE NOT ON FIRM GROUND HERE IN CALLING THIS "MARKET RESEARCH"
-Scholarship on quantification in education has focused on consumer-facing measures such as U.S.
-News Rankings [@RN5003] and UK "league tables" about school quality [@RN5005].
-To our knowledge, prior knowledge has not examined business-facing measures produced by third-party market research.
-Here, the desirable attributes of students and places they live are quantified with the purpose of informing colleges about where they should allocate recruiting resources to efficiently enroll the most desirable students.
-We reason that market research reproduces inequality by encouraging colleges to plan their recruiting efforts based on a snapshot of social stratification in student demand that is substantially a function of historic, structural inequality in educational opportunity.
-In contrast to @RN5005, we cannot demonstrate the effect of geomarkets on inequality.
-However, we show the mechanisms of how geomarkets reproduce inequality by simulating who is included/excluded when geomarkets are utilized within student list purchases as recommended by @RN4982.
--->
-```
-# Enrollment Management and Quantification
-
-**Enrollment Management**
-
-<!-- la la -->
-
-Enrollment management is simultaneously a profession, an administrative structure, and an industry.
-As a profession, enrollment management (EM) integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p.
-xiv].
-As an administrative structure, the EM office typically controls the activities of admissions, financial aid, and recruiting [@RN2406].
-<!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
--->
-
-@fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
-The funnel begins with a large pool of "prospects" (i.e., prospective students) that the university would like to enroll.
-"Leads" are prospects whose contact information has been obtained.
-"Inquiries" are prospects that contact the institution, including those that respond to an initial solicitation (e.g., email) and those that reach out on their own (e.g., sending SAT scores).
-<!--  Applicants consist of inquiries who apply plus "stealth applicants" who do not contact the university before applying.
-The funnel narrows at each successive stage in order to convey the assumption of "melt" (e.g., a subset of "inquiries" will apply).
--->The purpose of the enrollment funnel is to inform recruiting interventions that target one or more stages.
-These interventions seek to increase the probability of "conversion" across stages [@RN4322].
-At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz].
-Purchased leads are served emails, brochures, and targeted social media designed to solicit inquiries and applications [@RN4895].
-
-Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel it speaks to.
-The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
-Economists often investigate financial aid leveraging, which seeks to convert admits to enrolled students \[CITE\].
-<!--  Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935; @RN1814].
--->
-
-A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
-@RN4758 conceptualize recruiting behavior as an indicator of college enrollment priorities.
-Ethnographies by @RN3519 and @RN4407 identify connections between private school guidance counselors and college admissions officers as a mechanism for social reproduction.
-Recruiting visits to high schools are a means of maintaining ties with guidance counselors at feeder schools and establishing relationships with prospective students [@RN3519; @RN4895].
-Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly white public schools [@RN4758; @RN4759; @RN5023].
-From the student perspective, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
-@cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.
-<!--  Consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015], this literature reveals a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches segmented by class and race, a process consistent with the notion of that higher education is mechanism of social reproduction [@RN2377; @RN4873].
--->
-
-By focusing on the behaviors of individual colleges, scholarship on EM implicitly states that recruiting is something done by individual colleges.
-In addition to university personnel (e.g., admissions counselors, VP for enrollment management), the EM industry incorporates external stakeholders in the organizational field, including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
-We argue college access is structured by third-party servicers and products that interact with direct-providers (colleges).
-Although sociologists have hinted at the ways enrollment management contributes to inequality in college access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
-<!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
---> <!--
-[^1]: New institutional theory offers one explanation [@RN527].
-In highly developed organizational fields like higher education, forces of competitive and institutional isomorphism -- including institutional locics -- compel organizations adopt similar practices as their peers.
-Practices deemed advantageous spread quickly, for example, public research universities targeting nonresident prospects to compensate for declines in state appropriations.
-These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.
- -->
-
-Drawing from scholarship on organizational theory, enrollment management processes involve many "make or buy" [@RN4738] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
-EM consulting firms provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
-As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].
-In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
-
-Although the enrollment management industry is increasingly characterized by software-as-service products sold by private equity backed firms (e.g., for example, EAB's [Enroll 360](https://eab.com/solutions/enroll360/) product) [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
-The Student Search Service, created in 1972, became the ubiquitous means of identifying prospects BLAH BLAH.
-Second, EPS, launcched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
-Both products are applications of quantification.
-Student Search Service includes or excludes prospects based on selected attributes while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
-THEREFORE WE REVIEW SCHOLARSHIP ON QUANTIFICATION.
-
-```{=html}
-<!--
-
-The EM industry puts great stock in market research, in part because the EM profession emerged from marketing and market research [@RN4986; @RN2771].
-@RN4986 [p.
-17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].[^1] In recruiting, market research informs decisions about which prospective students to target with recruiting interventions.
-How to acquire market research is another make-or-buy decision.
-The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals.
-In 1983, College Board published *Applying Market Research in College Admissions* [@RN4986], which became the basis for their Admitted Student Questionnaire (ASQ) product, and published *The Structure of College Choice* [@RN4982], which became the basis for the Enrollment Planning Service (EPS) product/division.
-Today, EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
-Market research often distills large amounts of data into simpler indicators.
-Therefore, the quantification literature provides a useful lens to understand market research.
-
-[^1]: @RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
-In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited.
-In the chapter entitled "Travel," @RN3519 [p.
-69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."
--->
-```
-**Quantification**
-
-@RN5007 [p.
-402] define quantification as "the production and communication of numbers" (p.
-402).
-Reacivity, discipline, and authority are three interrelated themes that describe the effects of quantification.
-*Reactivity* is the idea that salient quantitative measures cause people and organizations to change their behavior.
-Quantification *disciplines* actors to react in particular ways.
-Quantification changes decision-making power *authority*, often weakening the discretion of local decision-makers.
-Scholarship about U.S.
-News & World Report (USNWR) Law School rankings demonstrate the effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
-Law school rankings were developed as a means of informing prospective students about the relative quality of different programs.
-Rankings affect (reactivity) where students apply and enroll \[CITE\].
-However, rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
-Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p.
-415].
-Rankings *disciplined* the behavior of law schools.
-For example, applicants with characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
-Rankings also weakened the *authority* of admissions directors and admissions officers to make independent decisions about which applicants to admit, ideal class size, and curricular offerinags [@RN5003].
-<!--
-In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.    -->
-
-The analysis of UK school "league tables" by @RN5005 shows how quantification can be a mechanism of social reproduction.
-In 1992, the UK government began publishing school league tables, which ranked schools based on student performance on national exams taken at age 16.
-League tables facilitate making evaluative comparisons between schools regardless of geographic proximity.
-They also discipline households and schools to conceive of school quality in terms of test scores, which are substantially a function of the class composition of schools.
-Using Census data measured at the local authority level (similar to a U.S.
-county), @RN5005 find that localities with higher performing schools experienced growth in the share of managerial/professional households following the adoption of league tables and a decline in the share of working-class households.
-Using longitudinal survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables.
-The introduction of league tables contributed to social reproduction because professional/managerial households were more aware of these consumer-facing metrics and had resources to respond by moving to more expensive neighbhorhoods, near higher performing schools.<!-- THIS IS RATCHET EFFECT STUFF.
-I CUT IT BECAUSE YOU HAVEN'T INTRODUCED THE IDEA OF RATCHET EFFECT
-Schools that performed well on initial league tables tended to be schools with a high share of professional/managerial families.
-Therefore, quantification takes a snapshot of existing social stratification and that snapshot becomes a power unto itself, amplifying the effect of previous strctural inequality on subsequent structural inequality.
--->
-
-```{=html}
-<!-- THIS IS THE CUT TRANSITION PARAGRAPH
-In the broader quantification literature, critical data studies [e.g., @RN4975; @RN4849; @RN4772] and sociological analyses of technology and inequality [e.g., @RN4778; @RN4786; @RN4877; @RN4810] introduce concepts salient to the analysis of market research.
-Our focus is on "selection devices," which utilize some combination of inputs and discretion to categorize places, people, and organizations for the purpose of allocating opportunities or interventions [@RN4778; @RN5012].
--->
-```
-**Correlation**.
-The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4772; @RN4849].
-In particular, the discussions of correlation (chapter 1) and homophily (chapter 2) by @RN4975 introduce ideas salient to the analysis the Market Segment Model and Geomarkets.
-Correlation measures the extent to which two or more variables move together.
-Predictive analytics are based on correlation and are developed in two steps.
-First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
-Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and to assign levels of risk to each case.
-@RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
-These models predict outcomes based on correlations without requiring knowledge about underlying causal relationships.
-@RN4975 [p.
-50] writes that "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"[^2]
-
-[^2]: Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
-For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
-@RN4849 [p.
-146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
-
-A bevy of studies show that predictions based on correlations reproduce structural inequality [for a review see @RN4794].
-The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point of time.
-The observed correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
-Reviewing scholarship about algorithms, @RN4794 [p.
-224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
-This phenomenon whereby has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^3] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
-They then recommended that selective colleges recruit from localities that have a high share of affluent, college-educated households.
-
-[^3]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
-Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
-Note that predictive models such as the LSI-R model are not merely used for social science purposes.
-Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.
-
-**Homophily**.
-Homophily is the idea actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together" [@mcpherson2001birds].
-Homophily is a core concept of network science (writ large), in which actors (nodes) are connected to one another directly and indirectly via network ties (edges).
-As discussed below, @RN4982 [chapter 4] *The company We Keep: Colleges and Their Competition* conducts a network analysis based on 1980 SAT score sending behavior whereby two colleges are defined as competitors if a large number of students send SAT scores to both colleges.
-Because network science models often draws from rational choice theory, @RN4975 argues, it assumes that homophily is the result of voluntary action by individuals.
-
-@RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
-Like correlational models discussed above, because most network science models do not observe how historical structures affect relationships, they "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
-95] that cause behavior consistent with homophily.
-Second, in commercial social networks, homopphily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
-Thus, "social networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@RN4975, p.
-YY].
-We observe similar processes in college access.
-Based on a snapshot of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
-In turn, @RN4982 reasons that colleges should target the Geomarkets and high schools targeted by peer colleges.
-This logic is subsequently programmed into EPS software that recommends which Geomarkets and high schools to recruit from.
-Thus, homophily observed on the demand side -- which is a function of historic structural inequality –- is programmed into the supply-side.
-
-Homophily is central to market research products that categorize customers.
-Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
-Market segments are subgroups within a larger market that have similar consumer demand.
-Early geodemographic classifications of consumers (e.g., PRIZM by Claritas Corporation) were derived from publicly available Census data, which disaggregated data to the zip code level.
-The Claritas Corporation had a financial incentive to argue that people living near one another share similar consumer preferences because geographic localities could then be categorized into market segments that would be useful for direct mail marketing campaigns [@RN4957].
-Later, the development of individual credit scores (e.g., FICO score) enabled merchants to classify consumers into many, fine-grained groups [@RN4877].
-@RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
-We observe similar processes in our case study.
-@RN4982 categorizes students into four market segments -- local, in-state, regional, and national -- based on their score sending behavior and then evaluates the attractiveness of each Geomarket based on how many "regional" and "national" students live there.
-<!--
-ORIGINAL TEXT ABOUT GEODEMOGRAPHIC CLASSIFICATION SYSTEMS
-
-Clusters were given evocative names like "Blue Blood Estates," "\[Archie\] Bunker's Neighbors", and "Share Croppers." Because U.S.
-publicly available U.S.
-Census data disaggregated data only to the zip code-level, the Claritas Corporation had a strong incentive to argue that consumer behavior within zip codes was very similar [@RN4957].
-@RN4565 emulated the behavior of Claritas by developing the *Segment Analysis Service: An Educationally Relevant Geodemographic Tagging Service*, which was an "add-on" to the *Student Search Service* student list product.
-Describing its conceptual underpinnings, @RN4565 [p.
-1] states, "the basic tenet of geodemography is that people with similar cultural backgrounds, means, and perspectives naturally gravitate toward one another or form relatively homogeneous communities; in other words, birds of a feather flock together." The Segment product allocated each high school and each census tract into a distinct "educational neighborhoods" thought to reflect useful market segments for colleges looking for students.
-
-PREVIOUS TEXT ABOUT CLASSIFICATION SITUATIONS
-Related to the creation of market segments, @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Historically, classifications were binary.
-Consumers with "good" credit were offered loans and those with bad credit were not.
-The innovation of individual FICO scores -- and, later, other credit scores -- enabled merchants to classify consumers into many, fine-grained groups or along a continuum.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].
-People previously denied credit were now granted credit on different terms.
-For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-Thus, at the bottom of the continuum, contemporary classification situations produce "predatory inclusion" -- exemplified by payday loans and for-profit colleges -- which is the logic of "including marginalized consumer-citizens into ostensibly democratizing mobility schemes on extractive terms" [@RN4774, p.
-443].
-
-Negative consequences of homophily
-Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
-
-Reviewing scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-U.S.
-law school rankings [@RN5003] and UK school league tables [@RN5005] exemplify this sort of salience.
-Studies of quantification in the sociology of education tend to study consumer-facing metrics.
-Law school rankings can be conceived as a business-to-consumer product because the goal is to inform consumers (prospective law school students) and consumers pay for the product.
-
--->
-
-[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-
-**Contribution**.
-As @RN3519 demonstrates in the chapter *Numbers*, enrollment management is fundamentally concerned with quantification.
-In the case of college rankings, sociology made visible how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges who are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
-Beyond rankings, sociological analyses of enrollment management focus on the behavior of colleges and their agents, yielding the implicit assumption that inequalities creates by enrollment management are a function of individual college behavior.
-We argue that third-party vendors and products structure the enrollment management behavior of colleges.
-These products take a snapshot of student demand -- without considering structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
-Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
-These products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociologists.
-
-This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
-The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
-Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
-Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification.
-However, this study provides inisight about the mechanism underlying the effect.
-By analyzing and simulating *Student Search Service* purchases that filter on Geomarkets we show how Geomarkets reproduce historical race-based inequality in college access.
-
-```{=html}
-<!--
-The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
-@RN5021 reports that in 1995, 37% of 4-year public universities and 49% of 4-year private universities used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
-Geomarkets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities.
-
-By contrast, many quantification projects -- for example, human rights indicators -- have "little oomph" [@RN5006, p.
-265].
-FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created."
--->
-```
-# The Market Segment Model and College Board Geomarkets
-
-```{=html}
-<!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.
-
-As a member organization, several College Board products developed from joint collaborations with administrators or professors from member institutions.
-During the 1970s, under the leadership of Larry Litten, then director of institutional research, Carleton College pionered applying market research to student recruiting.
-In the late 1970s, College Board provided Carleton College with funding and data for a six-year project that analyzed college choice decisions by households in six metropolitan areas.
-This project was the basis for the landmark book *Applying Market Research in College Admissions* [@RN4986], published by College Board, and was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll.
-Litten would later become Associate Director of the Consortium of Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
-
-The Market Segment Model was an effort to quantify and communicate the knowledge that admissions officers have about student demand. 
--->
-```
-**Creating Geomarkets and the Market Segment Model.**
-
-PUT THIS SOMEWHERE:
-
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
-9-10).
-
-In 1978, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
-x].
-Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
-<!--The geographic concepts inscribed in the Market Segment Model were based on conversations with college admissions officers.
---> The "initial task was to define enrollment markets in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982,p.
-4].
-Quoting an admissions officer, @RN4982 [p.
-11] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, [@RN4982] created three geographic [@RN4982, p.
-11], region, state, and community.
-The three regions were New England, Middle States, and the South.
-They "divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" (p.
-11).
-These enrollment markets, later called Geomarkets, were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
-@RN4982 [pp.
-11-12] briefly describes the creation of Geomarket borders:
-
-> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
-Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
-In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."
-
-The goal of the Market Segment Model was to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households in the Geomarket.
-Zemsky reached out to College Board because "we needed a database that described most institutions and most students" [@RN4982, p.
-x].
-<!--Zemsky reached out to College Board because "we needed a database that described most institutions and most students\ldots Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential" [@RN4982, p.
-x].
--->In 1979, College Board began providing funding and data on the score-sending behavior of SAT test-takers.
-
-> For our analysis, we sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
-The Market Segment model that we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
-The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways [@RN4982, p.
-4].
-
-SAT test-takers were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
-For a given student, each college that receives a score from the student can be defined as "local" (college located in the the same Geomarket as the student), "in-state" (same state but different Geomarket), "regional" (same region but different state), or "national" (different region).
-In turn, a test-taker is categorized in the "local" market segment if they submit more SAT scores to local institutions than they do to in-state, regional, or national institutions.
-An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
-<!--
-A "regional" student submits more SAT scores to regional institutions than they do to local, in-state, or national institutions.
-A "national" student submits more SAT scores to national institutions than they do to local, in-state, or regional institutions.
--->
-
-The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
-Appendix A describes these outputs in more detail.
-Both outputs are created separately for each Geomarket.
-For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
-A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
-These outputs became the basis for the Enrollment Planning Service (EPS) software.
-
-[^5]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
-
-[^6]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
-This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
-
-**Correlations**.
-@RN4982 [Chapter 3], *A Sense of Place: Students, Families, and Communities*, identifies the student characteristics correlated with being in the local, in-state, regional, or national market segment.
-The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
-These four variables "reflect the basic social patterns of the nation.
-It would have been surprising if these were not the four social variables that best explained the patterns of college choice" [@RN4982, p.
-33].<!--
-and "The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice" (pp.
-34-35).
-\[QUOTES REDUNDANT\]
--->
-
-The thesis of Chapter 3 is that student demand for higher education is a function of social origin.
-For example, @RN4982 \[p.
-33\] state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
-33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."[^7] @RN4982 \[p.
-42\] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
-Therefore, @RN4982 [p.
-44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:
-
-[^7]: The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
-Additionally, analyses reveal that the SAT score is the most important predictor of score-sending behavior, followed by parental education, family income, and educational aspirations.
-However, prior research shows that SAT scores are substantially a function of social origin [@RN1273].
-
-> On occasion, senior spokespersons for the profession worry that students outside the main \[Geo\]market areas remain forgotten and hence, unchallenged.
-Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
-The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
-
-**Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^8] Based on these analyses, @RN4982 \[p.
-46\] state that homophily describes the nature of competition between colleges: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and socioeconomic composition of institutions in competition with one another.
-Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-The authors argue that like-colleges compete for like-students as defined by socioeconomic characteristics.
-@RN4982 [p.
-72] describe these patterns as a natural process of homophily in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
-
-[^8]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
-The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
-Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
-For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
-54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
-
-> Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
-The layering of collegiate competition is primarily a socioeconomic layering.
-The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
-Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
-
-The discussion of competition by @RN4982 exemplifies the concerns about correlation and homophily described by [@RN4975].
-A correlational analysis of 1980 SAT score-sending patterns finds competition between colleges is defined by socioeconomic homophily.
-This homophily is presented as a naturally occurring phenomenon.
-Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer-coleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
-In itself, @RN4982's Market Segment Model is merely a social science depiction of the structure of student demand, one that does not consider historic, structural inequalities that cause observed patterns.
-However, by inscribing the Market Segment Model and Geomarkets into the EPS software, College Board amplified structural inequalities that contributed to homophily observed in 1980 SAT score-sending beahivor.
-In this way, commodification of the Market Segment Model and Geomarkets "engineers homophily" [@RN4975].
-
-**Enrollment Planning Service**.
-In 1984, College Board created the Enrollment Planning Service (EPS), based on the Market Segment Model [@RN4987; @RN4980].
-EPS was an early software-as-service platform that recreated the analyses of @RN4982.
-For each Geomarket, colleges could obtain Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors.
-Based on conversations with enrollment management professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket.
-Therefore, colleges used EPS software to decide which Geomarkets to recruit from and which high schools to visit within targeted Geomarkets.
-Typical @RN5019 marketing material describes EPS as,
-
-> The marketing software that pinpoints the schools and Geomarkets where your best prospects are most likely to be found.
-With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
-Focus your valuable time and resources on the right prospects.
-
-```{=html}
-<!--
-Promotional material highlights the themes of "locate the students you want" [@RN5019_2003] and "focus your valuable time and resources on the right prospects" [@RN5019].
-@RN5019_2010 promotional material describes EPS as:
-
-> the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
-You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, zip codes, high schools and international regions.
-EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition."
-
-Software documentation by @RN5017 states that "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
--->
-```
-Whereas @RN4982 identified 143 Geomarkets covering the New England, Middle States, and South region, EPS created Geomarkets for the remaining U.S.
-states, with 304 Geomarkets in total.
-@RN4988 shows the contemporary Geomarkets.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^9] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
-
-[^9]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
-Students from California don't all share the same college-going behaviors.
-We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."
-
-@RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
-Market research suggests that EPS was highly salient.
-@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
-
-EPS software may discipline colleges to approach recruiting in a manner consistent with the Market Segment Model.
-Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
-EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^10] Second, EPS users decide which high schools they will visit within selected Geomarkets.
-Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-
-[^10]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
-
-EPS may also weaken the authority of local decision-makers.
-@RN4982 sought to develop a concrete, data-driven framework -- The Market Segment Model -- that replicates the aggregate knowledge of local admissions officers.
-Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
-The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
-In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
-However, EPS software recommended visiting the same sets of affluent, high-achieving high schools that were receiving visits from other colleges.
-We were told that savvier, quantitatively-adept admissions offices used EPS to visit schools explicitly not recommended by EPS because there would be less competition for the good students that attended these high schools.
-
-Finally, we suspect that Geomarkets affect how admissions offices organize the recruiting territories assigned to each admissions officer.
-Geomarkets were created to mirror the territories of admissions officers [@RN4982].
-In turn, as Geomarkets became more salient, admissions offices often structured their territories around Geomarkets.
-On background, admissions officers and enrollment consultants told us it was common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket.
-In those states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
-This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX\], \[ADD TO APPENDIX\].
-<!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
-
-Although we cannot gain access to EPS software, Geomarkets are fundamental to EPS software and (arguably) to the organization of recruiting territories for many colleges.
-Therefore, we are interested in the extent to which this building block is associated with race and class.
-Research question 1 asks, what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
-We focus on metropolitan areas that are associated with three or more Geomarkets.
-
-We expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
-Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
-Moreover, @RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards distinguishing geographic areas that differed from one another in terms of class composition and college-going behavior.
-
-```{=html}
-<!-- 
-Additionally, @RN4982 created geomarkets in service of the Market Segment Model.
-The Market Segment Model argues that socioeconomic status drives whether students belong to the local, in-state, regional, or national market segment.
-In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status.
-The language of @RN4982 [pp.
-11-12, quoted above] suggests that geomarket borders within large metropolitan areas may have been drawn in ways that separate affluent from less affluent communities.
-We present the following hypothesis:
--->
-```
-We also expect substantial racial inequality between geomarkets in large metropolitan areas for several reasons.
-Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race once.
-U.S.
-cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779].
-Structures built upon racialized structures are racialized structures [@RN4786].
-Unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
-Second, geomarket borders may have been drawn along class divides.
-A strong correlation exists between race and wealth [@RN4884].
-Third, Geomarket borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
-Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
-
-**Geomarket Filter in Student Search Service**.
-By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service \[CITE\].
-A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
-
-[^11]: For public universities that purchased lists, 80% purchase more than 50,000 names annually.
-@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.
-
-Student lists are positively associated with student outcomes.
-A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
-After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^12] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^13]
-
-[^12]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
-Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
-
-[^13]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
-Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
-The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).
-
-However, @RN5012 argue that student list products exacerbate racial inequality in college access.
-@RN5012 conceptualize student list products as "selection devices" [@RN4778] that enable colleges to select which prospective students they target by incorporating search filters (e.g., high school graduating class, state).
-@RN4786 defines "racialized inputs" as ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
-@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) meet the criteria of racialized inputs.
-Using a national sample of high school students and using data from actual student lists purchased by public universities, @RN5012 show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
-
-Because of the extent of residential segregation in the U.S., geographic borders are a commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887].
-Student list products offer many geographic filters.
-Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^14] This manuscript focuses on Geomarkets.
-Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
-First, they are ostensibly race-neutral inputs in that neither @RN4982, nor subsequent promotional material [e.g., @RN1621], mention race.
-Second, Geomarkets may be systematically correlated with race.
-For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda Country, excluding Oakland." A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.
-This article extends @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
-
-[^14]: For example, the geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
-
-@RN4982 [pp. 42-44] discuss how the Market Segment Model and Geomarkets should be utilized in the context of purchasing student lists. The authors ask,
-
-> Imagine for the moment that you recruit for a college that draws most of its students from regional or national segments.
-Where would you concentrate your energies?
-Ideally, you would seek communities \[Geomarkets\] with a high proportion of students already predisposed toward institutions such as your own.
-The Market Segment Model would provide this information through segment percentages for the community in question.
-Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
-Your efforts would surely be better directed toward \ldots Manchester, Hartford, and Fairfield County \ldots Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own.
-
-In both mailings and off-campus recruiting visits, @RN4982 recommend focusing on Geomarkets with large populations of students who are deciding between "institutions like your own" (p.
-44).
-Selective colleges primarily draw from students in the "regional" and "national" segments.
-@RN4982 suggest that the student list purchases of selective colleges should focus on Geomarkets with large numbers of affluent, highly educated households.
-By contrast, the student list purchases of local state colleges and community colleges, which rely on enrollment from students in the "local" and "in-state," should focus on nearby Geomarkets with large numbers of low-SES students.
-Our analyses attempt to recreate the strategy recommended by @RN4982 by showing who is included/excluded when student list purchases filter on particular Geomarkets.
-Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access.
-Furthermore, we anticipate that the students of color from affluent Geomarkets will tend to have college-educated parents and be enrolled in affluent, predominantly white schools.
-
 # Data and Methods
@@
 
-**Research question 1, data**.
-RQ1 asks, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas?
-And, how does this variation change over time?
-To answer this question, we use census tract-level data from the US Census Bureau.
+Our study utilizes two main data sources based on each research question.
+The first research question is focused on understanding the socioeconomic and racial variation between Geomarkets in metropolitan areas and how such variation changes over time.
+To answer this question, we use census tract data from the U.S.
+Census Bureau.
+Utilizing a small geographic area is important for implementing a spatial merge between Census data and Geomarket shape files (described below).
 Census tract is the smallest geographic area for which measures of both race/ethnicity and socioeconomic characteristics are available.
-Utilizing a small geographic area is important for implementing a spatial merge between Census data and Geomarket shape files, described below.
-
-We utilize data from the 1980 Decennial Census, specifically variables from Summary Tape File 1 (STF1), which contains data from the about race and hispanic origin from the "short form" questions answered by all households and data from Summary Tape File 3 (STF3), which contains data about socioeconomic characteristics -- including income, education, and poverty -- derived from the "long form" questionnaire completed by a sample of households.
+Census tracts are statistical subdivisions of a county designed to capture relatively homogeneous demographic communities that are generally considered to be more accurate for population analyses than zip codes, which are spatial boundaries drawn to demarcate route efficiency for postal services \[[CITE](https://www.census.gov/newsroom/blogs/random-samplings/2014/07/understanding-geographic-relationships-counties-places-tracts-and-more.html)\] .
+
+For the first research question, we use data from the 1980 and 2000 Decennial Census as well as 5-year estimates from the 2020 American Community Survey (ACS).
+For the 1980 Decennial Census, we specifically draw on Summary Tape File 1 (STF1) for variables about race and Hispanic origin collected via “short form” questions answered by all households and Summary Tape File 3 (STF3) for variables about socioeconomic characteristics (income, education, and poverty) collected via the "long form" questionnaire completed by a sample of households.
 The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-In order to examine how the characteristics of Geomarkets changed over time, we utilize tract-level data from the 2000 Decennial Census, similarly using data from Summary File 1A (SF1A) to obtain measures of race and Hispanic origin and using data from Summary File 3A (SF3A) to obtain measures of income, education, and poverty.[^15] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
-These questions were replaced by the American Community Survey (ACS).
-We also utilized data from the 2020 5-year ACS, which includes data collected from 2016-2020.[^16] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
-
-[^15]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
-
-[^16]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
+To examine how Geomarket characteristics change over time, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty from the 2000 Decennial Census.[^1] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+These questions were replaced by ACS.
+Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.[^2] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+
+[^1]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
+
+[^2]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
 
@@
 Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^17] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
-Using these zip code-level data, we utilized the `aggregate()` function to create shapefile for each Geomarket.
-
-[^17]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
-
-Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated shapefiles.
-Census tract shapefiles that intersected with multiple Geomarkets were broken into smaller shapefiles wholly contained within a single Geomarket.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^3] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Using these zip code-level data, we utilized the `aggregate()` function to create shapefiles for each Geomarket.
+
+[^3]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
+
+Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
+Census tract shapefiles that intersected with multiple Geomarkets (i.e., a census tract's spatial boundaries were spread across two or more Geomarkets) broken into smaller shapefiles wholly contained within a single Geomarket.
 The variables in these shapefiles were then replaced with the original variable value multiplied by the proportion of land area of the census tract that was contained within the Geomarket.
@@
 We also created socioeconomic measures that approxmiate the predictors of market segment identified by @RN4982.
-These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^18] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
-
-[^18]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
+These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^4] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
+
+[^4]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
 
@@
 We draw from these data to answer RQ2.
-Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^19].
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^5].
 Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
 
-[^19]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
+[^5]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
 

### 2025-04-30 text/front_end_v3.qmd
finished editing front end

@@
 **Geomarket Filter in Student Search Service**.
-By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service \[CITE\].
-A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
+<!-- By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service \[CITE\].
+--> A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
 Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
@@
 
-However, @RN5012 argue that student list products exacerbate racial inequality in college access.
-@RN5012 conceptualize student list products as "selection devices" [@RN4778] that enable colleges to select which prospective students they target by incorporating search filters (e.g., high school graduating class, state).
+However, @RN5012 argue that student list products may exacerbate racial inequality in college access.
+The authors conceptualize student list products as "selection devices" [@RN4778] that enable colleges to select which prospective students they target by incorporating search filters (e.g., high school graduating class, state).
 @RN4786 defines "racialized inputs" as ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
-@RN5012 argue that several frequently utilized student list filters (zip code, AP test score, SAT score) meet the criteria of racialized inputs.
+@RN5012 argue that several frequently utilized student list filters (e.g., zip code, AP test score, SAT score) meet the criteria of racialized inputs.
 Using a national sample of high school students and using data from actual student lists purchased by public universities, @RN5012 show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
@@
 College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^14] This manuscript focuses on Geomarkets.
+In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
-First, they are ostensibly race-neutral inputs in that neither @RN4982, nor subsequent promotional material [e.g., @RN1621], mention race.
+First, they are ostensibly race-neutral inputs in that neither @RN4982 -- nor subsequent EPS promotional material [e.g., @RN1621] -- mention race.
 Second, Geomarkets may be systematically correlated with race.
@@
 
-[^14]: For example, the geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
+[^14]: Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
 
@@
 Ideally, you would seek communities \[Geomarkets\] with a high proportion of students already predisposed toward institutions such as your own.
-The Market Segment Model would provide this information through segment percentages for the community in question.
+The Market Segment Model would provide this information through segment percentages for the community [Geomarket] in question.
 Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.

### 2025-04-30 text/discussion.qmd
started revising discussion

@@
 
-The discussions of correlation and homophily by @RN4975 describe the logic of The Market Segment model and also how the enrollment management industry thinks about the process of recruiting students.
-The thesis of the Market Segment Model is that of the status attainment model, that student demand for college is a function of parental education and parental income.
-@RN4982 demonstrated this thesis by showing the correlation between student SAT score-sending behavior and measures of class.
-@RN4982 argued that households of particular class and college aspirations are likely to live in particular parts of a metropolitan area and that these geographic territories can be meaningfully captured by Geomarkets.
-This argument relies on the assumption of homophily, that birds of a feather flock together.
-Drawing from this logic, colleges should identify their core student market segment -- local, in-state, regional, or national -- identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
+The discussions of correlation and homophily by @RN4975 describe the logic of The Market Segment model but also the underlying institutional logic of the enrollment management industry.
+The logic of correlation argues that colleges should find students who look like students who enrolled at your institution in the past.
+The logic of homophily argues that there is a cultural match between institutions and students, for example, that  Catholic universities compete with Catholic universities for Catholic students [@RN4982, p.
+84].
+Therefore, @RN4982 recommends that colleges recruit from the same schools and communities that are popular with peer colleges.
+The operating logics of correlation and homophily are highly consistent with college recruiting behavior observed in the wild.
+For example, @RN5023 find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
+The also find great overlap in the schools visited by Villanova and Notre Dame, with a disproportionate focus on visits to Catholic schools.
+When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
+As @RN4975 argues, observed homophily is not the result of voluntary action; rather, homophily is programmed into algorithms that create connections.
+
+The Market Segment Model begins with the status attainment model, that student demand for college is a function of parental education and parental income.
+@RN4982 demonstrated this argument by showing the correlation between student SAT score-sending behavior and measures of class.
+The thesis of the Market Segment Model is that that households of particular class and college aspirations are likely to live in particular parts of a metropolitan area and that these geographic territories can be meaningfully captured by Geomarkets.
+Although they appear coarse by contemporary standards, Geomarkets were a landmark innovation in Geographic segmentation.
+The Market Segment Model recommends that colleges should identify their core student market segment -- local, in-state, regional, or national -- identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
 This information is contained in the Market Segment Profile (Appendix A), which is a standard output of EPS software.
-When considering new Geomarkets to target, colleges should consider which Geomarkets have large numbers of households that are interested in colleges that are direct competitors.
+When considering new Geomarkets to target, colleges should utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
 This information is contained in the Institutional Profile (Appendix A).
-Thus, the recommendations of the Market Segment Model can be boiled down to finding those Geomarkets with your desired class of student and targeting those high schools and communities that are popular with colleges like yours.
-[WHAT IS MISSING HERE: YOU DON'T TALK ABOUT HOW A SNAPSHOT OF THE PAST IS USED TO MAKE DECISIONS ABOUT FUTURE RECRUITING]
+Technologies that target target customers based on geography leverage racial and class segregation [@RN4775; @RN4849; @RN4975].
+The Market Segment Model and Geomarkets were created based on 1980 SAT score-sending patterns, contributing to a "ratchet effect" [@RN4851].
+The structural barriers (e.g., segregation, slavery, Jim Crow) that caused inequality in 1980 student demand are amplified because the Market Segment Model recommends that colleges focus recruiting resources on localities that already have high student demand.
 
-Interestingly, @RN4982 claim that the Market Segment Model merely formalizes the logic that admissions officers from selective colleges had been utilizing for decades.
-The idea is that good students from good families want to go to good colleges that are far from home.
-These good families are found in particular parts -- Geomarkets -- of metropolitan areas and students from other parts of the metro have the achievement and disposition to attend their local college.
-The Market Segment Model formalized this logic and commodified it in Enrollment Planning Service (EPS) software that helped colleges identify desirable Geomarkets and desirable high schools within those Geomarkets.
-As suggested by @RN5007, once commodified and quantified, the local knowledge of admissions officers becomes redundant.
-
-Technologies that target target customers based on geography utilize racial and class segregation [@RN4775; @RN4849; @RN4975].
-Geographic segmentation has become increasingly granular over time.
-Geomarkets were a landmark innovation in Geographic segmentation in the context of education.
-College Board provided data for @RN4982 because "the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" (p.
-x).
+@RN4982 state that the Market Segment Model merely formalizes the operating logic of college admissions officers.
+For admissions officers at selective colleges, the logic is that good students from good families are willing to attend good colleges that are far from home.
+These good families are found in particular parts -- Geomarkets -- of metropolitan areas.
+By contrast, students from other parts of the metro possess the disposition and achievement to attend their local college.
+The Market Segment Model formalized this logic and commodified it in EPS software that helped colleges identify desirable Geomarkets and desirable high schools within those Geomarkets.
+@RN5007 suggest that, once commodified and quantified, the local knowledge of admissions officers becomes redundant.
 
@@
 We find that Geomarkets are highly correlated with race and class.
-[KARINA SUMMARIZE FINDINGS] We show that in 1980, when Geomarkets were being created, that Black people tended to be highly concentrated in the poorest Geomarket in the metropolitan area.
+[KARINA SUMMARIZE FINDINGS IN A COUPLE SENTENCES; REVISE PARAGRAPH AS NECESSARY] We show that in 1980, when Geomarkets were being created, that Black people tended to be highly concentrated in the poorest Geomarket in the metropolitan area.
 Examples include, CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), TX19 -- City of Dallas, CA21 -- South & South Central Los Angeles, and PA5 -- Philadelphia County.
@@
 These findings are not surprising given the extent of race and class segregation in the US.
-Furthermore, the correlation between Geomarkets and class is expcted because geomarket borders were created in service of the Market Segment Model, a model which views demand for higher education as a function of class.
-Interestingly, @RN4982 literally have nothing to say about race and demand for higher education.
+Furthermore, the correlation between Geomarkets and class is expcted because Geomarket borders were created in service of conceptual model that views demand for higher education as a function of class.
 @RN4982 is written around New England, with a particular focus on Geomarkets in Connecticut and Massachusetts.
-Their silence is about race is conspicuous given that Busing was a political contentious issue in Boston and nearby metropolitan areas.
+@RN4982 do not discuss the intersection of student demand and race.
+Although unsurprising for early 1980s social science, this silence is conspicuous considering the salience of political controversy surrounding "busing" in Boston and nearby metropolitan areas.
 
@@
 
-These analyses contribute to scholarship about how students are sorted into colleges.
+These analyses contribute to scholarship about how students are sorted into colleges [IMPROVE TRANSITION].
 The status attainment model argues that college destination is a function of parental education and occupation [@RN1273; @RN1271; @RN1272].

### 2025-05-01 text/back_end_KS.qmd
methods

@@
 
+**RQ1**.
 Our study utilizes two main data sources based on each research question.
@@
 
+Across these data sources, we created measures of race and ethnicity, specifically the number and proportion of the census-tract population in the following categories: White, non-Hispanic; Black, non-Hispanic; Asian, non-Hispanic; Pacific Islander, non-Hispanic; American Indian and Alaskan Native, non-Hispanic; and two or more races, non-Hispanic.
+Measures of Asian, American Indian and Alaskan Native, two or more races, and Native Hawaiian and Pacific Islander (all non-Hispanic) were not available for the 1980 Decennial Census.
+We also created socioeconomic measures that approximate the predictors of market segment identified by @RN4982.
+These variables are median household income, percent of households below the poverty line, and people age 25+ with a BA.
+All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
+
 We assigned each census tract to College Board Geomarkets by implementing a partial spatial merge.
@@
 
-**Research question 1, variables**.
-RQ1 asks, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas?
-And, how does this variation change over time?
-We created measures of race and ethnicity, specifically the number and proportion of people (all ages) in the following categories: White, non-Hispanic; Black, non-Hispanic; Asian, non-Hispanic; Pacific Islander, non-Hispanic; American Indian and Alaskan Native, non-Hispanic; and two+ races, non-Hispanic.
-Measures of XXX were not available for 1980 Decennial Census.
-We also created socioeconomic measures that approxmiate the predictors of market segment identified by @RN4982.
-These variables are median household income, mean household income, percent of households below the poverty line, and people age 25+ with a BA.[^4] All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
-
-[^4]: STATE THAT MEAN HOUSEHOLD INCOME CREATED BY DIVIDING AGGREGATE INCOME BY NUMBER OF PEOPLE. NEED FIX THIS MEASURE THO.
-
-**Research question 2, data**.
-Geomarket is a search filter available in the College Board Student Search Service product.
-RQ2 asks, how does the racial and socioeconomic composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
+**RQ2**.
+Our study's second research question asks how does the racial and socioeconomic composition of included versus included prospects vary when student list purchases filter on Geomarkets.
 We answer this question using actual student lists purchased by public universities.
-
 In February 2020, we began issuing public records requests to public universities about student lists purchased from 2016 through 2020.
@@
 
-The data collection sample for this project was all public universities in CA, IL, MN, and TX.
+The data collection sample for this project was all public universities in California, Illinois, Minnesota, and Texas.
 Utilizing public records requests to obtain public records is a painstaking process.
-Initially, the majority of universities did not respond to our request or denied our request.
+Initially, the majority of universities did not respond or denied our request.
 Subsequently, we obtained pro bono representation from four law firms, which substantially increased the success of data collection.
-We collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
-However, we were unable to obtain legal representation for TX.
+We also collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
+However, we were unable to obtain legal representation for Texas.
 
@@
 Some universities provided records that were not usable for quantitative analyses (e.g., summary statistics across multiple orders; or data did not contain important fields).
-Some universities did not provide records based on legitimate grounds (e.g., data not in university possession; not required to create records that do not currently exist).
-We learned that many universities outsourced student list purchases to a third-party consulting firm.
+Some universities did not provide records based on legitimate legal grounds (e.g., data not in university possession; not required to create records that do not currently exist).
+We learned that many universities outsourced student list purchases to third-party consulting firms.
 Unfortunately, we were rarely able to obtain usable data from these universities.
@@
 
-This article analyzes student lists purchased from College Board.
-We exclude student list purchases by MN public universities from this report because Minnesota is predominantly an "ACT state." 14 public universities provided PDF order summaries containing the search criteria for 830 student list purchases.
-We utilized the XXXX Python/R package to convert these order summaries into tabular data.
-About 16% of student list purchases utilized the Geomarket filter.
-
-We received both (a) the order summary data and (b) the de-identified prospect-level student list data for 414 orders associated with 2,549,085 prospects.
-We draw from these data to answer RQ2.
-Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^5].
+The primary data source for our second research question is therefore student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
+This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Minnesota, and Arizona \[KS Does this still apply?\].
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^4] The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
+In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.
 Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
 
-[^5]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
+[^4]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
 
-In particular, we answer RQ2 by analyzing particular student list purchases that did *not* utilize the Geomarket search filter, but did purchase prospects from all Geomarkets in a particular metropolitan area.
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter, but did purchase prospects from all Geomarkets in a particular metropolitan area.
 For each prospect we know their home zip code and, therefore, which Geomarket they belong to.
-We answer RQ2 by simulating which prospective students would have been included or excluded from the student list purchase had the purchase filtered on partiulcar Geomarkets.
+We therefore simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
 
-**Research question 2, variables**.
-The variables of interest for RQ2 the characteristics of prospective students who would have been included/excluded by our simulations and the characteristics of their schools and neighborhoods.
-Prospect characteristics are derived from the pretest questionnaire administered to College Board test-takers.
-The de-identified student lists we received contain a standard subset of variables (e.g., HS code, HS GPA range, intended major).
-We create detailed measures of race and ethnicity.
-Information about student socioeconomic status is limited.
-However, we create a measure of whether the prospect is first-generation college student.
-We can also measure the socioeconomic characteristics of the students' home zip code and characteristics of their high school \[EXPAND OR DELETE THIS SENTENCE DEPENDING ON WHETHER WE USE THESE DATA\].
+The variables of interest for student list data are the characteristics of prospects who would have been included versus excluded by our simulations by Geomarket filter.
+Prospect characteristics are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
+College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity nonresponse) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016).
+From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016).[^5] Information about student socioeconomic status is limited.
+However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
+
+[^5]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
+Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”
 
@@
 
-**Research design**. We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
-
-Following XXX \[CITE\], we focus on a small number \[HOW MANY\] of metropolitan areas in order to provide sufficient analytic depth while also situating each case within the context of historic segregation and contemporary gentrification.
+**Research design**.
+We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
+The research questions of this study aim to explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
+Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarkets landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics.
+Such aims align with the purpose of a multiple case study design to gain a deeper understanding of the phenomenon of interest by comparing and contrasting across different settings \[CITE YIN 2001\].
+In doing so, multiple cases allow researchers to explore the "how" and "why" of the phenomenon in rich and nuanced contexts.
+We therefore focus on minimizing the number of metropolitan areas in order to provide sufficient analytic depth while also situating each case within its specific historic and contemporary geographical landscape \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
 Online Appendix XXX presents results for a larger number of metropolitan areas.
 
-Choice of metropolitan areas is informed by several factors.
+Case selection was informed by based on purposeful sampling to identify “information-rich” cases (Patton, 2002; Yin, 2009) that can provide in-depth knowledge about the phenomenon of interest.
+We first considered metropolitan areas as information rich cases across two broad dimensions: 1) Geographical and urban model structures and 2) variation across racial/ethnic and socioeconomic characteristics.
 Based on our read of @RN4982, we are interested in "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
@@
 Although Census data are available for all metropolitan areas (RQ1) we do not have good candidate student list purchases for all metropolitan areas (RQ2).
+
+Then, our purposeful sampling approach also took into consideration student list orders!
 \[SAY SOMETHING ABOUT WHAT IS A GOOD CANDIDATE STUDENT LIST PURCHASE -- PURCHASE WITH NOT TOO MANY CRITERIA THAT INCLUDE ALL GEOMARKETS IN A METRO\].

### 2025-05-01 text/back_end_KS.qmd
methods smoothing

@@
 We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
-The research questions of this study aim to explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
-Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarkets landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics.
-Such aims align with the purpose of a multiple case study design to gain a deeper understanding of the phenomenon of interest by comparing and contrasting across different settings \[CITE YIN 2001\].
-In doing so, multiple cases allow researchers to explore the "how" and "why" of the phenomenon in rich and nuanced contexts.
-We therefore focus on minimizing the number of metropolitan areas in order to provide sufficient analytic depth while also situating each case within its specific historic and contemporary geographical landscape \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
+Case studies are used to gain a deeper understanding of the phenomenon of interest by comparing and contrasting across different settings \[CITE YIN 2001\].
+Our study's research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
+Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics.
+Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts.
+We therefore focus on minimizing the number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
 Online Appendix XXX presents results for a larger number of metropolitan areas.
 
-Case selection was informed by based on purposeful sampling to identify “information-rich” cases (Patton, 2002; Yin, 2009) that can provide in-depth knowledge about the phenomenon of interest.
-We first considered metropolitan areas as information rich cases across two broad dimensions: 1) Geographical and urban model structures and 2) variation across racial/ethnic and socioeconomic characteristics.
-Based on our read of @RN4982, we are interested in "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
-11-12).
-Building on our conceptual framework, we are interested in racial and socioeconomic inequality between geomarkets within a metropolitan area (RQ1) and how this inequality contributes to inequality in which prospective students are targeted by student list purchases (RQ2).
-\[DISCUSSION OF WHICH CASES TO SELECT.
-KARINA\].
-Choice of metropolitan areas is informed by data availability.
-Although Census data are available for all metropolitan areas (RQ1) we do not have good candidate student list purchases for all metropolitan areas (RQ2).
+Case selection was therefore informed by purposeful sampling to identify “information-rich” cases (Patton, 2002; Yin, 2009) that can provide in-depth knowledge about variation across Geomarkets and overtime.
+We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
+Based on our read of @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
+11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, Philadelphia**,** Phoenix, and Pittsburgh.
+To compare and contrast with classic urban model structures, we also identified alternative urban models such as bimodal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
+The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S.
+regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations.
+Cumulatively related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2).
+We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.
 
-Then, our purposeful sampling approach also took into consideration student list orders!
-\[SAY SOMETHING ABOUT WHAT IS A GOOD CANDIDATE STUDENT LIST PURCHASE -- PURCHASE WITH NOT TOO MANY CRITERIA THAT INCLUDE ALL GEOMARKETS IN A METRO\].
+Lastly, choice of metropolitan cases was informed by student list data availability.
+Although Census data are available for all metropolitan areas nationally (RQ1), we do not have good candidate student list data for all metropolitan areas (RQ2).
+In order to simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets, student list data must contain orders that resulted in a large pool of prospects within the metropolitan areas considered above that would include all Geomarkets within metropolitan areas.
+This lead us to select on orders that filtered for an entire state (e.g., State filter is set to Pennsylvania), an entire metropolitan area without a Geomarket filter (CBSA filter is set to the Philadelphia metropolitan area) , or an entire metropolitan area with a Geomarket filter inclusive of all Geomarkets (CBSA filter is set to the Philadelphia metropolitan area and Geomarket filter is set to all five Geomarkets in the Philadelphia area).
+In order to isolate the impact of Geomarkets on the racial/ethnic and socioeconomic characteristics of prospect pools, a suitable metropolitan case candidate would also be contingent on student list orders that did not include too many filtering criteria beyond those detailed previously.
+Because nearly all orders we collected (Author, BLINDED) filter on some academic criteria, we only considered orders that included assessment threshold filters (e.g., PSAT, SAT) across a wide score range to explore whether and to what extent Geomarket contributions to prospect pools are consistent across all score ranges.
+
+Our case selection process resulted in three metropolitan cases: Chicago, Dallas-Fort Worth, and Los Angeles.
 

### 2025-05-01 text/back_end_KS.qmd
methods smoothing

@@
 We answer RQ1 by producing Geomarket-level tables and graphs that show how the Geomarkets in a selected metropolitan area vary on racial and socioeconomic characteristics and how they vary over time.
-For example, in the San Francisco Bay Area, how does the City of Oakland Geomarket (CA07) differ from the Alameda County Exluding Oakland Geomarket (CA08).
 We also produce interactive maps at the census tract-level to show more granular variation within and between Geomarkets.
@@
 Descriptive tables and graphs describe the racial and socioeconomic characteristics of prospects that would have been included/excluded had the purchase filtered on particular Geomarkets.
-Interactive maps provide a more granual visualization of where prospects lived and went to high school.
-For example, for purchased profiles who lived in the Chester County Geomarket (PA02) \[OR PHILADELPHIA PA05\] and identified as Black, where in the county did these students live and which high schools did they attend?
+We also create interactive maps of prospect pools by Geomarket contributions to provide a more granular visualization of analyses.
+
+```{=html}
+
+<!-- [DO WE NEED THIS LEVEL OF DETAIL???]
+
+For example, for purchased profiles who lived in the Chester County Geomarket (PA02) [OR PHILADELPHIA PA05] and identified as Black, where in the county did these students live and which high schools did they attend?
+
+ For example, for purchased profiles who lived in the Chester County Geomarket (PA02) [OR PHILADELPHIA PA05] and identified as Black, where in the county did these students live and which high schools did they attend?
 
@@
 In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.
+
+-->
+```
 

### 2025-05-02 text/discussion.qmd
working through discussion

@@
 
-The discussions of correlation and homophily by @RN4975 describe the logic of The Market Segment model but also the underlying institutional logic of the enrollment management industry.
-The logic of correlation argues that colleges should find students who look like students who enrolled at your institution in the past.
-The logic of homophily argues that there is a cultural match between institutions and students, for example, that  Catholic universities compete with Catholic universities for Catholic students [@RN4982, p.
-84].
-Therefore, @RN4982 recommends that colleges recruit from the same schools and communities that are popular with peer colleges.
-The operating logics of correlation and homophily are highly consistent with college recruiting behavior observed in the wild.
+The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the enrollment management industry.
+The logic of homophily argues that there is a natural cultural match between students, that colleges should find students who look like students who enrolled at their institution in the past, and that colleges should recruit from the same schools and communities that are popular with peer colleges.
+We suggest that homophily is the operating logic most consistent with college recruiting behavior observed in the wild.
 For example, @RN5023 find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
@@
 The Market Segment Model begins with the status attainment model, that student demand for college is a function of parental education and parental income.
-@RN4982 demonstrated this argument by showing the correlation between student SAT score-sending behavior and measures of class.
+@RN4982[chapter 3] demonstrated this argument by showing the correlation between student SAT score-sending behavior and measures of class.
 The thesis of the Market Segment Model is that that households of particular class and college aspirations are likely to live in particular parts of a metropolitan area and that these geographic territories can be meaningfully captured by Geomarkets.
-Although they appear coarse by contemporary standards, Geomarkets were a landmark innovation in Geographic segmentation.
+Although Geomarkets appear coarse by contemporary standards, they were a landmark innovation in Geographic segmentation.
 The Market Segment Model recommends that colleges should identify their core student market segment -- local, in-state, regional, or national -- identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
-This information is contained in the Market Segment Profile (Appendix A), which is a standard output of EPS software.
+This information is contained in the Market Segment Profile (Appendix A), a standard ouput in EPS software.
 When considering new Geomarkets to target, colleges should utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
-This information is contained in the Institutional Profile (Appendix A).
-Technologies that target target customers based on geography leverage racial and class segregation [@RN4775; @RN4849; @RN4975].
-The Market Segment Model and Geomarkets were created based on 1980 SAT score-sending patterns, contributing to a "ratchet effect" [@RN4851].
-The structural barriers (e.g., segregation, slavery, Jim Crow) that caused inequality in 1980 student demand are amplified because the Market Segment Model recommends that colleges focus recruiting resources on localities that already have high student demand.
+This information is contained in the Institutional Profile (Appendix A), also included in EPS software.
 
@@
 
-These analyses contribute to scholarship about how students are sorted into colleges [IMPROVE TRANSITION].
+This article adds a new perspective to existing explanations about how students are sorted into colleges.
 The status attainment model argues that college destination is a function of parental education and occupation [@RN1273; @RN1271; @RN1272].
-Indeed, the Market Segment Model can be seen as a market research application of the status attainment model.
-@RN5033 finds that Asian American students -- particularly children of Chinese, Indian, Korean, and Vietnamese immigrants -- tend to have high levels of educational achievement even if their parents do not, while the educational achievement of later-generation Asian Americans conforms to status attainment theory.
+The Market Segment Model is the status attainment model applied to geodemographic market research [CITE].
+@RN5033 finds that the educational achievement of Asian American students whose parents are immigrants tend to have high educational achievement even if their parents do not, while the educational achievement of later-generation Asian Americans conforms to status attainment theory.
 In our analyses of prospects included in student list purchases, Asian American prospects from poor Geomarkets tended to be first-generation students while Asian American prospects from affluent Geomarkets tended to have parents with a BA.
@@
 The cultural capital model explains a process by which upper and upper-middle class families sort themselves into selective colleges by providing their children with the pedigree (academic, extracurricular) valued by selective colleges, information about how to navigate the admissions gauntlet, and social networks that provide an inside track [@RN1228; @RN1241; @RN4991; @RN1814].
-The cultural capital model explains how affluent households maintain a disproportionate enrollment share at selective institutions in an era of holistic admissions that has the purported mission of increasing racial and class diversity.
-<!--Amidst concern about discrimination against Asian students at Ivy Plus institutions, @RN4991 shows how private admissions consultants coach Asian American students to......NEED TO REVISE THIS .-->
-
-Both the status attainment model and the cultural capital model are demand-side explanations for why education is a "social sieve" [@RN2377; @RN5046] that allows for a modicum of social mobility while maintaining a much larger flow of intergenerational class transmission [@RN861].
-The relative sizes of these flows can be observed in @RN4722, which obtained parental income from federal income tax returns for every college in the U.S.
-The disproportionate enrollment share of high income families at selective private colleges -- and even at most public research universities -- is staggering.
+This model explains how affluent households maintain a disproportionate enrollment share at selective institutions in an era of holistic admissions ostensibly designed to increase racial and class diversity [@RN4991].
+Both the status attainment model and the cultural capital model are demand-side explanations for why education is a "social sieve" [@RN2377; @RN5046] that allows for a modicum of mobility while maintaining a much larger flow of intergenerational class transmission [@RN861].
+@RN4722 -- who obtained parental income from federal income tax returns for every U.S.
+college -- show that the disproportionate enrollment share of high income families at selective private colleges -- and even at most public research universities -- is staggering.
 These patterns are consistent with the observation by @RN1297[p.
@@
 
-Complementing the cultural capital explanation of how students are sorted into colleges are the supply-side explanations of the credentialism literature and the more recent literature on enrollment management.
-The credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required for socially and economically advantageous positions [@RN865; @RN861; @RN880; @RN869].
-The enrollment management literature describes which prospects colleges want to enroll and what colleges do to get these students, at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758; @RN3544; @RN3547; @cottom2017lower].
+Supply-side explanations of credentialism and more recent scholarship on enrollment management complement the cultural capital model of sorting students to colleges.
+The credentialism literature recognizes that colleges have a financial incentive for educational credentials to determine the competition for socially and economically advantageous positions [@RN865; @RN861; @RN880; @RN869] and this credentials arms race favors affluent households.
+Scholarship on enrollment management describes which students colleges want to enroll and what colleges do to attract these students at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758; @RN3544; @RN3547; @cottom2017lower].
 @RN3519 and @RN4321 describe a tacit arrangement between high school guidance counselors on the demand side and college admissions officers on the supply side.
-High school guidance counselors at well-resourced schools, especially private schools, are motivated to give their students a competitive advantage in admissions.
+Counselors at well-resourced schools, especially private schools, are motivated to give their students a competitive advantage in admissions.
 Meanwhile, college admissions officers are motivated to enroll students who can afford full tuition price and are likely to donate in the future.
 These mutually beneficial desires are consummated by recognzing that upper and upper-middle class applicants satisfy the extracurricular needs of the college.
-The band needs oboists.
+The orchestra needs oboists.
 The lacrosse team needs players.
 Quantitative analyses of visits from colleges to high schools are consistent with @RN3519; selective private colleges and public research universities devote most of their recruiting resources on courting students from privileged schools and communities [@RN5023; @RN4758].
-However, the extant enrollment management literature is flawed in that it views recruiting as something that is done by individual colleges.
+However, these analyses conceive of recruiting as something that is done by individual colleges.
 
-We argue that existing explanations how students are sorted into colleges miss an important supply-side mechanism, third-party vendors that sort students on behalf of colleges.
-Based on a snapshot of student demand from 1980, @RN4982 created the Market Segment Model and Geomarkets.
-@RN4982 argues that demand for higher education is correlated with class, ingoring the historical structural barriers that produced class- and race-based inequality in student demand.
-In itself, this explanation is basically the same as the status attainment model.
-However, the College Board goes further by commodifying the Market Segment Model into products that tell colleges which Geomarkets to target.
-Therefore, third-party products encourage supply-side enrollment management behaviors to amplify historical structural inequalities observed on the demand-side.
-Geomarkets were incorporated into the Student Search Service product.
+Existing explanations of how students are sorted into colleges miss an important supply-side mechanism, third-party vendors that sort students on behalf of colleges.
+@RN4982 created the Market Segment Model and Geomarkets based on a snapshot of student demand from 1980.
+@RN4982 argues that demand for higher education is correlated with class, ignoring the historical structural barriers that produced class- and race-based inequality in 1980 student demand.
+In itself, the Market Segment Model is an unremarkable social science depiction.
+However, the College Board incscribed Geomarkets and the Market Segment Model into EPS software that told colleges which Geomarkets and high schools to target.
+Technologies that target target customers based on geography inevitably leverage racial and class segregation [@RN4775; @RN4849; @RN4975].
+Furthermore, technologies that use the logic of predictive analytics -- identify correlates in past cases to make predictions for future cases -- are prone to "pernicious feedback loops" [@RN4849].
+The structural barriers (e.g., segregation, slavery, Jim Crow) that caused historic place-based inequality in student demand are amplified because EPS recommends that colleges focus recruiting resources on localities that already have high student demand for peer colleges.
+Geomarkets were subsequently incorporated into the Student Search Service product.
 We show that that excluding low-income, non-white Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-white students with strong test scores.
@@
 
-The growing salience of third-party vendors warrants a reconfiguration of the organizational field that is salient to college access.
+The growing salience of third-party vendors warrants a reconfiguration of the organizational field salient to college access.
 @RN527[p.
-148] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
-@RN3547 shows how local external stakeholders intervene on admissions standards.
-@RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
-College access scholarship from economics analyzes the effects of policies (district/state/federal) and the effects of information interventions [CITE HOXBY TYPE].
-None of these explanations seriously consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.
+148] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.College access scholarship from economics often analyzes the effects of district/state/federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
+These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.^[ @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.]
 
@@
 First, EM consulting firms are central to the creation and implementation of recruiting campaigns [@RN5027].
-When we issued records requests to public universities about their student list purchases, at least 50\% of universities indicated that they outsourced student list purchases to an EM consultancy [].
-In many cases, the university was buying names each year but no university employee had knowledge about which student list vendors they were buying names from.
-This reliance on EM consultancies is partially explained by the high level of burnout and employee turnover in the EM profession [@hoover_em].
+The top two EM consultancies are Ruffalo Noel Levitz, which claims to serve 1,900 colleges annually [@rnlabout_2025], and EAB, which claims to serve 2,100 colleges [@about_eab].
+College reliance on EM consultancies is partially explained by the high level of burnout and employee turnover in the EM profession [@hoover_em].
+When we issued records requests to public universities about their student list purchases, at least 50\% of universities indicated that they outsourced student list purchases to an EM consultancy.
+In many cases, the university was buying names each year but no university employee had knowledge about which vendors they were buying from or which search filters they were utilizing.
 
-Second, technological advances created means of identifying and serving prospects.
+Second, technological advances creates new means of identifying and serving prospects.
 College Board began selling the contact information of test-takers in 1972 [@belkin2019-studata].
@@
 "Free" college search engines (e.g., Cappex, Niche) yielded news sources of student list data.
-Another source of student list data is college planning software purchased by high school districts and utilized by high school students and guidance counselors.
+Another data source is college planning software purchased by high school districts and utilized by high school students and guidance counselors.
 The most widely-known product is Naviance, which claims to be used by more than used by more than 10 million K-12 students and by 40% ofUS high schools [@powerschool].
 *Naviance* college planning software feeds into the *Intersect* recruiting platform, which allows colleges to target Naviance users while they are on the platform.
-*Intersect* is an example of behavioral-based -- as opposed to list-based -- targeting.
+*Intersect* is an example of behavioral-based targeting.
 
 The third dynamic in the EM industry has been the growth of test-optional movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT.
-ACT responded to uncertainty in their core testing business by attempting to become an edtech company, failing miserably, and being acquired by Nexus Capital.
-[ADD A COUPLE SENTENCES ABOUT COLLEGE BOARD; MAYBE GEODEMOGRAPHIC SEGMENT]
+ACT responded to uncertainty in their core testing business by attempting to become an edtech company [CITE], failing [], and being acquired by Nexus Capital in 2024 [CITE].
+Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters [CITE] and ACT's predicted probability of enrollment filter [CITE] -- developed free college search engines, and entered the enrollment management consulting market more aggessively.
+College Board has weathered the test-optional movement thanks to robust revenues from the Advanced Placement product, but the decline in PSAT/SAT/PreACT/ACT test-takers has undermined their oligopoly in the market for student list data.
 
-The fourth dynamic is the transformation from owner-operated to private-equity owned firms and the blurring between consultancy and software vendors.
-The 1980s and 1990s was a period of market entry in EM, mostly consistent of college admissions professionals or professors who decided to hang up a shingle [@RN5027].
+The fourth dynamic is the transformation from owner-operated to private-equity owned firms and erosion of the distinction between consultancy and software vendor.
+The 1980s and 1990s was a period of market entry in EM, when college admissions professionals or professors decided to hang up a shingle [@RN5027].
 The 2010s were a period of acquisitions and concentration.
-After regulatory scrutiny pushed private-equity interests out of for-profit college market [@RN5022], private equity investors saw value in acquiring firms that provided services to direct providers.
-Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions grow market share and vertical acquisitions designed to increase customer reliance.
+After regulatory scrutiny pushed private-equity interests out of for-profit college market [@RN5022], private equity investors find value in acquiring firms that provided services to direct providers.
+Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions to grow market share and vertical acquisitions designed to make customers more dependent on particular recruiting products.
 
 EAB exemplifies the the trend towards private equity ownership and industry concentration.
-EAB is the preeminent enrollment management (EM) consulting firm, serving more than 2,100 colleges and offering software-as-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
-The origins of EAB are traced to Bill Royall, who founded Royall & Company in 1983 to provide direct marketing and fundraising for Republican political campaigns [@jump_2020].
+EAB is the preeminent EM consultancy, offering software-as-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
+The origins of EAB trace to Bill Royall, who founded Royall & Company in 1983 to provide list-based marketing and fundraising for Republican political campaigns [@jump_2020].
 Royall did not sign its first colege client for several years, but EM became the primary focus by 1995.

### 2025-05-02 text/discussion.qmd
first run through of the discussion section finished

@@
 EAB is the preeminent EM consultancy, offering software-as-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
-The origins of EAB trace to Bill Royall, who founded Royall & Company in 1983 to provide list-based marketing and fundraising for Republican political campaigns [@jump_2020].
+The origins of EAB trace to Bill Royall, who founded Royall & Company in 1983 to provide direct marketing and fundraising for Republican political campaigns [@jump_2020].
 Royall did not sign its first colege client for several years, but EM became the primary focus by 1995.
@@
 
-As an owner-operated EM consultancy, Royall was the market leader in purchasing student lists on behalf of clients to help realize enrollment goals.
-Under Vista, EAB tranformed the market for student list data by completing a series of acquisitions that targeted the "leads" and "inquiries" stages of the enrollment funnel.
-In 2019, EAB acquired YouVisit, a leading provider of virtual tours, which can be conceptualized as an engine for "inquiries" in that taking a virtual tour shows interest in a college and leaves a trail of contact information for subsequent recruiting interventions.
-In 2020, EAB acquired Cappex, a leading provider of free college search engine, which is an engine for leads.
-In 2021, EAB and PowerSchool -- the leading provider of K12 student information systems and a subsidiary of Vista Eqity Partners -- jointly acquired the EM consultancy/software vendor Hobsons.
+As an owner-operated consultancy, Royall was the market leader in purchasing student lists from student list vendors on behalf of clients.
+Under Vista, EAB entered the market for student list data by completing a series of acquisitions that targeted the "leads" and "inquiries" stages of the enrollment funnel.
+In 2019, EAB acquired YouVisit, a leading provider of virtual tours.
+YouVisit is an engine for "inquiries" in that taking a virtual tour shows interest in a college and leaves a trail of contact information for subsequent recruiting interventions.
+In 2020, EAB acquired Cappex, a leading free college search engine, which is an engine for leads in that students provide contact details and other information.
+In 2021, EAB and PowerSchool -- the leading provider of K12 student information systems and a subsidiary of Vista -- jointly acquired the EM consultancy/software vendor Hobsons.
 EAB acquired Hobson's student success software, Starfish, while PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect.
-At the same time, EAB paid PowerSchool to become the exlusive reseller of Intersect.
+EAB paid PowerSchool to become the exlusive reseller of Intersect.
 Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.
-EAB combined all of these acquisitions into Enroll360:
+EAB combined these acquisitions into create [Enroll360](https://eab.com/solutions/enroll360/):
 
-> We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students as they pursue these increasingly digital journeys to college.
+> We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
 This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]
@@
 Historically, College Board and ACT sold prospect contact information at a price-per-prospect (e.g., \$0.50).
-Thus, the goal of lists was to generate revenue from the sale of names.
+The goal was to generate revenue from the sale of names.
 By contrast, Enroll360 wraps several proprietary databases of names (e.g., Cappex, Intersect) into a software-as-service platform that recruits those prospects at different stages of the enrollment funnel.
@@
 Colleges that are interested in recruiting these prospects must buy an Enroll360 subscription.
-Like College Board *Student Seardh Service* contemporary software-as-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
-For example, Investigative reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
+Like College Board *Student Search Service* contemporary software-as-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
+For example, reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
 Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and wheather students used Naviance to "research competitor institutions" [@markup_naviance].
-A document from the University of Utah to procure an Intersect subscription explains that the Intersect subscription is necessary to recruit the proprietary database of Naviance users: [@utah_procurement]:
+A procurement document from the University of Utah explains that the Intersect subscription is necessary to recruit the proprietary database of Naviance users: [@utah_procurement]:
 
@@
 
-We argue that an under-studied and under-regulated third-party EM industry has negative consequences for colleges and for students.
+An under-studied and under-regulated third-party EM industry has negative consequences for colleges and for students.
 Accross multiple student list products, we see a growing number of filters that recommend which prospects to target based on the beheavior of previous cohorts of students.
-The ACT/Encoura "Enrollment Predictor" filter enables colleges to filter prospects based on the predicted probability of enrollment, which is created based on analyzing the enrollment decisions of previous cohorts of students [schmidt_2019].
-Based on the Claritas/PRIZM market segmentation system that categorized zip codes into groups useful for merchants [@RN4957], The College Board Geodemographic Segment filters categorizes each Census Tract into one of about two dozen clusters based on the college-going behavior of previous cohorts [@RN4565].
-A @vista_eab_ml video promoting EAB's "Pipeline Analytics" product to investors states, "what we are doing is using artificial intelligence and machine learning to help schools build list of students may be more more likely to enroll and persist in their university." The product "identifies students that would be a good match for a school based on their similarity to students who have succeeded there before and then provides them [the students] information about those institutions so that they choose to apply and ultimately enroll." Each of these products are based on the underlying logic of correlations and homophily discussed by @RN4975 and applied by @RN4982.
+The ACT/Encoura "Enrollment Predictor" filter enables colleges to filter prospects based on the predicted probability of enrollment, which is created based on analyzing the enrollment decisions of previous cohorts of students [@schmidt_2019].
+College Board Geodemographic Segment filters categorizes each Census Tract into one of about two dozen clusters based on the college-going behavior of previous cohorts [@RN4565] and is based on the Claritas/PRIZM market segmentation system that categorized zip codes into groups useful for merchants [@RN4957].
+An investor-facing @vista_eab_ml video promoting EAB's "Pipeline Analytics" product states, "we are using artificial intelligence and machine learning to help schools build list of students that may be more more likely to enroll and persist in their university." The product "identifies students that would be a good match for a school based on their similarity to students who have succeeded there before and then provides them [the students] information about those institutions so that they choose to apply and ultimately enroll." Each of these products are based on the underlying logic of correlations and homophily discussed by @RN4975 and applied by @RN4982.
 These products identify correlations in a historic snapshot of students and apply those relationships to make recommendations about which prospects to target in the present.
@@
 
-Market dynamics in the EM industry may also affect competition, prices, and choice.
-To the extent that colleges need to buy expensive software to recruit prospects contained in an associated database, those costs will ultimately be passed on to students in the form of higher prices are reduced expenditure on education.
+Market dynamics in the EM industry raise policy concerns about tuition prices, student choice, and competition in the market for EM consultancies
+may also affect competition, prices, and student choice.
+To the extent that colleges must buy expensive software to recruit prospects contained in an associated database, this expenditure will ultimately be passed on to students in the form of higher prices or reduced expenditure on education.
 Unfettered horizontal and vertical acquisitions in the EM industry causes the EM industry to become more concentrated.
-Smaller consultancies that don't have vertical inputs -- e.g., control over proprietary pools of prospects -- cannot compete.
+Smaller consultancies that do not control veritial inputs may be unable to compete with companies that have acquired proprietary pools of prospects.
 Fewer EM consultancies means less competition and higher consulting fees, which will be passed on to consumers.
@@
 148] in the organizational field.
-The paucity of research about third-party vendos that structure college access is substantially due to difficulty obtaining data for empirical analyses.
+Unfortunatel, scholarship salient to education policy and regulation have ignored two, related economic transformations.
+First, third-party vendors offer software platforms that perform core tasks of organizations, including schools and colleges [@list_biz].
+Second, private firms comprise a growing share of economic activity [@RN5037; @RN5038].
+Most firms in EM and the broader edtech sector are private.
+However, private firms face fewer disclosure requirements than publicly traded firms.
+The paucity of research about third-party vendors that structure college access is substantially due to difficulty obtaining data for empirical analyses.
+The research community must figure out how to systematically investigate the role of private firms and investors in education, so that we can hold private interests accountable for education policy goals.
 
@@
 Student unit record data owned by district, state, and federal agencies is simultaneously granular -- can evaluate a particular policy -- and macro -- can speak to broad trends in the education sector.
-By contrast, obtaining granular data about third-party products in education is difficult because these data are owned by private interests that do not want researchers to interrogate their products.
+By contrast, obtaining granular data about third-party products in education is difficult because these data are owned by private interests that do not want researchers to interrogate their products [@RN4774; @RN4842].
+We see two broad avenues for research: case studies involving primary data collection -- including this study -- which tend to have a granular focus on particular products or vendors; and secondary data analysis, which may have a more macro focus.
 
-We see two broad avenues for research: case studies involving primary data collection -- including this study -- which tend to have a granular focus on particular products or vendors; and secondary data analysis, which may have a more macro focus.
-Case study research traditionally relies on interviews, ethnographies, and document analysis.
-Another approach to data collection is to issue records requests to public entities that contract with third-party entities.
+Although case study research often relies on interview and ethnographic data, another approach to data collection is to issue records requests to public entities that contract with third-party entities.
 For example, @RN5039 analyze public university contracts with online program managers (OPMs).
+The present manuscript is an example of case study research.
 We issued public records requests to obtain tabular data about student list purchases.
@@
 Public records requests seem better suited to obtaining contracts.
-Text analysis of publicly available websites and social media provides another avenue for data collection.
+Qualitative or computational text analysis of publicly available websites and social media provides another avenue for data collection.
 Here, data collection is efficient but data processing is often labor intensive.
-Data collected for case study research typically does not meet the knowledge infrastructure standard of repeated, ongoing data collection that can be the foundation for scalable research.
-Nevertheless, case study research is important for granular analyses of particular products -- like Intersect, Enroll360,  that yield important insights about underlying mechanisms about how vendors structure college access.
+Like investigative reporting [@markup_naviance], case study research can produce granular analyses of particular products (e.g., Intersect, Enroll360) that yield insight into the underlying mechanisms of how third-party vendors structure college access.
+However, data collected for case study research typically does not meet the knowledge infrastructure standard of repeated, ongoing data collection that can be the foundation for scalable research.
 
+We see subscription databases about private markets as having the potential to create a knowledge infrastructure for scalable research on the role of private firms and private equity funders in education, and also their role in other domains of interest to sociology.
 A growing set of "market intelligence" data providers serve the information needs of the investment community by providing data on private firms and deals involving private firms.
-These databases may provide the foundation for scalable research on the role of private firms and private equity investors in education.
-These databases tend to include three "types" of data: (1) *firm-level data* (e.g., location, employees, financials); (2) *acquisitions*, encompassing who acquired whom and resulting ownership structure); and (3) *investments*, which include amounts/valuations/stake percentages of private equity investments.
+These databases tend to include three broad "types" of data: (1) *firm-level data* (e.g., location, employees, financials); (2) *acquisitions*, encompassing who acquired whom and resulting ownership structure); and (3) *investments*, which include amounts/valuations/stake percentages of private equity investments.
 Some providers provide all three types of data (e.g., [*S&P Capital IQ*](https://www.spglobal.com/market-intelligence/en), [*LSEG Workspace*](https://www.lseg.com/en)).
 Other providers specialize.
-For example, [*PitchBook*](https://pitchbook.com/) is a leading provider of data about investments in private companies by private equity firms, while [*Orbis*](https://login.bvdinfo.com/R1/Orbis) is the leading provider of firm-level data for non-U.S.
-private firms [@RN5038].
+For example, [*PitchBook*](https://pitchbook.com/) is a leading provider of data about investments in private companies by private equity firms, while [*Orbis*](https://login.bvdinfo.com/R1/Orbis) is a leading provider of firm-level data for non-U.S.
+private firms.
+These databases are often utilized in journals of finance or management that consider private equity markets or patterns of acquisitions [e.g., @RN5044; @RN5038].
 Interestingly, the database subscriptions operate similar to contemporary student list products; an interactive, user-facing platform is wrapped around several sources of proprietary data.
 University libraries provides full or partial access to some subscriptions, while others must be purchased.
-These databases are often utilized for articles in management journals that focus on private equity markets [CITE], or patterns of acquisitions [CITE].
-Considering the growing share of economic activity that  [HOW THESE DATABASES HAVE BEEN USED; HOW THEY MAY BE USED]
+
+Although scholars cannot share subscription data, they can share code to process and analyze these data.
+We recommend that scholars develop panel datasets examining change over time in the firms operating in particular industries and acquisitions and private equity investments in those industries.
+For example, empirical scholarship can examine change over time in concentration in the enrolment management industry and its transformation from an owner-operated to a largely private-equity owned industry.
+Subsequent analyses could investigate the portfolio of large owners, their ties to other industries -- such as the direct-provider for-profit college market -- and the vertical inputs/technologies they are investing in.
 

### 2025-05-07 text/manuscript-full-KS.qmd
full version copy

--- 2025-05-01 text/back_end_KS.qmd
@@
 ```
-# Data and Methods
-
-## Data and Variables
-
-**RQ1**.
-Our study utilizes two main data sources based on each research question.
-The first research question is focused on understanding the socioeconomic and racial variation between Geomarkets in metropolitan areas and how such variation changes over time.
-To answer this question, we use census tract data from the U.S.
-Census Bureau.
-Utilizing a small geographic area is important for implementing a spatial merge between Census data and Geomarket shape files (described below).
-Census tract is the smallest geographic area for which measures of both race/ethnicity and socioeconomic characteristics are available.
-Census tracts are statistical subdivisions of a county designed to capture relatively homogeneous demographic communities that are generally considered to be more accurate for population analyses than zip codes, which are spatial boundaries drawn to demarcate route efficiency for postal services \[[CITE](https://www.census.gov/newsroom/blogs/random-samplings/2014/07/understanding-geographic-relationships-counties-places-tracts-and-more.html)\] .
-
-For the first research question, we use data from the 1980 and 2000 Decennial Census as well as 5-year estimates from the 2020 American Community Survey (ACS).
-For the 1980 Decennial Census, we specifically draw on Summary Tape File 1 (STF1) for variables about race and Hispanic origin collected via “short form” questions answered by all households and Summary Tape File 3 (STF3) for variables about socioeconomic characteristics (income, education, and poverty) collected via the "long form" questionnaire completed by a sample of households.
-The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-To examine how Geomarket characteristics change over time, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty from the 2000 Decennial Census.[^1] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
-These questions were replaced by ACS.
-Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.[^2] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
-
-[^1]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
-
-[^2]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
-
-Across these data sources, we created measures of race and ethnicity, specifically the number and proportion of the census-tract population in the following categories: White, non-Hispanic; Black, non-Hispanic; Asian, non-Hispanic; Pacific Islander, non-Hispanic; American Indian and Alaskan Native, non-Hispanic; and two or more races, non-Hispanic.
-Measures of Asian, American Indian and Alaskan Native, two or more races, and Native Hawaiian and Pacific Islander (all non-Hispanic) were not available for the 1980 Decennial Census.
-We also created socioeconomic measures that approximate the predictors of market segment identified by @RN4982.
-These variables are median household income, percent of households below the poverty line, and people age 25+ with a BA.
-All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
-
-We assigned each census tract to College Board Geomarkets by implementing a partial spatial merge.
-Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^3] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
-Using these zip code-level data, we utilized the `aggregate()` function to create shapefiles for each Geomarket.
-
-[^3]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
-
-Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
-Census tract shapefiles that intersected with multiple Geomarkets (i.e., a census tract's spatial boundaries were spread across two or more Geomarkets) broken into smaller shapefiles wholly contained within a single Geomarket.
-The variables in these shapefiles were then replaced with the original variable value multiplied by the proportion of land area of the census tract that was contained within the Geomarket.
-For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 1,000 people who identify as Hispanic non-white.
-The partial spatial merge splits this census tract into two observations.
-The observation assigned to Geomarket A reports $.60*1000 = 600$ Hispanic non-white people and the observation assigned to Geomarket B reports $.40*1000 = 400$ Hispanic non-white people.
-
-**RQ2**.
-Our study's second research question asks how does the racial and socioeconomic composition of included versus included prospects vary when student list purchases filter on Geomarkets.
-We answer this question using actual student lists purchased by public universities.
-In February 2020, we began issuing public records requests to public universities about student lists purchased from 2016 through 2020.
-We narrowed the scope of our request to student lists purchased from College Board, ACT, and the National Research Center for College and University Admissions (NRCCUA), the three largest student list vendors at the time.
-For each student list purchased from 2016 through 2020, we requested two related pieces of data: (1) the order summary, which specifies search criteria for the student list purchase; and (2) the de-identified prospect-level list produced from the search criteria.
-
-The data collection sample for this project was all public universities in California, Illinois, Minnesota, and Texas.
-Utilizing public records requests to obtain public records is a painstaking process.
-Initially, the majority of universities did not respond or denied our request.
-Subsequently, we obtained pro bono representation from four law firms, which substantially increased the success of data collection.
-We also collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
-However, we were unable to obtain legal representation for Texas.
-
-Even with firm representation, data collection remained difficult.
-Some universities provided records that were not usable for quantitative analyses (e.g., summary statistics across multiple orders; or data did not contain important fields).
-Some universities did not provide records based on legitimate legal grounds (e.g., data not in university possession; not required to create records that do not currently exist).
-We learned that many universities outsourced student list purchases to third-party consulting firms.
-Unfortunately, we were rarely able to obtain usable data from these universities.
-A small number of universities denied requests based on potentially questionable legal rationale, but we lacked the resources to litigate.
-
-The primary data source for our second research question is therefore student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
-This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Minnesota, and Arizona \[KS Does this still apply?\].
-Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^4] The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
-In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.
-Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
-
-[^4]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
-
-In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter, but did purchase prospects from all Geomarkets in a particular metropolitan area.
-For each prospect we know their home zip code and, therefore, which Geomarket they belong to.
-We therefore simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
-
-The variables of interest for student list data are the characteristics of prospects who would have been included versus excluded by our simulations by Geomarket filter.
-Prospect characteristics are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
-College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity nonresponse) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016).
-From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
-Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016).[^5] Information about student socioeconomic status is limited.
-However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
-
-[^5]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
-Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”
-
-## Methods
-
-**Research design**.
-We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
-The research questions of this study aim to explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
-Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarkets landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics.
-Such aims align with the purpose of a multiple case study design to gain a deeper understanding of the phenomenon of interest by comparing and contrasting across different settings \[CITE YIN 2001\].
-In doing so, multiple cases allow researchers to explore the "how" and "why" of the phenomenon in rich and nuanced contexts.
-We therefore focus on minimizing the number of metropolitan areas in order to provide sufficient analytic depth while also situating each case within its specific historic and contemporary geographical landscape \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
-Online Appendix XXX presents results for a larger number of metropolitan areas.
-
-Case selection was informed by based on purposeful sampling to identify “information-rich” cases (Patton, 2002; Yin, 2009) that can provide in-depth knowledge about the phenomenon of interest.
-We first considered metropolitan areas as information rich cases across two broad dimensions: 1) Geographical and urban model structures and 2) variation across racial/ethnic and socioeconomic characteristics.
-Based on our read of @RN4982, we are interested in "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
-11-12).
-Building on our conceptual framework, we are interested in racial and socioeconomic inequality between geomarkets within a metropolitan area (RQ1) and how this inequality contributes to inequality in which prospective students are targeted by student list purchases (RQ2).
-\[DISCUSSION OF WHICH CASES TO SELECT.
-KARINA\].
-Choice of metropolitan areas is informed by data availability.
-Although Census data are available for all metropolitan areas (RQ1) we do not have good candidate student list purchases for all metropolitan areas (RQ2).
-
-Then, our purposeful sampling approach also took into consideration student list orders!
-\[SAY SOMETHING ABOUT WHAT IS A GOOD CANDIDATE STUDENT LIST PURCHASE -- PURCHASE WITH NOT TOO MANY CRITERIA THAT INCLUDE ALL GEOMARKETS IN A METRO\].
-
-**Analyses**.
-Analyses are simultaneously descriptive and spatial.
-We answer RQ1 by producing Geomarket-level tables and graphs that show how the Geomarkets in a selected metropolitan area vary on racial and socioeconomic characteristics and how they vary over time.
-For example, in the San Francisco Bay Area, how does the City of Oakland Geomarket (CA07) differ from the Alameda County Exluding Oakland Geomarket (CA08).
-We also produce interactive maps at the census tract-level to show more granular variation within and between Geomarkets.
-
-We answer RQ2 by analyzing student list purchases that encompassed all Geomarkets in a selected metropolitan areas.
-Descriptive tables and graphs describe the racial and socioeconomic characteristics of prospects that would have been included/excluded had the purchase filtered on particular Geomarkets.
-Interactive maps provide a more granual visualization of where prospects lived and went to high school.
-For example, for purchased profiles who lived in the Chester County Geomarket (PA02) \[OR PHILADELPHIA PA05\] and identified as Black, where in the county did these students live and which high schools did they attend?
-
-NEXT STEPS: CREATE TABLE FOR PHILLY STUDENT LIST PURCHASES THAT FOCUSES ON FIRST GEN STATUS.
-
-Our conceptual framework suggests that racial and socioeconomic inequality is Geomarket regions is associated with the number of geomarkets in the metropolitan area.
-and is associated with the patterns of racial segregation in the metropolitan area.
-comment more on this.
-
-Below, we use simulations and actual student list purchases that filtered on geomarkets.
-In purchases that filtered on test-score and/or GPA thresholds, we can simluate who would be included and excluded had certain Geomarkets were been selected.
-In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.
-
-# Findings
-
-## Chicago
-
-### Shifting Demographics, Persistent Segregation in Geomarket Landscape
-
-@fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
-The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 in the development of The Market Segment Model.
-Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
-In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
-The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most densely populated Geomarket.
-
-These population shifts have contributed to changes in the racial/ethnic composition within and across Geomarkets over time.
-@fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarkets across the three census periods.
-In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
-In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
-By 2020, several Geomarkets that were previously predominantly White shifted to a more balanced racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
-For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
-Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
-
-@fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
-For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
-City of Chicago also experienced relatively modest declines in its Black population (37% in 1980 to 33% in 2020), alongside increases in its Hispanic (15% in 1980 to 26% in 2020) and Asian Pacific Islander populations (4% in 2000 to 6% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
-
-We also explore socioeconomic demographics - median income, percent living below poverty, and educational attainment - within and across Geomarkets over time in @fig-chicago-rq1-ses.
-Overall, median income both within and across Geomarkets in the Chicago metropolitan area remained relatively stable across the three census periods.
-For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$158,000 in the North Shore Geomarket to a low of \$66,000 in the City of Chicago Geomarket.
-However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence, such as \$120,000 in the Northwest Suburbs, \$119,000 in Evanston and Skokie, \$108,000 in Chain of Lakes, \$105,000 in the Western Suburbs, and \$99,000 in the South and Southwest Suburbs.
-While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$79,000) and decrease in the South and Southwest Suburbs Geomarket (\$93,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$14,000.
-
-@fig-chicago-rq1-ses also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
-Similar patterns are evident in educational attainment across Geomarkets.
-In 1980, the City of Chicago had the lowest percentage of residents with a bachelor's degree or higher (13%).
-However, by 2020, the Geomarket experienced the largest percentage point increase in college degree attainment, reaching 38%.
-In comparison, all other Geomarkets experienced smaller gains in the share of residents with a bachelor’s degree or higher over the same period.
-These trends contributed to a narrowing of the educational attainment gap across Geomarkets from 1980 to 2020.
-
-From 1980 to 2020, these demographic shifts have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by racial/ethnic and socioeconomic characteristics.
-The original Geomarket boundaries delineated significant racial and income segregation in the 1980s.
-This included Black, Hispanic, and Asian residents primarily in the City of Chicago Geomarket and White residents in Geomarkets forming the suburban ring of the metropolitan area.
-Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs.
-However, declining White populations, Hispanic and Asian population growth, and Black outmigration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining.
-The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents) while the the Evanston and North Shore Geomarkets remain predominantly White.
-While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains across Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
-
-### Composition of Students Lists by Geomarket Contributions
-
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospects across the entire state of Illinois.
-Two of these orders applied the same filters: SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
-The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with relatively middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
-Because all six orders included an Illinois state filter, we can utilize prospects from the Chicago metropolitan area within the resulting student lists to examine Geomarket contributions by simulating the application of a Geomarket filter across low-, middle-, and high-range SAT threshold orders.
-
-@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area.
-At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets generally reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
-For example, more than 8,000 of the 16,485 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
-Similar to the larger metropolitan area, the Northshore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students.
-Interestingly, while the City of Chicago is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-chicago-rq2-race includes nearly double the proportion of Hispanic (50%) prospects relative to the larger Geomarket resident population (26% Hispanic) in @fig-chicago-rq1-race with a lower proportion of White prospects (14%).
-
-However, the racial/ethnic composition of prospect pools become less similar to the overall metropolitan area as PSAT scores increase.
-Notably, the shares of White and Asian prospects within Geomarkets increase, while Black and Hispanic prospects decrease.
-For instance, Asian prospects over representation in Geomarkets relative to the overall resident population at the 1310-1600 SAT score threshold ranges from a minimum of 4 percentage points in the North Shore Geomarket (17% of prospects versus 13% of residents) to a maximum of 21 percentage points in the Chain of Lakes (28% of prospects versus 7% residents) Geomarket.
-The City of Chicago Geomarket exhibits the largest underrepresentation of Black prospects (7% of prospects versus 33% of residents) while Chain of Lakes exhibits the largest underrepresentation of Hispanic prospects (7% of prospects versus 24% of residents).
-
-We also explore contributions across Geomarkets to overall prospect pools by race/ethnicity.
-For instance, @fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle range SAT scores (1160-1300) by race/ethnicity.
-The top panel presents the full distribution of the 13,643 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% reside in Chain of Lakes, 13% in Northwest Suburbs, 5% in North Shore, 5% in Evanston and Skokie, 17% in City of Chicago, 28% in Western Suburbs, and 20% in South and Southwest Suburbs.
-These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-midSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
+# Introduction
+
+On January 9, 2019 a public research university purchased College Board *Student Search Service* order no.
+$448922$, thereby obtaining the contact information of out-of-state prospective students, who would subsequently receive brochures, postcards, email, and targeted social media.
+This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
+Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and has a higher share of Black residents.
+Similarly, in the DC/Maryland/Virginia metro area, the order selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets.
+
+[^1]: On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
 
 ```{=html}
-<!-- OLD RESULTS BASED ON LOW VERSUS HIGH SAT ORDERS
-Nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White.
-With the exception of the City of Chicago, @fig-chicago-rq2-lowSAT-race shows all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
-The City of Chicago Geomarket, however, contributed only 5% of White prospects, despite accounting for 20% of all included prospects.
-Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, exept for Evanston and Skokie (12% versus 4% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (10% versus 26% for all prospects) contributing smaller shares relative to their overall representation.
-
-In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race.
-Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 34%, respectively, compared to their overall contributions of 20% and 26% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
-The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
-
-@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores.
-Both figures demonstrate how representational Geomarket patterns across race/ethnicity remain relatively consistent even as SAT scores increase.
-For instance, @fig-chicago-rq2-midSAT-race shows that the Chain of Lakes, North Shore, and Western Suburbs Geomarkets contributed similar proportions of White and Asian students, yet smaller proportions of Black and Hispanic students, relative to their overall representation among all included prospects.
-The Northwest Suburbs and Evanston and Skokie similarly contributed a disproportionately higher share of Asian prospects, but lower shares of Black and Hispanic students.
-Conversely, the South and Southwest Suburbs Geomarket contributed a substantially larger proportion of Black students, while the City of Chicago yielded larger proportions of both Black and Hispanic students- again, relative to their overall representation among all included prospects targeted at middle range SAT scores.
-Despite declines in the overall number of Black (n=240) and Hispanic (n=950) prospects included in orders filtering for high-ranges of SAT scores (1310-1600), @fig-chicago-rq2-highSAT-race shows overall Geomarket contribution patterns persist.
-
-We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
-Similar to race/ethnicity figures above, @fig-chicago-rq2-lowSAT-ses presents the Geomarket distribution for the 16,060 total prospects (whose first-generations status is known) included in low-range SAT score orders (1020-1150).
-Among the 3,479 total prospects whose parents did not attend college, the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 34% in comparison to 20% for all included prospects in this SAT range.
-In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students relative to all included prospects, with such disparity ranging from a one-percentage-point difference (Chain of Lakes, Northwest Suburbs, North Shore) to a five-percentage-point difference (Western Suburbs).
-For prospects whose parents attended some college but did not complete their degrees, @fig-chicago-rq2-lowSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
-These representational patterns are reversed among prospects that are not first-generation college students.
-The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (13%) relative to all prospects (20%) in this SAT range, whereas all other Geomarkets contributed either comparable (Northwest Suburbs, South and Southwest Suburbs) or disproportionately larger shares (Chain of Lakes, North Sore, Evanston and Skokie, Western Suburbs).
-
-@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status become increasingly pronounced in orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores, respectively.
-At the middle SAT score range in @fig-chicago-rq2-midSAT-ses, the City of Chicago Geomarket contributes a substantially larger share of first-generation students whose parents did not attend college—38%, compared to 17% of all included prospects in this range, representing a 21-percentage-point disparity.
-This disproportionality increases to a 25-percentage-point difference at the highest SAT score thresholds in @fig-chicago-rq2-highSAT-ses, where the City of Chicago accounts for 40% of such first-generation students, despite representing only 15% of all included prospects in this score range.
-
-Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
-@fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150) orders.
-For instance, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% were first-generation college students whose parents attended some college but did not complete their degree, and 67% were not first-generation college students.
-@fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed degrees (29% and 28%, respectively).
-In contrast, Geomarkets such as Chain of Lakes (72%), North Shore (88%), Evanston and Skokie (73%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
-
-@fig-chicago-rq2-lowSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
-The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
-However, the magnitude of this disproportionality varied- from a 3-percentage-point difference for Black prospects to a 21 percentage-point difference for Asian prospects.
-Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
-Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.
-
-Geomarket contribution patterns to prospect pools by both first-generation college student status and race/ethnicity become more pronounced as SAT score thresholds increase.
-For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders.
-The City of Chicago Geomarket still contributes a disproportionately larger share of first-generation college students for both those whose parents did not attend college and those whose parents attended some college but did not complete a degree across most racial and ethnic groups.
-However, the magnitude of this disproportionately for City of Chicago becomes even greater for Asian students, with a nearly 30-percentage-point difference collectively for all first-generation college students (35% no college and 22% some college) relative to their overall Geomarket representation (28%).
-Similar to patterns observed at lower SAT score thresholds, the Geomarket contributions to first-generation Hispanic college students remain relatively balanced.
-All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation Hispanic college students across both parental education categories to the overall prospect pool.
-
-@fig-chicago-rq2-highSAT-combo shows the magnitude of Geomarket contribution disparities increases at the highest SAT score range (1310-1600), partly driven by the relatively fewer number of first-generation college students included within these orders.
-Again, the City of Chicago Geomarket disproportionately contributes a disproportionate share of first-generation Asian college students (43% across both parental education categories) and first-generation Hispanic college students (52% across both parental education categories) in comparison to the overall distribution of first-generation college students within each racial/ethnic category (13% for Asian and 40% for Hispanic, respectively).
-Conversely, nearly all other Geomarkets contribute a disproportionately larger share of non-first generation college students across all racial/ethnic categories.
--->
-```
-More than half (n=8,256) of all prospects whose contact information was purchased in these middle SAT score orders identified as White.
-@fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
-One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects.
-On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively.
-Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
-
-In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race.
-Among the nearly 1,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
-The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
-
-Appendix figures \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting lower-range (1020-1150) and high-range (1310-1600) SAT scores.
-Both figures demonstrate how representational Geomarket contributionsacross race/ethnicity, such as City of Chicago disproportionately contributing larger shares of Black and Hispanic prospects, remain consistent even at lower and higher SAT score thresholds.
-
-We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
-Similar to race/ethnicity figures above, @fig-chicago-rq2-midSAT-ses presents the Geomarket distribution for the 13,749 total prospects (whose first-generations status is known) included in middle-range SAT score orders (1160-1300).
-Among the 1,743 prospects whose parents did not attend college ("no college"), the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 38% in comparison to 17% for all included prospects in this SAT range.
-In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students whose parents did not attend college relative to all included prospects, with such disparity ranging from a one-percentage-point difference (South and Southwest Suburbs) to a nine-percentage-point difference (Western Suburbs).
-For prospects whose parents attended some college but did not complete their degrees ("some college"), @fig-chicago-rq2-midSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
-These representational patterns are reversed among prospects that are not first-generation college students.
-The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (12%) relative to all prospects (17%) in this SAT range, whereas all other Geomarkets contributed comparable shares.
-
-Relative to @fig-chicago-rq2-midSAT-ses, appendix figures \@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status are less pronounced in low-SAT range orders (1020-1150) and more pronounced in high-range SAT orders (1310-1600), respectively.
-
-Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
-@fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 12,981 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders.
-For instance, among all White prospects (n=8,155) whose contact information was purchased in this SAT score range, approximately 6% were first-generation college students whose parents did not attend college, 15% were first-generation college students whose parents attended some college but did not complete their degree, and 79% were not first-generation college students.
-@fig-chicago-rq2-midSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (10% and 9%, respectively).
-For first-generation college students whose parents attended some college but never completed degrees, only the South and Southwest Suburbs (25%) Geomarket contributed a substantially larger share of White prospects.
-In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
-
-@fig-chicago-rq2-midSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
-The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
-However, the magnitude of this disproportionality varied- from a 1-percentage-point difference for Black prospects to a 23 percentage-point difference for Asian prospects.
-Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
-Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs Geomarkets.
-Appendix figures @fig-chicago-rq2-lowSAT-combo and \@fig-chicago-rq2-highSAT-combo show the magnitude of these Geomarket contribution disparities become more pronounced at as SAT score thresholds increase.
-
-The enduring segregated structure of the Chicago metropolitan area from 1980 to 2020 analyzed above plays a significant role in how Geomarkets shape the racial/ethnic and socioeconomic composition of prospect pools.
-Reflecting the larger metropolitan area dynamics, central Geomarkets that are predominantly comprised of low-income communities of color such as the City of Chicago and South and Southwest Suburbs consistently contribute disproportionately larger shares of Black, Hispanic, and first-generation college prospects, especially at lower SAT score thresholds.
-On the other hand, predominantly White and affluent Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, as well as non-first generation college students.
-However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+<!--Geomarkets carve states and metroplitan areas into smaller geographic units and have names like,  On the same day, the university purchased two additional orders that were identical, except they filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$.
+On the same day, the university also purchased orders $\#448427$ ($57,537$ names) and $\#448440$ ($78,000$ names), which filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets as order $\#448922$.
+
+Geomarkets carve states and metroplitan areas into smaller geographic units.
+Upon closer inspection, the three orders selected on all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago", which is significantly poorer and less White.
+In the DC/Maryland/Virginia metro area, the orders selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets, which are poorer and have a higher share of Black residents.
+An analysis of 500 \[check\] student lists purchased by 9 public research universities found that 11% of purchases filtered on Geomarket [@RN5012].
+Recent research suggests that student lists are surprisingly important for college access outcomes and college graduation outcomes, particularly for first-generation and underrepresented minority students [@RN4739; @RN4894].
+To the extent that Geomarket borders are correlated with race and class, the Geomarket selections observed here raise concern that Geomarkets -- and other geographic search filters -- may systematically filter working class and non-white students from educational opportunities.
+-->
+```
+Sociology offers several perspectives to describe the process of sorting students into colleges as a process that tends toward social reproduction [@RN4873; @RN2377].
+On the student demand side, the status attainment model argues that student postsecondary destination is a function of family socioeconomic background, particularly parental education [e.g., @RN915; @RN1273; @RN1268; @RN5033; @RN1300].
+The cultural capital model argues that bourgeois households bestow the pedigree, social networks, and information necessary for their progeny to claim spots at selective colleges [@RN1228; @RN1241; @RN4991] [ADD CITES].
+@RN4873 describes schools as "sorting machines" that create, incorporate, and assign students to categories that determine educational opportunities, often reinforcing reinforcing racial and class inequality.
+On the supply-side, the credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required to claim advantageous social and economic positions [@RN1297], and this race for credentials benefits affluent households [@RN861] \[cite others\].
+Scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
+
+We argue that these perspectives ignore an important mechansim of social reproduction: third-party vendors create products that sort students on behalf of college.
+These products often amplify historic inequality in educational opportunity by using past snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
+In the contemporary era of private equity funded platform capitalism, third-party recruiting products have become increasingly powerful, granular, and ubiquitous.
+However, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.
+
+In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*.
+College Board underwrote the project -- providing data and funding -- and published the book as part of their nascent efforts "to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p.
+x].
+The broader project was conceptualized as an effort to to "capture and quantify" (p.
+11) the knowledge of admissions officers, based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel" (p.
+11).
+@RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of 1980 high school seniors.
+The thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
+Therefore, colleges should recruit from territories that contain large populations of your target social class.
+Geomarkets are geographic borders meant to define local recruiting markets, a territory of an admissions recruiter.
+Figure X, panel A shows Geomarkets in the Chicago-land area.
+As a planning tool for colleges, the Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
+
+Geomarkets have been incorporated into the supply-side structure of college access in three ways.
+Anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
+Figure X panel B shows that Chicago-land Geomarkets define recruiting territories for University of Chicago admissions recruiters (Appendix Figures BLAH BLAH).
+Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), which was in 1984 and remains active today.
+EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
+According to @RN5021, in 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis product, which was based on EPS.
+
+Third, Geomarkets were incorporated into College Board's student list product, Student Search Service.
+Student lists contain the contact information of prospective students and have been the primary source of lead generation in U.S.
+higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz].
+@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+The Student Search Service database consists of College Board test-takers.
+Colleges pay a fee for each prospect (e.g., \$0.50 in 2021).
+Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc.
+Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets.
+
+Scholarship on enrollment management has focused on college behavior [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
+Scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access.
+These mechanisms, which lie upstream of the behavior of individual colleges, are incorporated by individual colleges in ways that funnel certain kinds of students to certain kinds of institutions.
+This omission is surprising because sociology is concerned with sorting [@RN4873], particularly sorting into college.
+
+This manuscript analyzes College Board Geomarkets as a case study of quantification.
+In particular, we draw from the discussions of correlation and homophily from @RN4975.
+@RN4982 identified the correlates of 1980 SAT score-sending behavior and concluded that student demand for higher education is primarily a function of social class.
+The Market Segment Model argues that homophily -- actors that share characteristics form connections -- is the organizing principle of competition and college choice, stating that "the hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students" [@RN4982, p.
+72].
+Scholarship on quantification demonstrates that making recommendations based on past correlations amplifies the effects of historic structural inequality [@RN4794].
+The snapshot of student demand in 1980 -- itself a consequence historic structural inequality -- was programmed into recruiting products that colleges utilize to identify and target prospective students.
+The result is a supply-side that amplifies structural inequalities observed on the demand-side.
+
+Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS and within the Student Search Service student list product.
+First, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
+We address this question by spatially joining the geomarket shapefile to Census data about socioeconomic and racial characteristics.
+Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
+This research question is motivated by @RN4982, which recommends that selective colleges target affluent Geomarkets.
+Additionally, prior research on enrollment management behavior finds that selective colleges focus recruiting visits on high schools in affluent, predominantly white communities [@RN3519; @RN4758].
+We analyze this research question using data from actual student list purchases, which were collected using public records requests.
+We analyze student list purchases that include all Geomarkets in a metropolitan area in only to assess which prospects would be included/excluded by had the student list purchase filtered on particular Geomarkets.
+
+The following section introduces background and scholarship about enrollment management and introduces salient concepts from scholarship on quantification, particularly correlation and homophily.
+Second, we introduce the case, describing how @RN4982 developed Geomarkets and the Market Segment Model and how these concepts were incorporated into College Board recruiting products.
+Third, we describe data and methods.
+Fourth, we present results.
+[We find that\ldots ?PARAGRAPH ON RESULTS?] Finally, we discuss implications for scholarship and for policy.
+The sociology of education should move beyond the fixation on schools and policies as the primary sorting machines.
+@RN527 [p.
+148] remind us that the organizational field includes the totality of relevant actors." Third-party vendors have structured college access for decades and their influence is growing.
+The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product \[CITE\].
+By investigating high-leverage actors and products in the organizational field, sociology can develop important insights that shape future federal and state regulatory policy.
+\[REVISE THIS PARAGRAPH\]
+
+```{=html}
+\[ADD TEXT SOMEWHERE AROUND HERE ABOUT WHY THIS TOPIC IS SIGNIFICANT AND IMPORTANT FOR CURRENT/FUTURE POLICY DEBATES: Why market segment model important.
+Software as service products dominate em industry and structure college access.
+Look at eab machine learning product But rarely do we see the recipe underneath these products.
+The book structuring college access is that recipe for EPS software\]
+<!-- 
+THIS TEXT IS FINE, BUT TRYING TO SET UP THE CONTRIBUTION TO BROADER QUANTIFICATION LITERATURE, VIS-A-VIS NO-ONE HAS STUDIED MARKET RESEARCH; DON'T THINK THIS IS NECESSARY AND YOU ARE NOT ON FIRM GROUND HERE IN CALLING THIS "MARKET RESEARCH"
+Scholarship on quantification in education has focused on consumer-facing measures such as U.S.
+News Rankings [@RN5003] and UK "league tables" about school quality [@RN5005].
+To our knowledge, prior knowledge has not examined business-facing measures produced by third-party market research.
+Here, the desirable attributes of students and places they live are quantified with the purpose of informing colleges about where they should allocate recruiting resources to efficiently enroll the most desirable students.
+We reason that market research reproduces inequality by encouraging colleges to plan their recruiting efforts based on a snapshot of social stratification in student demand that is substantially a function of historic, structural inequality in educational opportunity.
+In contrast to @RN5005, we cannot demonstrate the effect of geomarkets on inequality.
+However, we show the mechanisms of how geomarkets reproduce inequality by simulating who is included/excluded when geomarkets are utilized within student list purchases as recommended by @RN4982.
+-->
+```
+# Enrollment Management and Quantification
+
+**Enrollment Management**
+
+<!-- la la -->
+
+Enrollment management is simultaneously a profession, an administrative structure, and an industry.
+As a profession, enrollment management (EM) integrates techniques from marketing and economics in order to "influence the characteristics and the size of enrolled student bodies" [@RN2771, p.
+xiv].
+As an administrative structure, the EM office typically controls the activities of admissions, financial aid, and recruiting [@RN2406].
+<!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
+-->
+
+@fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
+The funnel begins with a large pool of "prospects" (i.e., prospective students) that the university would like to enroll.
+"Leads" are prospects whose contact information has been obtained.
+"Inquiries" are prospects that contact the institution, including those that respond to an initial solicitation (e.g., email) and those that reach out on their own (e.g., sending SAT scores).
+<!--  Applicants consist of inquiries who apply plus "stealth applicants" who do not contact the university before applying.
+The funnel narrows at each successive stage in order to convey the assumption of "melt" (e.g., a subset of "inquiries" will apply).
+-->The purpose of the enrollment funnel is to inform recruiting interventions that target one or more stages.
+These interventions seek to increase the probability of "conversion" across stages [@RN4322].
+At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz].
+Purchased leads are served emails, brochures, and targeted social media designed to solicit inquiries and applications [@RN4895].
+
+Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel it speaks to.
+The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
+Scholarship from the economics of education often investigate financial aid leveraging, which seeks to convert admits to enrolled students [e.g., @RN3762; @RN3133].
+<!--  Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935; @RN1814].
+-->
+
+A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
+@RN4758 conceptualize recruiting behavior as an indicator of college enrollment priorities.
+Ethnographies by @RN3519 and @RN4407 identify connections between private school guidance counselors and college admissions officers as a mechanism for social reproduction.
+In turn, recruiting visits to high schools -- and subsequent phone calls -- are a means of maintaining ties with guidance counselors at feeder schools and establishing relationships with prospective students [@RN3519; @RN4895].
+Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly white public schools [@RN4758; @RN4759; @RN5023].
+From the student perspective, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
+@cottom2017lower shows that for-profit colleges found a niche in Black and Latinx communities because traditional colleges ignored these communities.
+<!--  Consistent with scholarship on political economy [@RN4517; @cottom2017lower; @RN5015], this literature reveals a vertical hierarchy of postsecondary institutions matching to a vertical hierarchy of customer niches segmented by class and race, a process consistent with the notion of that higher education is mechanism of social reproduction [@RN2377; @RN4873].
+-->
+
+By focusing on the behavior of colleges, scholarship on EM implicitly states that recruiting is something done by individual colleges.
+The broader EM industry includes relevant stakeholders in the organizational field [@RN527], including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
+We argue college access is structured by third-party servicers and products that interact with direct-providers (colleges).
+Although sociologists have hinted at the ways enrollment management contributes to inequality in college access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
+<!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
+--> <!--
+[^1]: New institutional theory offers one explanation [@RN527].
+In highly developed organizational fields like higher education, forces of competitive and institutional isomorphism -- including institutional locics -- compel organizations adopt similar practices as their peers.
+Practices deemed advantageous spread quickly, for example, public research universities targeting nonresident prospects to compensate for declines in state appropriations.
+These behaviors are conceived as autonomous decisions by individual colleges, albeit structured by institutional logics.
+ -->
+
+Drawing from scholarship on organizational theory, enrollment management processes require colleges to make many "make or buy" [@RN4738; @RN980] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
+Unfortunately, few empirical studies make EM vendors or products the object of empirical analysis.
+The most widely-known class of EM vendor is EM consulting firms, which provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
+As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
+The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].[^2]
+
+[^2]: In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
+
+While the contemporary EM industry is characterized by software-as-service products sold to colleges by private equity backed firms [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
+Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit.
+The Student Search Service -- created in 1972 and copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
+Second, the Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
+Both products are applications of quantification.
+Student Search Service includes or excludes prospects from a student list purchase based on value of selected variables while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
+The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on equality of opportunity.
+
+```{=html}
+<!--
+
+The EM industry puts great stock in market research, in part because the EM profession emerged from marketing and market research [@RN4986; @RN2771].
+@RN4986 [p.
+17] define market research as all "activities intended to increase understanding of a market and consumer behavior in that market." At the most basic level, in-house market research is a function of being intentional about record-keeping and using those records [@RN3519].[^1] In recruiting, market research informs decisions about which prospective students to target with recruiting interventions.
+How to acquire market research is another make-or-buy decision.
+The College Board played a pivotal role in transforming market research about enrollment from something done by individual universities to something done by third-party professionals.
+In 1983, College Board published *Applying Market Research in College Admissions* [@RN4986], which became the basis for their Admitted Student Questionnaire (ASQ) product, and published *The Structure of College Choice* [@RN4982], which became the basis for the Enrollment Planning Service (EPS) product/division.
+Today, EPS remains the market research arm of College Board, while Encoura Eduventures is the market research arm of ACT.
+Market research often distills large amounts of data into simpler indicators.
+Therefore, the quantification literature provides a useful lens to understand market research.
+
+[^1]: @RN3519 describes the archives of annual reports of admissions activities from a selective liberal arts college.
+In the 1942 report, "Section 2.5 is titled 'Secondary School Quartile Standings," a measure of where matriculants fell in the rankings of their high school classes" [@RN3519] while Section 3.11 describes the high schools visited.
+In the chapter entitled "Travel," @RN3519 [p.
+69] describes how The College used record-keeping to inform decisions about visits to local high schools in the 2000s: "The College keeps files on hundreds of schools across the country, and around the globe, with which it had contact \ldots The files enable officers to quickly discern"how we've done here lately," \ldots with the number, relative quality, and yield rate of applicants from a particular school."
+-->
+```
+**Quantification**
+
+@RN5007 [p.
+402] define quantification as "the production and communication of numbers" (p.
+402).
+Reacivity, discipline, and authority are three interrelated themes that describe the effects of quantification.
+*Reactivity* is the idea that salient quantitative measures cause people and organizations to change their behavior.
+Quantification *disciplines* actors to react in particular ways.
+Quantification changes decision-making power *authority*, often weakening the discretion of local decision-makers.
+Scholarship about U.S.
+News & World Report (USNWR) Law School rankings demonstrate the effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
+Law school rankings were developed as a means of informing prospective students about the relative quality of different programs.
+Rankings affect (reactivity) where students apply and enroll [@RN5003].
+Rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
+Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p.
+415].
+Rankings *disciplined* the behavior of law schools.
+For example, characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
+Rankings also weakened the *authority* of admissions personnel to make independent decisions class size and which applicants to admit [@RN5003].
+<!--
+In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.    -->
+
+The analysis of UK school "league tables" by @RN5005 shows how quantification can be a mechanism of social reproduction.
+In 1992, the UK government began publishing school league tables, which ranked schools based on student performance on national exams taken at age 16.
+League tables facilitate making evaluative comparisons between schools regardless of geographic proximity.
+They also discipline households and schools to conceive of school quality in terms of test scores, which are substantially a function of the class composition of schools.
+Using Census data measured at the local authority level (similar to a U.S.
+county), @RN5005 find that localities with higher performing schools experienced growth in the share of managerial/professional households following the adoption of league tables and a decline in the share of working-class households.
+Using longitudinal survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables.
+The introduction of league tables contributed to social reproduction because professional/managerial households were more aware of these consumer-facing metrics and had resources to respond by moving to more expensive neighbhorhoods, near higher performing schools.<!-- THIS IS RATCHET EFFECT STUFF.
+I CUT IT BECAUSE YOU HAVEN'T INTRODUCED THE IDEA OF RATCHET EFFECT
+Schools that performed well on initial league tables tended to be schools with a high share of professional/managerial families.
+Therefore, quantification takes a snapshot of existing social stratification and that snapshot becomes a power unto itself, amplifying the effect of previous strctural inequality on subsequent structural inequality.
+-->
+
+```{=html}
+<!-- THIS IS THE CUT TRANSITION PARAGRAPH
+In the broader quantification literature, critical data studies [e.g., @RN4975; @RN4849; @RN4772] and sociological analyses of technology and inequality [e.g., @RN4778; @RN4786; @RN4877; @RN4810] introduce concepts salient to the analysis of market research.
+Our focus is on "selection devices," which utilize some combination of inputs and discretion to categorize places, people, and organizations for the purpose of allocating opportunities or interventions [@RN4778; @RN5012].
+-->
+```
+**Correlation**.
+The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4772; @RN4849].
+In particular, the discussions of correlation (chapter 1) and homophily (chapter 2) by @RN4975 introduce ideas salient to the analysis the Market Segment Model and Geomarkets.
+Correlation measures the extent to which two or more variables move together.
+Predictive analytics are based on correlation and are developed in two steps.
+First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
+Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and/or to assign levels of risk to each case.
+@RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
+These models predict outcomes based on correlations without requiring knowledge about underlying causal relationships.
+@RN4975 [p.
+50] writes that
+
+> correlation grounds big data's so-called revolutionary potential.
+As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'
+
+Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
+For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
+@RN4849 [p.
+146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
+
+Many studies show that predictions based on correlations reproduce structural inequality [for a review see @RN4794].
+The correlations observed during the training data stage are a snapshot of relationships between variables at a particular point in time.
+Observed correlations may be a function of enduring structural inequality, but underlying causes are not considered by applications of predictive models.
+Reviewing scholarship about algorithms, @RN4794 [p.
+224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
+This phenomenon has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^3] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
+They then recommended that selective colleges recruit from localities that have a high share of affluent, college-educated households.
+
+[^3]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
+Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
+Note that predictive models such as the LSI-R model are not merely used for social science purposes.
+Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.
+
+**Homophily**.
+Homophily is the idea actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together" [@mcpherson2001birds].
+Homophily is a core concept of network science, in which actors (nodes) are connected to one another directly and indirectly via network ties (edges).
+In *The company We Keep: Colleges and Their Competition*, @RN4982 [chapter 4] conduct a network analysis whereby two colleges are defined as competitors if a large number of students submit SAT scores to both colleges.
+@RN4975 argues that because network science models often draw from rational choice theory, they assume that homophily is the result of voluntary action by individuals.
+
+@RN4975 problematizes the idea that homophily is a naturally occurring phenomenon.
+Because network science models -- like correlational models above -- typically do not observe how historical structures affect relationships, they "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p.
+95] that cause behavior consistent with homophily.
+Second, in commercial social networks, homopphily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
+Thus, "social networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@RN4975, p.
+YY].
+We observe similar phenomena in our case.
+Based on a snapshot of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
+In turn, @RN4982 reasons that colleges should target the Geomarkets and high schools targeted by peer colleges.
+This logic is subsequently programmed into EPS software that recommends which Geomarkets and high schools to recruit from.
+Thus, homophily observed on the demand side -- which is a function of historic structural inequality –- is programmed into the supply-side.
+
+Homophily is central to market research products that categorize customers.
+Geodemography emerged in the 1970s as a branch of market research that estimates the behavior of consumers based on where they live [@RN4878].
+Market segments are subgroups within a larger market that have similar consumer demand.
+Early geodemographic classifications of consumers (e.g., PRIZM by Claritas Corporation) were derived from publicly available Census data, which disaggregated data to the zip code level.
+The Claritas Corporation had a financial incentive to argue that people living near one another share similar consumer preferences because geographic localities could then be categorized into market segments that would be useful for direct mail marketing campaigns [@RN4957].
+Later, the development of individual credit scores (e.g., FICO score) enabled merchants to classify consumers into many, fine-grained groups [@RN4877].
+@RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
+We observe similar processes in our case study.
+@RN4982 categorizes students into four market segments -- local, in-state, regional, and national -- based on their score sending behavior and then evaluates the attractiveness of each Geomarket based on how many "regional" and "national" students live there.
+<!--
+ORIGINAL TEXT ABOUT GEODEMOGRAPHIC CLASSIFICATION SYSTEMS
+
+Clusters were given evocative names like "Blue Blood Estates," "\[Archie\] Bunker's Neighbors", and "Share Croppers." Because U.S.
+publicly available U.S.
+Census data disaggregated data only to the zip code-level, the Claritas Corporation had a strong incentive to argue that consumer behavior within zip codes was very similar [@RN4957].
+@RN4565 emulated the behavior of Claritas by developing the *Segment Analysis Service: An Educationally Relevant Geodemographic Tagging Service*, which was an "add-on" to the *Student Search Service* student list product.
+Describing its conceptual underpinnings, @RN4565 [p.
+1] states, "the basic tenet of geodemography is that people with similar cultural backgrounds, means, and perspectives naturally gravitate toward one another or form relatively homogeneous communities; in other words, birds of a feather flock together." The Segment product allocated each high school and each census tract into a distinct "educational neighborhoods" thought to reflect useful market segments for colleges looking for students.
+
+PREVIOUS TEXT ABOUT CLASSIFICATION SITUATIONS
+Related to the creation of market segments, @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
+Historically, classifications were binary.
+Consumers with "good" credit were offered loans and those with bad credit were not.
+The innovation of individual FICO scores -- and, later, other credit scores -- enabled merchants to classify consumers into many, fine-grained groups or along a continuum.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].
+People previously denied credit were now granted credit on different terms.
+For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
+Thus, at the bottom of the continuum, contemporary classification situations produce "predatory inclusion" -- exemplified by payday loans and for-profit colleges -- which is the logic of "including marginalized consumer-citizens into ostensibly democratizing mobility schemes on extractive terms" [@RN4774, p.
+443].
+
+Negative consequences of homophily
+Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
+
+Reviewing scholarship on quantification, @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
+U.S.
+law school rankings [@RN5003] and UK school league tables [@RN5005] exemplify this sort of salience.
+Studies of quantification in the sociology of education tend to study consumer-facing metrics.
+Law school rankings can be conceived as a business-to-consumer product because the goal is to inform consumers (prospective law school students) and consumers pay for the product.
+
+-->
+
+[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
+
+We make targeted contributions to scholarship on quantification in the sociology of education.
+First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
+As @RN3519 demonstrates in the chapter *Numbers*, EM is fundamentally concerned with quantification.
+However, sociological analyses of enrollment management focus on the behavior of colleges and their agents, yielding the implicit belief that inequalities created by EM are a function of individual colleges reacting to rankings and other macro structures in their environment.
+To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of producers.
+Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027; ].
+Drawing from scholarship from critical data studies [e.g., @RN4775; @RN4975; @RN4849], recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
+Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
+Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociologists.
+We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
+
+This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
+The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
+Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
+Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification on social reproduction.
+However, by showing which students are included/excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in educational opportunity.
+
+```{=html}
+<!--
+The market segment model became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
+@RN5021 reports that in 1995, 37% of 4-year public universities and 49% of 4-year private universities used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product, which was similar to EPS.
+Geomarkets later became a filter in the College Board *Student Search Service* student list product, which is utilized by the vast majority of public and private universities.
+
+By contrast, many quantification projects -- for example, human rights indicators -- have "little oomph" [@RN5006, p.
+265].
+FROM BERMAN & HIRSCHAN PAGE 263: "Merry demonstrates how human trafficking indicators largely sit on the shelf because no audience for them has been created."
+-->
+```
+# The Market Segment Model and College Board Geomarkets
+
+```{=html}
+<!--Historically, College Board partnered with Members to develop projects that later became revenue-generating products for College Board.
+
+As a member organization, several College Board products developed from joint collaborations with administrators or professors from member institutions.
+During the 1970s, under the leadership of Larry Litten, then director of institutional research, Carleton College pionered applying market research to student recruiting.
+In the late 1970s, College Board provided Carleton College with funding and data for a six-year project that analyzed college choice decisions by households in six metropolitan areas.
+This project was the basis for the landmark book *Applying Market Research in College Admissions* [@RN4986], published by College Board, and was the basis for the College Board Admitted Student Questionnaire (ASQ), a revenue-generating product that provides colleges with information about admitted students -- those who enrolled and who did not enroll.
+Litten would later become Associate Director of the Consortium of Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
+
+The Market Segment Model was an effort to quantify and communicate the knowledge that admissions officers have about student demand. 
+-->
+```
+**Creating Geomarkets and the Market Segment Model.**
+
+In 1978, set against the backdrop of impending college-age demographic decline, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
+x].
+To answer these questions, Zemsky began working with the Market Research Committee of the Consortium on Financing Higher Education (COFHE), a consortium of 30 selective private universities founded in the mid-1970s.
+The project goal became the creation of the Market Segment Model, which had the lofty ambition of predicting student demand for any college from any locality.
+However, "to gain a truly comprehensive view of the collegiate enrollment market, we needed a database that described most institutions and most students" [@RN4982, p.
+x].
+The group approached College Board.
+"Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest deomgraphic unit used in these analyses was the state" [@RN4982, p.
+x].
+
+Although the analyses of @RN4982 are based on the score-sending behavior of SAT test-takers, the authors had great respect for admissions officers and conceived of the project as an effort to "capture and quantify" (p.
+11) their knowledge:
+
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
+9-10).
+
+The "initial task" was to define geographic boundaries "in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982,p.
+4].
+Quoting an admissions officer, @RN4982 \[p.
+11\] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, @RN4982 [p.
+11] created "three types of boundaries -- region, state, and community.
+The initial regions in the project were New England, Middle States, and the South.
+Next, they "divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" (p.
+11).
+These enrollment markets, later called Geomarkets, were intended to be consistent with the conception of a catchment market from the perspective of admissions counselors.
+@RN4982 [pp.
+11-12] described the creation of Geomarket borders only briefly:
+
+> In many cases, the market boundaries match formal political and educational divisions, reflecting natural channels of communication.
+Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
+In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."
+
+Having defined geography, the Market Segment Model sought to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households in the Geomarket:
+
+> For our analysis, we sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
+The Market Segment model that we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
+The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways [@RN4982, p.
+4].
+
+SAT test-takers were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
+For a given student, each college that receives a score from the student can be defined as "local" (college located in the the same Geomarket as the student), "in-state" (same state but different Geomarket), "regional" (same region but different state), or "national" (different region).
+In turn, a test-taker is categorized in the "local" market segment if they submit more SAT scores to local institutions than they do to in-state, regional, or national institutions.
+An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
+<!--
+A "regional" student submits more SAT scores to regional institutions than they do to local, in-state, or national institutions.
+A "national" student submits more SAT scores to national institutions than they do to local, in-state, or regional institutions.
+-->
+
+The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
+Appendix A describes these outputs in more detail.
+Both outputs are created separately for each Geomarket.
+For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+These outputs became the basis for the Enrollment Planning Service (EPS) software.
+
+[^5]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
+
+[^6]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
+
+**Correlations**.
+@RN4982 [Chapter 3], *A Sense of Place: Students, Families, and Communities*, identifies the student characteristics correlated with being in the local, in-state, regional, or national market segment.
+The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
+Because these four variables "reflect the basic social patterns of the nation, it would have been surprising if these were not the four social variables that best explained the patterns of college choice" [@RN4982, p.
+33].<!--
+and "The information derived from the Market Segment Model, however, is a remarkably ordered set of data, consistent in its relationships, that reflects the basic social and economic patterning associated with the structure of college choice" (pp.
+34-35).
+\[QUOTES REDUNDANT\]
+-->
+
+The thesis of Chapter 3 is that student demand for higher education is a function of social origin.
+For the authors (p.
+33) state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and that "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 \[p.
+33\] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."\[\^7\] @RN4982 [p.
+42] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
+However, Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
+Therefore, @RN4982 [p.
+44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:
+
+> On occasion, senior spokespersons for the profession worry that students outside the main \[Geo\]market areas remain forgotten and hence, unchallenged.
+Inevitably, the increasing competition for students, the expense of travel and mailings, and internal political constraints compel institutions to concentrate their efforts where they will do the most good.
+The result is a natural reinforcing of the basic socioeconomic patterns that gave shape in the first place to the structure of college choice.
+
+**Homophily**.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^7] Based on these analyses, @RN4982 \[p.
+46\] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
+Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
+The authors argue that like-colleges compete for like-students as defined by socioeconomic characteristics.
+@RN4982 [p.
+72] describe observed patterns as a natural process in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+
+[^7]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
+Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
+These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
+For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
+54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
+For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
+
+> Students describe themselves socially simply by telling us the colleges and universities in which they are interested.
+The layering of collegiate competition is primarily a socioeconomic layering.
+The hierarchical structure of collegiate competition largely reflects the stratified social and economic dimensions of the communities from which colleges draw their students.
+Competition among colleges, as admissions officers have told us for so long, is in fact, a matter of keeping company with one's peers.
+
+The discussion of competition by @RN4982 exemplifies the concerns about correlation and homophily described by @RN4975.
+A correlational analysis of 1980 SAT score-sending patterns finds competition between colleges is defined by socioeconomic homophily.
+This homophily is presented as a naturally occurring phenomenon.
+Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer-coleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
+In itself, @RN4982's Market Segment Model is a social science depiction of student demand -- akin to the status attainment model -- that does not consider historic, structural inequalities that cause observed patterns.
+However, by inscribing the Market Segment Model and Geomarkets into the EPS software, College Board amplified structural inequalities that contributed to homophily observed in 1980 SAT score-sending beahivor.
+In this way, commodification of the Market Segment Model and Geomarkets "engineers homophily" [@RN4975].
+
+**Enrollment Planning Service**.
+In 1984, College Board created the Enrollment Planning Service (EPS), based on the Market Segment Model [@RN4987; @RN4980].
+EPS was an early software-as-service platform that recreated the analyses of @RN4982.
+For each Geomarket, colleges could obtain Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors.
+Based on background conversations with enrollment management professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket.
+Therefore, colleges used EPS software to decide which Geomarkets to recruit from and which high schools to visit within targeted Geomarkets.
+Typical @RN5019 marketing material describes EPS as,
+
+> The marketing software that pinpoints the schools and Geomarkets where your best prospects are most likely to be found.
+With the click of a mouse, EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition.
+Focus your valuable time and resources on the right prospects.
+
+```{=html}
+<!--
+Promotional material highlights the themes of "locate the students you want" [@RN5019_2003] and "focus your valuable time and resources on the right prospects" [@RN5019].
+@RN5019_2010 promotional material describes EPS as:
+
+> the analysis tool that pinpoints the schools and geomarkets where your best prospects are most likely to be found.
+You can research your existing (or new) markets using selection criteria and locate your top prospects in various ways -- by states, geomarkets, counties, zip codes, high schools and international regions.
+EPS provides you with comprehensive reports on your markets, your position in those markets, and your competition."
+
+Software documentation by @RN5017 states that "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas, mostly in the United States.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
+-->
+```
+Whereas @RN4982 identified 143 Geomarkets covering the New England, Middle States, and South region, EPS created Geomarkets for the remaining U.S.
+states, with 304 Geomarkets in total.
+@RN4988 shows the contemporary Geomarkets.
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^8] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
+
+[^8]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
+Students from California don't all share the same college-going behaviors.
+We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."
+
+@RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
+Market research suggests that EPS was highly salient.
+@RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
+
+EPS software may "discipline" [@RN5007] colleges to approach recruiting in a manner consistent with the Market Segment Model.
+Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
+EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
+Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
+
+[^9]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
+
+EPS may also weaken the authority of local decision-makers.
+@RN4982 sought to develop a concrete, data-driven framework -- The Market Segment Model -- that replicates the aggregate knowledge of local admissions officers.
+Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
+The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
+In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
+However, EPS software often recommended visiting the same sets of affluent, high-achieving high schools that were receiving visits from other colleges.
+We were told that savvier, quantitatively-adept admissions offices used EPS to visit schools that EPS recommended ignoring because there would be less competition for the good students who attended these high schools.
+
+Finally, we suspect that Geomarkets affect how admissions offices organize the recruiting territories assigned to each admissions officer.
+Geomarkets were created to mirror the territories of admissions officers [@RN4982].
+In turn, as Geomarkets became more salient, admissions offices often structured their territories around Geomarkets.
+On background, admissions officers and enrollment consultants told us it remains common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket.
+In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
+This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX CRYSTAL\] <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
+
+Although we cannot gain access to EPS software, Geomarkets are fundamental to EPS software and (arguably) to the organization of recruiting territories for many colleges.
+Therefore, we are interested in the extent to which this building block is associated with race and class.
+Research question 1 asks, what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
+We focus on metropolitan areas that are associated with three or more Geomarkets.
+
+We expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
+Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
+Moreover, @RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards identifying geographic areas that differed from one another in terms of class composition.
+
+```{=html}
+<!-- 
+Additionally, @RN4982 created geomarkets in service of the Market Segment Model.
+The Market Segment Model argues that socioeconomic status drives whether students belong to the local, in-state, regional, or national market segment.
+In turn, geomarkets are described by the relative numbers of local, in-state, regional, or national students, which is substantially a function of socioeconomic status.
+The language of @RN4982 [pp.
+11-12, quoted above] suggests that geomarket borders within large metropolitan areas may have been drawn in ways that separate affluent from less affluent communities.
+We present the following hypothesis:
+-->
+```
+We also expect substantial racial inequality between geomarkets in large metropolitan areas.
+Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race.
+U.S.
+cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779].
+Structures built upon racialized structures are racialized structures [@RN4786].
+Unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
+Second, although geomarket borders may have been drawn along class divides, a strong correlation exists between race and wealth [@RN4884].
+Third, Geomarket borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
+Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
+
+**Geomarket Filter in Student Search Service**.
+<!-- By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service \[CITE\].
+--> A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^10] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+
+[^10]: For public universities that purchased lists, 80% purchase more than 50,000 names annually.
+@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.
+
+Student lists are positively associated with student outcomes.
+A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^11] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^12]
+
+[^11]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
+Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
+
+[^12]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
+Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
+The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).
+
+However, @RN5012 argue that student list products may exacerbate racial inequality in college access.
+The authors conceptualize student list products as "selection devices" [@RN4778] that enable colleges to select which prospective students they target by incorporating search filters (e.g., high school graduating class, state).
+@RN4786 defines "racialized inputs" as ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
+@RN5012 argue that several frequently utilized student list filters (e.g., zip code, AP test score, SAT score) meet the criteria of racialized inputs.
+Using a national sample of high school students and using data from actual student lists purchased by public universities, @RN5012 show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
+
+Because of the extent of residential segregation in the U.S., geographic borders are a commonly studied racialized input in scholarship about algorithmic bias [@RN4775; @RN4849; @RN4887].
+Student list products offer many geographic filters.
+Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
+College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^13] This manuscript focuses on Geomarkets.
+In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
+Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
+First, they are ostensibly race-neutral inputs in that neither @RN4982 -- nor subsequent EPS promotional material [e.g., @RN1621] -- mention race.
+Second, Geomarkets may be systematically correlated with race.
+For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda Country, excluding Oakland." A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.
+This article extends @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
+
+[^13]: Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
+
+@RN4982 [pp. 42-44] discuss how the Market Segment Model and Geomarkets should be utilized in the context of purchasing student lists. The authors ask,
+
+> Imagine for the moment that you recruit for a college that draws most of its students from regional or national segments.
+Where would you concentrate your energies?
+Ideally, you would seek communities \[Geomarkets\] with a high proportion of students already predisposed toward institutions such as your own.
+The Market Segment Model would provide this information through segment percentages for the community \[Geomarket\] in question.
+Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
+Your efforts would surely be better directed toward \ldots Manchester, Hartford, and Fairfield County \ldots Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own.
+
+In both mailings and off-campus recruiting visits, @RN4982 recommend focusing on Geomarkets with large populations of students who are deciding between "institutions like your own" (p.
+44).
+Selective colleges primarily draw from students in the "regional" and "national" segments.
+@RN4982 suggest that the student list purchases of selective colleges should focus on Geomarkets with large numbers of affluent, highly educated households.
+By contrast, the student list purchases of local state colleges and community colleges, which rely on enrollment from students in the "local" and "in-state," should focus on nearby Geomarkets with large numbers of low-SES students.
+Our analyses attempt to recreate the strategy recommended by @RN4982 by showing who is included/excluded when student list purchases filter on particular Geomarkets.
+Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access.
+Furthermore, we anticipate that the students of color from affluent Geomarkets will tend to have college-educated parents and be enrolled in affluent, predominantly white schools.
 
@@
 
+# Findings
+
+## Chicago
+
+### Shifting Demographics, Persistent Segregation in Geomarket Landscape
+
+@fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
+The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 in the development of The Market Segment Model.
+Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
+In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
+The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most densely populated Geomarket.
+
+These population shifts have contributed to changes in the racial/ethnic composition within and across Geomarkets over time.
+@fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarkets across the three census periods.
+In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
+In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
+By 2020, several Geomarkets that were previously predominantly White shifted to a more balanced racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
+For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
+Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
+
+@fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
+For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
+City of Chicago also experienced relatively modest declines in its Black population (37% in 1980 to 33% in 2020), alongside increases in its Hispanic (15% in 1980 to 26% in 2020) and Asian Pacific Islander populations (4% in 2000 to 6% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
+
+We also explore socioeconomic demographics - median income, percent living below poverty, and educational attainment - within and across Geomarkets over time in @fig-chicago-rq1-ses.
+Overall, median income both within and across Geomarkets in the Chicago metropolitan area remained relatively stable across the three census periods.
+For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$158,000 in the North Shore Geomarket to a low of \$66,000 in the City of Chicago Geomarket.
+However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence, such as \$120,000 in the Northwest Suburbs, \$119,000 in Evanston and Skokie, \$108,000 in Chain of Lakes, \$105,000 in the Western Suburbs, and \$99,000 in the South and Southwest Suburbs.
+While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$79,000) and decrease in the South and Southwest Suburbs Geomarket (\$93,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$14,000.
+
+@fig-chicago-rq1-ses also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
+Similar patterns are evident in educational attainment across Geomarkets.
+In 1980, the City of Chicago had the lowest percentage of residents with a bachelor's degree or higher (13%).
+However, by 2020, the Geomarket experienced the largest percentage point increase in college degree attainment, reaching 38%.
+In comparison, all other Geomarkets experienced smaller gains in the share of residents with a bachelor’s degree or higher over the same period.
+These trends contributed to a narrowing of the educational attainment gap across Geomarkets from 1980 to 2020.
+
+From 1980 to 2020, these demographic shifts have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by racial/ethnic and socioeconomic characteristics.
+The original Geomarket boundaries delineated significant racial and income segregation in the 1980s.
+This included Black, Hispanic, and Asian residents primarily in the City of Chicago Geomarket and White residents in Geomarkets forming the suburban ring of the metropolitan area.
+Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs.
+However, declining White populations, Hispanic and Asian population growth, and Black outmigration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining.
+The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents) while the the Evanston and North Shore Geomarkets remain predominantly White.
+While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains across Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
+
+### Composition of Students Lists by Geomarket Contributions
+
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospects across the entire state of Illinois.
+Two of these orders applied the same filters: SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
+The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with relatively middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
+Because all six orders included an Illinois state filter, we can utilize prospects from the Chicago metropolitan area within the resulting student lists to examine Geomarket contributions by simulating the application of a Geomarket filter across low-, middle-, and high-range SAT threshold orders.
+
+@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area.
+At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets generally reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
+For example, more than 8,000 of the 16,485 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
+Similar to the larger metropolitan area, the Northshore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students.
+Interestingly, while the City of Chicago is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-chicago-rq2-race includes nearly double the proportion of Hispanic (50%) prospects relative to the larger Geomarket resident population (26% Hispanic) in @fig-chicago-rq1-race with a lower proportion of White prospects (14%).
+
+However, the racial/ethnic composition of prospect pools become less similar to the overall metropolitan area as PSAT scores increase.
+Notably, the shares of White and Asian prospects within Geomarkets increase, while Black and Hispanic prospects decrease.
+For instance, Asian prospects over representation in Geomarkets relative to the overall resident population at the 1310-1600 SAT score threshold ranges from a minimum of 4 percentage points in the North Shore Geomarket (17% of prospects versus 13% of residents) to a maximum of 21 percentage points in the Chain of Lakes (28% of prospects versus 7% residents) Geomarket.
+The City of Chicago Geomarket exhibits the largest underrepresentation of Black prospects (7% of prospects versus 33% of residents) while Chain of Lakes exhibits the largest underrepresentation of Hispanic prospects (7% of prospects versus 24% of residents).
+
+We also explore contributions across Geomarkets to overall prospect pools by race/ethnicity.
+For instance, @fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle range SAT scores (1160-1300) by race/ethnicity.
+The top panel presents the full distribution of the 13,643 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% reside in Chain of Lakes, 13% in Northwest Suburbs, 5% in North Shore, 5% in Evanston and Skokie, 17% in City of Chicago, 28% in Western Suburbs, and 20% in South and Southwest Suburbs.
+These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-midSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
+
+```{=html}
+<!-- OLD RESULTS BASED ON LOW VERSUS HIGH SAT ORDERS
+Nearly one half (n=7,728) of all prospects whose contact information was purchased in these lower SAT score orders identified as White.
+With the exception of the City of Chicago, @fig-chicago-rq2-lowSAT-race shows all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
+The City of Chicago Geomarket, however, contributed only 5% of White prospects, despite accounting for 20% of all included prospects.
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, exept for Evanston and Skokie (12% versus 4% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (10% versus 26% for all prospects) contributing smaller shares relative to their overall representation.
+
+In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-lowSAT-race.
+Among the more than 2,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 34%, respectively, compared to their overall contributions of 20% and 26% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
+The City of Chicago Geomarket also contributed larger shares of the nearly 5,000 Hispanic students (35% versus 20% for all prospects), while all other Geomarkets contributed comprable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
+
+@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores.
+Both figures demonstrate how representational Geomarket patterns across race/ethnicity remain relatively consistent even as SAT scores increase.
+For instance, @fig-chicago-rq2-midSAT-race shows that the Chain of Lakes, North Shore, and Western Suburbs Geomarkets contributed similar proportions of White and Asian students, yet smaller proportions of Black and Hispanic students, relative to their overall representation among all included prospects.
+The Northwest Suburbs and Evanston and Skokie similarly contributed a disproportionately higher share of Asian prospects, but lower shares of Black and Hispanic students.
+Conversely, the South and Southwest Suburbs Geomarket contributed a substantially larger proportion of Black students, while the City of Chicago yielded larger proportions of both Black and Hispanic students- again, relative to their overall representation among all included prospects targeted at middle range SAT scores.
+Despite declines in the overall number of Black (n=240) and Hispanic (n=950) prospects included in orders filtering for high-ranges of SAT scores (1310-1600), @fig-chicago-rq2-highSAT-race shows overall Geomarket contribution patterns persist.
+
+We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
+Similar to race/ethnicity figures above, @fig-chicago-rq2-lowSAT-ses presents the Geomarket distribution for the 16,060 total prospects (whose first-generations status is known) included in low-range SAT score orders (1020-1150).
+Among the 3,479 total prospects whose parents did not attend college, the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 34% in comparison to 20% for all included prospects in this SAT range.
+In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students relative to all included prospects, with such disparity ranging from a one-percentage-point difference (Chain of Lakes, Northwest Suburbs, North Shore) to a five-percentage-point difference (Western Suburbs).
+For prospects whose parents attended some college but did not complete their degrees, @fig-chicago-rq2-lowSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
+These representational patterns are reversed among prospects that are not first-generation college students.
+The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (13%) relative to all prospects (20%) in this SAT range, whereas all other Geomarkets contributed either comparable (Northwest Suburbs, South and Southwest Suburbs) or disproportionately larger shares (Chain of Lakes, North Sore, Evanston and Skokie, Western Suburbs).
+
+@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status become increasingly pronounced in orders targeting middle-range (1160-1300) and high-range (1310-1600) SAT scores, respectively.
+At the middle SAT score range in @fig-chicago-rq2-midSAT-ses, the City of Chicago Geomarket contributes a substantially larger share of first-generation students whose parents did not attend college—38%, compared to 17% of all included prospects in this range, representing a 21-percentage-point disparity.
+This disproportionality increases to a 25-percentage-point difference at the highest SAT score thresholds in @fig-chicago-rq2-highSAT-ses, where the City of Chicago accounts for 40% of such first-generation students, despite representing only 15% of all included prospects in this score range.
+
+Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
+@fig-chicago-rq2-lowSAT-combo presents the Geomarket distribution of the 15,183 total prospects (whose first-generation status and race/ethnicity is known) included in low-range SAT (120-1150) orders.
+For instance, among all White prospects (n=7,605) whose contact information was purchased in this SAT score range, approximately 11% were first-generation college students whose parents did not attend college, 22% were first-generation college students whose parents attended some college but did not complete their degree, and 67% were not first-generation college students.
+@fig-chicago-rq2-lowSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (22% and 15%, respectively) and whose parents attended some college but never completed degrees (29% and 28%, respectively).
+In contrast, Geomarkets such as Chain of Lakes (72%), North Shore (88%), Evanston and Skokie (73%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
+
+@fig-chicago-rq2-lowSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
+The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
+However, the magnitude of this disproportionality varied- from a 3-percentage-point difference for Black prospects to a 21 percentage-point difference for Asian prospects.
+Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
+Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-lowSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, and Evanston and Skokie Geomarkets.
+
+Geomarket contribution patterns to prospect pools by both first-generation college student status and race/ethnicity become more pronounced as SAT score thresholds increase.
+For instance, @fig-chicago-rq2-midSAT-combo presents prospect pools disaggregated by both first-generation college student status and race/ethnicity within the middle-range SAT score (1160–1300) orders.
+The City of Chicago Geomarket still contributes a disproportionately larger share of first-generation college students for both those whose parents did not attend college and those whose parents attended some college but did not complete a degree across most racial and ethnic groups.
+However, the magnitude of this disproportionately for City of Chicago becomes even greater for Asian students, with a nearly 30-percentage-point difference collectively for all first-generation college students (35% no college and 22% some college) relative to their overall Geomarket representation (28%).
+Similar to patterns observed at lower SAT score thresholds, the Geomarket contributions to first-generation Hispanic college students remain relatively balanced.
+All Geomarkets, with the exception of North Shore (22%) and Evanston and Skokie (32%), contributed more than 50% first-generation Hispanic college students across both parental education categories to the overall prospect pool.
+
+@fig-chicago-rq2-highSAT-combo shows the magnitude of Geomarket contribution disparities increases at the highest SAT score range (1310-1600), partly driven by the relatively fewer number of first-generation college students included within these orders.
+Again, the City of Chicago Geomarket disproportionately contributes a disproportionate share of first-generation Asian college students (43% across both parental education categories) and first-generation Hispanic college students (52% across both parental education categories) in comparison to the overall distribution of first-generation college students within each racial/ethnic category (13% for Asian and 40% for Hispanic, respectively).
+Conversely, nearly all other Geomarkets contribute a disproportionately larger share of non-first generation college students across all racial/ethnic categories.
+-->
+```
+More than half (n=8,256) of all prospects whose contact information was purchased in these middle SAT score orders identified as White.
+@fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
+One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects.
+On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively.
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+
+In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race.
+Among the nearly 1,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
+The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
+
+Appendix figures \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting lower-range (1020-1150) and high-range (1310-1600) SAT scores.
+Both figures demonstrate how representational Geomarket contributionsacross race/ethnicity, such as City of Chicago disproportionately contributing larger shares of Black and Hispanic prospects, remain consistent even at lower and higher SAT score thresholds.
+
+We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
+Similar to race/ethnicity figures above, @fig-chicago-rq2-midSAT-ses presents the Geomarket distribution for the 13,749 total prospects (whose first-generations status is known) included in middle-range SAT score orders (1160-1300).
+Among the 1,743 prospects whose parents did not attend college ("no college"), the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 38% in comparison to 17% for all included prospects in this SAT range.
+In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students whose parents did not attend college relative to all included prospects, with such disparity ranging from a one-percentage-point difference (South and Southwest Suburbs) to a nine-percentage-point difference (Western Suburbs).
+For prospects whose parents attended some college but did not complete their degrees ("some college"), @fig-chicago-rq2-midSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
+These representational patterns are reversed among prospects that are not first-generation college students.
+The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (12%) relative to all prospects (17%) in this SAT range, whereas all other Geomarkets contributed comparable shares.
+
+Relative to @fig-chicago-rq2-midSAT-ses, appendix figures \@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status are less pronounced in low-SAT range orders (1020-1150) and more pronounced in high-range SAT orders (1310-1600), respectively.
+
+Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
+@fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 12,981 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders.
+For instance, among all White prospects (n=8,155) whose contact information was purchased in this SAT score range, approximately 6% were first-generation college students whose parents did not attend college, 15% were first-generation college students whose parents attended some college but did not complete their degree, and 79% were not first-generation college students.
+@fig-chicago-rq2-midSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (10% and 9%, respectively).
+For first-generation college students whose parents attended some college but never completed degrees, only the South and Southwest Suburbs (25%) Geomarket contributed a substantially larger share of White prospects.
+In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
+
+@fig-chicago-rq2-midSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
+The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
+However, the magnitude of this disproportionality varied- from a 1-percentage-point difference for Black prospects to a 23 percentage-point difference for Asian prospects.
+Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
+Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs Geomarkets.
+Appendix figures @fig-chicago-rq2-lowSAT-combo and \@fig-chicago-rq2-highSAT-combo show the magnitude of these Geomarket contribution disparities become more pronounced at as SAT score thresholds increase.
+
+The enduring segregated structure of the Chicago metropolitan area from 1980 to 2020 analyzed above plays a significant role in how Geomarkets shape the racial/ethnic and socioeconomic composition of prospect pools.
+Reflecting the larger metropolitan area dynamics, central Geomarkets that are predominantly comprised of low-income communities of color such as the City of Chicago and South and Southwest Suburbs consistently contribute disproportionately larger shares of Black, Hispanic, and first-generation college prospects, especially at lower SAT score thresholds.
+On the other hand, predominantly White and affluent Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, as well as non-first generation college students.
+However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+
+## Dallas-Fort Worth
+
+### Suburban Geomarket Growth Eases Racial Disparities, Concentrated Wealth Persists
+
+The six Geomarkets within the Dallas-Fort Worth metropolitan area are illustrated in @fig-dallas-maps by total population across the three census periods.
+The Dallas-Fort Worth region represents a unique "metroplex" case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic urban core and suburban ring model showcased by Chicago.
+As depicted in @fig-dallas-maps, two distinct inner-core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
+These central Geomarkets are surrounded by the more conventional outer ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
+Worth Metroplex).
+
+From 1980 to 2020, the population of the overall Dallas-Fort Worth metropolitan area increased substantially from approximately 2.7 Million to 6.8 Million.
+@fig-dallas-maps showcases all six Geomarkets experienced population growth over the three census periods.
+However, the most pronounced increases occured in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
+These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers - City of Forth Worth and City of Dallas - into the surrounding suburban and rural areas of the metropolitan areas.
+This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket from 1980 to 2000.
+
+@fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
+In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
+By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
+In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
+Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.
+
+These shifts in racial and ethnic composition within and across Geomarkets became more pronounced between 2000 and 2020.
+The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents.
+However, both Geomarkets remained predominantly White (58% and 63%, respectively).
+In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian Pacific Islander), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian Pacific Islander).
+Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
+
+Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
+Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
+Between 1980 to 2000, affluence levels across Geomarkets were consistent, ranging from a low of \$82,000 in City of Fort Worth to a high of \$135,00 in the Collin & Rockwall Counties Geomarket.
+However, the substantial increase in median income for the Collin & Rockwall Counties Geomarket during this time period plateaued by 2020, reshuffling relatively affluence levels among Geomarkets.
+By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000).
+In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable and continued to rank at the lower levels of affluence within the metropolitan area.
+\[KS: NEED TO CHECK THIS; I THINK CITY OF DALLAS AND CITY OF FORT WORTH ARE MIXED UP BETWEEN GRAPH VERSUS TABLE\]
+
+@fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
+In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets.
+Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex, while increasing for all other Geomarkets.
+These divergent trends resulted in a widening gap in poverty levels across the Dallas-Fort Worth metropolitan area.
+A similar widening disparity is evident in educational attainment.
+Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
+
+The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s.
+While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
+Such growth is largely attributed to increases in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex.
+On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment across Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
+Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
+
+### Composition of Students Lists by Geomarkets
+
+We also analyze student list orders collected through public requests to assess how the racial/ethnic and socioeconomic composition of included versus excluded prospects varies when filtering on particular Geomarkets within the Dallas-Fort Worth metropolitan area.
+Specifically, we analyze three list orders placed by a research university targeting prospects across 87 different Geomarkets nationwide, including all six Geomarkets in the Dallas-Fort Worth Metropolitan area.
+Each order also filtered for prospects in the 2019-2020 graduating class and based on PSAT scores across three ranges: 1070-1180, 1190-1260, and 1270-1520.
+Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260) from the Dallas-Fort Worth metropolitan area to simulate Geomarket contributions and assess patterns of inclusion versus exclusion.
+
+@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area.
+At the lowest PSAT score thresholds, notable differences emerge in the racial/ethnic composition of prospects within Geomarkets when compared to the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
+For example, the largest number of prospects in the 1070-1180 PSAT range were from the Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian), both of which yield predominantly White prospect pools.
+On the other hand, the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarket pools have the greatest proportions of non-White prospects (although yielding fewer overall students).
+Interestingly, while the City of Dallas is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-dallas-rq2-race includes larger proportion of Hispanic (41%) prospects but lower proportions of Asian (4%) and Black (9%) students comparatively.
+
+These racial/ethnic composition patterns shift as PSAT scores increase.
+Shares of White prospects within Geomarkets remain relatively proportional to the overall 2020 metropolitan distribution in @fig-dallas-rq1-race.
+Howeover, prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+
+@fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
+@fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).
+Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
+The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+
+Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-dallas-rq2-midSAT-race.
+All Geomarkets besides Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex contributed equal or larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Dallas-Fort Woth metropolitan area.
+However, the City of Dallas Geomarket contributed the largest disproportionate share of Hispanic prospects (19% versus 11% for all prospects), whereas the Dallas Counties Geomarket contributed the largest disproportionate share of Black prospects (21% versus 13% for all prospects).
+
+Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses.
+Only 303 prospects of the 3,928 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
+Again, all Geomarkets besides Collin and Rockwall Counties (14% versus 36% for all prospects) and West of Dallas/Fort Worth Metroplex (15% versus 24% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool.
+Disproportioantely larger contributions range from six percentage points by the Dallas Counties Geomarket (19% versus 13% for all prospects) to 11 percentage points by the City of Dallas Geomarket (22% versus 11% for all prospects).
+Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree.
+Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportinate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
+
+Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Dallas-Fort Worth metropolitan area.
+Among all White prospects (n=2,094) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
+@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool.
+The only exception is the Irving, Arlington, & Grand Prairie Geomarket contributing a relatively larger share of White first-generation college students whose parents attended some college but did not complete their degree (21% versus 11% for all prospects).
+
+Similar patterns of one or two Geomarkets disproportionately contributing larger shares of first-generation college students was evident across other racial/ethnic categories in @fig-dallas-rq2-midSAT-combo.
+For instance, City of Dallas (18%) and City of Fort Worth (24%) contributed larger proportions of first-generation college students whose parents did not attend college relative to the overall pool of all prospects (8%).
+On the other hand, the Irving, Arlington, & Grand Prairie (23%) and Dallas Counties (26%) Geomarkets contributed disproportionately larger proportions of Black students whose parents attended some college but did not complete their degree relative to the overall pool of all prospects (17%).
+For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
+City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
+
+Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions.
+As Dallas-Fort Worth became increasingly diverse from 1980 to 2020, a more balanced racial/ethnic distribution emerged across the metropolitan area's Geomarket landscape.
+However, these patterns did not translate to racial/ethnic composition of student prospect pools.
+Suburban Geomarkets like Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex yielded a disproportionately larger share of Asian prospects, whereas central Geomarkets such as the City of Dallas, Dallas County, and Irving, Arlington, and Grand Prairie contributed a disproportionately larger share of Hispanic and Black prospects.
+These racial/ethnic disparities become significantly more pronounced, particularly in the disproportionate share of Asian prospects, as PSAT score thresholds increase (see Appendix @fig-dallas-rq2-lowSAT-race and @fig-dallas-rq2-highSAT-race).
+Similar Geomarket contribution patterns are also evident by first-generation college student status, with Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex providing larger shares of non-first generation college student prospects, and central Geomarkets (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie) providing disproportionately larger shares of first-generation college student prospects.
+Moreover, central Geomarkets are particularly important for including Black (City of Dallas and City of Fort Worth), Hispanic (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie), and Asian (City of Fort Worth and Irving, Arlington, and Grand Prairie) first-generation college students.
+While Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex Geomarkets still contribute a notable overall number of Black, Hispanic, and Asian prospects, such students are predominantly not first-generation college students.
+
 ## Los Angeles
@@
 
-In their analysis of quantifying school quality in England, @RN5005 [p.
-517] observe that "one problem with school league tables \ldots is that the measures of school quality often merely reflect the social origins of those who attend a particular school." Similarly, considering prior research showing that SAT scores are substantially a function of social origin [@RN1273], the Market Segment Model argues that student demand for higher education is mostly a function of social origin.
-
-Institutional theory defines the organizational field as "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services and products" [@RN527, p.
-143].
+The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the enrollment management industry.
+The logic of homophily argues that there is a natural cultural match between students, that colleges should find students who look like students who enrolled at their institution in the past, and that colleges should recruit from the same schools and communities that are popular with peer colleges.
+We suggest that homophily is the operating logic most consistent with college recruiting behavior observed in the wild.
+For example, @RN5023 find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
+The also find great overlap in the schools visited by Villanova and Notre Dame, with a disproportionate focus on visits to Catholic schools.
+When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
+As @RN4975 argues, observed homophily is not the result of voluntary action; rather, homophily is programmed into algorithms that create connections.
+
+The Market Segment Model begins with the status attainment model, that student demand for college is a function of parental education and parental income.
+@RN4982 [chapter 3] demonstrated this argument by showing the correlation between student SAT score-sending behavior and measures of class.
+The thesis of the Market Segment Model is that that households of particular class and college aspirations are likely to live in particular parts of a metropolitan area and that these geographic territories can be meaningfully captured by Geomarkets.
+Although Geomarkets appear coarse by contemporary standards, they were a landmark innovation in Geographic segmentation.
+The Market Segment Model recommends that colleges should identify their core student market segment -- local, in-state, regional, or national -- identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
+This information is contained in the Market Segment Profile (Appendix A), a standard ouput in EPS software.
+When considering new Geomarkets to target, colleges should utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
+This information is contained in the Institutional Profile (Appendix A), also included in EPS software.
+
+@RN4982 state that the Market Segment Model merely formalizes the operating logic of college admissions officers.
+For admissions officers at selective colleges, the logic is that good students from good families are willing to attend good colleges that are far from home.
+These good families are found in particular parts -- Geomarkets -- of metropolitan areas.
+By contrast, students from other parts of the metro possess the disposition and achievement to attend their local college.
+The Market Segment Model formalized this logic and commodified it in EPS software that helped colleges identify desirable Geomarkets and desirable high schools within those Geomarkets.
+@RN5007 suggest that, once commodified and quantified, the local knowledge of admissions officers becomes redundant.
+
+Research question 1 investigates racial and socioeconomic variation between and within Geomarkets over time.
+We find that Geomarkets are highly correlated with race and class.
+\[KARINA SUMMARIZE FINDINGS IN A COUPLE SENTENCES; REVISE PARAGRAPH AS NECESSARY\] We show that in 1980, when Geomarkets were being created, that Black people tended to be highly concentrated in the poorest Geomarket in the metropolitan area.
+Examples include, CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), TX19 -- City of Dallas, CA21 -- South & South Central Los Angeles, and PA5 -- Philadelphia County.
+In a handful of metropolitan areas (e.g., Cleveland), Black people remain conentrated in the poorest Geomarkets.
+In Chicago-Land, \[KARINA -- ONE OR TWO SENTENCES\]
+
+These findings are not surprising given the extent of race and class segregation in the US.
+Furthermore, the correlation between Geomarkets and class is expcted because Geomarket borders were created in service of conceptual model that views demand for higher education as a function of class.
+@RN4982 is written around New England, with a particular focus on Geomarkets in Connecticut and Massachusetts.
+@RN4982 do not discuss the intersection of student demand and race.
+Although unsurprising for early 1980s social science, this silence is conspicuous considering the salience of political controversy surrounding "busing" in Boston and nearby metropolitan areas.
+
+Research question 2 analyzes the racial and socioeconomic composition of student list purchases.
+We examine purchases that include all Geomarkets around a metropolitan area in order to assess the potential consequences if including/excluding particular Geomarkets from the purchase.
+This question is motivated by the @RN4982 recommendation that selective colleges that enroll students from the regional and national market segment should target affluent Geomarkets, while community colleges and non-selective 4-year colleges should focus on middle- and working-class Geomarkets.
+KARINA -- ADD TEXT SUMMARIZING RESULTS.
+
+This article adds a new perspective to existing explanations about how students are sorted into colleges.
+The status attainment model argues that college destination is a function of parental education and occupation [@RN1273; @RN1271; @RN1272].
+The Market Segment Model is the status attainment model applied to geodemographic market research \[CITE\].
+@RN5033 finds that the educational achievement of Asian American students whose parents are immigrants tend to have high educational achievement even if their parents do not, while the educational achievement of later-generation Asian Americans conforms to status attainment theory.
+In our analyses of prospects included in student list purchases, Asian American prospects from poor Geomarkets tended to be first-generation students while Asian American prospects from affluent Geomarkets tended to have parents with a BA.
+Drawing from @RN5033, one explanation for this finding is that Asian students living in low-income Geomarkets tend to have parents who immigrated to the U.S.
+whereas Asian students living in affluent Geomarkets are more likely to have parents who were born in the U.S.
+
+The cultural capital model explains a process by which upper and upper-middle class families sort themselves into selective colleges by providing their children with the pedigree (academic, extracurricular) valued by selective colleges, information about how to navigate the admissions gauntlet, and social networks that provide an inside track [@RN1228; @RN1241; @RN4991; @RN1814].
+This model explains how affluent households maintain a disproportionate enrollment share at selective institutions in an era of holistic admissions ostensibly designed to increase racial and class diversity [@RN4991].
+Both the status attainment model and the cultural capital model are demand-side explanations for why education is a "social sieve" [@RN2377; @RN5046] that allows for a modicum of mobility while maintaining a much larger flow of intergenerational class transmission [@RN861].
+@RN4722 -- who obtained parental income from federal income tax returns for every U.S.
+college -- show that the disproportionate enrollment share of high income families at selective private colleges -- and even at most public research universities -- is staggering.
+These patterns are consistent with the observation by @RN1297 [p.
+241-242] that educational credentials are property that "support their holders'\ldots claims to monopolize socially and economically advantageous positions."
+
+Supply-side explanations of credentialism and more recent scholarship on enrollment management complement the cultural capital model of sorting students to colleges.
+The credentialism literature recognizes that colleges have a financial incentive for educational credentials to determine the competition for socially and economically advantageous positions [@RN865; @RN861; @RN880; @RN869] and this credentials arms race favors affluent households.
+Scholarship on enrollment management describes which students colleges want to enroll and what colleges do to attract these students at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758; @RN3544; @RN3547; @cottom2017lower].
+@RN3519 and @RN4321 describe a tacit arrangement between high school guidance counselors on the demand side and college admissions officers on the supply side.
+Counselors at well-resourced schools, especially private schools, are motivated to give their students a competitive advantage in admissions.
+Meanwhile, college admissions officers are motivated to enroll students who can afford full tuition price and are likely to donate in the future.
+These mutually beneficial desires are consummated by recognzing that upper and upper-middle class applicants satisfy the extracurricular needs of the college.
+The orchestra needs oboists.
+The lacrosse team needs players.
+Quantitative analyses of visits from colleges to high schools are consistent with @RN3519; selective private colleges and public research universities devote most of their recruiting resources on courting students from privileged schools and communities [@RN5023; @RN4758].
+However, these analyses conceive of recruiting as something that is done by individual colleges.
+
+Existing explanations of how students are sorted into colleges miss an important supply-side mechanism, third-party vendors that sort students on behalf of colleges.
+@RN4982 created the Market Segment Model and Geomarkets based on a snapshot of student demand from 1980.
+@RN4982 argues that demand for higher education is correlated with class, ignoring the historical structural barriers that produced class- and race-based inequality in 1980 student demand.
+In itself, the Market Segment Model is an unremarkable social science depiction.
+However, the College Board incscribed Geomarkets and the Market Segment Model into EPS software that told colleges which Geomarkets and high schools to target.
+Technologies that target target customers based on geography inevitably leverage racial and class segregation [@RN4775; @RN4849; @RN4975].
+Furthermore, technologies that use the logic of predictive analytics -- identify correlates in past cases to make predictions for future cases -- are prone to "pernicious feedback loops" [@RN4849].
+The structural barriers (e.g., segregation, slavery, Jim Crow) that caused historic place-based inequality in student demand are amplified because EPS recommends that colleges focus recruiting resources on localities that already have high student demand for peer colleges.
+Geomarkets were subsequently incorporated into the Student Search Service product.
+We show that that excluding low-income, non-white Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-white students with strong test scores.
+
+## Enrollment Management Industry
+
+The growing salience of third-party vendors warrants a reconfiguration of the organizational field salient to college access.
+@RN527 \[p.
+148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.College access scholarship from economics often analyzes the effects of district/state/federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
+These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^14]
+
+[^14]:  @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
+
+@list_biz describes four key dynamics in the enrollment management (EM) industry, with a focus on the market for student list data.
+First, EM consulting firms are central to the creation and implementation of recruiting campaigns [@RN5027].
+The top two EM consultancies are Ruffalo Noel Levitz, which claims to serve 1,900 colleges annually [@rnlabout_2025], and EAB, which claims to serve 2,100 colleges [@about_eab].
+College reliance on EM consultancies is partially explained by the high level of burnout and employee turnover in the EM profession [@hoover_em].
+When we issued records requests to public universities about their student list purchases, at least 50% of universities indicated that they outsourced student list purchases to an EM consultancy.
+In many cases, the university was buying names each year but no university employee had knowledge about which vendors they were buying from or which search filters they were utilizing.
+
+Second, technological advances creates new means of identifying and serving prospects.
+College Board began selling the contact information of test-takers in 1972 [@belkin2019-studata].
+ACT followed suit and the two testing companies enjoyed a near duopoly for several decades.
+This business can be described as list-based lead generation based on the direct-mail model.
+"Free" college search engines (e.g., Cappex, Niche) yielded news sources of student list data.
+Another data source is college planning software purchased by high school districts and utilized by high school students and guidance counselors.
+The most widely-known product is Naviance, which claims to be used by more than used by more than 10 million K-12 students and by 40% ofUS high schools [@powerschool].
+*Naviance* college planning software feeds into the *Intersect* recruiting platform, which allows colleges to target Naviance users while they are on the platform.
+*Intersect* is an example of behavioral-based targeting.
+
+The third dynamic in the EM industry has been the growth of test-optional movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT.
+ACT responded to uncertainty in their core testing business by attempting to become an edtech company \[CITE\], failing \[\], and being acquired by Nexus Capital in 2024 \[CITE\].
+Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters \[CITE\] and ACT's predicted probability of enrollment filter \[CITE\] -- developed free college search engines, and entered the enrollment management consulting market more aggessively.
+College Board has weathered the test-optional movement thanks to robust revenues from the Advanced Placement product, but the decline in PSAT/SAT/PreACT/ACT test-takers has undermined their oligopoly in the market for student list data.
+
+The fourth dynamic is the transformation from owner-operated to private-equity owned firms and erosion of the distinction between consultancy and software vendor.
+The 1980s and 1990s was a period of market entry in EM, when college admissions professionals or professors decided to hang up a shingle [@RN5027].
+The 2010s were a period of acquisitions and concentration.
+After regulatory scrutiny pushed private-equity interests out of for-profit college market [@RN5022], private equity investors find value in acquiring firms that provided services to direct providers.
+Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions to grow market share and vertical acquisitions designed to make customers more dependent on particular recruiting products.
+
+EAB exemplifies the the trend towards private equity ownership and industry concentration.
+EAB is the preeminent EM consultancy, offering software-as-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
+The origins of EAB trace to Bill Royall, who founded Royall & Company in 1983 to provide direct marketing and fundraising for Republican political campaigns [@jump_2020].
+Royall did not sign its first colege client for several years, but EM became the primary focus by 1995.
+Royall was acquired by the health tech firm Advisory Board Company for \$850 million in 2015 and then by Vista Equity Partners for \$1.5 billion in 2017, renaming it EAB [@drew_hansen_2017].
+
+As an owner-operated consultancy, Royall was the market leader in purchasing student lists from student list vendors on behalf of clients.
+Under Vista, EAB entered the market for student list data by completing a series of acquisitions that targeted the "leads" and "inquiries" stages of the enrollment funnel.
+In 2019, EAB acquired YouVisit, a leading provider of virtual tours.
+YouVisit is an engine for "inquiries" in that taking a virtual tour shows interest in a college and leaves a trail of contact information for subsequent recruiting interventions.
+In 2020, EAB acquired Cappex, a leading free college search engine, which is an engine for leads in that students provide contact details and other information.
+In 2021, EAB and PowerSchool -- the leading provider of K12 student information systems and a subsidiary of Vista -- jointly acquired the EM consultancy/software vendor Hobsons.
+EAB acquired Hobson's student success software, Starfish, while PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect.
+EAB paid PowerSchool to become the exlusive reseller of Intersect.
+Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.
+EAB combined these acquisitions into create [Enroll360](https://eab.com/solutions/enroll360/):
+
+> We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
+This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]
+
+Enroll360 exemplifies a new kind of student list product that leverages proprietary data to sell software.
+Historically, College Board and ACT sold prospect contact information at a price-per-prospect (e.g., \$0.50).
+The goal was to generate revenue from the sale of names.
+By contrast, Enroll360 wraps several proprietary databases of names (e.g., Cappex, Intersect) into a software-as-service platform that recruits those prospects at different stages of the enrollment funnel.
+Here, the goal of lists is to create demand for software, which is more expensive than names.
+Colleges that are interested in recruiting these prospects must buy an Enroll360 subscription.
+Like College Board *Student Search Service* contemporary software-as-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
+For example, reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
+Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and wheather students used Naviance to "research competitor institutions" [@markup_naviance].
+A procurement document from the University of Utah explains that the Intersect subscription is necessary to recruit the proprietary database of Naviance users: [@utah_procurement]:
+
+> The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to.
+There is unique group of prospective students who are only in the PowerSchool Naviance platform.
+We have spoken with several vendors where we can reach out to prospective students.
+Each vendor has their own proprietary website or database where select students can be reached.
+
+An under-studied and under-regulated third-party EM industry has negative consequences for colleges and for students.
+Accross multiple student list products, we see a growing number of filters that recommend which prospects to target based on the beheavior of previous cohorts of students.
+The ACT/Encoura "Enrollment Predictor" filter enables colleges to filter prospects based on the predicted probability of enrollment, which is created based on analyzing the enrollment decisions of previous cohorts of students [@schmidt_2019].
+College Board Geodemographic Segment filters categorizes each Census Tract into one of about two dozen clusters based on the college-going behavior of previous cohorts [@RN4565] and is based on the Claritas/PRIZM market segmentation system that categorized zip codes into groups useful for merchants [@RN4957].
+An investor-facing @vista_eab_ml video promoting EAB's "Pipeline Analytics" product states, "we are using artificial intelligence and machine learning to help schools build list of students that may be more more likely to enroll and persist in their university." The product "identifies students that would be a good match for a school based on their similarity to students who have succeeded there before and then provides them \[the students\] information about those institutions so that they choose to apply and ultimately enroll." Each of these products are based on the underlying logic of correlations and homophily discussed by @RN4975 and applied by @RN4982.
+These products identify correlations in a historic snapshot of students and apply those relationships to make recommendations about which prospects to target in the present.
+As such, all of these technologies amplify the effect of previous inequality in student demand, which is itself a consequence of structural inequality in educational opportunity.
+
+Market dynamics in the EM industry raise policy concerns about tuition prices, student choice, and competition in the market for EM consultancies may also affect competition, prices, and student choice.
+To the extent that colleges must buy expensive software to recruit prospects contained in an associated database, this expenditure will ultimately be passed on to students in the form of higher prices or reduced expenditure on education.
+Unfettered horizontal and vertical acquisitions in the EM industry causes the EM industry to become more concentrated.
+Smaller consultancies that do not control veritial inputs may be unable to compete with companies that have acquired proprietary pools of prospects.
+Fewer EM consultancies means less competition and higher consulting fees, which will be passed on to consumers.
+Finally, the blurring of the lines between consultancy and student list vendor suggests that large consultancies will funnel prospects only to those colleges that are willing and able to pay for subscriptions, which raises policy concerns about student choice.
+
+## Developing a Knowledge Infrastructure
+
+Regulation of the higher education industry should be informed by research that investigates the "totality of relevant actors" [@RN527, p.
+148] in the organizational field.
+Unfortunatel, scholarship salient to education policy and regulation have ignored two, related economic transformations.
+First, third-party vendors offer software platforms that perform core tasks of organizations, including schools and colleges [@list_biz].
+Second, private firms comprise a growing share of economic activity [@RN5037; @RN5038].
+Most firms in EM and the broader edtech sector are private.
+However, private firms face fewer disclosure requirements than publicly traded firms.
+The paucity of research about third-party vendors that structure college access is substantially due to difficulty obtaining data for empirical analyses.
+The research community must figure out how to systematically investigate the role of private firms and investors in education, so that we can hold private interests accountable for education policy goals.
+
+Knowledge infrastructures are "systems of observation and measurement" [@RN5025, p.
+743] that provide the foundation for scalable research.
+They collect, process, and distribute data in ways that "enable certain kinds of knowledge production while simultaneously channeling researchers away from questions not readily answerable within\ldots that infrastructure" [@RN5025, p.
+742].
+The knowledge infrastructure of education research largely consists of data about students, schools, and colleges.
+It has been shaped by the ascendance of economics [@RN5036], which incorporates district, state, and federal administrative data to evaluate the effects of schools and policies on student outcomes [@RN5043].
+The research community mobilized to investigate for-profit colleges because for-profit colleges and students were included in ongoing data collection [@RN5022].
+By contrast, researchers ignored the growth of private, for-profit, third-party vendors (e.g., PowerSchool, Parchment) because these organizations are excluded from data collections known to education researchers.
+
+We argue that sociologists and adjacent scholars should expand the domain of the education research knowledge infrastructure to include third-party vendors in enrollment management, and edtech more broadly.
+This is a difficult task.
+Student unit record data owned by district, state, and federal agencies is simultaneously granular -- can evaluate a particular policy -- and macro -- can speak to broad trends in the education sector.
+By contrast, obtaining granular data about third-party products in education is difficult because these data are owned by private interests that do not want researchers to interrogate their products [@RN4774; @RN4842].
+We see two broad avenues for research: case studies involving primary data collection -- including this study -- which tend to have a granular focus on particular products or vendors; and secondary data analysis, which may have a more macro focus.
+
+Although case study research often relies on interview and ethnographic data, another approach to data collection is to issue records requests to public entities that contract with third-party entities.
+For example, @RN5039 analyze public university contracts with online program managers (OPMs).
+The present manuscript is an example of case study research.
+We issued public records requests to obtain tabular data about student list purchases.
+However, this process was labor intensive, requiring the pro bono efforts of several law firms.
+Public records requests seem better suited to obtaining contracts.
+Qualitative or computational text analysis of publicly available websites and social media provides another avenue for data collection.
+Here, data collection is efficient but data processing is often labor intensive.
+Like investigative reporting [@markup_naviance], case study research can produce granular analyses of particular products (e.g., Intersect, Enroll360) that yield insight into the underlying mechanisms of how third-party vendors structure college access.
+However, data collected for case study research typically does not meet the knowledge infrastructure standard of repeated, ongoing data collection that can be the foundation for scalable research.
+
+We see subscription databases about private markets as having the potential to create a knowledge infrastructure for scalable research on the role of private firms and private equity funders in education, and also their role in other domains of interest to sociology.
+A growing set of "market intelligence" data providers serve the information needs of the investment community by providing data on private firms and deals involving private firms.
+These databases tend to include three broad "types" of data: (1) *firm-level data* (e.g., location, employees, financials); (2) *acquisitions*, encompassing who acquired whom and resulting ownership structure); and (3) *investments*, which include amounts/valuations/stake percentages of private equity investments.
+Some providers provide all three types of data (e.g., [*S&P Capital IQ*](https://www.spglobal.com/market-intelligence/en), [*LSEG Workspace*](https://www.lseg.com/en)).
+Other providers specialize.
+For example, [*PitchBook*](https://pitchbook.com/) is a leading provider of data about investments in private companies by private equity firms, while [*Orbis*](https://login.bvdinfo.com/R1/Orbis) is a leading provider of firm-level data for non-U.S.
+private firms.
+These databases are often utilized in journals of finance or management that consider private equity markets or patterns of acquisitions [e.g., @RN5044; @RN5038].
+Interestingly, the database subscriptions operate similar to contemporary student list products; an interactive, user-facing platform is wrapped around several sources of proprietary data.
+University libraries provides full or partial access to some subscriptions, while others must be purchased.
+
+Although scholars cannot share subscription data, they can share code to process and analyze these data.
+We recommend that scholars develop panel datasets examining change over time in the firms operating in particular industries and acquisitions and private equity investments in those industries.
+For example, empirical scholarship can examine change over time in concentration in the enrolment management industry and its transformation from an owner-operated to a largely private-equity owned industry.
+Subsequent analyses could investigate the portfolio of large owners, their ties to other industries -- such as the direct-provider for-profit college market -- and the vertical inputs/technologies they are investing in.
 
@@
 \clearpage
-
-<!-- TABLES -->
-
-```{=tex}
-\appendix
-\renewcommand{\thetable}{A\arabic{table}}
-```
-# Appendix A
-
-@tbl-market-segment-characteristics reproduces a partial, simplified version of the bottom panel of @RN4982 [Figure 2.1] which represents the Market Segment Profile for "Connecticut Market 3: Fairfield County." @tbl-market-segment-characteristics has separate columns for each market segment (local, in-state, regional, national) and rows show the number of test-takers and characteristics of test-takers.
-For example, there were 550 "local" students in Fairfield county and these students submitted SAT scores to 2.8 postsecondary institutions on average.
-By contrast, there were 1,664 "regional" students who submitted test scores to 4.8 institutions on average.
-For local students, 11.2% had family income greater than \$35,000 and 9% had both parents with a BA.
-For regional students, 41.9% had family income greater than \$35,000 and 34.0% had both parents with a BA.
-Each Market Segment Profile also present information about the institutions that students from each segment sent scores to.
-
-@tbl-market-segment-institutions recreates the top panel of @RN4982 [Figure 2.1] for the Fairfield County, CT local market.
-For example, of the 550 local students, 277 (50.4%) sent scores to institution #1, which was a private master's granting institution.
-Of the 1,199 in-state students, 757 (63.1) sent scores to institution #1, a public doctoral granting institution, and 515 (43.0%) sent scores to institution #2, a public master's granting institution.
-
-The *Institutional Profile* describes students who send scores to a particular institution and which majors these students are interested in.
-For each institution, institutional profiles are created separately for students from a particular local market, for all students in a state, or all students in a region.
-@tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the institutional profile of an anonymous institution for students from Fairfield County, CT.
-@tbl-sample-institutional-profile shows that 58 in-state students submitted SAT scores to the institution.
-These 58 students represented 4.8% of the total 1,199 in-state students from Fairfield County.
-69 regional students sent scores to the institution, representing 4.1% of all 1,664 regional test-takers.
-Of these 69 regional students, 35 expressed interest in majoring in the liberal arts.
-These 35 students represent 7.0% of all regional students from Fairfield County who expressed interest in the liberal arts.
-
-@RN4982 argue that the Institutional Profile and the Market Segment Profile enable admissions officers to know where to look for students and which institutions are competing for those students (p.
-25):
-
-> The Institutional Profile and the Market Segment Profile quantify the admission officers' intuitive grasp of market structure.
-Structure here carries a dual meaning, connoting both the structure of student choice and the structure of institutional competition\ldots This two-sided interpretation furnishes the essential framework for planning by individual colleges and universities\ldots To draw effectively on its own natural constituency, a college not only must contact the "right" kind of students --- that is, students who are predisposed toward that type of institution -- but also must persuade them of its special character.
-This means knowing the competition as well as the clientele.
-
-```{r market-segment-characteristics, echo = F}
-#| label: tbl-market-segment-characteristics
-#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'
-
-market_segment <- data.frame(
-  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
-  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
-  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
-  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
-  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
-)
-
-kable(
-  x = market_segment,
-  col.names = c('Characteristic','Local','In-state','Regional','National'),
-  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```
-
-\clearpage
-
-```{r market-segment-institutions, echo = F}
-#| label: tbl-market-segment-institutions
-#| tbl-cap: 'Top 5 institutions in terms of number of scores sent by segment, Connecticut Market 3: Fairfield County'
-
-market_segment_inst <- data.frame(
-  rank = c(1,2,3,4,5),
-  local_num_scores_sent = c(277,261,183,103,100),
-  local_pct_scores_sent = c(50.4,47.5,33.3,18.7,18.2),
-  local_inst_type = c('priv ma','priv ma','priv ma','pub doct','pub ma'),
-  state_num_scores_sent = c(757,515,438,183,177),
-  state_pct_scores_sent = c(63.1,43.0,36.5,15.3,14.8),
-  state_inst_type = c('pub doct','pub ma','pub ma','pub ma','pub ma'),
-  regional_num_scores_sent = c(610,348,272,248,197),
-  regional_pct_scores_sent = c(36.7,20.9,16.3,14.9,11.8),
-  regional_inst_type = c('pub doct','priv doct','priv doct','pub doct','pub doct'),
-  national_num_scores_sent = c(1226,371,327,312,308),
-  national_pct_scores_sent = c(32.6,9.9,8.7,8.3,8.2),
-  national_inst_type = c('pub doc','priv doct','priv res','priv doct','priv doct')
-)
-
-kable(
-  x = market_segment_inst,
-  col.names = c('','Num','Pct', 'Type','Num','Pct','Type','Num','Pct','Type','Num','Pct','Type'),
-  #caption = 'Top 5 institutions in terms of number of scores sent, by segment \\\\Connecticut Market 3: Fairfield County'
-) %>% 
-  #kable_styling(full_width = F) %>% 
-  add_header_above(c(' '=1,'Local (N=550)' = 3, 'In-state (N=1,199)' = 3,'Regional (N=1,664)' = 3, 'National (N=3,766)' = 3))
-```
-
-\clearpage
-
-```{r sample-institutional-profile, echo = F}
-#| label: tbl-sample-institutional-profile
-#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'
-
-sample_inst <- data.frame(
-  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
-  local = c(1,0.2,0,0.0,0,0.0),
-  in_state = c(58,4.8,25,8.4,2,3.0),
-  regional= c(69,4.1,35,7.0,5,5.8),
-  national = c(109,2.9,61,5.3,3,0.8),
-  total = c(237,3.3,121,5.9,10,1.8)
-)
-
-kable(
-  x = sample_inst,
-  col.names = c('','Local','In-state','Regional','National','Total'),
-  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```

### 2025-05-07 text/manuscript-full-KS.qmd
full manuscriptcheck

@@
 In the contemporary era of private equity funded platform capitalism, third-party recruiting products have become increasingly powerful, granular, and ubiquitous.
-However, we argue that a watershed moment in their develeopment was the creation of Geomarkets by the College Board and UPenn professor, Robert Zemsky.
+However, we argue that a watershed moment in their development was the creation of Geomarkets.
+
+<!--#  by the College Board and UPenn professor, Robert Zemsky. -->
 
@@
 x].
-The broader project was conceptualized as an effort to to "capture and quantify" (p.
-11) the knowledge of admissions officers, based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel" (p.
-11).
-@RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of 1980 high school seniors.
+Conceptualized as an effort to to "capture and quantify" (p.
+11) the knowledge of admissions officers, @RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of high school seniors in 1980.
 The thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography.
@@
 Geomarkets are geographic borders meant to define local recruiting markets, a territory of an admissions recruiter.
-Figure X, panel A shows Geomarkets in the Chicago-land area.
+For instance, Figure X panel A shows Geomarkets in the Chicago-land area.
 As a planning tool for colleges, the Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
+
+<!--# based on the idea that "a good recruiter knows where to look for prospective applicants, as seen in the students' willingness or reluctance to travel" (p. 11). -->
 
@@
 Anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
-Figure X panel B shows that Chicago-land Geomarkets define recruiting territories for University of Chicago admissions recruiters (Appendix Figures BLAH BLAH).
-Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), which was in 1984 and remains active today.
+Figure X panel B shows that Chicago Geomarkets define recruiting territories for University of Chicago admissions recruiters (Appendix Figures BLAH BLAH).
+Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS).
 EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.
+
+<!--# , which was in 1984 and remains active today.
 According to @RN5021, in 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis product, which was based on EPS.
+-->
 
@@
 @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
-The Student Search Service database consists of College Board test-takers.
-Colleges pay a fee for each prospect (e.g., \$0.50 in 2021).
 Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc.
-Geomarket filters enable colleges to include/exclude prospects from particular Geomarkets.
+Geomarket filters enable colleges to include or exclude prospects from particular Geomarkets.
+
+<!--# The Student Search Service database consists of College Board test-takers. Colleges pay a fee for each prospect (e.g., $0.50 in 2021).  -->
 
 Scholarship on enrollment management has focused on college behavior [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
-Scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access.
+However, scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access.
 These mechanisms, which lie upstream of the behavior of individual colleges, are incorporated by individual colleges in ways that funnel certain kinds of students to certain kinds of institutions.
-This omission is surprising because sociology is concerned with sorting [@RN4873], particularly sorting into college.
+This omission is surprising given sociology is concerned with social sorting [@RN4873], particularly sorting into college.
 
@@
 Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS and within the Student Search Service student list product.
-First, what is the socioeconomic and racial variation between geomarkets in metropolitan areas and how does this variation change over time?
-We address this question by spatially joining the geomarket shapefile to Census data about socioeconomic and racial characteristics.
-Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular geomarkets?
+First, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas and how does this variation change over time?
+We address this question by spatially joining Geomarket shapefiles to Census data about socioeconomic and racial characteristics.
+Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular Geomarkets?
 This research question is motivated by @RN4982, which recommends that selective colleges target affluent Geomarkets.
@@
 We analyze this research question using data from actual student list purchases, which were collected using public records requests.
-We analyze student list purchases that include all Geomarkets in a metropolitan area in only to assess which prospects would be included/excluded by had the student list purchase filtered on particular Geomarkets.
-
-The following section introduces background and scholarship about enrollment management and introduces salient concepts from scholarship on quantification, particularly correlation and homophily.
+We analyze student list purchases that include all Geomarkets in a metropolitan area in order to simulate which prospects would be included or excluded had the student list purchase filtered on particular Geomarkets.
+
+The following section introduces background and scholarship about enrollment management and introduces salient concepts from scholarship on quantification.
 Second, we introduce the case, describing how @RN4982 developed Geomarkets and the Market Segment Model and how these concepts were incorporated into College Board recruiting products.
@@
 Fourth, we present results.
-[We find that\ldots ?PARAGRAPH ON RESULTS?] Finally, we discuss implications for scholarship and for policy.
+\[We find that\ldots ?PARAGRAPH ON RESULTS?\]
+
+Finally, we discuss implications for scholarship and for policy.
 The sociology of education should move beyond the fixation on schools and policies as the primary sorting machines.
-@RN527 [p.
-148] remind us that the organizational field includes the totality of relevant actors." Third-party vendors have structured college access for decades and their influence is growing.
-The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product \[CITE\].
+@RN527 \[p.
+148\] remind us that the organizational field includes the totality of relevant actors." Third-party vendors have structured college access for decades and their influence is growing.
+The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [CITE].
 By investigating high-leverage actors and products in the organizational field, sociology can develop important insights that shape future federal and state regulatory policy.
@@
 The Student Search Service -- created in 1972 and copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
-Second, the Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
+The Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
 Both products are applications of quantification.
-Student Search Service includes or excludes prospects from a student list purchase based on value of selected variables while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
+Student Search Service includes or excludes prospects from a student list purchase based on value of selected filters like Geomarket while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
 The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on equality of opportunity.
@@
 Quantification changes decision-making power *authority*, often weakening the discretion of local decision-makers.
-Scholarship about U.S.
-News & World Report (USNWR) Law School rankings demonstrate the effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
+For instance, scholarship about U.S.
+News & World Report (USNWR) law school rankings demonstrate the effects of quantification [@RN5003; @RN2356; @RN2376; @RN2375].
 Law school rankings were developed as a means of informing prospective students about the relative quality of different programs.
@@
 For example, characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
-Rankings also weakened the *authority* of admissions personnel to make independent decisions class size and which applicants to admit [@RN5003].
+Rankings also weakened the *authority* of admissions personnel to make independent decisions about class size and which applicants to admit [@RN5003].
 <!--
@@
 
-## Dallas-Fort Worth
-
-### Suburban Geomarket Growth Eases Racial Disparities, Concentrated Wealth Persists
-
-The six Geomarkets within the Dallas-Fort Worth metropolitan area are illustrated in @fig-dallas-maps by total population across the three census periods.
-The Dallas-Fort Worth region represents a unique "metroplex" case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic urban core and suburban ring model showcased by Chicago.
-As depicted in @fig-dallas-maps, two distinct inner-core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
-These central Geomarkets are surrounded by the more conventional outer ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
-Worth Metroplex).
-
-From 1980 to 2020, the population of the overall Dallas-Fort Worth metropolitan area increased substantially from approximately 2.7 Million to 6.8 Million.
-@fig-dallas-maps showcases all six Geomarkets experienced population growth over the three census periods.
-However, the most pronounced increases occured in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
-These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers - City of Forth Worth and City of Dallas - into the surrounding suburban and rural areas of the metropolitan areas.
-This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket from 1980 to 2000.
-
-@fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
-In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
-By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
-In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
-Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.
-
-These shifts in racial and ethnic composition within and across Geomarkets became more pronounced between 2000 and 2020.
-The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents.
-However, both Geomarkets remained predominantly White (58% and 63%, respectively).
-In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian Pacific Islander), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian Pacific Islander).
-Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
-
-Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
-Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
-Between 1980 to 2000, affluence levels across Geomarkets were consistent, ranging from a low of \$82,000 in City of Fort Worth to a high of \$135,00 in the Collin & Rockwall Counties Geomarket.
-However, the substantial increase in median income for the Collin & Rockwall Counties Geomarket during this time period plateaued by 2020, reshuffling relatively affluence levels among Geomarkets.
-By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000).
-In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable and continued to rank at the lower levels of affluence within the metropolitan area.
-\[KS: NEED TO CHECK THIS; I THINK CITY OF DALLAS AND CITY OF FORT WORTH ARE MIXED UP BETWEEN GRAPH VERSUS TABLE\]
-
-@fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
-In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets.
-Overtime, poverty rates declined in Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex, while increasing for all other Geomarkets.
-These divergent trends resulted in a widening gap in poverty levels across the Dallas-Fort Worth metropolitan area.
-A similar widening disparity is evident in educational attainment.
-Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
-
-The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s.
-While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
-Such growth is largely attributed to increases in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex.
-On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment across Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
-Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
-
-### Composition of Students Lists by Geomarkets
-
-We also analyze student list orders collected through public requests to assess how the racial/ethnic and socioeconomic composition of included versus excluded prospects varies when filtering on particular Geomarkets within the Dallas-Fort Worth metropolitan area.
-Specifically, we analyze three list orders placed by a research university targeting prospects across 87 different Geomarkets nationwide, including all six Geomarkets in the Dallas-Fort Worth Metropolitan area.
-Each order also filtered for prospects in the 2019-2020 graduating class and based on PSAT scores across three ranges: 1070-1180, 1190-1260, and 1270-1520.
-Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260) from the Dallas-Fort Worth metropolitan area to simulate Geomarket contributions and assess patterns of inclusion versus exclusion.
-
-@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area.
-At the lowest PSAT score thresholds, notable differences emerge in the racial/ethnic composition of prospects within Geomarkets when compared to the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
-For example, the largest number of prospects in the 1070-1180 PSAT range were from the Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian), both of which yield predominantly White prospect pools.
-On the other hand, the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarket pools have the greatest proportions of non-White prospects (although yielding fewer overall students).
-Interestingly, while the City of Dallas is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-dallas-rq2-race includes larger proportion of Hispanic (41%) prospects but lower proportions of Asian (4%) and Black (9%) students comparatively.
-
-These racial/ethnic composition patterns shift as PSAT scores increase.
-Shares of White prospects within Geomarkets remain relatively proportional to the overall 2020 metropolitan distribution in @fig-dallas-rq1-race.
-Howeover, prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
-
-@fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
-@fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).
-Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
-The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
-
-Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-dallas-rq2-midSAT-race.
-All Geomarkets besides Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex contributed equal or larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Dallas-Fort Woth metropolitan area.
-However, the City of Dallas Geomarket contributed the largest disproportionate share of Hispanic prospects (19% versus 11% for all prospects), whereas the Dallas Counties Geomarket contributed the largest disproportionate share of Black prospects (21% versus 13% for all prospects).
-
-Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses.
-Only 303 prospects of the 3,928 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
-Again, all Geomarkets besides Collin and Rockwall Counties (14% versus 36% for all prospects) and West of Dallas/Fort Worth Metroplex (15% versus 24% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool.
-Disproportioantely larger contributions range from six percentage points by the Dallas Counties Geomarket (19% versus 13% for all prospects) to 11 percentage points by the City of Dallas Geomarket (22% versus 11% for all prospects).
-Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree.
-Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportinate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
-
-Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Dallas-Fort Worth metropolitan area.
-Among all White prospects (n=2,094) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
-@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool.
-The only exception is the Irving, Arlington, & Grand Prairie Geomarket contributing a relatively larger share of White first-generation college students whose parents attended some college but did not complete their degree (21% versus 11% for all prospects).
-
-Similar patterns of one or two Geomarkets disproportionately contributing larger shares of first-generation college students was evident across other racial/ethnic categories in @fig-dallas-rq2-midSAT-combo.
-For instance, City of Dallas (18%) and City of Fort Worth (24%) contributed larger proportions of first-generation college students whose parents did not attend college relative to the overall pool of all prospects (8%).
-On the other hand, the Irving, Arlington, & Grand Prairie (23%) and Dallas Counties (26%) Geomarkets contributed disproportionately larger proportions of Black students whose parents attended some college but did not complete their degree relative to the overall pool of all prospects (17%).
-For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
-City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
-
-Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions.
-As Dallas-Fort Worth became increasingly diverse from 1980 to 2020, a more balanced racial/ethnic distribution emerged across the metropolitan area's Geomarket landscape.
-However, these patterns did not translate to racial/ethnic composition of student prospect pools.
-Suburban Geomarkets like Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex yielded a disproportionately larger share of Asian prospects, whereas central Geomarkets such as the City of Dallas, Dallas County, and Irving, Arlington, and Grand Prairie contributed a disproportionately larger share of Hispanic and Black prospects.
-These racial/ethnic disparities become significantly more pronounced, particularly in the disproportionate share of Asian prospects, as PSAT score thresholds increase (see Appendix @fig-dallas-rq2-lowSAT-race and @fig-dallas-rq2-highSAT-race).
-Similar Geomarket contribution patterns are also evident by first-generation college student status, with Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex providing larger shares of non-first generation college student prospects, and central Geomarkets (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie) providing disproportionately larger shares of first-generation college student prospects.
-Moreover, central Geomarkets are particularly important for including Black (City of Dallas and City of Fort Worth), Hispanic (City of Dallas, City of Fort Worth, Dallas County, and Irving, Arlington, and Grand Prairie), and Asian (City of Fort Worth and Irving, Arlington, and Grand Prairie) first-generation college students.
-While Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex Geomarkets still contribute a notable overall number of Black, Hispanic, and Asian prospects, such students are predominantly not first-generation college students.
+# Data and Methods
+
+## Data and Variables
+
+**RQ1**.
+Our study utilizes two main data sources based on each research question.
+The first research question is focused on understanding the socioeconomic and racial variation between Geomarkets in metropolitan areas and how such variation changes over time.
+To answer this question, we use census tract data from the U.S.
+Census Bureau.
+Utilizing a small geographic area is important for implementing a spatial merge between Census data and Geomarket shape files (described below).
+Census tract is the smallest geographic area for which measures of both race/ethnicity and socioeconomic characteristics are available.
+Census tracts are statistical subdivisions of a county designed to capture relatively homogeneous demographic communities that are generally considered to be more accurate for population analyses than zip codes, which are spatial boundaries drawn to demarcate route efficiency for postal services \[[CITE](https://www.census.gov/newsroom/blogs/random-samplings/2014/07/understanding-geographic-relationships-counties-places-tracts-and-more.html)\] .
+
+For the first research question, we use data from the 1980 and 2000 Decennial Census as well as 5-year estimates from the 2020 American Community Survey (ACS).
+For the 1980 Decennial Census, we specifically draw on Summary Tape File 1 (STF1) for variables about race and Hispanic origin collected via “short form” questions answered by all households and Summary Tape File 3 (STF3) for variables about socioeconomic characteristics (income, education, and poverty) collected via the "long form" questionnaire completed by a sample of households.
+The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
+To examine how Geomarket characteristics change over time, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty from the 2000 Decennial Census.[^14] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+These questions were replaced by ACS.
+Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.[^15] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+
+[^14]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
+
+[^15]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
+
+Across these data sources, we created measures of race and ethnicity, specifically the number and proportion of the census-tract population in the following categories: White, non-Hispanic; Black, non-Hispanic; Asian, non-Hispanic; Pacific Islander, non-Hispanic; American Indian and Alaskan Native, non-Hispanic; and two or more races, non-Hispanic.
+Measures of Asian, American Indian and Alaskan Native, two or more races, and Native Hawaiian and Pacific Islander (all non-Hispanic) were not available for the 1980 Decennial Census.
+We also created socioeconomic measures that approximate the predictors of market segment identified by @RN4982.
+These variables are median household income, percent of households below the poverty line, and people age 25+ with a BA.
+All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
+
+We assigned each census tract to College Board Geomarkets by implementing a partial spatial merge.
+Spatial manipulations utilized functions from the *sf* R package.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^16] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Using these zip code-level data, we utilized the `aggregate()` function to create shapefiles for each Geomarket.
+
+[^16]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
+
+Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
+Census tract shapefiles that intersected with multiple Geomarkets (i.e., a census tract's spatial boundaries were spread across two or more Geomarkets) broken into smaller shapefiles wholly contained within a single Geomarket.
+The variables in these shapefiles were then replaced with the original variable value multiplied by the proportion of land area of the census tract that was contained within the Geomarket.
+For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 1,000 people who identify as Hispanic non-white.
+The partial spatial merge splits this census tract into two observations.
+The observation assigned to Geomarket A reports $.60*1000 = 600$ Hispanic non-white people and the observation assigned to Geomarket B reports $.40*1000 = 400$ Hispanic non-white people.
+
+**RQ2**.
+Our study's second research question asks how does the racial and socioeconomic composition of included versus included prospects vary when student list purchases filter on Geomarkets.
+We answer this question using actual student lists purchased by public universities.
+In February 2020, we began issuing public records requests to public universities about student lists purchased from 2016 through 2020.
+We narrowed the scope of our request to student lists purchased from College Board, ACT, and the National Research Center for College and University Admissions (NRCCUA), the three largest student list vendors at the time.
+For each student list purchased from 2016 through 2020, we requested two related pieces of data: (1) the order summary, which specifies search criteria for the student list purchase; and (2) the de-identified prospect-level list produced from the search criteria.
+
+The data collection sample for this project was all public universities in California, Illinois, Minnesota, and Texas.
+Utilizing public records requests to obtain public records is a painstaking process.
+Initially, the majority of universities did not respond or denied our request.
+Subsequently, we obtained pro bono representation from four law firms, which substantially increased the success of data collection.
+We also collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
+However, we were unable to obtain legal representation for Texas.
+
+Even with firm representation, data collection remained difficult.
+Some universities provided records that were not usable for quantitative analyses (e.g., summary statistics across multiple orders; or data did not contain important fields).
+Some universities did not provide records based on legitimate legal grounds (e.g., data not in university possession; not required to create records that do not currently exist).
+We learned that many universities outsourced student list purchases to third-party consulting firms.
+Unfortunately, we were rarely able to obtain usable data from these universities.
+A small number of universities denied requests based on potentially questionable legal rationale, but we lacked the resources to litigate.
+
+The primary data source for our second research question is therefore student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
+This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Minnesota, and Arizona \[KS Does this still apply?\].
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^17] The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
+In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.
+Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
+
+[^17]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
+
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter, but did purchase prospects from all Geomarkets in a particular metropolitan area.
+For each prospect we know their home zip code and, therefore, which Geomarket they belong to.
+We therefore simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
+
+The variables of interest for student list data are the characteristics of prospects who would have been included versus excluded by our simulations by Geomarket filter.
+Prospect characteristics are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
+College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity nonresponse) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016).
+From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016).[^18] Information about student socioeconomic status is limited.
+However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
+
+[^18]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
+Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”
+
+## Methods
+
+**Research design**.
+We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
+Case studies are used to gain a deeper understanding of the phenomenon of interest by comparing and contrasting across different settings \[CITE YIN 2001\].
+Our study's research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
+Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics.
+Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts.
+We therefore focus on minimizing the number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
+Online Appendix XXX presents results for a larger number of metropolitan areas.
+
+Case selection was therefore informed by purposeful sampling to identify “information-rich” cases (Patton, 2002; Yin, 2009) that can provide in-depth knowledge about variation across Geomarkets and overtime.
+We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
+Based on our read of @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
+11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, Philadelphia**,** Phoenix, and Pittsburgh.
+To compare and contrast with classic urban model structures, we also identified alternative urban models such as bimodal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
+The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S.
+regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations.
+Cumulatively related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2).
+We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.
+
+Lastly, choice of metropolitan cases was informed by student list data availability.
+Although Census data are available for all metropolitan areas nationally (RQ1), we do not have good candidate student list data for all metropolitan areas (RQ2).
+In order to simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets, student list data must contain orders that resulted in a large pool of prospects within the metropolitan areas considered above that would include all Geomarkets within metropolitan areas.
+This lead us to select on orders that filtered for an entire state (e.g., State filter is set to Pennsylvania), an entire metropolitan area without a Geomarket filter (CBSA filter is set to the Philadelphia metropolitan area) , or an entire metropolitan area with a Geomarket filter inclusive of all Geomarkets (CBSA filter is set to the Philadelphia metropolitan area and Geomarket filter is set to all five Geomarkets in the Philadelphia area).
+In order to isolate the impact of Geomarkets on the racial/ethnic and socioeconomic characteristics of prospect pools, a suitable metropolitan case candidate would also be contingent on student list orders that did not include too many filtering criteria beyond those detailed previously.
+Because nearly all orders we collected (Author, BLINDED) filter on some academic criteria, we only considered orders that included assessment threshold filters (e.g., PSAT, SAT) across a wide score range to explore whether and to what extent Geomarket contributions to prospect pools are consistent across all score ranges.
+
+Our case selection process resulted in three metropolitan cases: Chicago, Dallas-Fort Worth, and Los Angeles.
+
+**Analyses**.
+Analyses are simultaneously descriptive and spatial.
+We answer RQ1 by producing Geomarket-level tables and graphs that show how the Geomarkets in a selected metropolitan area vary on racial and socioeconomic characteristics and how they vary over time.
+We also produce interactive maps at the census tract-level to show more granular variation within and between Geomarkets.
+
+We answer RQ2 by analyzing student list purchases that encompassed all Geomarkets in a selected metropolitan areas.
+Descriptive tables and graphs describe the racial and socioeconomic characteristics of prospects that would have been included/excluded had the purchase filtered on particular Geomarkets.
+We also create interactive maps of prospect pools by Geomarket contributions to provide a more granular visualization of analyses.
 
@@
 148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.College access scholarship from economics often analyzes the effects of district/state/federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
-These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^14]
-
-[^14]:  @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
+These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^19]
+
+[^19]:  @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
 

### 2025-05-08 text/manuscript-full-KS.qmd
front end revise

@@
 
-Scholarship at the nexus of enrollment management and college access can be categorized by which part(s) of the enrollment funnel it speaks to.
+Scholarship at the nexus of EM and college access can be categorized by which part(s) of the enrollment funnel it speaks to.
 The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
@@
 
-By focusing on the behavior of colleges, scholarship on EM implicitly states that recruiting is something done by individual colleges.
+By focusing on the behavior of colleges, scholarship on EM implicitly assumes that recruiting is something done by individual colleges.
 The broader EM industry includes relevant stakeholders in the organizational field [@RN527], including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
 We argue college access is structured by third-party servicers and products that interact with direct-providers (colleges).
-Although sociologists have hinted at the ways enrollment management contributes to inequality in college access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
+Although sociologists have hinted at the ways EM contributes to inequality in college access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
 <!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
@@
 
-Drawing from scholarship on organizational theory, enrollment management processes require colleges to make many "make or buy" [@RN4738; @RN980] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
+Drawing from scholarship on organizational theory, EM processes require colleges to make many "make or buy" [@RN4738; @RN980] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
 Unfortunately, few empirical studies make EM vendors or products the object of empirical analysis.
@@
 The Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
-Both products are applications of quantification.
-Student Search Service includes or excludes prospects from a student list purchase based on value of selected filters like Geomarket while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
-The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on equality of opportunity.
+Both products are applications of quantification that sort students into colleges.
+Student Search Service includes or excludes prospects from a student list purchases by colleges based on the value of selected filters like Geomarket while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
+The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on the inequitable sorting of students into colleges.
 
@@
 
-@RN5007 [p.
-402] define quantification as "the production and communication of numbers" (p.
-402).
-Reacivity, discipline, and authority are three interrelated themes that describe the effects of quantification.
-*Reactivity* is the idea that salient quantitative measures cause people and organizations to change their behavior.
-Quantification *disciplines* actors to react in particular ways.
-Quantification changes decision-making power *authority*, often weakening the discretion of local decision-makers.
+Quantification can be defined as "the production and communication of numbers" [@RN5007, p.
+402].
+While a growing interdisciplinary area of scholarship, sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, and personal life [@RN5008].
+Quantification has also been at the core of recent critical data scholarship exploring how algorithms and big data computation structure social opportunities [e.g., @RN4772; @RN4849].
+Our study is grounded in the systematic investigation of "interactions between different quantification regimes" across sociology and critical data studies, as called for by @RN5008 (p.
+239), to explore the role of third-party servicers and products in the sorting of students into colleges.
+Specifically, we draw on quantification scholarship investigating how correlation and homophily [@RN4975] encode social reproduction into our growingly dominant "algorithmic society" \[[CITE](https://medium.com/chatbots-bible-and-more/theres-an-algorithm-for-that-or-there-soon-will-be-d54406dc9bc1)\] and explore how such mechanisms are applied to the construction of Geomarkets in the Market Segment Model [@RN4982].
+
+<!--Reactivity, discipline, and authority are three interrelated themes that describe the effects of quantification.
+Reactivity is the idea that salient quantitative measures cause people and organizations to change their behavior.
+Quantification disciplines actors to react in particular ways.
+Quantification changes decision-making power authority, often weakening the discretion of local decision-makers.
 For instance, scholarship about U.S.
@@
 Rankings also affected hiring decisions of firms because clients evaluated firms based on the prestige of law schools attended by firm lawyers.
-Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- were using rankings to make decisions that had large consequences for schools \ldots \[then\] schools felt pressured to take them seriously" [@RN5007, p.
+Once law schools realized that "important groups of constituents --- students, faculty, trustees, employers, other media --- used rankings to make decisions that had large consequences for schools \ldots [then] schools felt pressured to take them seriously" [@RN5007, p.
 415].
-Rankings *disciplined* the behavior of law schools.
+Rankings disciplined the behavior of law schools.
 For example, characteristics valued by the rankings system (e.g., LSAT scores) became more important for decisions about admissions and merit aid [@RN5003].
-Rankings also weakened the *authority* of admissions personnel to make independent decisions about class size and which applicants to admit [@RN5003].
+Rankings also weakened the authority of admissions personnel to make independent decisions about class size and which applicants to admit [@RN5003].
 <!--
@@
 
-The analysis of UK school "league tables" by @RN5005 shows how quantification can be a mechanism of social reproduction.
+<!--The analysis of UK school "league tables" by @RN5005 shows how quantification can be a mechanism of social reproduction.
 In 1992, the UK government began publishing school league tables, which ranked schools based on student performance on national exams taken at age 16.
@@
 Using longitudinal survey data, managerial/professional households were more likely to move to localities with higher ranked schools following the introduction of school league tables.
-The introduction of league tables contributed to social reproduction because professional/managerial households were more aware of these consumer-facing metrics and had resources to respond by moving to more expensive neighbhorhoods, near higher performing schools.<!-- THIS IS RATCHET EFFECT STUFF.
+The introduction of league tables contributed to social reproduction because professional/managerial households were more aware of these consumer-facing metrics and had resources to respond by moving to more expensive neighbhorhoods, near higher performing schools.
+THIS IS RATCHET EFFECT STUFF.
 I CUT IT BECAUSE YOU HAVEN'T INTRODUCED THE IDEA OF RATCHET EFFECT
@@
 ```
+
+<!--# REMOVED FROM PARAGRAPH BELOW: The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4772; @RN4849].
+In particular, the discussions of correlation (chapter 1) and homophily (chapter 2) by @RN4975 introduce ideas salient to the analysis the Market Segment Model and Geomarkets.
+-->
+
 **Correlation**.
-The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4772; @RN4849].
-In particular, the discussions of correlation (chapter 1) and homophily (chapter 2) by @RN4975 introduce ideas salient to the analysis the Market Segment Model and Geomarkets.
-Correlation measures the extent to which two or more variables move together.
-Predictive analytics are based on correlation and are developed in two steps.
+Broadly, correlation measures the extent to which two or more variables move together.
+Predictive algorithms are based on correlation and are developed in two steps.
 First, apply statistical techniques to previous cases (training data) in order to identify factors positively and negatively associated with an outcome of interest.
@@
 
-Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with a variable of interest.
+Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with an variable of interest.
 For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
@@
 95] that cause behavior consistent with homophily.
-Second, in commercial social networks, homopphily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
+Second, in commercial social networks, homophily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
 Thus, "social networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@RN4975, p.
@@
 We make targeted contributions to scholarship on quantification in the sociology of education.
-First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
 As @RN3519 demonstrates in the chapter *Numbers*, EM is fundamentally concerned with quantification.
-However, sociological analyses of enrollment management focus on the behavior of colleges and their agents, yielding the implicit belief that inequalities created by EM are a function of individual colleges reacting to rankings and other macro structures in their environment.
+However, sociological analyses of EM focus on the behavior of colleges and their agents, yielding the implicit belief that inequalities created by EM are a function of individual colleges reacting to macro structures in their environment such as rankings and financial pressures.
 To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of producers.
-Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027; ].
-Drawing from scholarship from critical data studies [e.g., @RN4775; @RN4975; @RN4849], recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
-Based on a snapshot of existing social stratification, market research matches verticially categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
+Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027].
+Drawing from critical data studies [e.g., @RN4775; @RN4975; @RN4849], recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
+Based on a snapshot of existing social stratification, market research matches vertically categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
 Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociologists.
 We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
+
+<!--# REMOVED FROM PARAGRAPH ABOVE: First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
+-->
 
@@
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
-Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification on social reproduction.
-However, by showing which students are included/excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in educational opportunity.
+By showing which students are included versys excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in the sorting of students into colleges.
+
+<!--# removed from paragraph above: Unlike the analysis of UK school league tables, @RN5005 we cannot show the effect of quantification on social reproduction. However, b -->
 
@@
 
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of sotrytelling\ldots \[because\] we believe that the intuitios of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research this is based opn listening carefully to what admissions officers have to say (pp.
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of storytelling\ldots \[because\] we believe that the intuitions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research is based upon listening carefully to what admissions officers have to say (pp.
 9-10).
@@
 **Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one anothyer.[^7] Based on these analyses, @RN4982 \[p.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one another.[^7] Based on these analyses, @RN4982 \[p.
 46\] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
@@
 For each Geomarket, colleges could obtain Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors.
-Based on background conversations with enrollment management professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket.
+Based on background conversations with EM professionals, EPS software also provided information about the score-sending behavior of individual high schools within each Geomarket.
 Therefore, colleges used EPS software to decide which Geomarkets to recruit from and which high schools to visit within targeted Geomarkets.
@@
 Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
-The EPS product increases the ability of a college admissions leader -- working with College Board staff or an enrollment management consultant -- to plan recruiting efforts centrally.
-In background conversations, enrollment management professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
+The EPS product increases the ability of a college admissions leader -- working with College Board staff or an EM consultant -- to plan recruiting efforts centrally.
+In background conversations, EM professionals told us that EPS software enabled colleges to plan travel without relying on admissions officers having strong local knowledge of their territories.
 However, EPS software often recommended visiting the same sets of affluent, high-achieving high schools that were receiving visits from other colleges.
@@
 
-The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the enrollment management industry.
+The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the EM industry.
 The logic of homophily argues that there is a natural cultural match between students, that colleges should find students who look like students who enrolled at their institution in the past, and that colleges should recruit from the same schools and communities that are popular with peer colleges.

### 2025-05-09 text/back_end_KS.qmd
working through back end

@@
 
-```{=tex}
 \pagenumbering{roman}
 \newpage
-```
+
 **ABSTRACT**
@@
 
-```{=tex}
 \newpage
@@
 \pagenumbering{arabic}
-```
+
 # Data and Methods
@@
 **RQ1**.
-Our study utilizes two main data sources based on each research question.
-The first research question is focused on understanding the socioeconomic and racial variation between Geomarkets in metropolitan areas and how such variation changes over time.
+RQ1 primarily utilizes Census Data while RQ2 primarily utilizes data about student list purchases.
+Research question 1 is about socioeconomic and racial variation between Geomarkets in metropolitan areas and how such variation changes over time.
 To answer this question, we use census tract data from the U.S.
@@
 
-For the first research question, we use data from the 1980 and 2000 Decennial Census as well as 5-year estimates from the 2020 American Community Survey (ACS).
-For the 1980 Decennial Census, we specifically draw on Summary Tape File 1 (STF1) for variables about race and Hispanic origin collected via “short form” questions answered by all households and Summary Tape File 3 (STF3) for variables about socioeconomic characteristics (income, education, and poverty) collected via the "long form" questionnaire completed by a sample of households.
+We use data from the 1980 and the 2000 Decennial Census as well as 5-year estimates from the 2020 American Community Survey (ACS).
+For the 1980 Decennial Census, we draw from Summary Tape File 1 (STF1) for variables about race and Hispanic origin collected via "short form" questions answered by all households.
+We draw from Summary Tape File 3 (STF3) for variables about socioeconomic characteristics (income, education, and poverty) collected via the "long form" questionnaire completed by a sample of roughly 20% of all households.
 The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-To examine how Geomarket characteristics change over time, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty from the 2000 Decennial Census.[^1] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+For the 2000 Decennial Census, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty.[^1] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
 These questions were replaced by ACS.
@@
 These variables are median household income, percent of households below the poverty line, and people age 25+ with a BA.
-All measures were created at the census tract-level and the Geomarket-level by aggregating tract-level data to Geomarkets.
+All measures were created at the census tract-level and at the the Geomarket-level by aggregating tract-level data to Geomarkets.
 
@@
 Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
-Census tract shapefiles that intersected with multiple Geomarkets (i.e., a census tract's spatial boundaries were spread across two or more Geomarkets) broken into smaller shapefiles wholly contained within a single Geomarket.
-The variables in these shapefiles were then replaced with the original variable value multiplied by the proportion of land area of the census tract that was contained within the Geomarket.
-For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 1,000 people who identify as Hispanic non-white.
+We split census-tract shapefiles that intersected multiple Geomarkets (i.e., tracts whose boundaries spanned two or more Geomarkets) into smaller shapefiles, each entirely contained within a single Geomarket.
+We replaced each variable in these shapefiles with its original value scaled by the share of the tract’s land area that lay within the Geomarket.
+For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 100 people who identify as Asian, non-Hispanic.
 The partial spatial merge splits this census tract into two observations.
-The observation assigned to Geomarket A reports $.60*1000 = 600$ Hispanic non-white people and the observation assigned to Geomarket B reports $.40*1000 = 400$ Hispanic non-white people.
+The observation assigned to Geomarket A reports $.60*100 = 60$ Asian, non-Hispanic people and the observation assigned to Geomarket B reports $.40*100 = 40$ Asian, non-Hispanic people.
 
 **RQ2**.
-Our study's second research question asks how does the racial and socioeconomic composition of included versus included prospects vary when student list purchases filter on Geomarkets.
+Our second research question asks, how does the racial and socioeconomic composition of included versus excluded prospects vary when student list purchases filter on Geomarkets?
 We answer this question using actual student lists purchased by public universities.
@@
 We narrowed the scope of our request to student lists purchased from College Board, ACT, and the National Research Center for College and University Admissions (NRCCUA), the three largest student list vendors at the time.
-For each student list purchased from 2016 through 2020, we requested two related pieces of data: (1) the order summary, which specifies search criteria for the student list purchase; and (2) the de-identified prospect-level list produced from the search criteria.
+For each student list purchased from 2016 through 2020, we requested two related pieces of data: (1) the order summary, which specifies search criteria for the student list purchase; and (2) the de-identified prospect-level list produced from these search criteria.
 
 The data collection sample for this project was all public universities in California, Illinois, Minnesota, and Texas.
-Utilizing public records requests to obtain public records is a painstaking process.
-Initially, the majority of universities did not respond or denied our request.
-Subsequently, we obtained pro bono representation from four law firms, which substantially increased the success of data collection.
+We chose these states because they were priorities for the two foundations that funded the data collection.
+Utilizing public records requests to obtain quantitative data is a painstaking process, made more difficult by the onset of the Covid Pandemic.
+Initially, the majority of universities ignored or denied our request.
+We established a client relationship with a civil rights legal organization and leveraged this relationship to obtain pro bono representation from four law firms.
+Firm representation substantially increased the success of data collection.
 We also collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
-However, we were unable to obtain legal representation for Texas.
+However, we were unable to obtain legal representation for Texas because firms interested in representing us had conflicts of interest in Texas.
 
-Even with firm representation, data collection remained difficult.
+Firm representation substantially increased the success of data collection.
+Nevertheless, data collection remained difficult.
 Some universities provided records that were not usable for quantitative analyses (e.g., summary statistics across multiple orders; or data did not contain important fields).
 Some universities did not provide records based on legitimate legal grounds (e.g., data not in university possession; not required to create records that do not currently exist).
-We learned that many universities outsourced student list purchases to third-party consulting firms.
+We learned that many universities outsourced student list purchases to third-party consulting firms, particularly [EAB](https://eab.com/).
 Unfortunately, we were rarely able to obtain usable data from these universities.
-A small number of universities denied requests based on potentially questionable legal rationale, but we lacked the resources to litigate.
+Several universities denied requests based on questionable legal rationale, but we lacked the resources to litigate.
 
-The primary data source for our second research question is therefore student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
-This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Minnesota, and Arizona \[KS Does this still apply?\].
-Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^4] The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
-In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.
-Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
+The primary data source for our second research question is student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
+This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Texas, and Arizona.
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.
+The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
+In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.[^4] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
 
@@
 
-In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter, but did purchase prospects from all Geomarkets in a particular metropolitan area.
-For each prospect we know their home zip code and, therefore, which Geomarket they belong to.
-We therefore simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.^[For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
+Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
 
-The variables of interest for student list data are the characteristics of prospects who would have been included versus excluded by our simulations by Geomarket filter.
-Prospect characteristics are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
-College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity nonresponse) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016).
+The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
+College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) [CITE].
 From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
-Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016).[^5] Information about student socioeconomic status is limited.
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) [CITE].[^5] Information about student socioeconomic status is limited.
 However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
@@
 **Research design**.
-We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
-Case studies are used to gain a deeper understanding of the phenomenon of interest by comparing and contrasting across different settings \[CITE YIN 2001\].
-Our study's research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
-Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics.
+We utilize a multiple, quantitative case study design.
+Metropolitan areas are cases.
+Case studies deepen understanding of phenomena by comparing and contrasting across different settings \[CITE YIN 2001\].
+Our research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
+Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics [MAKE THIS MORE ABOUT EXPLORING THE MODEL.
+THIS SENTENCE SOUNDS FANCY BUT DOESN'T REALLY SAY ANYTHING].
 Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts.
-We therefore focus on minimizing the number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
+We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
 Online Appendix XXX presents results for a larger number of metropolitan areas.
 
-Case selection was therefore informed by purposeful sampling to identify “information-rich” cases (Patton, 2002; Yin, 2009) that can provide in-depth knowledge about variation across Geomarkets and overtime.
+Case selection was informed by purposeful sampling to identify "information-rich" cases (Patton, 2002; Yin, 2009)[CITE] that can provide knowledge about variation across Geomarkets and over time.
 We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
-Based on our read of @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
-11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, Philadelphia**,** Phoenix, and Pittsburgh.
+Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
+11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia.
 To compare and contrast with classic urban model structures, we also identified alternative urban models such as bimodal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
@@
 regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations.
-Cumulatively related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2).
+Related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2).
 We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.
@@
 Although Census data are available for all metropolitan areas nationally (RQ1), we do not have good candidate student list data for all metropolitan areas (RQ2).
-In order to simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets, student list data must contain orders that resulted in a large pool of prospects within the metropolitan areas considered above that would include all Geomarkets within metropolitan areas.
-This lead us to select on orders that filtered for an entire state (e.g., State filter is set to Pennsylvania), an entire metropolitan area without a Geomarket filter (CBSA filter is set to the Philadelphia metropolitan area) , or an entire metropolitan area with a Geomarket filter inclusive of all Geomarkets (CBSA filter is set to the Philadelphia metropolitan area and Geomarket filter is set to all five Geomarkets in the Philadelphia area).
-In order to isolate the impact of Geomarkets on the racial/ethnic and socioeconomic characteristics of prospect pools, a suitable metropolitan case candidate would also be contingent on student list orders that did not include too many filtering criteria beyond those detailed previously.
+In order to show which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets, the student list purchases we analyze must include all Geomarkets in a metropolitan area.
+This lead us to select on orders that filtered for an entire state (e.g., State filter is set to Pennsylvania), an entire metropolitan area without a Geomarket filter (CBSA filter is set to the Philadelphia metropolitan area), or an entire metropolitan area with a Geomarket filter inclusive of all Geomarkets (CBSA filter is set to the Philadelphia metropolitan area and Geomarket filter is set to all five Geomarkets in the Philadelphia area).
+Although we would like to analyze the D.C./Maryland/Virginia (DMV) area, the student list purchases we collected exclude "MD 5 -- Prince Georges Metropolitan" and "MD 7 -- Baltimore (Urban), which are substantially poorer and more Black than other Geomarkets in the DMV.
+Similarly, student list purchases targeting Cleveland exclude "OH 4 -- City of Cleveland (East)."
+
+In order to isolate the impact of Geomarkets on the racial/ethnic and socioeconomic characteristics of prospect pools, we did not consider purchases that utilized student preference (e.g., campus size, intended major), low-income household, or first-generation status filters.
 Because nearly all orders we collected (Author, BLINDED) filter on some academic criteria, we only considered orders that included assessment threshold filters (e.g., PSAT, SAT) across a wide score range to explore whether and to what extent Geomarket contributions to prospect pools are consistent across all score ranges.
@@
 @fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
-The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 in the development of The Market Segment Model.
+The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982[p.
+12].
 Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
@@
 
-These population shifts have contributed to changes in the racial/ethnic composition within and across Geomarkets over time.
-@fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarkets across the three census periods.
+These modest population sgufts belie substantial change in the racial/ethnic composition within and across Geomarkets over time.
+@fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarket across the three census periods.
 In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
 In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
-By 2020, several Geomarkets that were previously predominantly White shifted to a more balanced racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
+By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
 For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
 Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
-
 @fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
-For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
+<!-- For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
 City of Chicago also experienced relatively modest declines in its Black population (37% in 1980 to 33% in 2020), alongside increases in its Hispanic (15% in 1980 to 26% in 2020) and Asian Pacific Islander populations (4% in 2000 to 6% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
+-->
 
@@
 Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs.
-However, declining White populations, Hispanic and Asian population growth, and Black outmigration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining.
+However, declining White populations, Hispanic and Asian population growth, and Black out-migration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining.
 The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents) while the the Evanston and North Shore Geomarkets remain predominantly White.
 While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains across Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
+[I THINK SOME DETAIL CAN BE REDUCED HERE; ALSO, I THINK YOU ARE BURYING THE LEDE THAT NEEDS TO BE EMPHASIZED: THERE IS SUBSTANTIAL RACIAL AND SES INEQUALITY BETWEEN GEOMARKETS AND CHICAGO ESPECIALLY IS POOREST AND MOST NON-WHITE]
 
@@
 To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospects across the entire state of Illinois.
-Two of these orders applied the same filters: SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
-The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with relatively middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
+The first two orders filter on the same SAT thresholds (ranging from 1020 to 1150), the same GPA thresholds (ranging from B- to A+), and filter on the high school graduating classes of 2020 and 2021 graduating classes, respectively.
+The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds.
+Two orders targeted students with relatively middle-range scores (1160 to 1300) and two orders targeted those with relatively high scores (1310 to 1600).
 Because all six orders included an Illinois state filter, we can utilize prospects from the Chicago metropolitan area within the resulting student lists to examine Geomarket contributions by simulating the application of a Geomarket filter across low-, middle-, and high-range SAT threshold orders.
 
+DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES!
 @fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area.
@@
 For example, more than 8,000 of the 16,485 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
-Similar to the larger metropolitan area, the Northshore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students.
+Similar to the larger metropolitan area, the North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students.
 Interestingly, while the City of Chicago is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-chicago-rq2-race includes nearly double the proportion of Hispanic (50%) prospects relative to the larger Geomarket resident population (26% Hispanic) in @fig-chicago-rq1-race with a lower proportion of White prospects (14%).
@@
 The City of Chicago Geomarket exhibits the largest underrepresentation of Black prospects (7% of prospects versus 33% of residents) while Chain of Lakes exhibits the largest underrepresentation of Hispanic prospects (7% of prospects versus 24% of residents).
+[LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN]
 
@@
 ```
+
 More than half (n=8,256) of all prospects whose contact information was purchased in these middle SAT score orders identified as White.
@@
 On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively.
-Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
 
@@
 Both figures demonstrate how representational Geomarket contributionsacross race/ethnicity, such as City of Chicago disproportionately contributing larger shares of Black and Hispanic prospects, remain consistent even at lower and higher SAT score thresholds.
+[REVISE THESE SENTENCES TO HAMMER THE LEDE -- IF WE EXCLUDE CITY OF CHICAGO GEOMAREKT, THE WAY THE ASU ORDER DOES, WE LOSE A DISPROPORTIONATE SHARE OF BLACK AND HISPANIC STUDENTS]
 
@@
 In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
+[MY READ IS THAT THE PATTERNS FOR WHITE ARE NOT THE SAME AS THE PATTERNS FOR ASIAN AND LATINX -- SO I WOULDN'T SAY THESE PATTERNS PERSIST ACROSS OTHER RACIAL/ETHNIC CATEGORIES.
+I WOULD RECOMMEND STARTING WITH ASIAN AND WITH HISPANIC.
+THAT IS THE LEDE; ALSO, THIS IS WHERE YOU COULD REFER TO THE SIMILAR GRAPH FOR PHILADELPHIA AND SAY IT FOLLOWS THE SAME PATTERN; I THINK THIS IS THE CLIMAX AND COOLEST INSIGHT OF THE RESULTS]
 
@@
 However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+[SO WHAT HAPPENS IF A SCHOOL USES A GEOMARKET FILTER AND EXCLUDES CHICAGO.]
 
@@
 
-```{=tex}
 \appendix
 \renewcommand{\thetable}{A\arabic{table}}
-```
+
 # Appendix A

### 2025-05-12 text/manuscript-full-KS.qmd
full version

@@
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
-By showing which students are included versys excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in the sorting of students into colleges.
+By showing which students are included versus excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in the sorting of students into colleges.
 
@@
 
-Having defined geography, the Market Segment Model sought to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households in the Geomarket:
-
-> For our analysis, we sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
+Having defined geography, the Market Segment Model sought to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households.
+The model created a straightforward but consistent classification system that grouped SAT test-takers into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
+For a given student, each college that receives a score from the student can be defined as "local" (college located in the the same Geomarket as the student), "in-state" (same state but different Geomarket), "regional" (same region but different state), or "national" (different region).
+In turn, a test-taker is categorized in the "local" market segment if they submit more SAT scores to local institutions than they do to in-state, regional, or national institutions.
+An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
+
+<!--# For our analysis, we sought not a complex mathematical model, but a straightforward classification system that would track the pattern of SAT-score submissions to create a map of student choice.
 The Market Segment model that we developed was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
@@
 4].
-
-SAT test-takers were categorized into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
+-->
+
+<!-- SAT test-takers were categorized into one of four different market segments -- local, in-state, regional, and national -- based on SAT score-sending behavior.
 For a given student, each college that receives a score from the student can be defined as "local" (college located in the the same Geomarket as the student), "in-state" (same state but different Geomarket), "regional" (same region but different state), or "national" (different region).
@@
 An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
-<!--
 A "regional" student submits more SAT scores to regional institutions than they do to local, in-state, or national institutions.
@@
 For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
-A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a colleges which Geomarkets possess attractrive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a college which Geomarkets possess attractive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
 These outputs became the basis for the Enrollment Planning Service (EPS) software.
@@
 33\] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."\[\^7\] @RN4982 [p.
-42] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations." The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
+42] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations."
+
+The authors also find that these four variables predict student score-sending behavior at the Geomarket-level.
 However, Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
@@
 This homophily is presented as a naturally occurring phenomenon.
-Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer-coleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
+Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer-colleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
 In itself, @RN4982's Market Segment Model is a social science depiction of student demand -- akin to the status attainment model -- that does not consider historic, structural inequalities that cause observed patterns.

### 2025-05-13 text/back_end_KS.qmd
chicago revisions

@@
 \pagenumbering{arabic}
-
+```
 # Data and Methods
@@
 
-In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.^[For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.[^5] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
 Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
 
+[^5]: For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.
+
 The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
-College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) [CITE].
+College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) \[CITE\].
 From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
-Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) [CITE].[^5] Information about student socioeconomic status is limited.
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].[^6] Information about student socioeconomic status is limited.
 However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
 
-[^5]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
+[^6]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
 Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”
@@
 Our research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
-Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics [MAKE THIS MORE ABOUT EXPLORING THE MODEL.
-THIS SENTENCE SOUNDS FANCY BUT DOESN'T REALLY SAY ANYTHING].
+Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics \[MAKE THIS MORE ABOUT EXPLORING THE MODEL.
+THIS SENTENCE SOUNDS FANCY BUT DOESN'T REALLY SAY ANYTHING\].
 Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts.
@@
 We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
-Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
+Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area[s] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
 11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia.
@@
 ```
-
 # Findings
@@
 @fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
-The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982[p.
+The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 [p.
 12].
@@
 
-These modest population sgufts belie substantial change in the racial/ethnic composition within and across Geomarkets over time.
+These modest population shifts lead to substantial change in the racial/ethnic composition within and across Geomarkets over time.
 @fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarket across the three census periods.
@@
 
-From 1980 to 2020, these demographic shifts have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by racial/ethnic and socioeconomic characteristics.
+Demographic shifts overtime have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by substantial racial/ethnic and socioeconomic inequality between Geomarkets.
 The original Geomarket boundaries delineated significant racial and income segregation in the 1980s.
@@
 Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs.
-However, declining White populations, Hispanic and Asian population growth, and Black out-migration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining.
-The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents) while the the Evanston and North Shore Geomarkets remain predominantly White.
-While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains across Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
-[I THINK SOME DETAIL CAN BE REDUCED HERE; ALSO, I THINK YOU ARE BURYING THE LEDE THAT NEEDS TO BE EMPHASIZED: THERE IS SUBSTANTIAL RACIAL AND SES INEQUALITY BETWEEN GEOMARKETS AND CHICAGO ESPECIALLY IS POOREST AND MOST NON-WHITE]
+However, declining White populations, Hispanic and Asian population growth, and Black out-migration, as well as narrowing of education and poverty gaps within some Geomarkets, have contributed to a more stratified Geomarket dynamic in the metropolitan area with significant disparities remaining between Geomarkets.
+The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets now also include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents).
+While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains between Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
+
+<!--# [I THINK SOME DETAIL CAN BE REDUCED HERE; ALSO, I THINK YOU ARE BURYING THE LEDE THAT NEEDS TO BE EMPHASIZED: THERE IS SUBSTANTIAL RACIAL AND SES INEQUALITY BETWEEN GEOMARKETS AND CHICAGO ESPECIALLY IS POOREST AND MOST NON-WHITE ————-  DALLAS IS POOREST AND MOST NON-WHITE; NOT CHI] -->
 
@@
 
-DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES!
-@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area.
-At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets generally reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
-For example, more than 8,000 of the 16,485 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
-Similar to the larger metropolitan area, the North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students.
-Interestingly, while the City of Chicago is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-chicago-rq2-race includes nearly double the proportion of Hispanic (50%) prospects relative to the larger Geomarket resident population (26% Hispanic) in @fig-chicago-rq1-race with a lower proportion of White prospects (14%).
+<!--# DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES!
+--> @fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles.
+At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
+For example, more than 8,000 of the 15,981 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
+The North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects.
+Unlike any other Geomarket, nearly nine of every 10 prospects from the City of Chicago Geomarket were Hispanic (50%), Black (30%), Asian (5%), or Multiracial (2%).
 
-However, the racial/ethnic composition of prospect pools become less similar to the overall metropolitan area as PSAT scores increase.
-Notably, the shares of White and Asian prospects within Geomarkets increase, while Black and Hispanic prospects decrease.
-For instance, Asian prospects over representation in Geomarkets relative to the overall resident population at the 1310-1600 SAT score threshold ranges from a minimum of 4 percentage points in the North Shore Geomarket (17% of prospects versus 13% of residents) to a maximum of 21 percentage points in the Chain of Lakes (28% of prospects versus 7% residents) Geomarket.
-The City of Chicago Geomarket exhibits the largest underrepresentation of Black prospects (7% of prospects versus 33% of residents) while Chain of Lakes exhibits the largest underrepresentation of Hispanic prospects (7% of prospects versus 24% of residents).
-[LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN]
+The City of Chicago Geomarket increasingly becomes the only Geomarket with substantial shares of Black, and Hispanic prospects as PSAT scores increase.
+All Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline.
+For instance, the City of Chicago Geomarket pool is 18% Asian, 7% Black, 21% Hispanic, 5% Multiracial, making it the only Geomarket that is not predominantly White (49%).
+Prospect pools from all other Geomarkets in this high PSAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.4% (North Shore) to 5% (South and Southwest Suburbs) of prospects and 6% (Northwest Suburbs) to 10% (South and Southwest Suburbs) of prospects, respectively.
+<!--# [LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN] -->
 
@@
 ```
+For instance, @fig-chicago-rq2-midSAT-race shows the nearly 1,000 Black students whose contact information was purchased in these middle SAT score orders were largely contributed to the overall pool by the City of Chicago and South and Southwest Suburbs.
+These two Geomarkets contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects.
+On the other hand, all other Geomarkets contributed substantially smaller shares of Black prospects.
+The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
 
-More than half (n=8,256) of all prospects whose contact information was purchased in these middle SAT score orders identified as White.
-@fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
-One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects.
+In contrast, these representational patterns by Geomarket are reversed for White and Asian students.
+@fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates proportional to their overall contributions to all included prospects (except City of Chicago).
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects) and Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospectq8s) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+
+<!--# One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects.
 On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively.
-Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+-->
 
-In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race.
+<!--# MOVED THIS UP In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race.
 Among the nearly 1,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
 The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
+-->
 
-Appendix figures \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting lower-range (1020-1150) and high-range (1310-1600) SAT scores.
-Both figures demonstrate how representational Geomarket contributionsacross race/ethnicity, such as City of Chicago disproportionately contributing larger shares of Black and Hispanic prospects, remain consistent even at lower and higher SAT score thresholds.
-[REVISE THESE SENTENCES TO HAMMER THE LEDE -- IF WE EXCLUDE CITY OF CHICAGO GEOMAREKT, THE WAY THE ASU ORDER DOES, WE LOSE A DISPROPORTIONATE SHARE OF BLACK AND HISPANIC STUDENTS]
+These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Chicago metropolitan area using Geomarket filters that exclude the City of Chicago– as observed in some orders collected via our public requests – result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
+Appendix \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race show that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters.
+In both cases, the City of Chicago contributes disproportionately larger shares of Black and Hispanic prospects compared to all other Geomarkets.
+<!--# [REVISE THESE SENTENCES TO HAMMER THE LEDE -- IF WE EXCLUDE CITY OF CHICAGO GEOMAREKT, THE WAY THE ASU ORDER DOES, WE LOSE A DISPROPORTIONATE SHARE OF BLACK AND HISPANIC STUDENTS] -->
 
@@
 @fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 12,981 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders.
-For instance, among all White prospects (n=8,155) whose contact information was purchased in this SAT score range, approximately 6% were first-generation college students whose parents did not attend college, 15% were first-generation college students whose parents attended some college but did not complete their degree, and 79% were not first-generation college students.
+@fig-chicago-rq2-midSAT-combo shows the City of Chicago Geomarket consistently contributes a disproportionately larger share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
+For instance, among all Asian prospects (n=1,341) whose contact information was purchased in this SAT score range, approximately 12% were first-generation college students whose parents did not attend college, 16% were first-generation college students whose parents attended some college but did not complete their degree, and 72% were not first-generation college students.
+@fig-chicago-rq2-midSAT-combo shows the City of Chicago disproportionately contributes 35% of all Asian first-generation college students whose parents did not attend college.
+For first-generation college students whose parents attended some college but never completed degrees, City of Chicago (22%), South and Southwest Suburbs (20%), and Evanston and Skokie (20%) Geomarkets contributed larger shares of Asian prospects.
+
+Similar Geomarket contribution patterns are evident among Hispanic prospects.
+The City of Chicago Geomarket stands out as the only Geomarket contributing substantially larger shares of Hispanic first-generation college students across those whose parents did not attend college (46% versus 38% overall) and those whose parents attended some college but did not complete their degrees (29% versus 26% overall).
+Although Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and Western Surburbs Geomarkets.
+
+Lastly, although Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and Western Surburbs Geomarkets.
+Similarly, Geomarkets such as Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs contribute a disproportionately larger share of White prospects who are not first-generation college students, whereas City of Chicago and South and Southwest Suburbs contribute larger shares of White first-generation college students.
+
+<!--# OLD WHITE PROSPECT TEXT among all White prospects (n=8,155) whose contact information was purchased in this SAT score range, approximately 6% were first-generation college students whose parents did not attend college, 15% were first-generation college students whose parents attended some college but did not complete their degree, and 79% were not first-generation college students.
 @fig-chicago-rq2-midSAT-combo shows the City of Chicago and South and Southwest Suburbs Geomarkets disproportionately contribute larger shares of White first-generation college students who parents did not attend college (10% and 9%, respectively).
@@
 In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
-[MY READ IS THAT THE PATTERNS FOR WHITE ARE NOT THE SAME AS THE PATTERNS FOR ASIAN AND LATINX -- SO I WOULDN'T SAY THESE PATTERNS PERSIST ACROSS OTHER RACIAL/ETHNIC CATEGORIES.
+-->
+
+<!--# [MY READ IS THAT THE PATTERNS FOR WHITE ARE NOT THE SAME AS THE PATTERNS FOR ASIAN AND LATINX -- SO I WOULDN'T SAY THESE PATTERNS PERSIST ACROSS OTHER RACIAL/ETHNIC CATEGORIES.
 I WOULD RECOMMEND STARTING WITH ASIAN AND WITH HISPANIC.
-THAT IS THE LEDE; ALSO, THIS IS WHERE YOU COULD REFER TO THE SIMILAR GRAPH FOR PHILADELPHIA AND SAY IT FOLLOWS THE SAME PATTERN; I THINK THIS IS THE CLIMAX AND COOLEST INSIGHT OF THE RESULTS]
+THAT IS THE LEDE; ALSO, THIS IS WHERE YOU COULD REFER TO THE SIMILAR GRAPH FOR PHILADELPHIA AND SAY IT FOLLOWS THE SAME PATTERN; I THINK THIS IS THE CLIMAX AND COOLEST INSIGHT OF THE RESULTS] KS=== I THINK ITS" WEIRD TO INCLUDE PHILLY GRAPH HERE WHEN ITS NOT CASE...WE CAN SPEAK TO THIS IN THE DISCUSSION INSTEAD...
+WE SEE SIMILAR PATTERNS ACROSS OTHER METROS TOO -->
 
-@fig-chicago-rq2-midSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
-The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
-However, the magnitude of this disproportionality varied- from a 1-percentage-point difference for Black prospects to a 23 percentage-point difference for Asian prospects.
-Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
-Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs Geomarkets.
 Appendix figures @fig-chicago-rq2-lowSAT-combo and \@fig-chicago-rq2-highSAT-combo show the magnitude of these Geomarket contribution disparities become more pronounced at as SAT score thresholds increase.
@@
 On the other hand, predominantly White and affluent Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, as well as non-first generation college students.
-However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
-[SO WHAT HAPPENS IF A SCHOOL USES A GEOMARKET FILTER AND EXCLUDES CHICAGO.]
+However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian and Hispanic college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+Given the broader racial/ethnic distribution of the Chicago metropolitan area, student list orders that use Geomarket filters excluding the City of Chicago and including more affluent Geomarkets in targeting high-achieving students may still capture some Asian and Hispanic prospects.
+However, these students are far more likely to be non–first-generation students of color.
+
+<!--# [SO WHAT HAPPENS IF A SCHOOL USES A GEOMARKET FILTER AND EXCLUDES CHICAGO.] -->
 

### 2025-05-13 text/back_end_KS.qmd
dallas revisions

@@
 Given the broader racial/ethnic distribution of the Chicago metropolitan area, student list orders that use Geomarket filters excluding the City of Chicago and including more affluent Geomarkets in targeting high-achieving students may still capture some Asian and Hispanic prospects.
-However, these students are far more likely to be non–first-generation students of color.
+However, these students are far more unlikely to be first-generation students of color.
 
@@
 In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
-By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
+By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian).
 In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
-Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.
+Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian residents.
 
@@
 However, both Geomarkets remained predominantly White (58% and 63%, respectively).
-In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian Pacific Islander), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian Pacific Islander).
+In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian).
 Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
@@
 
-The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s.
+The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets.
 While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
 Such growth is largely attributed to increases in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex.
-On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment across Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
+On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment between Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
 Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
@@
 @fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area.
-At the lowest PSAT score thresholds, notable differences emerge in the racial/ethnic composition of prospects within Geomarkets when compared to the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
-For example, the largest number of prospects in the 1070-1180 PSAT range were from the Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian), both of which yield predominantly White prospect pools.
-On the other hand, the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarket pools have the greatest proportions of non-White prospects (although yielding fewer overall students).
+At the lowest PSAT score thresholds (1070-1180), notable tiers emerge based on the racial/ethnic composition of prospects between Geomarkets.
+For instance, the central suburban Geomarkets of Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects.
+While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportion of Hispanic prospects but lower proportions of Asian and Black students comparatively.
+The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest and predominantly White prospect pools.
+
+<!--# which most closely mirror the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
+For example,  the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarkets have the greatest proportions of non-White prospects although yielding smaller overall pools.
 Interestingly, while the City of Dallas is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-dallas-rq2-race includes larger proportion of Hispanic (41%) prospects but lower proportions of Asian (4%) and Black (9%) students comparatively.
+-->
 
-These racial/ethnic composition patterns shift as PSAT scores increase.
-Shares of White prospects within Geomarkets remain relatively proportional to the overall 2020 metropolitan distribution in @fig-dallas-rq1-race.
-Howeover, prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
 

### 2025-05-13 text/back_end_KS.qmd
dallas revisions

@@
 
-@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area.
-At the lowest PSAT score thresholds (1070-1180), notable tiers emerge based on the racial/ethnic composition of prospects between Geomarkets.
-For instance, the central suburban Geomarkets of Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects.
-While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportion of Hispanic prospects but lower proportions of Asian and Black students comparatively.
-The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest and predominantly White prospect pools.
+@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area, revealing a comparatively more distributed pattern of racial/ethnic diversity across Geomarkets unlike Chicago's concentrated diversity within a single Geomarket.
+This distribution is most notable at the lowest PSAT score threshold.
+For instance, at the 1070-1180 PSAT score range, the central suburban Geomarkets of Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects.
+While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportions of Hispanic prospects but lower proportions of Asian and Black students.
+The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest but predominantly White prospect pools.
 
@@
 Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+Such shifts also shift distribution patterns of racial/ethnic diversity across Geomarkets, making City of Dallas and City of Fort Worth critical Geomarkets for including Black and Hispanic students from the resulting prospect pools.
+\[KS LEFT OFF HERE}
 
 @fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
-@fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).
-Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
-The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+Disparities in Geomarket contributions were evident but differed for Asian, Black, and Hispanic students.
+For instance, The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+
+<!--# @fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).
+-->
 

### 2025-05-13 text/back_end_KS.qmd
working through results

@@
 From 1980 to 2020, the population of the overall Dallas-Fort Worth metropolitan area increased substantially from approximately 2.7 Million to 6.8 Million.
-@fig-dallas-maps showcases all six Geomarkets experienced population growth over the three census periods.
+@fig-dallas-maps shows all six Geomarkets experienced population growth over the three census periods.
 However, the most pronounced increases occured in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
@@
 
+[WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]
 @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
@@
 
-Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
+[CITY OF DALLAS!]Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
 Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
@@
 
+[DON'T KNOW IF WE NEED PARAGRAPH ON POVERTY + EDUCATIONAL ATTAINMENT].
 @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
@@
 
-The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets.
+[DON'T THINK YOU NEED SUMMARY PARAGRAPHS LIKE THIS WITHIN SECTIONS FOR THIS METRO]The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets.
 While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
@@
 Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+<!--OZAN CUT THIS TEXT 5/13.
+THINKING NOT ESSENTIAL.
+AND NOT SO SURE IF IT SHOWS THAT.
 Such shifts also shift distribution patterns of racial/ethnic diversity across Geomarkets, making City of Dallas and City of Fort Worth critical Geomarkets for including Black and Hispanic students from the resulting prospect pools.
-\[KS LEFT OFF HERE}
+\[KS LEFT OFF HERE} -->
 
 @fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
+The bars for "All (race known)" show that Collin & Rockwall (37%) and West of Dallas/Forth Worth (23%) contribute many more students than the other four Geomarkets.
 Disparities in Geomarket contributions were evident but differed for Asian, Black, and Hispanic students.
@@
 
-Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Dallas-Fort Worth metropolitan area.
+Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity, which we view as the most interesting results for the Dallas-Fort Worth metropolitan area.
 Among all White prospects (n=2,094) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
-@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool.
-The only exception is the Irving, Arlington, & Grand Prairie Geomarket contributing a relatively larger share of White first-generation college students whose parents attended some college but did not complete their degree (21% versus 11% for all prospects).
+@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool, the only modest exception being the Irving, Arlington, & Grand Prairie Geomarket.
 
+By contrast, for non-White groups, a larger share of prospects were first-generation students.
+Furthermore, first generation status differed substantially across Geomarkets, especially for Hispanic and Asian students.
+For Hispanic students, more than 50% of prospects from the City of Dallas, City of Forth Worth, Arlington \& Grand Prairie, and Dallas County are first-generation college students.
+By contrast, for the more affluent Collin & Rockwall Geomarket and the West of Dallas/Fort Worth Metroplex Geomarkets, less than 30% of prospects were first-generation.
+For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
+By contrast, the more affluent Collin & Rockwall sent more Asian prospects (285) than any other Geomarket, but about 87% of these students had parents with a BA.
+
+<!-- I ADDED THE ABOVE PARAGRAPH AND CUT THIS PARAGRAPH TO GIVE THE LEDE MORE PUNCH 
 Similar patterns of one or two Geomarkets disproportionately contributing larger shares of first-generation college students was evident across other racial/ethnic categories in @fig-dallas-rq2-midSAT-combo.
@@
 City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
+-->
 
@@
 
-### Decentralized Diversity and Income in a Multi-Nucleo in the Los Angeles Geomarket Landscape
+### Decentralized Diversity and Income in a Multi-Nucleo Landscape
 
+Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles -- 11 Geomarkets, results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), which are shown in the Online Appendix.
 The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
@@
 Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring of made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
-The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland.
+The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland, comprising the Inland Empire.
 
@@
 Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
-In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
+[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
 By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%).
@@
 
+[I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST. START TEHRE AND THEN GET TO OTHER RESULTS.]
 @fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020.
@@
 
-The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
+[THINK YOU CAN CUT THIS PARAGRAPH] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
 From 1980 to 2020, the region’s population expanded significantly leading to overall declines in the White population and increases in Hispanic and Asian populations across Geomarkets.
@@
 
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university filtered for prospects across the entire state of California, in the 2019-2020 high school graduating class, and by PSAT scores.
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2019-2020 [NOT JUST 2019-20.
+REVISE] high school graduating class, and by PSAT scores.
 Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.
 
-The racial/ethnic composition of purchased student profiles within Geomarkets in the Los Angeles metropolitan area is presented in @fig-la-rq2-race.
+@fig-la-rq2-race presents the racial/ethnic composition of purchased student profiles.
 The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
+[THE LEDE IS SOUTH CENTRAL AND EAST LA.
+START THERE!].
 However, the racial/ethnic composition of prospects become less proportional to overall metropolitan patterns by Geomarket as PSAT scores increase.
@@
 This pattern becomes most pronounced at the highest PSAT score prospect pools.
-This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
+[I DON'T THINK THAT BLACK AND LATINX STUDENTS HAVING LOW TEST SCORES IS THE STORY WE ARE TRYING TO TELL.
+THAT STORY HAS BEEN TOLD.
+THE MORE IMPORTANT THING IS, WITHIN SOME TEST THRESHOLD, HOW DOES THE RACIAL COMPOSITION VARY ACROSS GEOMARKETS.
+LIKE EVEN WITHIN THE MID RANGE, SOUTH CENTRAL, EAST LA, AND INLAND EMPIRE STAND OUT] <!-- This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
 @fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of residents within Geomarkets.
 Similar underrepresentation patterns are evident for Hispanic porspects are evident across all Geomarkets execept for a nearly equal representation in South and South Central Los Angeles (73% prospects versus 71% Hispanic residents).
+-->
 
@@
 For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.
-Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
-The Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
 
+[I THINK CLEANER IF YOU JUST HAVE A SMALL PARAGRAPH FOR HISPANIC AND A SMALL PARAGRAPH FOR BLACK.
+I THINK ONE LEDE ACROSS THESE RESULTS IS RIVERSIDE AS AN ENGINE FOR BLACK/LATINX AND FOR FIRST-GEN STUDENTS.
+I GET THAT THESE TWO POINTS CONTRADICT EACH OTHER].
 Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race.
@@
 
-Contributions of each Geomarket to the pool of prospects from the Los Angeles metropolitan area based on first-generation college student status are presented in @fig-la-rq2-midSAT-ses.
+[I MOVED ASIAN] For Asian students, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger shares, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
+
+Contributions of each Geomarket to the pool of prospects from the Los Angeles metropolitan area based on first-generation college student status are presented in @fig-la-rq2-midSAT-ses [THIS SENTENCE IS UGLY.
+MAKE IT MORE LIKE FIG X SHOWS THE CONTRIBUTIONS EACH GEOMARKET MAKES BY FIRST GENERATION STATUS GROUP].
 Only 889 of the more than 8,000 total prospects (11%) in this middle PSAT score range (1190-1260) pool were first-generation college students whose parents did not attend college.
+[I'M NOT SURE WHERE YOU ARE GETTING THESE NUMBERS FROM; THIS IS THE WRONG FIGURE I THINK.
+AND WHEN YOU REVISE, THE LEDES ARE RIVSERSIDE, EAST LA AND, SOUTH CENTRAL (ONLY 2% OF PROSPECTS BUT 10% OF NO-COLLEGE)].
 Only three of the 11 Geomarkets contributed substantially larger shares of these first-generation prospects: East Los Angeles (35%), South and South Central Los Angeles (47%), and Riverside, San Bernardino, and Ontario (15%).
@@
 
+[I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW]
 Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
-Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
+[TAKE A TIP FROM THE CRIT QUANT FOLKS; YOU DON'T ALWAYS NEED TO START WITH WHITE; REVISE THIS TEXT TO START WITH A PARAGRAPH ON HISPANIC AND A PARAGRAPH ON ASIAN -- OR VICE VERSA; THOSE TWO ARE THE LEAD; AND THEN CAN TALK ABOUT WHITE AND BLACK.
+]Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
 @fig-la-rq2-midSAT-combo shows all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students (across both no college and some college parents) relative to the overall pool.
-The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%).
+The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%) [DON'T MENTION THIS; THIS IS 6 STUDENTS!].
 Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).
 
-Geomarkets contributions to first-generation college students was mixed across other racial/ethnic categories in @fig-la-rq2-midSAT-combo.
+[THE LEAD IS EAST LA WHICH HAS 251 ASIAN STUDENTS AND IS WAY DIFFERENT THAN ANY OTHE IN FIRSTGEN; AND THEN WILSHIRE IS SORTA CLOSE] Geomarkts contributions to first-generation college students was mixed across other racial/ethnic categories in @fig-la-rq2-midSAT-combo.
 For Asian prospects (11% first-generation no college, 21% first-generation some college, and 68% not first-generation), the Hollywood & Wilshire (12%) and East Los Angeles (36%) Geomarkets contributed larger proportions of first-generation college students whose parents did not attend college.
-The San Fernando Valley- East (28%), Hollywood & Wilshire (35%), East Los Angeles (25%), South and South Central Los Angeles (44%), Long Beach (27%), and Covina and West Covina (25%) Geomarkets also contributed larger proportions of first-generation college students whose parents attended college but did not complete their their degree.
+The San Fernando Valley- East (28%), Hollywood & Wilshire (35%), East Los Angeles (25%), South and South Central Los Angeles (44%) [ONLY 9 STUDENTS!], Long Beach (27%), and Covina and West Covina (25%) Geomarkets also contributed larger proportions of first-generation college students whose parents attended college but did not complete their their degree.
 
@@
 
-Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
+[NOTE SURE YOU NEED THIS PARAGRAPH]Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
 Affluent Geomarkets like San Fernando Valley–West and West Los Angeles contribute an overrepresentation of White, Asian, and not-first generation college student prospects, the magnitude of which increases substantially as score ranges increase (see Appendix @fig-la-rq2-lowSAT-race and @fig-la-rq2-highSAT-race).

### 2025-05-13 text/manuscript-full-KS.qmd
front end revisions

@@
 
+```{=html}
 <!--Reactivity, discipline, and authority are three interrelated themes that describe the effects of quantification.
@@
 In a process akin to sports analytics in Moneyball \[CITE MICHAEL LEWIS\], we suggest that market research replaces the wisdom and authority of local admissions counselors.    -->
-
+```
+```{=html}
 <!--The analysis of UK school "league tables" by @RN5005 shows how quantification can be a mechanism of social reproduction.
@@
 -->
-
+```
 ```{=html}
@@
 ```
-
 <!--# REMOVED FROM PARAGRAPH BELOW: The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4772; @RN4849].
@@
 
+```{=html}
 <!-- SAT test-takers were categorized into one of four different market segments -- local, in-state, regional, and national -- based on SAT score-sending behavior.
@@
 -->
-
+```
 The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
@@
 
-@RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.
+Quantification has reactivity, discipline, and authority effects to the extent that stakeholders care about the numbers [@RN5006].
+Reactivity is the idea that salient quantitative measures cause people and organizations to change their behavior.
 Market research suggests that EPS was highly salient.
 @RN5021 reports that in 1995, 37% of 4-year publics and 49% of 4-year privates used EPS, while 41% of 4-year publics and 16% of 4-year privates used ACT's market analysis service product.
-
+<!--#  @RN5006 argue that quantification has effects to the extent that stakeholders care about the numbers.-->
+
+Quantification disciplines actors to react in particular ways.
 EPS software may "discipline" [@RN5007] colleges to approach recruiting in a manner consistent with the Market Segment Model.
@@
 
-EPS may also weaken the authority of local decision-makers.
+The effects of quantification also lead to changes in decision-making authority, frequently reducing the discretion of local decision-makers.
 @RN4982 sought to develop a concrete, data-driven framework -- The Market Segment Model -- that replicates the aggregate knowledge of local admissions officers.
@@
 
-Although we cannot gain access to EPS software, Geomarkets are fundamental to EPS software and (arguably) to the organization of recruiting territories for many colleges.
-Therefore, we are interested in the extent to which this building block is associated with race and class.
-Research question 1 asks, what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
-We focus on metropolitan areas that are associated with three or more Geomarkets.
-
-We expect substantial socioeconomic inequality between geomarkets in large metropolitan areas.
-Given the extent of class- and race-based residential segregation in the U.S., it would be surprising to not observe such inequality.
-Moreover, @RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards identifying geographic areas that differed from one another in terms of class composition.
+Since @RN4982 developed The Market Segment Model, Geomarkets have been central to EPS software and play a broader role in shaping how many colleges organize recruiting territories.
+We therefore frame our investigation of Geomarkets as a case study of quantification to explore the extent to which this geographical building block is associated with race and class.
+We first address (RQ1) what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time.
+<!--# Although we cannot gain access to EPS software, by focusing on metropolitan areas that are associated with three or more Geomarkets-->
+
+@RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards identifying geographic areas that differed from one another in terms of class composition.
+Given the extent of class-based residential segregation in the U.S., we therefore expect substantial socioeconomic inequality between Geomarkets in large metropolitan areas.
 
@@
 ```
-We also expect substantial racial inequality between geomarkets in large metropolitan areas.
+We also expect substantial racial inequality between Geomarkets within metropolitan areas.
 Interestingly, the Market Segment Model is explicitly based on socioeconomic stratification, but @RN4982 do not mention race.
 U.S.
-cities are characterized be extreme historic and contemporary residential racial segregation [@RN4779].
-Structures built upon racialized structures are racialized structures [@RN4786].
+cities are characterized by extreme historic and contemporary residential racial segregation [@RN4779].
 Unless designers intentionally consider racial segregation, selection devices that categorize people based on geographic location are likely to reproduce historical race-based inequality in opportunity [@RN4975].
@@
 Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
+<!--#  Structures built upon racialized structures are racialized structures [RN4786].-->
 
@@
 <!-- By 1984, College Board Geomarkets were incorporated as a search filter in the College Board Student Search Service \[CITE\].
---> A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^10] Historically, dominant list vendors are college board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+--> Geomarkets also function as a student list search filter for the College Board's Student Search Service.
+A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^10] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
 
@@
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the *Student Search Service*, for example the Geodemographic Segment filters.[^13] This manuscript focuses on Geomarkets.
-In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
+College Board also created geographic borders that subsequently became filters in the *Student Search Service*, such as Geomarkets and Geodemographic Segment filters.[^13] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda Country, excluding Oakland." A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.
-This article extends @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
+Our study therefore extends @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
+<!--# This manuscript focuses on Geomarkets.
+-->
 
@@
 By contrast, the student list purchases of local state colleges and community colleges, which rely on enrollment from students in the "local" and "in-state," should focus on nearby Geomarkets with large numbers of low-SES students.
-Our analyses attempt to recreate the strategy recommended by @RN4982 by showing who is included/excluded when student list purchases filter on particular Geomarkets.
-Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access.
-Furthermore, we anticipate that the students of color from affluent Geomarkets will tend to have college-educated parents and be enrolled in affluent, predominantly white schools.
+Our analyses attempt to recreate the strategy recommended by @RN4982 by showing who is included versus excluded when student list purchases filter on particular Geomarkets.
+Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access, and vice-versa.
+<!--#  Furthermore, we anticipate that the students of color from affluent Geomarkets will tend to have college-educated parents and be enrolled in affluent, predominantly white schools.-->
 

### 2025-05-14 text/back_end_KS.qmd
dallas revisions

@@
 
-<!--# DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES!
---> @fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles.
+<!--# DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES! -->
+
+@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles.
 At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
@@
 
-[WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]
-@fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
+<!--#  [WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]--> @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
 In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
@@
 In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
-Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian residents.
 
-These shifts in racial and ethnic composition within and across Geomarkets became more pronounced between 2000 and 2020.
+By 2000, the City of Dallas was no longer a predominantly White Geomarket, with a population that was approximately 24% Black, 30% Hispanic, and 3% Asian.
+Over the next two decades, the Hispanic population grew steadily, contributing to greater diversity.
+By 2020, the City of Dallas Geomarket had the lowest proportions of White and Asian residents across the Dallas-Fort Worth Metropolitan area, with approximately 21% Black, 38% Hispanic, and 2% Asian residents.
+
+These shifts in racial and ethnic composition also became more pronounced for other Geomarkets between 2000 and 2020.
 The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents.
@@
 In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian).
-Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
+<!--#  Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
+-->
 
-[CITY OF DALLAS!]Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
+<!--# [CITY OF DALLAS!] --> Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
 Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
@@
 By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000).
-In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable and continued to rank at the lower levels of affluence within the metropolitan area.
-\[KS: NEED TO CHECK THIS; I THINK CITY OF DALLAS AND CITY OF FORT WORTH ARE MIXED UP BETWEEN GRAPH VERSUS TABLE\]
 
-[DON'T KNOW IF WE NEED PARAGRAPH ON POVERTY + EDUCATIONAL ATTAINMENT].
-@fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
+In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable through 2020.
+As a result, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 median household incomes, respectively.
+@fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty and educational attainment within Geomarkets over time, revealing similar dynamics across Geomarkets to those observed in median household income.
+Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
+
+<!--#  [DON'T KNOW IF WE NEED PARAGRAPH ON POVERTY + EDUCATIONAL ATTAINMENT] @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
+Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
 In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets.
@@
 Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
+.-->
 
-[DON'T THINK YOU NEED SUMMARY PARAGRAPHS LIKE THIS WITHIN SECTIONS FOR THIS METRO]The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets.
+<!--#  [DON'T THINK YOU NEED SUMMARY PARAGRAPHS LIKE THIS WITHIN SECTIONS FOR THIS METRO]The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets.
 While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
@@
 Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
+-->
 
@@
 Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree.
-Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportinate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
+Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportionate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
 
@@
 Furthermore, first generation status differed substantially across Geomarkets, especially for Hispanic and Asian students.
-For Hispanic students, more than 50% of prospects from the City of Dallas, City of Forth Worth, Arlington \& Grand Prairie, and Dallas County are first-generation college students.
+For Hispanic students, more than 50% of prospects from the City of Dallas, City of Forth Worth, Arlington & Grand Prairie, and Dallas County are first-generation college students.
 By contrast, for the more affluent Collin & Rockwall Geomarket and the West of Dallas/Fort Worth Metroplex Geomarkets, less than 30% of prospects were first-generation.
 For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
-By contrast, the more affluent Collin & Rockwall sent more Asian prospects (285) than any other Geomarket, but about 87% of these students had parents with a BA.
+By contrast, the more affluent Collin & Rockwall sent more Asian prospects (285) than any other Geomarket, but about 87% of these prospects were not first-generation college students.
 
+```{=html}
 <!-- I ADDED THE ABOVE PARAGRAPH AND CUT THIS PARAGRAPH TO GIVE THE LEDE MORE PUNCH 
@@
 -->
-
+```
 Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions.
@@
 Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
-[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
+\[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA\] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
 By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%).
@@
 
-[I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST. START TEHRE AND THEN GET TO OTHER RESULTS.]
-@fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020.
+\[I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST.
+START TEHRE AND THEN GET TO OTHER RESULTS.\] @fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020.
 While median income increased across all of the metropolitan area over the three census periods, some shifts occurred in the relative socioeconomic levels across Geomarkets.
@@
 
-[THINK YOU CAN CUT THIS PARAGRAPH] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
+\[THINK YOU CAN CUT THIS PARAGRAPH\] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
 From 1980 to 2020, the region’s population expanded significantly leading to overall declines in the White population and increases in Hispanic and Asian populations across Geomarkets.
@@
 
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2019-2020 [NOT JUST 2019-20.
-REVISE] high school graduating class, and by PSAT scores.
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2019-2020 \[NOT JUST 2019-20.
+REVISE\] high school graduating class, and by PSAT scores.
 Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.
@@
 The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
-[THE LEDE IS SOUTH CENTRAL AND EAST LA.
-START THERE!].
+\[THE LEDE IS SOUTH CENTRAL AND EAST LA.
+START THERE!\].
 However, the racial/ethnic composition of prospects become less proportional to overall metropolitan patterns by Geomarket as PSAT scores increase.
@@
 This pattern becomes most pronounced at the highest PSAT score prospect pools.
-[I DON'T THINK THAT BLACK AND LATINX STUDENTS HAVING LOW TEST SCORES IS THE STORY WE ARE TRYING TO TELL.
+\[I DON'T THINK THAT BLACK AND LATINX STUDENTS HAVING LOW TEST SCORES IS THE STORY WE ARE TRYING TO TELL.
 THAT STORY HAS BEEN TOLD.
 THE MORE IMPORTANT THING IS, WITHIN SOME TEST THRESHOLD, HOW DOES THE RACIAL COMPOSITION VARY ACROSS GEOMARKETS.
-LIKE EVEN WITHIN THE MID RANGE, SOUTH CENTRAL, EAST LA, AND INLAND EMPIRE STAND OUT] <!-- This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
+LIKE EVEN WITHIN THE MID RANGE, SOUTH CENTRAL, EAST LA, AND INLAND EMPIRE STAND OUT\] <!-- This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
 @fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of residents within Geomarkets.
@@
 
-[I THINK CLEANER IF YOU JUST HAVE A SMALL PARAGRAPH FOR HISPANIC AND A SMALL PARAGRAPH FOR BLACK.
+\[I THINK CLEANER IF YOU JUST HAVE A SMALL PARAGRAPH FOR HISPANIC AND A SMALL PARAGRAPH FOR BLACK.
 I THINK ONE LEDE ACROSS THESE RESULTS IS RIVERSIDE AS AN ENGINE FOR BLACK/LATINX AND FOR FIRST-GEN STUDENTS.
-I GET THAT THESE TWO POINTS CONTRADICT EACH OTHER].
+I GET THAT THESE TWO POINTS CONTRADICT EACH OTHER\].
 Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race.
@@
 
-[I MOVED ASIAN] For Asian students, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger shares, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
+\[I MOVED ASIAN\] For Asian students, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger shares, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
 
@@
 Only 889 of the more than 8,000 total prospects (11%) in this middle PSAT score range (1190-1260) pool were first-generation college students whose parents did not attend college.
-[I'M NOT SURE WHERE YOU ARE GETTING THESE NUMBERS FROM; THIS IS THE WRONG FIGURE I THINK.
-AND WHEN YOU REVISE, THE LEDES ARE RIVSERSIDE, EAST LA AND, SOUTH CENTRAL (ONLY 2% OF PROSPECTS BUT 10% OF NO-COLLEGE)].
+\[I'M NOT SURE WHERE YOU ARE GETTING THESE NUMBERS FROM; THIS IS THE WRONG FIGURE I THINK.
+AND WHEN YOU REVISE, THE LEDES ARE RIVSERSIDE, EAST LA AND, SOUTH CENTRAL (ONLY 2% OF PROSPECTS BUT 10% OF NO-COLLEGE)\].
 Only three of the 11 Geomarkets contributed substantially larger shares of these first-generation prospects: East Los Angeles (35%), South and South Central Los Angeles (47%), and Riverside, San Bernardino, and Ontario (15%).
@@
 
-[I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW]
-Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
+[I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
 [TAKE A TIP FROM THE CRIT QUANT FOLKS; YOU DON'T ALWAYS NEED TO START WITH WHITE; REVISE THIS TEXT TO START WITH A PARAGRAPH ON HISPANIC AND A PARAGRAPH ON ASIAN -- OR VICE VERSA; THOSE TWO ARE THE LEAD; AND THEN CAN TALK ABOUT WHITE AND BLACK.
@@
 @fig-la-rq2-midSAT-combo shows all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students (across both no college and some college parents) relative to the overall pool.
-The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%) [DON'T MENTION THIS; THIS IS 6 STUDENTS!].
+The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%) \[DON'T MENTION THIS; THIS IS 6 STUDENTS!\].
 Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).
@@
 
-[NOTE SURE YOU NEED THIS PARAGRAPH]Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
+\[NOTE SURE YOU NEED THIS PARAGRAPH\]Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
 Affluent Geomarkets like San Fernando Valley–West and West Los Angeles contribute an overrepresentation of White, Asian, and not-first generation college student prospects, the magnitude of which increases substantially as score ranges increase (see Appendix @fig-la-rq2-lowSAT-race and @fig-la-rq2-highSAT-race).

### 2025-05-14 text/back_end_KS.qmd
LA revisions

@@
 As depicted in @fig-la-maps, Geomarkets closely follow this overall metropolitan model.
-Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring of made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
+Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
 The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland, comprising the Inland Empire.
@@
 The population growth in the Los Angeles metropolitan area increased from approximately 8 Million in 1980 to more than 12 Million in 2020.
-@fig-dallas-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model.
+@fig-la-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model.
 The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernadino, and Ontario.
@@
 
-\@fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020.
+@fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020.
 Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
@@
 
-\[I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST.
-START TEHRE AND THEN GET TO OTHER RESULTS.\] @fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020.
-While median income increased across all of the metropolitan area over the three census periods, some shifts occurred in the relative socioeconomic levels across Geomarkets.
-Many Geomarkets kept their relative levels of affluence from 1980 to 2020.
-For instance, the Covina and West Covina, South Bay, and San Fernando Valley remained among the most affluent, whereas the San Fernando Valley East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario Geomarkets remained at middle ranges of affluence and the Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets remained at the lowest levels of affluence across the metropolitan area during this period.
-The West Los Angeles and West Beach Geomarket demonstrated the most substantial increase in median household income, rising from \$112,000 in 1980 to \$132,000 in 2000, and reaching \$137,000 by 2020.
-This sustained growth positioned it as the most affluent Geomarket in the Los Angeles metropolitan area.
+<!--# [I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST.
+START TEHRE AND THEN GET TO OTHER RESULTS.] --> @fig-la-rq1-ses presents trends in income, poverty, and educational attainment in the Los Angeles metropolitan area from 1980 to 2020, demonstrating the South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire Geomarkets remained at the lowest socioeconomic levels across these indicators.
+For instance, the median household income in the South and South Central Los Angeles Geomarket increased modestly from \$57,000 in 1980 to \$59,000 in 2000, reaching \$67,000 by 2020.
+In comparison, East Los Angeles and Hollywood and Wilshire recorded slightly higher median incomes in 2020 at \$76,000 and \$84,000, respectively.
+These figures remain substantially lower than that of the most affluent Geomarket, West Los Angeles and West Beach, where the median household income reached \$137,000 in 2020.
 
-@fig-la-rq1-ses also illustrates shifting levels of poverty across Geomarkets in the Los Angeles metropolitan area over time.
-In 1980, poverty rates were relatively low (less than 12%) for all Geomarkets except for East Los Angeles (17%), South and South Central Los Angeles (24%), and Hollywood and Wilshire (18%).
+@fig-la-rq1-ses also illustrates parallel trends in poverty and educational attainment over time, with Hollywood and Wilshire, East Los Angeles, South and South Central Los Angeles Geomarkets consistenly ranking at lower levels across these socioeconomic indicators.
+Moreover, the disparities between these three Geomarkets and all others in the Los Angeles metropolitan area widened from 1980 to 2020.
+
+<!--# In 1980, poverty rates were relatively low (less than 12%) for all Geomarkets except for East Los Angeles (17%), South and South Central Los Angeles (24%), and Hollywood and Wilshire (18%).
 However, poverty rates declined in South and South Central Los Angeles (22%), remained stable in East Los Angeles and Hollywood and Wilshire, and increased in all other Geomarkets.
@@
 This includes four Geomarkets with poverty rates less than 10% (San Fernando Valley- West, West Los Angeles and West Beach, South Bay, and Covina and West Covina), four Geomarkets with poverty rates ranging from 11%-15% (San Fernando Valley- East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario), and three Geomarkets with poverty rates ranging from 16%-22% (Hollywood and Wilshire, East Los Angeles, South and South Central Los Angeles).
+-->
 
-Lastly, @fig-la-rq1-ses shows a widening disparity in educational attainment across the Los Angeles metropolitan area.
+<!--# Lastly, @fig-la-rq1-ses shows a widening disparity in educational attainment across the Los Angeles metropolitan area.
 Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the West Los Angeles and West Beach Geomarket experienced a substantially greater increase - from 35% in 1980 to 66% in 2020- widening the gap with other Geomarkets in terms of educational attainment.
+-->
 
-\[THINK YOU CAN CUT THIS PARAGRAPH\] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
+<!--#  [THINK YOU CAN CUT THIS PARAGRAPH] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
 From 1980 to 2020, the region’s population expanded significantly leading to overall declines in the White population and increases in Hispanic and Asian populations across Geomarkets.
@@
 Meanwhile, Geomarkets such as West Los Angeles and West Beach became increasingly affluent and highly educated, highlighting growing socioeconomic and educational divides.
-Given these demographic shifts within the context of a multi-nuclei urban model, the Los Angeles metropolitan area presents a distinctive Geomarket landscape case in which racial/ethnic enclaves across a range of various socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.
+Given these demographic shifts within the context of a multi-nuclei urban model, the Los Angeles metropolitan area presents a distinctive Geomarket landscape case in which racial/ethnic enclaves across a range of various socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.-->
 
@@
 
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2019-2020 \[NOT JUST 2019-20.
-REVISE\] high school graduating class, and by PSAT scores.
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2020-2022 high school graduating classes, and by PSAT scores.
 Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.
@@
 @fig-la-rq2-race presents the racial/ethnic composition of purchased student profiles.
-The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
-\[THE LEDE IS SOUTH CENTRAL AND EAST LA.
-START THERE!\].
+For the more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180), South and South Central Los Angeles (n=1,031) and East Los Angeles (n=871) Geomarket pools stand out as having an overwhelming majority of students of color, 98% and 95% respectively.
+Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (89%) and East Los Angeles (64%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (6%) and larger share of Asian prospects for East Los Angeles (30%).
+
+Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic and Black student prospects, while East Los Angeles stands out for its substantial share of Asian prospects.
+
+<!--#  very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
+[THE LEDE IS SOUTH CENTRAL AND EAST LA.
+START THERE!].
 However, the racial/ethnic composition of prospects become less proportional to overall metropolitan patterns by Geomarket as PSAT scores increase.
@@
 This pattern becomes most pronounced at the highest PSAT score prospect pools.
-\[I DON'T THINK THAT BLACK AND LATINX STUDENTS HAVING LOW TEST SCORES IS THE STORY WE ARE TRYING TO TELL.
+[I DON'T THINK THAT BLACK AND LATINX STUDENTS HAVING LOW TEST SCORES IS THE STORY WE ARE TRYING TO TELL.
 THAT STORY HAS BEEN TOLD.
 THE MORE IMPORTANT THING IS, WITHIN SOME TEST THRESHOLD, HOW DOES THE RACIAL COMPOSITION VARY ACROSS GEOMARKETS.
-LIKE EVEN WITHIN THE MID RANGE, SOUTH CENTRAL, EAST LA, AND INLAND EMPIRE STAND OUT\] <!-- This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
+LIKE EVEN WITHIN THE MID RANGE, SOUTH CENTRAL, EAST LA, AND INLAND EMPIRE STAND OUT] <!-- This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
 @fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of residents within Geomarkets.
@@
 
-@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area.
-For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.
+@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area, demonstrating South and South Central Los Angeles (7% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (25% versus 17% for all prospects) Geomarkets contributed larger proportions of Black prospects relative to their contributions to the overall prospect pool in the Los Angeles metropolitan area.
+<!--#  For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.-->
 
-\[I THINK CLEANER IF YOU JUST HAVE A SMALL PARAGRAPH FOR HISPANIC AND A SMALL PARAGRAPH FOR BLACK.
+For Hispanic prospects, South and South Central Los Angeles (8% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (24% versus 17% for all prospects) also contributed a disproportinately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
+
+<!--# [I THINK CLEANER IF YOU JUST HAVE A SMALL PARAGRAPH FOR HISPANIC AND A SMALL PARAGRAPH FOR BLACK.
 I THINK ONE LEDE ACROSS THESE RESULTS IS RIVERSIDE AS AN ENGINE FOR BLACK/LATINX AND FOR FIRST-GEN STUDENTS.
-I GET THAT THESE TWO POINTS CONTRADICT EACH OTHER\].
+I GET THAT THESE TWO POINTS CONTRADICT EACH OTHER].
 Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race.
@@
 All other Geomarkets contributed nearly equal shares of Black and Hispanic prospects.
+-->
 
-\[I MOVED ASIAN\] For Asian students, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger shares, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
+For Asian students, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger shares, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
+
+These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Los Angeles metropolitan area using Geomarket filters that exclude lower-income Geomarkets such as South and South Central Los Angeles and Riverside, San Bernadino, and Ontario Geomarkets would result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
+Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.
 

### 2025-05-14 text/back_end_KS.qmd
LA revisions

@@
 
-For Asian students, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger shares, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
+In comparison, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian prospects relative to their contributions to the overall prospect pool.
 
@@
 
-Contributions of each Geomarket to the pool of prospects from the Los Angeles metropolitan area based on first-generation college student status are presented in @fig-la-rq2-midSAT-ses [THIS SENTENCE IS UGLY.
-MAKE IT MORE LIKE FIG X SHOWS THE CONTRIBUTIONS EACH GEOMARKET MAKES BY FIRST GENERATION STATUS GROUP].
-Only 889 of the more than 8,000 total prospects (11%) in this middle PSAT score range (1190-1260) pool were first-generation college students whose parents did not attend college.
-\[I'M NOT SURE WHERE YOU ARE GETTING THESE NUMBERS FROM; THIS IS THE WRONG FIGURE I THINK.
-AND WHEN YOU REVISE, THE LEDES ARE RIVSERSIDE, EAST LA AND, SOUTH CENTRAL (ONLY 2% OF PROSPECTS BUT 10% OF NO-COLLEGE)\].
-Only three of the 11 Geomarkets contributed substantially larger shares of these first-generation prospects: East Los Angeles (35%), South and South Central Los Angeles (47%), and Riverside, San Bernardino, and Ontario (15%).
-However, disproportionate contributions become more spread across Geomarkets for the nearly 1,436 of the more than 8,000 total prospects (18%) who were first-generation college students whose parents attended but did not complete their degree.
-Overall, six of the 11 Geomarkets contributed larger shares of these prospects: Hollywood & Wilshire (25%), East Los Angeles (26%), South and South Central Los Angeles (28%), Long Beach (25%), Covina and West Covina (22%), and Riverside, San Bernardino, and Ontario (22%).
-This leads to the remaining Geomarkets - San Fernando Valley- West (80%), San Fernando Valley- East (78%), Glendale and Pasadena (77%), West Los Angeles & West Beach (87%), South Bay (81%) - contributing disproportionately larger shares of not first-generation college students relative to the overall prospect pool (71%).
+<!--# [I'M NOT SURE WHERE YOU ARE GETTING THESE NUMBERS FROM; THIS IS THE WRONG FIGURE I THINK.
+AND WHEN YOU REVISE, THE LEDES ARE RIVSERSIDE, EAST LA AND, SOUTH CENTRAL (ONLY 2% OF PROSPECTS BUT 10% OF NO-COLLEGE) KS-- weird...
+overall substantive rankings are correct but my %s are all off].
+--> @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
+For students whose parents did not attend college, East Los Angeles (16% versus 5% of all prospects), South and South Central Los Angeles (10% versus 2% of all prospects), and Riverside, San Bernardino, and Ontario (24% versus 17% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
+These three Geomarkets also made disproportionately larger contributions of first-generation college student prospects whose parents attended but did not complete their degree, as well as Hollywood & Wilshire (7% versus 5% of all prospects), Long Beach (7% versus 5% of all prospects), and Covina and West Covina (13% versus 10% of all prospects).
 
-[I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
-[TAKE A TIP FROM THE CRIT QUANT FOLKS; YOU DON'T ALWAYS NEED TO START WITH WHITE; REVISE THIS TEXT TO START WITH A PARAGRAPH ON HISPANIC AND A PARAGRAPH ON ASIAN -- OR VICE VERSA; THOSE TWO ARE THE LEAD; AND THEN CAN TALK ABOUT WHITE AND BLACK.
+<!--#  [I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] --> Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
+
+For Hispanic prospects, the South and South Central Los Angeles Geomarket stands out as providing the most substantial share of first-generation college students.
+Across all Hispanic prospects in the middle range PSAT orders, 25% were first-generation college students whose parents did not attend college, 25% were first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
+In comparison, the South and South Central Los Angeles Geomarket contributed much larger disproportionate share of Hispanic first-generation college students whose parents did not attend college (54%), with the Hollywood & Wilshire (30%), East Los Angeles (39%), and Riverside, San Bernardino, and Ontario (31%) also contributing larger but more modest disproportionate shares.
+The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
+
+<!--# [TAKE A TIP FROM THE CRIT QUANT FOLKS; YOU DON'T ALWAYS NEED TO START WITH WHITE; REVISE THIS TEXT TO START WITH A PARAGRAPH ON HISPANIC AND A PARAGRAPH ON ASIAN -- OR VICE VERSA; THOSE TWO ARE THE LEAD; AND THEN CAN TALK ABOUT WHITE AND BLACK.
 ]Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
 @fig-la-rq2-midSAT-combo shows all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students (across both no college and some college parents) relative to the overall pool.
-The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%) \[DON'T MENTION THIS; THIS IS 6 STUDENTS!\].
-Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).
+The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%) [DON'T MENTION THIS; THIS IS 6 STUDENTS!].
+Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).-->
 
-[THE LEAD IS EAST LA WHICH HAS 251 ASIAN STUDENTS AND IS WAY DIFFERENT THAN ANY OTHE IN FIRSTGEN; AND THEN WILSHIRE IS SORTA CLOSE] Geomarkts contributions to first-generation college students was mixed across other racial/ethnic categories in @fig-la-rq2-midSAT-combo.
-For Asian prospects (11% first-generation no college, 21% first-generation some college, and 68% not first-generation), the Hollywood & Wilshire (12%) and East Los Angeles (36%) Geomarkets contributed larger proportions of first-generation college students whose parents did not attend college.
-The San Fernando Valley- East (28%), Hollywood & Wilshire (35%), East Los Angeles (25%), South and South Central Los Angeles (44%) [ONLY 9 STUDENTS!], Long Beach (27%), and Covina and West Covina (25%) Geomarkets also contributed larger proportions of first-generation college students whose parents attended college but did not complete their their degree.
+<!--# [THE LEAD IS EAST LA WHICH HAS 251 ASIAN STUDENTS AND IS WAY DIFFERENT THAN ANY OTHE IN FIRSTGEN; AND THEN WILSHIRE IS SORTA CLOSE]--> Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 21% were first-generation college students whose parents attended some college but did not complete their degree, and 68% were not first-generation college students.
+@fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 251), a significant proportion of whom are first-generation college students (35%) whose parents did not attend college.
+The Hollywood & Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (35%).
 
-For Black prospects (5% first-generation no college, 20% first-generation some college, and 74% not first-generation), Geomarkets that disproportionately larger proportions of first-generation college students whose parents did not attend college include South Bay (11%), South and South Central Los Angeles (21%), Long Beach (7%), and Covina and West Covina (9%).
-The West Los Angeles and West Beach (26%), Hollywood & Wilshire (36%), East Los Angeles (33%), Long Beach (40%), and Covina and West Covina (27%) Geomarkets also contributed larger proportions of Black first-generation college students whose parents attended college but did not complete their their degree.
+While the pool of Black prospects is relatively small (n=186) in @fig-la-rq2-midSAT-combo , some patterns emerge by first-generation status.
+Overall, approximately 5% of Black prospects were first-generation college students-no college, 20% were first-generation college students-some college, and 74% not first-generation).
+The South and South Central Los Angeles Geomarket stands out for contributing a disproportionately high share (21%) of Black first-generation college students whose parents did not attend college, despite representing a relatively smaller overall number.
+Additionally, the Long Beach Geomarket is notable for its higher proportion (40%) of Black first-generation students whose parents attended college but did not complete a degree.
 
-Finally, Hispanic prospects in this middle range PSAT order were on average 25% first-generation college students whose parents did not attend college, 25% first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
-The Hollywood & Wilshire (30%), East Los Angeles (39%), South and South Central Los Angeles (54%), and Riverside, San Bernardino, and Ontario (31%) contributed disproportionate shares of Hispanic first-generation college students whose parents did not attend college.
-The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
-
-\[NOTE SURE YOU NEED THIS PARAGRAPH\]Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
+<!--#  [NOTE SURE YOU NEED THIS PARAGRAPH]Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
 Affluent Geomarkets like San Fernando Valley–West and West Los Angeles contribute an overrepresentation of White, Asian, and not-first generation college student prospects, the magnitude of which increases substantially as score ranges increase (see Appendix @fig-la-rq2-lowSAT-race and @fig-la-rq2-highSAT-race).
 On the other hand, more diverse and historically low-income Geomarkets, like South and South Central Los Angeles and Riverside, San Bernardino, and Ontario, contribute disproportionately more Black, Hispanic, and first-generation college student prospects.
-Similar to Chicago and Dallas-Fort Worth, Los Angeles further solidifies the pattern of some of the lowest-income Geomarkets consistently playing crucial roles in fostering college access for first-generation Asian (East Los Angeles), Black (South and South Central Los Angeles), and Hispanic college students (South and South Central Los Angeles, East Los Angeles, Hollywood and Wilshire, and Riverside, San Berdardino, and Ontario).
+Similar to Chicago and Dallas-Fort Worth, Los Angeles further solidifies the pattern of some of the lowest-income Geomarkets consistently playing crucial roles in fostering college access for first-generation Asian (East Los Angeles), Black (South and South Central Los Angeles), and Hispanic college students (South and South Central Los Angeles, East Los Angeles, Hollywood and Wilshire, and Riverside, San Berdardino, and Ontario).-->
 

### 2025-05-14 text/manuscript-full-KS.qmd
inserted new parts into whole, deleted old parts

@@
 \pagenumbering{arabic}
-```
+
 # Introduction
@@
 ```
+
 Sociology offers several perspectives to describe the process of sorting students into colleges as a process that tends toward social reproduction [@RN4873; @RN2377].
@@
 ```
+
 # Enrollment Management and Quantification
@@
 ```
+
 **Quantification**
@@
 ```
+
 ```{=html}
@@
 ```
+
 <!--# REMOVED FROM PARAGRAPH BELOW: The interdisciplinary literature on quantification [@RN5008] includes important contributions from the field of critical data studies [e.g., @RN4772; @RN4849].
@@
 ```
+
 # The Market Segment Model and College Board Geomarkets
@@
 ```
+
 **Creating Geomarkets and the Market Segment Model.**
@@
 ```
+
 The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
@@
 ```
+
 Whereas @RN4982 identified 143 Geomarkets covering the New England, Middle States, and South region, EPS created Geomarkets for the remaining U.S.
@@
 ```
+
 We also expect substantial racial inequality between Geomarkets within metropolitan areas.
@@
 
-For the first research question, we use data from the 1980 and 2000 Decennial Census as well as 5-year estimates from the 2020 American Community Survey (ACS).
-For the 1980 Decennial Census, we specifically draw on Summary Tape File 1 (STF1) for variables about race and Hispanic origin collected via “short form” questions answered by all households and Summary Tape File 3 (STF3) for variables about socioeconomic characteristics (income, education, and poverty) collected via the "long form" questionnaire completed by a sample of households.
+We use data from the 1980 and the 2000 Decennial Census as well as 5-year estimates from the 2020 American Community Survey (ACS).
+For the 1980 Decennial Census, we draw from Summary Tape File 1 (STF1) for variables about race and Hispanic origin collected via "short form" questions answered by all households.
+We draw from Summary Tape File 3 (STF3) for variables about socioeconomic characteristics (income, education, and poverty) collected via the "long form" questionnaire completed by a sample of roughly 20% of all households.
 The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-To examine how Geomarket characteristics change over time, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty from the 2000 Decennial Census.[^14] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+For the 2000 Decennial Census, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty.[^1] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
 These questions were replaced by ACS.
-Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.[^15] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
-
-[^14]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
-
-[^15]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
+Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.[^2] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+
+[^1]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
+
+[^2]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
 
@@
 Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^16] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^3] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
 Using these zip code-level data, we utilized the `aggregate()` function to create shapefiles for each Geomarket.
 
-[^16]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
+[^3]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
 
 Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
-Census tract shapefiles that intersected with multiple Geomarkets (i.e., a census tract's spatial boundaries were spread across two or more Geomarkets) broken into smaller shapefiles wholly contained within a single Geomarket.
-The variables in these shapefiles were then replaced with the original variable value multiplied by the proportion of land area of the census tract that was contained within the Geomarket.
-For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 1,000 people who identify as Hispanic non-white.
+We split census-tract shapefiles that intersected multiple Geomarkets (i.e., tracts whose boundaries spanned two or more Geomarkets) into smaller shapefiles, each entirely contained within a single Geomarket.
+We replaced each variable in these shapefiles with its original value scaled by the share of the tract’s land area that lay within the Geomarket.
+For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 100 people who identify as Asian, non-Hispanic.
 The partial spatial merge splits this census tract into two observations.
-The observation assigned to Geomarket A reports $.60*1000 = 600$ Hispanic non-white people and the observation assigned to Geomarket B reports $.40*1000 = 400$ Hispanic non-white people.
+The observation assigned to Geomarket A reports $.60*100 = 60$ Asian, non-Hispanic people and the observation assigned to Geomarket B reports $.40*100 = 40$ Asian, non-Hispanic people.
 
 **RQ2**.
-Our study's second research question asks how does the racial and socioeconomic composition of included versus included prospects vary when student list purchases filter on Geomarkets.
+Our second research question asks, how does the racial and socioeconomic composition of included versus excluded prospects vary when student list purchases filter on Geomarkets?
 We answer this question using actual student lists purchased by public universities.
@@
 We narrowed the scope of our request to student lists purchased from College Board, ACT, and the National Research Center for College and University Admissions (NRCCUA), the three largest student list vendors at the time.
-For each student list purchased from 2016 through 2020, we requested two related pieces of data: (1) the order summary, which specifies search criteria for the student list purchase; and (2) the de-identified prospect-level list produced from the search criteria.
+For each student list purchased from 2016 through 2020, we requested two related pieces of data: (1) the order summary, which specifies search criteria for the student list purchase; and (2) the de-identified prospect-level list produced from these search criteria.
 
 The data collection sample for this project was all public universities in California, Illinois, Minnesota, and Texas.
-Utilizing public records requests to obtain public records is a painstaking process.
-Initially, the majority of universities did not respond or denied our request.
-Subsequently, we obtained pro bono representation from four law firms, which substantially increased the success of data collection.
+We chose these states because they were priorities for the two foundations that funded the data collection.
+Utilizing public records requests to obtain quantitative data is a painstaking process, made more difficult by the onset of the Covid Pandemic.
+Initially, the majority of universities ignored or denied our request.
+We established a client relationship with a civil rights legal organization and leveraged this relationship to obtain pro bono representation from four law firms.
+Firm representation substantially increased the success of data collection.
 We also collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
-However, we were unable to obtain legal representation for Texas.
-
-Even with firm representation, data collection remained difficult.
+However, we were unable to obtain legal representation for Texas because firms interested in representing us had conflicts of interest in Texas.
+
+Firm representation substantially increased the success of data collection.
+Nevertheless, data collection remained difficult.
 Some universities provided records that were not usable for quantitative analyses (e.g., summary statistics across multiple orders; or data did not contain important fields).
 Some universities did not provide records based on legitimate legal grounds (e.g., data not in university possession; not required to create records that do not currently exist).
-We learned that many universities outsourced student list purchases to third-party consulting firms.
+We learned that many universities outsourced student list purchases to third-party consulting firms, particularly [EAB](https://eab.com/).
 Unfortunately, we were rarely able to obtain usable data from these universities.
-A small number of universities denied requests based on potentially questionable legal rationale, but we lacked the resources to litigate.
-
-The primary data source for our second research question is therefore student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
-This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Minnesota, and Arizona \[KS Does this still apply?\].
-Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.[^17] The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
-In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.
-Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
-
-[^17]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
-
-In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter, but did purchase prospects from all Geomarkets in a particular metropolitan area.
-For each prospect we know their home zip code and, therefore, which Geomarket they belong to.
-We therefore simulate which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
-
-The variables of interest for student list data are the characteristics of prospects who would have been included versus excluded by our simulations by Geomarket filter.
-Prospect characteristics are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
-College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity nonresponse) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016).
+Several universities denied requests based on questionable legal rationale, but we lacked the resources to litigate.
+
+The primary data source for our second research question is student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
+This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Texas, and Arizona.
+Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.
+The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
+In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.[^4] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
+
+[^4]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
+
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.[^5] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
+Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
+
+[^5]: For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.
+
+The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
+College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) \[CITE\].
 From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
-Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016).[^18] Information about student socioeconomic status is limited.
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].[^6] Information about student socioeconomic status is limited.
 However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
 
-[^18]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
+[^6]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
 Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”
@@
 **Research design**.
-We utilize a multiple, quantitative case study design in which metropolitan areas are cases.
-Case studies are used to gain a deeper understanding of the phenomenon of interest by comparing and contrasting across different settings \[CITE YIN 2001\].
-Our study's research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
-Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics.
+We utilize a multiple, quantitative case study design.
+Metropolitan areas are cases.
+Case studies deepen understanding of phenomena by comparing and contrasting across different settings \[CITE YIN 2001\].
+Our research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
+Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics \[MAKE THIS MORE ABOUT EXPLORING THE MODEL.
+THIS SENTENCE SOUNDS FANCY BUT DOESN'T REALLY SAY ANYTHING\].
 Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts.
-We therefore focus on minimizing the number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
+We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
 Online Appendix XXX presents results for a larger number of metropolitan areas.
 
-Case selection was therefore informed by purposeful sampling to identify “information-rich” cases (Patton, 2002; Yin, 2009) that can provide in-depth knowledge about variation across Geomarkets and overtime.
+Case selection was informed by purposeful sampling to identify "information-rich" cases (Patton, 2002; Yin, 2009)[CITE] that can provide knowledge about variation across Geomarkets and over time.
 We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
-Based on our read of @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
-11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, Philadelphia**,** Phoenix, and Pittsburgh.
+Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area[s] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
+11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia.
 To compare and contrast with classic urban model structures, we also identified alternative urban models such as bimodal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
@@
 
+```{=html}
+
+<!-- [DO WE NEED THIS LEVEL OF DETAIL???]
+
+For example, for purchased profiles who lived in the Chester County Geomarket (PA02) [OR PHILADELPHIA PA05] and identified as Black, where in the county did these students live and which high schools did they attend?
+
+ For example, for purchased profiles who lived in the Chester County Geomarket (PA02) [OR PHILADELPHIA PA05] and identified as Black, where in the county did these students live and which high schools did they attend?
+
+NEXT STEPS: CREATE TABLE FOR PHILLY STUDENT LIST PURCHASES THAT FOCUSES ON FIRST GEN STATUS.
+
+Our conceptual framework suggests that racial and socioeconomic inequality is Geomarket regions is associated with the number of geomarkets in the metropolitan area.
+and is associated with the patterns of racial segregation in the metropolitan area.
+comment more on this.
+
+Below, we use simulations and actual student list purchases that filtered on geomarkets.
+In purchases that filtered on test-score and/or GPA thresholds, we can simluate who would be included and excluded had certain Geomarkets were been selected.
+In purchases that filter on geomarkets, we can get an initial sense of who is included in geomarkets targeted by regional state colleges versus research universities.
+
+-->
+```
 # Findings
@@
 @fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
-The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 in the development of The Market Segment Model.
+The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 [p.
+12].
 Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
@@
 
-These population shifts have contributed to changes in the racial/ethnic composition within and across Geomarkets over time.
-@fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarkets across the three census periods.
+These modest population shifts lead to substantial change in the racial/ethnic composition within and across Geomarkets over time.
+@fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarket across the three census periods.
 In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
 In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
-By 2020, several Geomarkets that were previously predominantly White shifted to a more balanced racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
+By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
 For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
 Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
-
 @fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
-For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
+<!-- For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
 City of Chicago also experienced relatively modest declines in its Black population (37% in 1980 to 33% in 2020), alongside increases in its Hispanic (15% in 1980 to 26% in 2020) and Asian Pacific Islander populations (4% in 2000 to 6% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
+-->
 
@@
 Income, poverty, and educational attainment also created a static socioeconomic divide in 1980, with southern Geormarkets (City of Chicago and South and Southwest Suburbs) characterized by lower affluence and educational attainment compared to the higher socioeconomic standing of the northwestern suburbs.
-However, declining White populations, Hispanic and Asian population growth, and Black outmigration, as well as some narrowing of education and poverty gaps, have contributed to a more stratified Geomarket dynamic in the metropolitan area despite significant disparities remaining.
-The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents) while the the Evanston and North Shore Geomarkets remain predominantly White.
-While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains across Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
+However, declining White populations, Hispanic and Asian population growth, and Black out-migration, as well as narrowing of education and poverty gaps within some Geomarkets, have contributed to a more stratified Geomarket dynamic in the metropolitan area with significant disparities remaining between Geomarkets.
+The City of Chicago Geomarket remains the critical geographical hub across all Geomarkets for racial/ethnic diversity; however, other Geomarkets now also include substantive numbers of residents across racial/ethnic groups (such as South and Southwest Geomarket for Black residents, Chain of Lakes for Hispanic residents, and Northwest Suburbs for Asian residents).
+While some stark socioeconomic gaps within Geomarkets have narrowed since 1980, significant geographical income segregation remains between Geomarkets despite a more clearly defined middle-to-high socioeconomic Geomarket group (Chain of Lakes, Northwest Suburbs, Western Suburbs).
+
+<!--# [I THINK SOME DETAIL CAN BE REDUCED HERE; ALSO, I THINK YOU ARE BURYING THE LEDE THAT NEEDS TO BE EMPHASIZED: THERE IS SUBSTANTIAL RACIAL AND SES INEQUALITY BETWEEN GEOMARKETS AND CHICAGO ESPECIALLY IS POOREST AND MOST NON-WHITE ————-  DALLAS IS POOREST AND MOST NON-WHITE; NOT CHI] -->
 
@@
 To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospects across the entire state of Illinois.
-Two of these orders applied the same filters: SAT score thresholds ranging from 1020 to 1150, GPA ranging from B- to A+, and high school class targeting students from the 2019-2020 and 2020-2021 graduating classes, respectively.
-The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds- two targeted students with relatively middle-range scores (1160 to 1300) and two targeted those with relatively high scores (1310 to 1600).
+The first two orders filter on the same SAT thresholds (ranging from 1020 to 1150), the same GPA thresholds (ranging from B- to A+), and filter on the high school graduating classes of 2020 and 2021 graduating classes, respectively.
+The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds.
+Two orders targeted students with relatively middle-range scores (1160 to 1300) and two orders targeted those with relatively high scores (1310 to 1600).
 Because all six orders included an Illinois state filter, we can utilize prospects from the Chicago metropolitan area within the resulting student lists to examine Geomarket contributions by simulating the application of a Geomarket filter across low-, middle-, and high-range SAT threshold orders.
 
-@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area.
-At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets generally reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
-For example, more than 8,000 of the 16,485 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
-Similar to the larger metropolitan area, the Northshore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects, reflecting the Geomarket dynamics in the metropolitan area population although they yielded fewer overall students.
-Interestingly, while the City of Chicago is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-chicago-rq2-race includes nearly double the proportion of Hispanic (50%) prospects relative to the larger Geomarket resident population (26% Hispanic) in @fig-chicago-rq1-race with a lower proportion of White prospects (14%).
-
-However, the racial/ethnic composition of prospect pools become less similar to the overall metropolitan area as PSAT scores increase.
-Notably, the shares of White and Asian prospects within Geomarkets increase, while Black and Hispanic prospects decrease.
-For instance, Asian prospects over representation in Geomarkets relative to the overall resident population at the 1310-1600 SAT score threshold ranges from a minimum of 4 percentage points in the North Shore Geomarket (17% of prospects versus 13% of residents) to a maximum of 21 percentage points in the Chain of Lakes (28% of prospects versus 7% residents) Geomarket.
-The City of Chicago Geomarket exhibits the largest underrepresentation of Black prospects (7% of prospects versus 33% of residents) while Chain of Lakes exhibits the largest underrepresentation of Hispanic prospects (7% of prospects versus 24% of residents).
+<!--# DISCUSSION OF THIS FIGURE SOMEWHAT BURIES THE LEAD THAT CHICAGO LOOKS TOTALLY DIFFERENT THAN THE OTHERS IN TERMS OF RACIAL COMPOSITION OF PURCHASED PROFILES! -->
+
+@fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles.
+At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
+For example, more than 8,000 of the 15,981 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
+The North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects.
+Unlike any other Geomarket, nearly nine of every 10 prospects from the City of Chicago Geomarket were Hispanic (50%), Black (30%), Asian (5%), or Multiracial (2%).
+
+The City of Chicago Geomarket increasingly becomes the only Geomarket with substantial shares of Black, and Hispanic prospects as PSAT scores increase.
+All Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline.
+For instance, the City of Chicago Geomarket pool is 18% Asian, 7% Black, 21% Hispanic, 5% Multiracial, making it the only Geomarket that is not predominantly White (49%).
+Prospect pools from all other Geomarkets in this high PSAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.4% (North Shore) to 5% (South and Southwest Suburbs) of prospects and 6% (Northwest Suburbs) to 10% (South and Southwest Suburbs) of prospects, respectively.
+<!--# [LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN] -->
 
@@
 ```
-More than half (n=8,256) of all prospects whose contact information was purchased in these middle SAT score orders identified as White.
-@fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects.
-One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects.
+For instance, @fig-chicago-rq2-midSAT-race shows the nearly 1,000 Black students whose contact information was purchased in these middle SAT score orders were largely contributed to the overall pool by the City of Chicago and South and Southwest Suburbs.
+These two Geomarkets contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects.
+On the other hand, all other Geomarkets contributed substantially smaller shares of Black prospects.
+The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
+
+In contrast, these representational patterns by Geomarket are reversed for White and Asian students.
+@fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates proportional to their overall contributions to all included prospects (except City of Chicago).
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects) and Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospectq8s) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+
+<!--# One exception is the City of Chicago Geomarket, which contributed only 7% of White prospects, despite accounting for 17% of all included prospects.
 On the other hand, Western Suburbs and South and Southwest Suburbs Geomarkets 32% and 23% of White prospects while only accounting for 28% and 20% of all included prospects, respectively.
-Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with thier overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects), Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
-
-In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race.
+-->
+
+<!--# MOVED THIS UP In contrast, these representational patterns are reversed for Black and Hispanic prospects in @fig-chicago-rq2-midSAT-race.
 Among the nearly 1,000 Black students included in these orders, the City of Chicago and South and Southwest Suburbs contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects, whereas all other Geomarkets contributed substantially smaller proportions of Black prospects.
 The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
-
-Appendix figures \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race similarly depict Geomarket contributions for pools of prospects resulting from orders targeting lower-range (1020-1150) and high-range (1310-1600) SAT scores.
-Both figures demonstrate how representational Geomarket contributionsacross race/ethnicity, such as City of Chicago disproportionately contributing larger shares of Black and Hispanic prospects, remain consistent even at lower and higher SAT score thresholds.
+-->
+
+These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Chicago metropolitan area using Geomarket filters that exclude the City of Chicago– as observed in some orders collected via our public requests – result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
+Appendix \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race show that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters.
+In both cases, the City of Chicago contributes disproportionately larger shares of Black and Hispanic prospects compared to all other Geomarkets.
+<!--# [REVISE THESE SENTENCES TO HAMMER THE LEDE -- IF WE EXCLUDE CITY OF CHICAGO GEOMAREKT, THE WAY THE ASU ORDER DOES, WE LOSE A DISPROPORTIONATE SHARE OF BLACK AND HISPANIC STUDENTS] -->
 
@@
 In contrast, Geomarkets such as Chain of Lakes (82%), North Shore (94%), Evanston and Skokie (81%), and Western Surburbs (84%) contribute a disproportionately larger share of White prospects who are not first-generation college students.
-
-@fig-chicago-rq2-midSAT-combo indicates these Geomarket patterns persist across other racial/ethnic categories.
-The City of Chicago Geomarket consistently contributed a disproportionately large share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
-However, the magnitude of this disproportionality varied- from a 1-percentage-point difference for Black prospects to a 23 percentage-point difference for Asian prospects.
-Among Hispanic students, contributions to the first-generation prospect pool were notably balanced across Geomarkets, with the exception of North Shore and Evanston and Skokie Geomarkets, which contributed substantially lower shares of first-generation Hispanic students.
-Although Asian and Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs Geomarkets.
+-->
+
+<!--# [MY READ IS THAT THE PATTERNS FOR WHITE ARE NOT THE SAME AS THE PATTERNS FOR ASIAN AND LATINX -- SO I WOULDN'T SAY THESE PATTERNS PERSIST ACROSS OTHER RACIAL/ETHNIC CATEGORIES.
+I WOULD RECOMMEND STARTING WITH ASIAN AND WITH HISPANIC.
+THAT IS THE LEDE; ALSO, THIS IS WHERE YOU COULD REFER TO THE SIMILAR GRAPH FOR PHILADELPHIA AND SAY IT FOLLOWS THE SAME PATTERN; I THINK THIS IS THE CLIMAX AND COOLEST INSIGHT OF THE RESULTS] KS=== I THINK ITS" WEIRD TO INCLUDE PHILLY GRAPH HERE WHEN ITS NOT CASE...WE CAN SPEAK TO THIS IN THE DISCUSSION INSTEAD...
+WE SEE SIMILAR PATTERNS ACROSS OTHER METROS TOO -->
+
 Appendix figures @fig-chicago-rq2-lowSAT-combo and \@fig-chicago-rq2-highSAT-combo show the magnitude of these Geomarket contribution disparities become more pronounced at as SAT score thresholds increase.
@@
 On the other hand, predominantly White and affluent Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, as well as non-first generation college students.
-However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian and Hispanic college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+Given the broader racial/ethnic distribution of the Chicago metropolitan area, student list orders that use Geomarket filters excluding the City of Chicago and including more affluent Geomarkets in targeting high-achieving students may still capture some Asian and Hispanic prospects.
+However, these students are far more unlikely to be first-generation students of color.
+
+<!--# [SO WHAT HAPPENS IF A SCHOOL USES A GEOMARKET FILTER AND EXCLUDES CHICAGO.] -->
 
@@
 
-@fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
+<!--#  [WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]--> @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
 In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
-By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian Pacific Islander), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian Pacific Islander).
+By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian).
 In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
-Notably, by 2000, the City of Dallas was no longer a predominantly White Geomarket, with a racial/ethnic composition of approximately 24% Black, 30% Hispanic, and 3% Asian Pacific Islander residents.
-
-These shifts in racial and ethnic composition within and across Geomarkets became more pronounced between 2000 and 2020.
+
+By 2000, the City of Dallas was no longer a predominantly White Geomarket, with a population that was approximately 24% Black, 30% Hispanic, and 3% Asian.
+Over the next two decades, the Hispanic population grew steadily, contributing to greater diversity.
+By 2020, the City of Dallas Geomarket had the lowest proportions of White and Asian residents across the Dallas-Fort Worth Metropolitan area, with approximately 21% Black, 38% Hispanic, and 2% Asian residents.
+
+These shifts in racial and ethnic composition also became more pronounced for other Geomarkets between 2000 and 2020.
 The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents.
 However, both Geomarkets remained predominantly White (58% and 63%, respectively).
-In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian Pacific Islander), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian Pacific Islander), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian Pacific Islander).
-Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
-
-Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
+In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian).
+<!--#  Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
+-->
+
+<!--# [CITY OF DALLAS!] --> Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
 Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
@@
 By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000).
-In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable and continued to rank at the lower levels of affluence within the metropolitan area.
-\[KS: NEED TO CHECK THIS; I THINK CITY OF DALLAS AND CITY OF FORT WORTH ARE MIXED UP BETWEEN GRAPH VERSUS TABLE\]
-
-@fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
+
+In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable through 2020.
+As a result, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 median household incomes, respectively.
+@fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty and educational attainment within Geomarkets over time, revealing similar dynamics across Geomarkets to those observed in median household income.
+Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
+
+<!--#  [DON'T KNOW IF WE NEED PARAGRAPH ON POVERTY + EDUCATIONAL ATTAINMENT] @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty within Geomarkets over time, revealing shifting dynamics across Geomarkets similar to those observed in median household income.
+Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
 In 1980, poverty rates were comparably low in both Collin & Rockwall Counties and Dallas County, moderate in the Irving, Arlington, and Grand Prarie and West of Dallas/Fort Worth Metroplex Geomarkets, and highest for City of Dallas and City of Fort Worth Geomarkets.
@@
 Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the Collin & Rockwall Counties Geomarket experienced a substantially greater increase - from 23% in 1980 to 50% in 2020- maintaining the gap with other Geomarkets in terms of educational attainment.
-
-The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s.
+.-->
+
+<!--#  [DON'T THINK YOU NEED SUMMARY PARAGRAPHS LIKE THIS WITHIN SECTIONS FOR THIS METRO]The Dallas-Fort Worth metroplex experienced substantial demographic growth since Geomarket boundaries were dilineated by @RN4982 in the early 1980s, leading to an urban landscape with considerable racial/ethnic diversity across multiple Geomarkets (as opposed to concentrated within a single Geomarket like Chicago) but stark socioeconomic disparities between Geomarkets.
 While all six Geomarkets experienced population increases, the most significant growth occurred in the outer suburban areas, reflecting ongoing urban sprawl.
 Such growth is largely attributed to increases in Hispanic, Black, and Asian populations leading to only two of the six Dallas-Fort Worth Geomarkets remaining predominantly White (Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex.
-On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment across Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
+On the other hand, the metropolitan area has experienced widening disparities in income, poverty, and education attainment between Geomarkets, especially between the historically affluent Collin and Rockwall Counties and the less economically advantaged inner-core Geomarkets (City of Fort Worth and City of Dallas).
 Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
+-->
 
@@
 
-@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area.
-At the lowest PSAT score thresholds, notable differences emerge in the racial/ethnic composition of prospects within Geomarkets when compared to the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
-For example, the largest number of prospects in the 1070-1180 PSAT range were from the Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian), both of which yield predominantly White prospect pools.
-On the other hand, the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarket pools have the greatest proportions of non-White prospects (although yielding fewer overall students).
+@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area, revealing a comparatively more distributed pattern of racial/ethnic diversity across Geomarkets unlike Chicago's concentrated diversity within a single Geomarket.
+This distribution is most notable at the lowest PSAT score threshold.
+For instance, at the 1070-1180 PSAT score range, the central suburban Geomarkets of Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects.
+While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportions of Hispanic prospects but lower proportions of Asian and Black students.
+The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest but predominantly White prospect pools.
+
+<!--# which most closely mirror the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
+For example,  the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarkets have the greatest proportions of non-White prospects although yielding smaller overall pools.
 Interestingly, while the City of Dallas is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-dallas-rq2-race includes larger proportion of Hispanic (41%) prospects but lower proportions of Asian (4%) and Black (9%) students comparatively.
-
-These racial/ethnic composition patterns shift as PSAT scores increase.
-Shares of White prospects within Geomarkets remain relatively proportional to the overall 2020 metropolitan distribution in @fig-dallas-rq1-race.
-Howeover, prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+-->
+
+Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
+<!--OZAN CUT THIS TEXT 5/13.
+THINKING NOT ESSENTIAL.
+AND NOT SO SURE IF IT SHOWS THAT.
+Such shifts also shift distribution patterns of racial/ethnic diversity across Geomarkets, making City of Dallas and City of Fort Worth critical Geomarkets for including Black and Hispanic students from the resulting prospect pools.
+\[KS LEFT OFF HERE} -->
 
 @fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
-@fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).
-Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
-The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+The bars for "All (race known)" show that Collin & Rockwall (37%) and West of Dallas/Forth Worth (23%) contribute many more students than the other four Geomarkets.
+Disparities in Geomarket contributions were evident but differed for Asian, Black, and Hispanic students.
+For instance, The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+
+<!--# @fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).
+-->
 
@@
 Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree.
-Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportinate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
-
-Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Dallas-Fort Worth metropolitan area.
+Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportionate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
+
+Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity, which we view as the most interesting results for the Dallas-Fort Worth metropolitan area.
 Among all White prospects (n=2,094) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
-@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool.
-The only exception is the Irving, Arlington, & Grand Prairie Geomarket contributing a relatively larger share of White first-generation college students whose parents attended some college but did not complete their degree (21% versus 11% for all prospects).
-
+@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool, the only modest exception being the Irving, Arlington, & Grand Prairie Geomarket.
+
+By contrast, for non-White groups, a larger share of prospects were first-generation students.
+Furthermore, first generation status differed substantially across Geomarkets, especially for Hispanic and Asian students.
+For Hispanic students, more than 50% of prospects from the City of Dallas, City of Forth Worth, Arlington & Grand Prairie, and Dallas County are first-generation college students.
+By contrast, for the more affluent Collin & Rockwall Geomarket and the West of Dallas/Fort Worth Metroplex Geomarkets, less than 30% of prospects were first-generation.
+For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
+By contrast, the more affluent Collin & Rockwall sent more Asian prospects (285) than any other Geomarket, but about 87% of these prospects were not first-generation college students.
+
+```{=html}
+<!-- I ADDED THE ABOVE PARAGRAPH AND CUT THIS PARAGRAPH TO GIVE THE LEDE MORE PUNCH 
 Similar patterns of one or two Geomarkets disproportionately contributing larger shares of first-generation college students was evident across other racial/ethnic categories in @fig-dallas-rq2-midSAT-combo.
@@
 City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
-
+-->
+```
 Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions.
@@
 
-### Decentralized Diversity and Income in a Multi-Nucleo in the Los Angeles Geomarket Landscape
-
+### Decentralized Diversity and Income in a Multi-Nucleo Landscape
+
+Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles -- 11 Geomarkets, results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), which are shown in the Online Appendix.
 The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
@@
 As depicted in @fig-la-maps, Geomarkets closely follow this overall metropolitan model.
-Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring of made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
-The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland.
+Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
+The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland, comprising the Inland Empire.
 
 The population growth in the Los Angeles metropolitan area increased from approximately 8 Million in 1980 to more than 12 Million in 2020.
-@fig-dallas-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model.
+@fig-la-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model.
 The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernadino, and Ontario.
@@
 
-\@fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020.
+@fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020.
 Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
-In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
+\[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA\] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
 By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%).
@@
 
-@fig-la-rq1-ses presents trends in socioeconomic indicators for the Los Angeles metropolitan area from 1980 to 2020.
-While median income increased across all of the metropolitan area over the three census periods, some shifts occurred in the relative socioeconomic levels across Geomarkets.
-Many Geomarkets kept their relative levels of affluence from 1980 to 2020.
-For instance, the Covina and West Covina, South Bay, and San Fernando Valley remained among the most affluent, whereas the San Fernando Valley East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario Geomarkets remained at middle ranges of affluence and the Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets remained at the lowest levels of affluence across the metropolitan area during this period.
-The West Los Angeles and West Beach Geomarket demonstrated the most substantial increase in median household income, rising from \$112,000 in 1980 to \$132,000 in 2000, and reaching \$137,000 by 2020.
-This sustained growth positioned it as the most affluent Geomarket in the Los Angeles metropolitan area.
-
-@fig-la-rq1-ses also illustrates shifting levels of poverty across Geomarkets in the Los Angeles metropolitan area over time.
-In 1980, poverty rates were relatively low (less than 12%) for all Geomarkets except for East Los Angeles (17%), South and South Central Los Angeles (24%), and Hollywood and Wilshire (18%).
+<!--# [I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST.
+START TEHRE AND THEN GET TO OTHER RESULTS.] --> @fig-la-rq1-ses presents trends in income, poverty, and educational attainment in the Los Angeles metropolitan area from 1980 to 2020, demonstrating the South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire Geomarkets remained at the lowest socioeconomic levels across these indicators.
+For instance, the median household income in the South and South Central Los Angeles Geomarket increased modestly from \$57,000 in 1980 to \$59,000 in 2000, reaching \$67,000 by 2020.
+In comparison, East Los Angeles and Hollywood and Wilshire recorded slightly higher median incomes in 2020 at \$76,000 and \$84,000, respectively.
+These figures remain substantially lower than that of the most affluent Geomarket, West Los Angeles and West Beach, where the median household income reached \$137,000 in 2020.
+
+@fig-la-rq1-ses also illustrates parallel trends in poverty and educational attainment over time, with Hollywood and Wilshire, East Los Angeles, South and South Central Los Angeles Geomarkets consistenly ranking at lower levels across these socioeconomic indicators.
+Moreover, the disparities between these three Geomarkets and all others in the Los Angeles metropolitan area widened from 1980 to 2020.
+
+<!--# In 1980, poverty rates were relatively low (less than 12%) for all Geomarkets except for East Los Angeles (17%), South and South Central Los Angeles (24%), and Hollywood and Wilshire (18%).
 However, poverty rates declined in South and South Central Los Angeles (22%), remained stable in East Los Angeles and Hollywood and Wilshire, and increased in all other Geomarkets.
@@
 This includes four Geomarkets with poverty rates less than 10% (San Fernando Valley- West, West Los Angeles and West Beach, South Bay, and Covina and West Covina), four Geomarkets with poverty rates ranging from 11%-15% (San Fernando Valley- East, Glendale and Pasadena, Long Beach, and Riverside, San Bernardino, and Ontario), and three Geomarkets with poverty rates ranging from 16%-22% (Hollywood and Wilshire, East Los Angeles, South and South Central Los Angeles).
-
-Lastly, @fig-la-rq1-ses shows a widening disparity in educational attainment across the Los Angeles metropolitan area.
+-->
+
+<!--# Lastly, @fig-la-rq1-ses shows a widening disparity in educational attainment across the Los Angeles metropolitan area.
 Although all Geomarkets experienced increases in the percentage of residents with a bachelor's degree or higher between 1980 and 2020, the West Los Angeles and West Beach Geomarket experienced a substantially greater increase - from 35% in 1980 to 66% in 2020- widening the gap with other Geomarkets in terms of educational attainment.
-
-The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
+-->
+
+<!--#  [THINK YOU CAN CUT THIS PARAGRAPH] The Los Angeles metropolitan area exemplifies a complex multi-nuclei urban model, characterized by decentralized population growth and increasing demographic diversity across its 11 Geomarkets.
 From 1980 to 2020, the region’s population expanded significantly leading to overall declines in the White population and increases in Hispanic and Asian populations across Geomarkets.
@@
 
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university filtered for prospects across the entire state of California, in the 2019-2020 high school graduating class, and by PSAT scores.
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2020-2022 high school graduating classes, and by PSAT scores.
 Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.
 
-The racial/ethnic composition of purchased student profiles within Geomarkets in the Los Angeles metropolitan area is presented in @fig-la-rq2-race.
-The more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180) reflect very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
+@fig-la-rq2-race presents the racial/ethnic composition of purchased student profiles.
+For the more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180), South and South Central Los Angeles (n=1,031) and East Los Angeles (n=871) Geomarket pools stand out as having an overwhelming majority of students of color, 98% and 95% respectively.
+Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (89%) and East Los Angeles (64%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (6%) and larger share of Asian prospects for East Los Angeles (30%).
+
+Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic and Black student prospects, while East Los Angeles stands out for its substantial share of Asian prospects.
+
+<!--#  very similar racial/ethnic compositions within Geomarkets than overall demographic makeup of the metropolitan area in 2020 (see @fig-la-rq1-race).
+[THE LEDE IS SOUTH CENTRAL AND EAST LA.
+START THERE!].
 However, the racial/ethnic composition of prospects become less proportional to overall metropolitan patterns by Geomarket as PSAT scores increase.
@@
 This pattern becomes most pronounced at the highest PSAT score prospect pools.
-This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
+[I DON'T THINK THAT BLACK AND LATINX STUDENTS HAVING LOW TEST SCORES IS THE STORY WE ARE TRYING TO TELL.
+THAT STORY HAS BEEN TOLD.
+THE MORE IMPORTANT THING IS, WITHIN SOME TEST THRESHOLD, HOW DOES THE RACIAL COMPOSITION VARY ACROSS GEOMARKETS.
+LIKE EVEN WITHIN THE MID RANGE, SOUTH CENTRAL, EAST LA, AND INLAND EMPIRE STAND OUT] <!-- This disproportionate over representation for Asian prospects in comparison to the population of residents with Geomarkets ranges from 3 percentage points in the West Los Angeles and West Beach Geomarket (16% prospects versus 13% Asian residents) to 52 percentage points in the East Los Angeles Geomarket (79% prospects versus 27% residents).
 @fig-la-rq2-race illustrates the share of Black prospects within these high PSAT orders declines to less than 3% within all Geomarkets in comparison to Black residents making up a range of 3% (Covina and West Covina) to 18% (South and South Central Los Angeles) of residents within Geomarkets.
 Similar underrepresentation patterns are evident for Hispanic porspects are evident across all Geomarkets execept for a nearly equal representation in South and South Central Los Angeles (73% prospects versus 71% Hispanic residents).
-
-@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area.
-For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.
-Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students.
-The Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller or nearly equal shares.
-
+-->
+
+@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area, demonstrating South and South Central Los Angeles (7% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (25% versus 17% for all prospects) Geomarkets contributed larger proportions of Black prospects relative to their contributions to the overall prospect pool in the Los Angeles metropolitan area.
+<!--#  For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.-->
+
+For Hispanic prospects, South and South Central Los Angeles (8% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (24% versus 17% for all prospects) also contributed a disproportinately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
+
+<!--# [I THINK CLEANER IF YOU JUST HAVE A SMALL PARAGRAPH FOR HISPANIC AND A SMALL PARAGRAPH FOR BLACK.
+I THINK ONE LEDE ACROSS THESE RESULTS IS RIVERSIDE AS AN ENGINE FOR BLACK/LATINX AND FOR FIRST-GEN STUDENTS.
+I GET THAT THESE TWO POINTS CONTRADICT EACH OTHER].
 Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race.
@@
 All other Geomarkets contributed nearly equal shares of Black and Hispanic prospects.
-
-Contributions of each Geomarket to the pool of prospects from the Los Angeles metropolitan area based on first-generation college student status are presented in @fig-la-rq2-midSAT-ses.
-Only 889 of the more than 8,000 total prospects (11%) in this middle PSAT score range (1190-1260) pool were first-generation college students whose parents did not attend college.
-Only three of the 11 Geomarkets contributed substantially larger shares of these first-generation prospects: East Los Angeles (35%), South and South Central Los Angeles (47%), and Riverside, San Bernardino, and Ontario (15%).
-However, disproportionate contributions become more spread across Geomarkets for the nearly 1,436 of the more than 8,000 total prospects (18%) who were first-generation college students whose parents attended but did not complete their degree.
-Overall, six of the 11 Geomarkets contributed larger shares of these prospects: Hollywood & Wilshire (25%), East Los Angeles (26%), South and South Central Los Angeles (28%), Long Beach (25%), Covina and West Covina (22%), and Riverside, San Bernardino, and Ontario (22%).
-This leads to the remaining Geomarkets - San Fernando Valley- West (80%), San Fernando Valley- East (78%), Glendale and Pasadena (77%), West Los Angeles & West Beach (87%), South Bay (81%) - contributing disproportionately larger shares of not first-generation college students relative to the overall prospect pool (71%).
-
-Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
-Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
+-->
+
+In comparison, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian prospects relative to their contributions to the overall prospect pool.
+
+These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Los Angeles metropolitan area using Geomarket filters that exclude lower-income Geomarkets such as South and South Central Los Angeles and Riverside, San Bernadino, and Ontario Geomarkets would result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
+Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.
+
+<!--# [I'M NOT SURE WHERE YOU ARE GETTING THESE NUMBERS FROM; THIS IS THE WRONG FIGURE I THINK.
+AND WHEN YOU REVISE, THE LEDES ARE RIVSERSIDE, EAST LA AND, SOUTH CENTRAL (ONLY 2% OF PROSPECTS BUT 10% OF NO-COLLEGE) KS-- weird...
+overall substantive rankings are correct but my %s are all off].
+--> @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
+For students whose parents did not attend college, East Los Angeles (16% versus 5% of all prospects), South and South Central Los Angeles (10% versus 2% of all prospects), and Riverside, San Bernardino, and Ontario (24% versus 17% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
+These three Geomarkets also made disproportionately larger contributions of first-generation college student prospects whose parents attended but did not complete their degree, as well as Hollywood & Wilshire (7% versus 5% of all prospects), Long Beach (7% versus 5% of all prospects), and Covina and West Covina (13% versus 10% of all prospects).
+
+<!--#  [I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] --> Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
+
+For Hispanic prospects, the South and South Central Los Angeles Geomarket stands out as providing the most substantial share of first-generation college students.
+Across all Hispanic prospects in the middle range PSAT orders, 25% were first-generation college students whose parents did not attend college, 25% were first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
+In comparison, the South and South Central Los Angeles Geomarket contributed much larger disproportionate share of Hispanic first-generation college students whose parents did not attend college (54%), with the Hollywood & Wilshire (30%), East Los Angeles (39%), and Riverside, San Bernardino, and Ontario (31%) also contributing larger but more modest disproportionate shares.
+The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
+
+<!--# [TAKE A TIP FROM THE CRIT QUANT FOLKS; YOU DON'T ALWAYS NEED TO START WITH WHITE; REVISE THIS TEXT TO START WITH A PARAGRAPH ON HISPANIC AND A PARAGRAPH ON ASIAN -- OR VICE VERSA; THOSE TWO ARE THE LEAD; AND THEN CAN TALK ABOUT WHITE AND BLACK.
+]Among all White prospects (n=2,511) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
 @fig-la-rq2-midSAT-combo shows all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students (across both no college and some college parents) relative to the overall pool.
-The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%).
-Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).
-
-Geomarkets contributions to first-generation college students was mixed across other racial/ethnic categories in @fig-la-rq2-midSAT-combo.
-For Asian prospects (11% first-generation no college, 21% first-generation some college, and 68% not first-generation), the Hollywood & Wilshire (12%) and East Los Angeles (36%) Geomarkets contributed larger proportions of first-generation college students whose parents did not attend college.
-The San Fernando Valley- East (28%), Hollywood & Wilshire (35%), East Los Angeles (25%), South and South Central Los Angeles (44%), Long Beach (27%), and Covina and West Covina (25%) Geomarkets also contributed larger proportions of first-generation college students whose parents attended college but did not complete their their degree.
-
-For Black prospects (5% first-generation no college, 20% first-generation some college, and 74% not first-generation), Geomarkets that disproportionately larger proportions of first-generation college students whose parents did not attend college include South Bay (11%), South and South Central Los Angeles (21%), Long Beach (7%), and Covina and West Covina (9%).
-The West Los Angeles and West Beach (26%), Hollywood & Wilshire (36%), East Los Angeles (33%), Long Beach (40%), and Covina and West Covina (27%) Geomarkets also contributed larger proportions of Black first-generation college students whose parents attended college but did not complete their their degree.
-
-Finally, Hispanic prospects in this middle range PSAT order were on average 25% first-generation college students whose parents did not attend college, 25% first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
-The Hollywood & Wilshire (30%), East Los Angeles (39%), South and South Central Los Angeles (54%), and Riverside, San Bernardino, and Ontario (31%) contributed disproportionate shares of Hispanic first-generation college students whose parents did not attend college.
-The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
-
-Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
+The South and South Central Los Angeles Geomarket stands out as one exception contributing a relatively larger share of White first-generation college students whose parents did not attend college (17%) and whose parent attended some college but did not complete their degree (33%) [DON'T MENTION THIS; THIS IS 6 STUDENTS!].
+Other Geomarket that contributed signifcantly larger shares of White first-generation prospects whose parents did not complete their degree include Long Beach (17%), and Riverside, San Bernardino, and Ontario (19%).-->
+
+<!--# [THE LEAD IS EAST LA WHICH HAS 251 ASIAN STUDENTS AND IS WAY DIFFERENT THAN ANY OTHE IN FIRSTGEN; AND THEN WILSHIRE IS SORTA CLOSE]--> Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 21% were first-generation college students whose parents attended some college but did not complete their degree, and 68% were not first-generation college students.
+@fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 251), a significant proportion of whom are first-generation college students (35%) whose parents did not attend college.
+The Hollywood & Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (35%).
+
+While the pool of Black prospects is relatively small (n=186) in @fig-la-rq2-midSAT-combo , some patterns emerge by first-generation status.
+Overall, approximately 5% of Black prospects were first-generation college students-no college, 20% were first-generation college students-some college, and 74% not first-generation).
+The South and South Central Los Angeles Geomarket stands out for contributing a disproportionately high share (21%) of Black first-generation college students whose parents did not attend college, despite representing a relatively smaller overall number.
+Additionally, the Long Beach Geomarket is notable for its higher proportion (40%) of Black first-generation students whose parents attended college but did not complete a degree.
+
+<!--#  [NOTE SURE YOU NEED THIS PARAGRAPH]Despite Los Angeles's multi-nuclei urban structure contributing to a more decentralized Geomarket landscape with racial/ethnic diversity across a range of socioeconomic levels more spatially distributed across the metropolitan area, Geomarkets contributions to student prospect pools reflect racial and socioeconomic stratification patterns similar to other metropolitan cases in the study.
 Affluent Geomarkets like San Fernando Valley–West and West Los Angeles contribute an overrepresentation of White, Asian, and not-first generation college student prospects, the magnitude of which increases substantially as score ranges increase (see Appendix @fig-la-rq2-lowSAT-race and @fig-la-rq2-highSAT-race).
 On the other hand, more diverse and historically low-income Geomarkets, like South and South Central Los Angeles and Riverside, San Bernardino, and Ontario, contribute disproportionately more Black, Hispanic, and first-generation college student prospects.
-Similar to Chicago and Dallas-Fort Worth, Los Angeles further solidifies the pattern of some of the lowest-income Geomarkets consistently playing crucial roles in fostering college access for first-generation Asian (East Los Angeles), Black (South and South Central Los Angeles), and Hispanic college students (South and South Central Los Angeles, East Los Angeles, Hollywood and Wilshire, and Riverside, San Berdardino, and Ontario).
+Similar to Chicago and Dallas-Fort Worth, Los Angeles further solidifies the pattern of some of the lowest-income Geomarkets consistently playing crucial roles in fostering college access for first-generation Asian (East Los Angeles), Black (South and South Central Los Angeles), and Hispanic college students (South and South Central Los Angeles, East Los Angeles, Hollywood and Wilshire, and Riverside, San Berdardino, and Ontario).-->
 
@@
 
-The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the EM industry.
+The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the enrollment management industry.
 The logic of homophily argues that there is a natural cultural match between students, that colleges should find students who look like students who enrolled at their institution in the past, and that colleges should recruit from the same schools and communities that are popular with peer colleges.
@@
 148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.College access scholarship from economics often analyzes the effects of district/state/federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
-These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^19]
-
-[^19]: @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
+These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^1]
+
+[^1]:  @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
 
@@
 Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.
-EAB combined these acquisitions into create [Enroll360](https://eab.com/solutions/enroll360/):
-
-> We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
-This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]
+EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.^[The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
+This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]."]
 
@@
 
+\singlespacing
+
 ::: {#refs}
@@
 
+\doublespacing
 \clearpage

### 2025-05-14 text/manuscript-full-KS.qmd
went through introduction

@@
 title: "Structuring College Access: The Market Segment Model and College Board Geomarkets"
-author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)"]
+author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)", "Crystal Han (Stanford University)"]
 bibliography: ./../assets/bib/cb_geomarket_bib.bib
@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."^[On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
 Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and has a higher share of Black residents.
 Similarly, in the DC/Maryland/Virginia metro area, the order selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets.
-
-[^1]: On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
 
@@
 On the student demand side, the status attainment model argues that student postsecondary destination is a function of family socioeconomic background, particularly parental education [e.g., @RN915; @RN1273; @RN1268; @RN5033; @RN1300].
-The cultural capital model argues that bourgeois households bestow the pedigree, social networks, and information necessary for their progeny to claim spots at selective colleges [@RN1228; @RN1241; @RN4991] [ADD CITES].
-@RN4873 describes schools as "sorting machines" that create, incorporate, and assign students to categories that determine educational opportunities, often reinforcing reinforcing racial and class inequality.
-On the supply-side, the credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required to claim advantageous social and economic positions [@RN1297], and this race for credentials benefits affluent households [@RN861] \[cite others\].
-Scholarship on enrollment management offers an agentic depiction of organizational behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
+The cultural capital model argues that bourgeois households bestow the pedigree, social networks, and information necessary to claim spots at selective colleges [e.g., @RN1228; @RN1241; @RN4991].
+<!--@RN4873 describes schools as "sorting machines" that create, incorporate, and assign students to categories that determine educational opportunities, often reinforcing reinforcing racial and class inequality.
+-->On the supply-side, the credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required to claim advantageous social and economic positions [@RN1297], and this race for credentials benefits affluent households [e.g., @RN865; @RN861].
+Scholarship on enrollment management offers an agentic depiction of college behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
 
 We argue that these perspectives ignore an important mechansim of social reproduction: third-party vendors create products that sort students on behalf of college.
-These products often amplify historic inequality in educational opportunity by using past snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
+These products often amplify historic inequality in educational opportunity by using historical snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
 In the contemporary era of private equity funded platform capitalism, third-party recruiting products have become increasingly powerful, granular, and ubiquitous.
-However, we argue that a watershed moment in their development was the creation of Geomarkets.
+However, a watershed moment in their development was the creation of College Board Geomarkets.
 
@@
 Therefore, colleges should recruit from territories that contain large populations of your target social class.
-Geomarkets are geographic borders meant to define local recruiting markets, a territory of an admissions recruiter.
-For instance, Figure X panel A shows Geomarkets in the Chicago-land area.
+Geomarkets are geographic borders meant to define local recruiting markets.
+Figure X panel A shows Geomarkets in the Chicago-land area.
 As a planning tool for colleges, the Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
@@
 Anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
-Figure X panel B shows that Chicago Geomarkets define recruiting territories for University of Chicago admissions recruiters (Appendix Figures BLAH BLAH).
+For example, Figure X panel B shows that Chicago Geomarkets define recruiting territories for University of Chicago admissions recruiters.
 Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS).
@@
 Third, Geomarkets were incorporated into College Board's student list product, Student Search Service.
-Student lists contain the contact information of prospective students and have been the primary source of lead generation in U.S.
-higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz].
-@RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
-Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc.
+Student lists contain the contact information of prospective students and have been the primary source of "lead generation" in U.S.
+higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz; @RN4728].
+<!-- @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
+--> Colleges control which prospect profiles they purchase by selecting search filters, such as SAT score, AP score, state, etc.
 Geomarket filters enable colleges to include or exclude prospects from particular Geomarkets.
@@
 
-Scholarship on enrollment management has focused on college behavior [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
-However, scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products and consulting solutions -- as a set of mechanisms that structure college access.
+Scholarship on enrollment management has focused on the behavior of colleges [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
+However, scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products, their consulting solutions -- as a set of mechanisms that structure college access.
 These mechanisms, which lie upstream of the behavior of individual colleges, are incorporated by individual colleges in ways that funnel certain kinds of students to certain kinds of institutions.
-This omission is surprising given sociology is concerned with social sorting [@RN4873], particularly sorting into college.
+This omission is surprising because sociology is concerned with social sorting [@RN4873], particularly sorting into college.
 
@@
 
-Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS and within the Student Search Service student list product.
+Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS software (RQ1) and within the Student Search Service student list product (RQ2).
 First, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas and how does this variation change over time?
 We address this question by spatially joining Geomarket shapefiles to Census data about socioeconomic and racial characteristics.
-Second, how does the socioeconomic and racial composition of included versus included prospects vary when student list purchases filter on particular Geomarkets?
-This research question is motivated by @RN4982, which recommends that selective colleges target affluent Geomarkets.
-Additionally, prior research on enrollment management behavior finds that selective colleges focus recruiting visits on high schools in affluent, predominantly white communities [@RN3519; @RN4758].
-We analyze this research question using data from actual student list purchases, which were collected using public records requests.
-We analyze student list purchases that include all Geomarkets in a metropolitan area in order to simulate which prospects would be included or excluded had the student list purchase filtered on particular Geomarkets.
+Second, how does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
+This research question is motivated by @RN4982, which recommend that selective colleges target affluent Geomarkets, and by prior research on enrollment management behavior, which finds that selective colleges focus recruiting visits on high schools in affluent, predominantly white communities [@RN3519; @RN4758].
+We analyze this research question using data from actual student list purchases, which we collected using public records requests.
+We analyze student list purchases that include all Geomarkets in a metropolitan area in order to investigate which prospects would be included or excluded had the student list purchase filtered on particular Geomarkets.
 
@@
 Fourth, we present results.
-\[We find that\ldots ?PARAGRAPH ON RESULTS?\]
-
-Finally, we discuss implications for scholarship and for policy.
-The sociology of education should move beyond the fixation on schools and policies as the primary sorting machines.
-@RN527 \[p.
-148\] remind us that the organizational field includes the totality of relevant actors." Third-party vendors have structured college access for decades and their influence is growing.
-The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [CITE].
-By investigating high-leverage actors and products in the organizational field, sociology can develop important insights that shape future federal and state regulatory policy.
-\[REVISE THIS PARAGRAPH\]
+\[We find that\ldots ONE OR TWO SENTENCES\].
+Finally, we discuss implications for scholarship.
+Third-party vendors have structured college access for decades and their influence is growing.
+The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [vista_eab_ml].
+We propose two fruitful approaches to studying the role of third-party vendors in enrollment management.
+First, case studies -- like this manuscript and @RN5039 -- utilize primary data collection to conduct granular analyses of particular products or vendors.
+Second, a growing number of subscription databases (e.g., Orbis, Pitchbook, LSEG) serve the information needs of the investment community by providing data on private firms and acquisitions or investments in private firms.
+These data sources may provide the basis for a scalable "knowledge infrastructure" [@RN5025] that investigates the role of third-party vendors in education.
 
@@
 Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit.
-The Student Search Service -- created in 1972 and copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
+The Student Search Service -- copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
 The Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
@@
 Second, technological advances creates new means of identifying and serving prospects.
-College Board began selling the contact information of test-takers in 1972 [@belkin2019-studata].
-ACT followed suit and the two testing companies enjoyed a near duopoly for several decades.
+College Board and ACT leveraged data on test-takers to enjoy a near duopoly in market for student list data, one that lasted for several decades.
 This business can be described as list-based lead generation based on the direct-mail model.

### 2025-05-15 text/manuscript-full-KS.qmd
worked through quantification section on full manuscript

@@
 
-@fig-em-funnel depicts the "enrollment funnel," which modifies the traditional "marketing funnel" to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
+@fig-em-funnel depicts the "enrollment funnel," which modifies the traditional marketing funnel to depict broad stages in the process of recruiting students [@RN4986; @RN2004].
 The funnel begins with a large pool of "prospects" (i.e., prospective students) that the university would like to enroll.
@@
 The majority of scholarship focuses on the admissions stage, analyzing which admissions criteria are utilized and/or which applicants are admitted [e.g., @RN4874; @RN4993; @RN4131; @RN3522; @poon2022rethinking].
-Scholarship from the economics of education often investigate financial aid leveraging, which seeks to convert admits to enrolled students [e.g., @RN3762; @RN3133].
+Scholarship from the economics of education often investigate financial aid leveraging, which seeks to convert admits to enrolled students [e.g., @RN3133; @RN3762].
 <!--  Scholarship on college access often takes a demand-side perspective, whereby household differences in economic and cultural capital yield stratification in student preferences and the ability to gain admission [e.g., @RN4935; @RN1814].
@@
 By focusing on the behavior of colleges, scholarship on EM implicitly assumes that recruiting is something done by individual colleges.
-The broader EM industry includes relevant stakeholders in the organizational field [@RN527], including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
+The EM industry includes relevant stakeholders in the organizational field [@RN527], including professional associations (e.g., National Association for College Admission Counseling) and third-party servicers that supply products and consulting solutions to colleges (e.g., College Board, EAB).
 We argue college access is structured by third-party servicers and products that interact with direct-providers (colleges).
-Although sociologists have hinted at the ways EM contributes to inequality in college access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
+Although sociologists have hinted at the ways the broader EM industry contributes to inequality in college access [e.g., @RN2406], scholarship has failed to make third-party servicers and products the object of empirical analysis.
 <!--The enrollment management industry incorporates relevant actors in the organizational field, including professional associations (e.g., National Association for College Admission Counseling), and third-party servicers (e.g., College Board, EAB) that supply products and consulting solutions to colleges.
@@
 Drawing from scholarship on organizational theory, EM processes require colleges to make many "make or buy" [@RN4738; @RN980] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
-Unfortunately, few empirical studies make EM vendors or products the object of empirical analysis.
-The most widely-known class of EM vendor is EM consulting firms, which provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success.
+The most widely-known class of EM vendor is EM consulting firms, which provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success [@RN5027].
 As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].[^2]
-
-[^2]: In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
+The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].^[In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].]
 
@@
 Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit.
-The Student Search Service -- copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
-The Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
+The College Board Student Search Service -- copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
+The College Board Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
 Both products are applications of quantification that sort students into colleges.
@@
 
-Quantification can be defined as "the production and communication of numbers" [@RN5007, p.
-402].
-While a growing interdisciplinary area of scholarship, sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, and personal life [@RN5008].
-Quantification has also been at the core of recent critical data scholarship exploring how algorithms and big data computation structure social opportunities [e.g., @RN4772; @RN4849].
-Our study is grounded in the systematic investigation of "interactions between different quantification regimes" across sociology and critical data studies, as called for by @RN5008 (p.
-239), to explore the role of third-party servicers and products in the sorting of students into colleges.
-Specifically, we draw on quantification scholarship investigating how correlation and homophily [@RN4975] encode social reproduction into our growingly dominant "algorithmic society" \[[CITE](https://medium.com/chatbots-bible-and-more/theres-an-algorithm-for-that-or-there-soon-will-be-d54406dc9bc1)\] and explore how such mechanisms are applied to the construction of Geomarkets in the Market Segment Model [@RN4982].
+@RN5007[p.
+402] define quantification as "the production and communication of numbers." The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers).
+Sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, personal life and education, with a particular emphasis on rankings [@RN5008; @RN5003; @RN5005].
+In the field of critical data studies, quantification is central to scholarship that explores how computation and algorithms structure social opportunities [e.g., @RN4772; @RN4849].
+Our study is grounded in the systematic investigation of "interactions between different quantification regimes" across sociology and critical data studies, as called for by @RN5008[p.
+239], to explore the role of third-party servicers and products in the sorting of students into colleges.
+In particular, the discussions of correlation and homophily by @RN4975 show how social reproduction is encoded in algorithms that direct future resources.
+We explore how correlation and homophily were incorporated into the construction and application of Geomarkets within the Market Segment Model [@RN4982].
 
@@
 @RN4975 [p.
-50] writes that
-
-> correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'
-
-Due to data limitations, predictive analytics often utilize "proxy" variables [@RN4849; @RN4975], which are variables that are correlated with an variable of interest.
+50] writes that, "correlation grounds big data's so-called revolutionary potential.
+As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"
+
+Due to data limitations, predictive analytics often utilize "proxy" variables -- variables highly correlated with a variable of interest [@RN4849; @RN4975].
 For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
 @RN4849 [p.
-146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access, when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
+146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access [e.g., @RN4565], when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
 
@@
 224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
-This phenomenon has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^3] As we discuss below, @RN4982 analyzed 1980 score-sending behavior (training data) and inferred that demand for college is a function of class.
-They then recommended that selective colleges recruit from localities that have a high share of affluent, college-educated households.
-
-[^3]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
-Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
-Note that predictive models such as the LSI-R model are not merely used for social science purposes.
-Rather, they reproduce structural inequality because they direct the allocation of future resources based on a snapshot of historical inequality.
+This phenomenon has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].^[An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
+Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.] As we discuss below, @RN4982 analyzed 1980 SAT score-sending behavior (training data) and inferred that demand for college is a function of class.
+They then recommended that selective colleges recruit from localities with many affluent, college-educated households.
 
@@
 In *The company We Keep: Colleges and Their Competition*, @RN4982 [chapter 4] conduct a network analysis whereby two colleges are defined as competitors if a large number of students submit SAT scores to both colleges.
-@RN4975 argues that because network science models often draw from rational choice theory, they assume that homophily is the result of voluntary action by individuals.
+@RN4975 observes that network science models often draw from rational choice theory, which leads to the assumption that homophily is the result of voluntary action by individuals.
 
@@
 Thus, "social networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@RN4975, p.
-YY].
+94].
 We observe similar phenomena in our case.
-Based on a snapshot of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
+Based on a network analysis of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
 In turn, @RN4982 reasons that colleges should target the Geomarkets and high schools targeted by peer colleges.
@@
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].^[For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit.] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
 We observe similar processes in our case study.
@@
 
-[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit altogether.
-
 We make targeted contributions to scholarship on quantification in the sociology of education.
 As @RN3519 demonstrates in the chapter *Numbers*, EM is fundamentally concerned with quantification.
-However, sociological analyses of EM focus on the behavior of colleges and their agents, yielding the implicit belief that inequalities created by EM are a function of individual colleges reacting to macro structures in their environment such as rankings and financial pressures.
+However, sociological analyses of EM focus on the behavior of colleges and their agents, creating the sense that inequalities created by EM are a function of individual colleges reacting to macro structures in their environment such as rankings and financial pressures [e.g., @RN5003; @RN3519].
 To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of producers.
 Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027].
-Drawing from critical data studies [e.g., @RN4775; @RN4975; @RN4849], recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
+Recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand.
 Based on a snapshot of existing social stratification, market research matches vertically categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
-Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociologists.
-We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
-
-<!--# REMOVED FROM PARAGRAPH ABOVE: First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
--->
-
-This manuscript analyzes Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
+Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociology.
+
+<!--  We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
+# REMOVED FROM PARAGRAPH ABOVE: First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].
+-->
+
+This manuscript analyzes the Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
 The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
-By showing which students are included versus excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in the sorting of students into colleges.
+Unlike the analysis of UK school league tables by @RN5005, we cannot show the effect of quantification on social reproduction.
+However, by demonstrating which students are included versus excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in the sorting of students into colleges.
 
@@
 Both outputs are created separately for each Geomarket.
-For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
-A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a college which Geomarkets possess attractive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.^[For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+A university could obtain this for their own institution or for a competitor.^[Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.] Taken together, the Market Segment Profile shows a college which Geomarkets possess attractive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
 These outputs became the basis for the Enrollment Planning Service (EPS) software.
-
-[^5]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
-
-[^6]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
-This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
 
@@
 **Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one another.[^7] Based on these analyses, @RN4982 \[p.
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one another.^[This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
+Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
+These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
+For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
+54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
+For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."] Based on these analyses, @RN4982 \[p.
 46\] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
@@
 72] describe observed patterns as a natural process in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
-
-[^7]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
-The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
-Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
-For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
-54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
 
@@
 @RN4988 shows the contemporary Geomarkets.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^8] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
-
-[^8]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.^[@RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
 Students from California don't all share the same college-going behaviors.
-We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."
+We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
 
@@
 Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
-EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
+EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.^[Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."] Second, EPS users decide which high schools they will visit within selected Geomarkets.
 Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-
-[^9]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
 
@@
 A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^10] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
-
-[^10]: For public universities that purchased lists, 80% purchase more than 50,000 names annually.
-@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.^[For public universities that purchased lists, 80% purchase more than 50,000 names annually.
+@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
 
@@
 A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
-After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^11] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.[^12]
-
-[^11]: Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
-Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.
-
-[^12]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.^[Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
+Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.^[20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
 Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
-The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).
+The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).]
 
@@
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the *Student Search Service*, such as Geomarkets and Geodemographic Segment filters.[^13] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
+College Board also created geographic borders that subsequently became filters in the *Student Search Service*, such as Geomarkets and Geodemographic Segment filters.^[Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 -->
-
-[^13]: Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
 
@@
 The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-For the 2000 Decennial Census, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty.[^1] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+For the 2000 Decennial Census, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty.^[Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
 These questions were replaced by ACS.
-Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.[^2] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
-
-[^1]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
-
-[^2]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation
+Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.^[Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation.] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
 
@@
 Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post.[^3] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.^[The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
 Using these zip code-level data, we utilized the `aggregate()` function to create shapefiles for each Geomarket.
-
-[^3]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
 
@@
 The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
-In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.[^4] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
-
-[^4]: ADD CAVEAT THAT PURCHASES EXCLUDE PROSPECT PROFILES THAT WERE PREVIOUSLY PURCHASED
-
-In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.[^5] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
+In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.^[The caveat to this statement is that most purchases select to exclude prospect profiles that were previously purchased by the university.] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
+
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.^[For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
 Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
-
-[^5]: For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.
 
@@
 From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
-Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].[^6] Information about student socioeconomic status is limited.
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].^[Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
+Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as "two or more races, non-Hispanic."] Information about student socioeconomic status is limited.
 However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
-
-[^6]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
-Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as “two or more races, non-Hispanic.”
 

### 2025-05-16 text/manuscript-full-KS.qmd
worked through front end

@@
 @RN5007[p.
-402] define quantification as "the production and communication of numbers." The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers).
+402] define quantification as "the production and communication of numbers." <!-- The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers).--> The effects of quantification are described by the concepts reactivity, discipline, and authority [@RN5007].
 Sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, personal life and education, with a particular emphasis on rankings [@RN5008; @RN5003; @RN5005].
@@
 Second, in commercial social networks, homophily is more than an assumption; rather, it is programmed into algorithms that create connections between users.
-Thus, "social networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@RN4975, p.
-94].
+Thus, "networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@Chun2018Queerying, p.
+66].
 We observe similar phenomena in our case.
@@
 The group approached College Board.
-"Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest deomgraphic unit used in these analyses was the state" [@RN4982, p.
+"Coincidentally, the Board was reviewing its own efforts to help colleges estimate their enrollment potential" [@RN4982, p.
 x].
-
 Although the analyses of @RN4982 are based on the score-sending behavior of SAT test-takers, the authors had great respect for admissions officers and conceived of the project as an effort to "capture and quantify" (p.
@@
 
-> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting (p.9)\ldots We have begun with this celebration of storytelling\ldots \[because\] we believe that the intuitions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research is based upon listening carefully to what admissions officers have to say (pp.
+> Admissions officers invariably are tellers of stories -- about the colleges they represent, about the colleges they attended, about each other, and about the often vagabond life of college recruiting\ldots We have begun with this celebration of storytelling\ldots \[because\] we believe that the intuitions of admissions officers actually comprise a remarkably systematic body of knowledge about the college selection process\ldots Our research is based upon listening carefully to what admissions officers have to say (pp.
 9-10).
 
-The "initial task" was to define geographic boundaries "in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982,p.
+The "initial task" was to define geographic boundaries "in a manner consistent with admissions officers' intuitive understanding of student pools" [@RN4982, p.
 4].
-Quoting an admissions officer, @RN4982 \[p.
-11\] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, @RN4982 [p.
-11] created "three types of boundaries -- region, state, and community.
-The initial regions in the project were New England, Middle States, and the South.
+Quoting an admissions officer, @RN4982[p.
+11] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, @RN4982[p.
+11] created "three types of boundaries -- region, state, and community." The initial regions in the project were New England, Middle States, and the South.
 Next, they "divided each state into as few as two and as many as thirty community-based enrollment markets or pools, for a total of 143 separate markets" (p.
@@
 Having defined geography, the Market Segment Model sought to predict how student demand for a particular college varies by Geomarket, based on the characteristics of households.
-The model created a straightforward but consistent classification system that grouped SAT test-takers into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
+The model created a classification system that grouped SAT test-takers into one of four different *market segments* -- local, in-state, regional, and national -- based on SAT score-sending behavior.
 For a given student, each college that receives a score from the student can be defined as "local" (college located in the the same Geomarket as the student), "in-state" (same state but different Geomarket), "regional" (same region but different state), or "national" (different region).
@@
 An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
+@RN4982[p.
+4] comment that the Market Segment Model "was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
+The model worked because students, once so disaggregated, appeared to behave in remarkably consistent ways."
 
@@
 The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
-Appendix A describes these outputs in more detail.
+Appendix A describes these outputs in more detail [CRYSTAL].
 Both outputs are created separately for each Geomarket.
@@
 **Correlations**.
-@RN4982 [Chapter 3], *A Sense of Place: Students, Families, and Communities*, identifies the student characteristics correlated with being in the local, in-state, regional, or national market segment.
+@RN4982 [Chapter 3], *A Sense of Place: Students, Families, and Communities*, investigates the student characteristics correlated with being in the local, in-state, regional, or national market segment.
 The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
@@
 
-The thesis of Chapter 3 is that student demand for higher education is a function of social origin.
-For the authors (p.
-33) state that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and that "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 \[p.
-33\] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time."\[\^7\] @RN4982 [p.
-42] conclude that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations."
+The thesis of this chapter is that student demand for higher education is a function of social origin.
+For example, the authors state (p.
+33) that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and that "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982[p.
+33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time." They conclude (p.
+42) that "our research has simply demonstrated what everyone has always known: communities with high levels of family income and parental education are also communities in which students have higher than average SATs and more far-reaching aspirations."
 
@@
 However, Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
-Therefore, @RN4982 [p.
+Therefore, @RN4982[p.
 44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:
@@
 54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."] Based on these analyses, @RN4982 \[p.
-46\] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
+For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."] Based on these analyses, @RN4982[p.
+46] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
 Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
@@
 This homophily is presented as a naturally occurring phenomenon.
-Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer-colleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
+Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer colleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
 In itself, @RN4982's Market Segment Model is a social science depiction of student demand -- akin to the status attainment model -- that does not consider historic, structural inequalities that cause observed patterns.
 However, by inscribing the Market Segment Model and Geomarkets into the EPS software, College Board amplified structural inequalities that contributed to homophily observed in 1980 SAT score-sending beahivor.
-In this way, commodification of the Market Segment Model and Geomarkets "engineers homophily" [@RN4975].
+Drawing from @RN4975, commodification of the Market Segment Model and Geomarkets engineers homophily.
 
@@
 
-Quantification disciplines actors to react in particular ways.
-EPS software may "discipline" [@RN5007] colleges to approach recruiting in a manner consistent with the Market Segment Model.
+Quantification "disciplines" [@RN5007] actors to react in particular ways.
+EPS software may discipline colleges to approach recruiting in a manner consistent with the Market Segment Model.
 Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
@@
 Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-
-The effects of quantification also lead to changes in decision-making authority, frequently reducing the discretion of local decision-makers.
-@RN4982 sought to develop a concrete, data-driven framework -- The Market Segment Model -- that replicates the aggregate knowledge of local admissions officers.
+Anecdotally, the salience of Geomarkets and EPS software affected the decision-making authority of admissions officers and the territories that admissions officers were assigned to.^[Quantification often reduces the decision-making authority of local actors [@RN5007].
+The Market Segment Model sought to replicate the aggregate knowledge of local admissions officers.
 Once this knowledge was quantified and commodified onto a CD-ROM, the local expertise of admissions officers becomes less valuable.
@@
 We were told that savvier, quantitatively-adept admissions offices used EPS to visit schools that EPS recommended ignoring because there would be less competition for the good students who attended these high schools.
-
-Finally, we suspect that Geomarkets affect how admissions offices organize the recruiting territories assigned to each admissions officer.
-Geomarkets were created to mirror the territories of admissions officers [@RN4982].
+Finally, Geomarkets were created to mirror the territories of admissions officers [@RN4982].
 In turn, as Geomarkets became more salient, admissions offices often structured their territories around Geomarkets.
@@
 In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
-This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX CRYSTAL\] <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
-
-Since @RN4982 developed The Market Segment Model, Geomarkets have been central to EPS software and play a broader role in shaping how many colleges organize recruiting territories.
-We therefore frame our investigation of Geomarkets as a case study of quantification to explore the extent to which this geographical building block is associated with race and class.
-We first address (RQ1) what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time.
-<!--# Although we cannot gain access to EPS software, by focusing on metropolitan areas that are associated with three or more Geomarkets-->
-
-@RN4982 viewed demand for higher education as a function of class and developed Geomarkets with an eye towards identifying geographic areas that differed from one another in terms of class composition.
-Given the extent of class-based residential segregation in the U.S., we therefore expect substantial socioeconomic inequality between Geomarkets in large metropolitan areas.
+This can be seen in the Chicago-land recruiting territories of University of Chicago [ADD TO APPENDIX], of Lake Forest College, of Saint Mary's College [ADD TO APPENDIX CRYSTAL]] <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
+
+Because Geomarkets are foundational to EPS software, our first research question is interested in the extent to which this geographical building block is associated with race and class.
+We ask (RQ1), what is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
+@RN4982 viewed demand for higher education as a function of class.
+They developed 142 Geomarkets borders with an eye towards identifying geographic areas that differed from one another in terms of class composition, while [@RN5016] created the remaining Geomarket borders with an eye towards distinguishing geographies based on college-going behaviors.
+Considering the extent of class-based residential segregation in the U.S.
+[@reardon2011], we expect substantial socioeconomic inequality between Geomarkets in large metropolitan areas.
 
@@
 Second, although geomarket borders may have been drawn along class divides, a strong correlation exists between race and wealth [@RN4884].
-Third, Geomarket borders may have been drawn in a way that follows the countours of racial segregation in residential housing.
+Third, Geomarket borders may have been drawn in a way that follows the contours of racial segregation in residential housing.
 Examples include: the "South and South Central Los Angeles" geomarket (CA21); the "City of Oakland" geomarket (CA07), which is surrounded by the "Alameda County excluding Oakland" geomarket (CA08); and the "Wayne County Detroit" geomarket (MI01), which is surrounded by the "Detroit's Northern Suburbs" (MI02) and "Ann Arbor" (MI03) geomarkets.
@@
 Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.^[For public universities that purchased lists, 80% purchase more than 50,000 names annually.
-@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to purchasing names.] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to names.] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
 
@@
 A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
-After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.^[Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).
-Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and lowincome students.] @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years, particularly for sub-groups that were historically excluded from college.^[20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.^[Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and low-income students.] <!--^[Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).] --> @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years.^[20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.] <!-- , particularly for sub-groups that were historically excluded from college.^[.
 Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
-The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).]
+The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).]-->
 
@@
 Second, Geomarkets may be systematically correlated with race.
-For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda Country, excluding Oakland." A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.
-Our study therefore extends @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
+For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda Country, excluding Oakland." <!--# A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.--> We extend @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
 <!--# This manuscript focuses on Geomarkets.
@@
 
-> Imagine for the moment that you recruit for a college that draws most of its students from regional or national segments.
+> Imagine that you recruit for a college that draws most of its students from regional or national segments.
 Where would you concentrate your energies?
 Ideally, you would seek communities \[Geomarkets\] with a high proportion of students already predisposed toward institutions such as your own.
-The Market Segment Model would provide this information through segment percentages for the community \[Geomarket\] in question.
-Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen, while slightly less than half of the students with two or more attributes would be receptive\ldots.
-Your efforts would surely be better directed toward \ldots Manchester, Hartford, and Fairfield County \ldots Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own.
+The Market Segment Model would provide this information through segment percentages for the community in question.
+Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen \ldots Your efforts would surely be better directed toward \ldots Manchester, Hartford, and Fairfield County \ldots Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population.
+<!-- -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own.
+-->
 
@@
 Our analyses attempt to recreate the strategy recommended by @RN4982 by showing who is included versus excluded when student list purchases filter on particular Geomarkets.
-Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access, and vice-versa.
+Considering the correlation between race and wealth in segregated America, we suggest that a class-based list-buying strategy will increase racial stratification in college access.
 <!--#  Furthermore, we anticipate that the students of color from affluent Geomarkets will tend to have college-educated parents and be enrolled in affluent, predominantly white schools.-->

### 2025-05-19 text/discussion-KS-copy.qmd
discussion revisions

--- 2025-05-16 text/manuscript-full-KS.qmd
@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."^[On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
 Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and has a higher share of Black residents.
 Similarly, in the DC/Maryland/Virginia metro area, the order selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets.
+
+[^1]: On the same day, the university purchased two additional orders that filtered on PSAT score ranges of $1190 - 1260$ and $1270 - 1520$, respectively, but selected the same HS graduating class and Geomarkets.
 
@@
 ```
-
 Sociology offers several perspectives to describe the process of sorting students into colleges as a process that tends toward social reproduction [@RN4873; @RN2377].
@@
 Fourth, we present results.
-\[We find that\ldots ONE OR TWO SENTENCES\].
+[We find that\ldots ONE OR TWO SENTENCES].
 Finally, we discuss implications for scholarship.
@@
 As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
-The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].^[In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].]
+The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].[^2]
+
+[^2]: In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
 
@@
 ```
-
 **Quantification**
 
-@RN5007[p.
+@RN5007 [p.
 402] define quantification as "the production and communication of numbers." <!-- The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers).--> The effects of quantification are described by the concepts reactivity, discipline, and authority [@RN5007].
@@
 In the field of critical data studies, quantification is central to scholarship that explores how computation and algorithms structure social opportunities [e.g., @RN4772; @RN4849].
-Our study is grounded in the systematic investigation of "interactions between different quantification regimes" across sociology and critical data studies, as called for by @RN5008[p.
-239], to explore the role of third-party servicers and products in the sorting of students into colleges.
+Our study is grounded in the systematic investigation of "interactions between different quantification regimes" across sociology and critical data studies, as called for by @RN5008 \[p.
+239\], to explore the role of third-party servicers and products in the sorting of students into colleges.
 In particular, the discussions of correlation and homophily by @RN4975 show how social reproduction is encoded in algorithms that direct future resources.
@@
 ```
-
 ```{=html}
@@
 For example, "e-scores" utilize proxy variables to identify "people like you" and then predict your buying behavior based on the past buying behavior of people like you.
-@RN4849 [p.
-146] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access [e.g., @RN4565], when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
+@RN4849 \[p.
+146\] states that "the modelers for e-scores have to make do with trying to answer the question 'How have people like you behaved in the past?' when ideally they would ask, 'How have you behaved in the past?'" We observe similar behavior in market research about college access [e.g., @RN4565], when enrollment managers use 'which colleges did students near you consider' as a proxy for, 'which colleges would you consider?'
 
@@
 224] state that "predicting the future on the basis of the past threatens to reify and reproduce existing inequalities." Disproportionately targeted/excluded populations are predicted to have a higher risk of an outcome, which amplifies subsequent targeting/exclusion.
-This phenomenon has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].^[An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
-Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.] As we discuss below, @RN4982 analyzed 1980 SAT score-sending behavior (training data) and inferred that demand for college is a function of class.
+This phenomenon has been termed the "ratchet effect" [@RN4851] and "pernicious feedback loops" [@RN4849].[^3] As we discuss below, @RN4982 analyzed 1980 SAT score-sending behavior (training data) and inferred that demand for college is a function of class.
 They then recommended that selective colleges recruit from localities with many affluent, college-educated households.
+
+[^3]: An often cited example is the LSI-R recidivism model which predicts a prisoner's chances of re-arrest and is used by 24 states [@RN4849].
+Because the algorithm uses zip code as an input, people who live in highly policed neighborhoods have a higher predicted probability of being arrested, which leads to more policing in those neighborhoods, which perpetuates racialized inequality in arrests.
 
@@
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].^[For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit.] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
+Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
 We observe similar processes in our case study.
@@
 
+[^4]: For example, "payday loans" charge high interest rates to consumer groups that were previously denied credit.
+
 We make targeted contributions to scholarship on quantification in the sociology of education.
@@
 
+```{=html}
 <!--  We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
@@
 -->
-
+```
 This manuscript analyzes the Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
@@
 4].
-Quoting an admissions officer, @RN4982[p.
-11] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, @RN4982[p.
+Quoting an admissions officer, @RN4982 \[p.
+11\] write, "'There are only three kinds of college-bound students: those who want to live at home, those who want to live on campus but bring their laundry home, and those who want to go far enough from home that Mom and Dad can't visit without calling first.'" As such, @RN4982 [p.
 11] created "three types of boundaries -- region, state, and community." The initial regions in the project were New England, Middle States, and the South.
@@
 An "in-state" student submits more SAT scores to in-state institutions than they do to local, regional, or national institutions, etc.
-@RN4982[p.
+@RN4982 [p.
 4] comment that the Market Segment Model "was nothing more than a set of simple rules for disaggregating high school seniors into similar groups.
@@
 ```
-
 The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
-Appendix A describes these outputs in more detail [CRYSTAL].
+Appendix A describes these outputs in more detail \[CRYSTAL\].
 Both outputs are created separately for each Geomarket.
-For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.^[For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
-A university could obtain this for their own institution or for a competitor.^[Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
-This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.] Taken together, the Market Segment Profile shows a college which Geomarkets possess attractive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
+For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a college which Geomarkets possess attractive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
 These outputs became the basis for the Enrollment Planning Service (EPS) software.
+
+[^5]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
+
+[^6]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
 
@@
 For example, the authors state (p.
-33) that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and that "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982[p.
+33) that "these data allow us to say with considerable confidence that local and in-state students are not likely to come from families in which both parents have received college educations" and that "the implication is simply that college-educated parents instill in their children more wide-ranging educational aspirations." Commenting on family income, @RN4982 [p.
 33] write that "we could predict that all local students would come from moderate-income or low-income families and be wrong only 5.5 percent of the time." They conclude (p.
@@
 However, Geomarkets differ in the relative abundance of students with particular socioeconomic characteristics, which has practical implications for recruiting.
-Therefore, @RN4982[p.
+Therefore, @RN4982 [p.
 44] recommend that colleges target Geomarkets with desirable compositions of socioeconomic characteristics in order to reach students from desired student market segments:
@@
 **Homophily**.
-*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one another.^[This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
+*The Company We Keep: Colleges and Their Competition* [@RN4982, Chapter 4] conducts a network analysis to determine which institutions are in competition with one another.[^7] Based on these analyses, @RN4982 \[p.
+46\] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
+Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
+The authors argue that like-colleges compete for like-students as defined by socioeconomic characteristics.
+@RN4982 [p.
+72] describe observed patterns as a natural process in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+
+[^7]: This is a "two-mode" social network in which students (mode 1) send SAT scores -- the network tie -- to colleges (mode 2).
 The authors turn this into a one-mode college network that defines two institutions as being in competition with one another -- the network tie -- if at least 15% of students who sent SAT scores to one institution also sent scores to the other institution and vice-versa.
@@
 54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
-For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."] Based on these analyses, @RN4982[p.
-46] state that competition between colleges is characterized by homophily: "we draw a fundamental conclusion about the structure of college choice: collegiate competition occurs principally between like institutions." Subsequent analyses investigate the tuition price and the socioeconomic composition of institutions in competition with one another.
-Private selective colleges and private flagship universities compete directly for students, charge the highest prices, and enroll students with the highest socioeconomic status.
-The authors argue that like-colleges compete for like-students as defined by socioeconomic characteristics.
-@RN4982 [p.
-72] describe observed patterns as a natural process in which a vertical socioeconomic hierarchy of students is matched to a vertical hierarchy of universities:
+For example, public flagships compete primarily with other public flagships; private standard colleges, with other private standard colleges; Catholic institutions, with other Catholic institutions."
 
@@
 @RN4988 shows the contemporary Geomarkets.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.^[@RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
+Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^8] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
+
+[^8]: @RN5016 states that "geomarkets are areas within a state that represent a further segmentation of a population.
 Students from California don't all share the same college-going behaviors.
-We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
+We have accounted for this variance by segmenting the 50 states into 304 geomarkets to provide further insight into student behaviors within particular areas of individual states."
 
@@
 Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
-EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.^[Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
-Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."] Second, EPS users decide which high schools they will visit within selected Geomarkets.
+EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
 Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
-Anecdotally, the salience of Geomarkets and EPS software affected the decision-making authority of admissions officers and the territories that admissions officers were assigned to.^[Quantification often reduces the decision-making authority of local actors [@RN5007].
+Anecdotally, the salience of Geomarkets and EPS software affected the decision-making authority of admissions officers and the territories that admissions officers were assigned to.[^10] <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
+
+[^9]: Software documentation by @RN5017 states that, "EPS market codes are proprietary market codes owned by the College Board and are used to categorize external organizations and people into geographical areas\ldots.
+Some admissions offices use EPS market codes to focus their recruiting efforts in geographic areas in which they believe they will be the most successful."
+
+[^10]: Quantification often reduces the decision-making authority of local actors [@RN5007].
 The Market Segment Model sought to replicate the aggregate knowledge of local admissions officers.
@@
 In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
-This can be seen in the Chicago-land recruiting territories of University of Chicago [ADD TO APPENDIX], of Lake Forest College, of Saint Mary's College [ADD TO APPENDIX CRYSTAL]] <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
+This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX CRYSTAL\]
 
@@
 ```
-
 We also expect substantial racial inequality between Geomarkets within metropolitan areas.
@@
 A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.^[For public universities that purchased lists, 80% purchase more than 50,000 names annually.
-@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to names.] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+
+[^11]: For public universities that purchased lists, 80% purchase more than 50,000 names annually.
+@RN4896 reports that student lists were the top expenditure item in the undergraduate recruiting budget for both private and public institutions in 2022, with the average public institution allocating 15% of its budget to names.
 
@@
 A College Board research report by @RN4739 compared SAT test-takers who opted into the College Board Student Search Service -- allowing colleges to purchase their contact information -- to those who opted out.
-After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.^[Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and low-income students.] <!--^[Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).] --> @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years.^[20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.] <!-- , particularly for sub-groups that were historically excluded from college.^[.
+After controlling for covariates (e.g., SAT score, parental education, school fixed effects), 41.1% of students who participated in Search attended a 4-year college compared to 32.8% of students who opted out, an 8.3 percentage point difference and a 25.3 percent change -- $(41.1–32.8)/32.8$ -- in the relative probability.[^12] <!--^[Participating in Search was associated with a larger change in the relative probability of attending a 4-year college for Black students (24.5%) and Hispanic students (34.4%) than White students (21.6%), and a larger change for students whose parents did not attend college (40.6%) than those whose parents had a bachelor’s degree (18.9%).] --> @RN4739 also found that participating in Search was strongly associated with obtaining a BA in four years.[^13] <!-- , particularly for sub-groups that were historically excluded from college.^[.
 Furthermore, the relative increase in the probability of obtaining a BA was higher for Black (40.3%), Hispanic (43.3%), and Native American/Alaska Native students (48.3%) than it was for White (27.9%) and Asian (28.2%) students.
 The relative increase was also higher for students whose parents did not attend college (50.0%) than it was for students whose parents had a BA (25.3%).]-->
+
+[^12]: Leveraging a natural experiment in College Board student list purchases, @RN4894 find that purchasing a prospect profile increases the probability that the student will apply to and enroll at the purchasing college, with larger effects for Black, Hispanic, and low-income students.
+
+[^13]: 20.6% of students who participated in Search obtained a BA in four years compared to 15.7% of students who opted out, representing a 31.2% (=(20.6-15.7)/15.7) increase in the relative probability of graduation.
 
@@
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the *Student Search Service*, such as Geomarkets and Geodemographic Segment filters.^[Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
+College Board also created geographic borders that subsequently became filters in the *Student Search Service*, such as Geomarkets and Geodemographic Segment filters.[^14] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 -->
+
+[^14]: Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
 
@@
 The 1980 Decennial Census data collection period closely matches the data collection period of SAT score-sending behavior that @RN4982 utilized to create Geomarkets and the Market Segment Model.
-For the 2000 Decennial Census, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty.^[Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
+For the 2000 Decennial Census, we similarly use the Summary File 1A (SF1A) for census-tract measures of race and Hispanic origin and the Summary File 3A (SF3A) for census-tract measures of income, education, and poverty.[^15] Beginning in 2010, the Decennial Census no longer collected a "long form" questionnaire.
 These questions were replaced by ACS.
-Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.^[Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation.] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+Therefore, we also utilized the 2020 5-year ACS for both race and socioeconomic characteristics at the census tract level, which includes data collected from 2016-2020.[^16] This data collection period mirrors the period for our primary data collection about student list purchases used to analyze RQ2, as described below.
+
+[^15]: Data for the 1980 and 2000 Decennial Census were retrieved from the IPUMS National Historical Geographic Information System, created by @RN5028.
+
+[^16]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation.
 
@@
 Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.^[The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.[^17] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
 Using these zip code-level data, we utilized the `aggregate()` function to create shapefiles for each Geomarket.
+
+[^17]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
 
@@
 The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria.
-In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.^[The caveat to this statement is that most purchases select to exclude prospect profiles that were previously purchased by the university.] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
-
-In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.^[For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
+In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.[^18] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.
+
+[^18]: The caveat to this statement is that most purchases select to exclude prospect profiles that were previously purchased by the university.
+
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.[^19] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
 Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.
+
+[^19]: For analyses of Dallas, Detroit, Long Island, Northern New Jersey, and Philadelphia, we utilize data from student list purchases that did filter on Geomarkets but included all Geomarkets in the region.
 
@@
 From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
-Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].^[Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
-Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as "two or more races, non-Hispanic."] Information about student socioeconomic status is limited.
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].[^20] Information about student socioeconomic status is limited.
 However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
+
+[^20]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
+Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as "two or more races, non-Hispanic."
 
@@
 
-<!--#  [WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]--> @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
+<!--#  [WRITING IS QUITE GOOD!I THINK THERE IS A MODEST LEAD THAT THE NEXT COUPLE PARAGRAPHS DON'T STRESS ENOUGH , WHICH IS THAT TX19/CITY OF DALLAS IS THE MOST LEAST WHITE/ASIAN AND THE POOREST]-->
+
+@fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
 In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
@@
 
-<!--# [CITY OF DALLAS!] --> Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
+<!--# [CITY OF DALLAS!] -->
+
+Socioeconomic demographics of median income, percent living below poverty, and educational attainment over time are presented in @fig-dallas-rq1-ses for the Dallas-Fort Worth metropolitan area.
 Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
@@
 
-<!--# [I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST.
-START TEHRE AND THEN GET TO OTHER RESULTS.] --> @fig-la-rq1-ses presents trends in income, poverty, and educational attainment in the Los Angeles metropolitan area from 1980 to 2020, demonstrating the South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire Geomarkets remained at the lowest socioeconomic levels across these indicators.
+<!--# [I THINK THE LEADS ARE THAT SOUTH CENTRAL (ESPECIALLY), EAST LA, AND WILSHIRE ARE THE POOREST. START TEHRE AND THEN GET TO OTHER RESULTS.] -->
+
+@fig-la-rq1-ses presents trends in income, poverty, and educational attainment in the Los Angeles metropolitan area from 1980 to 2020, demonstrating the South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire Geomarkets remained at the lowest socioeconomic levels across these indicators.
 For instance, the median household income in the South and South Central Los Angeles Geomarket increased modestly from \$57,000 in 1980 to \$59,000 in 2000, reaching \$67,000 by 2020.
@@
 overall substantive rankings are correct but my %s are all off].
---> @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
+-->
+
+@fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
 For students whose parents did not attend college, East Los Angeles (16% versus 5% of all prospects), South and South Central Los Angeles (10% versus 2% of all prospects), and Riverside, San Bernardino, and Ontario (24% versus 17% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
@@
 
-<!--#  [I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] --> Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
+<!--#  [I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] -->
+
+Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
 
@@
 
-<!--# [THE LEAD IS EAST LA WHICH HAS 251 ASIAN STUDENTS AND IS WAY DIFFERENT THAN ANY OTHE IN FIRSTGEN; AND THEN WILSHIRE IS SORTA CLOSE]--> Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 21% were first-generation college students whose parents attended some college but did not complete their degree, and 68% were not first-generation college students.
+<!--# [THE LEAD IS EAST LA WHICH HAS 251 ASIAN STUDENTS AND IS WAY DIFFERENT THAN ANY OTHE IN FIRSTGEN; AND THEN WILSHIRE IS SORTA CLOSE]-->
+
+Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 21% were first-generation college students whose parents attended some college but did not complete their degree, and 68% were not first-generation college students.
 @fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 251), a significant proportion of whom are first-generation college students (35%) whose parents did not attend college.
@@
 
-The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the enrollment management industry.
-The logic of homophily argues that there is a natural cultural match between students, that colleges should find students who look like students who enrolled at their institution in the past, and that colleges should recruit from the same schools and communities that are popular with peer colleges.
+The discussion of homophily by @RN4975 describes not only the logic of The Market Segment model but also a broader institutional logic that characterizes the enrollment management industry.
+Homophily argues that there is a natural cultural match between students and colleges, suggesting colleges should find prospective students who resemble students who enrolled at their institution in the past and focus recruiting efforts on the same schools and communities that are popular with peer colleges.
 We suggest that homophily is the operating logic most consistent with college recruiting behavior observed in the wild.
 For example, @RN5023 find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
-The also find great overlap in the schools visited by Villanova and Notre Dame, with a disproportionate focus on visits to Catholic schools.
+The study also finds great overlap in the schools visited by Villanova University and the University of Notre Dame, with a disproportionate focus on visits to Catholic schools.
 When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
@@
 The Market Segment Model begins with the status attainment model, that student demand for college is a function of parental education and parental income.
-@RN4982 [chapter 3] demonstrated this argument by showing the correlation between student SAT score-sending behavior and measures of class.
-The thesis of the Market Segment Model is that that households of particular class and college aspirations are likely to live in particular parts of a metropolitan area and that these geographic territories can be meaningfully captured by Geomarkets.
+@RN4982 [chapter 3] illustrated this argument by demonstrating a correlation between student SAT score-sending behavior and measures of class.
+The thesis of the Market Segment Model is that that households of specific class characteristics and college aspirations are likely to live in particular communities within metropolitan areas and that these geographic territories can be meaningfully captured by Geomarkets.
 Although Geomarkets appear coarse by contemporary standards, they were a landmark innovation in Geographic segmentation.
-The Market Segment Model recommends that colleges should identify their core student market segment -- local, in-state, regional, or national -- identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
+The Market Segment Model recommends that colleges should identify their core student market segment (i.e., local, in-state, regional, or national), identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
 This information is contained in the Market Segment Profile (Appendix A), a standard ouput in EPS software.
@@
 @RN4982 state that the Market Segment Model merely formalizes the operating logic of college admissions officers.
-For admissions officers at selective colleges, the logic is that good students from good families are willing to attend good colleges that are far from home.
-These good families are found in particular parts -- Geomarkets -- of metropolitan areas.
-By contrast, students from other parts of the metro possess the disposition and achievement to attend their local college.
-The Market Segment Model formalized this logic and commodified it in EPS software that helped colleges identify desirable Geomarkets and desirable high schools within those Geomarkets.
+For admissions officers at selective colleges, the logic is that "good" students from "good" families are willing to attend "good" colleges that are far from home.
+These "good" families are found in particular parts -- Geomarkets -- of metropolitan areas.
+By contrast, students from other parts of the metropolitan area possess the disposition and achievement to attend their local college.
+The Market Segment Model formalized and commodified this logic within the EPS software to help colleges identify desirable Geomarkets and desirable high schools within those Geomarkets.
 @RN5007 suggest that, once commodified and quantified, the local knowledge of admissions officers becomes redundant.
@@
 148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.College access scholarship from economics often analyzes the effects of district/state/federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
-These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^1]
-
-[^1]:  @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
+These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^21]
+
+[^21]: @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
 
@@
 Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.
-EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.^[The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
-This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]."]
+EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.[^22]
+
+[^22]: The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
+This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]."
 
@@
 
+```{=tex}
 \doublespacing
 \clearpage
-
+```
 <!-- TABLES -->

### 2025-05-20 text/discussion-KS-copy.qmd
discussion notes/revisions

@@
 By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
-For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
+For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Hispanic population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
 Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
@@
 The study also finds great overlap in the schools visited by Villanova University and the University of Notre Dame, with a disproportionate focus on visits to Catholic schools.
-When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
-As @RN4975 argues, observed homophily is not the result of voluntary action; rather, homophily is programmed into algorithms that create connections.
+<!--# When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
+--> As @RN4975 argues, observed homophily is not the result of voluntary action; rather, homophily is programmed into algorithms that create connections.
 
@@
 This information is contained in the Market Segment Profile (Appendix A), a standard ouput in EPS software.
-When considering new Geomarkets to target, colleges should utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
-This information is contained in the Institutional Profile (Appendix A), also included in EPS software.
-
-@RN4982 state that the Market Segment Model merely formalizes the operating logic of college admissions officers.
+When considering new Geomarkets to target, EPS recommends colleges utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
+This information is contained in the Institutional Profile (Appendix A).
+
+<!--#  @RN4982 state that the Market Segment Model merely formalizes the operating logic of college admissions officers.
 For admissions officers at selective colleges, the logic is that "good" students from "good" families are willing to attend "good" colleges that are far from home.
@@
 @RN5007 suggest that, once commodified and quantified, the local knowledge of admissions officers becomes redundant.
+KS: NOT SURE THIS NEEDS REPEATING-->
 
@@
 We find that Geomarkets are highly correlated with race and class.
-\[KARINA SUMMARIZE FINDINGS IN A COUPLE SENTENCES; REVISE PARAGRAPH AS NECESSARY\] We show that in 1980, when Geomarkets were being created, that Black people tended to be highly concentrated in the poorest Geomarket in the metropolitan area.
-Examples include, CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), TX19 -- City of Dallas, CA21 -- South & South Central Los Angeles, and PA5 -- Philadelphia County.
-In a handful of metropolitan areas (e.g., Cleveland), Black people remain conentrated in the poorest Geomarkets.
-In Chicago-Land, \[KARINA -- ONE OR TWO SENTENCES\]
+<!--#  [KARINA SUMMARIZE FINDINGS IN A COUPLE SENTENCES; REVISE PARAGRAPH AS NECESSARY]--> Findings illustrate that in 1980, when Geomarkets were being created, Black and Hispanic residents tended to be highly concentrated in a single, usually the poorest Geomarket in the Chicago, Dallas-Fortworth, and to some extent the Los Angeles metropolitan areas.
+These Geomarket patterns extend to other metropolitan areas in the country, such as CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), and PA5 -- Philadelphia County.
+By 2020, various Geomarket landscapes emerged.
+In Chicago, Black residents remain concentrated in the poorest Geomarket.
+In contrast, in Dallas-Fortworth, Black and Hispanic communities are distributed across multiple Geomarkets; however, significant socioeconomic disparities persist between Geomarkets.
+The Los Angeles metropolitan area presents another distinctive Geomarket landscape in which Black, Hispanic, and Asian communities across a range of socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.
+<!--# [KARINA -- ONE OR TWO SENTENCES] -->
 
 These findings are not surprising given the extent of race and class segregation in the US.
-Furthermore, the correlation between Geomarkets and class is expcted because Geomarket borders were created in service of conceptual model that views demand for higher education as a function of class.
+Furthermore, the correlation between Geomarkets and class is expected because Geomarket borders were created in service of conceptual model that views demand for higher education as a function of class.
 @RN4982 is written around New England, with a particular focus on Geomarkets in Connecticut and Massachusetts.
 @RN4982 do not discuss the intersection of student demand and race.
-Although unsurprising for early 1980s social science, this silence is conspicuous considering the salience of political controversy surrounding "busing" in Boston and nearby metropolitan areas.
+Although unsurprising for early 1980s social science, this silence is conspicuous considering the salience of political controversy surrounding "busing" in Boston and nearby metropolitan areas.<!--# KS: COULD CUT THIS IF NEEDED; MAIN TAKEAWAY IS JUST THAT RQ1 IS UNSURPRINSING EMPIRICALLY VIA SEGREGATION AND THEORETICALLY VIA ZEMSKY & OEDEL-->
 
 Research question 2 analyzes the racial and socioeconomic composition of student list purchases.
-We examine purchases that include all Geomarkets around a metropolitan area in order to assess the potential consequences if including/excluding particular Geomarkets from the purchase.
+We examine purchases that include all Geomarkets in a metropolitan area in order to assess the potential consequences if including or excluding particular Geomarkets from the purchase.
 This question is motivated by the @RN4982 recommendation that selective colleges that enroll students from the regional and national market segment should target affluent Geomarkets, while community colleges and non-selective 4-year colleges should focus on middle- and working-class Geomarkets.
-KARINA -- ADD TEXT SUMMARIZING RESULTS.
+Findings across metropolitan cases suggest student list orders that use Geomarket filters to exclude middle- and working-class Geomarkets - such as IL11 -- City of Chicago, TX19 -- City of Dallas, TX20 -- City of Fort Worth, CA19 -- East Los Angeles, and CA21 -- South and South Central Los Angeles, in targeting high-achieving students may still capture some Black, Hispanic, and Asian prospects.
+However, these prospects are far more unlikely to be first-generation students of color.
+<!--# KARINA -- ADD TEXT SUMMARIZING RESULTS.-->
 
@@
 whereas Asian students living in affluent Geomarkets are more likely to have parents who were born in the U.S.
-
-The cultural capital model explains a process by which upper and upper-middle class families sort themselves into selective colleges by providing their children with the pedigree (academic, extracurricular) valued by selective colleges, information about how to navigate the admissions gauntlet, and social networks that provide an inside track [@RN1228; @RN1241; @RN4991; @RN1814].
+<!--# KARINA NOTES -- USE RESULTS TO SPEAK TO MAJOR THEORIES.-->
+
+The cultural capital model explains a process by which upper and upper-middle class families sort themselves into selective colleges by providing their children with the pedigree (academic opportunities, extracurriculars) valued by selective colleges, information about how to navigate the admissions gauntlet, and social networks that provide an inside track [@RN1228; @RN1241; @RN4991; @RN1814].
 This model explains how affluent households maintain a disproportionate enrollment share at selective institutions in an era of holistic admissions ostensibly designed to increase racial and class diversity [@RN4991].
@@
 Quantitative analyses of visits from colleges to high schools are consistent with @RN3519; selective private colleges and public research universities devote most of their recruiting resources on courting students from privileged schools and communities [@RN5023; @RN4758].
-However, these analyses conceive of recruiting as something that is done by individual colleges.
+However, these analyses conceive of recruiting as something that is done solely by individual colleges.
 
@@
 In itself, the Market Segment Model is an unremarkable social science depiction.
-However, the College Board incscribed Geomarkets and the Market Segment Model into EPS software that told colleges which Geomarkets and high schools to target.
-Technologies that target target customers based on geography inevitably leverage racial and class segregation [@RN4775; @RN4849; @RN4975].
+However, the College Board inscribed Geomarkets and the Market Segment Model into EPS software that tells colleges which Geomarkets and high schools to target.
+Technologies that target customers based on geography inevitably leverage racial and class segregation [@RN4775; @RN4849; @RN4975].
 Furthermore, technologies that use the logic of predictive analytics -- identify correlates in past cases to make predictions for future cases -- are prone to "pernicious feedback loops" [@RN4849].
@@
 We show that that excluding low-income, non-white Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-white students with strong test scores.
+<!--# KS Notes: THIS IS THE STRONGEST PARAGRAPH -->
 
@@
 @RN527 \[p.
-148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.College access scholarship from economics often analyzes the effects of district/state/federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
+148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
+College access scholarship from economics often analyzes the effects of district, state, and federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
 These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^21]
@@
 This business can be described as list-based lead generation based on the direct-mail model.
-"Free" college search engines (e.g., Cappex, Niche) yielded news sources of student list data.
+"Free" college search engines (e.g., Cappex, Niche) yielded new sources of student list data.
 Another data source is college planning software purchased by high school districts and utilized by high school students and guidance counselors.
-The most widely-known product is Naviance, which claims to be used by more than used by more than 10 million K-12 students and by 40% ofUS high schools [@powerschool].
+The most widely-known product is Naviance, which claims to be used by more than 10 million K-12 students and by 40% of US high schools [@powerschool].
 *Naviance* college planning software feeds into the *Intersect* recruiting platform, which allows colleges to target Naviance users while they are on the platform.
@@
 
-The third dynamic in the EM industry has been the growth of test-optional movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT.
+The third dynamic in the EM industry has been the growth of test-optional admissions movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT.
 ACT responded to uncertainty in their core testing business by attempting to become an edtech company \[CITE\], failing \[\], and being acquired by Nexus Capital in 2024 \[CITE\].
@@
 
-The fourth dynamic is the transformation from owner-operated to private-equity owned firms and erosion of the distinction between consultancy and software vendor.
+The fourth dynamic is the transformation from owner-operated to private-equity owned firms and the erosion of the distinction between consultancy and software vendor.
 The 1980s and 1990s was a period of market entry in EM, when college admissions professionals or professors decided to hang up a shingle [@RN5027].
 The 2010s were a period of acquisitions and concentration.
-After regulatory scrutiny pushed private-equity interests out of for-profit college market [@RN5022], private equity investors find value in acquiring firms that provided services to direct providers.
+After regulatory scrutiny pushed private-equity interests out of the for-profit college market [@RN5022], private equity investors found value in acquiring firms that provided services to direct providers.
 Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions to grow market share and vertical acquisitions designed to make customers more dependent on particular recruiting products.
@@
 Enroll360 exemplifies a new kind of student list product that leverages proprietary data to sell software.
-Historically, College Board and ACT sold prospect contact information at a price-per-prospect (e.g., \$0.50).
+Historically, College Board and ACT sold prospect contact information at a price-per-prospect (e.g., \$0.50 for each "name").
 The goal was to generate revenue from the sale of names.
 By contrast, Enroll360 wraps several proprietary databases of names (e.g., Cappex, Intersect) into a software-as-service platform that recruits those prospects at different stages of the enrollment funnel.
-Here, the goal of lists is to create demand for software, which is more expensive than names.
+Here, the goal of lists is to create demand for software, which is more expensive than purchasing names.
 Colleges that are interested in recruiting these prospects must buy an Enroll360 subscription.
-Like College Board *Student Search Service* contemporary software-as-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
+Like College Board's *Student Search Service,* contemporary software-as-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
 For example, reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
 Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and wheather students used Naviance to "research competitor institutions" [@markup_naviance].
-A procurement document from the University of Utah explains that the Intersect subscription is necessary to recruit the proprietary database of Naviance users: [@utah_procurement]:
-
-> The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to.
+<!--#  A procurement document from the University of Utah explains that the Intersect subscription is necessary to recruit the proprietary database of Naviance users: [@utah_procurement]:-->
+
+> <!--# The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to.
 There is unique group of prospective students who are only in the PowerSchool Naviance platform.
 We have spoken with several vendors where we can reach out to prospective students.
-Each vendor has their own proprietary website or database where select students can be reached.
+Each vendor has their own proprietary website or database where select students can be reached.-->
 
@@
 
-Market dynamics in the EM industry raise policy concerns about tuition prices, student choice, and competition in the market for EM consultancies may also affect competition, prices, and student choice.
+Market dynamics in the EM industry raise policy concerns about tuition prices, student choice, and competition in the market for EM consultancies that may also affect competition, prices, and student choice.
 To the extent that colleges must buy expensive software to recruit prospects contained in an associated database, this expenditure will ultimately be passed on to students in the form of higher prices or reduced expenditure on education.
 Unfettered horizontal and vertical acquisitions in the EM industry causes the EM industry to become more concentrated.
-Smaller consultancies that do not control veritial inputs may be unable to compete with companies that have acquired proprietary pools of prospects.
+Smaller consultancies that do not control vertical inputs may be unable to compete with companies that have acquired proprietary pools of prospects.
 Fewer EM consultancies means less competition and higher consulting fees, which will be passed on to consumers.
@@
 However, private firms face fewer disclosure requirements than publicly traded firms.
-The paucity of research about third-party vendors that structure college access is substantially due to difficulty obtaining data for empirical analyses.
+The paucity of research about third-party vendors that structure college access is partially due to difficulty obtaining data for empirical analyses.
 The research community must figure out how to systematically investigate the role of private firms and investors in education, so that we can hold private interests accountable for education policy goals.
@@
 We argue that sociologists and adjacent scholars should expand the domain of the education research knowledge infrastructure to include third-party vendors in enrollment management, and edtech more broadly.
-This is a difficult task.
+However, this is a difficult task.
 Student unit record data owned by district, state, and federal agencies is simultaneously granular -- can evaluate a particular policy -- and macro -- can speak to broad trends in the education sector.
@@
 University libraries provides full or partial access to some subscriptions, while others must be purchased.
+<!--# POTENTIALLY CUT THIS BASED ON OJs THOUGHTS/WANTS ; ALSO ENDING WITH BUILDING A KNOWLEDGE INFASTRUCTURE IS A BIT OF A BUZZKILL FROM ENDING ON MARKET DYNAMICS LEADING TO COMPETITION PRICE STUDENT CHOICE; but maybe fine for soc outlet-->
 

### 2025-05-20 text/manuscript-full-KS.qmd
working through results. again

@@
 ```
-# Findings
+# Results
+
+Results are presented below for Chicago, Dallas, and Los Angeles.
+The [Online Appendix](https://ozanj.github.io/cb_geo/#/results) provides a more convenient means of viewing results.
+The online appendix also contains interactive maps, additional figures and tables not contained in the main text, and results for additional metropolitan areas (e.g., Bay Area, Philadelphia).
+The table of contents for navigating between metro areas can be accessed by clicking on the three horizontal lines on the bottom left of the screen.
 
@@
 @fig-chicago-maps presents the seven Geomarkets in the Chicago metropolitan area by total population across the three census periods.
-The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982 [p.
+The maps in the figure show how Chicago Geomarkets reflect the classic urban geographical structure of an inner-city core surrounded by a ring of suburbs as described by @RN4982[p.
 12].
@@
 In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
-The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most densely populated Geomarket.
-
-These modest population shifts lead to substantial change in the racial/ethnic composition within and across Geomarkets over time.
+The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most populous Geomarket.
+
+These modest population shifts belie substantial change in the racial/ethnic composition within and across Geomarkets over time.
 @fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarket across the three census periods.
@@
 In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
-By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes (59% White, 8% Black, 24% Hispanic; 7% Asian Pacific Islander), Northwest Suburbs (62% White, 3% Black, 18% Hispanic; 15% Asian Pacific Islander), Evanston and Skokie (64% White, 7% Black, 10% Hispanic; 16% Asian Pacific Islander), Western Suburbs (56% White, 9% Black, 23% Hispanic; 9% Asian Pacific Islander), and South and Southwest Suburbs (53% White, 24% Black, 18% Hispanic; 3% Asian Pacific Islander).
+One data caveat is that the following race/ethnicity categories were not available in 1980: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic.
+By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes, Northwest Suburbs, Evanston and Skokie, Western Suburbs, and South and Southwest Suburbs.
 For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
@@
 
-We also explore socioeconomic demographics - median income, percent living below poverty, and educational attainment - within and across Geomarkets over time in @fig-chicago-rq1-ses.
+The interactive maps in the [Online Appendix](https://ozanj.github.io/cb_geo/#/research-question-1-15) can show these shifts at the Census Tract-level.
+For example, in 1980 Black people were concentrated in southern tracks of the City of Chicago Geomarket.
+As described by scholarship on gentrification in Chicago [@ewing2024_nofuture; @snidal2022_exodus].
+However, in 2000 and 2020 Black populations shift towards northeast tracks of the South and Southwest Suburbs Geomarket
+
+We explore socioeconomic demographics -- median income, percent living below poverty, and educational attainment -- within and across Geomarkets over time in @fig-chicago-rq1-ses.
+The primary takeaways are the the City of Chicago -- and to a somewhat lesser degree th South and Southwest Suburbs - stands out as having lower median income, higher poverty rates, and lower BA attainment than other Geomarkets.
+By contrast, the North Shore stands outas being particularly affluent and highly educated.
 Overall, median income both within and across Geomarkets in the Chicago metropolitan area remained relatively stable across the three census periods.
@@
 While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$79,000) and decrease in the South and Southwest Suburbs Geomarket (\$93,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$14,000.
-
 @fig-chicago-rq1-ses also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
@@
 However, by 2020, the Geomarket experienced the largest percentage point increase in college degree attainment, reaching 38%.
-In comparison, all other Geomarkets experienced smaller gains in the share of residents with a bachelor’s degree or higher over the same period.
-These trends contributed to a narrowing of the educational attainment gap across Geomarkets from 1980 to 2020.
-
-Demographic shifts overtime have transformed the Chicago metropolitan area's Geomarket landscape from largely a binary structure into a more tiered dynamic shaped by substantial racial/ethnic and socioeconomic inequality between Geomarkets.
+
+Overall, demographic shifts transformed the Chicago-land Geomarket landscape from largely a binary structure into a more tiered dynamic that retains by substantial racial/ethnic and socioeconomic inequality between Geomarkets.
 The original Geomarket boundaries delineated significant racial and income segregation in the 1980s.
@@
 
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders placed by a research university targeting prospects across the entire state of Illinois.
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders -- placed by a research university -- that target prospects across the entire state of Illinois.
 The first two orders filter on the same SAT thresholds (ranging from 1020 to 1150), the same GPA thresholds (ranging from B- to A+), and filter on the high school graduating classes of 2020 and 2021 graduating classes, respectively.
-The remaining four orders used similar GPA and high school class filters but differed in SAT thresholds.
+The remaining four orders used the same GPA and high school class filters but differed in SAT thresholds.
 Two orders targeted students with relatively middle-range scores (1160 to 1300) and two orders targeted those with relatively high scores (1310 to 1600).
@@
 
-The City of Chicago Geomarket increasingly becomes the only Geomarket with substantial shares of Black, and Hispanic prospects as PSAT scores increase.
-All Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline.
-For instance, the City of Chicago Geomarket pool is 18% Asian, 7% Black, 21% Hispanic, 5% Multiracial, making it the only Geomarket that is not predominantly White (49%).
+As PSAT score ranges increase, The City of Chicago Geomarket becomes the only Geomarket with substantial shares of Black, and Hispanic prospects.
+Moving from the top to the bottom panels, all Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline.
+For the highest SAT score range, the City of Chicago Geomarket pool is 18% Asian, 7% Black, 21% Hispanic, 5% Multiracial, making it the only Geomarket that is not predominantly White (49%).
 Prospect pools from all other Geomarkets in this high PSAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.4% (North Shore) to 5% (South and Southwest Suburbs) of prospects and 6% (Northwest Suburbs) to 10% (South and Southwest Suburbs) of prospects, respectively.
@@
 We also explore contributions across Geomarkets to overall prospect pools by race/ethnicity.
-For instance, @fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle range SAT scores (1160-1300) by race/ethnicity.
-The top panel presents the full distribution of the 13,643 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% reside in Chain of Lakes, 13% in Northwest Suburbs, 5% in North Shore, 5% in Evanston and Skokie, 17% in City of Chicago, 28% in Western Suburbs, and 20% in South and Southwest Suburbs.
+@fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle range SAT scores (1160-1300) by race/ethnicity.
+The top panel presents the full distribution of the 13,643 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% of all purchased prospect profiles reside in Chain of Lakes, 13% in Northwest Suburbs, 5% in North Shore, 5% in Evanston and Skokie, 17% in City of Chicago, 28% in Western Suburbs, and 20% in South and Southwest Suburbs.
 These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-midSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
@@
 -->
-```
-For instance, @fig-chicago-rq2-midSAT-race shows the nearly 1,000 Black students whose contact information was purchased in these middle SAT score orders were largely contributed to the overall pool by the City of Chicago and South and Southwest Suburbs.
-These two Geomarkets contributed disproportionately higher shares, 48% and 30%, respectively, compared to their overall contributions of 17% and 20% of all included prospects.
+
+```
+
+@fig-chicago-rq2-midSAT-race shows which Geomarkets send the bulk of students for a given race/ethnicity.
+The City of Chicago accounts for 17% of all prospects but 48% of Black students.
+Similarly, South \& Southwest Suburbs accounts for 20% of all prospects but 30% of all Black prospects.
+The upshot is that if the student list purchases had excluded these two Geomarkets, 78% of the nearly 1,000 Black students currently included in these purchases would be excluded.
 On the other hand, all other Geomarkets contributed substantially smaller shares of Black prospects.
+For example, the Northwest Suburbs accounts for about 13% of all prospects but only about 3% of Black prospects.
 The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
@@
 @fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates proportional to their overall contributions to all included prospects (except City of Chicago).
-Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects) and Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospectq8s) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects) and Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
 
@@
 
-Relative to @fig-chicago-rq2-midSAT-ses, appendix figures \@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status are less pronounced in low-SAT range orders (1020-1150) and more pronounced in high-range SAT orders (1310-1600), respectively.
+<!--KARINA -- I AM CUTTING THIS.
+# Relative to @fig-chicago-rq2-midSAT-ses, appendix figures \@fig-chicago-rq2-midSAT-ses and @fig-chicago-rq2-highSAT-ses demonstrate Geomarket contribution patterns by first-generation college student status are less pronounced in low-SAT range orders (1020-1150) and more pronounced in high-range SAT orders (1310-1600), respectively.
+-->
 
@@
 @fig-chicago-rq2-midSAT-combo shows the City of Chicago Geomarket consistently contributes a disproportionately larger share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
-For instance, among all Asian prospects (n=1,341) whose contact information was purchased in this SAT score range, approximately 12% were first-generation college students whose parents did not attend college, 16% were first-generation college students whose parents attended some college but did not complete their degree, and 72% were not first-generation college students.
-@fig-chicago-rq2-midSAT-combo shows the City of Chicago disproportionately contributes 35% of all Asian first-generation college students whose parents did not attend college.
-For first-generation college students whose parents attended some college but never completed degrees, City of Chicago (22%), South and Southwest Suburbs (20%), and Evanston and Skokie (20%) Geomarkets contributed larger shares of Asian prospects.
+For instance, among all Asian prospects (n=1,341) whose contact information was purchased in this SAT score range, approximately 12% were first-generation college students whose parents did not attend college, 16% were first-generation college students whose parents attended some college but did not complete their degree, and 72% had parents with a BA.
+By contrast, for 237 Asian students from the City of Chicago Geomarket, 35% had parents who did not attend colllege and only 43% had parents with a BA.
 
@@
 The City of Chicago Geomarket stands out as the only Geomarket contributing substantially larger shares of Hispanic first-generation college students across those whose parents did not attend college (46% versus 38% overall) and those whose parents attended some college but did not complete their degrees (29% versus 26% overall).
+By contrast, Hispanic students from the more affluent North Shore Geomarket and Evanston \& Skokie Geomarket were more likely to have parents with a BA.
+
 Although Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and Western Surburbs Geomarkets.
-
-Lastly, although Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and Western Surburbs Geomarkets.
 Similarly, Geomarkets such as Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs contribute a disproportionately larger share of White prospects who are not first-generation college students, whereas City of Chicago and South and Southwest Suburbs contribute larger shares of White first-generation college students.
@@
 THAT IS THE LEDE; ALSO, THIS IS WHERE YOU COULD REFER TO THE SIMILAR GRAPH FOR PHILADELPHIA AND SAY IT FOLLOWS THE SAME PATTERN; I THINK THIS IS THE CLIMAX AND COOLEST INSIGHT OF THE RESULTS] KS=== I THINK ITS" WEIRD TO INCLUDE PHILLY GRAPH HERE WHEN ITS NOT CASE...WE CAN SPEAK TO THIS IN THE DISCUSSION INSTEAD...
-WE SEE SIMILAR PATTERNS ACROSS OTHER METROS TOO -->
-
-Appendix figures @fig-chicago-rq2-lowSAT-combo and \@fig-chicago-rq2-highSAT-combo show the magnitude of these Geomarket contribution disparities become more pronounced at as SAT score thresholds increase.
-
-The enduring segregated structure of the Chicago metropolitan area from 1980 to 2020 analyzed above plays a significant role in how Geomarkets shape the racial/ethnic and socioeconomic composition of prospect pools.
-Reflecting the larger metropolitan area dynamics, central Geomarkets that are predominantly comprised of low-income communities of color such as the City of Chicago and South and Southwest Suburbs consistently contribute disproportionately larger shares of Black, Hispanic, and first-generation college prospects, especially at lower SAT score thresholds.
+WE SEE SIMILAR PATTERNS ACROSS OTHER METROS TOO Appendix figures @fig-chicago-rq2-lowSAT-combo and \@fig-chicago-rq2-highSAT-combo show the magnitude of these Geomarket contribution disparities become more pronounced at as SAT score thresholds increase.
+-->
+
+In summary, the enduring segregated structure of the Chicago metropolitan area from 1980 to 2020 plays a significant role in how Geomarkets shape the racial/ethnic and socioeconomic composition of prospect pools.
+Reflecting the larger metropolitan area dynamics, Geomarkets that are predominantly comprised of low-income communities of color such as the City of Chicago and South and Southwest Suburbs consistently contribute disproportionately larger shares of Black, Hispanic, and first-generation college prospects, especially at lower SAT score thresholds.
 On the other hand, predominantly White and affluent Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, as well as non-first generation college students.
@@
 
-From 1980 to 2020, the population of the overall Dallas-Fort Worth metropolitan area increased substantially from approximately 2.7 Million to 6.8 Million.
+From 1980 to 2020, the population of the overall Dallas-Fort Worth metropolitan area increased dramatically from approximately 2.7 Million to 6.8 Million.
 @fig-dallas-maps shows all six Geomarkets experienced population growth over the three census periods.
 However, the most pronounced increases occured in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
-These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers - City of Forth Worth and City of Dallas - into the surrounding suburban and rural areas of the metropolitan areas.
+These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers -- City of Forth Worth and City of Dallas -- into the surrounding suburban and rural areas of the metropolitan areas.
 This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket from 1980 to 2000.
@@
 In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
-By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County (58% White, 16% Black, 19% Hispanic; 6% Asian), Irving, Arlington, and Grand Prairie (54% White, 12% Black, 26% Hispanic; 16% Asian), and City of Fort Worth (57% White, 16% Black, 22% Hispanic; 3% Asian).
+By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County, Irving, Arlington, and Grand Prairie, and City of Fort Wort.
 In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
-
 By 2000, the City of Dallas was no longer a predominantly White Geomarket, with a population that was approximately 24% Black, 30% Hispanic, and 3% Asian.
@@
 Overall, median income both within and across Geomarkets remained relatively stable across the three census periods.
-Between 1980 to 2000, affluence levels across Geomarkets were consistent, ranging from a low of \$82,000 in City of Fort Worth to a high of \$135,00 in the Collin & Rockwall Counties Geomarket.
+<!-- Between 1980 to 2000, affluence levels across Geomarkets were consistent, ranging from a low of \$82,000 in City of Fort Worth to a high of \$135,00 in the Collin & Rockwall Counties Geomarket.
 However, the substantial increase in median income for the Collin & Rockwall Counties Geomarket during this time period plateaued by 2020, reshuffling relatively affluence levels among Geomarkets.
 By 2020, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000), and Dallas County (\$90,000).
-
 In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable through 2020.
-As a result, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 median household incomes, respectively.
+As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 [MISMATCH BETWEEN GRAPH AND TABLE?] median household incomes, respectively.
+By contrast, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000).
 @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty and educational attainment within Geomarkets over time, revealing similar dynamics across Geomarkets to those observed in median household income.
@@
 
-We also analyze student list orders collected through public requests to assess how the racial/ethnic and socioeconomic composition of included versus excluded prospects varies when filtering on particular Geomarkets within the Dallas-Fort Worth metropolitan area.
+We analyze how the racial/ethnic and socioeconomic composition of included versus excluded prospects varies when filtering on particular Geomarkets within the Dallas-Fort Worth metropolitan area.
 Specifically, we analyze three list orders placed by a research university targeting prospects across 87 different Geomarkets nationwide, including all six Geomarkets in the Dallas-Fort Worth Metropolitan area.
 Each order also filtered for prospects in the 2019-2020 graduating class and based on PSAT scores across three ranges: 1070-1180, 1190-1260, and 1270-1520.
-Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260) from the Dallas-Fort Worth metropolitan area to simulate Geomarket contributions and assess patterns of inclusion versus exclusion.
+Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260).
 
@@
 The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest but predominantly White prospect pools.
-
-<!--# which most closely mirror the overall demographic makeup of the metropolitan area in 2020 (see @fig-dallas-rq1-race).
-For example,  the Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and the Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) Geomarkets have the greatest proportions of non-White prospects although yielding smaller overall pools.
-Interestingly, while the City of Dallas is the most racially/ethnically diverse Geomarket in the larger metropolitan area, its prospect pool in @fig-dallas-rq2-race includes larger proportion of Hispanic (41%) prospects but lower proportions of Asian (4%) and Black (9%) students comparatively.
--->
-
 Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
-<!--OZAN CUT THIS TEXT 5/13.
-THINKING NOT ESSENTIAL.
-AND NOT SO SURE IF IT SHOWS THAT.
-Such shifts also shift distribution patterns of racial/ethnic diversity across Geomarkets, making City of Dallas and City of Fort Worth critical Geomarkets for including Black and Hispanic students from the resulting prospect pools.
-\[KS LEFT OFF HERE} -->
 
@@
 
-Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses.
+Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses [THE FIGURE IN THE PAPER IS WRONG].
 Only 303 prospects of the 3,928 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
@@
 City of Dallas (38%), City of Fort Worth (39%), Irving, Arlington, & Grand Prairie (32%), Dallas Counties (28%) all contributed larger shares of Hispanic students whose parents did not attend college relative to the overall prospect pool of these first-generation college students (23%), whereas only Dallas Counties contributed to larger shares of Hispanic students who parents attended some college but did not attain their degree (26% versus 20% for all prospects).
--->
-```
+
 Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions.
@@
 While Collin and Rockwall Counties and the West of Dallas/Fort Worth Metroplex Geomarkets still contribute a notable overall number of Black, Hispanic, and Asian prospects, such students are predominantly not first-generation college students.
+-->
+```
 
@@
 
-Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles -- 11 Geomarkets, results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), which are shown in the Online Appendix.
+[START HERE] Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles -- 11 Geomarkets, results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), which are shown in the Online Appendix.
 The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
@@
 
-The discussion of homophily by @RN4975 describe the logic of The Market Segment model but also an underlying institutional logic of the enrollment management industry.
-The logic of homophily argues that there is a natural cultural match between students, that colleges should find students who look like students who enrolled at their institution in the past, and that colleges should recruit from the same schools and communities that are popular with peer colleges.
+The discussion of homophily by @RN4975 describes not only the logic of The Market Segment model but also a broader institutional logic that characterizes the enrollment management industry.
+Homophily argues that there is a natural cultural match between students and colleges, suggesting colleges should find prospective students who resemble students who enrolled at their institution in the past and focus recruiting efforts on the same schools and communities that are popular with peer colleges.
 We suggest that homophily is the operating logic most consistent with college recruiting behavior observed in the wild.
 For example, @RN5023 find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
-The also find great overlap in the schools visited by Villanova and Notre Dame, with a disproportionate focus on visits to Catholic schools.
-When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
-As @RN4975 argues, observed homophily is not the result of voluntary action; rather, homophily is programmed into algorithms that create connections.
+The study also finds great overlap in the schools visited by Villanova University and the University of Notre Dame, with a disproportionate focus on visits to Catholic schools.
+<!--# When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
+--> As @RN4975 argues, observed homophily is not the result of voluntary action; rather, homophily is programmed into algorithms that create connections.
 
 The Market Segment Model begins with the status attainment model, that student demand for college is a function of parental education and parental income.
-@RN4982 [chapter 3] demonstrated this argument by showing the correlation between student SAT score-sending behavior and measures of class.
-The thesis of the Market Segment Model is that that households of particular class and college aspirations are likely to live in particular parts of a metropolitan area and that these geographic territories can be meaningfully captured by Geomarkets.
+@RN4982 [chapter 3] illustrated this argument by demonstrating a correlation between student SAT score-sending behavior and measures of class.
+The thesis of the Market Segment Model is that that households of specific class characteristics and college aspirations are likely to live in particular communities within metropolitan areas and that these geographic territories can be meaningfully captured by Geomarkets.
 Although Geomarkets appear coarse by contemporary standards, they were a landmark innovation in Geographic segmentation.
-The Market Segment Model recommends that colleges should identify their core student market segment -- local, in-state, regional, or national -- identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
+The Market Segment Model recommends that colleges should identify their core student market segment (i.e., local, in-state, regional, or national), identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
 This information is contained in the Market Segment Profile (Appendix A), a standard ouput in EPS software.
-When considering new Geomarkets to target, colleges should utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
-This information is contained in the Institutional Profile (Appendix A), also included in EPS software.
-
-@RN4982 state that the Market Segment Model merely formalizes the operating logic of college admissions officers.
-For admissions officers at selective colleges, the logic is that good students from good families are willing to attend good colleges that are far from home.
-These good families are found in particular parts -- Geomarkets -- of metropolitan areas.
-By contrast, students from other parts of the metro possess the disposition and achievement to attend their local college.
-The Market Segment Model formalized this logic and commodified it in EPS software that helped colleges identify desirable Geomarkets and desirable high schools within those Geomarkets.
+When considering new Geomarkets to target, EPS recommends colleges utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
+This information is contained in the Institutional Profile (Appendix A).
+
+<!--#  @RN4982 state that the Market Segment Model merely formalizes the operating logic of college admissions officers.
+For admissions officers at selective colleges, the logic is that "good" students from "good" families are willing to attend "good" colleges that are far from home.
+These "good" families are found in particular parts -- Geomarkets -- of metropolitan areas.
+By contrast, students from other parts of the metropolitan area possess the disposition and achievement to attend their local college.
+The Market Segment Model formalized and commodified this logic within the EPS software to help colleges identify desirable Geomarkets and desirable high schools within those Geomarkets.
 @RN5007 suggest that, once commodified and quantified, the local knowledge of admissions officers becomes redundant.
+KS: NOT SURE THIS NEEDS REPEATING-->
 
@@
 @RN527 \[p.
-148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.College access scholarship from economics often analyzes the effects of district/state/federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
-These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.[^1]
-
-[^1]:  @RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.
+148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
+College access scholarship from economics often analyzes the effects of district, state, and federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
+These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.^[@RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.]
 
@@
 
+\begingroup
+\small
+
 ::: {#refs}
@@
 
+\endgroup
 \doublespacing

### 2025-05-20 text/old_spring2024/discussion-KS-copy.qmd
working through results. again

--- 2025-05-20 text/discussion-KS-copy.qmd
@@
 ```
+
 This manuscript analyzes the Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
@@
 ```
+
 # Findings
@@
 ```
+
 For instance, @fig-chicago-rq2-midSAT-race shows the nearly 1,000 Black students whose contact information was purchased in these middle SAT score orders were largely contributed to the overall pool by the City of Chicago and South and Southwest Suburbs.
@@
 ```
+
 Analyses of student list orders across the Dallas-Fort Worth metropolitan area reveals patterns of racial/ethnic and socioeconomic stratification in Geomarket contributions.
@@
 
-```{=tex}
 \doublespacing
 \clearpage
-```
+
 <!-- TABLES -->

### 2025-05-21 text/manuscript-full-KS.qmd
went through results

@@
 
-### Decentralized Diversity and Income in a Multi-Nucleo Landscape
-
-[START HERE] Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles -- 11 Geomarkets, results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), which are shown in the Online Appendix.
+### Decentralized Diversity in a Multi-Nucleo Landscape
+
+Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles -- 11 Geomarkets, results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), results for which are shown in the [Online Appendix](https://ozanj.github.io/cb_geo/#/results).
 The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
@@
 The South and South Central Los Angeles Geomarket grew from approximately 1.2 to 1.6 Million from 1980 to 2020.
-In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets grew from approximately 700,000 to 1.2 Million and from 1.1 Million to nearly 3 Million during this time, respectively.
+In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets, respectively, grew from approximately 700,000 to 1.2 Million and from 1.1 Million to nearly 3 Million during this time.
 
@@
 Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
-\[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA\] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
+\[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA; KARINA -- MENTION THIS OR REVISE HOW YOU WRITE FINDINGS IN THIS PARAGRAPH, LIKE FOCUS MAINLY ON 2000-2020.
+YOU DECIDE.\] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
 By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%).
@@
 
+<!--# KARINA -- CUT THIS CUZ BASICALLY CONSISTENT WITH INCOME.
 @fig-la-rq1-ses also illustrates parallel trends in poverty and educational attainment over time, with Hollywood and Wilshire, East Los Angeles, South and South Central Los Angeles Geomarkets consistenly ranking at lower levels across these socioeconomic indicators.
@@
 
-<!--# In 1980, poverty rates were relatively low (less than 12%) for all Geomarkets except for East Los Angeles (17%), South and South Central Los Angeles (24%), and Hollywood and Wilshire (18%).
+ In 1980, poverty rates were relatively low (less than 12%) for all Geomarkets except for East Los Angeles (17%), South and South Central Los Angeles (24%), and Hollywood and Wilshire (18%).
 However, poverty rates declined in South and South Central Los Angeles (22%), remained stable in East Los Angeles and Hollywood and Wilshire, and increased in all other Geomarkets.
@@
 
-To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2020-2022 high school graduating classes, and by PSAT scores.
+To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets in the Los Angeles metropolitan areas, we analyze six orders placed by a research university that filtered for prospects across the entire state of California, in the 2020 through 2022 high school graduating classes, and by PSAT scores.
 Two orders indicated PSAT score thresholds ranging from 1070-1180, another two orders filtered for a 1190-1260 PSAT range, and the remaining two orders filtered for scores ranging from 1270-1520.
@@
 Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (89%) and East Los Angeles (64%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (6%) and larger share of Asian prospects for East Los Angeles (30%).
-
 Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic and Black student prospects, while East Los Angeles stands out for its substantial share of Asian prospects.
@@
 
-@fig-la-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions for the Los Angeles metropolitan area, demonstrating South and South Central Los Angeles (7% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (25% versus 17% for all prospects) Geomarkets contributed larger proportions of Black prospects relative to their contributions to the overall prospect pool in the Los Angeles metropolitan area.
-<!--#  For the more than 2,500 White prospects in this pool, we see the San Fernando Valley - West (28% versus 20% for all prospects) and the West Los Angeles and West Beach (12% versus 7% for all prospects)Geomarkets contributing disproportionately larger shares.-->
-
+@fig-la-rq2-midSAT-race shows how much each Geomarket accounts for the share of prospects in the middle PSAT score range by race/ethnicity.
+South and South Central Los Angeles accounts for only 2% of all purchased prospect profiles but accounts for 7% of Black prospects.
+The Reiverside, San Bernardino, \& Ontario Geomarket accounts for 17% of all prospects but 25% of Black prospects.
 For Hispanic prospects, South and South Central Los Angeles (8% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (24% versus 17% for all prospects) also contributed a disproportinately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
-
-<!--# [I THINK CLEANER IF YOU JUST HAVE A SMALL PARAGRAPH FOR HISPANIC AND A SMALL PARAGRAPH FOR BLACK.
-I THINK ONE LEDE ACROSS THESE RESULTS IS RIVERSIDE AS AN ENGINE FOR BLACK/LATINX AND FOR FIRST-GEN STUDENTS.
-I GET THAT THESE TWO POINTS CONTRADICT EACH OTHER].
-Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-la-rq2-midSAT-race.
-South and South Central Los Angeles and Riverside, San Bernadino, and Ontario Geomarkets contributed larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Los Angeles metropolitan area.
-On the other hand, the San Fernando Valley- West and the Glendale and Pasadena Geomarket contributed smaller proportions of Black and Hispanic Prospects.
-However, the West Los Angeles and West Beach (10% versus 7% for all prospects), Hollywood Wilshire (7% versus 5% for all prospects), and the Long Beach (8% versus 5% for all prospects) Geomarkets contributed a larger share of Black prospects, whereas Covina and West Covina contributed a larger share of Hispanic students.
-All other Geomarkets contributed nearly equal shares of Black and Hispanic prospects.
--->
-
-In comparison, the Glendale and Pasadena (17% versus 14% for all prospects), East Los Angeles (11% versus 5% for all prospects) and Covina and West Covina (17% versus 11% for all prospects) Geomarkets contributed a disproportionate larger share of Asian prospects relative to their contributions to the overall prospect pool.
-
-These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Los Angeles metropolitan area using Geomarket filters that exclude lower-income Geomarkets such as South and South Central Los Angeles and Riverside, San Bernadino, and Ontario Geomarkets would result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
-Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.
-
-<!--# [I'M NOT SURE WHERE YOU ARE GETTING THESE NUMBERS FROM; THIS IS THE WRONG FIGURE I THINK.
-AND WHEN YOU REVISE, THE LEDES ARE RIVSERSIDE, EAST LA AND, SOUTH CENTRAL (ONLY 2% OF PROSPECTS BUT 10% OF NO-COLLEGE) KS-- weird...
-overall substantive rankings are correct but my %s are all off].
---> @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
+For Asian students, the Glendale and Pasadena (17% of Asian prospects versus 14% for all prospects), East Los Angeles (11% of Asian prospects versus 5% of all prospects) and Covina and West Covina (17% of Asain prospects versus 11% of all prospects) Geomarkets contributed a disproportionate larger share of Asian prospects relative to their contributions to the overall prospect pool.
+For the more than 2,500 White prospects in this pool, the San Fernando Valley -- West (28% of White prospects versus 20% of all prospects) and the West Los Angeles and West Beach (12% of White prospects versus 7% for all prospects) Geomarkets contribute disproportionately larger shares.
+
+These results demonstrate that following the @RN4982 strategy of filtering on high-income Geomarkets and excluding low-income Geomarkets -- such as the South and South Central Los Angeles Geomarket and the Riverside, San Bernadino, and Ontario Geomarket -- would result in the disproportionate exclusiong of Black and Hispanic students from the resulting prospect pools.Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.
+
+[KARINA -- CHECK YOUR NUMBERS FOR THIS PARAGRAPH] @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
 For students whose parents did not attend college, East Los Angeles (16% versus 5% of all prospects), South and South Central Los Angeles (10% versus 2% of all prospects), and Riverside, San Bernardino, and Ontario (24% versus 17% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
+<!--#  KARINA -- I CUT THIS TEXT.
+OK?
 These three Geomarkets also made disproportionately larger contributions of first-generation college student prospects whose parents attended but did not complete their degree, as well as Hollywood & Wilshire (7% versus 5% of all prospects), Long Beach (7% versus 5% of all prospects), and Covina and West Covina (13% versus 10% of all prospects).
+-->
 
 <!--#  [I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] --> Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
-
-For Hispanic prospects, the South and South Central Los Angeles Geomarket stands out as providing the most substantial share of first-generation college students.
 Across all Hispanic prospects in the middle range PSAT orders, 25% were first-generation college students whose parents did not attend college, 25% were first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
-In comparison, the South and South Central Los Angeles Geomarket contributed much larger disproportionate share of Hispanic first-generation college students whose parents did not attend college (54%), with the Hollywood & Wilshire (30%), East Los Angeles (39%), and Riverside, San Bernardino, and Ontario (31%) also contributing larger but more modest disproportionate shares.
-The East Los Angeles (33%), South and South Central Los Angeles (28%), Long Beach (30%), and Riverside, San Bernardino, and Ontario (27%) also contributed larger Hispanic first-generation college students whose parents attended college but did not complete their their degree.
+By contrast, for the 156 Hispanic prospects from the South and South Central Geomarket, 54% had parents who never attended college, 28% had parents who attended some college, and only 18% had parents with a BA.
+Similarly, Hispanic prospects from East Los Angeles, from Hollywood & Wilshire, and from Riverside, San Bernardino, and Ontario were much more likely to have parents without a BA compared to all Hispanic prospects in the Los Angeles metro area.
 
@@
 
-While the pool of Black prospects is relatively small (n=186) in @fig-la-rq2-midSAT-combo , some patterns emerge by first-generation status.
+While the pool of Black prospects is relatively small (n=186) in @fig-la-rq2-midSAT-combo, some patterns emerge by first-generation status and Geomarket.
 Overall, approximately 5% of Black prospects were first-generation college students-no college, 20% were first-generation college students-some college, and 74% not first-generation).

### 2025-05-22 text/manuscript-full-KS.qmd
went through discussion after karina

@@
 format:
+  #docx:
   pdf:
@@
 
-The discussion of homophily by @RN4975 describes not only the logic of The Market Segment model but also a broader institutional logic that characterizes the enrollment management industry.
-Homophily argues that there is a natural cultural match between students and colleges, suggesting colleges should find prospective students who resemble students who enrolled at their institution in the past and focus recruiting efforts on the same schools and communities that are popular with peer colleges.
+The discussion of homophily by @RN4975 describes not only the logic of The Market Segment model but also a broader institutional logic of the enrollment management industry.
+The logic of homophily assumes that there is a natural cultural match between students and colleges, suggesting colleges should find prospective students who resemble students who enrolled at their institution in the past and focus recruiting efforts on the same schools and communities that are popular with peer colleges.
 We suggest that homophily is the operating logic most consistent with college recruiting behavior observed in the wild.
 For example, @RN5023 find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
-The study also finds great overlap in the schools visited by Villanova University and the University of Notre Dame, with a disproportionate focus on visits to Catholic schools.
-<!--# When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
+<!--# The study also finds great overlap in the schools visited by Villanova University and the University of Notre Dame, with a disproportionate focus on visits to Catholic schools.
+When we observe that students at selective private colleges tend to be rich and that out-of-state students at public research universities tend to be rich [@RN4722], this is partly because these universities devote most of their recruiting expenditures towards students from affluent schools and communities [@RN4758; @RN5023].
 -->  As @RN4975 argues, observed homophily is not the result of voluntary action; rather, homophily is programmed into algorithms that create connections.
 
-The Market Segment Model begins with the status attainment model, that student demand for college is a function of parental education and parental income.
-@RN4982 [chapter 3] illustrated this argument by demonstrating a correlation between student SAT score-sending behavior and measures of class.
-The thesis of the Market Segment Model is that that households of specific class characteristics and college aspirations are likely to live in particular communities within metropolitan areas and that these geographic territories can be meaningfully captured by Geomarkets.
+The Market Segment Model begins with the status attainment model idea that student demand for college is a function of parental education and parental income.
+@RN4982 illustrated this argument by demonstrating a correlation between student SAT score-sending behavior and measures of class.
+The thesis of the Market Segment Model is that that households of specific class characteristics and college aspirations are likely to live in particular communities and that these geographic territories can be meaningfully captured by Geomarkets.
 Although Geomarkets appear coarse by contemporary standards, they were a landmark innovation in Geographic segmentation.
@@
 We find that Geomarkets are highly correlated with race and class.
-<!--#  [KARINA SUMMARIZE FINDINGS IN A COUPLE SENTENCES; REVISE PARAGRAPH AS NECESSARY]--> Findings illustrate that in 1980, when Geomarkets were being created, Black and Hispanic residents tended to be highly concentrated in a single, usually the poorest Geomarket in the Chicago, Dallas-Fortworth, and to some extent the Los Angeles metropolitan areas.
+Findings illustrate that in 1980, when Geomarkets were being created, Black and Hispanic residents tended to be highly concentrated in the poorest Geomarket in the Chicago, Dallas-Fortworth, and to some extent the Los Angeles metropolitan areas.
 These Geomarket patterns extend to other metropolitan areas in the country, such as CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), and PA5 -- Philadelphia County.
-By 2020, various Geomarket landscapes emerged.
-In Chicago, Black residents remain concentrated in the poorest Geomarket.
-In contrast, in Dallas-Fortworth, Black and Hispanic communities are distributed across multiple Geomarkets; however, significant socioeconomic disparities persist between Geomarkets.
+In 2020 Chicago, Black residents remain concentrated in the poorest Geomarket but Gentrification also pushed Black residents to the South and Southwest Suburbs Geomarket.
+In Dallas-Fortworth, Black and Hispanic communities are distributed across multiple Geomarkets; however, significant socioeconomic disparities persist between Geomarkets.
 The Los Angeles metropolitan area presents another distinctive Geomarket landscape in which Black, Hispanic, and Asian communities across a range of socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.
-<!--# [KARINA -- ONE OR TWO SENTENCES] -->
-
-These findings are not surprising given the extent of race and class segregation in the US.
+
+<!--#  These findings are not surprising given the extent of race and class segregation in the US.
 Furthermore, the correlation between Geomarkets and class is expected because Geomarket borders were created in service of conceptual model that views demand for higher education as a function of class.
@@
 @RN4982 do not discuss the intersection of student demand and race.
-Although unsurprising for early 1980s social science, this silence is conspicuous considering the salience of political controversy surrounding "busing" in Boston and nearby metropolitan areas.<!--# KS: COULD CUT THIS IF NEEDED; MAIN TAKEAWAY IS JUST THAT RQ1 IS UNSURPRINSING EMPIRICALLY VIA SEGREGATION AND THEORETICALLY VIA ZEMSKY & OEDEL-->
+Although unsurprising for early 1980s social science, this silence is conspicuous considering the salience of political controversy surrounding "busing" in Boston and nearby metropolitan areas.KS: COULD CUT THIS IF NEEDED; MAIN TAKEAWAY IS JUST THAT RQ1 IS UNSURPRINSING EMPIRICALLY VIA SEGREGATION AND THEORETICALLY VIA ZEMSKY & OEDEL-->
 
 Research question 2 analyzes the racial and socioeconomic composition of student list purchases.
-We examine purchases that include all Geomarkets in a metropolitan area in order to assess the potential consequences if including or excluding particular Geomarkets from the purchase.
+We examine purchases that include all Geomarkets in a metropolitan area in order to assess the potential consequences of including or excluding particular Geomarkets from the purchase.
 This question is motivated by the @RN4982 recommendation that selective colleges that enroll students from the regional and national market segment should target affluent Geomarkets, while community colleges and non-selective 4-year colleges should focus on middle- and working-class Geomarkets.
-Findings across metropolitan cases suggest student list orders that use Geomarket filters to exclude middle- and working-class Geomarkets - such as IL11 -- City of Chicago, TX19 -- City of Dallas, TX20 -- City of Fort Worth, CA19 -- East Los Angeles, and CA21 -- South and South Central Los Angeles, in targeting high-achieving students may still capture some Black, Hispanic, and Asian prospects.
-However, these prospects are far more unlikely to be first-generation students of color.
-<!--# KARINA -- ADD TEXT SUMMARIZING RESULTS.-->
+Findings across metropolitan cases demonstrate that if student list purchases exclude working- and middle-class Geomarkets - such as IL11 -- City of Chicago, TX19 -- City of Dallas, TX20 -- City of Fort Worth, CA19 -- East Los Angeles, and CA21 -- South and South Central Los Angeles -- the resulting student list purchases disproportionately exclude Black, and Hispanic students who meet the requisite academic achievement criteria.
+Although these purchases would still capture some Black, Hispanic, and Asian prospects, these prospects are much less likely to be first-generation students.
 
@@
 The status attainment model argues that college destination is a function of parental education and occupation [@RN1273; @RN1271; @RN1272].
-The Market Segment Model is the status attainment model applied to geodemographic market research \[CITE\].
+We conceive of the the Market Segment Model is the status attainment model applied to geodemographic market research, akin to the rise of credit-scoring in UK retail banking [@RN4880].
 @RN5033 finds that the educational achievement of Asian American students whose parents are immigrants tend to have high educational achievement even if their parents do not, while the educational achievement of later-generation Asian Americans conforms to status attainment theory.
 In our analyses of prospects included in student list purchases, Asian American prospects from poor Geomarkets tended to be first-generation students while Asian American prospects from affluent Geomarkets tended to have parents with a BA.
-Drawing from @RN5033, one explanation for this finding is that Asian students living in low-income Geomarkets tend to have parents who immigrated to the U.S.
-whereas Asian students living in affluent Geomarkets are more likely to have parents who were born in the U.S.
-<!--# KARINA NOTES -- USE RESULTS TO SPEAK TO MAJOR THEORIES.-->
+Building on @RN5033, Asian students living in low-income Geomarkets may tend to have parents who immigrated to the U.S.
+whereas Asian students living in affluent Geomarkets are more likely to have parents born in the U.S.
 
@@
 This model explains how affluent households maintain a disproportionate enrollment share at selective institutions in an era of holistic admissions ostensibly designed to increase racial and class diversity [@RN4991].
-Both the status attainment model and the cultural capital model are demand-side explanations for why education is a "social sieve" [@RN2377; @RN5046] that allows for a modicum of mobility while maintaining a much larger flow of intergenerational class transmission [@RN861].
+Both the status attainment model and the cultural capital model are demand-side explanations for why education is a "social sieve" [@RN2377; @RN5046] that allows for a modicum of mobility while maintaining a larger flow of intergenerational class transmission [@RN861].
 @RN4722 -- who obtained parental income from federal income tax returns for every U.S.
@@
 The credentialism literature recognizes that colleges have a financial incentive for educational credentials to determine the competition for socially and economically advantageous positions [@RN865; @RN861; @RN880; @RN869] and this credentials arms race favors affluent households.
-Scholarship on enrollment management describes which students colleges want to enroll and what colleges do to attract these students at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758; @RN3544; @RN3547; @cottom2017lower].
+Scholarship on enrollment management describes which students colleges want and how colleges go about attracting these students at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758; @RN3544; @RN3547; @cottom2017lower].
 @RN3519 and @RN4321 describe a tacit arrangement between high school guidance counselors on the demand side and college admissions officers on the supply side.
-Counselors at well-resourced schools, especially private schools, are motivated to give their students a competitive advantage in admissions.
+Counselors at well-resourced -- especially private -- schools, are motivated to give their students a competitive advantage in admissions.
 Meanwhile, college admissions officers are motivated to enroll students who can afford full tuition price and are likely to donate in the future.
@@
 @RN4982 argues that demand for higher education is correlated with class, ignoring the historical structural barriers that produced class- and race-based inequality in 1980 student demand.
-In itself, the Market Segment Model is an unremarkable social science depiction.
+In itself, the Market Segment Model is a social science depiction.
 However, the College Board inscribed Geomarkets and the Market Segment Model into EPS software that tells colleges which Geomarkets and high schools to target.
@@
 Geomarkets were subsequently incorporated into the Student Search Service product.
-We show that that excluding low-income, non-white Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-white students with strong test scores.
-<!--# KS Notes: THIS IS THE STRONGEST PARAGRAPH -->
+We show that that excluding low-income, non-white Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-white students with strong academic achievement.
 
@@
 The growing salience of third-party vendors warrants a reconfiguration of the organizational field salient to college access.
-@RN527 \[p.
-148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
-College access scholarship from economics often analyzes the effects of district, state, and federal [e.g., @RN5043] policies and the effects of information interventions [@RN4018].
-These lines of research do not consider the growing role of third-party vendors, which are increasingly owned by private equity interests and offer software-as-service platforms that perform core functions for schools and colleges.^[@RN4991 describes the role of independent admissions consultants hired by families as third-party vendors that engage on the demand-side.]
+@RN527[p.
+148] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
+College access scholarship from economics often analyzes the effects of district, state, and federal policies [e.g., @RN5043].
+These literatures ignore the role of third-party vendors, which are often owned by private equity and offer software-as-service platforms that perform core functions for schools and colleges.^[@RN4991 studies independent admissions consultants, third-party vendors on the demand-side.]
 
@@
 When we issued records requests to public universities about their student list purchases, at least 50% of universities indicated that they outsourced student list purchases to an EM consultancy.
-In many cases, the university was buying names each year but no university employee had knowledge about which vendors they were buying from or which search filters they were utilizing.
+Often, the university was buying names each year but no employee had knowledge about which vendors they were buying from or which search filters they were utilizing.
 
@@
 College Board and ACT leveraged data on test-takers to enjoy a near duopoly in market for student list data, one that lasted for several decades.
-This business can be described as list-based lead generation based on the direct-mail model.
+This business can be described as list-based lead generation based on the direct-mail model [@singer1988].
 "Free" college search engines (e.g., Cappex, Niche) yielded new sources of student list data.
@@
 The most widely-known product is Naviance, which claims to be used by more than 10 million K-12 students and by 40% of US high schools [@powerschool].
-*Naviance* college planning software feeds into the *Intersect* recruiting platform, which allows colleges to target Naviance users while they are on the platform.
-*Intersect* is an example of behavioral-based targeting.
+*Naviance* college planning software feeds into the *Intersect* recruiting platform -- an example of behavioral-based targeting -- which allows colleges to target Naviance users while they are on the platform.
 
 The third dynamic in the EM industry has been the growth of test-optional admissions movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT.
-ACT responded to uncertainty in their core testing business by attempting to become an edtech company \[CITE\], failing \[\], and being acquired by Nexus Capital in 2024 \[CITE\].
-Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters \[CITE\] and ACT's predicted probability of enrollment filter \[CITE\] -- developed free college search engines, and entered the enrollment management consulting market more aggessively.
+ACT responded to uncertainty in their core testing business by attempting to become an edtech company [@blumenstyk_2019], failing [@molnar_2020a], and being acquired by Nexus Capital in 2024 [@knox_2024].
+Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters [@RN4565] -- developed free college search engines, and entered the enrollment management consulting market more aggessively [@list_biz].
 College Board has weathered the test-optional movement thanks to robust revenues from the Advanced Placement product, but the decline in PSAT/SAT/PreACT/ACT test-takers has undermined their oligopoly in the market for student list data.
 
-The fourth dynamic is the transformation from owner-operated to private-equity owned firms and the erosion of the distinction between consultancy and software vendor.
-The 1980s and 1990s was a period of market entry in EM, when college admissions professionals or professors decided to hang up a shingle [@RN5027].
+The fourth -- and most important -- dynamic is the transformation of EM from an owner-operated to private-equity owned industry and the resulting eroding distinction between consultancy and software vendor.
+The 1980s and 1990s was a period of market entry in EM, when college admissions professionals or professors hung up a shingle [@RN5027].
 The 2010s were a period of acquisitions and concentration.
+Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions to grow market share and vertical acquisitions designed to make customers more dependent on particular recruiting products.
 After regulatory scrutiny pushed private-equity interests out of the for-profit college market [@RN5022], private equity investors found value in acquiring firms that provided services to direct providers.
-Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions to grow market share and vertical acquisitions designed to make customers more dependent on particular recruiting products.
-
-EAB exemplifies the the trend towards private equity ownership and industry concentration.
+Direct providers are subsidized by state and federal funds (Title IV financial aid) and these funds are then paid to third-party vendors.
+Therefore, private equity ownership of direct providers and third-party vendors are both examples of government-subsidized rent-seeking.
+
+EAB exemplifies the trend towards private equity ownership and industry concentration.
 EAB is the preeminent EM consultancy, offering software-as-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
 The origins of EAB trace to Bill Royall, who founded Royall & Company in 1983 to provide direct marketing and fundraising for Republican political campaigns [@jump_2020].
-Royall did not sign its first colege client for several years, but EM became the primary focus by 1995.
+By 1995 EM became their primary focus.
 Royall was acquired by the health tech firm Advisory Board Company for \$850 million in 2015 and then by Vista Equity Partners for \$1.5 billion in 2017, renaming it EAB [@drew_hansen_2017].
 
-As an owner-operated consultancy, Royall was the market leader in purchasing student lists from student list vendors on behalf of clients.
+As an owner-operated consultancy, Royall was the market leader in purchasing names from student list vendors on behalf of clients.
 Under Vista, EAB entered the market for student list data by completing a series of acquisitions that targeted the "leads" and "inquiries" stages of the enrollment funnel.
@@
 In 2021, EAB and PowerSchool -- the leading provider of K12 student information systems and a subsidiary of Vista -- jointly acquired the EM consultancy/software vendor Hobsons.
-EAB acquired Hobson's student success software, Starfish, while PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect.
-EAB paid PowerSchool to become the exlusive reseller of Intersect.
-Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.
-EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.^[The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
+PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect and EAB paid PowerSchool to become the exlusive reseller of Intersect.
+<!--Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.--> EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.^[The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
 This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]."]
@@
 Enroll360 exemplifies a new kind of student list product that leverages proprietary data to sell software.
-Historically, College Board and ACT sold prospect contact information at a price-per-prospect (e.g., \$0.50 for each "name").
+Historically, College Board and ACT sold prospect contact information at a price-per-prospect (e.g., \$0.50 per name).
 The goal was to generate revenue from the sale of names.
 By contrast, Enroll360 wraps several proprietary databases of names (e.g., Cappex, Intersect) into a software-as-service platform that recruits those prospects at different stages of the enrollment funnel.
-Here, the goal of lists is to create demand for software, which is more expensive than purchasing names.
+Here, the goal of lists is to create demand for software.
 Colleges that are interested in recruiting these prospects must buy an Enroll360 subscription.
@@
 Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and wheather students used Naviance to "research competitor institutions" [@markup_naviance].
-<!--#  A procurement document from the University of Utah explains that the Intersect subscription is necessary to recruit the proprietary database of Naviance users: [@utah_procurement]:-->
-
-> <!--# The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to.
+A University of Utah procurment document states the Intersect subscription is necessary because "there is unique group of prospective students who are only in the PowerSchool Naviance platform" [@utah_procurement].
+
+<!--# The marketing services being purchased operates within the PowerSchool Naviance platform, which is a proprietary system, that no other company has access to.
 There is unique group of prospective students who are only in the PowerSchool Naviance platform.
@@
 These products identify correlations in a historic snapshot of students and apply those relationships to make recommendations about which prospects to target in the present.
-As such, all of these technologies amplify the effect of previous inequality in student demand, which is itself a consequence of structural inequality in educational opportunity.
-
-Market dynamics in the EM industry raise policy concerns about tuition prices, student choice, and competition in the market for EM consultancies that may also affect competition, prices, and student choice.
+As such, these technologies amplify the effect of previous inequality in student demand, which is itself a consequence of structural inequality in educational opportunity.
+
+Market dynamics in the EM industry raise policy concerns about tuition prices, student choice, and competition.
 To the extent that colleges must buy expensive software to recruit prospects contained in an associated database, this expenditure will ultimately be passed on to students in the form of higher prices or reduced expenditure on education.
@@
 Fewer EM consultancies means less competition and higher consulting fees, which will be passed on to consumers.
-Finally, the blurring of the lines between consultancy and student list vendor suggests that large consultancies will funnel prospects only to those colleges that are willing and able to pay for subscriptions, which raises policy concerns about student choice.
+Finally, the blurring of the lines between consultancy and student list vendor suggests that large consultancies will funnel prospects only to those colleges that pay for subscriptions, raising policy concerns about student choice.
 
@@
 Unfortunatel, scholarship salient to education policy and regulation have ignored two, related economic transformations.
-First, third-party vendors offer software platforms that perform core tasks of organizations, including schools and colleges [@list_biz].
+First, third-party vendors offer software platforms that perform core tasks of organizations, including schools and colleges.
 Second, private firms comprise a growing share of economic activity [@RN5037; @RN5038].
-Most firms in EM and the broader edtech sector are private.
-However, private firms face fewer disclosure requirements than publicly traded firms.
-The paucity of research about third-party vendors that structure college access is partially due to difficulty obtaining data for empirical analyses.
-The research community must figure out how to systematically investigate the role of private firms and investors in education, so that we can hold private interests accountable for education policy goals.
+Most firms in EM and the broader edtech sector are private and face fewer disclosure requirements than publicly traded firms.
+The lack of research about third-party vendors that structure college access is partially due to difficulty obtaining data for empirical analyses.
+In order to hold private interests accountable for education policy goals, the policy community needs the research community to figure out how to systematically investigate the role of private firms and investors in education.
 
@@
 The knowledge infrastructure of education research largely consists of data about students, schools, and colleges.
-It has been shaped by the ascendance of economics [@RN5036], which incorporates district, state, and federal administrative data to evaluate the effects of schools and policies on student outcomes [@RN5043].
-The research community mobilized to investigate for-profit colleges because for-profit colleges and students were included in ongoing data collection [@RN5022].
+It has been shaped by the ascendance of economics [@RN5036], which incorporates district, state, and federal administrative data to evaluate the effects of schools and policies on student outcomes [e.g., @RN5043].
+The research community mobilized to investigate for-profit colleges [e.g., @eaton_agile_predators; @RN5022] because for-profit colleges and students were included in ongoing data collection.
 By contrast, researchers ignored the growth of private, for-profit, third-party vendors (e.g., PowerSchool, Parchment) because these organizations are excluded from data collections known to education researchers.
 
-We argue that sociologists and adjacent scholars should expand the domain of the education research knowledge infrastructure to include third-party vendors in enrollment management, and edtech more broadly.
-However, this is a difficult task.
-Student unit record data owned by district, state, and federal agencies is simultaneously granular -- can evaluate a particular policy -- and macro -- can speak to broad trends in the education sector.
-By contrast, obtaining granular data about third-party products in education is difficult because these data are owned by private interests that do not want researchers to interrogate their products [@RN4774; @RN4842].
-We see two broad avenues for research: case studies involving primary data collection -- including this study -- which tend to have a granular focus on particular products or vendors; and secondary data analysis, which may have a more macro focus.
-
-Although case study research often relies on interview and ethnographic data, another approach to data collection is to issue records requests to public entities that contract with third-party entities.
+We argue that sociologists and adjacent scholars should expand the domain of the education research knowledge infrastructure to include third-party vendors in enrollment management and edtech more broadly.
+This is a challenging tasks.
+Student unit record data owned by district, state, and federal agencies is simultaneously granular (can evaluate a particular policy) and macro (can describe broad trends in the education sector).
+By contrast, obtaining granular data about third-party products is difficult because these data are owned by private interests that do not want researchers to interrogate their products [@RN4774; @RN4842].
+We see two avenues for research: case studies that require primary data collection and tend to have a granular focus on particular products or vendors; and secondary data analysis, which may have a more macro focus.
+
+Although case studies often relies on interview and ethnographic data, another approach is to issue records requests to public entities that contract with third-party entities.
 For example, @RN5039 analyze public university contracts with online program managers (OPMs).
@@
 Qualitative or computational text analysis of publicly available websites and social media provides another avenue for data collection.
-Here, data collection is efficient but data processing is often labor intensive.
-Like investigative reporting [@markup_naviance], case study research can produce granular analyses of particular products (e.g., Intersect, Enroll360) that yield insight into the underlying mechanisms of how third-party vendors structure college access.
-However, data collected for case study research typically does not meet the knowledge infrastructure standard of repeated, ongoing data collection that can be the foundation for scalable research.
-
-We see subscription databases about private markets as having the potential to create a knowledge infrastructure for scalable research on the role of private firms and private equity funders in education, and also their role in other domains of interest to sociology.
-A growing set of "market intelligence" data providers serve the information needs of the investment community by providing data on private firms and deals involving private firms.
+Here, data collection is efficient but data processing may be laborious.
+Like investigative reporting [@markup_naviance], case studies can produce granular analyses of particular products (e.g., Intersect, Enroll360) that yield insight into the mechanisms of how third-party vendors structure college access.
+However, case study data collections do not meet the standard of repeated, ongoing data collection that can be the basis for scalable research.
+
+Subscription databases have the potential to create a knowledge infrastructure for scalable research on the role of private firms and private equity funders in education, and also their role in other domains of interest to sociology [e.g., @eaton_pe_2025].
+A growing set of data providers serve the information needs of the investment community by providing data on private firms and deals involving private firms.
 These databases tend to include three broad "types" of data: (1) *firm-level data* (e.g., location, employees, financials); (2) *acquisitions*, encompassing who acquired whom and resulting ownership structure); and (3) *investments*, which include amounts/valuations/stake percentages of private equity investments.
@@
 We recommend that scholars develop panel datasets examining change over time in the firms operating in particular industries and acquisitions and private equity investments in those industries.
-For example, empirical scholarship can examine change over time in concentration in the enrolment management industry and its transformation from an owner-operated to a largely private-equity owned industry.
+For example, empirical scholarship can examine change over time in concentration in the enrollment management industry and its transformation from an owner-operated to a largely private-equity owned industry.
 Subsequent analyses could investigate the portfolio of large owners, their ties to other industries -- such as the direct-provider for-profit college market -- and the vertical inputs/technologies they are investing in.

### 2025-05-30 text/manuscript-full-KS.qmd
results revisions

@@
 ```
-
 @fig-chicago-rq2-midSAT-race shows which Geomarkets send the bulk of students for a given race/ethnicity.
 The City of Chicago accounts for 17% of all prospects but 48% of Black students.
-Similarly, South \& Southwest Suburbs accounts for 20% of all prospects but 30% of all Black prospects.
+Similarly, South & Southwest Suburbs accounts for 20% of all prospects but 30% of all Black prospects.
 The upshot is that if the student list purchases had excluded these two Geomarkets, 78% of the nearly 1,000 Black students currently included in these purchases would be excluded.
@@
 The City of Chicago Geomarket stands out as the only Geomarket contributing substantially larger shares of Hispanic first-generation college students across those whose parents did not attend college (46% versus 38% overall) and those whose parents attended some college but did not complete their degrees (29% versus 26% overall).
-By contrast, Hispanic students from the more affluent North Shore Geomarket and Evanston \& Skokie Geomarket were more likely to have parents with a BA.
+By contrast, Hispanic students from the more affluent North Shore Geomarket and Evanston & Skokie Geomarket were more likely to have parents with a BA.
 
@@
 In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable through 2020.
-As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 [MISMATCH BETWEEN GRAPH AND TABLE?] median household incomes, respectively.
+As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 \[MISMATCH BETWEEN GRAPH AND TABLE?\] median household incomes, respectively.
 By contrast, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000).
@@
 ```
-
 ## Los Angeles
@@
 @fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020.
-Over this period, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
-\[THE CAVEAT THAT WE DON'T HAVE ASIAN FOR 1980 IS MORE IMPORTANT FOR LA; KARINA -- MENTION THIS OR REVISE HOW YOU WRITE FINDINGS IN THIS PARAGRAPH, LIKE FOCUS MAINLY ON 2000-2020.
-YOU DECIDE.\] In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
+One data caveat here is that the Asian race/ethnicity category was not available in 1980.
+However, over the three decennial census periods, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
+In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
 By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%).
@@
 
+```{=html}
 <!--# KARINA -- CUT THIS CUZ BASICALLY CONSISTENT WITH INCOME.
@@
 -->
-
+```
 <!--# Lastly, @fig-la-rq1-ses shows a widening disparity in educational attainment across the Los Angeles metropolitan area.
@@
 South and South Central Los Angeles accounts for only 2% of all purchased prospect profiles but accounts for 7% of Black prospects.
-The Reiverside, San Bernardino, \& Ontario Geomarket accounts for 17% of all prospects but 25% of Black prospects.
+The Reiverside, San Bernardino, & Ontario Geomarket accounts for 17% of all prospects but 25% of Black prospects.
 For Hispanic prospects, South and South Central Los Angeles (8% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (24% versus 17% for all prospects) also contributed a disproportinately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
@@
 
-[KARINA -- CHECK YOUR NUMBERS FOR THIS PARAGRAPH] @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
+\[KARINA -- CHECK YOUR NUMBERS FOR THIS PARAGRAPH; KS: I DID; THEY ARE CORRECT FOR ORDERS 44820 AND 546946; AM I MISSING SOMETHING/WHY DO YOU THINK THEY ARE WRONG?\] @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
 For students whose parents did not attend college, East Los Angeles (16% versus 5% of all prospects), South and South Central Los Angeles (10% versus 2% of all prospects), and Riverside, San Bernardino, and Ontario (24% versus 17% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
@@
 
-<!--#  [I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] --> Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
+<!--#  [I THINK THE WAY YOU WRITE UP THIS FIGURE SHOULD BE CHANGED MORE SUBSTANTIALLY; I ADD SOME COMMENTS BELOW] -->
+
+Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
 Across all Hispanic prospects in the middle range PSAT orders, 25% were first-generation college students whose parents did not attend college, 25% were first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
@@
 The growing salience of third-party vendors warrants a reconfiguration of the organizational field salient to college access.
-@RN527[p.
-148] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
+@RN527 \[p.
+148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
 College access scholarship from economics often analyzes the effects of district, state, and federal policies [e.g., @RN5043].
-These literatures ignore the role of third-party vendors, which are often owned by private equity and offer software-as-service platforms that perform core functions for schools and colleges.^[@RN4991 studies independent admissions consultants, third-party vendors on the demand-side.]
+These literatures ignore the role of third-party vendors, which are often owned by private equity and offer software-as-service platforms that perform core functions for schools and colleges.[^21]
+
+[^21]: @RN4991 studies independent admissions consultants, third-party vendors on the demand-side.
 
@@
 PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect and EAB paid PowerSchool to become the exlusive reseller of Intersect.
-<!--Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.--> EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.^[The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
-This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]."]
+<!--Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.--> EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.[^22]
+
+[^22]: The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
+This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]."
 
@@
 
+```{=tex}
 \begingroup
 \small
-
+```
 ::: {#refs}
@@
 
+```{=tex}
 \endgroup
@@
 \clearpage
-
+```
 <!-- TABLES -->

### 2025-06-25 scripts/save_tracts.R
add tract to EPS mappings

@@ -1,0 +1,130 @@
+# https://stackoverflow.com/a/65935636
+
+library(tidyverse)
+library(shiny)
+library(leaflet)
+library(sf)
+
+
+data_dir <- file.path('..', 'data')
+eps_data_dir <- file.path(data_dir, 'eps_market')
+tracts_dir <- file.path(eps_data_dir, 'tracts')
+
+# Load county shapes for reference
+load(file.path(eps_data_dir, 'counties_2020.RData'))
+
+# Load tract shapes to trace
+load(file.path(eps_data_dir, 'tracts_1980.RData'))
+load(file.path(eps_data_dir, 'tracts_2000.RData'))
+load(file.path(eps_data_dir, 'tracts_2020.RData'))
+
+metro <- tracts_2020 %>% filter(str_detect(GEOID, '^17(097|031|043|197)'))
+# metro <- tracts_2000 %>% filter(str_detect(gisjoin2, '^170(097|031|043|197)'))
+# metro <- tracts_1980 %>% filter(str_detect(gisjoin2, '^170(097|031|043|197)'))
+
+# Load previously traced shapes for reference
+eps_2020 <- tracts_2020 %>%
+  mutate(eps = case_when(
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL 7.txt')), ',')[[1]] ~ 'IL 7',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL 8.txt')), ',')[[1]] ~ 'IL 8',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL 9.txt')), ',')[[1]] ~ 'IL 9',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL10.txt')), ',')[[1]] ~ 'IL10',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL11.txt')), ',')[[1]] ~ 'IL11',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL12.txt')), ',')[[1]] ~ 'IL12',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL13.txt')), ',')[[1]] ~ 'IL13'
+  )) %>%
+  filter(!is.na(eps)) %>%
+  group_by(eps) %>%
[…]

### 2025-06-25 scripts/save_eps_shapes.R
save EPS shapes based on tracts

@@ -1,0 +1,181 @@
+library(tidyverse)
+library(tidycensus)
+library(sf)
+library(leaflet)
+
+
+data_dir <- file.path('.', 'data')
+eps_data_dir <- file.path(data_dir, 'eps_market')
+tracts_dir <- file.path(eps_data_dir, 'tracts')
+
+scripts_dir <- file.path('.', 'scripts')
+
+shape_dir <- file.path('..', 'cb_geomarket_shape')
+
+
+source(file.path(scripts_dir, 'metro_eps_codes.R'))
+
+
+# ------------
+# Save shapes
+# ------------
+
+# tracts_2000.RData and zcta_2010.RData are too big to push to GitHub
+# See: https://www.dropbox.com/scl/fo/6a48fl8isgd3wuhxjiiaa/AOtyIPHmvLuixYU5_VyFmHs?rlkey=ucr4rkbtmoi2pg4hnbkcgsa1r&st=5q4x20sp&dl=0
+
+# Tract shapes (1980)
+# d1980_tract_sf <- st_read(file.path(shape_dir, 'nhgis0001_shape', 'nhgis0001_shapefile_tl2000_us_tract_1980', 'US_tract_1980.shp')) %>%
+#   rename_with(str_to_lower)
+# 
+# d1980_bna_sf <- st_read(file.path(shape_dir, 'nhgis0001_shape', 'nhgis0001_shapefile_tl2000_us_tract_1980', 'US_bna_1980.shp')) %>%
+#   rename_with(str_to_lower)
+# 
+# tracts_1980 <- bind_rows(d1980_tract_sf, d1980_bna_sf) %>%
+#   st_as_sf(sf_column_name = 'geometry') %>%
+#   st_transform(crs = 4326)
+# 
+# save(tracts_1980, file = file.path(eps_data_dir, 'tracts_1980.RData'))
+
+# Tract shapes (2000)
+# tracts_2000 <- st_read(file.path(shape_dir, '2000_decennial', 'US_tract_2000.shp')) %>%
+#   rename_with(str_to_lower) %>%
+#   st_as_sf(sf_column_name = 'geometry') %>%
+#   st_transform(crs = 4326)
+# 
+# save(tracts_2000, file = file.path(eps_data_dir, 'tracts_2000.RData'))
+
+# Tract shapes (2020)
+# tracts_2020 <- do.call(
+#   bind_rows,
[…]

### 2025-06-25 scripts/geomarket_shapes.Rmd
add EPS summary report

---
title: 'Geomarket Shapes'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    df_print: paged
---

```{r setup, include = F}
options(width = 10000)
knitr::opts_chunk$set(warning = F, message = F, echo = F, comment = '')

library(tidyverse)
library(readxl)
library(leaflet)

scripts_dir <- '.'

eps_data_dir <- file.path('..', 'data', 'eps_market')
eps_image_dir <- file.path('..', 'assets', 'images', 'eps')

shape_dir <- file.path('..', '..', 'cb_geomarket_shape')
```

```{r}
# 2010 zip code to EPS mapping
eps_zip <- read_excel(path = file.path(eps_data_dir, 'eps.xls'), sheet = 'Sheet 1', col_names = T) %>% 
  mutate(eps = if_else(str_sub(eps, 3, 3) == '0', str_c(str_sub(eps, 1, 2), ' ', str_sub(eps, 4, 4)), eps))

# Zip code shapes (2010)
load(file.path(eps_data_dir, 'zcta_2010.RData'))
eps_zcta <- zcta_2010 %>% merge(y = eps_zip, by.x = 'ZCTA5', by.y = 'zip', all.x = T) %>% 
  sf::st_as_sf()

# EPS shapes derived from zip codes
load(file.path(eps_data_dir, 'eps_shapes.RData'))

# Tract shapes
load(file.path(shape_dir, 'analysis_data', 'allyr_anal_tract_sf.RData'))

# EPS shapes (1980)
load(file.path(eps_data_dir, 'eps_shapes_1980.RData'))

# EPS shapes (2000)
load(file.path(eps_data_dir, 'eps_shapes_2000.RData'))

# EPS shapes (2020)
load(file.path(eps_data_dir, 'eps_shapes_2020.RData'))

# County shapes (2020)
load(file.path(eps_data_dir, 'counties_2020.RData'))
```

```{r}
plot_map <- function(state_fips, eps_codes, palette = RColorBrewer::brewer.pal(11, 'Spectral')) {

  m <- leaflet() %>% 
    addProviderTiles(providers$CartoDB.Positron) %>% 
    addPolygons(data = counties_2020 %>% filter(STATEFP %in% state_fips), label = ~NAMELSAD, color = 'gray', weight = 1, group = 'Counties')

  bbox <- sf::st_bbox(eps_geometry_zcta %>% filter(eps %in% eps_codes)) %>% as.vector()

  for (i in seq_along(eps_codes)) {
    m <- m %>% 
      addPolygons(data = eps_zcta %>% filter(eps == eps_codes[[i]]), label = ~paste0('<b>', eps, '</b>: ', NAME) %>% lapply(htmltools::HTML), color = if_else(i %in% 5:7, '#c4c4c4', palette[[i]]), fillColor = palette[[i]], weight = 1, group = 'EPS from 2010 zip codes') %>% 
      addPolygons(data = allyr_anal_tract_sf %>% filter(eps == eps_codes[[i]], year == '1980'), label = ~paste0('<b>', eps, '</b>: ', gisjoin) %>% lapply(htmltools::HTML), color = if_else(i %in% 5:7, '#c4c4c4', palette[[i]]), fillColor = palette[[i]], weight = 1, group = 'EPS from 1980 tracts') %>% 
      addPolygons(data = allyr_anal_tract_sf %>% filter(eps == eps_codes[[i]], year == '2000'), label = ~paste0('<b>', eps, '</b>: ', gisjoin) %>% lapply(htmltools::HTML), color = if_else(i %in% 5:7, '#c4c4c4', palette[[i]]), fillColor = palette[[i]], weight = 1, group = 'EPS from 2000 tracts') %>% 
      addPolygons(data = allyr_anal_tract_sf %>% filter(eps == eps_codes[[i]], year == '2020'), label = ~paste0('<b>', eps, '</b>: ', geoid) %>% lapply(htmltools::HTML), color = if_else(i %in% 5:7, '#c4c4c4', palette[[i]]), fillColor = palette[[i]], weight = 1, group = 'EPS from 2020 tracts')
  }

  m <- m %>% 
    addPolylines(data = eps_geometry_zcta %>% filter(eps %in% eps_codes), color = 'black', opacity = 1, weight = 2, group = 'EPS from 2010 zip codes') %>% 
    addPolylines(data = eps_1980 %>% filter(eps %in% eps_codes), color = 'black', opacity = 1, weight = 2, group = 'EPS from 1980 tracts') %>% 
    addPolylines(data = eps_2000 %>% filter(eps %in% eps_codes), color = 'black', opacity = 1, weight = 2, group = 'EPS from 2000 tracts') %>% 
    addPolylines(data = eps_2020 %>% filter(eps %in% eps_codes), color = 'black', opacity = 1, weight = 2, group = 'EPS from 2020 tracts') %>% 
    addLayersControl(
      position = 'bottomleft',
      overlayGroups = c('EPS from 2010 zip codes', 'EPS from 1980 tracts', 'EPS from 2000 tracts', 'EPS from 2020 tracts', 'Counties'),
      options = layersControlOptions(collapsed = F)
    ) %>% 
    fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
    hideGroup('EPS from 1980 tracts') %>% 
    hideGroup('EPS from 2000 tracts') %>% 
    hideGroup('EPS from 2020 tracts') %>% 
    hideGroup('Counties')

  m
}
```

```{r}
source(file.path(scripts_dir, 'metro_eps_codes.R'))

all_codes <- list(
  philadelphia = list(name = 'Philadelphia', eps = philly_eps_codes),
  dallas = list(name = 'Dallas', eps = dallas_eps_codes),
  atlanta = list(name = 'Atlanta', eps = atl_eps_codes),
  chicago = list(name = 'Chicago', eps = chi_eps_codes),
  cleveland = list(name = 'Cleveland', eps = cleveland_eps_codes),
  northern_new_jersey = list(name = 'Northern New Jersey', eps = nj_north_metro_eps_codes),
  houston = list(name = 'Houston', eps = houston_eps_codes),
  bay_area = list(name = 'Bay Area', eps = bay_area_eps_codes),
  long_island = list(name = 'Long Island', eps = long_island_eps_codes),
  new_york_city = list(name = 'New York, New York', eps = nyny_metro_eps_codes),
  detroit = list(name = 'Detroit', eps = detroit_eps_codes),
  boston = list(name = 'Boston', eps = boston_eps_codes),
  miami = list(name = 'Miami', eps = miami_eps_codes),
  dc_maryland_virginia = list(name = 'DMV', eps = dmv_eps_codes),
  orange_county = list(name = 'Orange County', eps = orange_county_eps_codes),
  san_diego = list(name = 'San Diego', eps = san_diego_eps_codes),
  los_angeles = list(name = 'Los Angeles', eps = los_angeles_eps_codes)
)

for (metro in all_codes[order(names(all_codes))]) {
  writeLines(str_c(metro$name, ': ', str_c(metro$eps, collapse = ', ')))
}
```

# Atlanta

- **GA 1**: Cherokee, Cobb, and Douglas Counties
- **GA 2**: Fulton County
- **GA 3**: DeKalb and Gwinnett Counties
- **GA 4**: Clayton, Fayette, Henry, and Rockdale Counties

```{r}
plot_map('13', all_codes$atlanta$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/georgia.png){height=300px}
  ![](../assets/images/eps/georgia_cb.png){height=300px}
</details><br>

# Bay Area

- **CA 4**: Marin County
- **CA 5**: San Francisco County
- **CA 6**: Contra Costa County
- **CA 7**: City of Oakland
- **CA 8**: Alameda County excluding Oakland
- **CA 9**: San Mateo County
- **CA10**: City of San Jose
- **CA11**: Santa Clara County excluding San Jose

```{r}
plot_map('06', all_codes$bay_area$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/bay_area.png){height=300px}
  ![](../assets/images/eps/california_cb.png){height=300px}
</details><br>

# Boston

- **MA 6**: Boston and Cambridge
- **MA 8**: Lowell, Concord, and Wellesley
- **MA10**: Milton, Lexington, and Waltham

```{r}
plot_map('25', all_codes$boston$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/massachusetts.png){height=250px}
  ![](../assets/images/eps/massachusetts_cb.png){height=250px}
</details><br>

# Chicago

- **IL 7**: Chain of Lakes
- **IL 8**: Northwest Suburbs
- **IL 9**: North Shore
- **IL10**: Evanston and Skokie
- **IL11**: City of Chicago
- **IL12**: Western Suburbs
- **IL13**: South and Southwest Suburbs

```{r}
plot_map('17', all_codes$chicago$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/chicago.png){height=350px}
  ![](../assets/images/eps/chicago_cb.png){height=350px}
</details><br>

# Cleveland

- **OH 2**: North Central Ohio
- **OH 3**: City of Cleveland (West)
- **OH 4**: City of Cleveland (East)
- **OH 5**: Cuyahoga, Geauga, and Lake Counties
- **OH 6**: Northeast Ohio

```{r}
plot_map('39', all_codes$cleveland$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/ohio.png){height=300px}
  ![](../assets/images/eps/ohio_cb.png){height=300px}
</details><br>

# Dallas

- **TX19**: City of Dallas
- **TX20**: City of Fort Worth
- **TX21**: Irving, Arlington, and Grand Prairie
- **TX22**: Dallas County excluding City of Dallas
- **TX23**: Collin and Rockwall Counties
- **TX24**: Counties West of Dallas/Ft. Worth Metroplex

```{r}
plot_map('48', all_codes$dallas$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/texas.png){height=250px}
  ![](../assets/images/eps/texas_dallas_cb.png){height=250px}
</details><br>

# DMV

- **DC 1**: District of Columbia
- **MD 2**: Montgomery Metropolitan
- **MD 3**: Central Maryland excluding Baltimore
- **MD 5**: Prince Georges Metropolitan
- **MD 7**: Baltimore (Urban)
- **VA 1**: Arlington and Alexandria
- **VA 2**: Fairfax County

```{r}
plot_map(c('11', '24', '51'), all_codes$dc_maryland_virginia$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/virginia.png){height=200px}
  ![](../assets/images/eps/maryland.png){height=200px}
  ![](../assets/images/eps/virginia_cb.png){height=200px}
  ![](../assets/images/eps/maryland_cb.png){height=200px}
</details><br>

# Detroit

- **MI 1**: Wayne County
- **MI 2**: Detroit's Northern Suburbs
- **MI 3**: Ann Arbor

```{r}
plot_map('26', all_codes$detroit$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/michigan.png){height=300px}
  ![](../assets/images/eps/michigan_cb.png){height=300px}
</details><br>

# Houston

- **TX15**: Northwest Houston and Conroe School District
- **TX16**: Southwest Houston Metro Area
- **TX17**: City of Houston (East)
- **TX18**: Galveston and East Harris Counties

```{r}
plot_map('48', all_codes$houston$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/texas.png){height=300px}
  ![](../assets/images/eps/texas_houston_cb.png){height=300px}
</details><br>

# Long Island

- **NY16**: Southern Nassau County
- **NY17**: Northern Nassau County
- **NY18**: Central Nassau County
- **NY19**: Northwest Suffolk County
- **NY20**: Southwest Suffolk County
- **NY21**: East Suffolk County

```{r}
plot_map('36', all_codes$long_island$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/new_york_city.png){width=400px}
</details><br>

# Los Angeles

- **CA14**: San Fernando Valley (West)
- **CA15**: San Fernando Valley (East)
- **CA16**: Glendale and Pasadena
- **CA17**: West Los Angeles and West Beach
- **CA18**: Hollywood and Wilshire
- **CA19**: East Los Angeles
- **CA20**: South Bay
- **CA21**: South and South Central Los Angeles
- **CA22**: Long Beach
- **CA23**: Covina and West Covina
- **CA27**: Riverside, San Bernardino, and Ontario

```{r}
plot_map('06', all_codes$los_angeles$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/los_angeles.png){height=300px}
  ![](../assets/images/eps/los_angeles_cb.png){height=300px}
  ![](../assets/images/eps/california.png){height=300px}
  ![](../assets/images/eps/california_cb.png){height=300px}
</details><br>

**Notes**:

- Check intersection of CA18, CA19, CA21 for 2000
- Check intersection of CA20 and CA22 for 1980 and 2000

# Miami

- **FL 5**: Broward, Martin, and Palm Beach Counties
- **FL 6**: Dade County
- **FL 7**: Collier, Hendry, and Monroe Counties

```{r}
plot_map('12', all_codes$miami$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/florida.png){height=300px}
  ![](../assets/images/eps/florida_cb.png){height=300px}
</details><br>

# New York

- **NY14**: Staten Island
- **NY15**: Westchester County
- **NY22**: Southeast Brooklyn
- **NY23**: West Brooklyn
- **NY24**: Northeast Brooklyn
- **NY25**: East Bronx
- **NY26**: West Bronx
- **NY27**: Manhattan
- **NY28**: South Queens
- **NY29**: Northwest Queens
- **NY30**: Northeast Queens

```{r}
plot_map('36', all_codes$new_york_city$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/new_york_city.png){width=400px}
</details><br>

# Northern New Jersey

- **NJ 4**: Middlesex County
- **NJ 6**: Somerset and Mercer Counties
- **NJ 7**: Union County
- **NJ 8**: Essex and Southern Passaic County
- **NJ 9**: Hudson County
- **NJ10**: Bergen County
- **NJ11**: Morris and Northern Passaic County

```{r}
plot_map('34', all_codes$northern_new_jersey$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/new_jersey.png){height=300px}
  ![](../assets/images/eps/new_jersey_north_cb.png){height=300px}
  ![](../assets/images/eps/new_jersey_south_cb.png){height=300px}
</details><br>

# Orange County

- **CA24**: Whittier and North Orange County
- **CA25**: Anaheim
- **CA26**: Santa Ana
- **CA28**: South Orange County

```{r}
plot_map('06', all_codes$orange_county$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/los_angeles.png){height=300px}
  ![](../assets/images/eps/california.png){height=300px}
  ![](../assets/images/eps/los_angeles_cb.png){height=300px}
  ![](../assets/images/eps/california_cb.png){height=300px}
</details><br>

# Philadelphia

- **PA 1**: Bucks County
- **PA 2**: Chester County
- **PA 3**: Delaware County
- **PA 4**: Montgomery County
- **PA 5**: Philadelphia County

```{r}
plot_map('42', all_codes$philadelphia$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/pennsylvania.png){height=250px}
  ![](../assets/images/eps/pennsylvania_cb.png){height=250px}
</details><br>

# San Diego

- **CA29**: North San Diego County excluding San Diego
- **CA30**: South San Diego County excluding San Diego
- **CA31**: City of San Diego

```{r}
plot_map('06', all_codes$san_diego$eps)
```

<details>
  <summary>**EPS geomarket map**</summary>
  ![](../assets/images/eps/california.png){height=300px}
  ![](../assets/images/eps/california_cb.png){height=300px}
</details><br>

### 2025-06-26 scripts/append_census.R
save combined EPS and tract data

@@ -119,6 +119,8 @@ options(scipen = 999)
   
   allyr_anal_eps_sf %>% glimpse()
   
+  save(allyr_anal_eps_sf, file = file.path(shape_dir, 'analysis_data', 'allyr_anal_eps_sf.RData'))
+  
   rm(d1980_stf1f3_anal_eps_sf,d2000_sf1a_anal_eps_sf,acs2020_ab_anal_eps_sf)
 
 ########### LOAD TRACT-LEVEL DATA
@@ -182,6 +184,8 @@ options(scipen = 999)
   allyr_anal_tract_sf <- st_transform(allyr_anal_tract_sf, crs = 4326)
   
   #allyr_anal_tract_sf <- st_make_valid(allyr_anal_tract_sf) # is the issue that geometries are not valid or that they are GEOMETRYCOLLECTION
+  
+  save(allyr_anal_tract_sf, file = file.path(shape_dir, 'analysis_data', 'allyr_anal_tract_sf.RData'))
   
   rm(d1980_stf1f3_anal_tract_sf,d2000_sf1a_anal_tract_sf,acs2020_ab_anal_tract_sf_fixed)
   

### 2025-06-26 scripts/create_1980_decennial.R

@@ -60,7 +60,8 @@ scripts_dir <- file.path('.','scripts') #
 
 ###### LOAD EPS SHAPE FILE DATA
   
-  load(file.path(eps_data_dir, 'eps_shapes.RData'))
+  load(file.path(eps_data_dir, 'eps_shapes_1980.RData'))
+  eps_geometry_zcta <- eps_1980
   
 ####### READ IN DATA
 
@@ -671,7 +672,7 @@ scripts_dir <- file.path('.','scripts') #
     households_tot = if_else(dig001==0,NA,dec001), # total number of households
     inc_house_med_all  = if_else(die001==0,NA,die001), # median household income
     inc_house_agg_all = if_else(dig001==0,NA,dig001), # aggregate household income
-    households_tot_calc = rowSums(select(., did001:did016), na.rm = TRUE), # calculated number of households that have values for household income bands
+    households_tot_calc = rowSums(select(., did001:did017), na.rm = TRUE), # calculated number of households that have values for household income bands
     #inc_house_mean_all = inc_house_agg_all/households_tot_calc, # mean income at tract-level, based on calculated number of houses that report income
     
     pov_denom = if_else(is.na(di8001) & is.na(di8002), NA, 

### 2025-06-26 scripts/create_2000_decennial.R

@@ -53,7 +53,8 @@ list.files(path = scripts_dir)
 
 ###### LOAD EPS SHAPE FILE DATA
 
-  load(file.path(eps_data_dir, 'eps_shapes.RData'))
+  load(file.path(eps_data_dir, 'eps_shapes_2000.RData'))
+  eps_geometry_zcta <- eps_2000
 
 ######## READ IN SHAPE FILE FOR 2000 CENSUS TRACTS
 
@@ -666,9 +667,9 @@ d2000_sf1a_tract_sf %>% select(gisjoin) %>% as.data.frame() %>% group_by(gisjoin
     st_transform(5070)
   
   # save in analysis data file
-  save(d2000_sf1a_anal_eps_sf, file = file.path(data_dir,'analysis_data', 'd2000_sf1a_anal_eps_sf.RData'))
+  save(d2000_sf1a_anal_eps_sf, file = file.path(shape_dir,'analysis_data', 'd2000_sf1a_anal_eps_sf.RData'))
   
-  load(file = file.path(data_dir,'analysis_data', 'd2000_sf1a_anal_eps_sf.RData'))
+  load(file = file.path(shape_dir,'analysis_data', 'd2000_sf1a_anal_eps_sf.RData'))
   
   # results for "important" vars
   d2000_sf1a_anal_eps %>% 

### 2025-06-26 scripts/create_2020_acs.R

@@ -53,7 +53,8 @@ list.files(path = scripts_dir)
 
 ###### LOAD EPS SHAPE FILE DATA
 
-  load(file.path(eps_data_dir, 'eps_shapes.RData'))
+load(file.path(eps_data_dir, 'eps_shapes_2020.RData'))
+eps_geometry_zcta <- eps_2020
 
 ######## READ IN SHAPE FILE FOR 2020 CENSUS TRACTS
 
@@ -130,28 +131,32 @@ acs2020_vars %>% filter(name %in% variables) %>% print()
 # READ IN TRACT-LEVEL 2016-2020 ACS DATA FROM TIDYCENSUS
 
   # List of all state abbreviations plus Washington, D.C.
-  states <- c(state.abb, "DC")
-  states
+  # states <- c(state.abb, "DC")
+  # states
+  # 
+  # # Initialize an empty list to store data
+  # all_data <- list()
+  # 
+  # # Loop through each state to get tract-level data
+  # for (state in states) {
+  #   state_data <- get_acs(
+  #     geography = "tract",
+  #     state = state,
+  #     variables = variables,
+  #     year = 2020,
+  #     survey = "acs5",
+  #     output = "wide"
+  #   )
+  #   all_data[[state]] <- state_data
+  # }
+  # 
+  # # Combine all the state data into one data frame
+  # acs2020_ab_tract <- bind_rows(all_data)
+  # rm(states,state,all_data,state_data)
   
-  # Initialize an empty list to store data
-  all_data <- list()
+  # save(acs2020_ab_tract, file = file.path(shape_dir, '2020', 'acs2020_ab_tract_raw.RData'))
   
-  # Loop through each state to get tract-level data
-  for (state in states) {
-    state_data <- get_acs(
-      geography = "tract",
-      state = state,
-      variables = variables,
-      year = 2020,
-      survey = "acs5",
-      output = "wide"
-    )
-    all_data[[state]] <- state_data
-  }
-
-  # Combine all the state data into one data frame
-  acs2020_ab_tract <- bind_rows(all_data)
-  rm(states,state,all_data,state_data)
+  load(file.path(shape_dir, '2020', 'acs2020_ab_tract_raw.RData'))
   
   acs2020_ab_tract %>% glimpse()
 
@@ -291,6 +296,7 @@ st_crs(eps_geometry_zcta) == st_crs(acs2020_ab_tract_sf)
[…]

### 2025-06-27 index.qmd
update rq1 graphs + tables

@@
   table_df <- readRDS(file.path(tables_dir, str_c('rq1_table_', metro, '.rds')))
-  table_df[is.na(table_df)] <- NA_real_
+  table_df$value[is.na(table_df$value)] <- NA_real_
+
+  var_order <- c(
+    '% White, non-Hispanic', '% Asian, non-Hispanic', '% Black, non-Hispanic', '% Hispanic',
+    '% Two+, non-Hispanic', '% NHPI, non-Hispanic', '% AIAN, non-Hispanic',
+    'Median income', 'Mean income', '% in poverty', '% with BA+'
+  )
+
+  race_df <- table_df %>% filter(statistic == 'sum', str_detect(variable, '^[^%]*Hispanic'), !str_detect(variable, 'API')) %>%
+    group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
+    mutate(
+      variable = str_c('% ', variable),
+      value = if_else((year == '1980' & str_detect(variable, 'Asian|NHPI|AIAN|Two')), NA_real_, value)
+    )
+
+  pov_df <- table_df %>% filter(statistic == 'sum', variable %in% c('Poverty', 'Not poverty')) %>%
+    group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
+    filter(variable == 'Poverty') %>% 
+    mutate(variable = '% in poverty')
+
+  edu_df <- table_df %>% filter(statistic == 'sum', variable %in% c('Less than HS', 'HS', 'Less than BA', 'BA+')) %>%
+    group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
+    filter(variable == 'BA+') %>% 
+    mutate(variable = '% with BA+')
+
+  table_df <- bind_rows(
+    race_df, pov_df, edu_df,
+    table_df %>% filter(variable %in% var_order, statistic != 'sum')
+  ) %>% 
+    mutate(
+      statistic = factor(statistic, levels = c('sum', 'mean', 'sd', 'p25', 'p50', 'p75')),
+      variable = factor(variable, levels = var_order),
+      value = sprintf('%.1f', value)
+    ) %>% 
+    arrange(desc(year), statistic) %>% 
+    pivot_wider(id_cols = c(eps_codename, variable), names_from = c(statistic, year), values_from = value) %>% 
+    arrange(variable, desc(eps_codename))
 
@@
 
-  kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Mean', 'SD', 'P25', 'P50', 'P75'), 3))) %>% 
+  kbl(table_df %>% select(-variable), escape = F, col.names = c('', rep(c('Wt. Mean', 'Mean', 'SD', 'P25', 'P50', 'P75'), 3))) %>% 
     pack_rows(index = headings_group, colnum = 1, label_row_css = '') %>% 
-    add_header_above(header = c(' ' = 1, '1980' = 5, '2000' = 5, '2020' = 5))
+    add_header_above(header = c(' ' = 1, '1980' = 6, '2000' = 6, '2020' = 6))
 }

### 2025-06-27 scripts/rq1_tables_graphs.R

@@ -46,9 +46,8 @@ source(file = file.path('scripts', 'directories.R'))
 
 ######## DATA PREP
 
-# run script that appends data
-list.files(path = file.path('.',scripts_dir))
-source(file = file.path(scripts_dir, 'append_census.R'))
+# load tract-level data
+load(file.path(shape_dir, 'analysis_data', 'allyr_anal_tract_sf.RData'))
 
 # script that creates character vectors for EPS codes
 source(file = file.path(scripts_dir, 'metro_eps_codes.R'))
@@ -228,13 +227,19 @@ graph_var_names <- list(
   "pct_nhisp_multi" = "% Two+, non-Hispanic",
   "mean_inc_house" = "Mean income",
   "med_inc_house" = "Median income",
+  "pov_yes" = "Poverty",
+  "pov_no" = "Not poverty",
   "pct_pov_yes" = "% in poverty",
+  "edu_lths_all" = "Less than HS",
+  "edu_hs_all" = "HS",
+  "edu_ltba_all" = "Less than BA",
+  "edu_baplus_all" = "BA+",
   "pct_edu_baplus_all" = "% with BA+"
 )
 graph_var_names
 
 # Define graph_vars in the desired order
-graph_varlist <- c("nhisp_white", "nhisp_black", "hisp_all", "nhisp_asian", "nhisp_nhpi", "nhisp_api", "nhisp_native", "nhisp_multi", "pct_nhisp_white", "pct_nhisp_black", "pct_hisp_all", "pct_nhisp_asian", "pct_nhisp_nhpi", "pct_nhisp_api", "pct_nhisp_native", "pct_nhisp_multi", "med_inc_house", "mean_inc_house", "pct_pov_yes", "pct_edu_baplus_all")
+graph_varlist <- c("nhisp_white", "nhisp_black", "hisp_all", "nhisp_asian", "nhisp_nhpi", "nhisp_api", "nhisp_native", "nhisp_multi", "pct_nhisp_white", "pct_nhisp_black", "pct_hisp_all", "pct_nhisp_asian", "pct_nhisp_nhpi", "pct_nhisp_api", "pct_nhisp_native", "pct_nhisp_multi", "med_inc_house", "mean_inc_house", "pov_no", "pov_yes", "pct_pov_yes", "edu_hs_all", "edu_ltba_all", "edu_baplus_all", "edu_lths_all", "pct_edu_baplus_all")
 
 create_eps_graph <- function(eps_codes, 
                              prefix = NULL,
[…]

### 2025-06-27 index.qmd
update rq2 graphs + tables

@@
   } else if (table_type == 'race_firstgen') {
+    x <- unique(table_df$eps_codename)  # Move "All" row to the end after rest of EPS rows
+    table_df$eps_codename <- factor(table_df$eps_codename, levels = c(x[x != 'All'], 'All'))
+    table_df <- table_df %>% arrange(order_ids, stu_race_cb, eps_codename)
+
     group_df <- table_df %>% group_by(order_ids, stu_race_cb) %>% summarise(n = n())

### 2025-06-27 scripts/create_stu_list.R

@@ -223,7 +223,8 @@ source(file = file.path(scripts_dir, 'metro_eps_codes.R'))
 ###### LOAD EPS SHAPE FILE DATA
 
 
-load(file.path(eps_data_dir, 'eps_shapes.RData'))
+load(file.path(eps_data_dir, 'eps_shapes_2020.RData'))
+eps_geometry_zcta <- eps_2020
 
 eps_geometry_zcta %>% class()
 eps_geometry_zcta %>% glimpse()

### 2025-06-27 index.qmd
remove mean income from rq1 table

@@
     '% Two+, non-Hispanic', '% NHPI, non-Hispanic', '% AIAN, non-Hispanic',
-    'Median income', 'Mean income', '% in poverty', '% with BA+'
+    'Median income', '% in poverty', '% with BA+'
   )

### 2025-07-02 text/manuscript-full-KS.qmd
methods revisions

@@
 Utilizing a small geographic area is important for implementing a spatial merge between Census data and Geomarket shape files (described below).
-Census tract is the smallest geographic area for which measures of both race/ethnicity and socioeconomic characteristics are available.
-Census tracts are statistical subdivisions of a county designed to capture relatively homogeneous demographic communities that are generally considered to be more accurate for population analyses than zip codes, which are spatial boundaries drawn to demarcate route efficiency for postal services \[[CITE](https://www.census.gov/newsroom/blogs/random-samplings/2014/07/understanding-geographic-relationships-counties-places-tracts-and-more.html)\] .
+Census tracts are the smallest geographic area for which measures of both race/ethnicity and socioeconomic characteristics are available.
+Defined as statistical subdivisions of counties, census tracts boundaries are reassessed every decennial census to reflect relatively homogeneous demographic communities, typically ranging in population size from 2,500 and 8,000 residents [@Bureau2014; @Bureau1994].
+Compared to zip codes, which are fixed spatial units designed primarily to demarcate route efficiency for postal services, census tracts are considered more reliable for population-based analyses.
 
@@
 
-We assigned each census tract to College Board Geomarkets by implementing a partial spatial merge.
-Spatial manipulations utilized functions from the *sf* R package.
-Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.[^17] Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
-Using these zip code-level data, we utilized the `aggregate()` function to create shapefiles for each Geomarket.
-
-[^17]: The author of the post reported that he found the zip codes associated with Geomarkets from a Google search.
-
-Next, we utilized the `st_intersection()` function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
+We assigned each census tract to a College Board Geomarket, which we constructed through manual digitization of maps published by the College Board in 2023 marketing materials promoting their Enrollment Planning Service (EPS) software products.
+Specifically, we georeferenced the College Board’s Geomarket maps by overlaying PDF images onto spatial coordinates aligned with census tract boundaries.
+We then extracted vector data by manually tracing the Geomarket boundaries and generating individual shapefiles for each Geomarket.
+
+This approach to constructing Geomarkets aligned precisely with 2020 Decennial Census tract boundaries; that is, each census tract corresponded uniquely to a Geomarket generated from the College Board images.
+However, this process was less precise when applied to the 1980 and 2000 Decennial Census tract boundaries, likely a factor of census tract delineations shifting over time based on demographic changes and our use of map images from 2023.
+For instance, census tracts that experience population growth exceeding the 8,000 resident threshold can be split into two or more smaller census tracts to maintain consistent population ranges.
+Conversely, two or more adjacent census tracts experiencing population declines can sometimes be merged to form a single census tract whose population falls within the 2,500 to 8,000 resident range.
+
+Some census tract boundaries from the 1980 and 2000 Decennial Censuses intersected multiple Geomarkets (i.e., tracts whose boundaries spanned two or more Geomarkets), resulting in spatial misalignment.
+For tracts divided across multiple Geomarkets, we assigned each tract to the Geomarket that encompassed the majority of its spatial area.
+This method introduced slight differences to Geomarket shapes across decennial censuses.
+However, to assess the robustness of our approach to spatial misalignments, we tested alternative methods for resolving ambiguous tract assignments to Geomarkets and found overall results remained relatively consistent across approaches.[^17]
+
+[^17]: An alternative approach to dealing with spatial misalignments included creating shapefiles for Geomarkets from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.
+Essentially, the data file was a Geomarket-zipcode crosswalk that held one observation per zip code and a column that assigned each zip code to one Geomarket.
+Using these zip code-level data, we utilized the aggregate() function to create shapefiles for each Geomarket.
+Next, we utilized the st_intersection() function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
+This spatial merge kept Geomarket boundaries consistent across all three Decennial Census and split census-tract shapefiles that intersected multiple Geomarkets into smaller shapefiles, each entirely contained within a single Geomarket.
+We replaced each variable in these shapefiles with its original value scaled by the share of the tract’s land area that lay within the Geomarket.
+For a particular census tract, imagine that 60% of its land area was contained in Geomarket A, 40% of its land area was contained in Geomarket B, and the census tract reports 100 people who identify as Asian, non-Hispanic.
+The partial spatial merge splits this census tract into two observations.
+The observation assigned to Geomarket A reports $.60*100 = 60$ Asian, non-Hispanic people and the observation assigned to Geomarket B reports $.40*100 = 40$ Asian, non-Hispanic people.
+Even under this most conservative approach to addressing spatial misalignment, overall patterns in the findings remained consistent.
+However, we opted to georeference the manually digitized maps directly sourced from the College Board, as the original source of the Geomarket-zipcode crosswalk cited in the the R-bloggers post could not be verified.
+
+```{=html}
+<!--#We assigned each census tract to College Board Geomarkets by implementing spatial merges.
+Spatial manipulations utilized functions from the sf R package.
+Data to create the shapefiles for the College Board Geomarkets were obtained from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.
+Essentially, the data file had one observation per zip code and a column that assigned each zip code to one Geomarket.
+Using these zip code-level data, we utilized the aggregate() function to create shapefiles for each Geomarket.
+
+Next, we utilized the st_intersection() function to assign each census tract to a Geomarket, based on the intersection of their associated spatial shapefiles.
 We split census-tract shapefiles that intersected multiple Geomarkets (i.e., tracts whose boundaries spanned two or more Geomarkets) into smaller shapefiles, each entirely contained within a single Geomarket.
@@
 The observation assigned to Geomarket A reports $.60*100 = 60$ Asian, non-Hispanic people and the observation assigned to Geomarket B reports $.40*100 = 40$ Asian, non-Hispanic people.
+--> 
+```
 
@@
 Metropolitan areas are cases.
-Case studies deepen understanding of phenomena by comparing and contrasting across different settings \[CITE YIN 2001\].
+Case studies deepen understanding of phenomena by comparing and contrasting across different settings [@RN4011].
 Our research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
@@
 -->
-
 ```

### 2025-07-03 index.qmd
update maps

@@
 
-      # Example special case for "los_angeles"
-      if (m == "los_angeles") {
-        rq1_map <- 'https://rq1-map-los-angeles.netlify.app'
+      # For maps too big to push to GitHub
+      if (!m %in% c("atlanta", "boston", "orange_county", "philadelphia", "san_diego")) {
+        rq1_map <- paste0('https://rq1-map-', str_replace_all(m, '_', '-'), '.netlify.app')
       }

### 2025-07-03 scripts/map_functions.R

@@ -1,32 +1,17 @@
 library(tidyverse)
 library(rmapshaper)
+library(formattable)
 
 
 source(file.path('scripts', 'directories.R'))
-source(file.path(scripts_dir, 'append_census.R'))
+# source(file.path(scripts_dir, 'append_census.R'))
 source(file.path(scripts_dir, 'metro_eps_codes.R'))
 
-all_codes <- list(
-  philadelphia = list(name = 'Philadelphia', eps = philly_eps_codes),
-  dallas = list(name = 'Dallas', eps = dallas_eps_codes),
-  atlanta = list(name = 'Atlanta', eps = atl_eps_codes),
-  chicago = list(name = 'Chicago', eps = chi_eps_codes),
-  cleveland = list(name = 'Cleveland', eps = cleveland_eps_codes),
-  northern_new_jersey = list(name = 'Northern New Jersey', eps = nj_north_metro_eps_codes),
-  houston = list(name = 'Houston', eps = houston_eps_codes),
-  bay_area = list(name = 'Bay Area', eps = bay_area_eps_codes),
-  long_island = list(name = 'Long Island', eps = long_island_eps_codes),
-  new_york_city = list(name = 'New York, New York', eps = nyny_metro_eps_codes),
-  detroit = list(name = 'Detroit', eps = detroit_eps_codes),
-  boston = list(name = 'Boston', eps = boston_eps_codes),
-  miami = list(name = 'Miami', eps = miami_eps_codes),
-  dc_maryland_virginia = list(name = 'DMV', eps = dmv_eps_codes),
-  orange_county = list(name = 'Orange County', eps = orange_county_eps_codes),
-  san_diego = list(name = 'San Diego', eps = san_diego_eps_codes),
-  los_angeles = list(name = 'Los Angeles', eps = los_angeles_eps_codes)
-)
+load(file.path(shape_dir, 'analysis_data', 'allyr_anal_tract_sf.RData'))
+load(file.path(shape_dir, 'analysis_data', 'allyr_anal_eps_sf.RData'))
 
-sf_use_s2(T)
+
+# sf_use_s2(T)
 
 regions_data <- data.frame(
   region = flatten_chr(map(names(all_codes), \(x) rep(x, length(all_codes[[x]]$eps)))),
[…]

### 2025-07-03 scripts/metro_eps_codes.R

@@ -141,3 +141,23 @@ san_diego_eps_codes
 # North San Diego County CA29
 # South San Diego County excluding San Diego CA30
 # City of San Diego CA31
+
+all_codes <- list(
+  philadelphia = list(name = 'Philadelphia', eps = philly_eps_codes),
+  dallas = list(name = 'Dallas', eps = dallas_eps_codes),
+  atlanta = list(name = 'Atlanta', eps = atl_eps_codes),
+  chicago = list(name = 'Chicago', eps = chi_eps_codes),
+  cleveland = list(name = 'Cleveland', eps = cleveland_eps_codes),
+  northern_new_jersey = list(name = 'Northern New Jersey', eps = nj_north_metro_eps_codes),
+  houston = list(name = 'Houston', eps = houston_eps_codes),
+  bay_area = list(name = 'Bay Area', eps = bay_area_eps_codes),
+  long_island = list(name = 'Long Island', eps = long_island_eps_codes),
+  new_york_city = list(name = 'New York, New York', eps = nyny_metro_eps_codes),
+  detroit = list(name = 'Detroit', eps = detroit_eps_codes),
+  boston = list(name = 'Boston', eps = boston_eps_codes),
+  miami = list(name = 'Miami', eps = miami_eps_codes),
+  dc_maryland_virginia = list(name = 'DMV', eps = dmv_eps_codes),
+  orange_county = list(name = 'Orange County', eps = orange_county_eps_codes),
+  san_diego = list(name = 'San Diego', eps = san_diego_eps_codes),
+  los_angeles = list(name = 'Los Angeles', eps = los_angeles_eps_codes)
+)

### 2025-07-03 scripts/rq1_maps.R

@@ -17,15 +17,11 @@ create_rq1_map <- function(metros, shared_legend = F) {
   choices <- list(region_choices = regions_data %>% filter(region %in% metros) %>% select(region, region_name, latitude, longitude))
   
   highlight_shp <- highlightOptions(weight = 2, color = '#606060', dashArray = '', bringToFront = T, sendToBack = T)
-  
+
   yrs <- c(1980, 2000, 2020)
   
-  m <- leaflet() %>%
-    addProviderTiles(providers$CartoDB.Positron) # %>%
-    
-    # addEasyButton(easyButton(
-    #   icon = 'fa-globe', title = 'Select Metro Area',
-    #   onClick = JS("function(btn, map){ $('.custom-control').not('#metro-control').slideUp(); $('#metro-control').slideToggle(); }")))
+  m <- leaflet(options = leafletOptions(zoomControl = T, zoomSnap = 0, zoomDelta = 0.5)) %>%
+    addProviderTiles(providers$CartoDB.Positron)
   
   for (metro in metros) {
     print(metro)
@@ -34,16 +30,16 @@ create_rq1_map <- function(metros, shared_legend = F) {
     eps <- eps_data %>% filter(eps %in% region$eps[[1]])
     tract <- tract_data %>% filter(eps %in% region$eps[[1]])
     
-    m <- m %>%
-      
-      # Metro outline
-      addPolygons(data = eps %>% filter(year == 2020), opacity = 1, color = 'purple', fillOpacity = 0, weight = 2, label = ~paste0('<b style="font-size:11px">', eps, ' - ', eps_name, '</b>') %>% lapply(htmltools::HTML), group = 'MSA', options = c(className = paste0('metro-shape metro-line metro-', metro)))
-    
     for (y in yrs) {
       m <- m %>%
         
+        # Metro outline
+        addPolygons(data = eps %>% filter(year == y), opacity = 1, color = 'purple', fillOpacity = 0, weight = 2, label = ~paste0('<b style="font-size:11px">', eps, ' - ', eps_name, '</b>') %>% lapply(htmltools::HTML), group = 'MSA', options = c(className = paste0('metro-shape metro-', metro, ' year-', y))) %>% 
+        
         # EPS outline
-        addPolylines(data = eps %>% filter(year == y), opacity = 1, color = 'purple', weight = 2, options = c(className = paste0('metro-shape metro-', metro, ' level-tract year-', y)))
[…]

### 2025-07-03 scripts/rq1_maps.js

@@ -16,26 +16,12 @@ function(el, x, choices) {
   $('.easy-button-button').css('width', 'auto');
   $('.easy-button-button .button-state .fa').css({'float': 'left', 'margin-top': '8px'});
   
-  // $('button[title="Select Metro Area"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Metro Area</span>');
-  
-  // metro selection options
-  
-  // let metroControlHTML = '<div id="metro-control" class="leaflet-control-layers leaflet-control leaflet-control-layers-expanded custom-control" style="width: auto; height: auto;">';
-  
-  // region_name.forEach(function(curr, idx) {
-  //   metroControlHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="metro-choice" data-region="' + region[idx] + '" data-lat="' + latitude[idx] + '" data-lng="' + longitude[idx] + '"><span> ' + curr + '</span></div>';
-  // });
-    
-  // metroControlHTML += '</div>';
-  
-  // $('button[title="Select Metro Area"]').parent().after(metroControlHTML);
-  
   // year selection options
   
   let yearControlHTML = '<p style="margin: 5px; font-weight: 600;">Year</p><div style="display: flex;">';
   
   ['1980', '2000', '2020'].forEach(function(curr, idx) {
-    yearControlHTML += '<div' + (idx === 0 ? '' : ' style="margin-left: 5px;"') + '><input type="radio" class="leaflet-control-layers-selector" name="year-choice" data-year="' + curr + '"><span>' + curr + '</span></div>';
+    yearControlHTML += '<div' + (idx === 0 ? '' : ' style="margin-left: 5px;"') + '><input type="radio" class="leaflet-control-layers-selector" name="year-choice" data-year="' + curr + '"><span> ' + curr + '</span></div>';
   });
   
   yearControlHTML += '</div><p style="margin: 5px; font-weight: 600;">Census data</p>';
@@ -47,12 +33,17 @@ function(el, x, choices) {
   let levelControlHTML = '<p style="margin: 5px; font-weight: 600;">Level</p><div style="display: flex; margin-bottom: 5px;">';
   
   ['EPS', 'Tract'].forEach(function(curr, idx) {
[…]

### 2025-07-03 scripts/rq2_maps.R

@@ -246,7 +246,7 @@ edu_vars <- data.frame(
   name = c('All', 'No college', 'Some college', 'No or some college', 'BA+', 'Unknown/Missing')
 )
 
-regions_data[regions_data$region == 'nyny', 'region'] <- 'nyny'  # for the AIAN graphs
+# regions_data[regions_data$region == 'new_york_city', 'region'] <- 'nyny'  # for the AIAN graphs
 
 
 # Map functions
@@ -266,12 +266,8 @@ create_rq2_map <- function(metros, map_type = 'rq2') {
   
   highlight_shp <- highlightOptions(weight = 1, color = '#606060', dashArray = '')
   
-  m <- leaflet() %>%
+  m <- leaflet(options = leafletOptions(zoomControl = T, zoomSnap = 0, zoomDelta = 0.5)) %>%
     addProviderTiles(providers$CartoDB.Positron) %>%
-    
-    # addEasyButton(easyButton(
-    #   icon = 'fa-globe', title = 'Select Metro Area',
-    #   onClick = JS("function(btn, map){ $('.custom-control').not('#metro-control').slideUp(); $('#metro-control').slideToggle(); }"))) %>% 
     
     addEasyButton(easyButton(
       icon = 'fa-file', title = 'Select Order Number',
@@ -294,7 +290,9 @@ create_rq2_map <- function(metros, map_type = 'rq2') {
       addPolygons(data = eps, opacity = 1, color = 'purple', fillOpacity = 0, weight = 2, label = ~paste0('<b style="font-size:11px">', eps, ' - ', eps_name, '</b>') %>% lapply(htmltools::HTML), group = 'MSA', options = c(className = paste0('metro-shape metro-', metro))) %>% 
       
       # EPS outline
-      addPolylines(data = eps, opacity = 1, color = 'purple', weight = 2, options = c(className = paste0('metro-shape metro-line-', metro)))
+      addPolylines(data = eps, opacity = 1, color = 'purple', weight = 2, options = c(className = paste0('metro-shape metro-line-', metro))) %>% 
+      
+      addLabelOnlyMarkers(data = st_point_on_surface(eps), label = ~paste0('<b style="font-size:11px">', eps, ' - ', eps_name, '</b>') %>% lapply(htmltools::HTML), labelOptions = labelOptions(noHide = TRUE, direction = 'top', className = 'label'))
     
     for (v in names(base_vars)) {
       

### 2025-07-03 scripts/rq2_maps.js

@@ -26,20 +26,7 @@ function(el, x, choices) {
   $('.easy-button-button').css('width', 'auto');
   $('.easy-button-button .button-state .fa').css({'float': 'left', 'margin-top': '8px'});
   
-  // $('button[title="Select Metro Area"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Metro Area</span>');
   $('button[title="Select Order Number"] .button-state').append('<span style="display: inline-block; float: left; padding-left: 5px;">Select Order Number</span>');
-  
-  // metro selection options
-  
-  // let metroControlHTML = '<div id="metro-control" class="leaflet-control-layers leaflet-control leaflet-control-layers-expanded custom-control" style="width: auto; height: auto;">';
-  
-  // region_name.forEach(function(curr, idx) {
-  //   metroControlHTML += '<div><input type="radio" class="leaflet-control-layers-selector" name="metro-choice" data-region="' + region[idx] + '" data-lat="' + latitude[idx] + '" data-lng="' + longitude[idx] + '"><span> ' + curr + '</span></div>';
-  // });
-    
-  // metroControlHTML += '</div>';
-  
-  // $('button[title="Select Metro Area"]').parent().after(metroControlHTML);
   
   // order selection options
   
@@ -85,6 +72,11 @@ function(el, x, choices) {
   $('#hs-race-container').slideUp(0);
   $('#hs-edu-container').slideUp(0);
   
+  // label toggle button
+  let labelHTML = '<div style="height: 0; border-top: 1px solid #ddd; margin: 5px -10px 5px -6px;"></div><div><input type="checkbox" class="leaflet-control-layers-selector" id="label-toggle"><span> Show EPS labels</span></div>'
+  
+  $('.leaflet-control-layers-overlays').append(labelHTML);
+  
   // handle collapsible selection
   
   $('.collapsible-label').on('click', function(e) {
@@ -115,25 +107,6 @@ function(el, x, choices) {
     $('input[data-order="' + active_attr.active_order + '"]').trigger('click');
   }
     
-  // $('input[name="metro-choice"]').on('change', function(e) {
-  //   let $this = $(this);
-    
-  //   let metro = $this.attr('data-region'),
-  //       lat = $this.attr('data-lat'),
-  //       lng = $this.attr('data-lng');
-        
-  //   active_attr.active_metro = metro;
-  //   update_base_layer();
-
[…]

### 2025-07-07 text/manuscript-full-KS.qmd
methods revisions

@@
 
-We assigned each census tract to a College Board Geomarket, which we constructed through manual digitization of maps published by the College Board in 2023 marketing materials promoting their Enrollment Planning Service (EPS) software products.
+We assigned each census tract to a College Board Geomarket, which we constructed through manual digitization of maps published by The College Board in 2023 marketing materials promoting their Enrollment Planning Service (EPS) software products.
 Specifically, we georeferenced the College Board’s Geomarket maps by overlaying PDF images onto spatial coordinates aligned with census tract boundaries.
 We then extracted vector data by manually tracing the Geomarket boundaries and generating individual shapefiles for each Geomarket.
-
-This approach to constructing Geomarkets aligned precisely with 2020 Decennial Census tract boundaries; that is, each census tract corresponded uniquely to a Geomarket generated from the College Board images.
+For instance, @fig-geomarket-methods shows The College Board's map of Geomarkets in the Chicago metropolitan area on the left and our created shapefiles on the right.
+
+This approach to constructing Geomarkets aligned precisely with 2020 census tract boundaries from ACS 5-year estimates; that is, each census tract corresponded uniquely to a Geomarket generated from the College Board images.
+For example, @fig-geomarket-misalign zooms into the northern Geomarkets of the Chicago metropolitan area from @fig-geomarket-methods.
+The center map showcases how our created Geomarket shapefiles (delineated in purple) align with the spatial coordinates of 2020 census tract boundaries (delineated in grey).
 However, this process was less precise when applied to the 1980 and 2000 Decennial Census tract boundaries, likely a factor of census tract delineations shifting over time based on demographic changes and our use of map images from 2023.
-For instance, census tracts that experience population growth exceeding the 8,000 resident threshold can be split into two or more smaller census tracts to maintain consistent population ranges.
+The map on the right in @fig-geomarket-misalign deilneates 1980 census tract boundaries, which are visibly larger and more spatially aggregated than the smaller, more granular 2020 tracts shown in the center map.
+
+<!--# census tracts that experience population growth exceeding the 8,000 resident threshold can be split into two or more smaller census tracts to maintain consistent population ranges.
 Conversely, two or more adjacent census tracts experiencing population declines can sometimes be merged to form a single census tract whose population falls within the 2,500 to 8,000 resident range.
+-->
 
 Some census tract boundaries from the 1980 and 2000 Decennial Censuses intersected multiple Geomarkets (i.e., tracts whose boundaries spanned two or more Geomarkets), resulting in spatial misalignment.
-For tracts divided across multiple Geomarkets, we assigned each tract to the Geomarket that encompassed the majority of its spatial area.
-This method introduced slight differences to Geomarket shapes across decennial censuses.
-However, to assess the robustness of our approach to spatial misalignments, we tested alternative methods for resolving ambiguous tract assignments to Geomarkets and found overall results remained relatively consistent across approaches.[^17]
+In such cases, we assigned each tract to the Geomarket that encompassed the majority of its spatial area.
+When the spatial majority was relatively balanced across multiple Geomarkets, we assigned tracts in a way that maintained Geomarket boundaries that most smoothly followed Geomarket delineations sourced from The College Board.
+While this approach minimized ambiguity, it introduced minor variations in Geomarket shapes across decennial censuses.
+We illustrate this process in @fig-geomarket-misalign, using a 1980 census tract in Chicago as an example, where our constructed shapefiles for the Chain of Lakes and North Shore Geomarkets do not align with the tract boundary (circled in red).
+In this case, because the majority of the tract’s spatial area fell within the Chain of Lakes Geomarket, we assigned the tract to Chain of Lakes rather than to North Shore.
+This assignment, along with others like it, slightly alters the overall boundaries of Geomarkets when compared to those based on 2020 census tract delineations.
+To assess the robustness of our approach to spatial misalignments, we tested alternative methods for resolving ambiguous tract assignments to Geomarkets and found overall patterns in the results remained relatively consistent across approaches.[^17]
 
@@
 ```
-
 **RQ2**.
@@
 
+```{r fig-geomarket-methods, echo=FALSE,out.width="49%", fig.cap="caption",fig.align='center',fig.cap = "Chicago, College Board and Recreated Geomarkets"}
+knitr::include_graphics('./../assets/images/eps/chicago_geos.png')
+
+```
+
+```{r fig-geomarket-misalign, echo=FALSE,out.width="100%", fig.cap="caption",fig.align='center',fig.cap = "Chicago, College Board and Recreated Geomarket Spatial Misalignment"}
+knitr::include_graphics('./../assets/images/eps/chicago_misalign.png')
+```
+
 ```{r fig-chicago-maps, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarkets, Total Population 1980-2020", out.width = "100%"}

### 2025-07-07 text/manuscript-full-KS.qmd
revisions rq1 Chicago

@@
 Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
-In contrast, the most significant population increases occured in the Western Suburbs (1.2 Million to 1.6 Million) and the South and Southwest Suburbs (less than 1 Million to 1.3 Million) Geomarkets.
-The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.1 Million in 1980 to 2.8 Million in 2020, despite remaining the most populous Geomarket.
+In contrast, the most significant population increases occured in the Western Suburbs (1.0 Million to 1.4 Million) and the South and Southwest Suburbs (less than 1 Million to 1.2 Million) Geomarkets.
+The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.5 Million in 1980 to 3.2 Million in 2020, despite remaining the most populous Geomarket.
 
@@
 @fig-chicago-rq1-race illustrates the racial and ethnic composition of each Geomarket across the three census periods.
-In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 83% in the South and Southwest Suburbs to approximately 93% in North Shore.
-In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 46% of its total population.
+In 1980, nearly all Geomarkets were predominantly White, with proportions ranging from 84% in the South and Southwest Suburbs to approximately 96% in North Shore.
+In contrast, the City of Chicago had a substantially more diverse population, with White residents comprising only 52% of its total population.
 One data caveat is that the following race/ethnicity categories were not available in 1980: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic.
 By 2020, several Geomarkets that were previously predominantly White shifted to a more heterogeneous racial/ethnic distribution, including Chain of Lakes, Northwest Suburbs, Evanston and Skokie, Western Suburbs, and South and Southwest Suburbs.
-For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 19 percentage point increase for Western Suburbs.
+For many Geomarkets, this balance in racial/ethnic composition was the result of large increases in the Latinx population from 1980 to 2020, which ranged from a 7 percentage point increase for the Evanston and Skokie Geomarket to a 20 percentage point increase for Chain of Lakes.
 Some Geomarkets also experienced a substantial increase in Black populations (South and Southwest Suburbs) and in Asian populations (Northwest Suburbs and Evanston and Skokie).
-@fig-chicago-rq1-race also indicates that, although the North Shore and City of Chicago Geomarkets underwent some changes in racial and ethnic composition, their overall demographic patterns remained relatively the same across the three census periods.
+@fig-chicago-rq1-race also indicates that, although the North Shore Geomarket underwent some changes in racial and ethnic composition, its overall demographic patterns remained relatively the same across the three census periods.
+City of Chicago experienced relatively modest declines in its Black population (35% in 1980 to 26% in 2020), alongside increases in its Hispanic (13% in 1980 to 30% in 2020) and Asian Pacific Islander populations (4% in 2000 to 7% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
 <!-- For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
@@
 We explore socioeconomic demographics -- median income, percent living below poverty, and educational attainment -- within and across Geomarkets over time in @fig-chicago-rq1-ses.
-The primary takeaways are the the City of Chicago -- and to a somewhat lesser degree th South and Southwest Suburbs - stands out as having lower median income, higher poverty rates, and lower BA attainment than other Geomarkets.
-By contrast, the North Shore stands outas being particularly affluent and highly educated.
+The primary takeaways are the City of Chicago -- and to a somewhat lesser degree th South and Southwest Suburbs - stands out as having lower median income, higher poverty rates, and lower BA attainment than other Geomarkets.
+By contrast, the North Shore stands out as being particularly affluent and highly educated.
 Overall, median income both within and across Geomarkets in the Chicago metropolitan area remained relatively stable across the three census periods.
-For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$158,000 in the North Shore Geomarket to a low of \$66,000 in the City of Chicago Geomarket.
-However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence, such as \$120,000 in the Northwest Suburbs, \$119,000 in Evanston and Skokie, \$108,000 in Chain of Lakes, \$105,000 in the Western Suburbs, and \$99,000 in the South and Southwest Suburbs.
-While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$79,000) and decrease in the South and Southwest Suburbs Geomarket (\$93,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$14,000.
+For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$168,000 in the North Shore Geomarket to a low of \$68,000 in the City of Chicago Geomarket.
+However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence, such as \$131,000 in Evanston and Skokie, \$120,000 in the Northwest Suburbs, \$111,000 in the Western Suburbs, \$109,000 in Chain of Lakes, and \$99,000 in the South and Southwest Suburbs.
+While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$80,000) and decrease in the South and Southwest Suburbs Geomarket (\$91,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$11,000.
 @fig-chicago-rq1-ses also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
 Similar patterns are evident in educational attainment across Geomarkets.
-In 1980, the City of Chicago had the lowest percentage of residents with a bachelor's degree or higher (13%).
-However, by 2020, the Geomarket experienced the largest percentage point increase in college degree attainment, reaching 38%.
+In 1980, the City of Chicago and South and Southwest Suburbabs had the lowest percentage of residents with a bachelor's degree or higher (14%).
+However, by 2020, the City of Chicago experienced the largest percentage point increase in college degree attainment, reaching 34%.
 
@@
 ```{r fig-chicago-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Chicago Area Geomarkets, 1980-2020", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq1_chicago_race.png')
+knitr::include_graphics('./../results/graphs/rq1/rq1_chicago_race.png')
+#knitr::include_graphics('./../assets/images/rq1_chicago_race.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-chicago-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Chicago Area Geomarkets, 1980-2020", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq1_chicago_ses.png')
+knitr::include_graphics('./../results/graphs/rq1/rq1_chicago_ses.png')
+#knitr::include_graphics('./../assets/images/rq1_chicago_ses.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-07-07 text/manuscript-full-KS.qmd
revisions rq2 Chicago

@@
 At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
-For example, more than 8,000 of the 15,981 total prospects in this SAT range order were concentrated in the Western Suburbs (56% White, 8% Asian, 6% Black, 26% Hispanic, 4% Multiracial) and South and Southwest Suburbs (55% White, 3% Asian, 17% Black, 22% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
-The North Shore (78% White, 8% Asian, 2% Black, 8% Hispanic) and the Northwest Suburbs (61% White, 11% Asian, 2% Black, 23% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects.
-Unlike any other Geomarket, nearly nine of every 10 prospects from the City of Chicago Geomarket were Hispanic (50%), Black (30%), Asian (5%), or Multiracial (2%).
-
-As PSAT score ranges increase, The City of Chicago Geomarket becomes the only Geomarket with substantial shares of Black, and Hispanic prospects.
+For example, more than 7,000 of the 15,964 total prospects in this SAT range order were concentrated in the Western Suburbs (59% White, 9% Asian, 6% Black, 23% Hispanic, 4% Multiracial) and South and Southwest Suburbs (54% White, 2% Asian, 19% Black, 21% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
+The North Shore (70% White, 12% Asian, 4% Black, 11% Hispanic) and the Northwest Suburbs (59% White, 11% Asian, 2% Black, 24% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects.
+Unlike any other Geomarket, nearly eight of every 10 prospects from the City of Chicago Geomarket were Hispanic (47%), Black (24%), Asian (5%), or Multiracial (3%).
+
+As PSAT score ranges increase, The City of Chicago Geomarket becomes the only Geomarket with substantial shares of Black and Hispanic prospects.
 Moving from the top to the bottom panels, all Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline.
-For the highest SAT score range, the City of Chicago Geomarket pool is 18% Asian, 7% Black, 21% Hispanic, 5% Multiracial, making it the only Geomarket that is not predominantly White (49%).
-Prospect pools from all other Geomarkets in this high PSAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.4% (North Shore) to 5% (South and Southwest Suburbs) of prospects and 6% (Northwest Suburbs) to 10% (South and Southwest Suburbs) of prospects, respectively.
+For the highest SAT score range, the City of Chicago is the only Geomarket where White prospects comprise a slim majority (55%), with the remaining prospect pool made up of 16% Asian, 6% Black, 18% Hispanic, 5% Multiracial students.
+Prospect pools from all other Geomarkets in this high PSAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.3% (North Shore) to 4% (South and Southwest Suburbs) of prospects and 5% (North Shore) to 10% (South and Southwest Suburbs) of prospects, respectively.
 <!--# [LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN] -->
@@
 @fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle range SAT scores (1160-1300) by race/ethnicity.
-The top panel presents the full distribution of the 13,643 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 11% of all purchased prospect profiles reside in Chain of Lakes, 13% in Northwest Suburbs, 5% in North Shore, 5% in Evanston and Skokie, 17% in City of Chicago, 28% in Western Suburbs, and 20% in South and Southwest Suburbs.
+The top panel presents the full distribution of the 13,729 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 10% of all purchased prospect profiles reside in Chain of Lakes, 11% in Northwest Suburbs, 6% in North Shore, 4% in Evanston and Skokie, 23% in City of Chicago, 29% in Western Suburbs, and 17% in South and Southwest Suburbs.
 These overall distributions are compared to those in the lower panels of @fig-chicago-rq2-midSAT-race, which illustrate the proportional representation of each race/ethnicity group across Geomarkets.
@@
 @fig-chicago-rq2-midSAT-race shows which Geomarkets send the bulk of students for a given race/ethnicity.
-The City of Chicago accounts for 17% of all prospects but 48% of Black students.
-Similarly, South & Southwest Suburbs accounts for 20% of all prospects but 30% of all Black prospects.
-The upshot is that if the student list purchases had excluded these two Geomarkets, 78% of the nearly 1,000 Black students currently included in these purchases would be excluded.
+The City of Chicago accounts for 23% of all prospects but 51% of Black students.
+Similarly, South & Southwest Suburbs accounts for 17% of all prospects but 26% of all Black prospects.
+The upshot is that if the student list purchases had excluded these two Geomarkets, 77% of the nearly 1,000 Black students currently included in these purchases would be excluded.
 On the other hand, all other Geomarkets contributed substantially smaller shares of Black prospects.
-For example, the Northwest Suburbs accounts for about 13% of all prospects but only about 3% of Black prospects.
-The City of Chicago Geomarket also contributed larger shares of the nearly 3,000 Hispanic students (39% versus 17% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
+For example, the Northwest Suburbs accounts for about 11% of all prospects but only about 3% of Black prospects.
+The City of Chicago Geomarket also contributed larger shares of the more than 2,500 Hispanic students (45% versus 23% for all prospects), while all other Geomarkets contributed comparable or relatively smaller proportions of Hispanic prospects than their overall representation among all included prospects.
 
@@
 @fig-chicago-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates proportional to their overall contributions to all included prospects (except City of Chicago).
-Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (19% versus 13% for all prospects) and Evanston and Skokie (10% versus 5% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (8% versus 10% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
+Similar patterns are observed for Asian students, with proportional contributions by Geomarket largely consistent with their overall distribution, except for Northwest Suburbs (17% versus 11% for all prospects) and Evanston and Skokie (9% versus 4% for all prospects) contributing disproportionately larger shares and the South and Southwest Suburbs (6% versus 17% for all prospects) contributing smaller shares relative to their overall contribution to the larger prospect pool.
 
@@
 These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Chicago metropolitan area using Geomarket filters that exclude the City of Chicago– as observed in some orders collected via our public requests – result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
-Appendix \@fig-chicago-rq2-midSAT-race and @fig-chicago-rq2-highSAT-race show that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters.
+The [Online Appendix](https://ozanj.github.io/cb_geo/#/results) shows that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters.
 In both cases, the City of Chicago contributes disproportionately larger shares of Black and Hispanic prospects compared to all other Geomarkets.
@@
 We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
-Similar to race/ethnicity figures above, @fig-chicago-rq2-midSAT-ses presents the Geomarket distribution for the 13,749 total prospects (whose first-generations status is known) included in middle-range SAT score orders (1160-1300).
-Among the 1,743 prospects whose parents did not attend college ("no college"), the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 38% in comparison to 17% for all included prospects in this SAT range.
-In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students whose parents did not attend college relative to all included prospects, with such disparity ranging from a one-percentage-point difference (South and Southwest Suburbs) to a nine-percentage-point difference (Western Suburbs).
+Similar to race/ethnicity figures above, @fig-chicago-rq2-midSAT-ses presents the Geomarket distribution for the 13,848 total prospects (whose first-generations status is known) included in middle-range SAT score orders (1160-1300).
+Among the 1,750 prospects whose parents did not attend college ("no college"), the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 43% in comparison to 23% for all included prospects in this SAT range.
+In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students whose parents did not attend college relative to all included prospects, with such disparity ranging from a one-percentage-point difference (Northwest Suburbs, South and Southwest Suburbs) to a ten-percentage-point difference (Western Suburbs).
 For prospects whose parents attended some college but did not complete their degrees ("some college"), @fig-chicago-rq2-midSAT-ses indicates both City of Chicago and South and Southwest Suburbs Geomarkets contribute a disproportionately larger share of first-generation students.
 These representational patterns are reversed among prospects that are not first-generation college students.
-The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (12%) relative to all prospects (17%) in this SAT range, whereas all other Geomarkets contributed comparable shares.
+The City of Chicago Geomarket contributed a substantially smaller share of not-first generation prospects (18%) relative to all prospects (23%) in this SAT range, whereas all other Geomarkets contributed comparable shares.
 
@@
 Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
-@fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 12,981 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders.
+@fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 13,059 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders.
 @fig-chicago-rq2-midSAT-combo shows the City of Chicago Geomarket consistently contributes a disproportionately larger share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
-For instance, among all Asian prospects (n=1,341) whose contact information was purchased in this SAT score range, approximately 12% were first-generation college students whose parents did not attend college, 16% were first-generation college students whose parents attended some college but did not complete their degree, and 72% had parents with a BA.
-By contrast, for 237 Asian students from the City of Chicago Geomarket, 35% had parents who did not attend colllege and only 43% had parents with a BA.
-
-Similar Geomarket contribution patterns are evident among Hispanic prospects.
-The City of Chicago Geomarket stands out as the only Geomarket contributing substantially larger shares of Hispanic first-generation college students across those whose parents did not attend college (46% versus 38% overall) and those whose parents attended some college but did not complete their degrees (29% versus 26% overall).
+For instance, among all Asian prospects (n=1,370) whose contact information was purchased in this SAT score range, approximately 12% were first-generation college students whose parents did not attend college, 16% were first-generation college students whose parents attended some college but did not complete their degree, and 73% had parents with a BA.
+By contrast, for 292 Asian students from the City of Chicago Geomarket, 30% had parents who did not attend college and less than 50% had parents with a BA.
+
+For Hispanic prospects, the City of Chicago stands out as the Geomarket contributing the largest overall number and share of Hispanic first-generation college students across those whose parents did not attend college (44% versus 38% overall) and those whose parents attended some college but did not complete their degrees (29% versus 26% overall).
+Other Geomarkets such as Northwest Suburbs and Chain of lakes also contributed substantially larger shares.
 By contrast, Hispanic students from the more affluent North Shore Geomarket and Evanston & Skokie Geomarket were more likely to have parents with a BA.
 
-Although Black prospects who are not first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and Western Surburbs Geomarkets.
+Although Black prospects who are first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and South and Southwest Surburbs Geomarkets.
 Similarly, Geomarkets such as Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs contribute a disproportionately larger share of White prospects who are not first-generation college students, whereas City of Chicago and South and Southwest Suburbs contribute larger shares of White first-generation college students.
@@
 ```{r fig-chicago-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Chicago Area", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_race.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_chicago_race_row_plot.png')
+#knitr::include_graphics('./../assets/images/rq2_chicago_race.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-chicago-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_race.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_chicago_race_col_plot_392834_488035.png')
+
+#knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_race.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-chicago-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_ses.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_chicago_firstgen_col_plot_392834_488035.png')
+#knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_ses.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-chicago-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_combo.png')
+#
+knitr::include_graphics('./../results/graphs/rq2b/rq2b_chicago_order_392834_488035_race_by_firstgen.png')
+#knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_combo.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-07-08 text/manuscript-full-KS.qmd
revisions dallas rq1

@@
 The Dallas-Fort Worth region represents a unique "metroplex" case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic urban core and suburban ring model showcased by Chicago.
-As depicted in @fig-dallas-maps, two distinct inner-core Geomarkets (City of Dallas and City of Fort Worth) are seperated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
+As depicted in @fig-dallas-maps, two distinct inner-core Geomarkets (City of Dallas and City of Fort Worth) are separated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
 These central Geomarkets are surrounded by the more conventional outer ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
@@
 @fig-dallas-maps shows all six Geomarkets experienced population growth over the three census periods.
-However, the most pronounced increases occured in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 340,000 to 1.4 Million and 170,000 to 1.2 Million, respectively.
+However, the most pronounced increases occurred in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 350,000 to 1.6 Million and 160,000 to 1.1 Million, respectively.
 These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers -- City of Forth Worth and City of Dallas -- into the surrounding suburban and rural areas of the metropolitan areas.
@@
 @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
-In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to nearly 92% in West of Dallas/Fort Worth Metroplex.
+In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to 93% in West of Dallas/Fort Worth Metroplex.
 By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County, Irving, Arlington, and Grand Prairie, and City of Fort Wort.
-In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 80%, respectively.
-By 2000, the City of Dallas was no longer a predominantly White Geomarket, with a population that was approximately 24% Black, 30% Hispanic, and 3% Asian.
+In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 79%, respectively.
+By 2000, the City of Dallas was no longer a predominantly White Geomarket, with a population that was approximately 25% Black, 36% Hispanic, and 2% Asian.
 Over the next two decades, the Hispanic population grew steadily, contributing to greater diversity.
-By 2020, the City of Dallas Geomarket had the lowest proportions of White and Asian residents across the Dallas-Fort Worth Metropolitan area, with approximately 21% Black, 38% Hispanic, and 2% Asian residents.
+By 2020, the City of Dallas Geomarket had the lowest proportions of White and Asian residents across the Dallas-Fort Worth Metropolitan area, with approximately 23% Black and 42% Hispanic residents.
 
@@
 The Collin and Rockwall and West of Dallas/Fort Worth Metroplex Geomarkets experienced continued declines in their share of White residents.
-However, both Geomarkets remained predominantly White (58% and 63%, respectively).
-In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (33% White, 23% Black, 32% Hispanic; 9% Asian), Irving, Arlington, and Grand Prairie (30% White, 20% Black, 36% Hispanic; 12% Asian), and City of Fort Worth (46% White, 15% Black, 33% Hispanic; 4% Asian).
+However, both Geomarkets remained predominantly White (57% and 63%, respectively).
+In contrast, all other Geomarkets experienced continued increases in the shares of Black and Hispanic residents, resulting in a diminished previously White-majority populations for Dallas County (31% White, 25% Black, 34% Hispanic; 8% Asian), Irving, Arlington, and Grand Prairie (29% White, 19% Black, 38% Hispanic; 11% Asian), and City of Fort Worth (42% White, 16% Black, 35% Hispanic; 4% Asian).
 <!--#  Lastly, continuous increases in the shares of Black and Hispanic residents further diversified the City of Dallas, making it the most racially/ethnically diverse Geomarket in the Dallas-Fort Worth metropolitan area.
@@
 In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable through 2020.
-As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$86,000 and \$87,000 \[MISMATCH BETWEEN GRAPH AND TABLE?\] median household incomes, respectively.
-By contrast, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$133,000, followed by West of Dallas/Fort Worth Metroplex (\$116,000).
+As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$81,000 and \$86,000 median household incomes, respectively.
+By contrast, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$129,000, followed by West of Dallas/Fort Worth Metroplex (\$120,000).
 @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty and educational attainment within Geomarkets over time, revealing similar dynamics across Geomarkets to those observed in median household income.
@@
 ```{r fig-dallas-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Dallas Area Geomarkets, 1980-2020", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq1-dallas-race.png')
+knitr::include_graphics('./../results/graphs/rq1/rq1_dallas_race.png')
+#knitr::include_graphics('./../assets/images/rq1-dallas-race.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-dallas-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Dallas Area Geomarkets, 1980-2020", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq1-dallas-ses.png')
+knitr::include_graphics('./../results/graphs/rq1/rq1_dallas_ses.png')
+
+#knitr::include_graphics('./../assets/images/rq1-dallas-ses.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-07-08 text/manuscript-full-KS.qmd
revisions dallas rq2

@@
 Overall, median income both within and across Geomarkets in the Chicago metropolitan area remained relatively stable across the three census periods.
-For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$168,000 in the North Shore Geomarket to a low of \$68,000 in the City of Chicago Geomarket.
-However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence, such as \$131,000 in Evanston and Skokie, \$120,000 in the Northwest Suburbs, \$111,000 in the Western Suburbs, \$109,000 in Chain of Lakes, and \$99,000 in the South and Southwest Suburbs.
-While overall income patterns across Geomarkets remained consistent through 2020, a median income increase for the City of Chicago Geomarket (\$80,000) and decrease in the South and Southwest Suburbs Geomarket (\$91,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$11,000.
+For instance, in 1980, median household income across the Chicago metropolitan area ranged from a high of \$165,000 in the North Shore Geomarket to a low of \$68,000 in the City of Chicago Geomarket.
+However, median incomes in all other Geomarkets were closer to the upper thresholds of affluence, such as \$125,000 in Evanston and Skokie, \$117,000 in the Northwest Suburbs, \$108,000 in the Western Suburbs, \$104,000 in Chain of Lakes, and \$99,000 in the South and Southwest Suburbs.
+While overall income patterns across Geomarkets remained consistent through 2020, a slight median income increase for the City of Chicago Geomarket (\$71,000) and decrease in the South and Southwest Suburbs Geomarket (\$84,000) narrowed the income gap between the lowest and second-lowest Geomarkets to just \$13,000.
 @fig-chicago-rq1-ses also shows that from 1980 to 2020, the percentage of residents living in poverty declined in the City of Chicago Geomarket while it increased in all other Geomarkets, resulting in a narrowing of poverty gaps across the Chicago metropolitan area.
 Similar patterns are evident in educational attainment across Geomarkets.
-In 1980, the City of Chicago and South and Southwest Suburbabs had the lowest percentage of residents with a bachelor's degree or higher (14%).
+In 1980, the City of Chicago and South and Southwest Suburbs had the lowest percentage of residents with a bachelor's degree or higher (14%).
 However, by 2020, the City of Chicago experienced the largest percentage point increase in college degree attainment, reaching 34%.
@@
 In contrast, median household income for City of Dallas and City of Fort Worth remained relatively stable through 2020.
-As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$81,000 and \$86,000 median household incomes, respectively.
-By contrast, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$129,000, followed by West of Dallas/Fort Worth Metroplex (\$120,000).
+As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$68,000 and \$74,000 median household incomes, respectively.
+By contrast, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$124,000, followed by West of Dallas/Fort Worth Metroplex (\$104,000).
 @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty and educational attainment within Geomarkets over time, revealing similar dynamics across Geomarkets to those observed in median household income.
@@
 This distribution is most notable at the lowest PSAT score threshold.
-For instance, at the 1070-1180 PSAT score range, the central suburban Geomarkets of Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects.
+For instance, at the 1070-1180 PSAT score range, the central suburban Geomarkets of Dallas County (37% White, 17% Asian, 15% Black, 27% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 33% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects.
 While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportions of Hispanic prospects but lower proportions of Asian and Black students.
-The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest but predominantly White prospect pools.
+The Collin & Rockwall Counties (58% White, 7% Black, 17% Hispanic, 13% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (60% White, 7% Black, 18% Hispanic, 9% Asian) both yield the largest but predominantly White prospect pools.
 Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
@@
 @fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
-The bars for "All (race known)" show that Collin & Rockwall (37%) and West of Dallas/Forth Worth (23%) contribute many more students than the other four Geomarkets.
+The bars for "All (race known)" show that Collin & Rockwall (30%) and West of Dallas/Forth Worth (27%) contribute many more students than the other four Geomarkets.
 Disparities in Geomarket contributions were evident but differed for Asian, Black, and Hispanic students.
-For instance, The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+For instance, The Dallas County (19% versus 12% for all prospects) and Colin and Rockwall Counties (39% versus 30% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
 
@@
 Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-dallas-rq2-midSAT-race.
-All Geomarkets besides Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex contributed equal or larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Dallas-Fort Woth metropolitan area.
-However, the City of Dallas Geomarket contributed the largest disproportionate share of Hispanic prospects (19% versus 11% for all prospects), whereas the Dallas Counties Geomarket contributed the largest disproportionate share of Black prospects (21% versus 13% for all prospects).
-
-Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses [THE FIGURE IN THE PAPER IS WRONG].
-Only 303 prospects of the 3,928 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
-Again, all Geomarkets besides Collin and Rockwall Counties (14% versus 36% for all prospects) and West of Dallas/Fort Worth Metroplex (15% versus 24% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool.
-Disproportioantely larger contributions range from six percentage points by the Dallas Counties Geomarket (19% versus 13% for all prospects) to 11 percentage points by the City of Dallas Geomarket (22% versus 11% for all prospects).
+All Geomarkets besides Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex contributed nearly equal or larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Dallas-Fort Woth metropolitan area.
+However, the City of Dallas Geomarket contributed the largest disproportionate share of Hispanic prospects (19% versus 12% for all prospects), whereas the Irving, Arlington, and Grand Prairie Geomarket contributed the largest disproportionate share of Black prospects (15% versus 7% for all prospects).
+
+Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses.
+Only 303 prospects of the 3,926 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
+Again, all Geomarkets besides Collin and Rockwall Counties (12% versus 30% for all prospects) and West of Dallas/Fort Worth Metroplex (16% versus 27% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool.
+Disproportionately larger contributions range from five percentage points by the City of Fort Worth (16% versus 11% for all prospects) to 10 percentage points by the City of Dallas Geomarket (22% versus 12% for all prospects).
 Contributions become more proportional for the pool of first-generation college students whose parents attended but did not complete their degree.
-Proportional balances across Geomarkets are also evident for the 3,069 prospects that are not first-generation college students, with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportionate share of not first-generation college students (40%) relative to the contributions to the overall prospect pool (36%).
+Proportional balances across Geomarkets are also evident for the 3,067 prospects that are not first-generation college students.
+<!--#  , with the exception of Collin and Rockwall Counties Geomarket contributing the largest disproportionate share of not first-generation college students (33%) relative to the contributions to the overall prospect pool (30%).-->
 
 Lastly, @fig-dallas-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity, which we view as the most interesting results for the Dallas-Fort Worth metropolitan area.
-Among all White prospects (n=2,094) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
-@fig-dallas-rq2-midSAT-combo shows nearly all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool, the only modest exception being the Irving, Arlington, & Grand Prairie Geomarket.
+Among all White prospects (n=2,093) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
+@fig-dallas-rq2-midSAT-combo shows almost all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool, the only modest exception being the Irving, Arlington, & Grand Prairie Geomarket.
 
@@
 For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
-By contrast, the more affluent Collin & Rockwall sent more Asian prospects (285) than any other Geomarket, but about 87% of these prospects were not first-generation college students.
+By contrast, the more affluent Collin & Rockwall sent more Asian prospects (234) than any other Geomarket, but about 89% of these prospects were not first-generation college students.
 
@@
 ```{r fig-chicago-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
-#
+
 knitr::include_graphics('./../results/graphs/rq2b/rq2b_chicago_order_392834_488035_race_by_firstgen.png')
@@
 ```{r fig-dallas-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Dallas Area", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2-dallas-race.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_dallas_race_row_plot.png')
+#knitr::include_graphics('./../assets/images/rq2-dallas-race.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-dallas-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_dallas_midSAT_race.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_dallas_race_col_plot_448427.png')
+#knitr::include_graphics('./../assets/images/rq2_dallas_midSAT_race.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-dallas-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_ses.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_dallas_firstgen_col_plot_448427.png')
+
+#knitr::include_graphics('./../assets/images/rq2_chicago_midSAT_ses.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-dallas-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_dallas_midSAT_combo.png')
+
+knitr::include_graphics('./../results/graphs/rq2b/rq2b_dallas_order_448427_race_by_firstgen.png')
+#knitr::include_graphics('./../assets/images/rq2_dallas_midSAT_combo.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-07-09 text/manuscript-full-KS.qmd
revisions rq1 LA

@@
 The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
-The Los Angeles metropolitan area is one of the only multi-nuclei models in the country that is formed when urban growth occurs around multiple centers that serve different functions rather than expanding from a single central business district core (CITE- Harris and Ullman 1945).
+The Los Angeles metropolitan area is one of the only multi-nuclei models in the country that is formed when urban growth occurs around multiple centers that serve different functions rather than expanding from a single central business district core [@Harris1945].
 As depicted in @fig-la-maps, Geomarkets closely follow this overall metropolitan model.
@@
 The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernadino, and Ontario.
-The South and South Central Los Angeles Geomarket grew from approximately 1.2 to 1.6 Million from 1980 to 2020.
-In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets, respectively, grew from approximately 700,000 to 1.2 Million and from 1.1 Million to nearly 3 Million during this time.
+The South and South Central Los Angeles Geomarket grew from approximately 1.1 to 1.6 Million from 1980 to 2020.
+In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets, respectively, grew from approximately 580,000 to nearly 1 Million and from 1.1 Million to 3.3 Million during this time.
 
@@
 However, over the three decennial census periods, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
-In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 62% White in Long Beach to 78% in the San Fernando Valleys (West and East) Geomarkets.
-By 2020, however, only the West Los Angeles and West Beach Geomarket remained predominantly White (62%).
-Remanining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (44% White, 4% Black, 39% Hispanic; 11% Asian Pacific Islander), San Fernando Valley -East (43% White, 4% Black, 41% Hispanic; 8% Asian Pacific Islander), Glendale and Pasadena (35% White, 3% Black, 37% Hispanic; 21% Asian Pacific Islander), South Bay (29% White, 12% Black, 33% Hispanic; 21% Asian Pacific Islander), Long Beach (26% White, 11% Black, 46% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 3% Black, 52% Hispanic; 24% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 54% Hispanic; 6% Asian Pacific Islander).
-However, the South and South Central Los Angeles and Hollywood and Wilshire Geomarkets experienced substantial declines in the share of Black residents from 1980 to 2020, dropping from 44% to 18% and 18% to 9%, respectively.
+In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 63% White in Covina and West Covina to 86% in the San Fernando Valley-East Geomarket.
+By 2020, however, only the West Los Angeles and West Beach (51%) and San Fernando Valley-East (59%) Geomarkets remained predominantly White.
+Remanining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (40% White, 4% Black, 40% Hispanic; 13% Asian Pacific Islander), Glendale and Pasadena (35% White, 4% Black, 41% Hispanic; 17% Asian Pacific Islander), South Bay (32% White, 10% Black, 32% Hispanic; 22% Asian Pacific Islander), Long Beach (27% White, 10% Black, 47% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 4% Black, 57% Hispanic; 19% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 53% Hispanic; 8% Asian Pacific Islander).
+However, the South and South Central Los Angeles and Hollywood and Wilshire Geomarkets experienced substantial declines in the share of Black residents from 1980 to 2020, dropping from 44% to 16% and 22% to 11%, respectively.
 
@@
 @fig-la-rq1-ses presents trends in income, poverty, and educational attainment in the Los Angeles metropolitan area from 1980 to 2020, demonstrating the South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire Geomarkets remained at the lowest socioeconomic levels across these indicators.
-For instance, the median household income in the South and South Central Los Angeles Geomarket increased modestly from \$57,000 in 1980 to \$59,000 in 2000, reaching \$67,000 by 2020.
-In comparison, East Los Angeles and Hollywood and Wilshire recorded slightly higher median incomes in 2020 at \$76,000 and \$84,000, respectively.
-These figures remain substantially lower than that of the most affluent Geomarket, West Los Angeles and West Beach, where the median household income reached \$137,000 in 2020.
+For instance, the median household income in the South and South Central Los Angeles Geomarket increased modestly from \$51,000 in 1980 to \$57,000 in 2000, reaching \$61,000 by 2020.
+In comparison, East Los Angeles and Hollywood and Wilshire recorded slightly higher median incomes in 2020 at \$74,000 and \$70,000, respectively.
+These figures remain substantially lower than that of the most affluent Geomarket, West Los Angeles and West Beach, where the median household income reached \$117,000 in 2020.
 
@@
 
-```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
+```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarkets, Total Population 1980-2020", out.width = "100%"}
 knitr::include_graphics('./../assets/images/la-maps.png')
@@
 ```{r fig-la-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Los Angeles Area Geomarkets, 1980-2020", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq1_la_race.png')
+
+knitr::include_graphics('./../results/graphs/rq1/rq1_los_angeles_race.png')
+
+#knitr::include_graphics('./../assets/images/rq1_la_race.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-la-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Los Angeles Area Geomarkets, 1980-2020", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq1_la_ses.png')
+knitr::include_graphics('./../results/graphs/rq1/rq1_los_angeles_ses.png')
+
+#knitr::include_graphics('./../assets/images/rq1_la_ses.png') [OLD RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-07-09 text/manuscript-full-KS.qmd
revisions rq2 LA

@@
 @fig-la-rq2-race presents the racial/ethnic composition of purchased student profiles.
-For the more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180), South and South Central Los Angeles (n=1,031) and East Los Angeles (n=871) Geomarket pools stand out as having an overwhelming majority of students of color, 98% and 95% respectively.
-Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (89%) and East Los Angeles (64%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (6%) and larger share of Asian prospects for East Los Angeles (30%).
-Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic and Black student prospects, while East Los Angeles stands out for its substantial share of Asian prospects.
+For the nearly 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180), South and South Central Los Angeles (n=1,035) and East Los Angeles (n=1,076) Geomarket pools stand out as having an overwhelming majority of students of color, 98% and 95% respectively.
+Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (88%) and East Los Angeles (61%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (5%) and larger share of Asian prospects for East Los Angeles (31%).
+Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic, Asian, and Black student prospects.
+<!--#  while East Los Angeles stands out for its substantial share of Asian prospects.-->
 
@@
 @fig-la-rq2-midSAT-race shows how much each Geomarket accounts for the share of prospects in the middle PSAT score range by race/ethnicity.
-South and South Central Los Angeles accounts for only 2% of all purchased prospect profiles but accounts for 7% of Black prospects.
-The Reiverside, San Bernardino, & Ontario Geomarket accounts for 17% of all prospects but 25% of Black prospects.
-For Hispanic prospects, South and South Central Los Angeles (8% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (24% versus 17% for all prospects) also contributed a disproportinately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
-For Asian students, the Glendale and Pasadena (17% of Asian prospects versus 14% for all prospects), East Los Angeles (11% of Asian prospects versus 5% of all prospects) and Covina and West Covina (17% of Asain prospects versus 11% of all prospects) Geomarkets contributed a disproportionate larger share of Asian prospects relative to their contributions to the overall prospect pool.
-For the more than 2,500 White prospects in this pool, the San Fernando Valley -- West (28% of White prospects versus 20% of all prospects) and the West Los Angeles and West Beach (12% of White prospects versus 7% for all prospects) Geomarkets contribute disproportionately larger shares.
-
-These results demonstrate that following the @RN4982 strategy of filtering on high-income Geomarkets and excluding low-income Geomarkets -- such as the South and South Central Los Angeles Geomarket and the Riverside, San Bernadino, and Ontario Geomarket -- would result in the disproportionate exclusiong of Black and Hispanic students from the resulting prospect pools.Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.
-
-\[KARINA -- CHECK YOUR NUMBERS FOR THIS PARAGRAPH; KS: I DID; THEY ARE CORRECT FOR ORDERS 44820 AND 546946; AM I MISSING SOMETHING/WHY DO YOU THINK THEY ARE WRONG?\] @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
-For students whose parents did not attend college, East Los Angeles (16% versus 5% of all prospects), South and South Central Los Angeles (10% versus 2% of all prospects), and Riverside, San Bernardino, and Ontario (24% versus 17% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
+South and South Central Los Angeles accounts for only 3% of all purchased prospect profiles but accounts for 8% of Black prospects.
+The Riverside, San Bernardino, & Ontario Geomarket accounts for 24% of all prospects but 29% of Black prospects.
+For Hispanic prospects, South and South Central Los Angeles (9% versus 3% for all prospects) and Riverside, San Bernadino, and Ontario (30% versus 24% for all prospects) also contributed a disproportionately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
+For Asian students, the Glendale and Pasadena (20% of Asian prospects versus 15% for all prospects), East Los Angeles (14% of Asian prospects versus 7% of all prospects) and Covina and West Covina (12% of Asain prospects versus 9% of all prospects) Geomarkets contributed a disproportionate larger share of prospects relative to their contributions to the overall prospect pool.
+For the more than 2,300 White prospects in this pool, the San Fernando Valley - West (17% of White prospects versus 11% of all prospects), West Los Angeles and West Beach (9% of White prospects versus 6% for all prospects), and Glendale and Pasadena (17% of White prospects versus 15% for all prospects) Geomarkets contribute disproportionately larger shares.
+
+These results demonstrate that following the @RN4982 strategy of filtering on high-income Geomarkets and excluding low-income Geomarkets -- such as the South and South Central Los Angeles Geomarket and the Riverside, San Bernardino, and Ontario Geomarket -- would result in the disproportionate exclusion of Black and Hispanic students from the resulting prospect pools.
+Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.
+
+@fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
+For students whose parents did not attend college, East Los Angeles (18% versus 7% of all prospects), South and South Central Los Angeles (11% versus 3% of all prospects), and Riverside, San Bernardino, and Ontario (27% versus 24% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
 <!--#  KARINA -- I CUT THIS TEXT.
@@
 Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
-Across all Hispanic prospects in the middle range PSAT orders, 25% were first-generation college students whose parents did not attend college, 25% were first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students.
-By contrast, for the 156 Hispanic prospects from the South and South Central Geomarket, 54% had parents who never attended college, 28% had parents who attended some college, and only 18% had parents with a BA.
-Similarly, Hispanic prospects from East Los Angeles, from Hollywood & Wilshire, and from Riverside, San Bernardino, and Ontario were much more likely to have parents without a BA compared to all Hispanic prospects in the Los Angeles metro area.
+Across all Hispanic prospects in the middle range PSAT orders, 26% were first-generation college students whose parents did not attend college, 26% were first-generation college students whose parents attended some college but did not complete their degrees, and 49% were not first-generation college students.
+By contrast, for the 175 Hispanic prospects from the South and South Central Geomarket, 51% had parents who never attended college, 28% had parents who attended some college, and only 21% had parents with a BA.
+Similarly, Hispanic prospects from East Los Angeles and from Riverside, San Bernardino, and Ontario were more likely to have parents without a BA compared to all Hispanic prospects in the Los Angeles metro area.
 
@@
 
-Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 21% were first-generation college students whose parents attended some college but did not complete their degree, and 68% were not first-generation college students.
-@fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 251), a significant proportion of whom are first-generation college students (35%) whose parents did not attend college.
-The Hollywood & Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (35%).
-
-While the pool of Black prospects is relatively small (n=186) in @fig-la-rq2-midSAT-combo, some patterns emerge by first-generation status and Geomarket.
-Overall, approximately 5% of Black prospects were first-generation college students-no college, 20% were first-generation college students-some college, and 74% not first-generation).
-The South and South Central Los Angeles Geomarket stands out for contributing a disproportionately high share (21%) of Black first-generation college students whose parents did not attend college, despite representing a relatively smaller overall number.
-Additionally, the Long Beach Geomarket is notable for its higher proportion (40%) of Black first-generation students whose parents attended college but did not complete a degree.
+Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 22% were first-generation college students whose parents attended some college but did not complete their degree, and 67% were not first-generation college students.
+@fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 276), a significant proportion of whom are first-generation college students (34%) whose parents did not attend college.
+The Hollywood & Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (34%).
+
+While the pool of Black prospects is relatively small (n=182) in @fig-la-rq2-midSAT-combo, some patterns emerge by first-generation status and Geomarket.
+Overall, approximately 6% of Black prospects were first-generation college students-no college, 21% were first-generation college students-some college, and 74% not first-generation).
+The South and South Central Los Angeles Geomarket stands out for contributing a disproportionately high share (13%) of Black first-generation college students whose parents did not attend college, despite representing a relatively smaller overall number.
+Additionally, the Long Beach Geomarket is notable for its higher proportion (42%) of Black first-generation students whose parents attended college but did not complete a degree.
 
@@
 ```{r fig-la-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Los Angeles Area", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_la_race.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_los_angeles_race_row_plot.png')
+
+#knitr::include_graphics('./../assets/images/rq2_la_race.png') [RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-la-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_la_midSAT_race.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_los_angeles_race_col_plot_448420_546946.png')
+#knitr::include_graphics('./../assets/images/rq2_la_midSAT_race.png')[RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-la-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
-  knitr::include_graphics('./../assets/images/rq2_la_midSAT_ses.png')
+knitr::include_graphics('./../results/graphs/rq2/rq2_los_angeles_firstgen_col_plot_448420_546946.png')
+ # knitr::include_graphics('./../assets/images/rq2_la_midSAT_ses.png') [RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
@@
 ```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
-knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')
+knitr::include_graphics('./../results/graphs/rq2b/rq2b_los_angeles_order_448420_546946_race_by_firstgen.png')
+#knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')[RESULTS BASED ON ZIP GEOMARKETS]
 #knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')

### 2025-07-09 text/manuscript-full-KS.qmd
methods revision + spell check

@@
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
-The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metroplitan areas into smaller geographic units.
+The order specified three "search filters" to control which names were purchased: first, the 2020 high school graduating class; second, under the heading "College Board Exams," PSAT score range of 1070 - 1180; and third, under the heading "Geography", the order filtered on 67 "Geomarkets," identified by titles like "IL08 - Northwest Suburbs" and "IL09 - North Shore."[^1] The Geomarket filter carves states and metropolitan areas into smaller geographic units.
 Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and has a higher share of Black residents.
@@
 
-We argue that these perspectives ignore an important mechansim of social reproduction: third-party vendors create products that sort students on behalf of college.
+We argue that these perspectives ignore an important mechanism of social reproduction: third-party vendors create products that sort students on behalf of college.
 These products often amplify historic inequality in educational opportunity by using historical snapshots of student demand to make recommendations about where colleges should prioritize recruiting efforts.
@@
 Next, the authors develop "tinker toy" diagrams that show which institutions are connected to one another.
-These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
+These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each Geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
 For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
@@
 In itself, @RN4982's Market Segment Model is a social science depiction of student demand -- akin to the status attainment model -- that does not consider historic, structural inequalities that cause observed patterns.
-However, by inscribing the Market Segment Model and Geomarkets into the EPS software, College Board amplified structural inequalities that contributed to homophily observed in 1980 SAT score-sending beahivor.
+However, by inscribing the Market Segment Model and Geomarkets into the EPS software, College Board amplified structural inequalities that contributed to homophily observed in 1980 SAT score-sending behavior.
 Drawing from @RN4975, commodification of the Market Segment Model and Geomarkets engineers homophily.
@@
 
-[^16]: Data for the 2020 5-year ACS were retreived from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation.
+[^16]: Data for the 2020 5-year ACS were retrieved from the [tidycensus](https://walker-data.com/tidycensus/) R package, created by @tidycensus_citation.
 
@@
 However, this process was less precise when applied to the 1980 and 2000 Decennial Census tract boundaries, likely a factor of census tract delineations shifting over time based on demographic changes and our use of map images from 2023.
-The map on the right in @fig-geomarket-misalign deilneates 1980 census tract boundaries, which are visibly larger and more spatially aggregated than the smaller, more granular 2020 tracts shown in the center map.
+The map on the right in @fig-geomarket-misalign delineates 1980 census tract boundaries, which are visibly larger and more spatially aggregated than the smaller, more granular 2020 tracts shown in the center map.
 
@@
 This assignment, along with others like it, slightly alters the overall boundaries of Geomarkets when compared to those based on 2020 census tract delineations.
-To assess the robustness of our approach to spatial misalignments, we tested alternative methods for resolving ambiguous tract assignments to Geomarkets and found overall patterns in the results remained relatively consistent across approaches.[^17]
-
-[^17]: An alternative approach to dealing with spatial misalignments included creating shapefiles for Geomarkets from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.
-Essentially, the data file was a Geomarket-zipcode crosswalk that held one observation per zip code and a column that assigned each zip code to one Geomarket.
+To assess the robustness of our approach to spatial misalignment, we tested alternative methods for resolving ambiguous tract assignments to Geomarkets and found overall patterns in the results remained relatively consistent across approaches.[^17]
+
+[^17]: An alternative approach to dealing with spatial misalignment included creating shapefiles for Geomarkets from a 2012 R-bloggers post written by a data scientist working in the enrollment management space.
+Essentially, the data file was a Geomarket-zip code crosswalk that held one observation per zip code and a column that assigned each zip code to one Geomarket.
 Using these zip code-level data, we utilized the aggregate() function to create shapefiles for each Geomarket.
@@
 Even under this most conservative approach to addressing spatial misalignment, overall patterns in the findings remained consistent.
-However, we opted to georeference the manually digitized maps directly sourced from the College Board, as the original source of the Geomarket-zipcode crosswalk cited in the the R-bloggers post could not be verified.
+However, we opted to georeference the manually digitized maps directly sourced from the College Board, as the original source of the Geomarket-zip code crosswalk cited in the the R-bloggers post could not be verified.
 
@@
 We chose these states because they were priorities for the two foundations that funded the data collection.
-Utilizing public records requests to obtain quantitative data is a painstaking process, made more difficult by the onset of the Covid Pandemic.
+Utilizing public records requests to obtain quantitative data is a painstaking process, made more difficult by the onset of the COVID19 Pandemic.
 Initially, the majority of universities ignored or denied our request.
@@
 Our research questions explore the socioeconomic and racial variation between Geomarkets, how such variation changes over time, and how Geomarket landscapes shape the composition of included versus excluded prospects.
-Our goal is not to generalize Geomarket variation nationally, but rather to explore how differences across Geomarket landscapes shape prospect pools based on the geographical model of metropolitan areas and their demographic characteristics \[MAKE THIS MORE ABOUT EXPLORING THE MODEL.
-THIS SENTENCE SOUNDS FANCY BUT DOESN'T REALLY SAY ANYTHING\].
+Our goal is not to make generalizations about Geomarket variation nationally, but to explore how different geographical models of metropolitan areas, shaped by their unique demographic characteristics, influence the composition of prospect pools.
 Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts.
-We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics \[TIE IN THEORETICAL FRAMEWORK HERE A BIT MORE\].
-Online Appendix XXX presents results for a larger number of metropolitan areas.
-
-Case selection was informed by purposeful sampling to identify "information-rich" cases (Patton, 2002; Yin, 2009)[CITE] that can provide knowledge about variation across Geomarkets and over time.
+We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics.
+The [Online Appendix](https://ozanj.github.io/cb_geo/#/results) presents results for a larger number of metropolitan areas.
+
+The central thesis of Zemsky & Oedel's (1983) Market Segment Model suggests that student demand for various colleges is a function of social class and geography, and that colleges strategically target recruiting in high socioeconomic areas.
+However, because Geomarkets vary significantly in their racial and class characteristics across Geomarkets, and in how colleges target these Geomarkets, a multiple case study enables us to compare how these dynamics unfold across different metropolitan areas.
+By treating each metropolitan area as a separate case, we explore of both within-case processes (e.g., how homophily structures recruitment in a single metropolitan area) and cross-case patterns (e.g., how the same recruitment strategy results in different racial/class outcomes across metropolitan areas).
+
+Case selection was informed by purposeful sampling to identify "information-rich" cases that can provide knowledge about variation across Geomarkets and over time \[@RN4011; @RN4549\] .
 We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
-Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area[s] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
+Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
 11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia.
-To compare and contrast with classic urban model structures, we also identified alternative urban models such as bimodal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
+To compare and contrast with classic urban model structures, we also identified alternative urban models such as bi-modal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
 The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S.
@@
 Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
-In contrast, the most significant population increases occured in the Western Suburbs (1.0 Million to 1.4 Million) and the South and Southwest Suburbs (less than 1 Million to 1.2 Million) Geomarkets.
+In contrast, the most significant population increases occurred in the Western Suburbs (1.0 Million to 1.4 Million) and the South and Southwest Suburbs (less than 1 Million to 1.2 Million) Geomarkets.
 The City of Chicago Geomarket, forming the metropolitan area's central core, was the only locality to experience a population decline, from 3.5 Million in 1980 to 3.2 Million in 2020, despite remaining the most populous Geomarket.
@@
 
-Although Black prospects who are first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and South and Southwest Surburbs Geomarkets.
-Similarly, Geomarkets such as Chain of Lakes, North Shore, Evanston and Skokie, and Western Surburbs contribute a disproportionately larger share of White prospects who are not first-generation college students, whereas City of Chicago and South and Southwest Suburbs contribute larger shares of White first-generation college students.
+Although Black prospects who are first-generation college students make up a relatively smaller portion of the overall pool, @fig-chicago-rq2-midSAT-combo suggests that such prospects are primarily contributed by the Evanston and Skokie, City of Chicago, and South and Southwest Suburbs Geomarkets.
+Similarly, Geomarkets such as Chain of Lakes, North Shore, Evanston and Skokie, and Western Suburbs contribute a disproportionately larger share of White prospects who are not first-generation college students, whereas City of Chicago and South and Southwest Suburbs contribute larger shares of White first-generation college students.
 
@@
 
-@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area, revealing a comparatively more distributed pattern of racial/ethnic diversity across Geomarkets unlike Chicago's concentrated diversity within a single Geomarket.
+@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forth Worth area, revealing a comparatively more distributed pattern of racial/ethnic diversity across Geomarkets unlike Chicago's concentrated diversity within a single Geomarket.
 This distribution is most notable at the lowest PSAT score threshold.
 For instance, at the 1070-1180 PSAT score range, the central suburban Geomarkets of Dallas County (37% White, 17% Asian, 15% Black, 27% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 33% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects.
-While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportions of Hispanic prospects but lower proportions of Asian and Black students.
+While the urban central Geomarkets of City of Dallas and City of Fort Worth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportions of Hispanic prospects but lower proportions of Asian and Black students.
 The Collin & Rockwall Counties (58% White, 7% Black, 17% Hispanic, 13% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (60% White, 7% Black, 18% Hispanic, 9% Asian) both yield the largest but predominantly White prospect pools.
@@
 Representational patterns across Geomarkets are reversed for Black and Hispanic prospects in @fig-dallas-rq2-midSAT-race.
-All Geomarkets besides Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex contributed nearly equal or larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Dallas-Fort Woth metropolitan area.
+All Geomarkets besides Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex contributed nearly equal or larger proportions of Black and Hispanic prospects relative to their contributions to the overall prospect pool in the Dallas-Fort Worth metropolitan area.
 However, the City of Dallas Geomarket contributed the largest disproportionate share of Hispanic prospects (19% versus 12% for all prospects), whereas the Irving, Arlington, and Grand Prairie Geomarket contributed the largest disproportionate share of Black prospects (15% versus 7% for all prospects).
@@
 
-### Decentralized Diversity in a Multi-Nucleo Landscape
+### Decentralized Diversity in a Multi-Nuclei Landscape
 
@@
 Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
-The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernadino, and Ontario Geomarkets are located inland, comprising the Inland Empire.
+The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernardino, and Ontario Geomarkets are located inland, comprising the Inland Empire.
 
@@
 @fig-la-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model.
-The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernadino, and Ontario.
+The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernardino, and Ontario.
 The South and South Central Los Angeles Geomarket grew from approximately 1.1 to 1.6 Million from 1980 to 2020.
-In contrast, the San Fernando Valley-West and Riverside and San Bernadino, and Ontario Geomarkets, respectively, grew from approximately 580,000 to nearly 1 Million and from 1.1 Million to 3.3 Million during this time.
+In contrast, the San Fernando Valley-West and Riverside and San Bernardino, and Ontario Geomarkets, respectively, grew from approximately 580,000 to nearly 1 Million and from 1.1 Million to 3.3 Million during this time.
 
@@
 By 2020, however, only the West Los Angeles and West Beach (51%) and San Fernando Valley-East (59%) Geomarkets remained predominantly White.
-Remanining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (40% White, 4% Black, 40% Hispanic; 13% Asian Pacific Islander), Glendale and Pasadena (35% White, 4% Black, 41% Hispanic; 17% Asian Pacific Islander), South Bay (32% White, 10% Black, 32% Hispanic; 22% Asian Pacific Islander), Long Beach (27% White, 10% Black, 47% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 4% Black, 57% Hispanic; 19% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 53% Hispanic; 8% Asian Pacific Islander).
+Remaining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (40% White, 4% Black, 40% Hispanic; 13% Asian Pacific Islander), Glendale and Pasadena (35% White, 4% Black, 41% Hispanic; 17% Asian Pacific Islander), South Bay (32% White, 10% Black, 32% Hispanic; 22% Asian Pacific Islander), Long Beach (27% White, 10% Black, 47% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 4% Black, 57% Hispanic; 19% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 53% Hispanic; 8% Asian Pacific Islander).
 However, the South and South Central Los Angeles and Hollywood and Wilshire Geomarkets experienced substantial declines in the share of Black residents from 1980 to 2020, dropping from 44% to 16% and 22% to 11%, respectively.
@@
 For the nearly 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180), South and South Central Los Angeles (n=1,035) and East Los Angeles (n=1,076) Geomarket pools stand out as having an overwhelming majority of students of color, 98% and 95% respectively.
-Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (88%) and East Los Angeles (61%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (5%) and larger share of Asian prospects for East Los Angeles (31%).
+Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (88%) and East Los Angeles (61%) having substantially greater proportions of Hispanic prospects, as well as a larger proportion of Black prospects for South and South Central Los Angeles (5%) and larger share of Asian prospects for East Los Angeles (31%).
 Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic, Asian, and Black student prospects.
@@
 The Riverside, San Bernardino, & Ontario Geomarket accounts for 24% of all prospects but 29% of Black prospects.
-For Hispanic prospects, South and South Central Los Angeles (9% versus 3% for all prospects) and Riverside, San Bernadino, and Ontario (30% versus 24% for all prospects) also contributed a disproportionately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
-For Asian students, the Glendale and Pasadena (20% of Asian prospects versus 15% for all prospects), East Los Angeles (14% of Asian prospects versus 7% of all prospects) and Covina and West Covina (12% of Asain prospects versus 9% of all prospects) Geomarkets contributed a disproportionate larger share of prospects relative to their contributions to the overall prospect pool.
+For Hispanic prospects, South and South Central Los Angeles (9% versus 3% for all prospects) and Riverside, San Bernardino, and Ontario (30% versus 24% for all prospects) also contributed a disproportionately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
+For Asian students, the Glendale and Pasadena (20% of Asian prospects versus 15% for all prospects), East Los Angeles (14% of Asian prospects versus 7% of all prospects) and Covina and West Covina (12% of Asian prospects versus 9% of all prospects) Geomarkets contributed a disproportionate larger share of prospects relative to their contributions to the overall prospect pool.
 For the more than 2,300 White prospects in this pool, the San Fernando Valley - West (17% of White prospects versus 11% of all prospects), West Los Angeles and West Beach (9% of White prospects versus 6% for all prospects), and Glendale and Pasadena (17% of White prospects versus 15% for all prospects) Geomarkets contribute disproportionately larger shares.
@@
 
-@fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets.
+@fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernardino, and Ontario Geomarkets.
 For students whose parents did not attend college, East Los Angeles (18% versus 7% of all prospects), South and South Central Los Angeles (11% versus 3% of all prospects), and Riverside, San Bernardino, and Ontario (27% versus 24% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students.
@@
 The Market Segment Model recommends that colleges should identify their core student market segment (i.e., local, in-state, regional, or national), identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
-This information is contained in the Market Segment Profile (Appendix A), a standard ouput in EPS software.
+This information is contained in the Market Segment Profile (Appendix A), a standard output in EPS software.
 When considering new Geomarkets to target, EPS recommends colleges utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
@@
 We find that Geomarkets are highly correlated with race and class.
-Findings illustrate that in 1980, when Geomarkets were being created, Black and Hispanic residents tended to be highly concentrated in the poorest Geomarket in the Chicago, Dallas-Fortworth, and to some extent the Los Angeles metropolitan areas.
+Findings illustrate that in 1980, when Geomarkets were being created, Black and Hispanic residents tended to be highly concentrated in the poorest Geomarket in the Chicago, Dallas-Fort Worth, and to some extent the Los Angeles metropolitan areas.
 These Geomarket patterns extend to other metropolitan areas in the country, such as CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), and PA5 -- Philadelphia County.
 In 2020 Chicago, Black residents remain concentrated in the poorest Geomarket but Gentrification also pushed Black residents to the South and Southwest Suburbs Geomarket.
-In Dallas-Fortworth, Black and Hispanic communities are distributed across multiple Geomarkets; however, significant socioeconomic disparities persist between Geomarkets.
+In Dallas-Fort Worth, Black and Hispanic communities are distributed across multiple Geomarkets; however, significant socioeconomic disparities persist between Geomarkets.
 The Los Angeles metropolitan area presents another distinctive Geomarket landscape in which Black, Hispanic, and Asian communities across a range of socioeconomic levels are distributed across multiple Geomarkets rather than concentrated in one or two.
@@
 Meanwhile, college admissions officers are motivated to enroll students who can afford full tuition price and are likely to donate in the future.
-These mutually beneficial desires are consummated by recognzing that upper and upper-middle class applicants satisfy the extracurricular needs of the college.
+These mutually beneficial desires are consummated by recognizing that upper and upper-middle class applicants satisfy the extracurricular needs of the college.
 The orchestra needs oboists.
@@
 ACT responded to uncertainty in their core testing business by attempting to become an edtech company [@blumenstyk_2019], failing [@molnar_2020a], and being acquired by Nexus Capital in 2024 [@knox_2024].
-Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters [@RN4565] -- developed free college search engines, and entered the enrollment management consulting market more aggessively [@list_biz].
+Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters [@RN4565] -- developed free college search engines, and entered the enrollment management consulting market more aggressively [@list_biz].
 College Board has weathered the test-optional movement thanks to robust revenues from the Advanced Placement product, but the decline in PSAT/SAT/PreACT/ACT test-takers has undermined their oligopoly in the market for student list data.
@@
 In 2021, EAB and PowerSchool -- the leading provider of K12 student information systems and a subsidiary of Vista -- jointly acquired the EM consultancy/software vendor Hobsons.
-PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect and EAB paid PowerSchool to become the exlusive reseller of Intersect.
+PowerSchool acquired Hobson's Naviance product and its sister recruiting product Intersect and EAB paid PowerSchool to become the exclusive re-seller of Intersect.
 <!--Finally, in 2022, EAB acquired Concourse, the leading "direct admissions" platform.--> EAB integrated these acquisitions to create the [Enroll360](https://eab.com/solutions/enroll360/) recruiting platform.[^22]
@@
 For example, reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
-Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and wheather students used Naviance to "research competitor institutions" [@markup_naviance].
-A University of Utah procurment document states the Intersect subscription is necessary because "there is unique group of prospective students who are only in the PowerSchool Naviance platform" [@utah_procurement].
+Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and whether students used Naviance to "research competitor institutions" [@markup_naviance].
+A University of Utah procurement document states the Intersect subscription is necessary because "there is unique group of prospective students who are only in the PowerSchool Naviance platform" [@utah_procurement].
 
@@
 An under-studied and under-regulated third-party EM industry has negative consequences for colleges and for students.
-Accross multiple student list products, we see a growing number of filters that recommend which prospects to target based on the beheavior of previous cohorts of students.
+Across multiple student list products, we see a growing number of filters that recommend which prospects to target based on the behavior of previous cohorts of students.
 The ACT/Encoura "Enrollment Predictor" filter enables colleges to filter prospects based on the predicted probability of enrollment, which is created based on analyzing the enrollment decisions of previous cohorts of students [@schmidt_2019].
@@
 148] in the organizational field.
-Unfortunatel, scholarship salient to education policy and regulation have ignored two, related economic transformations.
+Unfortunately, scholarship salient to education policy and regulation have ignored two, related economic transformations.
 First, third-party vendors offer software platforms that perform core tasks of organizations, including schools and colleges.

### 2025-07-09 scripts/rq1_tables_graphs.R
update rq1 ses graphs

@@ -372,13 +372,28 @@ create_eps_graph <- function(eps_codes,
     
   } else {
     # SES approach
-    eps_table <- eps_table %>%
+    inc_table <- eps_table %>%
       filter(
         statistic == stat,
-        variable %in% c("Median income", "% in poverty", "% with BA+")
+        variable == "Median income"
       )
     
-    plot <- eps_table %>%
+    pov_table <- eps_table %>% 
+      filter(statistic == 'sum', variable %in% c('Poverty', 'Not poverty')) %>%
+      group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
+      filter(variable == 'Poverty') %>% 
+      mutate(variable = '% in poverty')
+    
+    edu_table <- eps_table %>%
+      filter(statistic == 'sum', variable %in% c('Less than HS', 'HS', 'Less than BA', 'BA+')) %>%
+      group_by(eps_codename, year, statistic) %>% mutate(value = value / sum(value) * 100) %>% ungroup() %>% 
+      filter(variable == 'BA+') %>% 
+      mutate(variable = '% with BA+')
+    
+    plot <- bind_rows(inc_table, pov_table, edu_table) %>% 
+      mutate(
+        variable = factor(variable, levels = c('Median income', '% in poverty', '% with BA+')),
+      ) %>%
       ggplot(aes(x = eps_codename, y = value)) +
       geom_col(fill = "grey50", position = position_dodge(width = 0.8), width = 0.7) +
       coord_flip() +

### 2025-07-09 text/manuscript-full-KS.qmd
revisions + copy edits

@@
 Geomarkets are geographic borders meant to define local recruiting markets.
-Figure X panel A shows Geomarkets in the Chicago-land area.
+The left panel of @fig-chicago-recruiting shows Geomarkets in the Chicago-land area.
 As a planning tool for colleges, the Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.
@@
 Anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories.
-For example, Figure X panel B shows that Chicago Geomarkets define recruiting territories for University of Chicago admissions recruiters.
+For example, @fig-chicago-recruiting shows that Chicago Geomarkets define recruiting territories for University of Chicago admissions recruiters.
 Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS).
@@
 <!-- @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
---> Colleges control which prospect profiles they purchase by selecting search filters, such as SAT score, AP score, state, etc.
+--> Colleges control which prospect profiles they purchase by selecting search filters, such as SAT score, AP score, and state.
 Geomarket filters enable colleges to include or exclude prospects from particular Geomarkets.
@@
 
-Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS software (RQ1) and within the Student Search Service student list product (RQ2).
+Our analyses address two research questions, which speak to how Geomarkets are utilized within EPS software (RQ1) and within the Student Search Service student list product (RQ2).
 First, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas and how does this variation change over time?
@@
 Second, how does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
-This research question is motivated by @RN4982, which recommend that selective colleges target affluent Geomarkets, and by prior research on enrollment management behavior, which finds that selective colleges focus recruiting visits on high schools in affluent, predominantly white communities [@RN3519; @RN4758].
+This research question is motivated by @RN4982, which recommend that selective colleges target affluent Geomarkets, and by prior research on enrollment management behavior, which finds that selective colleges focus recruiting visits on high schools in affluent, predominantly White communities [@RN3519; @RN4758].
 We analyze this research question using data from actual student list purchases, which we collected using public records requests.
@@
 Fourth, we present results.
-[We find that\ldots ONE OR TWO SENTENCES].
+Findings for RQ1 reveal that Geomarkets mirror deep-rooted patterns of racial and class segregation, with Black and Latinx communities historically and often still concentrated in the poorest Geomarkets, though these dynamics vary by metropolitan area.
+Findings for RQ2 indicate student list purchases that exclude working- and middle-class Geomarkets are likely to systematically overlook high-achieving Black and Hispanic prospective students, especially those who are first-generation college students, thereby highlighting how third-party recruitment tools serve as mechanisms of social reproduction by sorting students in ways that reinforce historic inequalities in educational opportunity.
+
 Finally, we discuss implications for scholarship.
 Third-party vendors have structured college access for decades and their influence is growing.
-The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [vista_eab_ml].
+The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [@vista_eab_ml].
 We propose two fruitful approaches to studying the role of third-party vendors in enrollment management.
@@
 Both products are applications of quantification that sort students into colleges.
-Student Search Service includes or excludes prospects from a student list purchases by colleges based on the value of selected filters like Geomarket while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
+Student Search Service includes or excludes prospects from student list purchases by colleges based on the value of selected filters like Geomarket while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households.
 The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on the inequitable sorting of students into colleges.
@@
 @RN5007 [p.
-402] define quantification as "the production and communication of numbers." <!-- The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers).--> The effects of quantification are described by the concepts reactivity, discipline, and authority [@RN5007].
-Sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, personal life and education, with a particular emphasis on rankings [@RN5008; @RN5003; @RN5005].
+402] define quantification as "the production and communication of numbers." <!-- The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers).
+The effects of quantification are described by the concepts reactivity, discipline, and authority [@RN5007].
+--> Sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, personal life, and education, with a particular emphasis on rankings [@RN5008; @RN5003; @RN5005].
 In the field of critical data studies, quantification is central to scholarship that explores how computation and algorithms structure social opportunities [e.g., @RN4772; @RN4849].
@@
 This logic is subsequently programmed into EPS software that recommends which Geomarkets and high schools to recruit from.
-Thus, homophily observed on the demand side -- which is a function of historic structural inequality –- is programmed into the supply-side.
+Thus, homophily observed on the demand side of college choice-- which is a function of historic structural inequality –- is programmed into the supply-side.
 
@@
 However, sociological analyses of EM focus on the behavior of colleges and their agents, creating the sense that inequalities created by EM are a function of individual colleges reacting to macro structures in their environment such as rankings and financial pressures [e.g., @RN5003; @RN3519].
-To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of producers.
+To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of colleges.
 Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027].
@@
 This manuscript analyzes the Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
-The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target.
+The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advises colleges which Geomarkets to target.
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
@@
 The primary data source for our second research question is student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
-This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Texas, and Arizona.
+This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 public universities located in California, Illinois, Texas, and Arizona.
 Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.
@@
 The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major).
-College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) \[CITE\].
+College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want [@RN0010].
 From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S.
-Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].[^20] Information about student socioeconomic status is limited.
-However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].
+Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic [@RN0010].[^20] Information about student socioeconomic status is limited.
+However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire.
 
@@
 
-Case selection was informed by purposeful sampling to identify "information-rich" cases that can provide knowledge about variation across Geomarkets and over time \[@RN4011; @RN4549\] .
+Case selection was informed by purposeful sampling to identify "information-rich" cases that can provide knowledge about variation across Geomarkets and over time [@RN4011; @RN4549] .
 We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
@@
 
+```{r fig-chicago-recruiting, echo = FALSE, fig.align = 'center', fig.cap = "The College Board Enrollment Planning Service Chicago Geomarkets (left) and University of Chicago Recruiting Territories (right)", out.width = "85%"}
+knitr::include_graphics('./../assets/images/chicago_recruiting.png')
+#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
+#![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
 ```{r fig-em-funnel, echo = FALSE, fig.align = 'center', fig.cap = "The Enrollment Funnel", out.width = "45%"}
@@
 \clearpage
+
+\newpage
+
+# Appendix
+
+```{r market-segment-characteristics, echo = F}
+#| label: tbl-market-segment-characteristics
+#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'
+
+market_segment <- data.frame(
+  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
+  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
+  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
+  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
+  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
+)
+
+kable(
+  x = market_segment,
+  col.names = c('Characteristic','Local','In-state','Regional','National'),
+  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
+)
+```
+
+```{r sample-institutional-profile, echo = F}
+#| label: tbl-sample-institutional-profile
+#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'
+
+sample_inst <- data.frame(
+  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
+  local = c(1,0.2,0,0.0,0,0.0),
+  in_state = c(58,4.8,25,8.4,2,3.0),
+  regional= c(69,4.1,35,7.0,5,5.8),
+  national = c(109,2.9,61,5.3,3,0.8),
+  total = c(237,3.3,121,5.9,10,1.8)
+)
+
+kable(
+  x = sample_inst,
+  col.names = c('','Local','In-state','Regional','National','Total'),
+  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
+)
+
+```

### 2025-07-10 text/manuscript-full-KS.qmd
copy edits

@@
 On January 9, 2019 a public research university purchased College Board *Student Search Service* order no.
-$448922$, thereby obtaining the contact information of out-of-state prospective students, who would subsequently receive brochures, postcards, email, and targeted social media.
+$448922$, thereby obtaining the contact information of out-of-state prospective students, who would subsequently receive brochures, postcards, emails, and targeted social media.
 This was a large student list purchase, yielding $122,426$ "names." At $\$0.45$ per name, the order cost $\$55,091.70$.
@@
 x].
-Conceptualized as an effort to to "capture and quantify" (p.
+Conceptualized as an effort to "capture and quantify" (p.
 11) the knowledge of admissions officers, @RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of high school seniors in 1980.
@@
 Scholarship on quantification demonstrates that making recommendations based on past correlations amplifies the effects of historic structural inequality [@RN4794].
-The snapshot of student demand in 1980 -- itself a consequence historic structural inequality -- was programmed into recruiting products that colleges utilize to identify and target prospective students.
+The snapshot of student demand in 1980 -- itself a consequence of historic structural inequality -- was programmed into recruiting products that colleges utilize to identify and target prospective students.
 The result is a supply-side that amplifies structural inequalities observed on the demand-side.
@@
 Third-party vendors have structured college access for decades and their influence is growing.
-The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [@vista_eab_ml].
+The contemporary enrollment management industry is characterized by private-equity owned firms selling software-as-a-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [@vista_eab_ml].
 We propose two fruitful approaches to studying the role of third-party vendors in enrollment management.
@@
 In turn, recruiting visits to high schools -- and subsequent phone calls -- are a means of maintaining ties with guidance counselors at feeder schools and establishing relationships with prospective students [@RN3519; @RN4895].
-Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly white public schools [@RN4758; @RN4759; @RN5023].
+Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly White public schools [@RN4758; @RN4759; @RN5023].
 From the student perspective, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
@@
 The most widely-known class of EM vendor is EM consulting firms, which provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success [@RN5027].
-As "creating a class" has becomes complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
+As "creating a class" has become complicated and high-stakes [@RN3519], many colleges hire EM consulting firms to develop and/or implement recruiting campaigns.
 The two largest firms -- Ruffalo Noel Levitz and EAB -- claim to serve more than 3,000 colleges and universities collectively [@rnlabout; @eabcomm].[^2]
@@
 
-While the contemporary EM industry is characterized by software-as-service products sold to colleges by private equity backed firms [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
+While the contemporary EM industry is characterized by software-as-a-service products sold to colleges by private equity backed firms [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
 Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit.
 The College Board Student Search Service -- copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions.
-The College Board Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
+The College Board Enrollment Planning Service (EPS), launched in 1984, is an early software-as-a-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets.
 Both products are applications of quantification that sort students into colleges.
@@
 Second, apply these results (e.g., regression coefficients) to future cases in order to make predictions and/or to assign levels of risk to each case.
-@RN4975 provides the example of @RN4989, who develop a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
+@RN4975 provides the example of @RN4989, who developed a method to predict sensitive personal attributes (e.g., gender, political party) based on Facebook Likes.
 These models predict outcomes based on correlations without requiring knowledge about underlying causal relationships.
@@
 50] writes that, "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"
+As Wired editor Chris Anderson infamously declared [\ldots], big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"
 
@@
 **Homophily**.
-Homophily is the idea actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together" [@mcpherson2001birds].
+Homophily is the idea that actors who share common characteristics are likely to form connections with one another, or "birds of a feather flock together" [@mcpherson2001birds].
 Homophily is a core concept of network science, in which actors (nodes) are connected to one another directly and indirectly via network ties (edges).
-In *The company We Keep: Colleges and Their Competition*, @RN4982 [chapter 4] conduct a network analysis whereby two colleges are defined as competitors if a large number of students submit SAT scores to both colleges.
+In *The Company We Keep: Colleges and Their Competition*, @RN4982 [chapter 4] conduct a network analysis whereby two colleges are defined as competitors if a large number of students submit SAT scores to both colleges.
 @RN4975 observes that network science models often draw from rational choice theory, which leads to the assumption that homophily is the result of voluntary action by individuals.
@@
 We observe similar phenomena in our case.
-Based on a network analysis of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin).
-In turn, @RN4982 reasons that colleges should target the Geomarkets and high schools targeted by peer colleges.
+Based on a network analysis of 1980 SAT score-sending, @RN4982 conclude that like-colleges compete for like-students (i.e., students of similar social origin).
+In turn, @RN4982 reason that colleges should target the Geomarkets and high schools targeted by peer colleges.
 This logic is subsequently programmed into EPS software that recommends which Geomarkets and high schools to recruit from.
-Thus, homophily observed on the demand side of college choice-- which is a function of historic structural inequality –- is programmed into the supply-side.
+Thus, homophily observed on the demand side of college choice -- which is a function of historic structural inequality -- is programmed into the supply-side.
 
@@
 @RN4810 introduce the concept "classification situations" to describe the expansion of actuarial techniques to categorize customers into many, ordinally ranked groups.
-Merchants and lenders began tying these classifications to tiered products that targets different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
+Merchants and lenders began tying these classifications to tiered products that target different consumer groups with different levels of benefits and costs [@RN5015].[^4] Classification situations engender markets where a vertical hierarchy of products are matched to a vertical hierarchy of consumers.
 We observe similar processes in our case study.
-@RN4982 categorizes students into four market segments -- local, in-state, regional, and national -- based on their score sending behavior and then evaluates the attractiveness of each Geomarket based on how many "regional" and "national" students live there.
+@RN4982 categorize students into four market segments -- local, in-state, regional, and national -- based on their score-sending behavior and then evaluate the attractiveness of each Geomarket based on how many "regional" and "national" students live there.
 <!--
@@
 Based on a snapshot of existing social stratification, market research matches vertically categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification.
-Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociology.
+Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-a-service platforms, but have remained invisible to sociology.
 
@@
 This manuscript analyzes the Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments.
-The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advises colleges which Geomarkets to target.
+The Market Segment Model and Geomarkets became the basis for the College Board Enrollment Planning Service (EPS), which advises colleges which Geomarkets to target.
 Later, Geomarkets were incorporated into the College Board student list product named Student Search Service.
@@
 @RN4982 [Chapter 3], *A Sense of Place: Students, Families, and Communities*, investigates the student characteristics correlated with being in the local, in-state, regional, or national market segment.
-The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score sending behavior, both individually and in combination.
+The analyses identify four variables -- educational aspirations, parental education, scholastic aptitude, and family income -- that predict score-sending behavior, both individually and in combination.
 Because these four variables "reflect the basic social patterns of the nation, it would have been surprising if these were not the four social variables that best explained the patterns of college choice" [@RN4982, p.
@@
 In 1984, College Board created the Enrollment Planning Service (EPS), based on the Market Segment Model [@RN4987; @RN4980].
-EPS was an early software-as-service platform that recreated the analyses of @RN4982.
+EPS was an early software-as-a-service platform that recreated the analyses of @RN4982.
 For each Geomarket, colleges could obtain Market Segment Report for each local market and the Institutional Profile -- their own and that of competitors.
@@
 Geomarkets were subsequently incorporated into the Student Search Service product.
-We show that that excluding low-income, non-white Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-white students with strong academic achievement.
+We show that that excluding low-income, non-White Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-White students with strong academic achievement.
 
@@
 College access scholarship from economics often analyzes the effects of district, state, and federal policies [e.g., @RN5043].
-These literatures ignore the role of third-party vendors, which are often owned by private equity and offer software-as-service platforms that perform core functions for schools and colleges.[^21]
+These literatures ignore the role of third-party vendors, which are often owned by private equity and offer software-as-a-service platforms that perform core functions for schools and colleges.[^21]
 
@@
 EAB exemplifies the trend towards private equity ownership and industry concentration.
-EAB is the preeminent EM consultancy, offering software-as-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
+EAB is the preeminent EM consultancy, offering software-as-a-service products along the domains of recruitment, pricing and financial aid, student success, and advancement [@about_eab].
 The origins of EAB trace to Bill Royall, who founded Royall & Company in 1983 to provide direct marketing and fundraising for Republican political campaigns [@jump_2020].
@@
 The goal was to generate revenue from the sale of names.
-By contrast, Enroll360 wraps several proprietary databases of names (e.g., Cappex, Intersect) into a software-as-service platform that recruits those prospects at different stages of the enrollment funnel.
+By contrast, Enroll360 wraps several proprietary databases of names (e.g., Cappex, Intersect) into a software-as-a-service platform that recruits those prospects at different stages of the enrollment funnel.
 Here, the goal of lists is to create demand for software.
 Colleges that are interested in recruiting these prospects must buy an Enroll360 subscription.
-Like College Board's *Student Search Service,* contemporary software-as-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
+Like College Board's *Student Search Service,* contemporary software-as-a-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
 For example, reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.

### 2025-07-10 text/manuscript-full-KS.qmd
abstract revisions

@@
 
-*The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Model." Geomarkets carve states and metropolitan areas into smaller geographic units, meant to define local recruiting markets.
-The Market Segment Model predicts how student demand for a particular college varies by Geomarket, based on the socioeconomic characteristics of households.
-Geomarkets became an input for two College Board products that help colleges recruit students.
-First, the Enrollment Planning Service (EPS) software recommends specific Geomarkets and high schools from which colleges should recruit.
-Second, the Student Search Service sells the contact information of prospective students -- referred to as "student lists" -- and colleges can filter by Geomarket to determine which prospect profiles they purchase.
-We draw from scholarship on quantification, particularly the discussions of correlation and homophily by @RN4975, to conceptualize how recruiting products incorporate Geomarkets.
-We address two research questions: What is the socioeconomic and racial variation between Geomarkets and how does this variation change over time?
-How does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
+*The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Mode" (MSM).
+Geomarkets carve states and metropolitan areas into smaller geographic units to define local recruiting markets.
+MSM predicts how student demand for a college varies by Geomarket, based on household socioeconomic characteristics.
+Geomarkets became an input for two College Board recruiting products.
+First, the Enrollment Planning Service informs which Geomarkets and high schools a college recruits from.
+Second, the Student Search Service sells prospective student contact information -- or "student lists" -- which colleges filter by Geomarket to determine which prospect profiles to purchase.
+Drawing from scholarship on quantification, particularly discussions of correlation and homophily, we conceptualize how recruiting products incorporate Geomarkets.
+We address two research questions: What is the socioeconomic and racial variation between Geomarkets?
+How does the socioeconomic and racial composition of included versus excluded prospects vary when student lists filter on Geomarkets?
 We answer RQ1 by analyzing Census data from 1980, 2000, and 2020.
-We answer RQ2 using data on student lists purchased by public universities, which we collected by issuing public records requests.
-We utilize a quantitative case study design.
-Metropolitan areas are cases.
-Analyses consist of descriptive statistics and interactive maps.
+RQ2 analyzes data on student lists purchased by public universities, collected through public records requests.
+Using a quantitative case study design with metropolitan areas as cases, descriptive statistics and interactive maps suggest Geomarkets mirror deeply-rooted patterns of racial and class segregation.
+Moreover, student lists that exclude low-income Geomarkets systematically overlook non-white students, first-generation students, and especially first-generation non-white students.
 
@@
 Findings for RQ1 reveal that Geomarkets mirror deep-rooted patterns of racial and class segregation, with Black and Latinx communities historically and often still concentrated in the poorest Geomarkets, though these dynamics vary by metropolitan area.
-Findings for RQ2 indicate student list purchases that exclude working- and middle-class Geomarkets are likely to systematically overlook high-achieving Black and Hispanic prospective students, especially those who are first-generation college students, thereby highlighting how third-party recruitment tools serve as mechanisms of social reproduction by sorting students in ways that reinforce historic inequalities in educational opportunity.
+Findings for RQ2 indicate student list purchases that exclude low-income Geomarkets are likely to systematically overlook high-achieving Black, Hispanic, and Asian prospective students, especially those who are first-generation college students, thereby highlighting how third-party recruitment tools serve as mechanisms of social reproduction by sorting students in ways that reinforce historic inequalities in educational opportunity.
 
@@
 50] writes that, "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared [\ldots], big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"
+As Wired editor Chris Anderson infamously declared \[\ldots\], big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"
 

### 2025-07-11 text/manuscript-full-KS.qmd
copy edits

@@
 
-*The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Mode" (MSM).
+*The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Model" (MSM).
 Geomarkets carve states and metropolitan areas into smaller geographic units to define local recruiting markets.
@@
 Using a quantitative case study design with metropolitan areas as cases, descriptive statistics and interactive maps suggest Geomarkets mirror deeply-rooted patterns of racial and class segregation.
-Moreover, student lists that exclude low-income Geomarkets systematically overlook non-white students, first-generation students, and especially first-generation non-white students.
+Moreover, student lists that exclude low-income Geomarkets systematically overlook non-White students, first-generation students, and especially first-generation non-White students.
 
@@
 
-On January 9, 2019 a public research university purchased College Board *Student Search Service* order no.
+On January 9, 2019 a public research university purchased College Board Student Search Service order no.
 $448922$, thereby obtaining the contact information of out-of-state prospective students, who would subsequently receive brochures, postcards, emails, and targeted social media.
@@
 
-In 1978, set against the backdrop of impending college-age demographic decline, University of Pennsylvania professor Robert Zemsky, was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?" [@RN4982, p.
+In 1978, set against the backdrop of impending college-age demographic decline, University of Pennsylvania professor Robert Zemsky was tasked by the President to figure out, "'Who thinks about Penn?'" and "'What other institutions do they think about when they think about us?'" [@RN4982, p.
 x].
@@
 Each major metropolitan area is composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs.
-In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful."
+In more sparsely populated areas, communities are sometimes combined in order to make the analysis meaningful.
 
@@
 The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
-Appendix A describes these outputs in more detail \[CRYSTAL\].
+Appendix A describes these outputs in more detail.
 Both outputs are created separately for each Geomarket.
-For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family in come greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
+For each market segment (local, in-state, regional, national), the Market Segment Profile shows the number of students, their average SAT score, the percent aspiring to a BA+, percent with family income greater than \$35,000, percent with both parents having a BA, etc.[^5] The Institutional Profile shows the number of students who send scores to a particular institution -- separately by market segment -- and which majors these students are interested in.
 A university could obtain this for their own institution or for a competitor.[^6] Taken together, the Market Segment Profile shows a college which Geomarkets possess attractive student market segments while the Institutional Profile shows the extent to which students in a particular Geomarket are interested in your college or a competitor college.
@@
 These diagrams are drawn separately for each student segment -- local, in-state, regional, and national -- and separately for each Geomarket, such that the analyses convey which institutions compete with one another for which student segments in each local market.
-For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," [@RN4982, p.
+For example, describing the Figure 4.4 "Structure of Fairfield County Regional Market," @RN4982 [p.
 54] state that "competitive overlap, moreover, is often confined to institutions belonging to the same \[Carnegie\] type as well sector.
@@
 @RN4988 shows the contemporary Geomarkets.
-Documentation and promotional material suggests that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^8] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
+Documentation and promotional material suggest that geomarket borders were chosen based on a combination of formal geographic borders (e.g., counties) as well as proprietary College Board data designed to identify geographic areas with different college-going behaviors.[^8] However, Geomarkets for New England, Middle States, and the South are identical to those developed by @RN4982.
 
@@
 Drawing from promotional literature [e.g., @RN5019; @RN1621], @RN4980, and background conversations with enrollment managers, the practical purpose of EPS was to inform the "travel schedule" of admissions recruiters.
-EPS promotional material and user guides encourages users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
-Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college educated households, while low-income communities are left to local four-year and community colleges.
+EPS promotional material and user guides encourage users to begin by identifying which Geomarkets they will recruit from.[^9] Second, EPS users decide which high schools they will visit within selected Geomarkets.
+Based on the principle of homophily, the Market Segment Model suggests that selective colleges should target Geomarkets with large numbers of affluent, college-educated households, while low-income communities are left to local four-year and community colleges.
 Anecdotally, the salience of Geomarkets and EPS software affected the decision-making authority of admissions officers and the territories that admissions officers were assigned to.[^10] <!-- uc san diego appears to use geomarkets to organize high school visits: inspect the html in this page https://beatriton.ucsd.edu/register/requestschoolvisit -->
@@
 In turn, as Geomarkets became more salient, admissions offices often structured their territories around Geomarkets.
-On background, admissions officers and enrollment consultants told us it remains common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket.
-In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
+On background, admissions officers and enrollment consultants told us it remains common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket." In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
 This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX CRYSTAL\]
@@
 @RN4982 viewed demand for higher education as a function of class.
-They developed 142 Geomarkets borders with an eye towards identifying geographic areas that differed from one another in terms of class composition, while [@RN5016] created the remaining Geomarket borders with an eye towards distinguishing geographies based on college-going behaviors.
+They developed 143 Geomarkets borders with an eye towards identifying geographic areas that differed from one another in terms of class composition, while [@RN5016] created the remaining Geomarket borders with an eye towards distinguishing geographies based on college-going behaviors.
 Considering the extent of class-based residential segregation in the U.S.
@@
 The authors conceptualize student list products as "selection devices" [@RN4778] that enable colleges to select which prospective students they target by incorporating search filters (e.g., high school graduating class, state).
-@RN4786 defines "racialized inputs" as ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input [@RN4786].
+@RN4786 defines "racialized inputs" as ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input.
 @RN5012 argue that several frequently utilized student list filters (e.g., zip code, AP test score, SAT score) meet the criteria of racialized inputs.
@@
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the *Student Search Service*, such as Geomarkets and Geodemographic Segment filters.[^14] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
+College Board also created geographic borders that subsequently became filters in the Student Search Service, such as Geomarkets and Geodemographic Segment filters.[^14] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 Second, Geomarkets may be systematically correlated with race.
-For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda Country, excluding Oakland." <!--# A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.--> We extend @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
+For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda County, excluding Oakland." <!--# A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.--> We extend @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
 <!--# This manuscript focuses on Geomarkets.
@@
 
-[^14]: Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based past college enrollment and other factors [@RN4565].
+[^14]: Geodemographic segment filters utilize cluster analysis to allocate each census tract and each high school into different categories based on past college enrollment and other factors [@RN4565].
 
@@
 The Market Segment Model would provide this information through segment percentages for the community in question.
-Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen \ldots Your efforts would surely be better directed toward \ldots Manchester, Hartford, and Fairfield County \ldots Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population.
+Further classification of students by social attributes allows you to identify a group for mailings or recruiting\ldots If you were to recruit in Boston, only about two out of every ten students with fewer than two attributes would likely listen\ldots Your efforts would surely be better directed toward\ldots Manchester, Hartford, and Fairfield County\ldots Indeed, in Fairfield County alone you could reach more than 40 percent of your "primary target" population.
 <!-- -- that is, students with a greater than 75-percent probability of concentrating their college choices among institutions like your own.
@@
 
-# Appendix
+# Appendix A
 

### 2025-07-13 text/manuscript-full-KS.qmd
copy edits

@@
 Upon closer inspection, this student list purchase selected all Geomarkets in the Chicago-land area except for "IL10 - City of Chicago," which is significantly poorer and has a higher share of Black residents.
-Similarly, in the DC/Maryland/Virginia metro area, the order selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington & Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets.
+Similarly, in the DC/Maryland/Virginia metro area, the order selected the affluent "MD02 - Montgomery Metropolitan," "MD03 - Central Maryland (w/o Baltimore)," "VA01 - Arlington and Alexandria," and "VA02 - Fairfax County" Geomarkets, but excluded the "MD05 - Prince Georges Metropolitan," "MD07 - Baltimore (Urban)," and "DC01 - District of Columbia" Geomarkets.
 
@@
 50] writes that, "correlation grounds big data's so-called revolutionary potential.
-As Wired editor Chris Anderson infamously declared \[\ldots\], big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"
+As Wired editor Chris Anderson infamously declared\ldots, big data proved that 'correlation supersedes causation, and science can advance even without coherent models, unified theories, or really any mechanistic explanation at all.'"
 
@@
 Negative consequences of homophily
-Chun (2021) discusses the negative consequences arising from homophily.Second, when we “engineer homophily” – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
+Chun (2021) discusses the negative consequences arising from homophily.Second, when we "engineer homophily" – through algorithms or policies – we make it less likely for heterophily to emerge, eliminating possibilities – of friendships, of matches – that may have occurred in the absence of such engineering.
 
@@
 A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the “enrollment funnel.” @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the "enrollment funnel." @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
 
@@
 Census tracts are the smallest geographic area for which measures of both race/ethnicity and socioeconomic characteristics are available.
-Defined as statistical subdivisions of counties, census tracts boundaries are reassessed every decennial census to reflect relatively homogeneous demographic communities, typically ranging in population size from 2,500 and 8,000 residents [@Bureau2014; @Bureau1994].
+Defined as statistical subdivisions of counties, census tracts boundaries are reassessed every decennial census to reflect relatively homogeneous demographic communities, typically ranging in population size from 2,500 to 8,000 residents [@Bureau2014; @Bureau1994].
 Compared to zip codes, which are fixed spatial units designed primarily to demarcate route efficiency for postal services, census tracts are considered more reliable for population-based analyses.
@@
 
-Across these data sources, we created measures of race and ethnicity, specifically the number and proportion of the census-tract population in the following categories: White, non-Hispanic; Black, non-Hispanic; Asian, non-Hispanic; Pacific Islander, non-Hispanic; American Indian and Alaskan Native, non-Hispanic; and two or more races, non-Hispanic.
-Measures of Asian, American Indian and Alaskan Native, two or more races, and Native Hawaiian and Pacific Islander (all non-Hispanic) were not available for the 1980 Decennial Census.
+Across these data sources, we created measures of race and ethnicity, specifically the number and proportion of the census-tract population in the following categories: White, non-Hispanic; Black, non-Hispanic; Hispanic; Asian, non-Hispanic; Native Hawaiian and Pacific Islander, non-Hispanic; American Indian and Alaskan Native, non-Hispanic; and two or more races, non-Hispanic.
+Measures of Asian, Native Hawaiian and Pacific Islander, American Indian and Alaskan Native, and two or more races (all non-Hispanic) were not available for the 1980 Decennial Census.
 We also created socioeconomic measures that approximate the predictors of market segment identified by @RN4982.
 These variables are median household income, percent of households below the poverty line, and people age 25+ with a BA.
-All measures were created at the census tract-level and at the the Geomarket-level by aggregating tract-level data to Geomarkets.
+All measures were created at the census tract-level and at the Geomarket-level by aggregating tract-level data to Geomarkets.
 
@@
 Even under this most conservative approach to addressing spatial misalignment, overall patterns in the findings remained consistent.
-However, we opted to georeference the manually digitized maps directly sourced from the College Board, as the original source of the Geomarket-zip code crosswalk cited in the the R-bloggers post could not be verified.
+However, we opted to georeference the manually digitized maps directly sourced from the College Board, as the original source of the Geomarket-zip code crosswalk cited in the R-bloggers post could not be verified.
 
@@
 We chose these states because they were priorities for the two foundations that funded the data collection.
-Utilizing public records requests to obtain quantitative data is a painstaking process, made more difficult by the onset of the COVID19 Pandemic.
+Utilizing public records requests to obtain quantitative data is a painstaking process, made more difficult by the onset of the COVID-19 Pandemic.
 Initially, the majority of universities ignored or denied our request.
 We established a client relationship with a civil rights legal organization and leveraged this relationship to obtain pro bono representation from four law firms.
-Firm representation substantially increased the success of data collection.
 We also collected data from Arizona State University and Northern Arizona University, because we were able to obtain pro-bono legal representation for these two universities.
@@
 Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic [@RN0010].[^20] Information about student socioeconomic status is limited.
-However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire.
+However, we create a measure of whether the prospect is a first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire.
 
 [^20]: Any student who selects a Hispanic ethnicity category is defined as Hispanic/Latino, regardless of the race categories they select.
-Additionally, non-Hispanic students who check “American Indian or Alaska Native” and another race group are defined as "two or more races, non-Hispanic."
+Additionally, non-Hispanic students who check "American Indian or Alaska Native" and another race group are defined as "two or more races, non-Hispanic."
 
@@
 Our goal is not to make generalizations about Geomarket variation nationally, but to explore how different geographical models of metropolitan areas, shaped by their unique demographic characteristics, influence the composition of prospect pools.
-Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomenon in rich and nuanced contexts.
+Such aims align with the purpose of case studies, which drive the exploration of research questions focused on the "how" and "why" of phenomena in rich and nuanced contexts.
 We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics.
@@
 The central thesis of Zemsky & Oedel's (1983) Market Segment Model suggests that student demand for various colleges is a function of social class and geography, and that colleges strategically target recruiting in high socioeconomic areas.
-However, because Geomarkets vary significantly in their racial and class characteristics across Geomarkets, and in how colleges target these Geomarkets, a multiple case study enables us to compare how these dynamics unfold across different metropolitan areas.
-By treating each metropolitan area as a separate case, we explore of both within-case processes (e.g., how homophily structures recruitment in a single metropolitan area) and cross-case patterns (e.g., how the same recruitment strategy results in different racial/class outcomes across metropolitan areas).
-
-Case selection was informed by purposeful sampling to identify "information-rich" cases that can provide knowledge about variation across Geomarkets and over time [@RN4011; @RN4549] .
-We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
+However, because Geomarkets vary significantly in their racial and class characteristics, and in how colleges target these Geomarkets, a multiple case study enables us to compare how these dynamics unfold across different metropolitan areas.
+By treating each metropolitan area as a separate case, we explore both within-case processes (e.g., how homophily structures recruitment in a single metropolitan area) and cross-case patterns (e.g., how the same recruitment strategy results in different racial/class outcomes across metropolitan areas).
+
+Case selection was informed by purposeful sampling to identify "information-rich" cases that can provide knowledge about variation across Geomarkets and over time [@RN4011; @RN4549].
+We considered metropolitan areas as information rich cases across four broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data.
 Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p.
 11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia.
-To compare and contrast with classic urban model structures, we also identified alternative urban models such as bi-modal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
+To compare and contrast with classic urban model structures, we also identified alternative urban models such as bi-modal or metroplex models that contain two centralized cores surrounded by the more classic suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City).
 The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S.
 regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations.
-Related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2).
+Related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan area (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2).
 We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.
@@
 In order to show which prospective students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets, the student list purchases we analyze must include all Geomarkets in a metropolitan area.
-This lead us to select on orders that filtered for an entire state (e.g., State filter is set to Pennsylvania), an entire metropolitan area without a Geomarket filter (CBSA filter is set to the Philadelphia metropolitan area), or an entire metropolitan area with a Geomarket filter inclusive of all Geomarkets (CBSA filter is set to the Philadelphia metropolitan area and Geomarket filter is set to all five Geomarkets in the Philadelphia area).
-Although we would like to analyze the D.C./Maryland/Virginia (DMV) area, the student list purchases we collected exclude "MD 5 -- Prince Georges Metropolitan" and "MD 7 -- Baltimore (Urban), which are substantially poorer and more Black than other Geomarkets in the DMV.
+This lead us to select on orders that filtered for an entire state (e.g., State filter is set to Pennsylvania), an entire metropolitan area without a Geomarket filter (e.g., CBSA filter is set to the Philadelphia metropolitan area), or an entire metropolitan area with a Geomarket filter inclusive of all Geomarkets (e.g., CBSA filter is set to the Philadelphia metropolitan area and Geomarket filter is set to all five Geomarkets in the Philadelphia area).
+Although we would like to analyze the D.C./Maryland/Virginia (DMV) area, the student list purchases we collected exclude "MD 5 -- Prince George's Metropolitan" and "MD 7 -- Baltimore (Urban)," which are substantially poorer and more Black than other Geomarkets in the DMV.
 Similarly, student list purchases targeting Cleveland exclude "OH 4 -- City of Cleveland (East)."
@@
 
-We answer RQ2 by analyzing student list purchases that encompassed all Geomarkets in a selected metropolitan areas.
+We answer RQ2 by analyzing student list purchases that encompassed all Geomarkets in a selected metropolitan area.
 Descriptive tables and graphs describe the racial and socioeconomic characteristics of prospects that would have been included/excluded had the purchase filtered on particular Geomarkets.
@@
 12].
-Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area - Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie - experienced relatively little population growth from 1980 through 2020.
+Based on the color gradient scale representing total population density, @fig-chicago-maps indicates northern suburb Geomarkets in the metropolitan area -- Chain of Lakes, Northwest Suburbs, Northshore, Evanston and Skokie -- experienced relatively little population growth from 1980 through 2020.
 In contrast, the most significant population increases occurred in the Western Suburbs (1.0 Million to 1.4 Million) and the South and Southwest Suburbs (less than 1 Million to 1.2 Million) Geomarkets.
@@
 @fig-chicago-rq1-race also indicates that, although the North Shore Geomarket underwent some changes in racial and ethnic composition, its overall demographic patterns remained relatively the same across the three census periods.
-City of Chicago experienced relatively modest declines in its Black population (35% in 1980 to 26% in 2020), alongside increases in its Hispanic (13% in 1980 to 30% in 2020) and Asian Pacific Islander populations (4% in 2000 to 7% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
+City of Chicago experienced relatively modest declines in its Black population (35% in 1980 to 26% in 2020), alongside increases in its Hispanic (13% in 1980 to 30% in 2020) and Asian populations (4% in 2000 to 7% in 2020), further substantiating the Geomarket as the most racially/ethnic diverse in the metropolitan area.
 <!-- For instance, North Shore experienced some increases in Hispanic (2% in 1980 to 8% in 2020), Asian Pacific Islander (7% in 2000 to 10% in 2020), and Black (2% in 1980 to 3% in 2020) populations yet remained a predominantly White Geomarket (74% in 2020).
@@
 
-The interactive maps in the [Online Appendix](https://ozanj.github.io/cb_geo/#/research-question-1-15) can show these shifts at the Census Tract-level.
-For example, in 1980 Black people were concentrated in southern tracks of the City of Chicago Geomarket.
-As described by scholarship on gentrification in Chicago [@ewing2024_nofuture; @snidal2022_exodus].
-However, in 2000 and 2020 Black populations shift towards northeast tracks of the South and Southwest Suburbs Geomarket
+The interactive maps in the [Online Appendix](https://ozanj.github.io/cb_geo/#/research-question-1-15) can show these shifts at the census tract-level.
+For example, in 1980 Black people were concentrated in southern tracts of the City of Chicago Geomarket, as described by scholarship on gentrification in Chicago [@ewing2024_nofuture; @snidal2022_exodus].
+However, in 2000 and 2020 Black populations shift towards northeast tracts of the South and Southwest Suburbs Geomarket
 
 We explore socioeconomic demographics -- median income, percent living below poverty, and educational attainment -- within and across Geomarkets over time in @fig-chicago-rq1-ses.
-The primary takeaways are the City of Chicago -- and to a somewhat lesser degree th South and Southwest Suburbs - stands out as having lower median income, higher poverty rates, and lower BA attainment than other Geomarkets.
+The primary takeaways are the City of Chicago -- and to a somewhat lesser degree the South and Southwest Suburbs -- stands out as having lower median income, higher poverty rates, and lower BA attainment than other Geomarkets.
 By contrast, the North Shore stands out as being particularly affluent and highly educated.
@@
 In 1980, the City of Chicago and South and Southwest Suburbs had the lowest percentage of residents with a bachelor's degree or higher (14%).
-However, by 2020, the City of Chicago experienced the largest percentage point increase in college degree attainment, reaching 34%.
-
-Overall, demographic shifts transformed the Chicago-land Geomarket landscape from largely a binary structure into a more tiered dynamic that retains by substantial racial/ethnic and socioeconomic inequality between Geomarkets.
+However, by 2020, the City of Chicago experienced the largest percentage point increase in college degree attainment, reaching nearly 40%.
+
+Overall, demographic shifts transformed the Chicago-land Geomarket landscape from largely a binary structure into a more tiered dynamic that retains substantial racial/ethnic and socioeconomic inequality between Geomarkets.
 The original Geomarket boundaries delineated significant racial and income segregation in the 1980s.
@@
 
-### Composition of Students Lists by Geomarket Contributions
+### Composition of Student Lists by Geomarkets
 
 To examine how the socioeconomic and racial composition of included versus excluded prospects varies when student list purchases filter on particular Geomarkets, we analyze six orders -- placed by a research university -- that target prospects across the entire state of Illinois.
-The first two orders filter on the same SAT thresholds (ranging from 1020 to 1150), the same GPA thresholds (ranging from B- to A+), and filter on the high school graduating classes of 2020 and 2021 graduating classes, respectively.
+The first two orders filter on the same SAT thresholds (ranging from 1020 to 1150), the same GPA thresholds (ranging from B- to A+), and filter on the high school graduating classes of 2019/2020 and 2020/2021, respectively.
 The remaining four orders used the same GPA and high school class filters but differed in SAT thresholds.
@@
 @fig-chicago-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Chicago metropolitan area, notably revealing that the City of Chicago stands out distinctly from all other Geomarkets in terms of the substantial racial and ethnic diversity of its purchased profiles.
-At the lowest PSAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
-For example, more than 7,000 of the 15,964 total prospects in this SAT range order were concentrated in the Western Suburbs (59% White, 9% Asian, 6% Black, 23% Hispanic, 4% Multiracial) and South and Southwest Suburbs (54% White, 2% Asian, 19% Black, 21% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
+At the lowest SAT score thresholds (top panel, 1020-1150) the racial/ethnic distribution of prospects within Geomarkets most closely reflects the demographic makeup of the metropolitan area in 2020 (see @fig-chicago-rq1-race).
+For example, more than 7,000 of the 15,964 total prospects (whose race/ethnicity is known) in this SAT range order were concentrated in the Western Suburbs (59% White, 9% Asian, 6% Black, 23% Hispanic, 4% Multiracial) and South and Southwest Suburbs (54% White, 2% Asian, 19% Black, 21% Hispanic, and 3% Multiracial), both of which yield majority White prospect pools comparable to the racial/ethnic makeup of their overall Geomarket populations.
 The North Shore (70% White, 12% Asian, 4% Black, 11% Hispanic) and the Northwest Suburbs (59% White, 11% Asian, 2% Black, 24% Hispanic) Geomarket pools exhibit the greatest proportions of White prospects.
@@
 
-As PSAT score ranges increase, The City of Chicago Geomarket becomes the only Geomarket with substantial shares of Black and Hispanic prospects.
-Moving from the top to the bottom panels, all Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline.
-For the highest SAT score range, the City of Chicago is the only Geomarket where White prospects comprise a slim majority (55%), with the remaining prospect pool made up of 16% Asian, 6% Black, 18% Hispanic, 5% Multiracial students.
-Prospect pools from all other Geomarkets in this high PSAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.3% (North Shore) to 4% (South and Southwest Suburbs) of prospects and 5% (North Shore) to 10% (South and Southwest Suburbs) of prospects, respectively.
+As SAT score ranges increase, the City of Chicago Geomarket becomes the only Geomarket with substantial shares of Black and Hispanic prospects.
+Moving from the top to the bottom panels in @fig-chicago-rq2-race, all Geomarkets experience growth in the shares of White and Asian prospects, while Black and Hispanic prospect shares decline.
+For the highest SAT score range, the City of Chicago is the only Geomarket where White prospects comprise a slim majority (56%), with the remaining prospect pool made up of 16% Asian, 6% Black, 18% Hispanic, and 5% Multiracial students.
+Prospect pools from all other Geomarkets in this high SAT order are predominantly White and Asian, with Black and Hispanic shares ranging from 0.3% (North Shore) to 4% (South and Southwest Suburbs) of prospects and 5% (North Shore) to 10% (South and Southwest Suburbs) of prospects, respectively.
 <!--# [LEDE -- BUT IN BOTH THE MEDIUM AND HIGH THRESHOULD, CHICAGO IS THE ONLY GEOMARKET WHERE A SUBSTANTIAL SHARE OF PROSPECTS ARE NOT WHITE OR ASIAN] -->
@@
 We also explore contributions across Geomarkets to overall prospect pools by race/ethnicity.
-@fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle range SAT scores (1160-1300) by race/ethnicity.
+@fig-chicago-rq2-midSAT-race shows the contributions of each Geomarket to the pool of prospects from the two orders targeting middle-range SAT scores (1160-1300) by race/ethnicity.
 The top panel presents the full distribution of the 13,729 total prospects (whose race/ethnicity is known) included in these orders by Geomarkets: 10% of all purchased prospect profiles reside in Chain of Lakes, 11% in Northwest Suburbs, 6% in North Shore, 4% in Evanston and Skokie, 23% in City of Chicago, 29% in Western Suburbs, and 17% in South and Southwest Suburbs.
@@
 The City of Chicago accounts for 23% of all prospects but 51% of Black students.
-Similarly, South & Southwest Suburbs accounts for 17% of all prospects but 26% of all Black prospects.
+Similarly, South and Southwest Suburbs accounts for 17% of all prospects but 26% of all Black prospects.
 The upshot is that if the student list purchases had excluded these two Geomarkets, 77% of the nearly 1,000 Black students currently included in these purchases would be excluded.
@@
 
-These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Chicago metropolitan area using Geomarket filters that exclude the City of Chicago– as observed in some orders collected via our public requests – result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
+These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Chicago metropolitan area using Geomarket filters that exclude the City of Chicago -- as observed in some orders collected via our public requests -- result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
 The [Online Appendix](https://ozanj.github.io/cb_geo/#/results) shows that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters.
@@
 We also examine the contributions of each Geomarket to the pool of prospects based on first-generation college student status.
-Similar to race/ethnicity figures above, @fig-chicago-rq2-midSAT-ses presents the Geomarket distribution for the 13,848 total prospects (whose first-generations status is known) included in middle-range SAT score orders (1160-1300).
-Among the 1,750 prospects whose parents did not attend college ("no college"), the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students - 43% in comparison to 23% for all included prospects in this SAT range.
+Similar to race/ethnicity figures above, @fig-chicago-rq2-midSAT-ses presents the Geomarket distribution for the 13,848 total prospects (whose first-generation status is known) included in the middle-range SAT score orders (1160-1300).
+Among the 1,750 prospects whose parents did not attend college ("no college"), the City of Chicago stands out as the only Geomarket contributing a disproportionately larger share of first-generation college students -- 43% in comparison to 23% for all included prospects in this SAT range.
 In contrast, all other Geomarkets contributed disproportionately smaller shares of first-generation college students whose parents did not attend college relative to all included prospects, with such disparity ranging from a one-percentage-point difference (Northwest Suburbs, South and Southwest Suburbs) to a ten-percentage-point difference (Western Suburbs).
@@
 Finally, we examine Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Chicago metropolitan area.
-@fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 13,059 total prospects (whose first-generation status and race/ethnicity is known) included in middle-range SAT (1160-1300) orders.
+@fig-chicago-rq2-midSAT-combo presents the Geomarket distribution of the 13,059 total prospects (whose first-generation status and race/ethnicity is known) included in the middle-range SAT (1160-1300) orders.
 @fig-chicago-rq2-midSAT-combo shows the City of Chicago Geomarket consistently contributes a disproportionately larger share of first-generation college students whose parents did not attend college across all racial/ethnic categories.
 For instance, among all Asian prospects (n=1,370) whose contact information was purchased in this SAT score range, approximately 12% were first-generation college students whose parents did not attend college, 16% were first-generation college students whose parents attended some college but did not complete their degree, and 73% had parents with a BA.
-By contrast, for 292 Asian students from the City of Chicago Geomarket, 30% had parents who did not attend college and less than 50% had parents with a BA.
+By contrast, for 292 Asian students from the City of Chicago Geomarket, 30% had parents who did not attend college and approximately 50% had parents with a BA.
 
 For Hispanic prospects, the City of Chicago stands out as the Geomarket contributing the largest overall number and share of Hispanic first-generation college students across those whose parents did not attend college (44% versus 38% overall) and those whose parents attended some college but did not complete their degrees (29% versus 26% overall).
-Other Geomarkets such as Northwest Suburbs and Chain of lakes also contributed substantially larger shares.
-By contrast, Hispanic students from the more affluent North Shore Geomarket and Evanston & Skokie Geomarket were more likely to have parents with a BA.
+Other Geomarkets such as Northwest Suburbs and Chain of Lakes also contributed substantially larger shares.
+By contrast, Hispanic students from the more affluent North Shore Geomarket and Evanston and Skokie Geomarket were more likely to have parents with a BA.
 
@@
 On the other hand, predominantly White and affluent Geomarkets in the suburban ring of the metropolitan area like the Chain of Lakes, North Shore, and Evanston and Skokie contribute larger proportions of White and Asian prospects, as well as non-first generation college students.
-However, the City of Chicago stands out in playing crucial role in fostering college access for first-generation Asian and Hispanic college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
+However, the City of Chicago stands out in playing a crucial role in fostering college access for first-generation Asian and Hispanic college students, contributing such prospects to the overall pool at significantly higher, disproportionate rates than any other Geomarket.
 Given the broader racial/ethnic distribution of the Chicago metropolitan area, student list orders that use Geomarket filters excluding the City of Chicago and including more affluent Geomarkets in targeting high-achieving students may still capture some Asian and Hispanic prospects.
-However, these students are far more unlikely to be first-generation students of color.
+However, these students are far less likely to be first-generation students of color.
 
@@
 The Dallas-Fort Worth region represents a unique "metroplex" case formed by the convergence of two separate metropolitan areas, which creates a unique urban spatial structure that differs from the classic urban core and suburban ring model showcased by Chicago.
-As depicted in @fig-dallas-maps, two distinct inner-core Geomarkets (City of Dallas and City of Fort Worth) are separated by a centrally located suburban Geomarket encompassing Irving, Arlington, & Grand Prairie.
-These central Geomarkets are surrounded by the more conventional outer ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Ft.
-Worth Metroplex).
+As depicted in @fig-dallas-maps, two distinct inner-core Geomarkets (City of Dallas and City of Fort Worth) are separated by a centrally located suburban Geomarket encompassing Irving, Arlington, and Grand Prairie.
+These central Geomarkets are surrounded by the more conventional outer ring of suburban Geomarkets (Dallas County, Collin and Rockwall Counties, and West of Dallas/Fort Worth Metroplex).
 
@@
 @fig-dallas-maps shows all six Geomarkets experienced population growth over the three census periods.
-However, the most pronounced increases occurred in the West of Dallas/Fort Worth Metroplex and Collin & Rockwall Counties Geomarkets, which grew from approximately 350,000 to 1.6 Million and 160,000 to 1.1 Million, respectively.
-These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers -- City of Forth Worth and City of Dallas -- into the surrounding suburban and rural areas of the metropolitan areas.
+However, the most pronounced increases occurred in the West of Dallas/Fort Worth Metroplex and Collin and Rockwall Counties Geomarkets, which grew from approximately 350,000 to 1.6 Million and 160,000 to 1.1 Million, respectively.
+These population trends, as shown in @fig-dallas-maps, suggest population growth has been partly driven by urban sprawl from the metropolitan centers -- City of Fort Worth and City of Dallas -- into the surrounding suburban and rural areas of the metropolitan areas.
 This outward expansion may also account for the comparatively larger geographic size of the West of Dallas/Fort Worth Metroplex Geomarket from 1980 to 2000.
@@
 @fig-dallas-rq1-race presents the racial and ethnic composition of each Geomarket within the Dallas-Fort Worth metropolitan area across the three census periods.
-In 1980, all Geomarkets were predominantly White, with White populations ranging from 62% in City of Dallas to 93% in West of Dallas/Fort Worth Metroplex.
-By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County, Irving, Arlington, and Grand Prairie, and City of Fort Wort.
-In contrast, Collin & Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 78% and 79%, respectively.
+In 1980, all Geomarkets were predominantly White, with White populations ranging from 61% in City of Dallas to 93% in West of Dallas/Fort Worth Metroplex.
+By 2000, however, Geomarkets located in the central part of the metropolitan area exhibited a more balanced racial/ethnic distribution, including Dallas County, Irving, Arlington, and Grand Prairie, and City of Fort Worth.
+In contrast, Collin and Rockwall Counties and West of Dallas/Fort Worth Metroplex Geomarkets remained predominantly White in 2000 at 77% and 80%, respectively.
 By 2000, the City of Dallas was no longer a predominantly White Geomarket, with a population that was approximately 25% Black, 36% Hispanic, and 2% Asian.
@@
 As a result, --> In 2020, both City of Dallas and City of Fort Worth continued to rank at the lowest levels of affluence within the metropolitan area at \$68,000 and \$74,000 median household incomes, respectively.
-By contrast, Collin & Rockwall Counties remained the most affluent Geomarket with a median household income of \$124,000, followed by West of Dallas/Fort Worth Metroplex (\$104,000).
+By contrast, Collin and Rockwall Counties remained the most affluent Geomarket with a median household income of \$124,000, followed by West of Dallas/Fort Worth Metroplex (\$104,000).
 @fig-dallas-rq1-ses also illustrates changes in the percentage of residents living in poverty and educational attainment within Geomarkets over time, revealing similar dynamics across Geomarkets to those observed in median household income.
-Together, these trends underscore the growingly racial/ethnic diverse but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
+Together, these trends underscore the growing racial/ethnic diversity but widening economic inequality within the evolving Geomarket landscape of the Dallas-Fort Worth metroplex.
 
@@
 
-### Composition of Students Lists by Geomarkets
+### Composition of Student Lists by Geomarkets
 
@@
 Specifically, we analyze three list orders placed by a research university targeting prospects across 87 different Geomarkets nationwide, including all six Geomarkets in the Dallas-Fort Worth Metropolitan area.
-Each order also filtered for prospects in the 2019-2020 graduating class and based on PSAT scores across three ranges: 1070-1180, 1190-1260, and 1270-1520.
-Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260).
-
-@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forth Worth area, revealing a comparatively more distributed pattern of racial/ethnic diversity across Geomarkets unlike Chicago's concentrated diversity within a single Geomarket.
+Each order also filtered for prospects in the 2020 graduating class and on PSAT scores across three ranges: 1070-1180, 1190-1260, and 1270-1520.<!-- Similar to Chicago analyses, we focus on prospects within the middle PSAT score range (1190-1260).-->
+
+@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Fort Worth area, revealing a comparatively more distributed pattern of racial/ethnic diversity across Geomarkets unlike Chicago's concentrated diversity within a single Geomarket.
 This distribution is most notable at the lowest PSAT score threshold.
@@
 While the urban central Geomarkets of City of Dallas and City of Fort Worth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportions of Hispanic prospects but lower proportions of Asian and Black students.
-The Collin & Rockwall Counties (58% White, 7% Black, 17% Hispanic, 13% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (60% White, 7% Black, 18% Hispanic, 9% Asian) both yield the largest but predominantly White prospect pools.
+The Collin and Rockwall Counties (58% White, 7% Black, 17% Hispanic, 13% Asian) and West of Dallas/Fort Worth Metroplex Geomarkets (60% White, 7% Black, 18% Hispanic, 9% Asian) both yield the largest but predominantly White prospect pools.
 Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.
@@
 @fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions.
-The bars for "All (race known)" show that Collin & Rockwall (30%) and West of Dallas/Forth Worth (27%) contribute many more students than the other four Geomarkets.
+The bars for "All (race known)" show that Collin and Rockwall (30%) and West of Dallas/Fort Worth (27%) contribute many more students than the other four Geomarkets.
 Disparities in Geomarket contributions were evident but differed for Asian, Black, and Hispanic students.
-For instance, The Dallas County (19% versus 12% for all prospects) and Colin and Rockwall Counties (39% versus 30% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
+For instance, The Dallas County (19% versus 12% for all prospects) and Collin and Rockwall Counties (39% versus 30% for all prospects) Geomarkets contributed a disproportionately larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.
 
@@
 Contributions of each Geomarket to the pool of prospects based on first-generation college student status from the Dallas-Fort Worth metropolitan area are presented in @fig-dallas-rq2-midSAT-ses.
-Only 303 prospects of the 3,926 total in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
+Only 303 prospects of the 3,926 total (whose race/ethnicity is known) in this middle PSAT score range pool were first-generation college students whose parents did not attend college.
 Again, all Geomarkets besides Collin and Rockwall Counties (12% versus 30% for all prospects) and West of Dallas/Fort Worth Metroplex (16% versus 27% for all prospects) contributed equal or larger proportions of first-generation college students (no college) relative to their contributions to the overall prospect pool.
@@
 Among all White prospects (n=2,093) whose contact information was purchased in this middle PSAT score range, approximately 3% were first-generation college students whose parents did not attend college, 11% were first-generation college students whose parents attended some college but did not complete their degree, and 86% were not first-generation college students.
-@fig-dallas-rq2-midSAT-combo shows almost all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool, the only modest exception being the Irving, Arlington, & Grand Prairie Geomarket.
+@fig-dallas-rq2-midSAT-combo shows almost all Geomarkets contribute a nearly proportionate share of prospects by first-generation college students relative to the overall pool, the only modest exception being the Irving, Arlington, and Grand Prairie Geomarket.
 
@@
 Furthermore, first generation status differed substantially across Geomarkets, especially for Hispanic and Asian students.
-For Hispanic students, more than 50% of prospects from the City of Dallas, City of Forth Worth, Arlington & Grand Prairie, and Dallas County are first-generation college students.
-By contrast, for the more affluent Collin & Rockwall Geomarket and the West of Dallas/Fort Worth Metroplex Geomarkets, less than 30% of prospects were first-generation.
-For Asian prospects, City of Fort Worth and Irving, Arlington, & Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
-By contrast, the more affluent Collin & Rockwall sent more Asian prospects (234) than any other Geomarket, but about 89% of these prospects were not first-generation college students.
+For Hispanic students, more than 50% of prospects from the City of Dallas, City of Fort Worth, Irving, Arlington and Grand Prairie, and Dallas County are first-generation college students.
+By contrast, for the more affluent Collin and Rockwall Geomarket and the West of Dallas/Fort Worth Metroplex Geomarkets, less than 30% of prospects were first-generation.
+For Asian prospects, City of Fort Worth and Irving, Arlington, and Grand Prairie Geomarkets contributed disproportionately larger shares of first-generation college students across both parents with no college and some college.
+By contrast, the more affluent Collin and Rockwall sent more Asian prospects (234) than any other Geomarket, but about 89% of these prospects were not first-generation college students.
 
@@
 
-Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles -- 11 Geomarkets, results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), results for which are shown in the [Online Appendix](https://ozanj.github.io/cb_geo/#/results).
+Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles (11 Geomarkets) -- results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), results for which are shown in the [Online Appendix](https://ozanj.github.io/cb_geo/#/results).
 The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
@@
 As depicted in @fig-la-maps, Geomarkets closely follow this overall metropolitan model.
-Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring made up by the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
+Hollywood and Wilshire, South and South Central Los Angeles, and East Los Angeles Geomarkets are centrally located in the metropolitan area followed by an outer core ring made up of the Glendale and Pasadena, San Fernando Valley-East, West Los Angeles and West Beach, South Bay, and Long Beach Geomarkets.
 The San Fernando Valley-West Geomarket is located in the northwest part of the region, whereas Covina and West Covina as well as the Riverside, San Bernardino, and Ontario Geomarkets are located inland, comprising the Inland Empire.
@@
 The population growth in the Los Angeles metropolitan area increased from approximately 8 Million in 1980 to more than 12 Million in 2020.
-@fig-la-maps showcases population growth by Geomarkets over the three census periods reflect the urban sprawl forming the multi-nuclei model.
-The most pronounced increases occurred in South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West, Riverside and San Bernardino, and Ontario.
+@fig-la-maps showcases population growth by Geomarkets over the three census periods, reflecting the urban sprawl forming the multi-nuclei model.
+The most pronounced increases occurred in the South and South Central Los Angeles Geomarket, as well as the outer/inland Geomarkets of San Fernando Valley-West and Riverside, San Bernardino, and Ontario.
 The South and South Central Los Angeles Geomarket grew from approximately 1.1 to 1.6 Million from 1980 to 2020.
-In contrast, the San Fernando Valley-West and Riverside and San Bernardino, and Ontario Geomarkets, respectively, grew from approximately 580,000 to nearly 1 Million and from 1.1 Million to 3.3 Million during this time.
+In contrast, the San Fernando Valley-West and Riverside, San Bernardino, and Ontario Geomarkets, respectively, grew from approximately 580,000 to nearly 1 Million and from 1.1 Million to 3.3 Million during this time.
 
 @fig-la-rq1-race presents the racial and ethnic composition of each Geomarket within the Los Angeles metropolitan area from 1980 to 2020.
-One data caveat here is that the Asian race/ethnicity category was not available in 1980.
+As mentioned previously, a data caveat here is that the Asian race/ethnicity category was not available in 1980.
 However, over the three decennial census periods, the proportion of White residents declined across all Geomarkets, while Asian and Hispanic populations generally increased.
-In 1980, all Geomarkets - except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire- were predominantly White, with percentages ranging from 63% White in Covina and West Covina to 86% in the San Fernando Valley-East Geomarket.
+In 1980, all Geomarkets -- except for East Los Angeles, South and South Central Los Angeles, and Hollywood and Wilshire -- were predominantly White, with percentages ranging from 63% White in Covina and West Covina to 86% in the San Fernando Valley-East Geomarket.
 By 2020, however, only the West Los Angeles and West Beach (51%) and San Fernando Valley-East (59%) Geomarkets remained predominantly White.
-Remaining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West- (40% White, 4% Black, 40% Hispanic; 13% Asian Pacific Islander), Glendale and Pasadena (35% White, 4% Black, 41% Hispanic; 17% Asian Pacific Islander), South Bay (32% White, 10% Black, 32% Hispanic; 22% Asian Pacific Islander), Long Beach (27% White, 10% Black, 47% Hispanic; 13% Asian Pacific Islander), Covina and West Covina (19% White, 4% Black, 57% Hispanic; 19% Asian Pacific Islander), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 53% Hispanic; 8% Asian Pacific Islander).
+The remaining Geomarkets exhibited a more racially/ethnic diverse population by 2020, including San Fernando Valley-West (40% White, 4% Black, 40% Hispanic; 13% Asian), Glendale and Pasadena (35% White, 4% Black, 41% Hispanic; 17% Asian), South Bay (32% White, 10% Black, 32% Hispanic; 22% Asian), Long Beach (27% White, 10% Black, 47% Hispanic; 13% Asian), Covina and West Covina (19% White, 4% Black, 57% Hispanic; 19% Asian), and Riverside, San Bernardino, and Ontario (28% White, 7% Black, 53% Hispanic; 8% Asian).
 However, the South and South Central Los Angeles and Hollywood and Wilshire Geomarkets experienced substantial declines in the share of Black residents from 1980 to 2020, dropping from 44% to 16% and 22% to 11%, respectively.
@@
 @fig-la-rq1-ses presents trends in income, poverty, and educational attainment in the Los Angeles metropolitan area from 1980 to 2020, demonstrating the South and South Central Los Angeles, East Los Angeles, and Hollywood and Wilshire Geomarkets remained at the lowest socioeconomic levels across these indicators.
-For instance, the median household income in the South and South Central Los Angeles Geomarket increased modestly from \$51,000 in 1980 to \$57,000 in 2000, reaching \$61,000 by 2020.
+For instance, the median household income in the South and South Central Los Angeles Geomarket increased modestly from \$51,000 in 1980 to \$56,000 in 2000, reaching \$61,000 by 2020.
 In comparison, East Los Angeles and Hollywood and Wilshire recorded slightly higher median incomes in 2020 at \$74,000 and \$70,000, respectively.
@@
 South and South Central Los Angeles accounts for only 3% of all purchased prospect profiles but accounts for 8% of Black prospects.
-The Riverside, San Bernardino, & Ontario Geomarket accounts for 24% of all prospects but 29% of Black prospects.
-For Hispanic prospects, South and South Central Los Angeles (9% versus 3% for all prospects) and Riverside, San Bernardino, and Ontario (30% versus 24% for all prospects) also contributed a disproportionately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
-For Asian students, the Glendale and Pasadena (20% of Asian prospects versus 15% for all prospects), East Los Angeles (14% of Asian prospects versus 7% of all prospects) and Covina and West Covina (12% of Asian prospects versus 9% of all prospects) Geomarkets contributed a disproportionate larger share of prospects relative to their contributions to the overall prospect pool.
-For the more than 2,300 White prospects in this pool, the San Fernando Valley - West (17% of White prospects versus 11% of all prospects), West Los Angeles and West Beach (9% of White prospects versus 6% for all prospects), and Glendale and Pasadena (17% of White prospects versus 15% for all prospects) Geomarkets contribute disproportionately larger shares.
+The Riverside, San Bernardino, and Ontario Geomarket accounts for 24% of all prospects but 29% of Black prospects.
+For Hispanic prospects, South and South Central Los Angeles (9% versus 3% for all prospects) and Riverside, San Bernardino, and Ontario (30% versus 24% for all prospects) also contributed a disproportionately larger proportion of Hispanic prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area.
+For Asian students, the Glendale and Pasadena (20% versus 15% for all prospects), East Los Angeles (14% versus 7% for all prospects) and Covina and West Covina (12% versus 9% for all prospects) Geomarkets contributed a disproportionately larger share of Asian prospects relative to their contributions to the overall prospect pool.
+For the more than 2,300 White prospects in this pool, the San Fernando Valley - West (17% versus 12% for all prospects), West Los Angeles and West Beach (9% versus 6% for all prospects), and Glendale and Pasadena (17% versus 15% for all prospects) Geomarkets contribute disproportionately larger shares of White prospects.
 
@@
 
-Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area.
-Across all Hispanic prospects in the middle range PSAT orders, 26% were first-generation college students whose parents did not attend college, 26% were first-generation college students whose parents attended some college but did not complete their degrees, and 49% were not first-generation college students.
+Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area for the middle range SAT orders.
+Across all Hispanic prospects, 26% were first-generation college students whose parents did not attend college, 26% were first-generation college students whose parents attended some college but did not complete their degrees, and 49% were not first-generation college students.
 By contrast, for the 175 Hispanic prospects from the South and South Central Geomarket, 51% had parents who never attended college, 28% had parents who attended some college, and only 21% had parents with a BA.
@@
 @fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 276), a significant proportion of whom are first-generation college students (34%) whose parents did not attend college.
-The Hollywood & Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (34%).
+The Hollywood and Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (34%).
 
 While the pool of Black prospects is relatively small (n=182) in @fig-la-rq2-midSAT-combo, some patterns emerge by first-generation status and Geomarket.
-Overall, approximately 6% of Black prospects were first-generation college students-no college, 21% were first-generation college students-some college, and 74% not first-generation).
+Overall, approximately 6% of Black prospects were first-generation college students whose parents did not attend college, 21% were first-generation college students whose parents attended some college, and 73% were not first-generation.
 The South and South Central Los Angeles Geomarket stands out for contributing a disproportionately high share (13%) of Black first-generation college students whose parents did not attend college, despite representing a relatively smaller overall number.
@@
 
-The discussion of homophily by @RN4975 describes not only the logic of The Market Segment model but also a broader institutional logic of the enrollment management industry.
+The discussion of homophily by @RN4975 describes not only the logic of the Market Segment Model but also a broader institutional logic of the enrollment management industry.
 The logic of homophily assumes that there is a natural cultural match between students and colleges, suggesting colleges should find prospective students who resemble students who enrolled at their institution in the past and focus recruiting efforts on the same schools and communities that are popular with peer colleges.
@@
 @RN4982 illustrated this argument by demonstrating a correlation between student SAT score-sending behavior and measures of class.
-The thesis of the Market Segment Model is that that households of specific class characteristics and college aspirations are likely to live in particular communities and that these geographic territories can be meaningfully captured by Geomarkets.
+The thesis of the Market Segment Model is that households of specific class characteristics and college aspirations are likely to live in particular communities and that these geographic territories can be meaningfully captured by Geomarkets.
 Although Geomarkets appear coarse by contemporary standards, they were a landmark innovation in Geographic segmentation.
@@
 Findings illustrate that in 1980, when Geomarkets were being created, Black and Hispanic residents tended to be highly concentrated in the poorest Geomarket in the Chicago, Dallas-Fort Worth, and to some extent the Los Angeles metropolitan areas.
-These Geomarket patterns extend to other metropolitan areas in the country, such as CA7 -- City of Oakland, MA 6 -- Boston & Cambridge, OH 4 -- City of Cleveland (East), TX17 -- City of Houston (East), and PA5 -- Philadelphia County.
+These Geomarket patterns extend to other metropolitan areas in the country, such as CA7 - City of Oakland, MA 6 - Boston and Cambridge, OH 4 - City of Cleveland (East), TX17 - City of Houston (East), and PA5 - Philadelphia County.
 In 2020 Chicago, Black residents remain concentrated in the poorest Geomarket but Gentrification also pushed Black residents to the South and Southwest Suburbs Geomarket.
@@
 This question is motivated by the @RN4982 recommendation that selective colleges that enroll students from the regional and national market segment should target affluent Geomarkets, while community colleges and non-selective 4-year colleges should focus on middle- and working-class Geomarkets.
-Findings across metropolitan cases demonstrate that if student list purchases exclude working- and middle-class Geomarkets - such as IL11 -- City of Chicago, TX19 -- City of Dallas, TX20 -- City of Fort Worth, CA19 -- East Los Angeles, and CA21 -- South and South Central Los Angeles -- the resulting student list purchases disproportionately exclude Black, and Hispanic students who meet the requisite academic achievement criteria.
+Findings across metropolitan cases demonstrate that if student list purchases exclude working- and middle-class Geomarkets -- such as IL11 - City of Chicago, TX19 - City of Dallas, TX20 - City of Fort Worth, CA19 - East Los Angeles, and CA21 - South and South Central Los Angeles -- the resulting student list purchases would disproportionately exclude Black and Hispanic students who meet the requisite academic achievement criteria.
 Although these purchases would still capture some Black, Hispanic, and Asian prospects, these prospects are much less likely to be first-generation students.
@@
 The status attainment model argues that college destination is a function of parental education and occupation [@RN1273; @RN1271; @RN1272].
-We conceive of the the Market Segment Model is the status attainment model applied to geodemographic market research, akin to the rise of credit-scoring in UK retail banking [@RN4880].
-@RN5033 finds that the educational achievement of Asian American students whose parents are immigrants tend to have high educational achievement even if their parents do not, while the educational achievement of later-generation Asian Americans conforms to status attainment theory.
+We conceive of the Market Segment Model as the status attainment model applied to geodemographic market research, akin to the rise of credit-scoring in UK retail banking [@RN4880].
+@RN5033 finds that Asian American students whose parents are immigrants tend to have high educational achievement even if their parents do not, while the educational achievement of later-generation Asian Americans conforms to status attainment theory.
 In our analyses of prospects included in student list purchases, Asian American prospects from poor Geomarkets tended to be first-generation students while Asian American prospects from affluent Geomarkets tended to have parents with a BA.
@@
 Both the status attainment model and the cultural capital model are demand-side explanations for why education is a "social sieve" [@RN2377; @RN5046] that allows for a modicum of mobility while maintaining a larger flow of intergenerational class transmission [@RN861].
-@RN4722 -- who obtained parental income from federal income tax returns for every U.S.
+@RN4722 -- who obtained parental income data from federal income tax returns for every U.S.
 college -- show that the disproportionate enrollment share of high income families at selective private colleges -- and even at most public research universities -- is staggering.
@@
 @RN3519 and @RN4321 describe a tacit arrangement between high school guidance counselors on the demand side and college admissions officers on the supply side.
-Counselors at well-resourced -- especially private -- schools, are motivated to give their students a competitive advantage in admissions.
+Counselors at well-resourced -- especially private -- schools are motivated to give their students a competitive advantage in admissions.
 Meanwhile, college admissions officers are motivated to enroll students who can afford full tuition price and are likely to donate in the future.
@@
 
-Existing explanations of how students are sorted into colleges miss an important supply-side mechanism, third-party vendors that sort students on behalf of colleges.
+Existing explanations of how students are sorted into colleges miss an important supply-side mechanism -- third-party vendors that sort students on behalf of colleges.
 @RN4982 created the Market Segment Model and Geomarkets based on a snapshot of student demand from 1980.
-@RN4982 argues that demand for higher education is correlated with class, ignoring the historical structural barriers that produced class- and race-based inequality in 1980 student demand.
+@RN4982 argue that demand for higher education is correlated with class, ignoring the historical structural barriers that produced class- and race-based inequality in 1980 student demand.
 In itself, the Market Segment Model is a social science depiction.
@@
 Technologies that target customers based on geography inevitably leverage racial and class segregation [@RN4775; @RN4849; @RN4975].
-Furthermore, technologies that use the logic of predictive analytics -- identify correlates in past cases to make predictions for future cases -- are prone to "pernicious feedback loops" [@RN4849].
+Furthermore, technologies that use the logic of predictive analytics -- identifying correlates in past cases to make predictions for future cases -- are prone to "pernicious feedback loops" [@RN4849].
 The structural barriers (e.g., segregation, slavery, Jim Crow) that caused historic place-based inequality in student demand are amplified because EPS recommends that colleges focus recruiting resources on localities that already have high student demand for peer colleges.
 Geomarkets were subsequently incorporated into the Student Search Service product.
-We show that that excluding low-income, non-White Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-White students with strong academic achievement.
+We show that excluding low-income, non-White Geomarkets from student list purchases results in the disproportionate exclusion of first-generation, non-White students with strong academic achievement.
 
@@
 @RN527 \[p.
-148\] define organizational fields as, "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
+148\] define organizational fields as "those organizations that, in the aggregate, constitute a recognized area of institutional life: key suppliers, resource and product consumers, regulatory agencies, and other organizations that produce similar services or products" and that "the virtue of this unit of analysis is that it directs our attention\ldots to the totality of relevant actors." Sociological literatures salient to college access devote substantive attention to schools, colleges, families, and communities.
 College access scholarship from economics often analyzes the effects of district, state, and federal policies [e.g., @RN5043].
@@
 
-Second, technological advances creates new means of identifying and serving prospects.
-College Board and ACT leveraged data on test-takers to enjoy a near duopoly in market for student list data, one that lasted for several decades.
+Second, technological advances create new means of identifying and serving prospects.
+College Board and ACT leveraged data on test-takers to enjoy a near duopoly in the market for student list data, one that lasted for several decades.
 This business can be described as list-based lead generation based on the direct-mail model [@singer1988].
@@
 The most widely-known product is Naviance, which claims to be used by more than 10 million K-12 students and by 40% of US high schools [@powerschool].
-*Naviance* college planning software feeds into the *Intersect* recruiting platform -- an example of behavioral-based targeting -- which allows colleges to target Naviance users while they are on the platform.
-
-The third dynamic in the EM industry has been the growth of test-optional admissions movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT.
+Naviance college planning software feeds into the Intersect recruiting platform -- an example of behavioral-based targeting -- which allows colleges to target Naviance users while they are on the platform.
+
+The third dynamic in the EM industry has been the growth of the test-optional admissions movement, which reduces both the number of paying test-takers and the coverage of student list products sold by College Board and ACT.
 ACT responded to uncertainty in their core testing business by attempting to become an edtech company [@blumenstyk_2019], failing [@molnar_2020a], and being acquired by Nexus Capital in 2024 [@knox_2024].
@@
 [^22]: The Enroll360 press release states, "We spent the last couple of years creating a connected recruitment ecosystem that allows enrollment leaders to keep pace with students pursuing increasingly digital journeys to college.
-This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals [@koppenheffer_2021]."
+This work led us to join forces with several leading companies: Cappex, Intersect, Wisr, and YouVisit\ldots Individually, each solution can solve important challenges at various stages of the enrollment funnel\ldots By bringing these capabilities together, our vision is to reinvent how enrollment leaders reach their goals" [@koppenheffer_2021].
 
@@
 Colleges that are interested in recruiting these prospects must buy an Enroll360 subscription.
-Like College Board's *Student Search Service,* contemporary software-as-a-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
-For example, reporting @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
+Like College Board's Student Search Service, contemporary software-as-a-service recruiting products incorporate search filters that enable colleges to choose which prospects they recruit.
+For example, @markup_naviance shows how the Intersect recruiting platform allows colleges to send paid advertisements to Naviance users.
 Intersect customers (colleges) control which Naviance users will receive recruiting messages by filtering on filters such as "academic ability," intended majors, and whether students used Naviance to "research competitor institutions" [@markup_naviance].
-A University of Utah procurement document states the Intersect subscription is necessary because "there is unique group of prospective students who are only in the PowerSchool Naviance platform" [@utah_procurement].
+A University of Utah procurement document states the Intersect subscription is necessary because "there is a unique group of prospective students who are only in the PowerSchool Naviance platform" [@utah_procurement].
 
@@
 The ACT/Encoura "Enrollment Predictor" filter enables colleges to filter prospects based on the predicted probability of enrollment, which is created based on analyzing the enrollment decisions of previous cohorts of students [@schmidt_2019].
-College Board Geodemographic Segment filters categorizes each Census Tract into one of about two dozen clusters based on the college-going behavior of previous cohorts [@RN4565] and is based on the Claritas/PRIZM market segmentation system that categorized zip codes into groups useful for merchants [@RN4957].
-An investor-facing @vista_eab_ml video promoting EAB's "Pipeline Analytics" product states, "we are using artificial intelligence and machine learning to help schools build list of students that may be more more likely to enroll and persist in their university." The product "identifies students that would be a good match for a school based on their similarity to students who have succeeded there before and then provides them \[the students\] information about those institutions so that they choose to apply and ultimately enroll." Each of these products are based on the underlying logic of correlations and homophily discussed by @RN4975 and applied by @RN4982.
+College Board Geodemographic Segment filters categorize each census tract into one of about two dozen clusters based on the college-going behavior of previous cohorts [@RN4565] and is based on the Claritas/PRIZM market segmentation system that categorized zip codes into groups useful for merchants [@RN4957].
+An investor-facing @vista_eab_ml video promoting EAB's "Pipeline Analytics" product states, "we are using artificial intelligence and machine learning to help schools build lists of students that may be more likely to enroll and persist in their university." The product "identifies students that would be a good match for a school based on their similarity to students who have succeeded there before and then provides them \[the students\] information about those institutions so that they choose to apply and ultimately enroll." Each of these products is based on the underlying logic of correlations and homophily discussed by @RN4975 and applied by @RN4982.
 These products identify correlations in a historic snapshot of students and apply those relationships to make recommendations about which prospects to target in the present.
@@
 To the extent that colleges must buy expensive software to recruit prospects contained in an associated database, this expenditure will ultimately be passed on to students in the form of higher prices or reduced expenditure on education.
-Unfettered horizontal and vertical acquisitions in the EM industry causes the EM industry to become more concentrated.
+Unfettered horizontal and vertical acquisitions in the EM industry cause the EM industry to become more concentrated.
 Smaller consultancies that do not control vertical inputs may be unable to compete with companies that have acquired proprietary pools of prospects.
@@
 We argue that sociologists and adjacent scholars should expand the domain of the education research knowledge infrastructure to include third-party vendors in enrollment management and edtech more broadly.
-This is a challenging tasks.
+This is a challenging task.
 Student unit record data owned by district, state, and federal agencies is simultaneously granular (can evaluate a particular policy) and macro (can describe broad trends in the education sector).
@@
 
-Although case studies often relies on interview and ethnographic data, another approach is to issue records requests to public entities that contract with third-party entities.
+Although case studies often rely on interview and ethnographic data, another approach is to issue records requests to public entities that contract with third-party entities.
 For example, @RN5039 analyze public university contracts with online program managers (OPMs).
@@
 A growing set of data providers serve the information needs of the investment community by providing data on private firms and deals involving private firms.
-These databases tend to include three broad "types" of data: (1) *firm-level data* (e.g., location, employees, financials); (2) *acquisitions*, encompassing who acquired whom and resulting ownership structure); and (3) *investments*, which include amounts/valuations/stake percentages of private equity investments.
+These databases tend to include three broad "types" of data: (1) *firm-level data* (e.g., location, employees, financials); (2) *acquisitions*, encompassing who acquired whom and resulting ownership structure; and (3) *investments*, which include amounts/valuations/stake percentages of private equity investments.
 Some providers provide all three types of data (e.g., [*S&P Capital IQ*](https://www.spglobal.com/market-intelligence/en), [*LSEG Workspace*](https://www.lseg.com/en)).
@@
 Interestingly, the database subscriptions operate similar to contemporary student list products; an interactive, user-facing platform is wrapped around several sources of proprietary data.
-University libraries provides full or partial access to some subscriptions, while others must be purchased.
+University libraries provide full or partial access to some subscriptions, while others must be purchased.
 <!--# POTENTIALLY CUT THIS BASED ON OJs THOUGHTS/WANTS ; ALSO ENDING WITH BUILDING A KNOWLEDGE INFASTRUCTURE IS A BIT OF A BUZZKILL FROM ENDING ON MARKET DYNAMICS LEADING TO COMPETITION PRICE STUDENT CHOICE; but maybe fine for soc outlet-->
@@
 
-```{r fig-chicago-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
+```{r fig-chicago-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group, Middle-Range SAT orders", out.width = "100%"}
 knitr::include_graphics('./../results/graphs/rq2/rq2_chicago_race_col_plot_392834_488035.png')
@@
 
-```{r fig-chicago-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+```{r fig-chicago-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group and First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
 
@@
 
-```{r fig-dallas-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
+```{r fig-dallas-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group, Middle-Range SAT orders", out.width = "100%"}
 knitr::include_graphics('./../results/graphs/rq2/rq2_dallas_race_col_plot_448427.png')
@@
 
-```{r fig-dallas-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+```{r fig-dallas-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group and First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
 
@@
 
-```{r fig-la-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
+```{r fig-la-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group, Middle-Range SAT orders", out.width = "100%"}
 knitr::include_graphics('./../results/graphs/rq2/rq2_los_angeles_race_col_plot_448420_546946.png')
@@
 
-```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
+```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group and First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
 knitr::include_graphics('./../results/graphs/rq2b/rq2b_los_angeles_order_448420_546946_race_by_firstgen.png')

### 2025-07-14 text/manuscript-full-KS.qmd
update appendix

@@
 The two primary outputs of the Market Segment Model are the (1) Market Segment Profile and (2) the Institutional Profile.
-Appendix A describes these outputs in more detail.
+The [Online Appendix](https://structuring-college-access.github.io/) describes these outputs in more detail.
 Both outputs are created separately for each Geomarket.
@@
 
-[^5]: For example, Appendix @tbl-market-segment-characteristics shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
-
-[^6]: Appendix @tbl-sample-institutional-profile reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
+[^5]: For example, the Market Segment Profile in the [Online Appendix](https://structuring-college-access.github.io/#/market-segment-profile) shows how many students in "CT3 -- Fairfield County" are defined as regional or national based on their SAT score-sending behavior.
+
+[^6]: The [Online Appendix](https://structuring-college-access.github.io/#/institutional-profile) reproduces a partial, simplified version of @RN4982 [Figure 2.3], the Institutional Profile of an anonymous college for students from Fairfield County, CT.
 This table shows that 69 Regional students and 109 National students from Fairfield County sent SAT scores to the college.
@@
 This homophily is presented as a naturally occurring phenomenon.
-Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer colleges, information that can be discerned from the Institutional Profiles (Appendix @tbl-sample-institutional-profile).
+Given these findings, @RN4982 recommend that colleges should target Geomarkets that contain a critical mass of students interested in peer colleges, information that can be discerned from the Institutional Profiles (see [Online Appendix](https://structuring-college-access.github.io/#/institutional-profile)).
 In itself, @RN4982's Market Segment Model is a social science depiction of student demand -- akin to the status attainment model -- that does not consider historic, structural inequalities that cause observed patterns.
@@
 On background, admissions officers and enrollment consultants told us it remains common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket." In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
-This can be seen in the Chicago-land recruiting territories of University of Chicago \[ADD TO APPENDIX\], of Lake Forest College, of Saint Mary's College \[ADD TO APPENDIX CRYSTAL\]
+This can be seen in the Chicago-land recruiting territories of University of Chicago, Lake Forest College, and Saint Mary's College.
+<!--\[ADD TO APPENDIX CRYSTAL\]-->
 
@@
 We, therefore, select a small number of metropolitan areas in order to provide sufficient analytic depth while situating each case within its specific historic and contemporary sociodemographic characteristics.
-The [Online Appendix](https://ozanj.github.io/cb_geo/#/results) presents results for a larger number of metropolitan areas.
+The [Online Appendix](https://structuring-college-access.github.io/#/results) presents results for a larger number of metropolitan areas.
 
@@
 Results are presented below for Chicago, Dallas, and Los Angeles.
-The [Online Appendix](https://ozanj.github.io/cb_geo/#/results) provides a more convenient means of viewing results.
+The [Online Appendix](https://structuring-college-access.github.io/#/results) provides a more convenient means of viewing results.
 The online appendix also contains interactive maps, additional figures and tables not contained in the main text, and results for additional metropolitan areas (e.g., Bay Area, Philadelphia).
@@
 
-The interactive maps in the [Online Appendix](https://ozanj.github.io/cb_geo/#/research-question-1-15) can show these shifts at the census tract-level.
+The interactive maps in the [Online Appendix](https://structuring-college-access.github.io/#/research-question-1-15) can show these shifts at the census tract-level.
 For example, in 1980 Black people were concentrated in southern tracts of the City of Chicago Geomarket, as described by scholarship on gentrification in Chicago [@ewing2024_nofuture; @snidal2022_exodus].
@@
 These contribution patterns by race/ethnicity demonstrate that student list orders targeting prospects in the Chicago metropolitan area using Geomarket filters that exclude the City of Chicago -- as observed in some orders collected via our public requests -- result in the loss of substantially disproportionate shares of Black and Hispanic students from the resulting prospect pools.
-The [Online Appendix](https://ozanj.github.io/cb_geo/#/results) shows that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters.
+The [Online Appendix](https://structuring-college-access.github.io/#/results) shows that this pattern holds across both lower-range (1020-1150) and higher-range (1310-1600) SAT score filters.
 In both cases, the City of Chicago contributes disproportionately larger shares of Black and Hispanic prospects compared to all other Geomarkets.
@@
 
-Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles (11 Geomarkets) -- results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), results for which are shown in the [Online Appendix](https://ozanj.github.io/cb_geo/#/results).
+Due to the large number of Geomarkets in urban Southern California, we distinguish Los Angeles (11 Geomarkets) -- results shown here -- from Orange County (four Geomarkets) and San Diego (three Geomarkets), results for which are shown in the [Online Appendix](https://structuring-college-access.github.io/#/results).
 The 11 Geomarkets within the Los Angeles metropolitan area are illustrated in @fig-la-maps by total population across the three census periods.
@@
 The Market Segment Model recommends that colleges should identify their core student market segment (i.e., local, in-state, regional, or national), identify Geomarkets that contain large numbers of households from this market segment, and then target high schools and communities within these Geomarkets.
-This information is contained in the Market Segment Profile (Appendix A), a standard output in EPS software.
+This information is contained in the Market Segment Profile ([Online Appendix](https://structuring-college-access.github.io/#/market-segment-profile)), a standard output in EPS software.
 When considering new Geomarkets to target, EPS recommends colleges utilize the logic of homophily by identifying Geomarkets that are popular with peer colleges.
-This information is contained in the Institutional Profile (Appendix A).
+This information is contained in the Institutional Profile ([Online Appendix](https://structuring-college-access.github.io/#/institutional-profile)).
 
@@
 ```
-
-\clearpage
-
-\newpage
-
-# Appendix A
-
-```{r market-segment-characteristics, echo = F}
-#| label: tbl-market-segment-characteristics
-#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'
-
-market_segment <- data.frame(
-  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
-  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
-  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
-  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
-  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
-)
-
-kable(
-  x = market_segment,
-  col.names = c('Characteristic','Local','In-state','Regional','National'),
-  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
-)
-```
-
-```{r sample-institutional-profile, echo = F}
-#| label: tbl-sample-institutional-profile
-#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'
-
-sample_inst <- data.frame(
-  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
-  local = c(1,0.2,0,0.0,0,0.0),
-  in_state = c(58,4.8,25,8.4,2,3.0),
-  regional= c(69,4.1,35,7.0,5,5.8),
-  national = c(109,2.9,61,5.3,3,0.8),
-  total = c(237,3.3,121,5.9,10,1.8)
-)
-
-kable(
-  x = sample_inst,
-  col.names = c('','Local','In-state','Regional','National','Total'),
-  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
-)
-
-```

### 2025-07-14 text/manuscript-full-KS.qmd
update figure + landscape page

@@
 \clearpage
+\pagenumbering{gobble}
 ```
@@
 
-```{r fig-chicago-recruiting, echo = FALSE, fig.align = 'center', fig.cap = "The College Board Enrollment Planning Service Chicago Geomarkets (left) and University of Chicago Recruiting Territories (right)", out.width = "85%"}
+::: {.landscape}
+
+```{r fig-chicago-recruiting, echo = FALSE, fig.align = 'center', fig.cap = "The College Board Enrollment Planning Service Chicago Geomarkets (left) and University of Chicago Recruiting Territories (right)", out.width = "100%"}
 knitr::include_graphics('./../assets/images/chicago_recruiting.png')
@@
 ```
+
+:::
 
@@
 
-```{r fig-geomarket-methods, echo=FALSE,out.width="49%", fig.cap="caption",fig.align='center',fig.cap = "Chicago, College Board and Recreated Geomarkets"}
+::: {.landscape}
+
+```{r fig-geomarket-methods, echo=FALSE,out.width="100%", fig.cap="caption",fig.align='center',fig.cap = "Chicago, College Board and Recreated Geomarkets"}
 knitr::include_graphics('./../assets/images/eps/chicago_geos.png')
@@
 
+::: {.landscape}
+
 ```{r fig-dallas-maps, echo = FALSE, fig.align = 'center', fig.cap = "Dallas-Forth Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
@@
 
+::: {.landscape}
+
 ```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarkets, Total Population 1980-2020", out.width = "100%"}

### 2025-07-15 text/manuscript-full-KS.qmd
update references

@@
 bibliography: ./../assets/bib/cb_geomarket_bib.bib
-csl: ./../assets/bib/apa.csl
+csl: ./../assets/bib/chicago-author-date.csl
 format:

### 2025-07-22 text/manuscript-full-KS.qmd
last to dos

@@
 On background, admissions officers and enrollment consultants told us it remains common parlance to hear an admissions officer say something like, "I recruit 'PA 2'," which refers to the "Chester County, PA Geomarket." In states a college recruits heavily from, admissions officers are often assigned specific Geomarkets as their territory.
-This can be seen in the Chicago-land recruiting territories of University of Chicago, Lake Forest College, and Saint Mary's College.
+This can be seen in the Chicago-land recruiting territories of University of Chicago in @fig-chicago-recruiting.
 <!--\[ADD TO APPENDIX CRYSTAL\]-->
@@
 
-The primary data source for our second research question is student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data.
-This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 public universities located in California, Illinois, Texas, and Arizona.
+The primary data source for our second research question is student lists purchased from College Board.
+Originally collected from a larger project sample (Author, BLINDED), the data for this study's sample includes a total of 414 orders associated with 2,549,085 prospects made by 11 public universities located in California, Illinois, Texas, and Arizona.
 Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase.
@@
 
-::: {.landscape}
-
+::: landscape
 ```{r fig-chicago-recruiting, echo = FALSE, fig.align = 'center', fig.cap = "The College Board Enrollment Planning Service Chicago Geomarkets (left) and University of Chicago Recruiting Territories (right)", out.width = "100%"}
@@
 ```
-
 :::
@@
 
-::: {.landscape}
-
+::: landscape
 ```{r fig-geomarket-methods, echo=FALSE,out.width="100%", fig.cap="caption",fig.align='center',fig.cap = "Chicago, College Board and Recreated Geomarkets"}
@@
 
-::: {.landscape}
-
+::: landscape
 ```{r fig-dallas-maps, echo = FALSE, fig.align = 'center', fig.cap = "Dallas-Forth Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
@@
 
-::: {.landscape}
-
+::: landscape
 ```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarkets, Total Population 1980-2020", out.width = "100%"}

### 2025-07-22 scripts/create_stu_list.R

@@ -63,7 +63,7 @@ acs_zip <- acs_income_zip %>%
 ################### FINAL SAMPLE FOR EMPIRICAL REPORT
 
 #remove extra dataframes
-rm(lists_orders_zip_df)
+#rm(lists_orders_zip_df)
 
 #remove MN universities from ordersdf
 orders_df %>% count(univ_id, univ_name)
@@ -121,6 +121,28 @@ lists_orders_zip_hs_df <- lists_orders_zip_hs_df %>% mutate(
                        univ_id=="228723" , "research", "regional"))
 
 lists_orders_zip_hs_df %>% count(univ_name, univ_type)
+
+
+################### FINAL SAMPLE FOR AJS MANUSCRIPT
+
+# keep only orders that have accompanying students lists 
+
+# How many total orders did we get across all universities?
+orders_df %>% count(univ_name) # 14 universities (+2 MN univs removed above)
+orders_df %>% summarise(n=n_distinct(order_num)) # 835 total lists from 14 universities
+orders_df %>% summarise(sum(num_students, na.rm = TRUE))#total prospects from order sum 4669973
+
+# How many total lists we get across all universities?
+lists_orders_zip_hs_df %>% count(univ_name) # total lists from 13 universities (-NAU only gave us orders)
+lists_orders_zip_hs_df %>% summarise(n=n_distinct(ord_num)) #596 lists
+lists_orders_zip_hs_df %>% nrow() #3665455 prospects
+
+#How many orders + accompanying lists do we have across universities?
+final_sample <- subset(lists_orders_zip_hs_df, ord_num %in% orders_df$order_num)
+final_sample %>% count(univ_name) # from 11 universities 
+final_sample %>% summarise(n=n_distinct(ord_num)) #414 lists
+final_sample %>% nrow() #2,549,085
+
 
 
 ################### CREATE FILTER DUMMIES   

### 2025-07-22 text/manuscript-full-KS.qmd
add College Board 2023 PDF

@@
 
-We assigned each census tract to a College Board Geomarket, which we constructed through manual digitization of maps published by The College Board in 2023 marketing materials promoting their Enrollment Planning Service (EPS) software products.
+We assigned each census tract to a College Board Geomarket, which we constructed through manual digitization of maps published by The College Board in 2023 marketing materials promoting their Enrollment Planning Service (EPS) software products [@collegeboard_connections_geographies].
 Specifically, we georeferenced the College Board’s Geomarket maps by overlaying PDF images onto spatial coordinates aligned with census tract boundaries.

### 2025-07-22 text/manuscript-full-KS.qmd
update rq2 data sentence

@@
 
-In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.[^19] For each prospect, we know their home zip code and, therefore, which Geomarket they belong to.
+In particular, we answer the second research question by analyzing particular student list purchases that did *not* utilize the Geomarket search filter and purchased prospects from all Geomarkets in a particular metropolitan area.[^19] For each prospect, we know their high school, and the exact longitude and latitude of their high school, which we use to assign high school students to Geomarkets.
 Our analyses then show which students would have been included or excluded from the student list purchase had the purchase filtered on particular Geomarkets.

### 2025-07-22 text/manuscript-full-KS.qmd
blind authors

@@
 -->On the supply-side, the credentialism literature recognizes that colleges have a financial incentive for educational credentials to be required to claim advantageous social and economic positions [@RN1297], and this race for credentials benefits affluent households [e.g., @RN865; @RN861].
-Scholarship on enrollment management offers an agentic depiction of college behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758; @cottom2017lower; @RN4324].
+Scholarship on enrollment management offers an agentic depiction of college behavior, showing that colleges expend considerable resources recruiting desirable prospects [@RN3519; @RN4758_blinded; @cottom2017lower; @RN4324].
 
@@
 Student lists contain the contact information of prospective students and have been the primary source of "lead generation" in U.S.
-higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz; @RN4728].
+higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz_blinded; @RN4728].
 <!-- @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists.
@@
 
-Scholarship on enrollment management has focused on the behavior of colleges [e.g., @cottom2017lower; @RN4758; @RN4759; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
+Scholarship on enrollment management has focused on the behavior of colleges [e.g., @cottom2017lower; @RN4758_blinded; @RN4759_blinded; @RN3519], contributing to the belief that recruiting is a function of individual colleges.
 However, scholarship has failed to investigate the broader enrollment management industry -- consisting of third-party vendors, their products, their consulting solutions -- as a set of mechanisms that structure college access.
@@
 Second, how does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets?
-This research question is motivated by @RN4982, which recommend that selective colleges target affluent Geomarkets, and by prior research on enrollment management behavior, which finds that selective colleges focus recruiting visits on high schools in affluent, predominantly White communities [@RN3519; @RN4758].
+This research question is motivated by @RN4982, which recommend that selective colleges target affluent Geomarkets, and by prior research on enrollment management behavior, which finds that selective colleges focus recruiting visits on high schools in affluent, predominantly White communities [@RN3519; @RN4758_blinded].
 We analyze this research question using data from actual student list purchases, which we collected using public records requests.
@@
 These interventions seek to increase the probability of "conversion" across stages [@RN4322].
-At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz].
+At the top of the enrollment funnel, purchasing student lists is the primary means of converting prospects to leads [@list_biz_blinded].
 Purchased leads are served emails, brochures, and targeted social media designed to solicit inquiries and applications [@RN4895].
@@
 A growing literature analyzes the earlier "recruiting" stages of identifying leads, soliciting inquiries, and soliciting applications.
-@RN4758 conceptualize recruiting behavior as an indicator of college enrollment priorities.
+@RN4758_blinded conceptualize recruiting behavior as an indicator of college enrollment priorities.
 Ethnographies by @RN3519 and @RN4407 identify connections between private school guidance counselors and college admissions officers as a mechanism for social reproduction.
 In turn, recruiting visits to high schools -- and subsequent phone calls -- are a means of maintaining ties with guidance counselors at feeder schools and establishing relationships with prospective students [@RN3519; @RN4895].
-Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly White public schools [@RN4758; @RN4759; @RN5023].
+Quantitative case-studies of off-campus recruiting visits by public research universities and by selective private universities reveal a preference for visiting private schools and affluent, predominantly White public schools [@RN4758_blinded; @RN4759_blinded; @RN5023_blinded].
 From the student perspective, @RN4324 finds that underrepresented students were drawn to colleges that made them feel wanted, often attending institutions with lower graduation rates and requiring larger loans than other college options.
@@
 
-Drawing from scholarship on organizational theory, EM processes require colleges to make many "make or buy" [@RN4738; @RN980] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz].
+Drawing from scholarship on organizational theory, EM processes require colleges to make many "make or buy" [@RN4738; @RN980] decisions about whether to perform a given task in-house or outsource it to a third-party vendor [@list_biz_blinded].
 The most widely-known class of EM vendor is EM consulting firms, which provide advice and implementation in the areas of marketing, recruiting, pricing and financial aid, and student success [@RN5027].
@@
 
-[^2]: In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz; @list_empirics].
-
-While the contemporary EM industry is characterized by software-as-a-service products sold to colleges by private equity backed firms [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
+[^2]: In our data collection, attempting to obtain data about student list purchases from all public universities in four states, at least half of these universities outsourced student list purchases to an EM consultancy [@list_biz_blinded; @list_empirics_blinded].
+
+While the contemporary EM industry is characterized by software-as-a-service products sold to colleges by private equity backed firms [@list_biz_blinded], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products.
 Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit.
@@
 A student list contains the contact information of prospective students who meet the search filter criteria (e.g., test score, GPA) specified by the university.
-Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the "enrollment funnel." @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz].
+Student lists are the fundamental input for undergraduate recruiting campaigns because purchased names -- alongside prospects who reach out on their own -- constitute the set of prospects who receive subsequent recruiting interventions (e.g., mail, email) designed to push them toward the application and enrollment stages of the "enrollment funnel." @RN4895 reports that 86% of public colleges and 87% of private colleges purchase student lists.[^11] Historically, dominant list vendors are the College Board and ACT, which derived student list data from test-takers, but the test-optional movement, advances in technology, and surging private equity investments have contributed to new sources of student list data [@list_biz_blinded].
 
@@
 
-However, @RN5012 argue that student list products may exacerbate racial inequality in college access.
+However, @RN5012_blinded argue that student list products may exacerbate racial inequality in college access.
 The authors conceptualize student list products as "selection devices" [@RN4778] that enable colleges to select which prospective students they target by incorporating search filters (e.g., high school graduating class, state).
 @RN4786 defines "racialized inputs" as ostensibly race-neutral inputs that are systematically correlated with race because marginalized racial/ethnic groups have historically been excluded from the input.
-@RN5012 argue that several frequently utilized student list filters (e.g., zip code, AP test score, SAT score) meet the criteria of racialized inputs.
-Using a national sample of high school students and using data from actual student lists purchased by public universities, @RN5012 show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
+@RN5012_blinded argue that several frequently utilized student list filters (e.g., zip code, AP test score, SAT score) meet the criteria of racialized inputs.
+Using a national sample of high school students and using data from actual student lists purchased by public universities, @RN5012_blinded show that racialized search filters have a strong negative relationship with the selection of Black and Latinx prospects.
 
@@
 Some geographic filters are based on known geographic borders (e.g., zip code, county, CBSA, state).
-College Board also created geographic borders that subsequently became filters in the Student Search Service, such as Geomarkets and Geodemographic Segment filters.[^14] In their sample, @RN5012 found that 11% of student lists purchased by public research universities filtered on Geomarkets.
+College Board also created geographic borders that subsequently became filters in the Student Search Service, such as Geomarkets and Geodemographic Segment filters.[^14] In their sample, @RN5012_blinded found that 11% of student lists purchased by public research universities filtered on Geomarkets.
 Drawing from @RN4786, we argue that geomarkets satisfy the two criteria of racialized inputs.
@@
 Second, Geomarkets may be systematically correlated with race.
-For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda County, excluding Oakland." <!--# A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.--> We extend @RN5012 by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
+For example, Geomarket "CA07" is "City of Oakland" and "CA08" is "Alameda County, excluding Oakland." <!--# A university might purchase prospect contact information by filtering on CA08 in combination with additional filters, such as year of high school graduation and SAT test score.--> We extend @RN5012_blinded by examining how the racial composition of actual student list purchases changes if universities filter on particular Geomarkets.
 <!--# This manuscript focuses on Geomarkets.
@@
 We suggest that homophily is the operating logic most consistent with college recruiting behavior observed in the wild.
-For example, @RN5023 find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
+For example, @RN5023_blinded find great overlap in the sets of private schools visited by the University of South Carolina and the University of Alabama.
 <!--# The study also finds great overlap in the schools visited by Villanova University and the University of Notre Dame, with a disproportionate focus on visits to Catholic schools.
@@
 The credentialism literature recognizes that colleges have a financial incentive for educational credentials to determine the competition for socially and economically advantageous positions [@RN865; @RN861; @RN880; @RN869] and this credentials arms race favors affluent households.
-Scholarship on enrollment management describes which students colleges want and how colleges go about attracting these students at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758; @RN3544; @RN3547; @cottom2017lower].
+Scholarship on enrollment management describes which students colleges want and how colleges go about attracting these students at different stages of the enrollment funnel [@RN3519; @RN4324; @RN3522; @RN4407; @RN4758_blinded; @RN3544; @RN3547; @cottom2017lower].
 @RN3519 and @RN4321 describe a tacit arrangement between high school guidance counselors on the demand side and college admissions officers on the supply side.
@@
 The lacrosse team needs players.
-Quantitative analyses of visits from colleges to high schools are consistent with @RN3519; selective private colleges and public research universities devote most of their recruiting resources on courting students from privileged schools and communities [@RN5023; @RN4758].
+Quantitative analyses of visits from colleges to high schools are consistent with @RN3519; selective private colleges and public research universities devote most of their recruiting resources on courting students from privileged schools and communities [@RN5023_blinded; @RN4758_blinded].
 However, these analyses conceive of recruiting as something that is done solely by individual colleges.
@@
 
-@list_biz describes four key dynamics in the enrollment management (EM) industry, with a focus on the market for student list data.
+@list_biz_blinded describes four key dynamics in the enrollment management (EM) industry, with a focus on the market for student list data.
 First, EM consulting firms are central to the creation and implementation of recruiting campaigns [@RN5027].
@@
 ACT responded to uncertainty in their core testing business by attempting to become an edtech company [@blumenstyk_2019], failing [@molnar_2020a], and being acquired by Nexus Capital in 2024 [@knox_2024].
-Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters [@RN4565] -- developed free college search engines, and entered the enrollment management consulting market more aggressively [@list_biz].
+Over the last two decades, the testing companies developed new search filters that explicitly incorporated predictive analytics -- for example College Board's Geodemographic Segment filters [@RN4565] -- developed free college search engines, and entered the enrollment management consulting market more aggressively [@list_biz_blinded].
 College Board has weathered the test-optional movement thanks to robust revenues from the Advanced Placement product, but the decline in PSAT/SAT/PreACT/ACT test-takers has undermined their oligopoly in the market for student list data.
@@
 The 2010s were a period of acquisitions and concentration.
-Drawing from resource dependence theory, @list_biz describe the proliferation of horizontal acquisitions to grow market share and vertical acquisitions designed to make customers more dependent on particular recruiting products.
+Drawing from resource dependence theory, @list_biz_blinded describe the proliferation of horizontal acquisitions to grow market share and vertical acquisitions designed to make customers more dependent on particular recruiting products.
 After regulatory scrutiny pushed private-equity interests out of the for-profit college market [@RN5022], private equity investors found value in acquiring firms that provided services to direct providers.

### 2025-07-23 text/manuscript-full-KS.qmd
update map figures

@@
 \clearpage
-\pagenumbering{gobble}
 ```
@@
 
-::: landscape
 ```{r fig-chicago-recruiting, echo = FALSE, fig.align = 'center', fig.cap = "The College Board Enrollment Planning Service Chicago Geomarkets (left) and University of Chicago Recruiting Territories (right)", out.width = "100%"}
@@
 ```
-:::
 
@@
 
-::: landscape
 ```{r fig-geomarket-methods, echo=FALSE,out.width="100%", fig.cap="caption",fig.align='center',fig.cap = "Chicago, College Board and Recreated Geomarkets"}
@@
 
-::: landscape
 ```{r fig-dallas-maps, echo = FALSE, fig.align = 'center', fig.cap = "Dallas-Forth Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
@@
 
-::: landscape
 ```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarkets, Total Population 1980-2020", out.width = "100%"}

### 2025-07-23 text/manuscript-full-KS.qmd
add cover page

@@
 title: "Structuring College Access: The Market Segment Model and College Board Geomarkets"
-author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)", "Crystal Han (Stanford University)"]
+# author: ["Ozan Jaquette (UCLA)", "Karina Salazar (University of Arizona)", "Crystal Han (Stanford University)"]
 bibliography: ./../assets/bib/cb_geomarket_bib.bib

### 2025-07-25 text/manuscript-full-KS.qmd
add figure notes

@@
 \endgroup
-\doublespacing
-\clearpage
+\clearpage
+\singlespacing
 ```
@@
 #![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+```{=tex}
+\clearpage
 ```
@@
 
+**Note**: Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-chicago-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Chicago Area Geomarkets, 1980-2020", out.width = "100%"}
@@
 
+**Note**: Household income measured using 2024 CPI
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-chicago-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Chicago Area", out.width = "100%"}
@@
 
+**Note**: Excludes students with missing values for race.
+Order number 392833 (ordered 7/10/2018): HS class 2019 OR 2020; state = IL; SAT 1020 - 1150; GPA B- to A+.
+Order number 487984 (ordered 7/19/2019): HS class 2020 OR 2021; state = IL; SAT 1020 - 1150; GPA B- to A+.
+Order number 392834 (ordered 7/10/2018): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300.
+Order number 488035 (ordered 7/19/2019): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300; GPA B- to A+.
+Order number 392835 (ordered 7/9/2018): HS class 2019 OR 2020; state = IL; SAT 1310 - 1600; GPA B- to A+.
+Order number 488053 (ordered 7/19/2019): HS class 2020 OR 2021; state = IL; SAT 1310 - 1600; GPA B- to A+.
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-chicago-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group, Middle-Range SAT orders", out.width = "100%"}
@@
 #![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+**Note**: Excludes students with missing values for race.
+Order number 392834 (ordered 7/10/2018): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300.
+Order number 488035 (ordered 7/19/2019): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300; GPA B- to A+.
+
+```{=tex}
+\clearpage
 ```
@@
 
+**Note**: Excludes students with missing values for firstgen.
+Order number 392834 (ordered 7/10/2018): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300.
+Order number 488035 (ordered 7/19/2019): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300; GPA B- to A+.
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-chicago-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Chicago Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group and First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
@@
 
-```{r fig-dallas-maps, echo = FALSE, fig.align = 'center', fig.cap = "Dallas-Forth Worth Geomarkets, Total Population 1980-2020", out.width = "100%"}
+**Note**: Excludes students who have missing values for race or parental education.
+Order number 392834 (ordered 7/10/2018): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300.
+Order number 488035 (ordered 7/19/2019): HS class 2019 OR 2020; state = IL; SAT 1160 - 1300; GPA B- to A+.
+
+```{=tex}
+\clearpage
+```
+
+```{r fig-dallas-maps, echo = FALSE, fig.align = 'center', fig.cap = "Dallas-Forth Worth Geomarkets, Total Population 1980-2020", out.width = "95%"}
 knitr::include_graphics('./../assets/images/dallas_maps.png')
@@
 
+**Note**: Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-dallas-rq1-ses, echo = FALSE, fig.align = 'center', fig.cap = "Socioeconomic Characteristics of Dallas Area Geomarkets, 1980-2020", out.width = "100%"}
@@
 #![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+**Note**: Household income measured using 2024 CPI
+
+```{=tex}
+\clearpage
 ```
@@
 
+**Note**: Excludes students with missing values for race.
+Order number 448922 (ordered 1/9/2019): HS class 2020; selected Geomarkets; PSAT 1070 - 1180.
+Order number 448427 (ordered 1/8/2019): HS class 2020; selected Geomarkets; PSAT 1190 - 1260.
+Order number 448440 (ordered 1/8/2019): HS class 2020; selected Geomarkets; PSAT 1270 - 1520.
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-dallas-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group, Middle-Range SAT orders", out.width = "100%"}
@@
 
+**Note**: Excludes students with missing values for race. Order number 448427 (ordered 1/8/2019): HS class 2020; selected Geomarkets; PSAT 1190 - 1260.
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-dallas-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
@@
 
+**Note**: Excludes students with missing values for firstgen. Order number 448427 (ordered 1/8/2019): HS class 2020; selected Geomarkets; PSAT 1190 - 1260.
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-dallas-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Dallas Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group and First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
@@
 
-```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarkets, Total Population 1980-2020", out.width = "100%"}
+**Note**: Excludes students who have missing values for race or parental education. Order number 448427 (ordered 1/8/2019): HS class 2020; selected Geomarkets; PSAT 1190 - 1260.
+
+```{=tex}
+\clearpage
+```
+
+```{r fig-la-maps, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarkets, Total Population 1980-2020", out.width = "95%"}
 knitr::include_graphics('./../assets/images/la-maps.png')
@@
 
+```{=tex}
+\clearpage
+```
+
 ```{r fig-la-rq1-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Los Angeles Area Geomarkets, 1980-2020", out.width = "100%"}
@@
 #![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+**Note**: Race/ethnicity categories not available in 1980 Census: Asian, non-Hispanic; Two+ races, non-Hispanic; NHPI, non-Hispanic; AIAN non-Hispanic
+
+```{=tex}
+\clearpage
 ```
@@
 
+**Note**: Household income measured using 2024 CPI
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-la-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Los Angeles Area", out.width = "100%"}
@@
 #![The enrollment funnel](assets/images/enrollment_funnel.png)
+```
+
+**Note**: Excludes students with missing values for race.
+Order number 448375 (ordered 1/8/2019): HS class 2020; State = CA; PSAT 1070 - 1180.
+Order number 546954 (ordered 1/6/2020): HS class 2021; State = CA; PSAT 1070 - 1180.
+Order number 448420 (ordered 1/8/2019): HS class 2020 OR 2021; State = CA; PSAT 1190 - 1260.
+Order number 546946 (ordered 1/6/2020): HS class 2021 OR 2022; State = CA; PSAT 1190 - 1260.
+Order number 448374 (ordered 1/8/2019): HS class 2020 OR 2021; State = CA; PSAT 1270 - 1520.
+Order number 546945 (ordered 1/6/2020): HS class 2021 OR 2022; State = CA; PSAT 1270 - 1520.
+
+```{=tex}
+\clearpage
 ```
@@
 
+**Note**: Excludes students with missing values for race.
+Order number 448420 (ordered 1/8/2019): HS class 2020 OR 2021; State = CA; PSAT 1190 - 1260.
+Order number 546946 (ordered 1/6/2020): HS class 2021 OR 2022; State = CA; PSAT 1190 - 1260.
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-la-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
@@
 
+**Note**: Excludes students with missing values for firstgen.
+Order number 448420 (ordered 1/8/2019): HS class 2020 OR 2021; State = CA; PSAT 1190 - 1260.
+Order number 546946 (ordered 1/6/2020): HS class 2021 OR 2022; State = CA; PSAT 1190 - 1260.
+
+```{=tex}
+\clearpage
+```
+
 ```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/Ethnic Group and First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
@@
 ```
+
+**Note**: Excludes students who have missing values for race or parental education.
+Order number 448420 (ordered 1/8/2019): HS class 2020 OR 2021; State = CA; PSAT 1190 - 1260.
+Order number 546946 (ordered 1/6/2020): HS class 2021 OR 2022; State = CA; PSAT 1190 - 1260.

### 2025-07-28 text/manuscript-full-KS.qmd
update abstract

@@
 
-*The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Model" (MSM).
+The Structure of College Choice (1983) created "Geomarkets" and the "Market Segment Model." Geomarkets carve metropolitan areas into smaller geographies that define local recruiting markets.
+The Market Segment Model predicts demand for a college in a Geomarket as a function of socioeconomic characteristics.
+Geomarkets became an input for College Board recruiting products, including their "student list" product, which provides colleges with the contact information of prospective students who meet desired search filter criteria.
+We draw from scholarship on quantification and address two questions: What is the socioeconomic and racial variation between Geomarkets?
+How does the socioeconomic and racial composition of included versus excluded prospects vary if student list purchases filter on Geomarkets?
+We answer RQ1 using Census data and RQ2 using data from actual student list purchases.
+Geomarket borders are highly correlated with race and class.
+Student list purchases that exclude low-income Geomarkets systematically overlook high-scoring non-White and first-generation students.
+
+<!-- *The Structure of College Choice* [@RN4982] created "Geomarkets" and the "Market Segment Model" (MSM).
 Geomarkets carve states and metropolitan areas into smaller geographic units to define local recruiting markets.
@@
 Moreover, student lists that exclude low-income Geomarkets systematically overlook non-White students, first-generation students, and especially first-generation non-White students.
+-->
 

### 2025-10-09 aefp/aefp_geomarket.qmd
added cerpp proposal

@@
 
+```{=html}
 <!--  

### 2025-10-09 cerpp/cerpp_geomarket.qmd
added cerpp proposal

--- 2025-01-15 soc_of_ed/soc_of_ed_geomarket.qmd
@@
 
+# ABSTRACT
+
+College Board (CB) "Geomarkets" carve metropolitan areas into smaller geographies.
+Geomarkets are a "search filter" in the CB "student list" product that enables colleges to acquire prospect contact information.
+If Geomarket borders are correlated with race and class, using Geomarkets to target prospects may amplify inequality.
+We address two questions: What is the socioeconomic and racial variation between Geomarkets?
+How does the socioeconomic and racial composition of included versus excluded prospects vary if student list purchases filter on Geomarkets?
+We answer RQ1 using Census data and RQ2 using data from actual student list purchases.
+
 # PRECIS
@@
 **BACKGROUND.** In 1983, Zemsky and Oedel authored *The Structure of College Choice*.
-Based on an analysis of 1980 SAT score-sending behavior and published by *The College Board*, @RN4982 developed the Market Segment Model and created "Geomarkets." The Market Segment Model predicts how student demand for a particular college varies across local geographic markets based on the characteristics of students in those local markets.
+Based on an analysis of 1980 SAT score-sending behavior and published by *The College Board* (CB), @RN4982 developed the Market Segment Model and created "Geomarkets." The Market Segment Model predicts how student demand for a particular college varies across local geographic markets based on the characteristics of students in those local markets.
 
@@
 14] for recruiting purposes.
-@fig-nyc-geomarkets shows New York City area Geomarkets.
-The Market Segment Model and Geomarkets are the basis for the College Board Enrollment Planning Service (EPS), founded in 1984 and still active today.
-EPS software recommends which Geomarkets a college should recruit from and which high schools within targeted Geomarkets a college should visit.[^1]
-
-[^1]: In 1995, 37% of 4-year public institutions and 49% of 4-year private institutions used EPS [@RN5021].
-
-Geomarkets were incorporated into the *College Board* "student list" product named Student Search Service.
+Figure 1 shows New York City area Geomarkets.
+Geomarkets were incorporated into the CB "student list" product named Student Search Service.
 Student lists have been the primary source of lead generation in U.S.
-higher education for over 50 years [@belkin2019-studata; @list_biz].[^2] Lists contain contact information for prospective students.
-The Student Search Service database consists of College Board test-takers.
-Colleges pay a fee for each prospect (e.g., \$0.50).
+higher education for over 50 years [@belkin2019-studata; @list_biz].
+Lists contain contact information for prospective students.
+The Student Search Service database consists of CB test-takers.
 Colleges control which prospect profiles they purchase by selecting search filters, such as high school graduation year, SAT score, AP score, state, etc.
-Geomarket filters enable colleges to target prospects living in particular Geomarkets.^[An analysis of 830 student lists purchased by 14 public universities found that 16% of purchases filtered on Geomarket [@RN5012].]
-
-[^2]: @RN4895 reported that 87% of private and 86% of public four-year institutions purchase lists.
+Geomarket filters enable colleges to target prospects living in particular Geomarkets.
 
@@
 The overarching thesis of the Market Segment Model is homophily, the idea that similar colleges compete with one another for similar students.
-@RN4982[p.
-42-45] states that selective colleges primarily draw from students in the "regional" and "national" segments and, therefore, their recruiting visits and student list purchases should focus on Geomarkets with large numbers of affluent, highly educated households.
+@RN4982 [p.
+42-45] states that selective colleges primarily draw from students in the "regional" and "national" segments and, therefore, their student list purchases should focus on Geomarkets with large numbers of affluent, highly educated households.
 
 When we use data on past correlations to make recommendations about the future, we amplify the effects of historic structural inequality [@RN4794].
-A snapshot of student demand in 1980, itself a consequence of historical structural inequality, became inscribed into the EPS and Student Search Service products used by colleges to plan and implement recruiting strategy.
+A snapshot of student demand in 1980, itself a consequence of historical structural inequality, became inscribed into the Student Search Service product.
 The result is a supply-side that reinforces structural inequalities observed on the demand-side.
@@
 
-**SIGNIFICANCE.** Scholarship on quantification in the sociology of education [e.g.,  @RN2356;@RN5005] has not analyzed third-party products that colleges use to classify and identify prospective students.
-More broadly, the concept of "organizational fields" includes all organizations that "constitute a recognized area of institutional life"  [@RN527, p.
-143].
-In the enrollment management industry, third-party vendors and their products structure college access in ways that funnel certain kinds of students to certain kinds of institutions.
-Higher education policy over the last 20 years -- dominated by economists -- has focused on regulating consumers (students) and direct providers (colleges).
-However, federal regulatory agencies (e.g., FTC, CFPB) are increasingly interested in regulating the entire organizational field because they know that third-party oligopolies can wield market power over consumers and providers [@philippon2019great].
-Therefore, scholarship from sociology should guide which problems and actors become the focus of future federal regulation.
+**SIGNIFICANCE.** Academic scholarship on enrollment management tends to focus on students (micro), policies (macro), and organizations (meso).
+Scholars have not analyzed third-party classification systems and products that colleges utilize to classify prospective students.
+CB Geomarkets, utilized in the CB Student Search Service, offers one such case.
 
@@
 Variables of interest include: race/ethnicity; mean and median household income; percent of households below poverty line; and BA attainment of people aged 25+.
-We obtained shapefiles for College Board Geomarkets from a 2012 *R-bloggers* post.
-We assigned census tracts to Geomarkets by implementing a partial spatial join.
+We obtained shapefiles for CB Geomarkets from a 2012 *R-bloggers* post.
+We assigned census tracts to Geomarkets.
 
 To answer RQ2, we issued public records requests about student lists purchased by 14 public universities from 2016 through 2020.
-For 414 College Board student list purchases, associated with about 2.6 million prospects, we received both (1) the order summary -- showing which search filters were utilized -- and (2) the de-identified prospect-level data, including race, ethnicity, zip code, and high school.
-@fig-asu-order-448922 shows the partial order summary, which we converted to tabular data, for a purchase by Arizona State University that filtered on the 2020 high school graduating class, PSAT scores of 1070 to 1180, and selected Geomarkets.
+For 414 CB student list purchases, associated with about 2.6 million prospects, we received both (1) the order summary -- showing which search filters were utilized -- and (2) the de-identified prospect-level data, including race, ethnicity, zip code, and high school.
+Figure 2 shows the partial order summary for a purchase that filtered on the 2020 high school graduating class, PSAT scores of 1070 to 1180, and selected Geomarkets.
 
@@
 **PRELIMINARY FINDINGS**.
-For RQ1, Table \ref{tab:chi_eps_table_summary} shows racial and socioeconomic characteristics for Geomarkets around Chicago.
-@fig-socal-hispanic is a screenshot of an interactive choropleth map of Los Angeles, showing Geomarket borders and the tract-level %Hispanic in 2020.
-@fig-socal-med-inc shows tract-level median household income around Los Angeles.
+For RQ1, Table 1 shows racial and socioeconomic characteristics for Geomarkets around Chicago.
+Figure 3 is a screenshot of an interactive choropleth map of Los Angeles, showing Geomarket borders and the tract-level %Hispanic in 2020.
+Figure 4 shows tract-level median household income around Los Angeles.
 
@@
 These prospects would have been excluded had the purchase filtered on Geomarkets and not selected that Geomarket.
-@tbl-philly-sim-count, @tbl-philly-sim-col-pct, and @tbl-philly-sim-row-pct, respectively, show counts, column percentages, and row percentages for the Philadelphia metropolitan area, based on a purchase that filtered on students from the 2020 high school class who scored between 1070 and 1180 on the PSAT.
-@tbl-philly-sim-col-pct shows that "PA5--Philadelphia County" contains 15% of all purchased prospect profiles but 44% of Black prospects, indicating that opportunity for Black students is sensitive to the inclusion of this Geomarket.
-@fig-philly-black-prospects is a screenshot of a (very) preliminary interactive map for this purchase, with marker size indicating the number of purchased prospect profiles who attended that high school and separate markers by race/ethnicity.
+Table 2, Table 3, and Table 4, respectively, show counts, column percentages, and row percentages for the Philadelphia metropolitan area, based on a purchase that filtered on students from the 2020 high school class who scored between 1070 and 1180 on the PSAT.
+Table 3 shows that "PA5--Philadelphia County" contains 15% of all purchased prospect profiles but 44% of Black prospects, indicating that opportunity for Black students is sensitive to the inclusion of this Geomarket.
+Figure 5 is a screenshot of a (very) preliminary interactive map for this purchase, with marker size indicating the number of purchased prospect profiles who attended that high school and separate markers by race/ethnicity.
+
+**PRESENTATION PLAN**.
+I will provide a succinct background about CB Geomarkets and the Student Search Service product, but I will omit discussion of theory.
+I will show actual student list purchases that target the Chicago-land area but exclude the City of Chicago Geomarket.
+Next, I will show which prospects are excluded by this decision.
+Finally, I will engage participants in a discussion about contemporary third-party products that classify students on behalf of colleges.
+How can practitioners and policymakers effectively interrogate these products to avoid unintentional exclusion along racial and class dimensions?
 
@@
 :::
-
-\newpage
-\setcounter{page}{1}
-\renewcommand{\thepage}{A-\arabic{page}}
-
-# Appendix
-
-```{r socal_eps_table_summary, echo=FALSE, results='asis', eval = FALSE}
-# Generate the Los Angeles (socal) table with Bookdown label
-socal_table <- format_eps_table(
-  df = socal_eps_table, 
-  years = c(2020), 
-  metro_area = "Los Angeles", 
-  format = "latex",
-  label = "socal_eps_table_summary"  # This label matches the reference in your text
-)
-
-# Output the table directly
-cat(socal_table)
-```
-
-```{r philly_eps_table_summary, echo=FALSE, results='asis', eval = FALSE}
-# Wrap the table in the landscape environment
-cat("\\begin{landscape}\n")
-
-# Create the table for Philadelphia
-philly_table <- format_eps_table(
-  df = philly_eps_table, 
-  years = c(2000, 2020), 
-  metro_area = "Philadelphia metropolitan area", 
-  format = "latex",
-  label = "philly_eps_table_summary"  # This label matches the reference in your text  
-)
-
-# Output the tables directly
-cat(philly_table)
-cat("\n\\end{landscape}")
-```
-
-```{r chi_eps_table_summary, echo=FALSE, results='asis'}
-# Wrap the table in the landscape environment
-cat("\\begin{landscape}\n")
-
-# Create the table for Philadelphia
-chi_table <- format_eps_table(
-  df = chi_eps_table, 
-  years = c(2000, 2020), 
-  metro_area = "Chicago-land", 
-  format = "latex",
-  label = "chi_eps_table_summary"  # This label matches the reference in your text  
-)
-
-# Output the tables directly
-cat(chi_table)
-cat("\n\\end{landscape}")
-```
-
-\newpage
-
-```{r create_eps_sim_tables, echo = FALSE, results='asis', message = FALSE, warning = FALSE}
-# Load required packages and input data
-load(file = file.path('.', '..', 'results', 'tables', 'philly_asu_eps_lists.RData'))
-# str(philly_order448440_psat1270_1520)
-# str(philly_order448922_psat1070_1180)
-
-library(dplyr)
-library(knitr)
-library(kableExtra)
-library(scales)  # For adding commas to large numbers
-
-# Function to format three separate tables
-format_three_tables <- function(input_list) {
-  # Extract individual data frames from the input list
-  count_table <- input_list$count_table %>% 
-    rename(all = stu_all, race_known = stu_race_known, white = stu_white, asian = stu_asian, black = stu_black, hispanic = stu_hispanic, two_races = stu_tworaces) %>%
-    mutate(across(where(is.numeric), ~ if_else(. > 1000, comma(.), as.character(.))))  # Add commas to large numbers
-
-  # Modify race_known to be a column percentage of the race_known total
-  col_pct_table <- input_list$col_pct_table %>%
-    rename(all = stu_all, race_known = stu_race_known, white = col_pct_white, asian = col_pct_asian, black = col_pct_black, hispanic = col_pct_hispanic, two_races = col_pct_tworaces) %>%
-    select(-all) %>%
-    mutate(
-      race_known = (race_known / sum(race_known)) * 100,  # Convert race_known to a column percentage of the total
-      across(-eps_codename, ~ round(., 1))  # Round percentage values to one decimal point
-    )
-
-  # Set race_known to always be 100% in row_pct_table
-  row_pct_table <- input_list$row_pct_table %>%  
-    rename(all = stu_all, race_known = stu_race_known, white = row_pct_white, asian = row_pct_asian, black = row_pct_black, hispanic = row_pct_hispanic, two_races = row_pct_tworaces) %>%
-    select(-all) %>%
-    mutate(
-      race_known = 100.0,  # Set race_known to 100%
-      across(-eps_codename, ~ round(., 1))  # Round percentage values to one decimal point
-    )
-
-  # Format and display count table with caption
-  count_table_output <- count_table %>%
-    kbl(format = "latex", booktabs = TRUE, align = c("l", rep("r", ncol(count_table) - 1))) %>%
-    kable_styling(latex_options = c("striped", "hold_position"), position = "left", font_size = 10) %>%
-    row_spec(0, bold = TRUE, background = "#D3D3D3")
-
-  # Format and display column percentage table with caption
-  col_pct_table_output <- col_pct_table %>%
-    kbl(format = "latex", booktabs = TRUE, align = c("l", rep("r", ncol(col_pct_table) - 1))) %>%
-    kable_styling(latex_options = c("striped", "hold_position"), position = "left", font_size = 10) %>%
-    row_spec(0, bold = TRUE, background = "#D3D3D3")
-
-  # Format and display row percentage table with caption
-  row_pct_table_output <- row_pct_table %>%
-    kbl(format = "latex", booktabs = TRUE, align = c("l", rep("r", ncol(row_pct_table) - 1))) %>%
-    kable_styling(latex_options = c("striped", "hold_position"), position = "left", font_size = 10) %>%
-    row_spec(0, bold = TRUE, background = "#D3D3D3")
-
-  # Return formatted tables
-  list(count_table_output, col_pct_table_output, row_pct_table_output)
-}
-
-# Example usage
-formatted_tables <- format_three_tables(philly_order448922_psat1070_1180)
-
-# Print formatted tables with landscape mode in LaTeX
-#cat("\\begin{landscape}\n")
-#cat(as.character(formatted_tables[[1]]))
-#cat("\n\n")
-#cat(as.character(formatted_tables[[2]]))
-#cat("\n\n")
-#cat(as.character(formatted_tables[[3]]))
-#cat("\n\\end{landscape}")
-
-```
-
-\newpage
-
-```{r philly-sim-count, echo = FALSE, results='asis'}
-#| label: tbl-philly-sim-count
-#| tbl-cap: 'Count of Philadelphia area students included in student list purchase that filtered on 2020 high school class and PSAT scores 1070 -- 1180 (Not shown: American Indian/Alaska Native (AI/AN); and Native Hawaiian and other Pacific Islander (NHPI))'
-
-cat(as.character(formatted_tables[[1]]))
-```
-
-```{r philly-sim-col-pct, echo = FALSE, results='asis'}
-#| label: tbl-philly-sim-col-pct
-#| tbl-cap: 'Percent of total prospects purchased from each geomarket, Philadelphia area students included in student list purchase that filtered on 2020 high school class and PSAT scores 1070 -- 1180 (Not shown: AI/AN, NHPI)'
-
-cat(as.character(formatted_tables[[2]]))
-```
-
-```{r philly-sim-row-pct, echo = FALSE, results='asis'}
-#| label: tbl-philly-sim-row-pct
-#| tbl-cap: 'Racial composition of purchased prospects from each Geomarket, Philadelphia area students included in student list purchase that filtered on 2020 high school class and PSAT scores 1070 -- 1180 (Not shown: AI/AN, NHPI)'
-
-cat(as.character(formatted_tables[[3]]))
-```
-
-\newpage
-
-```{r fig-nyc-geomarkets, echo=FALSE, fig.cap="Geomarkets in New York City Area, NY state", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
-knitr::include_graphics(path = file.path('.','..','results','figures','nyc_geomarkets.png'))
-```
-
-\newpage
-
-```{r fig-asu-order-448922, echo=FALSE, fig.cap="Partial order summary from a College Board student list purchase", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
-knitr::include_graphics(path = file.path('.','..','results','figures','asu_order_448922_Redacted.pdf'))
-```
-
-\newpage
-
-```{r fig-socal-hispanic, echo=FALSE, fig.cap="Screenshot of Los Angeles area Geomarkets, showing 2020 percent Hispanic by census tract", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
-knitr::include_graphics(path = file.path('.','..','results','figures','socal_hispanic.png'))
-```
-
-\newpage
-
-```{r fig-socal-med-inc, echo=FALSE, fig.cap="Screenshot of Los Angeles area Geomarkets, showing 2020 median household income by census tract", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
-knitr::include_graphics(path = file.path('.','..','results','figures','socal_med_inc.png'))
-```
-
-\newpage
-
-```{r fig-philly-black-prospects, echo=FALSE, fig.cap="Screenshot of Philadelphia area Geomarkets, markers are high schools and marker size a function of number of Black student profiles purchased (2020 HS class, PSAT scores 1070 -- 1180)", fig.width=10, fig.height=7, out.extra='angle=90', out.width='110%'}
-knitr::include_graphics(path = file.path('.','..','results','figures','philly_black_prospects.png'))
-```

### 2025-10-09 text/manuscript-full-KS.qmd
added cerpp proposal

@@
 ```
+
 **RQ2**.
@@
 ```
+
 # Results
@@
 ```
+
 @fig-chicago-rq2-midSAT-race shows which Geomarkets send the bulk of students for a given race/ethnicity.
@@
 ```
+
 ## Los Angeles
@@
 ```
+
 <!--# Lastly, @fig-la-rq1-ses shows a widening disparity in educational attainment across the Los Angeles metropolitan area.
@@
 
-```{=tex}
 \begingroup
 \small
-```
+
 ::: {#refs}
@@
 
-```{=tex}
 \endgroup
@@
 \singlespacing
-```
+
 <!-- TABLES -->
@@
 
-```{=tex}
-\clearpage
-```
+\clearpage
 

### 2025-10-27 scripts/geomarket_cbsa.Rmd
add EPS shapes

---
title: '2020 Geomarkets + CBSA'
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    df_print: paged
---

```{r setup, include = F}
options(width = 10000)
knitr::opts_chunk$set(warning = F, message = F, echo = F, comment = '')

library(tidyverse)
library(leaflet)

scripts_dir <- '.'

eps_data_dir <- file.path('..', 'data', 'eps_market')
```

```{r}
# EPS shapes (2020)
load(file.path(eps_data_dir, 'eps_shapes_2020.RData'))

# County shapes (2020)
load(file.path(eps_data_dir, 'counties_2020.RData'))

# CBSA shapes (2020)
load(file.path(eps_data_dir, 'cbsa_2020.RData'))
```

```{r}
plot_map <- function(eps_codes, palette = c(RColorBrewer::brewer.pal(11, 'Spectral'), 'black', 'white')) {
  state <- unique(str_match(eps_codes, '^[A-Z]{2}')[,1])

  if (length(state) > 1) print(state)

  cbsas <- cbsa_2020 %>% filter(str_detect(NAME, str_c(', ', state)) | str_detect(NAME, str_c('-', state)))
  metros <- cbsas %>% filter(LSAD == 'M1')
  micros <- cbsas %>% filter(LSAD == 'M2')

  m <- leaflet() %>% 
    addProviderTiles(providers$CartoDB.Positron) %>% 
    addPolygons(data = counties_2020 %>% filter(STUSPS %in% state) %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, color = 'gray', weight = 1, group = 'Counties')

  if (nrow(metros) > 0) {
    m <- m %>%
      addPolygons(data = metros %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, color = 'gray', weight = 1, group = 'CBSA (Metro Area)')
  }

  if (nrow(micros) > 0) {
    m <- m %>%
      addPolygons(data = micros %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, color = 'gray', weight = 1, group = 'CBSA (Micro Area)')
  }

  bbox <- sf::st_bbox(eps_2020 %>% filter(eps %in% eps_codes)) %>% as.vector()

  for (i in seq_along(eps_codes)) {
    m <- m %>% 
      addPolygons(data = eps_2020 %>% filter(eps == eps_codes[[i]]), label = ~paste0('<b>', eps, '</b>') %>% lapply(htmltools::HTML), color = if_else(i %in% 5:7, '#c4c4c4', palette[[i]]), fillColor = palette[[i]], weight = 1, group = 'EPS from 2020 tracts')
  }

  m <- m %>% 
    addPolylines(data = eps_2020 %>% filter(eps %in% eps_codes), color = 'black', opacity = 1, weight = 2, group = 'EPS from 2020 tracts') %>% 
    addLayersControl(
      position = 'bottomleft',
      overlayGroups = c('EPS from 2020 tracts', 'CBSA (Metro Area)', 'CBSA (Micro Area)', 'Counties'),
      options = layersControlOptions(collapsed = F)
    ) %>% 
    fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
    hideGroup('CBSA (Metro Area)') %>% 
    hideGroup('CBSA (Micro Area)') %>% 
    hideGroup('Counties')

  m
}
```

# All CBSAs

- **Metro Areas**: Pink
- **Micro Areas**: Gray

```{r}
leaflet() %>% addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(data = cbsa_2020 %>% filter(LSAD == 'M1') %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, weight = 1, opacity = 1, color = 'pink') %>%
  addPolygons(data = cbsa_2020 %>% filter(LSAD == 'M2') %>% sf::st_transform(crs = 4326), label = ~NAMELSAD, weight = 1, opacity = 1, color = '#cccccc')
```

# 1) ME, NH, VT, CT and MA (excluding Boston)

## Connecticut (CT)

- **CT 1**: New London and Windham County
- **CT 2**: New Haven and Middlesex County
- **CT 3**: Fairfield County
- **CT 4**: Waterbury and Litchfield County
- **CT 5**: Hartford and Tolland County

```{r}
plot_map(
  c('CT 1', 'CT 2', 'CT 3', 'CT 4', 'CT 5')
)
```

## Maine (ME)

- **ME 1**: Portland and Southern Maine
- **ME 2**: Augusta and Central Maine
- **ME 3**: Bangor and Norther Maine

```{r}
plot_map(
  c('ME 1', 'ME 2', 'ME 3')
)
```

## Massachusetts (MA)

- **MA 1**: Berkshire and Franklin Counties
- **MA 2**: Springfield and Hampshire County
- **MA 3**: Fitchburg and North Worcester County
- **MA11**: Worcester

```{r}
plot_map(
  c('MA 1', 'MA 2', 'MA 3', 'MA11')
)
```

## New Hampshire

- **NH 1**: Seacost
- **NH 2**: Merrimack Valley (Manchester)
- **NH 3**: Monadnock and Lake Sunapee
- **NH 4**: Lakes and White Mountain

```{r}
plot_map(
  c('NH 1', 'NH 2', 'NH 3', 'NH 4')
)
```

## Vermont

- **VT 1**: Burlington
- **VT 2**: Southern Vermont
- **VT 3**: Northern and Eastern Vermont

```{r}
plot_map(
  c('VT 1', 'VT 2', 'VT 3')
)
```

# 2) RI and Greater Boston

## Rhode Island

- **RI 1**: Providence and Northern Rhode Island
- **RI 2**: Southern Rhode Island

```{r}
plot_map(
  c('RI 1', 'RI 2')
)
```

## Greater Boston

- **MA 4**: Essex County
- **MA 5**: Cape Cod and Islands
- **MA 6**: Boston and Cambridge
- **MA 7**: Quincy and Plymouth County
- **MA 8**: Lowell, Concord and Wellesley
- **MA 9**: Norfolk and Bristol County
- **MA10**: Milton, Lexington and Waltham

```{r}
plot_map(
  c('MA 4', 'MA 5', 'MA 6', 'MA 7', 'MA 8', 'MA 9', 'MA10')
)
```

# 6) NJ (North)

## New Jersey (North)

- **NJ 7**: Union County
- **NJ 8**: Essex and Souther Passaic County
- **NJ 9**: Hudson County
- **NJ10**: Bergen County
- **NJ11**: Morris and Norther Passaic County
- **NJ12**: Sussex, Warren and Hunterdon Counties

```{r}
plot_map(
  c('NJ 7', 'NJ 8', 'NJ 9', 'NJ10', 'NJ11', 'NJ12')
)
```

# 7) NJ (South)

## New Jersey (South)

- **NJ 1**: Southern Jersey
- **NJ 2**: Camden and Burlington County
- **NJ 3**: Jersey Shore and Pinelands
- **NJ 4**: Middlesex County
- **NJ 5**: Monmouth County
- **NJ 6**: Somerset and Mercer Counties

```{r}
plot_map(
  c('NJ 1', 'NJ 2', 'NJ 3', 'NJ 4', 'NJ 5', 'NJ 6')
)
```

# 8) DE and Greater Philadeliphia

## Delaware

- **DE 1**: New Castle County
- **DE 2**: Kent and Sussex Counties

```{r}
plot_map(
  c('DE 1', 'DE 2')
)
```

## Greater Philadelphia

- **PA 1**: Bucks County
- **PA 2**: Chester County
- **PA 3**: Delaware County
- **PA 4**: Montgomery County
- **PA 5**: Philadelphia County

```{r}
plot_map(
  c('PA 1', 'PA 2', 'PA 3', 'PA 4', 'PA 5')
)
```

# 9) PA (Excluding Philadelphia)

## Pennsylvania (excluding Philadelphia)

- **PA 6**: Lehigh Valley
- **PA 7**: Northeastern Pennsylvania
- **PA 8**: North Central Pennsylvania
- **PA 9**: Northwestern Pennsylvania
- **PA10**: Southern Pennsylvania (East)
- **PA11**: Southern Pennsylvania (West)
- **PA12**: Allegheny County
- **PA13**: Unnamed

```{r}
plot_map(
  c('PA 6', 'PA 7', 'PA 8', 'PA 9', 'PA10', 'PA11', 'PA12', 'PA13')
)
```

# 10) MD

## Maryland

- **MD 1**: Western Maryland
- **MD 2**: Montgomery Metropolitan
- **MD 3**: Central Maryland excluding Baltimore
- **MD 4**: Eastern Shore
- **MD 5**: Prince Georges Metropolitan
- **MD 6**: Southern Maryland
- **MD 7**: Baltimore (Urban)

```{r}
plot_map(
  c('MD 1', 'MD 2', 'MD 3', 'MD 4', 'MD 5', 'MD 6', 'MD 7')
)
```

# 11) DC and Northern VA

## District of Columbia

- **DC 1**: District of Columbia

```{r}
plot_map(
  c('DC 1')
)
```

## Northern Virginia

- **VA 1**: Arlington and Alexandria
- **VA 2**: Fairfax County
- **VA 3**: North Central Virginia

```{r}
plot_map(
  c('VA 1', 'VA 2', 'VA 3')
)
```

# 12) KY, TN, WV and VA (excluding Northern VA)

## Kentucky

- **KY 1**: Lexington and Fayette
- **KY 2**: Louisville and Western Kentucky

```{r}
plot_map(
  c('KY 1', 'KY 2')
)
```

## Tennessee

- **TN 1**: Chattanooga
- **TN 2**: Knoxville
- **TN 3**: Memphis
- **TN 4**: Nashville and Davidson

```{r}
plot_map(
  c('TN 1', 'TN 2', 'TN 3', 'TN 4')
)
```

## West Virginia

- **WV 1**: Charleston and Huntington
- **WV 2**: Northern West Virginia

```{r}
plot_map(
  c('WV 1', 'WV 2')
)
```

## Virginia (excluding Northern VA)

- **VA 4**: Northern Neck
- **VA 5**: Central Virginia
- **VA 6**: Richmond
- **VA 7**: Southside Virginia
- **VA 8**: Tidewater
- **VA 9**: Shenandoah
- **VA10**: Southwest Virginia

```{r}
plot_map(
  c('VA 4', 'VA 5', 'VA 6', 'VA 7', 'VA 8', 'VA 9', 'VA10')
)
```

# 13) NC and SC

## North Carolina

- **NC 1**: Coastal Plains
- **NC 2**: East Central
- **NC 3**: Research Triangle
- **NC 4**: Sand Hills
- **NC 5**: North Piedmont
- **NC 6**: South Piedmont
- **NC 7**: Western North Carolina

```{r}
plot_map(
  c('NC 1', 'NC 2', 'NC 3', 'NC 4', 'NC 5', 'NC 6', 'NC 7')
)
```

## South Carolina

- **SC 1**: Pee Dee
- **SC 2**: Low Country
- **SC 3**: Mid Lands
- **SC 4**: East Piedmont
- **SC 5**: West Piedmont

```{r}
plot_map(
  c('SC 1', 'SC 2', 'SC 3', 'SC 4', 'SC 5')
)
```

# 14) Northern GA

## Northern Georgia

- **GA 1**: Cherokee, Cobb and Douglass Counties
- **GA 2**: Fulton County
- **GA 3**: DeKalb and Gwinnett Counties
- **GA 4**: Clayton, Fayette, Henry and Rockdale Counties
- **GA 5**: Northeast Georgia
- **GA 8**: Northwest Georgia

```{r}
plot_map(
  c('GA 1', 'GA 2', 'GA 3', 'GA 4', 'GA 5', 'GA 8')
)
```

# 15) LA, MS, AL, GA (South) and FL (North)

## Louisiana

- **LA 1**: Baton Rouge
- **LA 2**: New Orleans
- **LA 3**: Shreveport

```{r}
plot_map(
  c('LA 1', 'LA 2', 'LA 3')
)
```

## Mississippi

- **MS 1**: Jackson
- **MS 2**: Northern Mississippi

```{r}
plot_map(
  c('MS 1', 'MS 2')
)
```

## Alabama

- **AL 1**: Birmingham and Tuscaloosa
- **AL 2**: Huntsville and Florence
- **AL 3**: Mobile
- **AL 4**: Montgomery

```{r}
plot_map(
  c('AL 1', 'AL 2', 'AL 3', 'AL 4')
)
```

## Georgia (South)

- **GA 6**: Southeast Georgia
- **GA 7**: Southwest Georgia

```{r}
plot_map(
  c('GA 6', 'GA 7')
)
```

## Florida (North)

- **FL 1**: Panhandle
- **FL 2**: Crown
- **FL 3**: East Central

```{r}
plot_map(
  c('FL 1', 'FL 2', 'FL 3')
)
```

# 16) FL (South), AS, GU, MP, PR, UM and VI

## Florida (South)

- **FL 4**: West Central
- **FL 5**: Broward, Martin and Palm Beach Counties
- **FL 6**: Dade County
- **FL 7**: Collier, Hendry and Monroe Counties

```{r}
plot_map(
  c('FL 4', 'FL 5', 'FL 6', 'FL 7')
)
```

# 17) OH

## Ohio

- **OH 1**: Northwest Ohio
- **OH 2**: North Central Ohio
- **OH 3**: City of Cleveland (West)
- **OH 4**: City of Cleveland (East)
- **OH 5**: Cuyahoga, Geauga, and Lake Counties
- **OH 6**: Northeast Ohio
- **OH 7**: West Central Ohio
- **OH 8**: Central Ohio
- **OH 9**: Greater Cincinnati
- **OH10**: Southeast Ohio

```{r}
plot_map(
  c('OH 1', 'OH 2', 'OH 3', 'OH 4', 'OH 5', 'OH 6', 'OH 7', 'OH 8', 'OH 9', 'OH10')
)
```

# 18) MI and WI

## Michigan

- **MI 1**: Wayne County
- **MI 2**: Detroit's Northern Suburbs
- **MI 3**: Ann Arbor
- **MI 4**: Capital District
- **MI 5**: Kalamazoo and Grand Rapids
- **MI 6**: "The Thumb"
- **MI 7**: Northern Michigan

```{r}
plot_map(
  c('MI 1', 'MI 2', 'MI 3', 'MI 4', 'MI 5', 'MI 6', 'MI 7')
)
```

## Wisconsin

- **WI 1**: Madison and Janesville
- **WI 2**: Milwaukee and Racine
- **WI 3**: Northern Wisconsin

```{r}
plot_map(
  c('WI 1', 'WI 2', 'WI 3')
)
```

# 19) Greater Chicago

## Greater Chicago

- **IL 7**: Chain of Lakes
- **IL 8**: Northwest Suburbs
- **IL 9**: North Shore
- **IL10**: Evanston and Skokie
- **IL11**: City of Chicago
- **IL12**: Western Suburbs
- **IL13**: South and Southwest Suburbs

```{r}
plot_map(
  c('IL 7', 'IL 8', 'IL 9', 'IL10', 'IL11', 'IL12', 'IL13')
)
```

# 20) IL and IN (excluding Chicago)

## Illinois

- **IL 1**: Rockford
- **IL 2**: Quad Cities
- **IL 3**: Peoria
- **IL 4**: Springfield
- **IL 5**: Decatur and Champaign
- **IL 6**: Southern Illinois

```{r}
plot_map(
  c('IL 1', 'IL 2', 'IL 3', 'IL 4', 'IL 5', 'IL 6')
)
```

## Indiana

- **IN 1**: "The Region"
- **IN 2**: Northwest Indiana
- **IN 3**: South Bend and Elkhart
- **IN 4**: Northeast Indiana
- **IN 5**: West Central Indiana
- **IN 6**: East Central Indiana
- **IN 7**: Greater Indianapolis
- **IN 8**: West Indiana
- **IN 9**: South Central Indiana
- **IN10**: East Indiana
- **IN11**: Southwest Indiana
- **IN12**: Southeast Indiana

```{r}
plot_map(
  c('IN 1', 'IN 2', 'IN 3', 'IN 4', 'IN 5', 'IN 6', 'IN 7', 'IN 8', 'IN 9', 'IN10', 'IN11', 'IN12')
)
```

# 21) ND, SD, MN, IA, NE, KS, MO, AR and OK

## North Dakota

- **ND 1**: Fargo and Eastern North Dakota
- **ND 2**: Western North Dakota

```{r}
plot_map(
  c('ND 1', 'ND 2')
)
```

## South Dakota

- **SD 1**: Souix Falls and Eastern South Dakota
- **SD 2**: Western South Dakota

```{r}
plot_map(
  c('SD 1', 'SD 2')
)
```

## Minnesota

- **MN 1**: Twin Cities
- **MN 2**: Northern Minnesota

```{r}
plot_map(
  c('MN 1', 'MN 2')
)
```

## Iowa

- **IA 1**: Cedar Rapids and Eastern Iowa
- **IA 2**: Des Moines and Western Iowa

```{r}
plot_map(
  c('IA 1', 'IA 2')
)
```

## Nebraska

- **NE 1**: Lincoln
- **NE 2**: Omaha
- **NE 3**: Western Nebraska

```{r}
plot_map(
  c('NE 1', 'NE 2', 'NE 3')
)
```

## Kansas

- **KS 1**: Kansas City and Topeka
- **KS 2**: Wichita and Western Kansas

```{r}
plot_map(
  c('KS 1', 'KS 2')
)
```

## Missouri

- **MO 1**: Kansas City and St. Joseph
- **MO 2**: St. Louis and Eastern Missouri
- **MO 3**: Springfield and Southern Missouri

```{r}
plot_map(
  c('MO 1', 'MO 2', 'MO 3')
)
```

## Arkansas

- **AR 1**: Little Rock
- **AR 2**: Northern Arkansas

```{r}
plot_map(
  c('AR 1', 'AR 2')
)
```

## Oklahoma

- **OK 1**: Oklahoma City and Western Oklahoma
- **OK 2**: Tulsa and Western Oklahoma

```{r}
plot_map(
  c('OK 1', 'OK 2')
)
```

# 22) TX (excluding Dallas-Ft. Worth and Houston)

## Texas (excluding D-FW and Houston)

- **TX 1**: Amarillo, Panhandle and South Plains
- **TX 2**: El Paso
- **TX 3**: Midland, Odessa, and Trans Pecos
- **TX 4**: Abilene and San Angelo
- **TX 5**: Red River Area
- **TX 6**: Austin and Central Texas
- **TX 7**: Waco, Temple and Killeen
- **TX 8**: East Texas
- **TX 9**: Beaumont and Port Arthur
- **TX10**: Central Gulf Coast, Wharton County and Victoria County
- **TX11**: South Texas Valley
- **TX13**: Del Rio, Uvalde County, and Bexar County Area
- **TX14**: City of San Antonio

```{r}
plot_map(
  c('TX 1', 'TX 2', 'TX 3', 'TX 4', 'TX 5', 'TX 6', 'TX 7', 'TX 8', 'TX 9', 'TX10', 'TX11', 'TX13', 'TX14')
)
```

# 23) Greater Dallas-Ft.Worth

## Greater Dallas-Ft.Worth

- **TX19**: City of Dallas
- **TX20**: City of Ft. Worth
- **TX21**: Irving, Arlington and Grand Prairie
- **TX22**: Dallas County excluding City of Dallas
- **TX23**: Collin and Rockwall Counties
- **TX24**: Counties West of Dallas/Ft. Worth Metroplex

```{r}
plot_map(
  c('TX19', 'TX20', 'TX21', 'TX22', 'TX23', 'TX24')
)
```

# 24) Greater Houston

## Greater Houston

- **TX12**: Brazos and Trinity Valley
- **TX15**: Northwest Houston and Conroe School District
- **TX16**: Southwest Houston Metro Area
- **TX17**: City of Houston (East)
- **TX18**: Galveston and East Harris Counties

```{r}
plot_map(
  c('TX12', 'TX15', 'TX16', 'TX17', 'TX18')
)
```

# 25) AZ, NM, CO & UT

## Arizona

- **AZ 1**: Phoenix
- **AZ 2**: Tuscon
- **AZ 3**: Northern Arizona

```{r}
plot_map(
  c('AZ 1', 'AZ 2', 'AZ 3')
)
```

## New Mexico

- **NM 1**: Albuquerque and Northern New Mexico
- **NM 2**: Southern New Mexico

```{r}
plot_map(
  c('NM 1', 'NM 2')
)
```

## Colorado

- **CO 1**: Colorado Springs and Southeastern Colorado
- **CO 2**: Metro and Northeastern Colorado
- **CO 3**: Mountain Western Colorado

```{r}
plot_map(
  c('CO 1', 'CO 2', 'CO 3')
)
```

## Utah

- **UT 1**: Salt Lake City, Ogden & Provo
- **UT 2**: Southern Utah

```{r}
plot_map(
  c('UT 1', 'UT 2')
)
```

# 26) HI and Southern CA (excluding Los Angeles)

## Southern California

- **CA12**: Central Coast
- **CA13**: Santa Barbara and West Ventura Counties
- **CA27**: Riverside, San Bernardino and Ontario
- **CA28**: South Orange County
- **CA29**: North San Diego County excluding San Diego
- **CA30**: South San Diego County excluding San Diego
- **CA31**: City of San Diego
- **CA32**: Central Valley North
- **CA33**: Central Valley South
- **CA34**: Greater Imperial Valley

```{r}
plot_map(
  c('CA12', 'CA13', 'CA27', 'CA28', 'CA29', 'CA30', 'CA31', 'CA32', 'CA33', 'CA34')
)
```

# 27) Greater Los Angeles

## Greater Los Angeles

- **CA14**: San Fernando Valley (West)
- **CA15**: San Fernando Valley (East)
- **CA16**: Glendale and Pasadena
- **CA17**: West Los Angeles and West Beach
- **CA18**: Hollywood and Wilshire
- **CA19**: East Los Angeles
- **CA20**: South Bay
- **CA21**: South and Central Los Angeles
- **CA22**: Long Beach
- **CA23**: Covina and West Covina
- **CA24**: Whittier and North Orange County
- **CA25**: Anaheim
- **CA26**: Santa Ana

```{r}
plot_map(
  c('CA14', 'CA15', 'CA16', 'CA17', 'CA18', 'CA19', 'CA20', 'CA21', 'CA22', 'CA23', 'CA24', 'CA25', 'CA26')
)
```

# 28) NV and Northern CA

## Nevada

- **NV 1**: Las Vegas
- **NV 2**: Reno

```{r}
plot_map(
  c('NV 1', 'NV 2')
)
```

## Northern California

- **CA 1**: Far Northern California
- **CA 2**: Valley of the Moon
- **CA 3**: Sacramento County
- **CA 4**: Marin County
- **CA 5**: San Francisco County
- **CA 6**: Contra Costa County
- **CA 7**: City of Oakland
- **CA 8**: Alameda County excluding Oakland
- **CA 9**: San Mateo County
- **CA10**: City of San Jose
- **CA11**: Santa Clara County excluding San Jose

```{r}
plot_map(
  c('CA 1', 'CA 2', 'CA 3', 'CA 4', 'CA 5', 'CA 6', 'CA 7', 'CA 8', 'CA 9', 'CA10', 'CA11')
)
```

# 29) WA, OR, ID, MT, WY and AK

## Washington

- **WA 1**: Greater Seattle
- **WA 2**: South Sound
- **WA 3**: Greater Spokane
- **WA 4**: Greater Washington (East)
- **WA 5**: Greater Washington (West)
- **WA 6**: Bellingham

```{r}
plot_map(
  c('WA 1', 'WA 2', 'WA 3', 'WA 4', 'WA 5', 'WA 6')
)
```

## Oregon

- **OR 1**: Greater Portland (West)
- **OR 2**: Greater Portland (East)
- **OR 3**: Northern Valley
- **OR 4**: Southern Valley
- **OR 5**: Southwest Valley
- **OR 6**: East Oregon

```{r}
plot_map(
  c('OR 1', 'OR 2', 'OR 3', 'OR 4', 'OR 5', 'OR 6')
)
```

## Idaho

- **ID 1**: Boise City
- **ID 2**: Northern Idaho

```{r}
plot_map(
  c('ID 1', 'ID 2')
)
```

## Montana

- **MT 1**: Billings and Eastern Montana
- **MT 2**: Western Montana

```{r}
plot_map(
  c('MT 1', 'MT 2')
)
```

## Wyoming

- **WY 1**: Casper and Chayenne
- **WY 2**: Western Wyoming

```{r}
plot_map(
  c('WY 1', 'WY 2')
)
```

## Alaska

- **AK 1**: Anchorage, Kenai, and Mat-su District
- **AK 2**: Greater Alaska

```{r}
plot_map(
  c('AK 1', 'AK 2')
)
```

### 2025-10-27 scripts/save_eps_shapes.R

@@ -1,5 +1,6 @@
 library(tidyverse)
 library(tidycensus)
+library(tigris)
 library(sf)
 library(leaflet)
 
@@ -45,12 +46,13 @@ source(file.path(scripts_dir, 'metro_eps_codes.R'))
 # save(tracts_2000, file = file.path(eps_data_dir, 'tracts_2000.RData'))
 
 # Tract shapes (2020)
+# states <- c(state.abb, 'DC')
 # tracts_2020 <- do.call(
 #   bind_rows,
-#   lapply(c('GA', 'CA', 'MA', 'IL', 'OH', 'TX', 'DC', 'MD', 'VA', 'MI', 'NY', 'FL', 'NJ', 'PA'), function(st) {
+#   lapply(states, function(st) {
 #     tracts(
 #       state = st,
-#       cb = TRUE, 
+#       cb = TRUE,
 #       year = 2020
 #     )
 #   })
@@ -88,6 +90,14 @@ source(file.path(scripts_dir, 'metro_eps_codes.R'))
 # 
 # save(county_subdivisions_2020, file = file.path(eps_data_dir, 'county_subdivisions_2020.RData'))
 
+# CBSA shapes (2020)
+# cbsa_2020 <- core_based_statistical_areas(
+#   cb = TRUE,
+#   year = 2020
+# )
+# 
+# save(cbsa_2020, file = file.path(eps_data_dir, 'cbsa_2020.RData'))
+
 # Place shapes (2020)
 # places_2020 <- do.call(
 #   bind_rows,
@@ -121,8 +131,8 @@ load(file.path(eps_data_dir, 'tracts_2020.RData'))
 # Save shapes
 # ------------
 
-get_shapes <- function(tracts_df, geo_col, yr, eps) {
-  eps_files <- str_c(eps, '.txt')
+get_shapes <- function(tracts_df, geo_col, yr, eps_files) {
+  eps <- str_split(eps_files, '\\.txt', simplify = T)[,1]
   tracts_df %>% 
     mutate(eps = case_when(!!!rlang::parse_exprs(str_c(geo_col, ' %in% str_split(read.table("', file.path(tracts_dir, yr, eps_files), '"), ",")[[1]] ~ "', eps, '"')))) %>% 
     filter(!is.na(eps)) %>% 
@@ -132,18 +142,18 @@ get_shapes <- function(tracts_df, geo_col, yr, eps) {
 }
 
[…]

### 2025-10-27 scripts/save_tracts.R

@@ -18,20 +18,48 @@ load(file.path(eps_data_dir, 'tracts_1980.RData'))
 load(file.path(eps_data_dir, 'tracts_2000.RData'))
 load(file.path(eps_data_dir, 'tracts_2020.RData'))
 
-metro <- tracts_2020 %>% filter(str_detect(GEOID, '^17(097|031|043|197)'))
+counties <- '^48'
+metro <- tracts_2020 %>% filter(str_detect(GEOID, counties))
 # metro <- tracts_2000 %>% filter(str_detect(gisjoin2, '^170(097|031|043|197)'))
 # metro <- tracts_1980 %>% filter(str_detect(gisjoin2, '^170(097|031|043|197)'))
 
 # Load previously traced shapes for reference
 eps_2020 <- tracts_2020 %>%
   mutate(eps = case_when(
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL 7.txt')), ',')[[1]] ~ 'IL 7',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL 8.txt')), ',')[[1]] ~ 'IL 8',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL 9.txt')), ',')[[1]] ~ 'IL 9',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL10.txt')), ',')[[1]] ~ 'IL10',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL11.txt')), ',')[[1]] ~ 'IL11',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL12.txt')), ',')[[1]] ~ 'IL12',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'IL13.txt')), ',')[[1]] ~ 'IL13'
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 1.txt')), ',')[[1]] ~ 'X1',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 2.txt')), ',')[[1]] ~ 'X2',
+    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 3.txt')), ',')[[1]] ~ 'X3',
[…]

### 2025-11-19 scripts/geomarket_cbsa.Rmd
add missing EPS shapes

@@
 ```
+
+# Other
+
+Not in `connections-geographies-cb.pdf`, so base the shapes off the original zip code EPS geomarkets. (See: `cb_geo/data/eps_market/eps_shapes.RData` and `cb_geo/data/eps_market/eps_names.csv`)
+
+## Hawaii
+
+- **HI 1**: Island of Oahu
+- **HI 2**: Remaining Hawaiian Islands
+
+```{r}
+plot_map(
+  c('HI 1', 'HI 2')
+)
+```
+
+## New York
+
+- **NY 1**: Southern Tier West
+- **NY 2**: Erie Co
+- **NY 3**: Genesee Valley & N. Frontier
+- **NY 4**: Rochester & Monroe Co
+- **NY 5**: Finger Lakes Region
+- **NY 6**: Central New York
+- **NY 7**: Saint Lawrence Valley
+- **NY 8**: Adirondacks
+- **NY 9**: Tri Cities
+- **NY10**: Central Hudson Valley
+- **NY11**: Catskills
+- **NY12**: Southern Tier East
+- **NY13**: Rockland Co
+
+```{r}
+plot_map(
+  c('NY 1', 'NY 2', 'NY 3', 'NY 4', 'NY 5', 'NY 6', 'NY 7', 'NY 8', 'NY 9', 'NY10', 'NY11', 'NY12', 'NY13')
+)
+```
+
+## Long Island
+
+- **NY16**: Southern Nassau Co
+- **NY17**: Northern Nassau Co
+- **NY18**: Central Nassau Co
+- **NY19**: Northwest Suffolk Co
+- **NY20**: Southwest Suffolk Co
+- **NY21**: East Suffolk Co
+
+```{r}
+plot_map(
+  c('NY16', 'NY17', 'NY18', 'NY19', 'NY20', 'NY21')
+)
+```
+
+## NY, NY
+
+- **NY14**: Staten Island
+- **NY15**: Westchester Co
+- **NY22**: Southeast Brooklyn
+- **NY23**: West Brooklyn
+- **NY24**: Northeast Brooklyn
+- **NY25**: East Bronx
+- **NY26**: West Bronx
+- **NY27**: Manhattan
+- **NY28**: South Queens
+- **NY29**: Northwest Queens
+- **NY30**: Northeast Queens
+
+```{r}
+plot_map(
+  c('NY14', 'NY15', 'NY22', 'NY23', 'NY24', 'NY25', 'NY26', 'NY27', 'NY28', 'NY29', 'NY30')
+)
+```

### 2025-11-19 scripts/save_tracts.R

@@ -18,7 +18,7 @@ load(file.path(eps_data_dir, 'tracts_1980.RData'))
 load(file.path(eps_data_dir, 'tracts_2000.RData'))
 load(file.path(eps_data_dir, 'tracts_2020.RData'))
 
-counties <- '^48'
+counties <- '^36'
 metro <- tracts_2020 %>% filter(str_detect(GEOID, counties))
 # metro <- tracts_2000 %>% filter(str_detect(gisjoin2, '^170(097|031|043|197)'))
 # metro <- tracts_1980 %>% filter(str_detect(gisjoin2, '^170(097|031|043|197)'))
@@ -26,36 +26,36 @@ metro <- tracts_2020 %>% filter(str_detect(GEOID, counties))
 # Load previously traced shapes for reference
 eps_2020 <- tracts_2020 %>%
   mutate(eps = case_when(
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 1.txt')), ',')[[1]] ~ 'X1',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 2.txt')), ',')[[1]] ~ 'X2',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 3.txt')), ',')[[1]] ~ 'X3',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 4.txt')), ',')[[1]] ~ 'X4',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 5.txt')), ',')[[1]] ~ 'X5',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 6.txt')), ',')[[1]] ~ 'X6',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 7.txt')), ',')[[1]] ~ 'X7',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 8.txt')), ',')[[1]] ~ 'X8',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX 9.txt')), ',')[[1]] ~ 'X9',
-    GEOID %in% str_split(read.table(file.path(tracts_dir, '2020', 'TX10.txt')), ',')[[1]] ~ 'X10',
[…]
